### 2017-12-27 manuscript/manuscript.Rmd
initialized manuscript

---
title: "A comprehensive analysis of natural sequence variation within the <i>Arabidopsis thaliana</i> nuclear auxin signaling pathway"
author: "R. Clay Wright"
date: "`r Sys.Date()`"
output: 
   pdf_document: 
     toc: true
     number_sections: true
bibliography: references.bib
csl: apa.csl
---

```{r, echo = FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(warning = FALSE, collapse = TRUE, comment = "#>", tidy = TRUE, out.width = "90%", fig.height=4, fig.width=7, out.extra='style="margin: auto; display: block; padding-top: 15px;"')
library(r1001genomes)
library(DT)
library(dplyr)

```

<style>
  .using-app-insert {
     background-color: #bfbfbf;
     padding: 15px;
     border-style: solid;
     border-color:  #8c8c8c;
     border-radius: 7px 7px 0 0;
     border-width: 1px;
  }

  .using-R-insert {
     background-color: #e6e6e6;
     padding: 15px;
     border-style: solid;
     border-color:  #8c8c8c;
     border-radius: 0 0 7px 7px;
     border-width: 1px;
     border-top: 0;
  }

  .using-app-insert h2, .using-app-insert h3, .using-app-insert h4 {
     border-bottom: 0;
  }

  .using-R-insert h2, .using-R-insert h3, .using-R-insert h4 {
     border-bottom: 0;
  }

  img {
    display: block;
    margin: auto;
    width: 90%;
    height: auto;

  }
</style>

# Introduction

This web app and open source software package for the R programming language was built to facilitate exploration of the naturally occuring sequence diversity for your favorite gene or gene family. We hope this product will help plant biologists without bioinformatics experience access the 1001 genomes project’s rich catalog of natural variation, formulate hypotheses and identify new interesting alleles in their favorite gene family.

The 1001 genomes dataset was collected with the intention of genome scale analyses, such as genome wide association studies, in which a particular phenotype is quantified for all of the accessions and sequence variants that are associated with phenotypic variation are identified. [In recent work](https://doi.org/10.1534/genetics.117.3000), we have utilized this rich dataset in a different manner to measure the effects of coding sequence variation on protein functional variation and phenotypic variation . This work has expanded the sequence-function-phenotype map for the auxin-signaling F-box gene family, and this product will facilitate similar analysis other genes, gene families, and genomic regions.

So pick out your favorite gene, gene family or genomic region and follow the steps in this guide to analyze how it varies in plants collected from all over the world. For this example we will be using the auxin-signaling F-box genes, *TIR1* and *AFB2* which have the TAIR locus IDs [AT3G62980](https://www.arabidopsis.org/servlets/TairObject?id=39952&type=locus) and [AT3G26810](https://www.arabidopsis.org/servlets/TairObject?id=38086&type=locus). Click on one of the previous links to search TAIR for the TAIR IDs of your favorite genes.

***
Before digging into the data, make some hypotheses about what genes, domains and/or postions in your 
favorite gene family might be under different evolutionary pressures. We can make some inferences about the evolutionary pressure acting on particular genes based on the sequence and geographic diversity observed in the 1001 genomes dataset. Thomas Mitchell-Olds and Johanna Schmitt published an excellent overview of the evolutionary inference from natural genetic diversity in *A. thaliana* [@mitchell-olds_genetic_2006]. We can calculate genetic diversity as the average number of nucleotide differences per site between two DNA sequences in all possible pairs in the sample population, and is denoted by $\pi$.

However, evolutionary pressures potentially exert influence across all levels of biology, nucleotides, genes, genomes, phenotypes, organisms and populations.  This makes any inference from study of genetic diversity contingent on potentially unknown effects of epistasis between genes, population structure, environment, and many other effectors. Additionally, as an agricultural tag-along, much of the genetic variation in *A. thaliana* is due to admixture between populations as the result of human influence migration. So keep in mind throughout this analysis the limitations of examining a single snapshot in time of a potentially biased collection of frequently self-fertizilizing organisms that were dispersed by human migration. Because of these limitations, infering theories about how evolutionary forces are acting on your favorite genes is very uncertain business. However, if we assume a few things we can make some speculations.

Assumptions:  
1. Polymorphisms in the dataset are not *completely* detrimental to organismal fitness.  
2. Most nonsynonymous polymorphisms in a gene result in a reduction in function.

Now, taking a gene-centric view of evolution, here are some of the evolutionary pressures, or types of natural selection, that we might observe:

### Purifying selection  
If a gene's sequence is highly conserved (low diversity) it may be under [purifying/negative/stabilizing selection](https://en.wikipedia.org/wiki/Negative_selection_(natural_selection)). 
If a genes' function is crucial to the fitness of the organism, and most polymorphisms result in loss of function, we would expect a low level of nonsynonymous polymorphism in this gene. This is because deleterious polymorphisms that do arise in this gene within the population will be outcompeted by the more-fit parent. This results in the decreased genetic diversity in genes under purifying selection.

### Genetic Drift
If evolutionary pressure is very weak on a gene its sequence might change with no effect on organismal fitness. This allows a gene to accumulate many mutations that may persist in the population or be lost at random. This results in random fluctuations in allele frequencies, which when combined with an appreciable rate of mutation and a sufficient population size yields increased genetic diversity in the population.

Beyond purifying selection and drift it is very difficult to make any strong inferences on the population as a whole. However there are two more types of selective pressure, balancing or neutral selection and positive or directional selection.

### Balancing selection
Genes of high diversity may be under balancing selection, also called neutral divergence. Balancing selection maintains multiple variants for long evolutionary times because different alleles provide selective advantages in different environments or genetic backgrounds. This results in high diversity at the particular site under selection and increased diversity at closely linked neutral sites due to genetic hitchhiking with this allele under balancing selction.

### Directional selection  
If a polymorphism (or allele) is associated with a phenotype that provides an organims with a fitness advantage this phenotype and associated allele may be experiencing directional or positive selection. Positive/directional selection results in an increase in the prevalence of the advantageous allele within the population over time. Directional selection is one of the most difficult things to infer and is a very active area of research. Directional selection, like stabilizing selection is associated with reduced genetic diversity within the population under selection. However, unlike stabilizing selection the gene under directional selection will potentially have two major alleles that provide adaptive potential to different environments or genetic backgrounds.

# Hypothesize
Now go read/collect the literature about your family and make some hypotheses!  
Here are some questions to help you get started:

What structure/function knowledge is there?  
What is each domain in the protein doing?  
Are there any postions that are of vital importance?

Is there a published protein sequence alignment of your gene family?  
What postions in the protein family are highly conserved?  
What postions are highly variable?

Is there a published phylogenetic tree of your gene family?  
What are the most recently duplicated sister gene pairs?  
What do you know about the phenotypes of mutants in these pairs?  
Do the genetics suggest these pairs are redundant?

## For example
In TIR1 and AFB2 there are 2 domains. The F-box domain interacts with the SCF components [@gray_identification_1999], and the LRR domain binds auxin and Aux/IAA proteins [@kepinski_auxininduced_2004]. The F-box is roughly 40 amino acids and wraps tightly around the SKP1 component of the SCF complex. The LRR is greater than 500 amino acids and binds auxin in a pocket formed by a horseshoe shaped chain of LRRs [@tan_mechanism_2007]. The outside of this horseshoe shape does not have any known interactions other than a dimerization interface [@dezfulian_oligomerization_2016]. Because the F-box domain must interact with the SCF complex for TIR1/AFB2 to perform their known function, we might expect the F-box to be highly conserved. The small size of the F-box also helps strengthen this hypothesis. The LRR domain is much larger and may have differing specificities for Aux/IAAs, so we might expect it to be more diverse. 
We would also expect the residues of the LRR that contact auxin and the Aux/IAA degron to be highly conserved.

***

# A note on the format of this document
<!--=========================================================================-->
<div  class="using-app-insert"><!--panel start--------------------------------->
## Using the App

In addition to calling the functions of the r1001genomes package in R directly. there is also a GUI provided in the form of a Shiny app. This app can perform some of the analysis tasks the package is capable of performing, without the need to have knowledge of programming in R.

In this vignette panels like this one are used to show examples of using the app to achieve similar results as using  the R function calls directly in this vignette. Screenshots will usually be provided to make it clear what section or tab we are talking about

```{r, echo = FALSE}
knitr::include_graphics("Screenshots/App_Initial.PNG")
knitr::asis_output("
                   ") # This newline required for headings to render after image
```
## Starting the web app

to start the Shiny app, run the following command:
```{r, eval=FALSE}
run1001genomes()
```
The app should appear in a new window, and look similar to the image above.
</div><!--panel end------------------------------------------------------------>
<div  class="using-R-insert"><!--panel start----------------------------------->
## Using R
in these lighter colored panels, the equivalent R code will be displayed.
</div><!--panel end------------------------------------------------------------>
<!--=========================================================================-->

# Download the chromosome positions and gene models 
The first step in retrieving the natural variation in our genes is to retrieve the information about where the genes are in the genome of Col-0, the laboratory reference accession, and how the mRNA and coding sequence is laid out. This information is called the gene model.

<!--=========================================================================-->
<div  class="using-app-insert"><!--panel start--------------------------------->
## Loading Gene Data in the App
```{r, echo=FALSE}
knitr::include_graphics("Screenshots/Tab_1_Select_Genes.PNG")
knitr::asis_output("
                   ") # This newline required for headings to render after image
```
### Selecting Genes and Generating the Gene Model Table
At the top of the first tab (the "SNP Stats" tab) is a text input field where you can input a list of TAIR loci.

The app will download the gene model information as soon as you click submit. Copy and paste your TAIR loci in the box, separated by commas, and click submit.

The app may take a little while to process after clicking the submit button on the first tab, especially if you have a long list of genes, but once this process is complete on the first tab, the data for these genes will be already loaded for the subsequent tabs.

Once the loading process has finished, the first table you should see scrolling down the SNP Stats tab is the Gene Information table. it should look similar to the table produced in the "Using R" section below.
</div><!--panel end------------------------------------------------------------>
<div  class="using-R-insert"><!--panel start----------------------------------->
## Using R
The code the app uses to download this table is the `getGeneInfo` function. This function is just a simplifying wrapper for the `BioMart` function `getBM`. This downloads the gene model from ENSEMBL plants.

```{r}
tair_ids <- c("AT3G62980", "AT3G26810")
geneInfo <- getGeneInfo(genes = tair_ids, firstOnly = TRUE, inputType = "tair_locus", useCache = TRUE)
```

Now we can give this table a look. We can also freshen up the table a bit with `datatable`. 
```{r}
datatable(geneInfo[, -c(5,6,9)], colnames = c("tair locus", "symbol", "transcript", 
                                              "Chr", "transcript \nstart", "transcript \nend", 
                                              "transcript \nlength"), rownames = FALSE, 
          options=list(paging=FALSE, searching=FALSE))
```
</div><!--panel end------------------------------------------------------------>
<!--=========================================================================-->

Do you have a spreasheet of loci?

Here is some quick code to make a text string you can copy and paste into the app gene input box.  
```{r}
AFB_loci <- read.csv(system.file("extdata", "AFB_gene_ids.csv", package = "r1001genomes"))
paste(AFB_loci$tair_locus, collapse = ",")
```

Occasionally there are multiple models for a single gene if there are multiple possible transcripts or splice variants. Does your gene have multiple models?

# Download the variants in the 1001 genomes dataset for these genes
The 1001 genomes dataset is stored as a VCF (Variant Call Format) file. This is 
a very efficient format for storing such a large dataset (the 1135 genomes are ~17gb), 
and for some analyses its matrix format is very useful.

<!--=========================================================================-->
<div  class="using-app-insert"><!--panel start--------------------------------->
## Downloading and processing of VCFs in the App
Upon clicking the "Submit" button on the "SNP Stats" tab, the web app automatically creates a tidy dataframe and calculates the number of unique and nonunique polymorphisms in each region of the gene model as well as calculating the diversity in the protein coding sequence. The diversity is the average number of nucleotide differences per site between all possible pairs of sequences. Diversity is also calculated for only the set of synonymous and nonsynonymous polymorphisms.

The results of these calculations are displayed in the three tables below the Gene Information table.

</div><!--panel end------------------------------------------------------------>
<div  class="using-R-insert"><!--panel start----------------------------------->
## Using R
We can download a set of VCF files for our favorite genes using the `VCFList` function.

```{r, eval=FALSE}
YFG_VCF <- VCFList(geneInfo = geneInfo, by = "transcript", tidy = FALSE)
```

However, for some filtering, summarizing and plotting operations it is more efficient to 
convert the matrix into a very long data frame. Most of the functions in `r1001genomes` 
operate on a long data frame, often called a tidy data frame. We can automatically covert 
our VCF to a tidy data frame by setting the `tidy` argument to `TRUE`.

```{r, cache=TRUE}
YFG_VCF <- VCFList(geneInfo = geneInfo, by = "transcript", tidy = TRUE)
```

We now have a list of very long data frames with all the information on SNPs inside, but much of the details on the effects of these polymorphisms are contained in a single column labeld 'EFF' as a long text string, making it hard to filter and process the data by the individual attributes within that text string.
the `parseEFF` function will separate indiual fields of the EFF string into separate columns on a single dataframe, and can be applied to our list of data frames with the `llply` function from the plyr package
```{r}
YFG_VCF <- llply(YFG_VCF, parseEFF)
```

To generate the three tables of polymorphism stats the app calls the function `polymorphRow`. in R we can do the same thing.
```{r, cache=TRUE}
#TIR1_AFB2_table <- polymorphTable(geneInfo = geneInfo, strains = strains)
YFG_VCF <- llply(YFG_VCF, Nucleotide_diversity)
TIR1_AFB2_table <- ldply(YFG_VCF, polymorphRow, geneInfo = geneInfo, .id = "Transcript_ID")

```

Before displaing this table we will again clean it up a bit.

```{r}
formatRound(datatable(TIR1_AFB2_table[,c(1:6,8)],
                  colnames = c("transcript", "5' UTR", "intron", "3' UTR",
                               "coding \n synonymous", "coding \n missense",
                               "coding \n total"), 
                  escape = FALSE, rownames = FALSE,
    options=list(paging=FALSE, searching=FALSE)), columns = 2:7, digits = 0)
```
The nucleotide diversity for the chosen genes is also calculated in the call to `polymorphTable`.

```{r}
formatRound(datatable(TIR1_AFB2_table[, c(1,13,9:12)],
                  colnames = c("transcript",
                               "&pi; transcript",
                               "&pi; coding",
                               "&pi;<sub>N</sub>",
                               "&pi;<sub>S</sub>",
                               "&pi;<sub>N</sub>/&pi;<sub>S</sub>"), 
                  escape = FALSE,
                  options = list(paging=FALSE, searching=FALSE)),
                columns = 1:6, digits = 6)
```

</div><!--panel end------------------------------------------------------------>
<!--=========================================================================-->

## Results
Do these values match your hypotheses? Is there anything unexpected?

For our TIR1 and AFB2 example, we can make several interesting observations based on this dataset.  
1. *AFB2* had higher number of total polymorphisms than *TIR1*, across all subsets of sites except for introns. 
2. *AFB2* also has a higher total coding diversity and higher ratio of diversities at nonsynonymous to synonymous sites.

This was expected, as *TIR1* mutants have a strong loss of auxin sensitivity whereas *AFB2* mutants have a much weaker effect. Therefore, polymorphisms in *TIR1* are much more likely to have an effect on phenotype than polymorphisms in *AFB2*.

# Diversity within each gene

Now we'll move on to the second "Diversity Plot" tab. This tab will calculate the diversity at each position within the coding sequence of each gene. In addition to calculating the nucleotide diversity of a whole sequence, like the transcipt of a gene, we can also calculate the diversity statistic for each position along the reference genome. This could be useful if you suspect a particular region of your gene may be more important to function or more sensitive to nucelotide or amino acid changes.

<!--=========================================================================-->
<div  class="using-app-insert"><!--panel start--------------------------------->
## Diversity Plot
The second tab of the app, labeld "Diversity Plot" allows you to plot the nucleotide diversity for each of the genes you selected in the "SNP Stats" tab.  
```{r}
imagePath <- "Screenshots/Tab_2_Select_Gene.PNG"
knitr::include_graphics(imagePath)
knitr::asis_output("
                   ") # This newline required for headings to render after image
```
After clicking to the "Diversity Plot" tab, select the first gene you would like to plot then click the submit button.  
![](Screenshots/Tab_2_Plot_Selection_Box.PNG)

You should see a plot like the one above. The plot shows Nei's nucleotide diversity by *nucleotide position* of the protein coding region of the gene, though the X-axis is marked by codon number from the N-terminus for convenience. The symbols are differentiated by whether the site has a synonymous or missense variant. If multiple variant types are present at the same position, both markers will be plotted with a slight jitter, to prevent over plotting.

Below the plot is a grey box. If you click and drag across some points in the plot, this will create a selection box. The data for the selected points will appear in the grey box.

Below the plot is a data table with the data provided for all points on the plot. You can click the "Download Content..." button above the table to download a .csv file of this complete table.

</div><!--panel end------------------------------------------------------------>
<div  class="using-R-insert"><!--panel start----------------------------------->

## Using R
The app uses the function `get_coding_diversity` to calculate the diversity stat at each nucleotide position of the coding sequence, then uses the function `plot_coding_diversity` to plot that information for a single gene.

```{r}
YFG_Coding_Diversity <- llply(YFG_VCF, get_coding_diversity)
lapply(YFG_Coding_Diversity, plot_coding_diversity)

```

</div><!--panel end------------------------------------------------------------>

## Results
Do these values match your hypotheses? Is there anything unexpected?  
In *AFB2* there are five alleles with diversity greater than 1E-2 and 20 alleles with diversity greater than 3E-3. *TIR1* has two alleles with diversity greater than 1E-2 and nine allese with diversity greater than 3E-3. This high number of high frequency alleles in the population suggests that *AFB2* is more likely to be under drift or balancing selection relative to *TIR1* or that *TIR1* is under stronger purifying selection than *AFB2*. It is important to consider however that sampling bias in certain geographic populations could affect these hypotheses. To get some idea of if this is true we will look at the geographic distribution of these alleles.

<!--================== TEMPLATE START =======================================-->
# Mapping the collection location of accessions by alleles   
As an initial visualization of the population structure that may be influencing the allele frequency and prevalence of alleles in a gene, we can map the collection locations from which the accessions were collected. If the accessions sharing a similar polymorphism are all closely clustered then it is likely several members of the same population derived from a common ancestor are represented. If there are many distantly located accessions sharing the same polymorphism this may result from human-associated migration and admixture or possibly from independent mutation events. In addition to these patterns, look for potential environmental gradients of particular alleles (e.g. is a particular allele only found above/below a certain latitude or only in coastal/desert environments?).

<!--=========================================================================-->
<div  class="using-app-insert"><!--panel start--------------------------------->
## SNP mapping
When you first click on the "SNP mapping" panel you will be faced with a choice of which genes you would like to map accessions for. Click in the checkboxes of the gene or genes you would like map variant accessions for and click submit. Now you can use the "Log nucleotide diversity filter limit" slider to set the minimum diversity of the alleles to map. You can also use the "Type of SNP to mark" radio buttons to select the type of SNPs you would like to show on your map.  
![](Screenshots/Tab_3_Select_Genes_and_Filters.PNG)  
This slider and radio buttons will automatically update the list of alleles you can add to your map in the "Allele selection" section. Select any number of alleles and click "Update map" to highlight the location of the accessions containing these alleles on the map.  
![](Screenshots/Tab_3_Mutation_Select.PNG)  
This will reveal the map with accessions containing the selected alleles shown as colored transparent points. All other accessions will be shown as grey points.  
![](Screenshots/Tab_3_Map.PNG)  
You can change the layering of the points, to remove the other (n/a) accessions or to plot one allele on top using the "Layers" button in the bottom left of the map. 
![](Screenshots/Tab_3_Map_w_Layers.PNG)
You can view the full list of accessions in the map along with the alleles those accessions contain directly below the map in the "Map Data" section. You can navigate through this datatable or search for particular alleles using the box at the top right. You can also download the contents of this datatable.
![](Screenshots/Tab_3_Map_Data_Table.PNG)

</div><!--panel end------------------------------------------------------------>
<div  class="using-R-insert"><!--panel start----------------------------------->
## Using R
To repeat the mapping of accessions in R, we will first have to filter by diversity the VCF of mutations. We also want to exclude the reference variants (i.e. when GT is 0|0). We'll use some tools from the [tidyverse](http://tidyverse.org/) to do this. 
```{r}
AFB2_missense <- YFG_VCF$AT3G26810.1 %>% 
  dplyr::filter(Diversity > 3*10^(-3) & gt_GT != "0|0" & 
                  Effect == "missense_variant")
head(AFB2_missense)
```
The accession (as ecotype ID) is in the "Indiv" column of this data frame. Now we will need to attach the accession name, latitude, longitude and other data to this data frame. We'll do this with the `add_ecotype_details` function. The metadata is also available as an included dataset in the r1001genomes package called `accessions`.

```{r}
AFB2_missense <- label_bySNPs(AFB2_missense)
head(AFB2_missense)
```
Now we can plot the variants. In the app we use the Leaflet package to provide an interactive plot. We can also do this within R.
```{r}
AFB2_missense <- rbind(AFB2_missense[is.na(AFB2_missense$SNPs), ],
                       AFB2_missense[!is.na(AFB2_missense$SNPs), ])

    # make a field with text to be displayed when clicking on a marker
    AFB2_missense$popup <- paste("EcoID:",  AFB2_missense$Indiv,"Name:", AFB2_missense$Name, " SNPs:", AFB2_missense$SNPs)

    # create the color pallet for the map points
    pal <- brewer.pal(8, "Set1")
    pallet <- colorFactor(palette=pal, domain=AFB2_missense$SNPs)
map <- leaflet()
    map <- addProviderTiles(map, providers$Stamen.TonerLite,
                     options = providerTileOptions(noWrap = TRUE))

    # groupnames to be used by draw groups of points as separate layers below
    groupnames <- unique(AFB2_missense$SNPs)
    groupnames <- groupnames[!is.na(groupnames)]

    # add markers for NA points first so they are furthest back layer
    map <- addCircleMarkers(map, data=AFB2_missense[is.na(AFB2_missense$SNPs), ], color= "#9b9b9b", group="NA",
                            radius=6, stroke=FALSE, fillOpacity=0.6)

    # for each of the group names, add a set of markers
    for (SNP in groupnames){
          map <- addCircleMarkers(map, data=AFB2_missense[AFB2_missense$SNPs == SNP, ], color= ~pallet(SNPs), group= SNP,
                            radius=6, popup= ~popup, stroke=FALSE, fillOpacity=0.85)
    }

    # add the legend to the map
    map <- addLegend(map, position="bottomright", pal=pallet,
                     values=AFB2_missense$SNPs, title="Marker Colors", opacity=1)

    # add layer control to map to turn on or off groups of points
    map <- addLayersControl(map, overlayGroups=c(groupnames, "NA"),
                            options = layersControlOptions(collapsed = TRUE),
                            position="bottomleft")

    map
```
You can also use the `ggmap` function from the package of the same name to create maps. The "ggrepel" package also helps to handle overlapping labels.

```{r}
europe <- get_map(location = 'europe', zoom = 4)
michigan <- get_map(location = 'michigan', zoom = 5)
library(ggrepel)

ggmap(europe) +  xlab('longitude') + ylab('latitude') + geom_point(data = AFB2_missense, aes(x = Long, y = Lat), alpha = 0.5, shape = 18, color = "darkgrey") + geom_label_repel(data = subset(AFB2_missense, grepl(pattern = "204", AFB2_missense$SNPs)), mapping = aes(label = Name, x = Long, y = Lat), fill = "white", box.padding = unit(.4, "lines"), point.padding = unit(0.5, "lines"), label.padding = unit(.15, "lines"), segment.size = 1, nudge_x = 0.5, nudge_y = 0.5)

ggmap(michigan) +  xlab('longitude') + ylab('latitude') + geom_point(data = AFB2_missense, aes(x = Long, y = Lat), alpha = 0.5, shape = 18, color = "grey") + geom_label_repel(data = subset(AFB2_missense, grepl(pattern = "516", AFB2_missense$SNPs)), mapping = aes(label = Name, x = Long, y = Lat), fill = "white", box.padding = unit(.4, "lines"), point.padding = unit(0.5, "lines"), label.padding = unit(.15, "lines"), segment.size = 1, nudge_x = 0.5, nudge_y = 0.5)

```

</div><!--panel end------------------------------------------------------------>
<!--=========================================================================-->

<!--================== TEMPLATE START =======================================--
# heading   
description

<!--=========================================================================--
<div  class="using-app-insert"><!--panel start---------------------------------
## Tab title
some text and picture

</div><!--panel end------------------------------------------------------------
<div  class="using-R-insert"><!--panel start-----------------------------------
## Using R
some text and code snippet
</div><!--panel end------------------------------------------------------------
<!--=========================================================================-->

### 2017-12-28 r1001genomes/R/VCF_Utils.R
Resolved subtree merge

@@ -541,12 +541,15 @@ add_ecotype_details <- function(data, Ecotype_column="Indiv") {
 #   return(indivGT)
 # }
 
-#' Title
+
+#' Label
 #'
-#' @param data
-#' @param collapse
+#' @param data a data frame of variants with a column named "Indiv"
+#' @param collapse should each accession will be a single line
 #'
-#' @return
+#' @return a df with a single row per accession, with a new column "SNPs" that
+#'  has a single text string detailing all the variants from data within that
+#'  accession
 #' @export
 #'
 #' @examples

### 2017-12-28 r1001genomes/R/Strains_and_Gene_Families.R
Removed r1001genomes subtree, too many conflicts for it to be worth it. r1001genomes will be maintained independently and installed from github.com/wrightrc/r1001genomes

@@ -1,126 +1,0 @@
-### CODE INPUTS ================================================================
-
-strains <-
-  c(88,108,139,159,265,350,351,403,410,424,428,430,470,476,484,504,506,531,544,
-    546,628,630,680,681,685,687,728,742,763,765,766,768,772,801,853,854,867,868,
-    870,915,932,991,992,997,1002,1006,1061,1062,1063,1066,1070,1158,1166,1254,
-    1257,1313,1317,1552,1612,1622,1651,1652,1676,1684,1739,1741,1756,1757,1793,
-    1797,1819,1820,1829,1834,1835,1851,1852,1853,1872,1890,1925,1942,1943,1954,
-    2016,2017,2031,2053,2057,2081,2091,2106,2108,2141,2159,2166,2171,2191,2202,
-    2212,2239,2240,2276,2278,2285,2286,2317,2370,2412,4779,4807,4826,4840,4857,
-    4884,4900,4939,4958,5023,5104,5151,5165,5210,5236,5249,5253,5276,5279,5349,
-    5353,5395,5486,5577,5644,5651,5717,5718,5720,5726,5741,5748,5757,5768,5772,
-    5776,5779,5784,5798,5800,5811,5822,5830,5831,5832,5836,5837,5856,5860,5865,
-    5867,5874,5890,5893,5907,5921,5950,5984,5993,6008,6009,6010,6011,6012,6013,
-    6016,6017,6019,6020,6021,6022,6023,6024,6025,6030,6034,6035,6036,6038,6039,
-    6040,6041,6042,6043,6046,6064,6069,6070,6071,6073,6074,6076,6077,6085,6086,
-    6087,6088,6090,6091,6092,6094,6095,6096,6097,6098,6099,6100,6101,6102,6104,
-    6105,6106,6107,6108,6109,6111,6112,6113,6114,6115,6118,6119,6122,6123,6124,
-    6125,6126,6128,6131,6132,6133,6134,6136,6137,6138,6140,6141,6142,6145,6148,
-    6149,6150,6151,6153,6154,6163,6166,6169,6172,6173,6174,6177,6180,6184,6188,
-    6189,6191,6192,6193,6194,6195,6198,6201,6202,6203,6209,6210,6214,6216,6217,
-    6218,6220,6221,6231,6235,6237,6238,6240,6241,6242,6243,6244,6252,6255,6258,
-    6268,6276,6284,6296,6390,6396,6413,6424,6434,6445,6680,6739,6740,6744,6749,
[…]

### 2017-12-28 r1001genomes/R/VCF_Utils.R

@@ -1,627 +1,0 @@
-### LIBRARIES ==================================================================
-
-#library(VariantAnnotation)
-#library(plyr)
-#library(stringr)
-#library(biomaRt)
-#library(ggplot2)
-#library(reshape2)
-#library(vcfR)
-#library(dplyr)
-#library(ggmap)
-#library(ggthemes)
-
-### APP FUNCTIONS =============================================================
-
-
-#' Run 1001 genomes browser shiny app
-#'
-#' @return Runs the app
-#' @export
-#'
-#' @examples
-run1001genomes <- function() {
-  appDir <- system.file("shiny-app", package = "r1001genomes")
-  if (appDir == "") {
-    stop("Could not find example directory. Try re-installing `mypackage`.",
-         call. = FALSE)
-  }
-
-  shiny::runApp(appDir, display.mode = "normal")
-}
-
-### UTILITY FUNCTIONS =========================================================
-
-
-#' Make a string containing the desired genomic ranges
-#' of the format "chrom:start-stop"
-#' @param data a data.frame with columns `chromosome_name`, `transcript_start`, and `transcript_end`
-#'
-#' @return a vector of strings with the format "chrom:start-stop"
-#' @export
-#'
-#' @examples
-#'
-makeRegionString <- function (data) {
-  # format "chrom:start-stop"
-  return(paste(c(as.character(data$chromosome_name),
-                 ":",
-                 as.character(data$transcript_start),
-                 "-",
-                 as.character(data$transcript_end)), collapse=""))
-}
-
-
-#' Download and save a .VCF file from 1001genomes.org.
-#'
-#' @param fName the file name the downloaded vcf will be saved as eg.
-#'    "data1.vcf"
-#' @param strainStr A character string of comma separated strain or
-#'    "Ecotype ID"s used in the construction of the URL
-#' @param regionStr A character string of the format "[chrom]:[start]-[stop]"
-#'    where [chrom], [start], and [stop] are numbers. this string is used
-#'    directly in the construction of the URL
-#' @param download A logical, if TRUE (default) the file will be downloaded, if
-#'     FALSE, the file will not be downloaded, the URL will still be returned
-#'
-#' @return The URL used for the download
-#' @export
[…]

### 2017-12-30 TUTORIAL.Rmd
intro to Docker and R markdown drafted

---
title: "Tutorial"
author: "R Clay Wright"
date: "12/28/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This tutorial will (hopefully) gently guide you through how to use the tools docker and Rmarkdown to analyze the natural variation in your set of genes and collaborate on our manuscript, tentatively titled A comprehensive analysis of natural sequence variation within the <i>Arabidopsis thaliana</i> nuclear auxin response pathway" (running title: "nuclear auxin response natural variation""). I assume that you have a basic working knowledge of how to download and install software as well as R and RStudio, 
but if you feel you could use a refresher check out [this introduction](https://github.com/wrightrc/Intro_to_R_RStudio_and_ANOVA). We'll use docker to share with you a version of R and Rstudio that already has all of the necessary packages to do the analysis. We'll use Rmarkdown to write the manuscript and make figures and tables by integrating prose, R code and results from the code.

So in this tutorial I will guide you through in as abstracted a way as possible:
1. Installing docker
2. Starting the version of RStudio in docker 
3. Working in docker/RStudio
4. Saving your work in docker/RStudio
5. Using Rmarkdown to format text
6. Using Rmarkdown to insert R code and results into text
7. Citations in Rmarkdown

If you have any troubles with anything, find something unclear or even find a typo please email me and/or make the change in your version of these documents and I will integrate them in the next version.

Ready to get started?

## docker

Docker is a program that allows software configurations to be shared across computers and time. We'll use docker to share the RStudio configuration that will allow you to run the `r1001genomes` package and shiny app without having to upgrade your computer, your R and Rstudio or deal with installing the packages `r1001genomes` depends on. Docker is kind of like a virtual machine that you may have used in the past to run Windows on your Mac or *visa versa*, but docker is better, faster, stronger than virtual machines.

First things first, let's install docker!

### Installing docker

Docker requires a 64bit processor and operating system. For Macs, you can check your processor by clicking the "apple" in the menu bar and selecting "About this Mac". If "Processor" says Core Duo, you unfortunately have a 32 bit processor, otherwise anything else is a new 64bit processor. On a machine running Windows, go to the start menu and settings, then click on "System" and finally "About". If under "System type" you see "32-bit operating system,..." you can't install docker. If you see "32-bit operating system, x64-based processor", you *can* update Windows to 64-bit and then install docker, but this update requires a clean install, meaning all of your files will be deleted.

Anyway, if you have 32-bit woes let me know and we'll come up with a way to allow you to collaborate. I've installed docker on the Macs in the lab so you can always work those.

#### Mac  
If you are using a 2010 or newer Mac with OS X Yosemite (10.10) or newer you can simply [download the Community Edition of docker](https://download.docker.com/mac/stable/Docker.dmg)(click the link to begin download) and install it.

If you are using an older Mac or OS you will need to install the older [docker Toolbox](https://download.docker.com/mac/stable/DockerToolbox.pkg)(click the link to begin download). [Follow these directions](https://docs.docker.com/toolbox/toolbox_install_mac/) to help you get everything installed.

#### Windows  
If you are using Windows 10 Pro, Enterprise or Education you can simply [download the Community Edition of docker](https://download.docker.com/win/stable/Docker%20for%20Windows%20Installer.exe)(click the link to begin download) and install it.

If you are using an older version of Windows you will need to install the older [docker Toolbox](https://download.docker.com/mac/stable/DockerToolbox.pkg)(click the link to begin download). [Follow these directions](https://docs.docker.com/toolbox/toolbox_install_windows/) to help you get everything installed.

### Using docker

Before we jump in let's learn some docker vocabulary. The initial software setup that is a stable snapshot and can be shared is called an "image". Whenever you run that image, you create a "container", which is just an instance of that container. From the docker website:

>An **image** is a lightweight, stand-alone, executable package that includes everything needed to run a piece of software, including the code, a runtime, libraries, environment variables, and config files.  
>  
>A container is a runtime instance of an image—what the image becomes in memory when actually executed. It runs completely isolated from the host environment by default, only accessing host files and ports if configured to do so.

So I have shared with you (and the rest of the world actually) the docker image that will allow you to easily run the r1001genomes package and Shiny app. When you run that image you will create your own container of that image that you can do work in, change settings, build on, save or erase.

Docker has recently built a graphical interface for managing images and containers called kitematic. We'll use this to simplify setting up your container from the image and allowing you to transfer files to and from the container.

#### Installing kitematic

If you click on the docker icon in the menu bar you will see kitematic listed as an option.  
![](TUTORIAL_files/Screenshot 2017-12-29 16.01.58.png)  
Select "Kitematic" to open a window with a link to download and install kitematic. Once you have installed kitematic selecting the "Kitematic" link will open the kitematic interface.  
![](TUTORIAL_files/Screenshot 2017-12-29 16.50.28.png)  
The containers that you currently have running on your system are listed in the left panel. You can see I have a "nuclear-auxin-natural-variation" container running, that docker has given the creative name "distracted-nobel". In the main pane of kitematic there are several popular images that we can create new containers from.

The image that we are going to work with here is called `r1001genomes`. Docker images build from previous images (like all programming does, and science for that matter). I built the `r1001genomes` image on top of a previous image called `verse` from the `rocker` project. The `verse` image contains R, Rstudio, the [tidyverse](www.tidyverse.org) packages, Rmarkdown and all of the dependencies for these things. To this, I added  the necessary packages that `r1001genomes` builds from and `r1001genomes`. Let's make a new container for this image!

You can search in the box at the top of kitematic for "r1001genomes", then click "create" to download the image and create a container. This will take a **while**. You are downloading linux, R, Rstudio and all of those packages. Fortunately docker will reuse all of these components whenever you download a related image.  
![](TUTORIAL_files/Screenshot 2017-12-29 21.08.01.png)

Once the image is downloaded and the container is created you'll see a log and some other information "IP & PORTS" and "VOLUMES".  
![](TUTORIAL_files/Screenshot 2017-12-29 21.13.24.png)

As it says: "You can access this container using the following IP address and port". This means that if you point your browser to the "ACCESS URL" you will see the version of Rstudio that we've built.

Go ahead give it a try!  
![](TUTORIAL_files/Screenshot 2017-12-29 21.34.22.png)  
Now we need some data to play with in our Rstudio container! You'll notice that the only folder available in the "Files" pane is "kitematic", which matches the folder in the "VOLUMES" pane back in kitematic. To make use of this folder we need to map it to a folder on our computer. Click on the "gear" settings icon in the "VOLUMES" pane, then click "CHANGE" and select a folder to be able to access it within your container. Go ahead and select the "natural-variation-webtool" folder that this tutorial is in.

Kitematic will then restart the container and the access URL might be changed. So go back to the "Home" tab and paste the new URL into your browser or, if the access URL didn't change, reload the page.

Now when you click on the "kitematic" folder you will see the contents of this folder. Set the working directory to the "manuscript" folder (using code, or the "More" button). Click on "manuscript.Rmd" to open the manuscript markdown file.

Time for some R markdown!

## R Markdown

The idea behind markdown and R markdown was to make html (hyper-text markUP language) more low key. So the goal of markDOWN, is to be easy to write and easy to read but still have a lot of the powers of HTML (text formatting, layout and templating).

This document was is actually written in R Markdown. If you haven't already, open the "code" markdown version of this document ("TUTORIAL.Rmd") in Rstudio, and navigate down to the "R Markdown" section using the navigation bar at the bottom of the "Source" pane.  
![](TUTORIAL_files/Screenshot 2017-12-29 22.45.46.png)

You can follow along with the code and the document side-by-side to learn how markdown text formatting works. Here's some example markdown from the R markdown file template.

Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

I would highly recommend checking out the "more details" link above and don't miss the [cheatsheet](http://rmarkdown.rstudio.com/lesson-15.html)! You can also find this and other cheatsheets and a Markdown reference in the RStudio "Help" menu.  
![](TUTORIAL_files/Screenshot 2017-12-29 22.57.15.png)

### 2017-12-31 TUTORIAL.Rmd
Finished tutorial

@@
 ---
-title: "Tutorial"
+title: "Collaborative analysis and writing in R"
 author: "R Clay Wright"
@@
 
-This tutorial will (hopefully) gently guide you through how to use the tools docker and Rmarkdown to analyze the natural variation in your set of genes and collaborate on our manuscript, tentatively titled A comprehensive analysis of natural sequence variation within the <i>Arabidopsis thaliana</i> nuclear auxin response pathway" (running title: "nuclear auxin response natural variation"").
+<!--This is a CSS (Cascading Style Sheet).
+It controls how panels and images are formatted.></!-->
+<style>
+  img {
+    display: block;
+    margin: auto;
+    width: 90%;
+    height: auto;
+
+  }
+</style>
+
+This tutorial will (hopefully) gently guide you through how to use the tools docker and Rmarkdown to analyze the natural variation in your set of genes and collaborate on our manuscript, tentatively titled: *A comprehensive analysis of natural sequence variation within the Arabidopsis thaliana nuclear auxin response pathway* (running title: " natural variation in the nuclear auxin response").
 I assume that you have a basic working knowledge of how to download and install software as well as R and RStudio,
@@
 
-So in this tutorial I will guide you through in as abstracted a way as possible:
+So in this tutorial I will guide you through, in as abstracted a way as possible:  
 1. Installing docker  
@@
 If you click on the docker icon in the menu bar you will see kitematic listed as an option.  
-![](TUTORIAL_files/Screenshot 2017-12-29 16.01.58.png)  
+![](TUTORIAL_screenshots/Screenshot 2017-12-29 16.01.58.png)  
 Select "Kitematic" to open a window with a link to download and install kitematic. Once you have installed kitematic selecting the "Kitematic" link will open the kitematic interface.  
-![](TUTORIAL_files/Screenshot 2017-12-29 16.50.28.png)  
+![](TUTORIAL_screenshots/Screenshot 2017-12-29 16.50.28.png)  
 The containers that you currently have running on your system are listed in the left panel.
@@
 Fortunately docker will reuse all of these components whenever you download a related image.
-![](TUTORIAL_files/Screenshot 2017-12-29 21.08.01.png)
+![](TUTORIAL_screenshots/Screenshot 2017-12-29 21.08.01.png)
 
 Once the image is downloaded and the container is created you'll see a log and some other information "IP & PORTS" and "VOLUMES".  
-![](TUTORIAL_files/Screenshot 2017-12-29 21.13.24.png)
+![](TUTORIAL_screenshots/Screenshot 2017-12-29 21.13.24.png)
 
@@
 Go ahead give it a try!  
-![](TUTORIAL_files/Screenshot 2017-12-29 21.34.22.png)  
+![](TUTORIAL_screenshots/Screenshot 2017-12-29 21.34.22.png)  
 Now we need some data to play with in our Rstudio container!
@@
 If you haven't already, open the "code" markdown version of this document ("TUTORIAL.Rmd") in Rstudio, and navigate down to the "R Markdown" section using the navigation bar at the bottom of the "Source" pane.
-![](TUTORIAL_files/Screenshot 2017-12-29 22.45.46.png)
+![](TUTORIAL_screenshots/Screenshot 2017-12-29 22.45.46.png)
 
@@
 
-## Including Plots
+### Including Plots
 
@@
 
-I would highly recommend checking out the "more details" link above and don't miss the [cheatsheet](http://rmarkdown.rstudio.com/lesson-15.html)!
+Don't get bogged down in formatting the text, but I would highly recommend checking out the "more details" link above and don't miss the [cheatsheet](http://rmarkdown.rstudio.com/lesson-15.html)!
 You can also find this and other cheatsheets and a Markdown reference in the RStudio "Help" menu.
-![](TUTORIAL_files/Screenshot 2017-12-29 22.57.15.png)
+Focus foremost should be on the content.
+We'll worry about formatting in revisions.
+![](TUTORIAL_screenshots/Screenshot 2017-12-29 22.57.15.png)
+
+In the manuscript skeleton and the example analysis you can also see some examples of how to use R markdown, DT (data table) and kable to format and insert tables into our document.
+
+### Citations in R markdown  
+You can include citations in markdown documents using the following syntax: `[@citation_key]`.
+The citation key refers to the unique identifier generated when exporting a bibliography from a reference manager.
+For the manuscript you will need to export the references from your reference editor in one of many formats, but I would recommend BibTex (.bib).
+For other formats see [the markdown documentation](http://rmarkdown.rstudio.com/authoring_pandoc_markdown.html#citations).
+
+Let's take a look at an example, open the "references.bib" file in the "manuscript folder".
+You'll notice each reference begins with an @-symbol and the type of reference.
+Then within curly braces are all of the data about the reference.
+The first item is the citation key, which in this case is "gray_identification_1999".
+So to cite this paper in markdown you would add `[@gray_identification_1999]` to the end of a sentence.
+My process for citing-while-writing is to first write, then when I want to cite something, find the paper in my reference manager and copy it to a folder for the project I'm working on.
+This will allow me to export the folder as a .bib file later).
+Then I copy the citation key for the reference and paste it into my writing.
+When I'm ready to knit the document, I'll export the folder of references as a .bib (with Better Bibtex you can sync this file so that it is automatically updated).
+This bibliography file must be in the same folder with the .Rmd file, along with a .csl file that specifies the format of the citations and bibliography in the final document (available for most journals [via Zotero](https://www.zotero.org/styles)).
+These files also must be specified in the header of the document.
+![](TUTORIAL_screenshots/Screenshot 2017-12-30 14.04.14.png)
+
+Most common reference managers allow you to export a bibliography as a bibtex file, and copy and paste citation keys, however I would highly recommend downloading [Zotero](www.zotero.org) and importing your reference library from whatever reference editor you currently use.
+Zotero is free, open-source and integrates very well with all popular web browsers and writing tools.
+If you want to work in a different citation manager, here are some links to help you create and copy and paste citation keys as well as export BibTex bibliography files.
+Note that most of these tutorials are for Latex, which uses a slightly different syntax from Rmarkdown.
+In Rmarkdown insert citations via `[@citation_key]`.
+The `citation_key` string does have some stipulations
+>The citation key must begin with a letter, digit, or \_, and may contain alphanumerics, _, and internal punctuation characters (:.#$%&-+?<>~/).
+
+*To insert multiple citations separate the two citations by semicolons, e.g.* `Blah blah [@smith04; @doe99].`
+
+[Endnote](https://www.reed.edu/cis/help/LaTeX/EndNote.html): Export as [bibtex with labels](http://www.endnote.com/downloads/style/bibtex-export-using-en-label-field).  
+Endnote seems to have some issues generating unique citation keys for each reference.
+It might be best to export your library as an Endnote XML and import the whole thing into Zotero, if you have issues with the above solutions.
+[Mendeley](https://blog.mendeley.com/2011/10/25/howto-use-mendeley-to-create-citations-using-latex-and-bibtex/)  
+[Zotero](https://github.com/bbest/rmarkdown-example#writing-with-rmarkdown)
+
+Once you are done writing and citing and have exported your BibTex file, you can add your bibliography file to the manuscript folder and add the name of your file to the document header (comma separated).
+This should allow your version of the manuscript to knit, with perhaps some errors if there are duplicate citation keys.
+Don't worry we'll sort out duplicates later, once I have merged everyones sections' together.
+
+#### Also print out your key citations and we will make a binder for everyones use in the computer room! Also don't forget to send me any changes or suggestions you have for this tutorial.
+
+Thanks and happy collaborative writing!

### 2017-12-31 manuscript/manuscript.Rmd
Finished tutorial

@@
 title: "A comprehensive analysis of natural sequence variation within the <i>Arabidopsis thaliana</i> nuclear auxin signaling pathway"
-author: "R. Clay Wright"
+author: 
+  affiliation: University of Washington
+  name: "R. Clay Wright"
+author:
+
 date: "`r Sys.Date()`"
+
+output:
+  html_document:
+     fig_caption: yes
+#  word_document:
+#    fig_caption: yes
+#    fig_height: 6
+#    fig_width: 6
+#    highlight: haddock
+keywords: r markdown
+fontsize: 11pt
+fontfamily: Helvetica
 output: 
@@
      number_sections: true
-bibliography: references.bib
+bibliography: ["references.bib"]
 csl: apa.csl
+link-citations: true
 ---
@@
 ```
-
+<!--This is a CSS (Cascading Style Sheet).
+It controls how panels and images are formatted.></!-->
 <style>
-  .using-app-insert {
-     background-color: #bfbfbf;
-     padding: 15px;
-     border-style: solid;
-     border-color:  #8c8c8c;
-     border-radius: 7px 7px 0 0;
-     border-width: 1px;
-  }
-
-  .using-R-insert {
-     background-color: #e6e6e6;
-     padding: 15px;
-     border-style: solid;
-     border-color:  #8c8c8c;
-     border-radius: 0 0 7px 7px;
-     border-width: 1px;
-     border-top: 0;
-  }
-
-  .using-app-insert h2, .using-app-insert h3, .using-app-insert h4 {
-     border-bottom: 0;
-  }
-
-  .using-R-insert h2, .using-R-insert h3, .using-R-insert h4 {
-     border-bottom: 0;
-  }
-
   img {
@@
     height: auto;
-
   }
@@
 
-# Introduction
+#List of working titles  
+You should keep a list of titles and constantly be refining and tweaking it. A title helps focus you. The title should be in the same style as papers you plan to cite.
+You might also think of snappy titles or sub titles for press releases. Find some press releases for paper similar to yours to see how journalists interpret the results.
 
-This web app and open source software package for the R programming language was built to facilitate exploration of the naturally occuring sequence diversity for your favorite gene or gene family.
-We hope this product will help plant biologists without bioinformatics experience access the 1001 genomes project’s rich catalog of natural variation, formulate hypotheses and identify new interesting alleles in their favorite gene family.
+#Abstract  
+Your abstract should be in the following form, more or less exactly:
 
-The 1001 genomes dataset was collected with the intention of genome scale analyses, such as genome wide association studies, in which a particular phenotype is quantified for all of the accessions and sequence variants that are associated with phenotypic variation are identified.
-[In recent work](https://doi.org/10.1534/genetics.117.3000), we have utilized this rich dataset in a different manner to measure the effects of coding sequence variation on protein functional variation and phenotypic variation .
-This work has expanded the sequence-function-phenotype map for the auxin-signaling F-box gene family, and this product will facilitate similar analysis other genes, gene families, and genomic regions.
+* Motivating, high level statement  
+* Statement defining the cutting edge of your research area  
+* A statement of the form: "However, it remains unclear ..." that defines what problem you are  
+addressing.  
+* A statement of the form:"Here, we ..."  that concisely states your contribution.  
+* A series of 3?5 statement about your results.  
+* A concluding statement.  
+When presenting your abstract to your advisor, these sentences should be clearly laid out, in a bullet list preferably.
 
-So pick out your favorite gene, gene family or genomic region and follow the steps in this guide to analyze how it varies in plants collected from all over the world.
-For this example we will be using the auxin-signaling F-box genes, *TIR1* and *AFB2* which have the TAIR locus IDs [AT3G62980](https://www.arabidopsis.org/servlets/TairObject?id=39952&type=locus) and [AT3G26810](https://www.arabidopsis.org/servlets/TairObject?id=38086&type=locus).
-Click on one of the previous links to search TAIR for the TAIR IDs of your favorite genes.
+#Related work  
+You need a list of about 30 papers that are the most relevant.
+They should be weighted toward recent, well cited work.
+You should group them into two to four areas.
+For each paper, you should have a sentence that says what the paper is about and how it relates to your work.
 
-***
-Before digging into the data, make some hypotheses about what genes, domains and/or postions in your 
-favorite gene family might be under different evolutionary pressures.
-We can make some inferences about the evolutionary pressure acting on particular genes based on the sequence and geographic diversity observed in the 1001 genomes dataset.
-Thomas Mitchell-Olds and Johanna Schmitt published an excellent overview of the evolutionary inference from natural genetic diversity in *A.
-thaliana* [@mitchell-olds_genetic_2006].
-We can calculate genetic diversity as the average number of nucleotide differences per site between two DNA sequences in all possible pairs in the sample population, and is denoted by $\pi$.
+#Glossary  
+Make a list of all of the new terms that you have invented to talk about different parts of your work.
+This list will eventually be put into paragraph format, but for now it should look like a glossary.
+You can also add terms that from the literature that you will use in case you want to make sure that the reader (or you or your advisor) really knows what you mean when you use that term.
 
-However, evolutionary pressures potentially exert influence across all levels of biology, nucleotides, genes, genomes, phenotypes, organisms and populations.
-This makes any inference from study of genetic diversity contingent on potentially unknown effects of epistasis between genes, population structure, environment, and many other effectors.
-Additionally, as an agricultural tag-along, much of the genetic variation in *A.
-thaliana* is due to admixture between populations as the result of human influence migration.
-So keep in mind throughout this analysis the limitations of examining a single snapshot in time of a potentially biased collection of frequently self-fertizilizing organisms that were dispersed by human migration.
-Because of these limitations, infering theories about how evolutionary forces are acting on your favorite genes is very uncertain business.
-However, if we assume a few things we can make some speculations.
+#Results  
+This should be a list of 3-5  statements  that very exactly say what you are claiming as results, and then the support for that claim.
+Statements will be then supported by experimental, theoretical, or numerical evidence.
 
-Assumptions:  
-1. Polymorphisms in the dataset are not *completely* detrimental to organismal fitness.  
-2. Most nonsynonymous polymorphisms in a gene result in a reduction in function.
+Note that it is very important that you do not use mere titles for the sections of your paper.
+If you do, you end up with a laundry list.
+So do not say "transcription initiation" and list all of the things you know about the subject.
+Instead say, say "Transcription initiation is regulated by proteins that bind to or near promoters", and then list the evidence you have (or plan to get) that says why you suspect the sentence to be true.
 
-Now, taking a gene-centric view of evolution, here are some of the evolutionary pressures, or types of natural selection, that we might observe:
+Another way to think of these statements is that they are essentially your hypotheses, if you are a scientist, or they are statements of capabilities if you are an engineer.
 
-### Purifying selection  
-If a gene's sequence is highly conserved (low diversity) it may be under [purifying/negative/stabilizing selection](https://en.wikipedia.org/wiki/Negative_selection_(natural_selection)). 
-If a genes' function is crucial to the fitness of the organism, and most polymorphisms result in loss of function, we would expect a low level of nonsynonymous polymorphism in this gene.
-This is because deleterious polymorphisms that do arise in this gene within the population will be outcompeted by the more-fit parent.
-This results in the decreased genetic diversity in genes under purifying selection.
+#Figures and captions
 
-### Genetic Drift
-If evolutionary pressure is very weak on a gene its sequence might change with no effect on organismal fitness.
-This allows a gene to accumulate many mutations that may persist in the population or be lost at random.
-This results in random fluctuations in allele frequencies, which when combined with an appreciable rate of mutation and a sufficient population size yields increased genetic diversity in the population.
+You should have mock ups of three to five figures (with sub?figures) that you plan to get the data for.
+These figures are incredibly important to think about from an early stage.
+For example, if you know that you will have a certain data figure with certain x and y axes and certain error bars, then you have a specification of what data you will need to generate and what it ought to look like.
 
-Beyond purifying selection and drift it is very difficult to make any strong inferences on the population as a whole.
-However there are two more types of selective pressure, balancing or neutral selection and positive or directional selection.
+You should also have working drafts of all of the captions for each sub figure. Captions have  titles and then descriptions.
 
-### Balancing selection
-Genes of high diversity may be under balancing selection, also called neutral divergence.
-Balancing selection maintains multiple variants for long evolutionary times because different alleles provide selective advantages in different environments or genetic backgrounds.
-This results in high diversity at the particular site under selection and increased diversity at closely linked neutral sites due to genetic hitchhiking with this allele under balancing selction.
+Finally, you should know what glossary terms or results statements the figures correspond to and roughly how you will use them in your paper.
 
-### Directional selection  
-If a polymorphism (or allele) is associated with a phenotype that provides an organims with a fitness advantage this phenotype and associated allele may be experiencing directional or positive selection.
-Positive/directional selection results in an increase in the prevalence of the advantageous allele within the population over time.
-Directional selection is one of the most difficult things to infer and is a very active area of research.
-Directional selection, like stabilizing selection is associated with reduced genetic diversity within the population under selection.
-However, unlike stabilizing selection the gene under directional selection will potentially have two major alleles that provide adaptive potential to different environments or genetic backgrounds.
+#List of things to do
 
-# Hypothesize
-Now go read/collect the literature about your family and make some hypotheses!  
-Here are some questions to help you get started:
+Keep a running list of everything you need to do to finish the paper.
+Don’t put anything extraneous on the list.
+Make sure every unfinished result has a corresponding sub?list of things to do.
+Make sure that there are always items on your list that you can and will do t  his week,  as well as longer term items that help you plan ahead.
 
-What structure/function knowledge is there?  
-What is each domain in the protein doing?  
-Are there any postions that are of vital importance?
-
-Is there a published protein sequence alignment of your gene family?  
-What postions in the protein family are highly conserved?  
-What postions are highly variable?
-
-Is there a published phylogenetic tree of your gene family?  
-What are the most recently duplicated sister gene pairs?  
-What do you know about the phenotypes of mutants in these pairs?  
-Do the genetics suggest these pairs are redundant?
-
-## For example
-In TIR1 and AFB2 there are 2 domains.
-The F-box domain interacts with the SCF components [@gray_identification_1999], and the LRR domain binds auxin and Aux/IAA proteins [@kepinski_auxininduced_2004].
-The F-box is roughly 40 amino acids and wraps tightly around the SKP1 component of the SCF complex.
-The LRR is greater than 500 amino acids and binds auxin in a pocket formed by a horseshoe shaped chain of LRRs [@tan_mechanism_2007].
-The outside of this horseshoe shape does not have any known interactions other than a dimerization interface [@dezfulian_oligomerization_2016].
-Because the F-box domain must interact with the SCF complex for TIR1/AFB2 to perform their known function, we might expect the F-box to be highly conserved.
-The small size of the F-box also helps strengthen this hypothesis.
-The LRR domain is much larger and may have differing specificities for Aux/IAAs, so we might expect it to be more diverse.
-We would also expect the residues of the LRR that contact auxin and the Aux/IAA degron to be highly conserved.
-
-***
-
-# A note on the format of this document
-<!--=========================================================================-->
-<div  class="using-app-insert"><!--panel start--------------------------------->
-## Using the App
-
-In addition to calling the functions of the r1001genomes package in R directly.
-there is also a GUI provided in the form of a Shiny app.
-This app can perform some of the analysis tasks the package is capable of performing, without the need to have knowledge of programming in R.
-
-In this vignette panels like this one are used to show examples of using the app to achieve similar results as using  the R function calls directly in this vignette.
-Screenshots will usually be provided to make it clear what section or tab we are talking about
-
-```{r, echo = FALSE}
-knitr::include_graphics("Screenshots/App_Initial.PNG")
-knitr::asis_output("
-                   ") # This newline required for headings to render after image
-```
-## Starting the web app
-
-to start the Shiny app, run the following command:
-```{r, eval=FALSE}
-run1001genomes()
-```
-The app should appear in a new window, and look similar to the image above.
-</div><!--panel end------------------------------------------------------------>
-<div  class="using-R-insert"><!--panel start----------------------------------->
-## Using R
-in these lighter colored panels, the equivalent R code will be displayed.
-</div><!--panel end------------------------------------------------------------>
-<!--=========================================================================-->
-
-# Download the chromosome positions and gene models 
-The first step in retrieving the natural variation in our genes is to retrieve the information about where the genes are in the genome of Col-0, the laboratory reference accession, and how the mRNA and coding sequence is laid out.
-This information is called the gene model.
-
-<!--=========================================================================-->
-<div  class="using-app-insert"><!--panel start--------------------------------->
-## Loading Gene Data in the App
-```{r, echo=FALSE}
-knitr::include_graphics("Screenshots/Tab_1_Select_Genes.PNG")
-knitr::asis_output("
-                   ") # This newline required for headings to render after image
-```
-### Selecting Genes and Generating the Gene Model Table
-At the top of the first tab (the "SNP Stats" tab) is a text input field where you can input a list of TAIR loci.
-
-The app will download the gene model information as soon as you click submit. Copy and paste your TAIR loci in the box, separated by commas, and click submit.
-
-The app may take a little while to process after clicking the submit button on the first tab, especially if you have a long list of genes, but once this process is complete on the first tab, the data for these genes will be already loaded for the subsequent tabs.
-
-Once the loading process has finished, the first table you should see scrolling down the SNP Stats tab is the Gene Information table.
-it should look similar to the table produced in the "Using R" section below.
-</div><!--panel end------------------------------------------------------------>
-<div  class="using-R-insert"><!--panel start----------------------------------->
-## Using R
-The code the app uses to download this table is the `getGeneInfo` function.
-This function is just a simplifying wrapper for the `BioMart` function `getBM`.
-This downloads the gene model from ENSEMBL plants.
-
-```{r}
-tair_ids <- c("AT3G62980", "AT3G26810")
-geneInfo <- getGeneInfo(genes = tair_ids, firstOnly = TRUE, inputType = "tair_locus", useCache = TRUE)
-```
-
-Now we can give this table a look. We can also freshen up the table a bit with `datatable`. 
-```{r}
-datatable(geneInfo[, -c(5,6,9)], colnames = c("tair locus", "symbol", "transcript", 
-                                              "Chr", "transcript \nstart", "transcript \nend", 
-                                              "transcript \nlength"), rownames = FALSE, 
-          options=list(paging=FALSE, searching=FALSE))
-```
-</div><!--panel end------------------------------------------------------------>
-<!--=========================================================================-->
-
-Do you have a spreasheet of loci?
-
-Here is some quick code to make a text string you can copy and paste into the app gene input box.  
-```{r}
-AFB_loci <- read.csv(system.file("extdata", "AFB_gene_ids.csv", package = "r1001genomes"))
-paste(AFB_loci$tair_locus, collapse = ",")
-```
-
-Occasionally there are multiple models for a single gene if there are multiple possible transcripts or splice variants. Does your gene have multiple models?
-
-# Download the variants in the 1001 genomes dataset for these genes
-The 1001 genomes dataset is stored as a VCF (Variant Call Format) file. This is 
-a very efficient format for storing such a large dataset (the 1135 genomes are ~17gb), 
-and for some analyses its matrix format is very useful.
-
-<!--=========================================================================-->
-<div  class="using-app-insert"><!--panel start--------------------------------->
-## Downloading and processing of VCFs in the App
-Upon clicking the "Submit" button on the "SNP Stats" tab, the web app automatically creates a tidy dataframe and calculates the number of unique and nonunique polymorphisms in each region of the gene model as well as calculating the diversity in the protein coding sequence.
-The diversity is the average number of nucleotide differences per site between all possible pairs of sequences.
-Diversity is also calculated for only the set of synonymous and nonsynonymous polymorphisms.
-
-The results of these calculations are displayed in the three tables below the Gene Information table.
-
-</div><!--panel end------------------------------------------------------------>
-<div  class="using-R-insert"><!--panel start----------------------------------->
-## Using R
-We can download a set of VCF files for our favorite genes using the `VCFList` function.
-
-```{r, eval=FALSE}
-YFG_VCF <- VCFList(geneInfo = geneInfo, by = "transcript", tidy = FALSE)
-```
-
-However, for some filtering, summarizing and plotting operations it is more efficient to 
-convert the matrix into a very long data frame. Most of the functions in `r1001genomes` 
-operate on a long data frame, often called a tidy data frame. We can automatically covert 
-our VCF to a tidy data frame by setting the `tidy` argument to `TRUE`.
-
-```{r, cache=TRUE}
-YFG_VCF <- VCFList(geneInfo = geneInfo, by = "transcript", tidy = TRUE)
-```
-
-We now have a list of very long data frames with all the information on SNPs inside, but much of the details on the effects of these polymorphisms are contained in a single column labeld 'EFF' as a long text string, making it hard to filter and process the data by the individual attributes within that text string.
-the `parseEFF` function will separate indiual fields of the EFF string into separate columns on a single dataframe, and can be applied to our list of data frames with the `llply` function from the plyr package
-```{r}
-YFG_VCF <- llply(YFG_VCF, parseEFF)
-```
-
-To generate the three tables of polymorphism stats the app calls the function `polymorphRow`. in R we can do the same thing.
-```{r, cache=TRUE}
-#TIR1_AFB2_table <- polymorphTable(geneInfo = geneInfo, strains = strains)
-YFG_VCF <- llply(YFG_VCF, Nucleotide_diversity)
-TIR1_AFB2_table <- ldply(YFG_VCF, polymorphRow, geneInfo = geneInfo, .id = "Transcript_ID")
-
-```
-
-Before displaing this table we will again clean it up a bit.
-
-```{r}
-formatRound(datatable(TIR1_AFB2_table[,c(1:6,8)],
-                  colnames = c("transcript", "5' UTR", "intron", "3' UTR",
-                               "coding \n synonymous", "coding \n missense",
-                               "coding \n total"), 
-                  escape = FALSE, rownames = FALSE,
-    options=list(paging=FALSE, searching=FALSE)), columns = 2:7, digits = 0)
-```
-The nucleotide diversity for the chosen genes is also calculated in the call to `polymorphTable`.
-
-```{r}
-formatRound(datatable(TIR1_AFB2_table[, c(1,13,9:12)],
-                  colnames = c("transcript",
-                               "&pi; transcript",
-                               "&pi; coding",
-                               "&pi;<sub>N</sub>",
-                               "&pi;<sub>S</sub>",
-                               "&pi;<sub>N</sub>/&pi;<sub>S</sub>"), 
-                  escape = FALSE,
-                  options = list(paging=FALSE, searching=FALSE)),
-                columns = 1:6, digits = 6)
-```
-
-</div><!--panel end------------------------------------------------------------>
-<!--=========================================================================-->
-
-## Results
-Do these values match your hypotheses? Is there anything unexpected?
-
-For our TIR1 and AFB2 example, we can make several interesting observations based on this dataset.  
-1. *AFB2* had higher number of total polymorphisms than *TIR1*, across all subsets of sites except for introns. 
-2. *AFB2* also has a higher total coding diversity and higher ratio of diversities at nonsynonymous to synonymous sites.
-
-This was expected, as *TIR1* mutants have a strong loss of auxin sensitivity whereas *AFB2* mutants have a much weaker effect.
-Therefore, polymorphisms in *TIR1* are much more likely to have an effect on phenotype than polymorphisms in *AFB2*.
-
-# Diversity within each gene
-
-Now we'll move on to the second "Diversity Plot" tab.
-This tab will calculate the diversity at each position within the coding sequence of each gene.
-In addition to calculating the nucleotide diversity of a whole sequence, like the transcipt of a gene, we can also calculate the diversity statistic for each position along the reference genome.
-This could be useful if you suspect a particular region of your gene may be more important to function or more sensitive to nucelotide or amino acid changes.
-
-<!--=========================================================================-->
-<div  class="using-app-insert"><!--panel start--------------------------------->
-## Diversity Plot
-The second tab of the app, labeld "Diversity Plot" allows you to plot the nucleotide diversity for each of the genes you selected in the "SNP Stats" tab.  
-```{r}
-imagePath <- "Screenshots/Tab_2_Select_Gene.PNG"
-knitr::include_graphics(imagePath)
-knitr::asis_output("
-                   ") # This newline required for headings to render after image
-```
-After clicking to the "Diversity Plot" tab, select the first gene you would like to plot then click the submit button.  
-![](Screenshots/Tab_2_Plot_Selection_Box.PNG)
-
-You should see a plot like the one above.
-The plot shows Nei's nucleotide diversity by *nucleotide position* of the protein coding region of the gene, though the X-axis is marked by codon number from the N-terminus for convenience.
-The symbols are differentiated by whether the site has a synonymous or missense variant.
-If multiple variant types are present at the same position, both markers will be plotted with a slight jitter, to prevent over plotting.
-
-Below the plot is a grey box. If you click and drag across some points in the plot, this will create a selection box. The data for the selected points will appear in the grey box.
-
-Below the plot is a data table with the data provided for all points on the plot. You can click the "Download Content..." button above the table to download a .csv file of this complete table.
-
-</div><!--panel end------------------------------------------------------------>
-<div  class="using-R-insert"><!--panel start----------------------------------->
-
-## Using R
-The app uses the function `get_coding_diversity` to calculate the diversity stat at each nucleotide position of the coding sequence, then uses the function `plot_coding_diversity` to plot that information for a single gene.
-
-```{r}
-YFG_Coding_Diversity <- llply(YFG_VCF, get_coding_diversity)
-lapply(YFG_Coding_Diversity, plot_coding_diversity)
-
-```
-
-</div><!--panel end------------------------------------------------------------>
-
-## Results
-Do these values match your hypotheses? Is there anything unexpected?  
-In *AFB2* there are five alleles with diversity greater than 1E-2 and 20 alleles with diversity greater than 3E-3.
-*TIR1* has two alleles with diversity greater than 1E-2 and nine allese with diversity greater than 3E-3.
-This high number of high frequency alleles in the population suggests that *AFB2* is more likely to be under drift or balancing selection relative to *TIR1* or that *TIR1* is under stronger purifying selection than *AFB2*.
-It is important to consider however that sampling bias in certain geographic populations could affect these hypotheses.
-To get some idea of if this is true we will look at the geographic distribution of these alleles.
-
-<!--================== TEMPLATE START =======================================-->
-# Mapping the collection location of accessions by alleles   
-As an initial visualization of the population structure that may be influencing the allele frequency and prevalence of alleles in a gene, we can map the collection locations from which the accessions were collected.
-If the accessions sharing a similar polymorphism are all closely clustered then it is likely several members of the same population derived from a common ancestor are represented.
-If there are many distantly located accessions sharing the same polymorphism this may result from human-associated migration and admixture or possibly from independent mutation events.
-In addition to these patterns, look for potential environmental gradients of particular alleles (e.g.
-is a particular allele only found above/below a certain latitude or only in coastal/desert environments?).
-
-<!--=========================================================================-->
-<div  class="using-app-insert"><!--panel start--------------------------------->
-## SNP mapping
-When you first click on the "SNP mapping" panel you will be faced with a choice of which genes you would like to map accessions for.
-Click in the checkboxes of the gene or genes you would like map variant accessions for and click submit.
-Now you can use the "Log nucleotide diversity filter limit" slider to set the minimum diversity of the alleles to map.
-You can also use the "Type of SNP to mark" radio buttons to select the type of SNPs you would like to show on your map.
-![](Screenshots/Tab_3_Select_Genes_and_Filters.PNG)  
-This slider and radio buttons will automatically update the list of alleles you can add to your map in the "Allele selection" section.
-Select any number of alleles and click "Update map" to highlight the location of the accessions containing these alleles on the map.
-![](Screenshots/Tab_3_Mutation_Select.PNG)  
-This will reveal the map with accessions containing the selected alleles shown as colored transparent points. All other accessions will be shown as grey points.  
-![](Screenshots/Tab_3_Map.PNG)  
-You can change the layering of the points, to remove the other (n/a) accessions or to plot one allele on top using the "Layers" button in the bottom left of the map. 
-![](Screenshots/Tab_3_Map_w_Layers.PNG)
-You can view the full list of accessions in the map along with the alleles those accessions contain directly below the map in the "Map Data" section.
-You can navigate through this datatable or search for particular alleles using the box at the top right.
-You can also download the contents of this datatable.
-![](Screenshots/Tab_3_Map_Data_Table.PNG)
-
-</div><!--panel end------------------------------------------------------------>
-<div  class="using-R-insert"><!--panel start----------------------------------->
-## Using R
-To repeat the mapping of accessions in R, we will first have to filter by diversity the VCF of mutations.
-We also want to exclude the reference variants (i.e.
-when GT is 0|0).
-We'll use some tools from the [tidyverse](http://tidyverse.org/) to do this.
-```{r}
-AFB2_missense <- YFG_VCF$AT3G26810.1 %>% 
-  dplyr::filter(Diversity > 3*10^(-3) & gt_GT != "0|0" & 
-                  Effect == "missense_variant")
-head(AFB2_missense)
-```
-The accession (as ecotype ID) is in the "Indiv" column of this data frame.
-Now we will need to attach the accession name, latitude, longitude and other data to this data frame.
-We'll do this with the `add_ecotype_details` function.
-The metadata is also available as an included dataset in the r1001genomes package called `accessions`.
-
-```{r}
-AFB2_missense <- label_bySNPs(AFB2_missense)
-head(AFB2_missense)
-```
-Now we can plot the variants. In the app we use the Leaflet package to provide an interactive plot. We can also do this within R.
-```{r}
-AFB2_missense <- rbind(AFB2_missense[is.na(AFB2_missense$SNPs), ],
-                       AFB2_missense[!is.na(AFB2_missense$SNPs), ])
-
-    # make a field with text to be displayed when clicking on a marker
-    AFB2_missense$popup <- paste("EcoID:",  AFB2_missense$Indiv,"Name:", AFB2_missense$Name, " SNPs:", AFB2_missense$SNPs)
-
-    # create the color pallet for the map points
-    pal <- brewer.pal(8, "Set1")
-    pallet <- colorFactor(palette=pal, domain=AFB2_missense$SNPs)
-map <- leaflet()
-    map <- addProviderTiles(map, providers$Stamen.TonerLite,
-                     options = providerTileOptions(noWrap = TRUE))
-
-    # groupnames to be used by draw groups of points as separate layers below
-    groupnames <- unique(AFB2_missense$SNPs)
-    groupnames <- groupnames[!is.na(groupnames)]
-
-    # add markers for NA points first so they are furthest back layer
-    map <- addCircleMarkers(map, data=AFB2_missense[is.na(AFB2_missense$SNPs), ], color= "#9b9b9b", group="NA",
-                            radius=6, stroke=FALSE, fillOpacity=0.6)
-
-    # for each of the group names, add a set of markers
-    for (SNP in groupnames){
-          map <- addCircleMarkers(map, data=AFB2_missense[AFB2_missense$SNPs == SNP, ], color= ~pallet(SNPs), group= SNP,
-                            radius=6, popup= ~popup, stroke=FALSE, fillOpacity=0.85)
-    }
-
-    # add the legend to the map
-    map <- addLegend(map, position="bottomright", pal=pallet,
-                     values=AFB2_missense$SNPs, title="Marker Colors", opacity=1)
-
-    # add layer control to map to turn on or off groups of points
-    map <- addLayersControl(map, overlayGroups=c(groupnames, "NA"),
-                            options = layersControlOptions(collapsed = TRUE),
-                            position="bottomleft")
-
-    map
-```
-You can also use the `ggmap` function from the package of the same name to create maps. The "ggrepel" package also helps to handle overlapping labels.
-
-```{r}
-europe <- get_map(location = 'europe', zoom = 4)
-michigan <- get_map(location = 'michigan', zoom = 5)
-library(ggrepel)
-
-ggmap(europe) +  xlab('longitude') + ylab('latitude') + geom_point(data = AFB2_missense, aes(x = Long, y = Lat), alpha = 0.5, shape = 18, color = "darkgrey") + geom_label_repel(data = subset(AFB2_missense, grepl(pattern = "204", AFB2_missense$SNPs)), mapping = aes(label = Name, x = Long, y = Lat), fill = "white", box.padding = unit(.4, "lines"), point.padding = unit(0.5, "lines"), label.padding = unit(.15, "lines"), segment.size = 1, nudge_x = 0.5, nudge_y = 0.5)
-
-ggmap(michigan) +  xlab('longitude') + ylab('latitude') + geom_point(data = AFB2_missense, aes(x = Long, y = Lat), alpha = 0.5, shape = 18, color = "grey") + geom_label_repel(data = subset(AFB2_missense, grepl(pattern = "516", AFB2_missense$SNPs)), mapping = aes(label = Name, x = Long, y = Lat), fill = "white", box.padding = unit(.4, "lines"), point.padding = unit(0.5, "lines"), label.padding = unit(.15, "lines"), segment.size = 1, nudge_x = 0.5, nudge_y = 0.5)
-
-```
-
-</div><!--panel end------------------------------------------------------------>
-<!--=========================================================================-->
-
-<!--================== TEMPLATE START =======================================--
-# heading   
-description
-
-<!--=========================================================================--
-<div  class="using-app-insert"><!--panel start---------------------------------
-## Tab title
-some text and picture
-
-</div><!--panel end------------------------------------------------------------
-<div  class="using-R-insert"><!--panel start-----------------------------------
-## Using R
-some text and code snippet
-</div><!--panel end------------------------------------------------------------
-<!--=========================================================================-->
+# Abstract

### 2018-01-01 TUTORIAL.Rmd
edited tutorial, made manuscript skeleton

@@
 date: "12/28/2017"
-output: html_document
+output: 
+  html_document:
+    toc: true
+
 ---
@@
 
-This tutorial will (hopefully) gently guide you through how to use the tools docker and Rmarkdown to analyze the natural variation in your set of genes and collaborate on our manuscript, tentatively titled: *A comprehensive analysis of natural sequence variation within the Arabidopsis thaliana nuclear auxin response pathway* (running title: " natural variation in the nuclear auxin response").
+This tutorial will (hopefully) gently guide you through how to use the tools docker and Rmarkdown to analyze the natural variation in your set of genes and collaborate on our manuscript, tentatively titled: *A comprehensive analysis of natural sequence variation within the Arabidopsis thaliana nuclear auxin response pathway* (running title: "natural variation in the nuclear auxin response").
 I assume that you have a basic working knowledge of how to download and install software as well as R and RStudio,
@@
 
-So in this tutorial I will guide you through, in as abstracted a way as possible:  
+So in this tutorial I will guide you through:  
 1. Installing docker  
-2. Starting the version of RStudio in docker  
+2. Starting RStudio and r1001genomes in docker  
 3. Working in docker/RStudio  
@@
 
-## docker
+## Docker
 
@@
 >  
->A container is a runtime instance of an image—what the image becomes in memory when actually executed.
+>A **container** is a runtime instance of an image—what the image becomes in memory when actually executed.
 It runs completely isolated from the host environment by default, only accessing host files and ports if configured to do so.
@@
 The image that we are going to work with here is called `r1001genomes`.
-Docker images build from previous images (like all programming does, and science for that matter).
+Docker images build on previous images (like all programming does, and all science does, for that matter).
 I built the `r1001genomes` image on top of a previous image called `verse` from the `rocker` project.
-The `verse` image contains R, Rstudio, the [tidyverse](www.tidyverse.org) packages, Rmarkdown and all of the dependencies for these things.
+(The rocker project and R Open Science also have a [great docker tutorial](http://ropenscilabs.github.io/r-docker-tutorial/).) The `verse` image contains R, Rstudio, the [tidyverse](www.tidyverse.org) packages, Rmarkdown and all of the dependencies for these things.
 To this, I added  the necessary packages that `r1001genomes` builds from and `r1001genomes`.
@@
 Kitematic will then restart the container and the access URL might be changed.
-So go back to the "Home" tab and paste the new URL into your browser or, if the access URL didn't change, reload the page.
+So go back to the "Home" tab and paste the new URL into your browser or, if the access URL didn't change just reload the page.
 
 Now when you click on the "kitematic" folder you will see the contents of this folder.
-Set the working directory to the "manuscript" folder (using code, or the "More" button).
+Any files you save in this folder will be saved to the folder you mapped on your computer.
+You can also open files from this folder in RStudio.
+In the RStudio "Files" pane, navigate to the "manuscript" folder and set it as your working directory (using the "More" button).
 Click on "manuscript.Rmd" to open the manuscript markdown file.
@@
 
-The idea behind markdown and R markdown was to make html (hyper-text markUP language) more low key.
-So the goal of markDOWN, is to be easy to write and easy to read but still have a lot of the powers of HTML (text formatting, layout and templating).
+The idea behind the creation of markdown and R markdown was to make html (hyper-text markUP language) more low key.
+So the goals of markDOWN are to be easy to write and easy to read but to still have a lot of the powers of HTML (text formatting, layout and templating).
+R Markdown also allows us to generate PDF, Word, and Presentation documents in addition to HTML.
 
 This document was is actually written in R Markdown.
-If you haven't already, open the "code" markdown version of this document ("TUTORIAL.Rmd") in Rstudio, and navigate down to the "R Markdown" section using the navigation bar at the bottom of the "Source" pane.
+If you haven't already, open the "code" markdown version of this document ("TUTORIAL.Rmd") in Rstudio, and navigate down to the "R Markdown" section using the navigation bar at the bottom of the "Source" pane (see the figure).
 ![](TUTORIAL_screenshots/Screenshot 2017-12-29 22.45.46.png)
@@
 
-### Including Plots
+You can insert code chunks easily using the "Insert" button at the top of the source window, Code>Insert Chunk in the menu bar or hotkeys specified next to this menu bar command.
+
+### Including Plots and Code
 
@@
 
-Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
+Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot (check out the markdown code in TUTORIAL.Rmd).
 
-Don't get bogged down in formatting the text, but I would highly recommend checking out the "more details" link above and don't miss the [cheatsheet](http://rmarkdown.rstudio.com/lesson-15.html)!
+You can also incorporate results from code chunks in your prose. First, we need a result .
+
+```{r}
+set.seed(42)
+group1 <- rnorm(n = 10, mean = 1, sd = 0.5)
+group2 <- rnorm(n = 10, mean = 2, sd = 0.5)
+t_test <- t.test(x = group1, y = group2)
+t_test
+```
+
+Now we can refer to these results in our code by creating an inline code chunk with syntax `r #code`.
+So we could say that we have some support that the means of the two groups we simulated are different (`r t_test$method`, t-value = `r t_test$statistic`, p = `r t_test$p.value`)
+
+Don't get bogged down in formatting the text, but I would highly recommend checking out the <http://rmarkdown.rstudio.com> link and don't miss the [cheatsheet](http://rmarkdown.rstudio.com/lesson-15.html)!
 You can also find this and other cheatsheets and a Markdown reference in the RStudio "Help" menu.
-Focus foremost should be on the content.
-We'll worry about formatting in revisions.
 ![](TUTORIAL_screenshots/Screenshot 2017-12-29 22.57.15.png)
 
-In the manuscript skeleton and the example analysis you can also see some examples of how to use R markdown, DT (data table) and kable to format and insert tables into our document.
+A few tips to help your R Markdown experience stay low key:  
+1. Turn on whitespace characters so you know where spaces, tabs and returns are. This is in Preferences>Code>Display.  
+2. Include 2 spaces at the end of each paragraph or whenever you want to start a new line (like in this numbered list).  
+3.
+Use and empty line between paragraphs, headings, offset quotes and any other sort of formatted elements like lists, tables and code chunks.
+This offset helps improve readability and makes sure that these elements are formatted the way you want them.
+
+**However, keep in mind your foremost focus should be on the content. We'll worry about formatting in revisions.**
+
+In the example analysis you can also see some examples of how to use R markdown, DT (data table) and kable to format and insert tables into our document.
+But, one last thing before you move on to the example analysis and your own analysis...
 
 ### Citations in R markdown
+
 You can include citations in markdown documents using the following syntax: `[@citation_key]`.
-The citation key refers to the unique identifier generated when exporting a bibliography from a reference manager.
+The citation key refers to a unique identifier generated when exporting a bibliography from a reference manager.
 For the manuscript you will need to export the references from your reference editor in one of many formats, but I would recommend BibTex (.bib).
-For other formats see [the markdown documentation](http://rmarkdown.rstudio.com/authoring_pandoc_markdown.html#citations).
+For other possible formats see [the markdown documentation](http://rmarkdown.rstudio.com/authoring_pandoc_markdown.html#citations).
 
-Let's take a look at an example, open the "references.bib" file in the "manuscript folder".
+Let's take a look at an example, open the "references.bib" file.
 You'll notice each reference begins with an @-symbol and the type of reference.
@@
 In Rmarkdown insert citations via `[@citation_key]`.
-The `citation_key` string does have some stipulations
+The `citation_key` string does have some stipulations,
+
 >The citation key must begin with a letter, digit, or \_, and may contain alphanumerics, _, and internal punctuation characters (:.#$%&-+?<>~/).
+
+This can cause some issues with BibTex files exported from some reference managers, that is why I recommend Zotero with the Better BibTex add-on.
 
@@
 
+Now you are ready to work through example.Rmd (and/or example.html) and begin thinking about and analyzing natural variation in your gene family!
+
 Thanks and happy collaborative writing!

### 2018-01-01 example.Rmd
edited tutorial, made manuscript skeleton

---
title: "Example gene-wise analysis of the 1001 genomes data set"
author: "R. Clay Wright"
date: "`r Sys.Date()`"
output: 
   html_document: 
     toc: true
     number_sections: true
bibliography: references.bib
csl: apa.csl
---

```{r, echo = FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(warning = FALSE, collapse = TRUE, comment = "#>", tidy = TRUE, out.width = "90%", fig.height=4, fig.width=7, out.extra='style="margin: auto; display: block; padding-top: 15px;"')
library(r1001genomes)
library(DT)
library(dplyr)

```

<style>
  .using-app-insert {
     background-color: #bfbfbf;
     padding: 15px;
     border-style: solid;
     border-color:  #8c8c8c;
     border-radius: 7px 7px 0 0;
     border-width: 1px;
  }

  .using-R-insert {
     background-color: #e6e6e6;
     padding: 15px;
     border-style: solid;
     border-color:  #8c8c8c;
     border-radius: 0 0 7px 7px;
     border-width: 1px;
     border-top: 0;
  }

  .using-app-insert h2, .using-app-insert h3, .using-app-insert h4 {
     border-bottom: 0;
  }

  .using-R-insert h2, .using-R-insert h3, .using-R-insert h4 {
     border-bottom: 0;
  }

  img {
    display: block;
    margin: auto;
    width: 90%;
    height: auto;

  }
</style>

# Introduction

This web app and open source software package for the R programming language was built to facilitate exploration of the naturally occurring sequence diversity for your favorite gene or gene family. We hope this product will help plant biologists without bioinformatics experience access the 1001 genomes project’s rich catalog of natural variation, formulate hypotheses and identify new interesting alleles in their favorite gene family.

The 1001 genomes dataset was collected with the intention of genome scale analyses, such as genome wide association studies, in which a particular phenotype is quantified for all of the accessions and sequence variants that are associated with phenotypic variation are identified. [In recent work](https://doi.org/10.1534/genetics.117.3000), we have utilized this rich dataset in a different manner to measure the effects of coding sequence variation on protein functional variation and phenotypic variation. This work has expanded the sequence-function-phenotype map for the auxin-signaling F-box gene family, and this product will facilitate similar analysis other genes, gene families, and genomic regions.

So pick out your favorite gene, gene family or genomic region and follow the steps in this guide to analyze how it varies in plants collected from all over the world. For this example we will be using the auxin-signaling F-box genes, *TIR1* and *AFB2* which have the TAIR locus IDs [AT3G62980](https://www.arabidopsis.org/servlets/TairObject?id=39952&type=locus) and [AT3G26810](https://www.arabidopsis.org/servlets/TairObject?id=38086&type=locus). Click on one of the previous links to search TAIR for the TAIR IDs of your favorite genes.

***
Before digging into the data, make some hypotheses about what genes, domains and/or positions in your 
favorite gene family might be under different evolutionary pressures. We can make some inferences about the evolutionary pressure acting on particular genes based on the sequence and geographic diversity observed in the 1001 genomes dataset. Thomas Mitchell-Olds and Johanna Schmitt published an excellent overview of the evolutionary inference from natural genetic diversity in *A. thaliana* [@mitchell-olds_genetic_2006]. We can calculate genetic diversity as the average number of nucleotide differences per site between two DNA sequences in all possible pairs in the sample population, and is denoted by $\pi$.

However, evolutionary pressures potentially exert influence across all levels of biology, nucleotides, genes, genomes, phenotypes, organisms and populations.  This makes any inference from study of genetic diversity contingent on potentially unknown effects of epistasis between genes, population structure, environment, and many other effectors. Additionally, as an agricultural tag-along, much of the genetic variation in *A. thaliana* is due to admixture between populations as the result of human influence migration. So keep in mind throughout this analysis the limitations of examining a single snapshot in time of a potentially biased collection of frequently self-fertilizing organisms that were dispersed by human migration. Because of these limitations, inferring theories about how evolutionary forces are acting on your favorite genes is very uncertain business. However, if we assume a few things we can make some speculations.

Assumptions:  
1. Polymorphisms in the dataset are not *completely* detrimental to organismal fitness.  
2. Most nonsynonymous polymorphisms in a gene result in a reduction in function.

Now, taking a gene-centric view of evolution, here are some of the evolutionary pressures, or types of natural selection, that we might observe:

### Purifying selection  
If a gene's sequence is highly conserved (low diversity) it may be under [purifying/negative/stabilizing selection](https://en.wikipedia.org/wiki/Negative_selection_(natural_selection)). 
If a genes' function is crucial to the fitness of the organism, and most polymorphisms result in loss of function, we would expect a low level of nonsynonymous polymorphism in this gene. This is because deleterious polymorphisms that do arise in this gene within the population will be outcompeted by the more-fit parent. This results in the decreased genetic diversity in genes under purifying selection.

### Genetic Drift
If evolutionary pressure is very weak on a gene its sequence might change with no effect on organismal fitness. This allows a gene to accumulate many mutations that may persist in the population or be lost at random. This results in random fluctuations in allele frequencies, which when combined with an appreciable rate of mutation and a sufficient population size yields increased genetic diversity in the population.

Beyond purifying selection and drift it is very difficult to make any strong inferences on the population as a whole. However there are two more types of selective pressure, balancing or neutral selection and positive or directional selection.

### Balancing selection
Genes of high diversity may be under balancing selection, also called neutral divergence. Balancing selection maintains multiple variants for long evolutionary times because different alleles provide selective advantages in different environments or genetic backgrounds. This results in high diversity at the particular site under selection and increased diversity at closely linked neutral sites due to genetic hitchhiking with this allele under balancing selection.

### Directional selection  
If a polymorphism (or allele) is associated with a phenotype that provides an organism with a fitness advantage this phenotype and associated allele may be experiencing directional or positive selection. Positive/directional selection results in an increase in the prevalence of the advantageous allele within the population over time. Directional selection is one of the most difficult things to infer and is a very active area of research. Directional selection, like stabilizing selection is associated with reduced genetic diversity within the population under selection. However, unlike stabilizing selection the gene under directional selection will potentially have two major alleles that provide adaptive potential to different environments or genetic backgrounds.

# Hypothesize
Now go read/collect the literature about your family and make some hypotheses!  
Here are some questions to help you get started:

What structure/function knowledge is there?  
What is each domain in the protein doing?  
Are there any positions that are of vital importance?

Is there a published protein sequence alignment of your gene family?  
What positions in the protein family are highly conserved?  
What positions are highly variable?

Is there a published phylogenetic tree of your gene family?  
What are the most recently duplicated sister gene pairs?  
What do you know about the phenotypes of mutants in these pairs?  
Do the genetics suggest these pairs are redundant?

## For example
In TIR1 and AFB2 there are 2 domains. The F-box domain interacts with the SCF components [@gray_identification_1999], and the LRR domain binds auxin and Aux/IAA proteins [@kepinski_auxininduced_2004]. The F-box is roughly 40 amino acids and wraps tightly around the SKP1 component of the SCF complex. The LRR is greater than 500 amino acids and binds auxin in a pocket formed by a horseshoe shaped chain of LRRs [@tan_mechanism_2007]. The outside of this horseshoe shape does not have any known interactions other than a dimerization interface [@dezfulian_oligomerization_2016]. Because the F-box domain must interact with the SCF complex for TIR1/AFB2 to perform their known function, we might expect the F-box to be highly conserved. The small size of the F-box also helps strengthen this hypothesis. The LRR domain is much larger and may have differing specificities for Aux/IAAs, so we might expect it to be more diverse. 
We would also expect the residues of the LRR that contact auxin and the Aux/IAA degron to be highly conserved.

***

# A note on the format of this document
<!--=========================================================================-->
<div  class="using-app-insert"><!--panel start--------------------------------->
## Using the App

In addition to calling the functions of the r1001genomes package in R directly. There is also a GUI provided in the form of a Shiny app. This app can perform some of the analysis tasks the package is capable of performing, without the need to have knowledge of programming in R.

In this vignette panels like this one are used to show examples of using the app to achieve similar results as using  the R function calls directly in this vignette. Screenshots will usually be provided to make it clear what section or tab we are talking about

```{r, echo = FALSE}
knitr::include_graphics("Screenshots/App_Initial.PNG")
knitr::asis_output("
                   ") # This newline required for headings to render after image
```
## Starting the web app

To start the Shiny app, run the following command:
```{r, eval=FALSE}
run1001genomes()
```
The app should appear in a new window, and look similar to the image above.
</div><!--panel end------------------------------------------------------------>
<div  class="using-R-insert"><!--panel start----------------------------------->
## Using R
In these lighter colored panels, the equivalent R code will be displayed.
</div><!--panel end------------------------------------------------------------>
<!--=========================================================================-->

# Download the chromosome positions and gene models 
The first step in retrieving the natural variation in our genes is to retrieve the information about where the genes are in the genome of Col-0, the laboratory reference accession, and how the mRNA and coding sequence is laid out. This information is called the gene model.

<!--=========================================================================-->
<div  class="using-app-insert"><!--panel start--------------------------------->
## Loading Gene Data in the App
```{r, echo=FALSE}
knitr::include_graphics("Screenshots/Tab_1_Select_Genes.PNG")
knitr::asis_output("
                   ") # This newline required for headings to render after image
```
### Selecting Genes and Generating the Gene Model Table
At the top of the first tab (the "SNP Stats" tab) is a text input field where you can input a list of TAIR loci.

The app will download the gene model information as soon as you click submit. Copy and paste your TAIR loci in the box, separated by commas, and click submit.

The app may take a little while to process after clicking the submit button on the first tab, especially if you have a long list of genes, but once this process is complete on the first tab, the data for these genes will be already loaded for the subsequent tabs.

Once the loading process has finished, the first table you should see scrolling down the SNP Stats tab is the Gene Information table. It should look similar to the table produced in the "Using R" section below.
</div><!--panel end------------------------------------------------------------>
<div  class="using-R-insert"><!--panel start----------------------------------->
## Using R
The code the app uses to download this table is the `getGeneInfo` function. This function is just a simplifying wrapper for the `BioMart` function `getBM`. This downloads the gene model from ENSEMBL plants.

```{r}
tair_ids <- c("AT3G62980", "AT3G26810")
geneInfo <- getGeneInfo(genes = tair_ids, firstOnly = TRUE, inputType = "tair_locus", useCache = TRUE)
```

Now we can give this table a look. We can also freshen up the table a bit with `datatable`. 
```{r}
datatable(geneInfo[, -c(5,6,9)], colnames = c("tair locus", "symbol", "transcript", 
                                              "Chr", "transcript \nstart", "transcript \nend", 
                                              "transcript \nlength"), rownames = FALSE, 
          options=list(paging=FALSE, searching=FALSE))
```
</div><!--panel end------------------------------------------------------------>
<!--=========================================================================-->

Do you have a spreadsheet of loci?

Here is some quick code to make a text string you can copy and paste into the app gene input box.  
```{r}
AFB_loci <- read.csv(system.file("extdata", "AFB_gene_ids.csv", package = "r1001genomes"))
paste(AFB_loci$tair_locus, collapse = ",")
```

Occasionally there are multiple models for a single gene if there are multiple possible transcripts or splice variants. Does your gene have multiple models?

# Download the variants in the 1001 genomes dataset for these genes
The 1001 genomes dataset is stored as a VCF (Variant Call Format) file. This is 
a very efficient format for storing such a large dataset (the 1135 genomes are ~17gb), 
and for some analyses its matrix format is very useful.

<!--=========================================================================-->
<div  class="using-app-insert"><!--panel start--------------------------------->
## Downloading and processing of VCFs in the App
Upon clicking the "Submit" button on the "SNP Stats" tab, the web app automatically creates a tidy dataframe and calculates the number of unique and nonunique polymorphisms in each region of the gene model as well as calculating the diversity in the protein coding sequence. The diversity is the average number of nucleotide differences per site between all possible pairs of sequences. Diversity is also calculated for only the set of synonymous and nonsynonymous polymorphisms.

The results of these calculations are displayed in the three tables below the Gene Information table.

</div><!--panel end------------------------------------------------------------>
<div  class="using-R-insert"><!--panel start----------------------------------->
## Using R
We can download a set of VCF files for our favorite genes using the `VCFList` function.

```{r, eval=FALSE}
YFG_VCF <- VCFList(geneInfo = geneInfo, by = "transcript", tidy = FALSE)
```

However, for some filtering, summarizing and plotting operations it is more efficient to 
convert the matrix into a very long data frame. Most of the functions in `r1001genomes` 
operate on a long data frame, often called a tidy data frame. We can automatically covert 
our VCF to a tidy data frame by setting the `tidy` argument to `TRUE`.

```{r, cache=TRUE}
YFG_VCF <- VCFList(geneInfo = geneInfo, by = "transcript", tidy = TRUE)
```

We now have a list of very long data frames with all the information on SNPs inside, but much of the details on the effects of these polymorphisms are contained in a single column labeled 'EFF' as a long text string, making it hard to filter and process the data by the individual attributes within that text string.
The `parseEFF` function will separate individual fields of the EFF string into separate columns on a single dataframe, and can be applied to our list of data frames with the `llply` function from the plyr package
```{r}
YFG_VCF <- llply(YFG_VCF, parseEFF)
```

To generate the three tables of polymorphism stats the app calls the function `polymorphRow`. In R we can do the same thing.
```{r, cache=TRUE}
#TIR1_AFB2_table <- polymorphTable(geneInfo = geneInfo, strains = strains)
YFG_VCF <- llply(YFG_VCF, Nucleotide_diversity)
TIR1_AFB2_table <- ldply(YFG_VCF, polymorphRow, geneInfo = geneInfo, .id = "Transcript_ID")

```

Before displaying this table we will again clean it up a bit.

```{r}
formatRound(datatable(TIR1_AFB2_table[,c(1:6,8)],
                  colnames = c("transcript", "5' UTR", "intron", "3' UTR",
                               "coding \n synonymous", "coding \n missense",
                               "coding \n total"), 
                  escape = FALSE, rownames = FALSE,
    options=list(paging=FALSE, searching=FALSE)), columns = 2:7, digits = 0)
```
The nucleotide diversity for the chosen genes is also calculated in the call to `polymorphTable`.

```{r}
formatRound(datatable(TIR1_AFB2_table[, c(1,13,9:12)],
                  colnames = c("transcript",
                               "&pi; transcript",
                               "&pi; coding",
                               "&pi;<sub>N</sub>",
                               "&pi;<sub>S</sub>",
                               "&pi;<sub>N</sub>/&pi;<sub>S</sub>"), 
                  escape = FALSE,
                  options = list(paging=FALSE, searching=FALSE)),
                columns = 1:6, digits = 6)
```

</div><!--panel end------------------------------------------------------------>
<!--=========================================================================-->

## Results
Do these values match your hypotheses? Is there anything unexpected?

For our TIR1 and AFB2 example, we can make several interesting observations based on this dataset.  
1. *AFB2* had higher number of total polymorphisms than *TIR1*, across all subsets of sites except for introns. 
2. *AFB2* also has a higher total coding diversity and higher ratio of diversities at nonsynonymous to synonymous sites.

This was expected, as *TIR1* mutants have a strong loss of auxin sensitivity whereas *AFB2* mutants have a much weaker effect. Therefore, polymorphisms in *TIR1* are much more likely to have an effect on phenotype than polymorphisms in *AFB2*.

# Diversity within each gene

Now we'll move on to the second "Diversity Plot" tab. This tab will calculate the diversity at each position within the coding sequence of each gene. In addition to calculating the nucleotide diversity of a whole sequence, like the transcript of a gene, we can also calculate the diversity statistic for each position along the reference genome. This could be useful if you suspect a particular region of your gene may be more important to function or more sensitive to nucleotide  or amino acid changes.

<!--=========================================================================-->
<div  class="using-app-insert"><!--panel start--------------------------------->
## Diversity Plot
The second tab of the app, labeled "Diversity Plot" allows you to plot the nucleotide diversity for each of the genes you selected in the "SNP Stats" tab.  
```{r, echo=FALSE}
imagePath <- "Screenshots/Tab_2_Select_Gene.PNG"
knitr::include_graphics(imagePath)
knitr::asis_output("
                   ") # This newline required for headings to render after image
```
After clicking to the "Diversity Plot" tab, select the first gene you would like to plot then click the submit button.  
![](Screenshots/Tab_2_Plot_Selection_Box.PNG)

You should see a plot like the one above. The plot shows Nei's nucleotide diversity by *nucleotide position* of the protein coding region of the gene, though the X-axis is marked by codon number from the N-terminus for convenience. The symbols are differentiated by whether the site has a synonymous or missense variant. If multiple variant types are present at the same position, both markers will be plotted with a slight jitter, to prevent over plotting.

Below the plot is a grey box. If you click and drag across some points in the plot, this will create a selection box. The data for the selected points will appear in the grey box.

Below the plot is a data table with the data provided for all points on the plot. You can click the "Download Content..." button above the table to download a .csv file of this complete table.

</div><!--panel end------------------------------------------------------------>
<div  class="using-R-insert"><!--panel start----------------------------------->

## Using R
The app uses the function `getCodingDiv` to calculate the diversity stat at each nucleotide position of the coding sequence, then uses the function `plotCodingDiv` to plot that information for a single gene.

```{r}
YFG_Coding_Diversity <- llply(YFG_VCF, getCodingDiv)
lapply(YFG_Coding_Diversity, plotCodingDiv)

```

</div><!--panel end------------------------------------------------------------>

## Results
Do these values match your hypotheses? Is there anything unexpected?  
In *AFB2* there are five alleles with diversity greater than 1E-2 and 20 alleles with diversity greater than 3E-3. *TIR1* has two alleles with diversity greater than 1E-2 and nine alleles with diversity greater than 3E-3. This high number of high frequency alleles in the population suggests that *AFB2* is more likely to be under drift or balancing selection relative to *TIR1* or that *TIR1* is under stronger purifying selection than *AFB2*. It is important to consider however that sampling bias in certain geographic populations could affect these hypotheses. To get some idea of if this is true we will look at the geographic distribution of these alleles.

<!--================== TEMPLATE START =======================================-->
# Mapping the collection location of accessions by alleles   
As an initial visualization of the population structure that may be influencing the allele frequency and prevalence of alleles in a gene, we can map the collection locations from which the accessions were collected. If the accessions sharing a similar polymorphism are all closely clustered then it is likely several members of the same population derived from a common ancestor are represented. If there are many distantly located accessions sharing the same polymorphism this may result from human-associated migration and admixture or possibly from independent mutation events. In addition to these patterns, look for potential environmental gradients of particular alleles (e.g. is a particular allele only found above/below a certain latitude or only in coastal/desert environments?).

<!--=========================================================================-->
<div  class="using-app-insert"><!--panel start--------------------------------->
## SNP mapping
When you first click on the "SNP mapping" panel you will be faced with a choice of which genes you would like to map accessions for. Click in the checkboxes of the gene or genes you would like map variant accessions for and click submit. Now you can use the "Log nucleotide diversity filter limit" slider to set the minimum diversity of the alleles to map. You can also use the "Type of SNP to mark" radio buttons to select the type of SNPs you would like to show on your map.  
![](Screenshots/Tab_3_Select_Genes_and_Filters.PNG)  
This slider and radio buttons will automatically update the list of alleles you can add to your map in the "Allele selection" section. Select any number of alleles and click "Update map" to highlight the location of the accessions containing these alleles on the map.  
![](Screenshots/Tab_3_Mutation_Select.PNG)  
This will reveal the map with accessions containing the selected alleles shown as colored transparent points. All other accessions will be shown as grey points.  
![](Screenshots/Tab_3_Map.PNG)  
You can change the layering of the points, to remove the other (n/a) accessions or to plot one allele on top using the "Layers" button in the bottom left of the map. 
![](Screenshots/Tab_3_Map_w_Layers.PNG)
You can view the full list of accessions in the map along with the alleles those accessions contain directly below the map in the "Map Data" section. You can navigate through this datatable or search for particular alleles using the box at the top right. You can also download the contents of this datatable.
![](Screenshots/Tab_3_Map_Data_Table.PNG)

</div><!--panel end------------------------------------------------------------>
<div  class="using-R-insert"><!--panel start----------------------------------->
## Using R
To repeat the mapping of accessions in R, we will first have to filter by diversity the VCF of mutations. We also want to exclude the reference variants (i.e. when GT is 0|0). We'll use some tools from the [tidyverse](http://tidyverse.org/) to do this. 
```{r}
AFB2_missense <- YFG_VCF$AT3G26810.1 %>% 
  dplyr::filter(Diversity > 3*10^(-3) & gt_GT != "0|0" & 
                  Effect == "missense_variant")
head(AFB2_missense)
```
The accession (as ecotype ID) is in the "Indiv" column of this data frame. Now we will need to attach the accession name, latitude, longitude and other data to this data frame. We'll do this with the `addAccDetails` function. The metadata is also available as an included dataset in the r1001genomes package called `accessions`.

```{r}
AFB2_missense <- labelBySNPs(AFB2_missense)
head(AFB2_missense)
```
Now we can plot the variants. In the app we use the Leaflet package to provide an interactive plot. We can also do this within R.
```{r}
#load the leaflet package 
library(leaflet)
AFB2_missense <- rbind(AFB2_missense[is.na(AFB2_missense$SNPs), ],
                       AFB2_missense[!is.na(AFB2_missense$SNPs), ])

    # make a field with text to be displayed when clicking on a marker
    AFB2_missense$popup <- paste("EcoID:",  AFB2_missense$Indiv,"Name:", AFB2_missense$Name, " SNPs:", AFB2_missense$SNPs)

    # create the color pallet for the map points
    pal <- RColorBrewer::brewer.pal(8, "Set1")
    pallet <- colorFactor(palette=pal, domain=AFB2_missense$SNPs)
map <- leaflet()
    map <- addProviderTiles(map, providers$Stamen.TonerLite,
                     options = providerTileOptions(noWrap = TRUE))

    # groupnames to be used by draw groups of points as separate layers below
    groupnames <- unique(AFB2_missense$SNPs)
    groupnames <- groupnames[!is.na(groupnames)]

    # add markers for NA points first so they are furthest back layer
    map <- addCircleMarkers(map, data=AFB2_missense[is.na(AFB2_missense$SNPs), ], color= "#9b9b9b", group="NA",
                            radius=6, stroke=FALSE, fillOpacity=0.6)

    # for each of the group names, add a set of markers
    for (SNP in groupnames){
          map <- addCircleMarkers(map, data=AFB2_missense[AFB2_missense$SNPs == SNP, ], color= ~pallet(SNPs), group= SNP,
                            radius=6, popup= ~popup, stroke=FALSE, fillOpacity=0.85)
    }

    # add the legend to the map
    map <- addLegend(map, position="bottomright", pal=pallet,
                     values=AFB2_missense$SNPs, title="Marker Colors", opacity=1)

    # add layer control to map to turn on or off groups of points
    map <- addLayersControl(map, overlayGroups=c(groupnames, "NA"),
                            options = layersControlOptions(collapsed = TRUE),
                            position="bottomleft")

    map
```
You can also use the `ggmap` function from the package of the same name to create maps. The "ggrepel" package also helps to handle overlapping labels.

```{r}
library(ggplot2)
library(ggmap)
library(ggrepel)

europe <- get_map(location = 'europe', zoom = 4)
michigan <- get_map(location = 'michigan', zoom = 5)

ggmap(europe) + xlab('longitude') + ylab('latitude') + geom_point(data = AFB2_missense, aes(x = Long, y = Lat), alpha = 0.5, shape = 18, color = "darkgrey") + geom_label_repel(data = subset(AFB2_missense, grepl(pattern = "204", AFB2_missense$SNPs)), mapping = aes(label = Name, x = Long, y = Lat), fill = "white", point.padding = unit(0.5, "lines"), segment.size = 1)

ggmap(michigan) + xlab('longitude') + ylab('latitude') + geom_point(data = AFB2_missense, aes(x = Long, y = Lat), alpha = 0.5, shape = 18, color = "darkgrey") + geom_label_repel(data = subset(AFB2_missense, grepl(pattern = "516", AFB2_missense$SNPs)), mapping = aes(label = Name, x = Long, y = Lat), point.padding = unit(0.5, "lines"), segment.size = 1)

```

</div><!--panel end------------------------------------------------------------>
<!--=========================================================================-->

# References

<!--================== TEMPLATE START =======================================--
# heading   
description

<!--=========================================================================--
<div  class="using-app-insert"><!--panel start---------------------------------
## Tab title
some text and picture

</div><!--panel end------------------------------------------------------------
<div  class="using-R-insert"><!--panel start-----------------------------------
## Using R
some text and code snippet
</div><!--panel end------------------------------------------------------------
<!--=========================================================================-->

### 2018-01-01 manuscript/manuscript.Rmd
edited tutorial, made manuscript skeleton

@@
   name: "R. Clay Wright"
-author:
 
@@
     html_document:
+      toc: true
       fig_caption: yes
-#  word_document:
-#    fig_caption: yes
-#    fig_height: 6
-#    fig_width: 6
-#    highlight: haddock
+
 keywords: r markdown
@@
 fontfamily: Helvetica
-output: 
-   pdf_document: 
-     toc: true
-     number_sections: true
-bibliography: ["references.bib"]
+
+bibliography: ["natural_variation.bib", "TIR1_AFB.bib"]
 csl: apa.csl
@@
 
-#List of working titles  
-You should keep a list of titles and constantly be refining and tweaking it. A title helps focus you. The title should be in the same style as papers you plan to cite.
-You might also think of snappy titles or sub titles for press releases. Find some press releases for paper similar to yours to see how journalists interpret the results.
+# List of working titles  
+* A comprehensive analysis of natural sequence variation within the <i>Arabidopsis thaliana</i> nuclear auxin signaling pathway  
+* *Add your suggestions*
 
-#Abstract  
-Your abstract should be in the following form, more or less exactly:
+# Abstract
 
@@
 * A concluding statement.
-When presenting your abstract to your advisor, these sentences should be clearly laid out, in a bullet list preferably.
 
-#Related work  
-You need a list of about 30 papers that are the most relevant.
-They should be weighted toward recent, well cited work.
-You should group them into two to four areas.
-For each paper, you should have a sentence that says what the paper is about and how it relates to your work.
+# Related work  
+What and how have we, as a field, learned from natural variation?  
+[@hollister_genomic_2014; @bielawski_detecting_2013; @kryazhimskiy_population_2008; @schmid_multilocus_2005; @nieduszynski_sequence_2011; @melamed_combining_2015; @weigel_population_2015; @schmitz_patterns_2013; @weigel_natural_2005; @hardtke_case_2010; @koornneef_naturally_2004; @mitchell-olds_genetic_2006; @joly-lopez_developing_2016; @wright_insights_2017a; @park_distributions_2017; @_135_2016a]
 
-#Glossary  
-Make a list of all of the new terms that you have invented to talk about different parts of your work.
-This list will eventually be put into paragraph format, but for now it should look like a glossary.
-You can also add terms that from the literature that you will use in case you want to make sure that the reader (or you or your advisor) really knows what you mean when you use that term.
+# Methods
 
-#Results  
-This should be a list of 3-5  statements  that very exactly say what you are claiming as results, and then the support for that claim.
-Statements will be then supported by experimental, theoretical, or numerical evidence.
+# Results  
+## TIR1/AFB genes  
+[@gray_identification_1999; @kepinski_auxininduced_2004; @tan_mechanism_2007; @dezfulian_oligomerization_2016; @wu_evolution_2017b]
 
-Note that it is very important that you do not use mere titles for the sections of your paper.
-If you do, you end up with a laundry list.
-So do not say "transcription initiation" and list all of the things you know about the subject.
-Instead say, say "Transcription initiation is regulated by proteins that bind to or near promoters", and then list the evidence you have (or plan to get) that says why you suspect the sentence to be true.
+## Aux/IAA genes
 
-Another way to think of these statements is that they are essentially your hypotheses, if you are a scientist, or they are statements of capabilities if you are an engineer.
+### Group A Aux/IAAs
 
-#Figures and captions
+### Group B Aux/IAAs
 
-You should have mock ups of three to five figures (with sub?figures) that you plan to get the data for.
-These figures are incredibly important to think about from an early stage.
-For example, if you know that you will have a certain data figure with certain x and y axes and certain error bars, then you have a specification of what data you will need to generate and what it ought to look like.
+### Group C Aux/IAAs
 
-You should also have working drafts of all of the captions for each sub figure. Captions have  titles and then descriptions.
+## TPL/TPR genes
 
-Finally, you should know what glossary terms or results statements the figures correspond to and roughly how you will use them in your paper.
+## ARF genes
 
-#List of things to do
+### Class A ARFs
 
-Keep a running list of everything you need to do to finish the paper.
-Don’t put anything extraneous on the list.
-Make sure every unfinished result has a corresponding sub?list of things to do.
-Make sure that there are always items on your list that you can and will do t  his week,  as well as longer term items that help you plan ahead.
+### Class B ARFs
 
-# Abstract
+### Class C ARFs

### 2018-01-02 TUTORIAL.Rmd
a few more edits

@@
 
-As it says: "You can access this container using the following IP address and port". This means that if you point your browser to the "ACCESS URL" you will see the version of Rstudio that we've built.
+As it says: "You can access this container using the following IP address and port".
+This means that if you point your browser to the "ACCESS URL" you will see the version of Rstudio that we've built.
+The username and password are both "rstudio".
 
@@
 You can also open files from this folder in RStudio.
-In the RStudio "Files" pane, navigate to the "manuscript" folder and set it as your working directory (using the "More" button).
-Click on "manuscript.Rmd" to open the manuscript markdown file.
+You can save your data and the state of your RStudio session into this folder and it will stay on your computer.
+If you quit your browser you can navigate back to the same access url, because the docker container will be accessible unless you stop it in kitematic.
+
+In the RStudio "Files" pane, navigate to the "manuscript" folder and set it as your working directory (using the "More" button). Click on "manuscript.Rmd" to open the manuscript markdown file.
 

### 2018-01-02 manuscript/manuscript.Rmd
a few more edits

@@
 ## TIR1/AFB genes  
+Romi
 [@gray_identification_1999; @kepinski_auxininduced_2004; @tan_mechanism_2007; @dezfulian_oligomerization_2016; @wu_evolution_2017b]
@@
 ### Group A Aux/IAAs
+Britney
 
 ### Group B Aux/IAAs
+Mollye/Alex
 
 ### Group C Aux/IAAs
+Mega/Alex
 
 ## TPL/TPR genes  
+Alex
 
@@
 ### Class A ARFs
+Amy
 
 ### Class B ARFs
+Andrew/Arjun
 
 ### Class C ARFs
+Hardik

### 2018-01-02 VCF_downloader_playground.R
Deleted a few unnecessary files for coauthors

@@ -1,209 +1,0 @@
-
-#source("Natural_Variation_Shiny_APP/VCF_downloader.R")
-require(BSgenome.Athaliana.TAIR.TAIR9)
-
-# this section is mainly a playground for calling functions and testing workflows
-#
-
-#geneInfo <- getGeneInfo(IAAgenes)
-#geneInfo <- geneInfo[-c(2,6,8,13,15,17,18,19,20,21,23,28,31,32,34,35,37,39,40,42,44,46,48,53),]
-
-geneInfo <- getGeneInfo(TPLgenes)
-#geneInfo <- geneInfo[c(3,6,8,12,14), ]
-
-#run loadData on a single transcript_ID, useful for debuging
-SNPData = loadData(geneInfo[2,], strains)
-#diversity <- Nucleotide_diversity(SNPData)
-plotData <- plotPi(SNPData)
-
-syn_loci <- filtR(SNPData,split_var="names",col_name="Effect",value="synonymous_variant")
-#NucDiv(syn_loci)
-
-missense_loci <- filtR(SNPData,split_var="names",col_name="Effect",value="missense_variant")
-#NucDiv(missense_loci)
-
-tableData <- rbind(simplifySNP(missense_loci), simplifySNP(syn_loci))
-
-#geneInfo <- getGeneInfo(genes)
-#tableData = polymorphTable(geneInfo, strains)
-#write.csv(tableData,"output\\table.csv")
-
-
-#SNPData[SNPData$names=="3:9869957_C/T",]
-
-data1 <- readVcf(file="data1.vcf", genome="Athaliana")
-data1Granges <- rowRanges(data1)
-myStringSet <- data1Granges[271]$ALT[[1]]
-
-Athaliana<-BSgenome.Athaliana.TAIR.TAIR9
-subseq(Athaliana$Chr1, start(data1Granges[271:272]), end(data1Granges[271:272]))
-
-tair10 <- useMart("plants_mart", host="plants.ensembl.org", dataset="athaliana_eg_gene")
-attributePages(tair10)
-listAttributes(tair10, page="feature_page")
-
-output <- getBM(attributes=c("ensembl_transcript_id", "chromosome_name", "start_position",
-                             "end_position", "strand", "transcript_start", "transcript_end", "pfam", "pfam_start", "pfam_end",
-                             "gene3d", "gene3d_start", "gene3d_end"
-), filters="tair_locus", values=IAAgenes[1:5], mart=tair10)
-
-
[…]

### 2018-01-05 example.Rmd
Added section about creating annotation tables.

@@
 library(dplyr)
+library(reshape2)
 
@@
 We would also expect the residues of the LRR that contact auxin and the Aux/IAA degron to be highly conserved.
+
+## Creating annotations
+
+We can build a table of all of these structure function annotations in Excel as a CSV or within R. For the AFBs we will annotate the protein domains and a micro RNA binding site.
+
+```{r}
+# create matrix for F-box/LRR division in the coding sequence (CDS), 
+#  based on Tan et al. supplementary alignment 
+domains<-matrix(data = c(c(1,1,1,1,151,151,1), # Fbox start in CDS
+                         3*c(51,47,46,46,96,96,58), # Fbox end in AA * 3 = CDS 
+                         3*c(51,47,46,46,96,96,58)+1, # LRR start 
+                         c(1785,1758,1728,1734,1872,1860,1779)), # LRR end 
+             nrow = 4, ncol = 7, byrow = TRUE, 
+             dimnames = list(c('Fbox_start', 'Fbox_end', 
+                               'LRR_start', 'LRR_end'), 
+                           c("TIR1","AFB1","AFB2", "AFB3",
+                             "AFB4", "AFB5", "COI1")))
+
+#create columns for miR393
+miR393_start <- c(1519,1507,1504,1510,1657,1657,1551)
+miR393_end <- miR393_start+18
+
+# bind these columns to the domains
+domains<-rbind(domains,miR393_start,miR393_end)
+
+# convert to a data frame and convert values to AA numbering from CDS
+domains <- as.data.frame(t(round(domains/3))) 
+#domains
+write.csv(domains, "manuscript/annotations_wide.csv")
+```
+
+You can used the `annotations_wide.csv` file as a basis to build annotations for your gene family.
+However because we can make alignments of our genes because they are so closely related it is not necessary to create an annotation for each gene in you family.
+Also this wide format doesn't make much sense if we are trying to annotate something unique to one gene.
+Instead I would recommend using a long format annotation.
+This will also make plotting these annotation on top of plots and alignments later much easier.
+Here is how we can convert the wide format annotation data frame into a long format.
+
+```{r}
+# create a column for the genes, which are currently the row names 
+domains$gene <- row.names(domains)
+
+# We'll order the gene names so that they will always be plotted in the 
+# conventional order
+domains$gene <- factor(x= domains$gene, levels = c('TIR1','AFB1','AFB2','AFB3','AFB4','AFB5','COI1'), ordered = T)
+
+# Now we'll use melt() from the reshape2 package to "melt" the data frame into 
+# long format. It's best to look at the output to figure out what melt does.
+domains_long <- reshape2::melt(data = domains, 
+                               id.vars = "gene", 
+                               variable.name = "annotation", # column name for 
+                               # all of the column names in the wide format
+                               value.name = "position") # what is the name 
+                               # of all of the values in the table
+domains_long
+
+# Let's add another column to indicate that these annotations are for 
+# amino acid positions
+domains_long$position_type <- "AA"
+#write.csv(domains_long, "manuscript/annotation_long.csv")
+```
+
+I would use this `annotation_long.csv` to keep track of the structure/function information on individual genes you come across as you are reading the literature.
+Then you can map this information on top of the diversity plots and alignments that we will create below to see if these annotations apply to the whole family or just the single genes cited in the literature.
 

### 2018-03-07 manuscript/manuscript.Rmd
Amy's ARF-A secton and some ARF-B

@@
 [@gray_identification_1999; @kepinski_auxininduced_2004; @tan_mechanism_2007; @dezfulian_oligomerization_2016; @wu_evolution_2017b]
+1) Plant biologist of all types should use this web tool!
 
+>Makes it really easy to look at the evolution of genes in arabidopsis.
+>Makes it really easy to look at how families of genes are evolving.
+>Easy way to find accessions that might be interesting to you.
+
+2) This web tool can be used to identify domains or residues of potential functional importance in any gene or gene family. (E.g. stretches of conservation in regions of unknown function).
+
+>I was able to find that domains of iaa and phytate interaction are highly conserved, which is what we would expect.
+>I was able to show that extra flanking domains of AFB4 and 5 are very mutated and therefore probably not important.
+
+3) It can help identify sites of relevant functional divergence between sister pairs and more broadly within gene families (E.g.
+one sister has much higher diversity than the other.
+What are the shared regions of variation?
+What regions are highly variable in the more diverse sister?)
+
+>AFB4 the most diverse sister. All regions highly variable. Iaa binding mutated. likely not functional at all. All have higher variation before fbox binding domain and before dimerization domain.
+
+4) It can also identify accessions containing multiple nonsynonymous genes in a family or network that can be used for studying network evolution.
+(E.g.
+there is a nonsynonymous polymorphism in a key functional residue of a gene that you expect will alter its function, however there are many other polymorphisms in other genes in the pathway that may or may not compensate for this change.
+We can assess this by backcrossing this accession with Col-0 or other distantly related accessions and resequencing phenotypic segregants).
+
+>Did not find this yet. None seem to be in a critical site, so I will be looking later on.
 ## Aux/IAA genes
@@
 
+Auxin response is mediated by the auxin responsive transcription factors, or ARFs.
+There are 23 ARFs in Arabidopsis that are divided into three phylogenetic classes.
+Class A ARFs, ARF5, ARF6, ARF7, ARF8, and ARF19, activate transcription.
+These ARFs have a Q-rich region in the middle of the protein that may mediate activation (Guilefoyle, 2007).
+It has recently been shown that the middle region of ARF5 interacts with the SWI/SNF chromatin remodeling ATPases BRAMA and SPLAYED, possibly to reduce nucleosome occupancy and allow for the recruitment of transcription machinery (Wu, 2016).
+Additionally ARF7 can interact with Mediator subunits, directly tethering transcriptional proteins to its binding regions (Ito, 2016).
+Class B and C ARFs are historically categorized as repressor ARFs, though the mechanism through which they confer repression has not been identified.
+Their middle regions tend to be proline- and serine-rich.
+
+ARFs are comprised of three major domains, recent structures of which have informed structure-function analysis of the ARFs (Boer, 2014; Nanao, 2014; Korasick, 2014).
+These domains are conserved throughout land plants.
+ARFs share an N-terminal B3 DNA binding domain.
+Flanking this DNA-binding domain is a dimerization domain, which folds up into a single “taco-shaped” domain to allow for dimerization between ARFs.
+There is an auxillary domain that immediately follows and interacts with the dimerization domain.
+The middle region is the most variable between ARFs, as mentioned above, but is characterized by repetitive units of glutamine (class A), serine, or proline residues (classes B and C).
+On the C-terminal end of the protein a protein-protein interaction domain mediates interactions among ARFs, between ARFs and other transcription factors, and between ARFs and the Aux/IAA repressors.
+This interaction domain was recently characterized as a Phox and Bem1 (PB1) domain, which is comprised of a positive and negative face with conserved basic and acidic residues, respectively.
+Theoretically the dipolar nature of the PB1 domain can mediate multimerization by the pairwise interaction of these faces on different proteins, and indeed the ARF7 PB1 domain was crystallized as a multimer.
+It is unclear whether this multimerization occurs or plays a developmental significant role in vivo.
+Interfering with ARF dimerization in either the DNA-binding proximal dimerization domain or the PB1 domain decreases the ability of class A ARFs to activate transcription in a heterologous yeast system (Pierre-Jerome, 2016).
+
+While domain architecture is broadly conserved among the ARFs, there are exceptional cases.
+Three ARFs do not contain a PB1 domain at all, ARF3, ARF13, and ARF17, and several more have lost the conserved acidic or basic residues in the PB1 domain, suggesting they may be reduced to a single interaction domain.
+Several ARFs additionally have an expanded conserved region within the DNA-binding domain, of unknown function.
+The majority of domain variation among ARFs occurs in the large B-class subfamily.
+The liverwort Marcantia polymorpha has a single representative ARF of each class (Flores-Sandoval, 2015).
+The much expanded classes of the flowering plants arise from both whole genome and tandem duplication events (Remington, 2004), and may have allowed for the expansion of the quantity and complexity of loci regulated by the ARFs and a consequent expansion in their regulation of developmental processes.
+Here we examine natural variation in the three ARF classes and how this variation may be informative of conservation and function among the ARFs.
+
 ### Class A ARFs
@@
 
+Class A ARFs are the most well-studied ARF subfamily--the five family members all act as transcriptional activators and have well-characterized, distinct developmental targets.
+Class A ARFs all contain the canonical B3 DNA-binding domain and PB1 interaction domain, and all have glutamine-rich middle regions (Guilefoyle, 2007).
+Analysis of class A ARF nucleotide diversity showed that the majority of these ARFs had low nucleotide diveristy scores, all ${\pi_N/\pi_S}$ values substantially lower than 1 with the exception of ARF19, with a relatively high ${\pi_N/\pi_S}$ value of 1.8.
+Comparing diversity within sister pairs, the two class A ARF sister pairs had divergent diversity scores, with ARF19 and ARF8 having scores at least triple those of ARF7 and ARF6 respectively.
+This may suggest that ARF6 and ARF7 have more conserved function whereas ARF19 and ARF8 are more diverse.
+Overall the diversity of class A ARFs was generally low, especially compared to the class B and C ARFs, suggesting that class A ARFs are central to auxin signal transduction and plant development.
+
+For all class A ARFs, the middle region of the protein was the predominant high diversity region.
+The DNA-binding domain had very few, low-diversity missense mutations, as did the PB1 domain.
+Considering the necessity of their conserved functions, the low level of variation in these key DNA and protein-protein interaction domains is expected.
+
+Within the middle region, ARF7 had several expansions of polyglutamine sequences compared to the other class A ARFs.
+Polyglutamine regions are known to readily expand and contract throughout evolutionary time due to replication error, and variation in polyQ length can be acted on by natural selection and have phenotypic consequences (Press, 2014).
+
+A few missense variations in ARF5 highlighted potential functional variation among ecotypes.
+Two high-diversity changes at the positive face of the PB1 domain, very close to the conserved lysine residue, to phenylalanine were interstingly always found together in all ecotypes in which they occured: Tu-B2-3,Ru-2, Haes-1, Rd-0, 627ME-13Y1, Le-0,Bu-0, Bsch-0.
+These mutations to a rather bulky side chain may change the folding of the PB1 domain and interfere with interactions on the positive face.
+A high diversity polymorphism that results in change in the 17th amino acid of ARF5 from a leucine to a methionine is also intriguing.
+ARF5 has an N-terminal extension of unknown function relative to the other class A ARFs *how many amino acids*.
+It is possible that this methionine acts as an alternative start site in the accessions in which it is found yielding a truncated ARF5 variant without this N-terminal extension.
+
 ### Class B ARFs
-Andrew/Arjun
+Andrew
+
+ARF23, which is truncated in its DNA-binding region, has a high ${\pi_N/\pi_S}$ value of 4.1, and has many high-diversity valued nonsense variants, as may be expected of a gene that is pseudogenizing.
+
+ARF3, ARF13, and ARF17 all lack a PB1 C-terminal domain, which suggests they may be unable to interact with Aux/IAA repressors and may act as an auxin-insensitive mechanism of adding competitive inhibition to ARF-mediated transcription.
+These ARFs span the range of ${\pi_N/\pi_S}$ ratios relative to the other ARFs, ranging from 0.85 to 4.1.
+ARF13, with the highest ${\pi_N/\pi_S}$ of 4.1 has many nonsense variants of high diversity, perhaps suggesting it is pseudogenizing.
+
+ARF14, ARF15, ARF16, ARF20, and ARF21 all do not have one or multiple of the conserved acidic residues in the PB1 domain, and themselves form a subclade of the B ARFs.
+Interestingly, no non-synonymous variation is seen in these amino acid residues of these ARFs, suggesting that the amino acids that are in the regions may still be important for PB1 function.
 

### 2018-03-07 manuscript/manuscript.Rmd
Hardik's ARF_C section

--- 2018-01-02 manuscript/manuscript.Rmd
@@
 ### Class B ARFs
-Andrew/Arjun
+Andrew
 
 ### Class C ARFs
-Hardik
+Hardik (All figure cite below are in folder "/Dropbox/natural-variation-webtool/manuscript_ARF_C/FIGURES")
+
+Class C ARFs shows low nucleotide diveristy scores, all pi_N/pi_S values substantially lower than 1.
+Comparing diversity within coding region of class C ARF, ARF16  has least diversity whereas its family members (ARF10, ARF17) shows  scores at least four times higher (PMCID: PMC519086).
+Structurally, all three members of Class C ARFs contain both canonical B3 DNA-binding domain and whereas only ARF10 and ARF16 contains PB1 interaction domain.
+DNA binding domain exhibit overall low diversity and presence of largely synonymous changes for all the ARFs.
+Only ARF10, shows mis-sense variants (p.Ser147Thr/c.439T>A  and p.Ala157Ser/c.469G>T) DNA binding domains with higher diversity score for conservative variation (codon 147) compared to non conservation variation (codon 157) (Fig.
+ARF10_DBD Variants distribution).
+Of the PB1 domain containing class C ARFs, ARF16 exhibits several naturally occurring non-synonymous variants in contrast to ARF10 (Fig.
+Arf16_PB_variants distribution) and is sporadically distributed in location.
+
+Few members of class B ARFs (ARF 2,3,4) and C ARFs (ARF10,16, 17) distinguish from other ARFs on the basis of its ability to be post transcriptionally regulated by trans-acting small interfering RNA (tasiRNA) and microRNA (miRNA) (PMID: 12202040, PMID: 15851028, ).
+(@AMY Arf 6 and 8 targeted by miRNA 167 - check for conservation).
+Interestingly, miRNA binding motif of the class C ARFs show high degree of conservation across the natural variation of 1001 genomes queried here (Fig.
+ARF10_16_17_Alignment).
+This is consistent with perturbation studies that has shown drastic developmental phenotype for class C ARFs (PMID: 16006581, PMID: 17672844, PMID: 15829600).
+For the class B ARFs (2,3,4) tasiRNA binding motif is highly conserved and show no variation suggesting strong regulatory constrains at the locus (Fig ARF2_3_4_Alignment).
+DNA binding domain of ARF4 exhibits non-synonymous variation in 4 accession and would be interesting to carry out phenotypic characterize them for developmental growth defects (Fig ARF4_DBD Variants distribution).

### 2018-03-07 manuscript/manuscript.Rmd
Updated bib and citations for ARF_C

@@
 
-bibliography: ["natural_variation.bib", "TIR1_AFB.bib"]
+bibliography: ["natural_variation.bib", "TIR1_AFB.bib", "ARF_C.bib"]
 csl: apa.csl
@@
 Class C ARFs shows low nucleotide diveristy scores, all pi_N/pi_S values substantially lower than 1.
-Comparing diversity within coding region of class C ARF, ARF16  has least diversity whereas its family members (ARF10, ARF17) shows  scores at least four times higher (PMCID: PMC519086).
+Comparing diversity within coding region of class C ARF, ARF16  has least diversity whereas its family members (ARF10, ARF17) shows  scores at least four times higher [@remington_contrasting_2004].
 Structurally, all three members of Class C ARFs contain both canonical B3 DNA-binding domain and whereas only ARF10 and ARF16 contains PB1 interaction domain.
@@
 
-Few members of class B ARFs (ARF 2,3,4) and C ARFs (ARF10,16, 17) distinguish from other ARFs on the basis of its ability to be post transcriptionally regulated by trans-acting small interfering RNA (tasiRNA) and microRNA (miRNA) (PMID: 12202040, PMID: 15851028, ).
+Few members of class B ARFs (ARF 2,3,4) and C ARFs (ARF10,16, 17) distinguish from other ARFs on the basis of its ability to be post transcriptionally regulated by trans-acting small interfering RNA (tasiRNA) and microRNA (miRNA) [@rhoades_prediction_2002; @allen_micrornadirected_2005].
 (@AMY Arf 6 and 8 targeted by miRNA 167 - check for conservation).
@@
 ARF10_16_17_Alignment).
-This is consistent with perturbation studies that has shown drastic developmental phenotype for class C ARFs (PMID: 16006581, PMID: 17672844, PMID: 15829600).
+This is consistent with perturbation studies that has shown drastic developmental phenotype for class C ARFs [@wang_control_2005; @mallory_micrornadirected_2005; @liu_repression_2007].
 For the class B ARFs (2,3,4) tasiRNA binding motif is highly conserved and show no variation suggesting strong regulatory constrains at the locus (Fig ARF2_3_4_Alignment).

### 2018-03-07 manuscript/manuscript.Rmd

--- 2018-03-07 manuscript/manuscript.Rmd
@@
 ### Class C ARFs
-Hardik
+Hardik (All figure cite below are in folder "/Dropbox/natural-variation-webtool/manuscript_ARF_C/FIGURES")
+
+Class C ARFs shows low nucleotide diveristy scores, all pi_N/pi_S values substantially lower than 1.
+Comparing diversity within coding region of class C ARF, ARF16  has least diversity whereas its family members (ARF10, ARF17) shows  scores at least four times higher (PMCID: PMC519086).
+Structurally, all three members of Class C ARFs contain both canonical B3 DNA-binding domain and whereas only ARF10 and ARF16 contains PB1 interaction domain.
+DNA binding domain exhibit overall low diversity and presence of largely synonymous changes for all the ARFs.
+Only ARF10, shows mis-sense variants (p.Ser147Thr/c.439T>A  and p.Ala157Ser/c.469G>T) DNA binding domains with higher diversity score for conservative variation (codon 147) compared to non conservation variation (codon 157) (Fig.
+ARF10_DBD Variants distribution).
+Of the PB1 domain containing class C ARFs, ARF16 exhibits several naturally occurring non-synonymous variants in contrast to ARF10 (Fig.
+Arf16_PB_variants distribution) and is sporadically distributed in location.
+
+Few members of class B ARFs (ARF 2,3,4) and C ARFs (ARF10,16, 17) distinguish from other ARFs on the basis of its ability to be post transcriptionally regulated by trans-acting small interfering RNA (tasiRNA) and microRNA (miRNA) (PMID: 12202040, PMID: 15851028, ).
+(@AMY Arf 6 and 8 targeted by miRNA 167 - check for conservation).
+Interestingly, miRNA binding motif of the class C ARFs show high degree of conservation across the natural variation of 1001 genomes queried here (Fig.
+ARF10_16_17_Alignment).
+This is consistent with perturbation studies that has shown drastic developmental phenotype for class C ARFs (PMID: 16006581, PMID: 17672844, PMID: 15829600).
+For the class B ARFs (2,3,4) tasiRNA binding motif is highly conserved and show no variation suggesting strong regulatory constrains at the locus (Fig ARF2_3_4_Alignment).
+DNA binding domain of ARF4 exhibits non-synonymous variation in 4 accession and would be interesting to carry out phenotypic characterize them for developmental growth defects (Fig ARF4_DBD Variants distribution).

### 2018-03-07 manuscript/manuscript.Rmd
AuxIAA team section

--- 2018-03-07 manuscript/manuscript.Rmd
@@
 
-bibliography: ["natural_variation.bib", "TIR1_AFB.bib", "ARF_C.bib"]
+bibliography: ["natural_variation.bib", "TIR1_AFB.bib", "AuxIAA_citations.bib"]
 csl: apa.csl
@@
 
-### Group A Aux/IAAs
-Britney
+###Introduction
+The Aux/IAA proteins (IAAs) mediate both auxin perception and transcriptional co-repression of auxin responsive genes.
+Auxin perception is carried out along with TIR1/AFB proteins, which together act as a coreceptor to perceive intracellular auxin.
+Interaction of IAAs with auxin and TIR1/AFBs is conferred by the highly conserved 13 amino acid degron (DII) motif [@tan_mechanism_2007].
+Stabilizing mutations within the degron can lead to dominant phenotypes [@berleth_auxin_2004; @yang_iaa1_2004].
+Degradation of IAAs relieves their repression upon ARF transcription factors.
+IAA degradation rate is influenced both by the degron sequence and by rate motifs found outside the degron, such as the KR motif [@moss_rate_2015; @dreher_arabidopsis_2006].
+IAA-mediated repression of ARFs depends upon an EAR domain that recruits co-repressors of the TOPLESS (TPL) family, and the PB1 domain that enables interaction with ARF PB1 domains [@nanao_structural_2014; @kagale_ear_2011].
+The Arabidopsis Aux/IAA gene family has 29 members, most of which possess a canonical degron.
+Notable exceptions are the Auxin insensitive IAA20 and IAA30 that lack a degron sequence entirely, while the closely related IAA31 retains a semi-functional degron sequence with a very slow degradation rate [@dreher_arabidopsis_2006; @havens_synthetic_2012].
+Using the Natural Variation Webtool, we found that for the canonical Aux/IAAs, no missense mutations were present in the KR motif or PB1 interaction faces, and very few missense mutations were observed in the EAR domain and degron (Supp.
+Fig.
+Alignment of canonical IAAs).
+However, as discussed below, there were several interesting missense mutations found in EAR domains and degrons.
+The majority of nonsynonymous mutations observed were in the variable regions between the conserved domains (Supp.
+Fig.
+Alignment of canonical IAAs).
 
-### Group B Aux/IAAs
-Mollye/Alex
+###Sister Pair Diversification
+Previous research has pointed to genetic redundancy within closely related pairs or families of IAA proteins [@overvoorde_functional_2005].
+For example, the iaa8-1, iaa9-1 double mutant and the iaa5-1, iaa6-1, iaa19-1 triple mutant have wild-type phenotypes [@overvoorde_functional_2005].
+Despite this, most Arabidopsis IAA sister pairs have one member exhibiting higher diversity while the other sister is more conserved (Figure IAA phylogeny heat map).
+For example, despite having only minor differences in sequence (61.4% sequence identity), the IAA6/19 sister pair has significant differences in expression patterns, protein abundances and functions, suggesting they have undergone functional specialization since their divergence [@winkler_variation_2017].
+IAA19 was expressed at higher levels than IAA6 and showed less sequence divergence; IAA6 showed regions of increased divergence upstream of the degron and in the N-terminal half of the PB1 domain [@winkler_variation_2017].
+Consistent with these results, the Natural Variation Webtool (Figure IAA phylogeny heat map) revealed a lower natural variation for IAA19 (ᴨN/ᴨS = 0.55) compared to IAA6 (ᴨN/ᴨS = 2.3), and also detected high natural variation diversity within the same regions of IAA6 as seen in Winkler et al.
+(Supp Fig IAA6 diversity plot).
+Within IAA Group B, there are four sister pairs--IAA10 and IAA11, IAA12/BDL and IAA13, IAA18 and IAA26/PAP1, and IAA20 and IAA30 (Fig X) [@remington_contrasting_2004].
+Interestingly, among these sister pairs only IAA12 and 13 have noticeable differences in diversity.
+IAA10 and IAA11 show little difference in ℼN/ℼS values, with ℼN/ℼS of 0.80 and 0.67 respectively, implicating the possibility that IAA10 and 11 play redundant roles.
+In rice however, it has been shown that IAA11 gain of function mutants block lateral root initiation [@zhu_gain--function_2012].
+The Arabidopsis EFP browser shows that IAA10 and 11 have almost identical expression patterns, with the exception of IAA10 being more highly expressed in the developing seed, and IAA11 being more highly expressed in root tissue during lateral root development.
+As discussed below, the Natural Variation Webtool provides a potentially expedited route to conducting experiments meant to elucidate the extent of redundancy between family members such as IAA10 and 11.
 
-### Group C Aux/IAAs
-Mega/Alex
+###Identification of new Loss-of-Function/Gain-of-Function mutations
+To test the extent of redundancy/specialization between sister pairs, the Natural Variation Webtool allows for identifications of accessions that can be potentially used to investigate the phenotypic effects in loss of function variants.
+For example, it has been shown that IAA10 has a missense mutation in the EAR domain, which converts the second leucine in the LxLxL motif to a serine (L49S), a mutation likely to reduce the ability of the IAA to recruit TOPLESS co-repressor [@lee_biological_2016].
+This mutation is found in three accessions in the southern UK.
+A second potentially stronger mutation, represented in one accession from northeastern Spain is a missense resulting in an early stop codon in the PB1 domain (need Supp.
+Fig?).
+Such a mutation is likely to prevent IAA10 from interacting with an ARF, rendering it impotent.
+These alleles would be strong candidates for loss-of-function mutants, allowing researchers to discover the function of IAA10.
+In particular, it would make possible a straightforward path for creating an iaa10/11 double mutant using CRISPR/Cas9 directed mutagenesis, allowing for examination of the extent of redundancy between these family members.
 
+Several examples of possible gain-of-function degron missense mutations were found in IAA2, IAA3, IAA6, IAA7, IAA8, IAA13, IAA15, IAA18, IAA19, IAA27, IAA28, IAA29.
+Most degron mutations were found in one or two accessions represented in the database (Supplemental Table or Map?).
+Interestingly, almost all of the IAAs found to have accessions with degron missense mutations have a closely related sister, which may indicate a redundant function within the sister-pairs.
+IAA6 natural variation has two potentially dominant missense mutations found within the degron (Supp Fig.
+alignment), one previously believed to lead to the shy1-1 phenotype [@kim_two_1996].
+However, this variant is widely represented in 31 accessions across europe (Supplemental Map Fig).
+This finding supports the proposition that iaa6[C78R] may not be the mutation causing the shy1-1 phenotype but rather represents natural variation in the degron sequence ([note from Mike Prigge in TAIR](https://www.arabidopsis.org/servlets/TairObject?type=notepad&id=11247)).
+These accessions could be used to confirm whether the shy1-1 phenotype exists within these plants.
+Additionally, an accession was identified with a V77M mutation in the IAA6 degron that may be useful to explore potential iaa6 gain-of-function phenotypes.
+
+###Identifying novel functional sequences in IAAs
+An exciting possible use for the Natural Variation Webtool is to search for conserved, low-diversity sites that may represent possible sites of post-translational modification or protein-protein interaction.
+For example, recent detailed biochemical analysis identified putative lysine ubiquitination sites in members of the IAA6/19 sister pairs [@winkler_variation_2017]; Supp Fig IAA6/19 alignment).
+These sites show no natural sequence variation in Arabidopsis, while several other lysines are represented in the natural sequence diversity (Supp Figs alignment and map).
+In IAA6, there are 17 lysines of which 6 are putative ubiquitination sites, one is in the KR motif, while two others show natural variation.
+In IAA19, there are 15 total lysines, with 8 as putative ubiquitination sites, 2 in the KR motif and degron, and 1 that exhibits natural variation.
+Natural variation sequence diversity may be used to narrow down a subset of lysines to assess for functional ubiquitination.
+
+The members of the IAA8/9/27 group have a unique N-terminal extension with a highly conserved region common to orthologs in several other species [@dreher_arabidopsis_2006].
+IAA8/9 have  redundant roles in leaf serration [@koenig_auxin_2009], while IAA8 important in lateral roots [@arase_iaa8_2012].
+IAA8 has been reported to uniquely localize in the cytosol [@arase_iaa8_2012] where it may interact with LSD1 protein [@coll_programmed_2011].
+The N-terminal extension has a conserved region shared between IAA8 and IAA9 with low natural sequence diversity (Supp Fig IAA8/9 alignment) that may represent a site of novel functionality, possibly mediating interaction with LSD1.
+There are a few accessions with missense alleles in this region in both IAA8 and IAA9 that may be valuable in probing the functionality of this region (Supplemental Fig IAA8/9 map).
+
+Members of the IAA17/7/14/16 group have a second EAR motif (LxLxL) with no observed missense mutations in IAA7/16/17 (Supp Fig IAA7/14/16/17 alignment).
+This observation supports the experimental evidence for IAA7 that the second EAR motif is functional: it can interact with TPR1 corepressor and has a minor repressive role in auxin-related developmental phenotypes [lee_biological_2016].
+For IAA14, we did observe two missense mutations that may be of interest to probe functionality of the IAA14 EAR motif: one in the central “L” (L41F) that is represented in only one accession, and the other in the second “x” (K40N) that is found in 12 accessions around Copenhagen (Supp Fig.
+IAA14 map).
 ## TPL/TPR genes  
@@
 ### Class C ARFs
-Hardik (All figure cite below are in folder "/Dropbox/natural-variation-webtool/manuscript_ARF_C/FIGURES")
-
-Class C ARFs shows low nucleotide diveristy scores, all pi_N/pi_S values substantially lower than 1.
-Comparing diversity within coding region of class C ARF, ARF16  has least diversity whereas its family members (ARF10, ARF17) shows  scores at least four times higher [@remington_contrasting_2004].
-Structurally, all three members of Class C ARFs contain both canonical B3 DNA-binding domain and whereas only ARF10 and ARF16 contains PB1 interaction domain.
-DNA binding domain exhibit overall low diversity and presence of largely synonymous changes for all the ARFs.
-Only ARF10, shows mis-sense variants (p.Ser147Thr/c.439T>A  and p.Ala157Ser/c.469G>T) DNA binding domains with higher diversity score for conservative variation (codon 147) compared to non conservation variation (codon 157) (Fig.
-ARF10_DBD Variants distribution).
-Of the PB1 domain containing class C ARFs, ARF16 exhibits several naturally occurring non-synonymous variants in contrast to ARF10 (Fig.
-Arf16_PB_variants distribution) and is sporadically distributed in location.
-
-Few members of class B ARFs (ARF 2,3,4) and C ARFs (ARF10,16, 17) distinguish from other ARFs on the basis of its ability to be post transcriptionally regulated by trans-acting small interfering RNA (tasiRNA) and microRNA (miRNA) [@rhoades_prediction_2002; @allen_micrornadirected_2005].
-(@AMY Arf 6 and 8 targeted by miRNA 167 - check for conservation).
-Interestingly, miRNA binding motif of the class C ARFs show high degree of conservation across the natural variation of 1001 genomes queried here (Fig.
-ARF10_16_17_Alignment).
-This is consistent with perturbation studies that has shown drastic developmental phenotype for class C ARFs [@wang_control_2005; @mallory_micrornadirected_2005; @liu_repression_2007].
-For the class B ARFs (2,3,4) tasiRNA binding motif is highly conserved and show no variation suggesting strong regulatory constrains at the locus (Fig ARF2_3_4_Alignment).
-DNA binding domain of ARF4 exhibits non-synonymous variation in 4 accession and would be interesting to carry out phenotypic characterize them for developmental growth defects (Fig ARF4_DBD Variants distribution).
+Hardik

### 2018-03-07 manuscript/manuscript.Rmd

--- 2018-03-07 manuscript/manuscript.Rmd
@@
 
-bibliography: ["natural_variation.bib", "TIR1_AFB.bib", "ARF_C.bib"]
+bibliography: ["natural_variation.bib", "TIR1_AFB.bib", "ARF_C.bib", "AuxIAA_citations.bib"]
+
 csl: apa.csl

### 2018-03-07 manuscript/manuscript.Rmd
Alex's manuscript changes

--- 2018-03-07 manuscript/manuscript.Rmd
@@
 
-bibliography: ["natural_variation.bib", "TIR1_AFB.bib", "AuxIAA_citations.bib"]
+bibliography: ["natural_variation.bib", "TIR1_AFB.bib"]
 csl: apa.csl
@@
 
-###Introduction
-The Aux/IAA proteins (IAAs) mediate both auxin perception and transcriptional co-repression of auxin responsive genes.
-Auxin perception is carried out along with TIR1/AFB proteins, which together act as a coreceptor to perceive intracellular auxin.
-Interaction of IAAs with auxin and TIR1/AFBs is conferred by the highly conserved 13 amino acid degron (DII) motif [@tan_mechanism_2007].
-Stabilizing mutations within the degron can lead to dominant phenotypes [@berleth_auxin_2004; @yang_iaa1_2004].
-Degradation of IAAs relieves their repression upon ARF transcription factors.
-IAA degradation rate is influenced both by the degron sequence and by rate motifs found outside the degron, such as the KR motif [@moss_rate_2015; @dreher_arabidopsis_2006].
-IAA-mediated repression of ARFs depends upon an EAR domain that recruits co-repressors of the TOPLESS (TPL) family, and the PB1 domain that enables interaction with ARF PB1 domains [@nanao_structural_2014; @kagale_ear_2011].
-The Arabidopsis Aux/IAA gene family has 29 members, most of which possess a canonical degron.
-Notable exceptions are the Auxin insensitive IAA20 and IAA30 that lack a degron sequence entirely, while the closely related IAA31 retains a semi-functional degron sequence with a very slow degradation rate [@dreher_arabidopsis_2006; @havens_synthetic_2012].
-Using the Natural Variation Webtool, we found that for the canonical Aux/IAAs, no missense mutations were present in the KR motif or PB1 interaction faces, and very few missense mutations were observed in the EAR domain and degron (Supp.
-Fig.
-Alignment of canonical IAAs).
-However, as discussed below, there were several interesting missense mutations found in EAR domains and degrons.
-The majority of nonsynonymous mutations observed were in the variable regions between the conserved domains (Supp.
-Fig.
-Alignment of canonical IAAs).
+### Group A Aux/IAAs
+Britney
 
-###Sister Pair Diversification
-Previous research has pointed to genetic redundancy within closely related pairs or families of IAA proteins [@overvoorde_functional_2005].
-For example, the iaa8-1, iaa9-1 double mutant and the iaa5-1, iaa6-1, iaa19-1 triple mutant have wild-type phenotypes [@overvoorde_functional_2005].
-Despite this, most Arabidopsis IAA sister pairs have one member exhibiting higher diversity while the other sister is more conserved (Figure IAA phylogeny heat map).
-For example, despite having only minor differences in sequence (61.4% sequence identity), the IAA6/19 sister pair has significant differences in expression patterns, protein abundances and functions, suggesting they have undergone functional specialization since their divergence [@winkler_variation_2017].
-IAA19 was expressed at higher levels than IAA6 and showed less sequence divergence; IAA6 showed regions of increased divergence upstream of the degron and in the N-terminal half of the PB1 domain [@winkler_variation_2017].
-Consistent with these results, the Natural Variation Webtool (Figure IAA phylogeny heat map) revealed a lower natural variation for IAA19 (ᴨN/ᴨS = 0.55) compared to IAA6 (ᴨN/ᴨS = 2.3), and also detected high natural variation diversity within the same regions of IAA6 as seen in Winkler et al.
-(Supp Fig IAA6 diversity plot).
-Within IAA Group B, there are four sister pairs--IAA10 and IAA11, IAA12/BDL and IAA13, IAA18 and IAA26/PAP1, and IAA20 and IAA30 (Fig X) [@remington_contrasting_2004].
-Interestingly, among these sister pairs only IAA12 and 13 have noticeable differences in diversity.
-IAA10 and IAA11 show little difference in ℼN/ℼS values, with ℼN/ℼS of 0.80 and 0.67 respectively, implicating the possibility that IAA10 and 11 play redundant roles.
-In rice however, it has been shown that IAA11 gain of function mutants block lateral root initiation [@zhu_gain--function_2012].
-The Arabidopsis EFP browser shows that IAA10 and 11 have almost identical expression patterns, with the exception of IAA10 being more highly expressed in the developing seed, and IAA11 being more highly expressed in root tissue during lateral root development.
-As discussed below, the Natural Variation Webtool provides a potentially expedited route to conducting experiments meant to elucidate the extent of redundancy between family members such as IAA10 and 11.
+### Group B Aux/IAAs
+Mollye/Alex
 
-###Identification of new Loss-of-Function/Gain-of-Function mutations
-To test the extent of redundancy/specialization between sister pairs, the Natural Variation Webtool allows for identifications of accessions that can be potentially used to investigate the phenotypic effects in loss of function variants.
-For example, it has been shown that IAA10 has a missense mutation in the EAR domain, which converts the second leucine in the LxLxL motif to a serine (L49S), a mutation likely to reduce the ability of the IAA to recruit TOPLESS co-repressor [@lee_biological_2016].
-This mutation is found in three accessions in the southern UK.
-A second potentially stronger mutation, represented in one accession from northeastern Spain is a missense resulting in an early stop codon in the PB1 domain (need Supp.
-Fig?).
-Such a mutation is likely to prevent IAA10 from interacting with an ARF, rendering it impotent.
-These alleles would be strong candidates for loss-of-function mutants, allowing researchers to discover the function of IAA10.
-In particular, it would make possible a straightforward path for creating an iaa10/11 double mutant using CRISPR/Cas9 directed mutagenesis, allowing for examination of the extent of redundancy between these family members.
+### Group C Aux/IAAs
+Mega/Alex
 
-Several examples of possible gain-of-function degron missense mutations were found in IAA2, IAA3, IAA6, IAA7, IAA8, IAA13, IAA15, IAA18, IAA19, IAA27, IAA28, IAA29.
-Most degron mutations were found in one or two accessions represented in the database (Supplemental Table or Map?).
-Interestingly, almost all of the IAAs found to have accessions with degron missense mutations have a closely related sister, which may indicate a redundant function within the sister-pairs.
-IAA6 natural variation has two potentially dominant missense mutations found within the degron (Supp Fig.
-alignment), one previously believed to lead to the shy1-1 phenotype [@kim_two_1996].
-However, this variant is widely represented in 31 accessions across europe (Supplemental Map Fig).
-This finding supports the proposition that iaa6[C78R] may not be the mutation causing the shy1-1 phenotype but rather represents natural variation in the degron sequence ([note from Mike Prigge in TAIR](https://www.arabidopsis.org/servlets/TairObject?type=notepad&id=11247)).
-These accessions could be used to confirm whether the shy1-1 phenotype exists within these plants.
-Additionally, an accession was identified with a V77M mutation in the IAA6 degron that may be useful to explore potential iaa6 gain-of-function phenotypes.
+## TPL/TPR genes  
+  The Auxin signaling pathway utilizes the TOPLESS (TPL) family of Gro/TLE/TUP1 type co-repressor proteins to maintain auxin responsive genes in a transcriptionally-repressed state in the absence of auxin (cite).
+In Arabidopsis the TPL family includes TOPLESS and four TPL-related proteins (TPR1-4), which are comprised of three structural domains: an N-terminal TPL domain and two beta-propellar domains.
+TPL is recruited to the AUX/IAA proteins through interaction with a conserved protein motif known as an Ethylene-responsive element binding factor-associated amphiphilic repression (EAR) domain (cite).
+Canonical EAR domains have the amino acid sequence LxLxL, as is found in domain I of the AUX/IAAs (figure xA, cite).
+TPL co-repressors bind EAR domains through direct interaction with the TPL N-terminal C-terminal to LisH (CTLH) domain (citations of pre-structure founding papers/reviews).
+Recent structural analyses of the TPL N-terminal domain have highlighted the precise interaction interface between TPL and AUX/IAA EAR domains, as well as the TPL-TPL dimerization and tetramerization motifs (Avarelilo et al 2017, Ke et al, 2015), and even the residues required for higher-order multimers of TPL tetramers (Ma et al, 2017).
+The two tandem beta propellar domains of TPL/TPRs have as yet not been crystallized, however numerous high resolution crystal structures of beta propellar domains allow for structural prediction of the TPL c-terminal beta propellar domains and the locations of likely interaction faces (Figure, X).
 
-###Identifying novel functional sequences in IAAs
-An exciting possible use for the Natural Variation Webtool is to search for conserved, low-diversity sites that may represent possible sites of post-translational modification or protein-protein interaction.
-For example, recent detailed biochemical analysis identified putative lysine ubiquitination sites in members of the IAA6/19 sister pairs [@winkler_variation_2017]; Supp Fig IAA6/19 alignment).
-These sites show no natural sequence variation in Arabidopsis, while several other lysines are represented in the natural sequence diversity (Supp Figs alignment and map).
-In IAA6, there are 17 lysines of which 6 are putative ubiquitination sites, one is in the KR motif, while two others show natural variation.
-In IAA19, there are 15 total lysines, with 8 as putative ubiquitination sites, 2 in the KR motif and degron, and 1 that exhibits natural variation.
-Natural variation sequence diversity may be used to narrow down a subset of lysines to assess for functional ubiquitination.
+  The TOPLESS co-repressor family generally exhibits a high level of sequence conservation at the amino acid sequence level across the Arabidopsis accessions with all ${\pi_N/\pi_S}$ values below 1 (Table X).
+The closely related TPL and TPR1 have the highest ${\pi_N/\pi_S}$ values (TPL-0.425, TPR1-0.739), suggesting that these these two related genes tolerate a higher degree of sequence diversity compared to TPR2-4.
+The TPL/TPR family demonstrates a high level of sequence conservation across the N-terminal TPL domain, with the lowest of observed ${\pi_N/\pi_S}$ values (Figure X).
+Indeed, all non-synonymous mutations observed in this region are either in linker domains between alpha helices, or are highly conservative mutations within helices (i.e.
+Valine to Isoleucine), which would be predicted to exhibit little effect on TPL folding and activity.
 
-The members of the IAA8/9/27 group have a unique N-terminal extension with a highly conserved region common to orthologs in several other species [@dreher_arabidopsis_2006].
-IAA8/9 have  redundant roles in leaf serration [@koenig_auxin_2009], while IAA8 important in lateral roots [@arase_iaa8_2012].
-IAA8 has been reported to uniquely localize in the cytosol [@arase_iaa8_2012] where it may interact with LSD1 protein [@coll_programmed_2011].
-The N-terminal extension has a conserved region shared between IAA8 and IAA9 with low natural sequence diversity (Supp Fig IAA8/9 alignment) that may represent a site of novel functionality, possibly mediating interaction with LSD1.
-There are a few accessions with missense alleles in this region in both IAA8 and IAA9 that may be valuable in probing the functionality of this region (Supplemental Fig IAA8/9 map).
+  The increased N-terminal conservation underscores the importance of the N-terminal TPL domain.
+For example, the initial tpl-1 mutation (N176H) in the ninth helix is a dominant gain-of-function allele (Long et al, 2002, PMID12050130), which is capable of binding wild-type TPL protein and inducing protein agreggation (Ma, 2017, pmid28630893).
+It is therefore understandable that very little diversity is observed in the ninth alpha helix of TPL/TPRs as mutations in this domain could act in a dominant negative fashion, rendering any de novo mutation in this region embryo lethal.
+However, the entire N-terminal domain exhibits very low rates of nucleotide diversity, suggesting that each domain in this region is critical at the population level.
+Indeed, the crystal structure of TPL has revealed that the TPL domain is responsible for the multiple roles of TPL/TPR proteins: The LisH domain and CRA domain coordinate TPL dimerization and tetramerization respectively (26601214, 28698367), the CTLH domain is critical for interaction with DNA-binding transcription factors (26601214), and helix 9 within the CRA domain may be critical for the formation of higher-order TPL complexes (multimers of TPL tetramers, 28630893).
 
-Members of the IAA17/7/14/16 group have a second EAR motif (LxLxL) with no observed missense mutations in IAA7/16/17 (Supp Fig IAA7/14/16/17 alignment).
-This observation supports the experimental evidence for IAA7 that the second EAR motif is functional: it can interact with TPR1 corepressor and has a minor repressive role in auxin-related developmental phenotypes [lee_biological_2016].
-For IAA14, we did observe two missense mutations that may be of interest to probe functionality of the IAA14 EAR motif: one in the central “L” (L41F) that is represented in only one accession, and the other in the second “x” (K40N) that is found in 12 accessions around Copenhagen (Supp Fig.
-IAA14 map).
-## TPL/TPR genes  
-Alex
+  We then compared the frequency of SNPs that cause nonsynonymous mutations across the TPL domains.
+Compared to the TPL domain, both the proline rich and WD-40 domains exhibit a higher rate of missense mutations accross both of the WD-40 repeat regions, here labeled WD40-1 and WD40-2 respectively.
+We utilized the Phyre2.0 modeling software to generate a rough model of the TPL/TPR WD-40 repeats stricly for the purpose of mapping non-synonymous mutations onto a visible protein chassis (Figure X).
+Here we can see both faces of each predicted beta-propellar domains, which are connected by a short linker region.
+Through utilization of protein modelling, it becomes clear that we observe conservation of amino acid sequence at the center of the beta-propellar domains.
+These observations could be used as the basis for forward mutagensis strategies to test whether this region of TPL functions similarly to its homolog TLE, which binds its target genes EH1 and the C-terminal WRPW/Y motif (Hairy/Hes/Runx) in the center of the beta-propeller domain (pmid 16762837).
+
+  It is interesting to note that despite the presence of 5 TPL/TPR genes in the genome, which are considered to be redundant (pmid 16763149), very few SNPs are observed that are predicted to result in loss of function for any one TPL/TPR gene.
+In fact, only one deletion was observed (665delG) that creates a frameshift mutation in TPR4 in a single accession (Aiell-4 in Italy).
+This mutation occurs in codon 222, just downstream of the TPL N-terminal domain, and would generate an early stop codon 11 codons downstream of the deletion site.
+It is unclear whether this N-terminal fragment would be expressed at any appreciable level, as it is possible that the generation of an early stop codon would trigger nonsense-mediated decay.
+However, if expressed, it would be fascinating to examine whether this naturally occuring N-terminal truncation retains activity as a co-repressor.
+Given the observation that the TPL N-terminal domain is sufficient to bind histones (28630893), and repress transcription via synthetic assays in yeast (24979769), it is possible that this represents a semi-functional TPL variant.
+There has been no evidence for priotization of individual TPL/TPR family members in the auxin response pathway, but experiments modulating the gene copy number of TPR1 revealed the importance of maintaing the proper dosage of TPL/TPR in order to accurately trigger the salicyclic acid innate immune response (20647385).
+These results, taken together with the absence of TPL/TPR loss-of-function alleles in the Arabidopsis accessions suggest that perhaps the entire family is being conserved at the nucleotide level in the population to meet a gene dosage requirement, with little opportunity for neo- or sub-functionalization.
 

### 2018-03-07 manuscript/manuscript.Rmd

--- 2018-03-07 manuscript/manuscript.Rmd
@@
 ## TPL/TPR genes  
-Alex
+  The Auxin signaling pathway utilizes the TOPLESS (TPL) family of Gro/TLE/TUP1 type co-repressor proteins to maintain auxin responsive genes in a transcriptionally-repressed state in the absence of auxin (cite).
+In Arabidopsis the TPL family includes TOPLESS and four TPL-related proteins (TPR1-4), which are comprised of three structural domains: an N-terminal TPL domain and two beta-propellar domains.
+TPL is recruited to the AUX/IAA proteins through interaction with a conserved protein motif known as an Ethylene-responsive element binding factor-associated amphiphilic repression (EAR) domain (cite).
+Canonical EAR domains have the amino acid sequence LxLxL, as is found in domain I of the AUX/IAAs (figure xA, cite).
+TPL co-repressors bind EAR domains through direct interaction with the TPL N-terminal C-terminal to LisH (CTLH) domain (citations of pre-structure founding papers/reviews).
+Recent structural analyses of the TPL N-terminal domain have highlighted the precise interaction interface between TPL and AUX/IAA EAR domains, as well as the TPL-TPL dimerization and tetramerization motifs (Avarelilo et al 2017, Ke et al, 2015), and even the residues required for higher-order multimers of TPL tetramers (Ma et al, 2017).
+The two tandem beta propellar domains of TPL/TPRs have as yet not been crystallized, however numerous high resolution crystal structures of beta propellar domains allow for structural prediction of the TPL c-terminal beta propellar domains and the locations of likely interaction faces (Figure, X).
+
+  The TOPLESS co-repressor family generally exhibits a high level of sequence conservation at the amino acid sequence level across the Arabidopsis accessions with all ${\pi_N/\pi_S}$ values below 1 (Table X).
+The closely related TPL and TPR1 have the highest ${\pi_N/\pi_S}$ values (TPL-0.425, TPR1-0.739), suggesting that these these two related genes tolerate a higher degree of sequence diversity compared to TPR2-4.
+The TPL/TPR family demonstrates a high level of sequence conservation across the N-terminal TPL domain, with the lowest of observed ${\pi_N/\pi_S}$ values (Figure X).
+Indeed, all non-synonymous mutations observed in this region are either in linker domains between alpha helices, or are highly conservative mutations within helices (i.e.
+Valine to Isoleucine), which would be predicted to exhibit little effect on TPL folding and activity.
+
+  The increased N-terminal conservation underscores the importance of the N-terminal TPL domain.
+For example, the initial tpl-1 mutation (N176H) in the ninth helix is a dominant gain-of-function allele (Long et al, 2002, PMID12050130), which is capable of binding wild-type TPL protein and inducing protein agreggation (Ma, 2017, pmid28630893).
+It is therefore understandable that very little diversity is observed in the ninth alpha helix of TPL/TPRs as mutations in this domain could act in a dominant negative fashion, rendering any de novo mutation in this region embryo lethal.
+However, the entire N-terminal domain exhibits very low rates of nucleotide diversity, suggesting that each domain in this region is critical at the population level.
+Indeed, the crystal structure of TPL has revealed that the TPL domain is responsible for the multiple roles of TPL/TPR proteins: The LisH domain and CRA domain coordinate TPL dimerization and tetramerization respectively (26601214, 28698367), the CTLH domain is critical for interaction with DNA-binding transcription factors (26601214), and helix 9 within the CRA domain may be critical for the formation of higher-order TPL complexes (multimers of TPL tetramers, 28630893).
+
+  We then compared the frequency of SNPs that cause nonsynonymous mutations across the TPL domains.
+Compared to the TPL domain, both the proline rich and WD-40 domains exhibit a higher rate of missense mutations accross both of the WD-40 repeat regions, here labeled WD40-1 and WD40-2 respectively.
+We utilized the Phyre2.0 modeling software to generate a rough model of the TPL/TPR WD-40 repeats stricly for the purpose of mapping non-synonymous mutations onto a visible protein chassis (Figure X).
+Here we can see both faces of each predicted beta-propellar domains, which are connected by a short linker region.
+Through utilization of protein modelling, it becomes clear that we observe conservation of amino acid sequence at the center of the beta-propellar domains.
+These observations could be used as the basis for forward mutagensis strategies to test whether this region of TPL functions similarly to its homolog TLE, which binds its target genes EH1 and the C-terminal WRPW/Y motif (Hairy/Hes/Runx) in the center of the beta-propeller domain (pmid 16762837).
+
+  It is interesting to note that despite the presence of 5 TPL/TPR genes in the genome, which are considered to be redundant (pmid 16763149), very few SNPs are observed that are predicted to result in loss of function for any one TPL/TPR gene.
+In fact, only one deletion was observed (665delG) that creates a frameshift mutation in TPR4 in a single accession (Aiell-4 in Italy).
+This mutation occurs in codon 222, just downstream of the TPL N-terminal domain, and would generate an early stop codon 11 codons downstream of the deletion site.
+It is unclear whether this N-terminal fragment would be expressed at any appreciable level, as it is possible that the generation of an early stop codon would trigger nonsense-mediated decay.
+However, if expressed, it would be fascinating to examine whether this naturally occuring N-terminal truncation retains activity as a co-repressor.
+Given the observation that the TPL N-terminal domain is sufficient to bind histones (28630893), and repress transcription via synthetic assays in yeast (24979769), it is possible that this represents a semi-functional TPL variant.
+There has been no evidence for priotization of individual TPL/TPR family members in the auxin response pathway, but experiments modulating the gene copy number of TPR1 revealed the importance of maintaing the proper dosage of TPL/TPR in order to accurately trigger the salicyclic acid innate immune response (20647385).
+These results, taken together with the absence of TPL/TPR loss-of-function alleles in the Arabidopsis accessions suggest that perhaps the entire family is being conserved at the nucleotide level in the population to meet a gene dosage requirement, with little opportunity for neo- or sub-functionalization.
 

### 2018-03-07 manuscript/manuscript.Rmd
add Alex's figure template

@@
 IAA14 map).
+
 ## TPL/TPR genes
+
   The Auxin signaling pathway utilizes the TOPLESS (TPL) family of Gro/TLE/TUP1 type co-repressor proteins to maintain auxin responsive genes in a transcriptionally-repressed state in the absence of auxin (cite).

### 2018-03-07 manuscript/manuscript.Rmd
Romi's AFB section

--- 2018-03-07 manuscript/manuscript.Rmd
@@
 [@gray_identification_1999; @kepinski_auxininduced_2004; @tan_mechanism_2007; @dezfulian_oligomerization_2016; @wu_evolution_2017b]
+@Terrile_Nitric_2012 nitrosylation residues C140 important in SCFTIR1 dimerization and interaction with Aux/IAAs
+@Wright_Insights_2017 residues in TIR1 and AFB2 relevant in their degradation of Aux/IAAs.
+@Yu_Untethering_2015 E12K E15K residues important in TIR1s interaction with CUL1 (SCF complex).
+@tan_mechanism_2007 R401 R484 K485 K74 K113 R114 R403 R344 All important residues binding phytate.
+R403 hydrogen bonds with auxin compounds
+E487 R489 I488 residues grab the non-polar rings of auxin in place. convex wall part.
+I107 P409 F465 F380 C405 P347 P350 F351 mostly non-polar residues interacting with IAAs. Part of Loop 14, 12, and 2.
+G142A, F143A, I151A, V117A and G147A mutants dimerization
 
+Auxin acts by binding to receptors that in turn target co-repressors for degradation.
+Auxin receptors (six in the model plant Arabidopsis thaliana) evolved through gene duplication and diversification early in the history of vascular plants @Parry_Complex_2009.
+The Nemhauser lab has recently shown that the rate of co-repressor degradation is determined by the identity of both the receptor and co-repressor @Havens_Synthetic_2012, and that this rate sets the pace of development3.
+Although it is unknown whether different receptors play different roles at the organismal level, a number of studies have pointed out differences in the ways the six different receptors in A.
+thaliana differ in their expression.
+The expression of AFB5 is induced in the presence of auxin, making this a circadian-regulated gene @Covington_Circadian_2007.
+It also appears AFB3 is more highly expressed in the roots in the presence of nitrate, allowing increased lateral root formation @Vidal_Nitrate_2010.
+Although TIR1/AFBs are expressed ubiquitously in A.
+thaliana tissues, TIR1, AFB2, and AFB3 have been shown to accumulate in the shoot and root meristems and leaf tissues, with slightly different expression patterns in TIR1 @Parry_Complex_2009.
+AFB1 has been found to have a mutation drastically reducing its ability to assemble into an SCF complex, contributing to it’s increased accumulation, as well as preventing it from degrading Aux/IAAs.
+Higher order TIR1/AFB mutants show AFB1 to have inconclusive organismal function.
+However, AFB1 is shown to be able to accumulate ubiquitously in all tissues @Yu_Untethering_2015.
+When it comes to binding auxins, AFB4 and AFB5 have been shown to preferentially and functionally bind the synthetic auxin picloram, while others have a preference for naturally occurring auxin 3-IAA @Prigge_Arabidopsis_2016.
+Interestingly, the ways in which TIR1/AFBs are able to bind and mark Aux/IAAs for degradation are variable.
+TIR1 and AFB2 have slightly different speed at which they perform this function, with AFB2 being generally quicker, as shown by our lab’s yeast degradation system @Havens_Synthetic_2012.
+Finally, higher order TIR1/AFB mutants show AFB1 to have inconclusive organismal function.
+  Using the Natural Variation Webtool developed in our lab, I will characterize the natural variation in the TIR1/AFB family of auxin receptors by looking at the genomes of 1001 accessions of A.
+thaliana.
+This webtool allows me to	quantify and compare the nucleotide diversity of different genes in this family, as well as find relevant potentially function-altering variants.
+Using its diversity statistic function has shown me TIR1 and AFB1 both seem to be low diversity, hinting at their likely functional importance.
+AFB3 and 4 appear to be higher diversity, while their sisters are more conserved.
+This matches our current understanding of AFB3 and 4 as being minor players in the auxin signaling pathway, but brings up questions about the function of AFB1, which is inconclusive.
+Using this webtool, the AFB1 sequence can be aligned to the highly characterized TIR1 AA sequence showing locations of known non-synonymous variants.
+Although the most important functional regions do not contain variants in AFB1, there is a low diversity mutant, only found in the Can-1 accession, which is in the highly conserved TIR1/AFB oligomerization domain.
+Because variants eliciting significant changes in function often are not selected for, the fact that this variant is extremely low diversity holds promise that it is associated with a strong auxin signaling phenotype.
+I will order and characterize this accession in order to identify potential auxin signaling phenotypes.
+This would support that AFB1 functions through oligomerization with other TIR1/AFBs.
+  The AFB4 and AFB4 receptors are known to have an extension in their N-terminal domains.
+Using the alignment feature on this webtool has indicated this region to be the highest in diversity in these proteins, containing lots of missense variants (Fig.
+4).
+This suggests that this extension does not play an important functional role in these proteins.
+
+Figure 4: Alignment of A.
+thaliana TIR1/AFB F-box domains showing variants.
+Cyan marks the Fbox binding domain.
+Yellow, green, and magenta, mark missense and synonymous, synonymous only, and missense only variant locations respectively.
+Yellow squared E’s show Cul1-interacting residues, differing in AFB1.
+
+@Parry_Complex_2009
+1.	Parry G, Calderon-Villalobos LI, Prigge M, et al.
+Complex regulation of the TIR1/AFB family of auxin receptors.
+Proceedings of the National Academy of Sciences of the United States of America.
+2009;106(52):22540-22545.
+doi:10.1073/pnas.0911967106.
+@Havens_Synthetic_2012
+2.	Kyle A. Havens, et al. A Synthetic Approach Reveals Extensive Tunability of Auxin Signaling. Plant Physiology Sep 2012, 160 (1) 135-142; doi: 10.1104/pp.112.202184
+@Guseman_Auxin-induced_2015
+3.	Guseman JM, et al. Auxin-induced degradation dynamics set the pace for lateral root development. Development (Cambridge, England). 2015;142(5):905-909. doi:10.1242/dev.117234.
+@Covington_Circadian_2007
+8.	Covington MF, Harmer SL. The Circadian Clock Regulates Auxin Signaling and Responses in Arabidopsis . Weigel D, ed. PLoS Biology. 2007;5(8):e222. doi:10.1371/journal.pbio.0050222.
+@Vidal_Nitrate_2010
+9.	Vidal EA, Araus V, Lu C, et al.
+Nitrate-responsive miR393/AFB3 regulatory module controls root system architecture in Arabidopsis thaliana.
+Proceedings of the National Academy of Sciences of the United States of America.
+2010;107(9):4477-4482.
+doi:10.1073/pnas.0909571107.
+@Prigge_Arabidopsis_2016
+10.	Prigge MJ, Greenham K, Zhang Y, et al.
+The Arabidopsis Auxin Receptor F-Box Proteins AFB4 and AFB5 Are Required for Response to the Synthetic Auxin Picloram.
+G3: Genes|Genomes|Genetics.
+2016;6(5):1383-1390.
+doi:10.1534/g3.115.025585.
+@Yu_Untethering_2015
+11.	Yu H, Zhang Y, Moss BL, et al.
+Untethering the TIR1 auxin receptor from the SCF complex increases its stability and inhibits auxin response.
+Nature plants.
+2015;1(3):14030.
+doi:10.1038/nplants.2014.30.
 ## Aux/IAA genes
@@
 ## TPL/TPR genes  
-  The Auxin signaling pathway utilizes the TOPLESS (TPL) family of Gro/TLE/TUP1 type co-repressor proteins to maintain auxin responsive genes in a transcriptionally-repressed state in the absence of auxin (cite).
-In Arabidopsis the TPL family includes TOPLESS and four TPL-related proteins (TPR1-4), which are comprised of three structural domains: an N-terminal TPL domain and two beta-propellar domains.
-TPL is recruited to the AUX/IAA proteins through interaction with a conserved protein motif known as an Ethylene-responsive element binding factor-associated amphiphilic repression (EAR) domain (cite).
-Canonical EAR domains have the amino acid sequence LxLxL, as is found in domain I of the AUX/IAAs (figure xA, cite).
-TPL co-repressors bind EAR domains through direct interaction with the TPL N-terminal C-terminal to LisH (CTLH) domain (citations of pre-structure founding papers/reviews).
-Recent structural analyses of the TPL N-terminal domain have highlighted the precise interaction interface between TPL and AUX/IAA EAR domains, as well as the TPL-TPL dimerization and tetramerization motifs (Avarelilo et al 2017, Ke et al, 2015), and even the residues required for higher-order multimers of TPL tetramers (Ma et al, 2017).
-The two tandem beta propellar domains of TPL/TPRs have as yet not been crystallized, however numerous high resolution crystal structures of beta propellar domains allow for structural prediction of the TPL c-terminal beta propellar domains and the locations of likely interaction faces (Figure, X).
-
-  The TOPLESS co-repressor family generally exhibits a high level of sequence conservation at the amino acid sequence level across the Arabidopsis accessions with all ${\pi_N/\pi_S}$ values below 1 (Table X).
-The closely related TPL and TPR1 have the highest ${\pi_N/\pi_S}$ values (TPL-0.425, TPR1-0.739), suggesting that these these two related genes tolerate a higher degree of sequence diversity compared to TPR2-4.
-The TPL/TPR family demonstrates a high level of sequence conservation across the N-terminal TPL domain, with the lowest of observed ${\pi_N/\pi_S}$ values (Figure X).
-Indeed, all non-synonymous mutations observed in this region are either in linker domains between alpha helices, or are highly conservative mutations within helices (i.e.
-Valine to Isoleucine), which would be predicted to exhibit little effect on TPL folding and activity.
-
-  The increased N-terminal conservation underscores the importance of the N-terminal TPL domain.
-For example, the initial tpl-1 mutation (N176H) in the ninth helix is a dominant gain-of-function allele (Long et al, 2002, PMID12050130), which is capable of binding wild-type TPL protein and inducing protein agreggation (Ma, 2017, pmid28630893).
-It is therefore understandable that very little diversity is observed in the ninth alpha helix of TPL/TPRs as mutations in this domain could act in a dominant negative fashion, rendering any de novo mutation in this region embryo lethal.
-However, the entire N-terminal domain exhibits very low rates of nucleotide diversity, suggesting that each domain in this region is critical at the population level.
-Indeed, the crystal structure of TPL has revealed that the TPL domain is responsible for the multiple roles of TPL/TPR proteins: The LisH domain and CRA domain coordinate TPL dimerization and tetramerization respectively (26601214, 28698367), the CTLH domain is critical for interaction with DNA-binding transcription factors (26601214), and helix 9 within the CRA domain may be critical for the formation of higher-order TPL complexes (multimers of TPL tetramers, 28630893).
-
-  We then compared the frequency of SNPs that cause nonsynonymous mutations across the TPL domains.
-Compared to the TPL domain, both the proline rich and WD-40 domains exhibit a higher rate of missense mutations accross both of the WD-40 repeat regions, here labeled WD40-1 and WD40-2 respectively.
-We utilized the Phyre2.0 modeling software to generate a rough model of the TPL/TPR WD-40 repeats stricly for the purpose of mapping non-synonymous mutations onto a visible protein chassis (Figure X).
-Here we can see both faces of each predicted beta-propellar domains, which are connected by a short linker region.
-Through utilization of protein modelling, it becomes clear that we observe conservation of amino acid sequence at the center of the beta-propellar domains.
-These observations could be used as the basis for forward mutagensis strategies to test whether this region of TPL functions similarly to its homolog TLE, which binds its target genes EH1 and the C-terminal WRPW/Y motif (Hairy/Hes/Runx) in the center of the beta-propeller domain (pmid 16762837).
-
-  It is interesting to note that despite the presence of 5 TPL/TPR genes in the genome, which are considered to be redundant (pmid 16763149), very few SNPs are observed that are predicted to result in loss of function for any one TPL/TPR gene.
-In fact, only one deletion was observed (665delG) that creates a frameshift mutation in TPR4 in a single accession (Aiell-4 in Italy).
-This mutation occurs in codon 222, just downstream of the TPL N-terminal domain, and would generate an early stop codon 11 codons downstream of the deletion site.
-It is unclear whether this N-terminal fragment would be expressed at any appreciable level, as it is possible that the generation of an early stop codon would trigger nonsense-mediated decay.
-However, if expressed, it would be fascinating to examine whether this naturally occuring N-terminal truncation retains activity as a co-repressor.
-Given the observation that the TPL N-terminal domain is sufficient to bind histones (28630893), and repress transcription via synthetic assays in yeast (24979769), it is possible that this represents a semi-functional TPL variant.
-There has been no evidence for priotization of individual TPL/TPR family members in the auxin response pathway, but experiments modulating the gene copy number of TPR1 revealed the importance of maintaing the proper dosage of TPL/TPR in order to accurately trigger the salicyclic acid innate immune response (20647385).
-These results, taken together with the absence of TPL/TPR loss-of-function alleles in the Arabidopsis accessions suggest that perhaps the entire family is being conserved at the nucleotide level in the population to meet a gene dosage requirement, with little opportunity for neo- or sub-functionalization.
+Alex
 

### 2018-03-07 manuscript/manuscript.Rmd

--- 2018-03-07 manuscript/manuscript.Rmd
@@
 ## TIR1/AFB genes
-Romi
-[@gray_identification_1999; @kepinski_auxininduced_2004; @tan_mechanism_2007; @dezfulian_oligomerization_2016; @wu_evolution_2017b]
-1) Plant biologist of all types should use this web tool!
 
->Makes it really easy to look at the evolution of genes in arabidopsis.
->Makes it really easy to look at how families of genes are evolving.
->Easy way to find accessions that might be interesting to you.
+Auxin acts by binding to receptors that in turn target co-repressors for degradation.
+Auxin receptors (six in the model plant Arabidopsis thaliana) evolved through gene duplication and diversification early in the history of vascular plants @Parry_Complex_2009.
+The Nemhauser lab has recently shown that the rate of co-repressor degradation is determined by the identity of both the receptor and co-repressor @Havens_Synthetic_2012, and that this rate sets the pace of development3.
+Although it is unknown whether different receptors play different roles at the organismal level, a number of studies have pointed out differences in the ways the six different receptors in A.
+thaliana differ in their expression.
+The expression of AFB5 is induced in the presence of auxin, making this a circadian-regulated gene @Covington_Circadian_2007.
+It also appears AFB3 is more highly expressed in the roots in the presence of nitrate, allowing increased lateral root formation @Vidal_Nitrate_2010.
+Although TIR1/AFBs are expressed ubiquitously in A.
+thaliana tissues, TIR1, AFB2, and AFB3 have been shown to accumulate in the shoot and root meristems and leaf tissues, with slightly different expression patterns in TIR1 @Parry_Complex_2009.
+AFB1 has been found to have a mutation drastically reducing its ability to assemble into an SCF complex, contributing to it’s increased accumulation, as well as preventing it from degrading Aux/IAAs.
+Higher order TIR1/AFB mutants show AFB1 to have inconclusive organismal function.
+However, AFB1 is shown to be able to accumulate ubiquitously in all tissues @Yu_Untethering_2015.
+When it comes to binding auxins, AFB4 and AFB5 have been shown to preferentially and functionally bind the synthetic auxin picloram, while others have a preference for naturally occurring auxin 3-IAA @Prigge_Arabidopsis_2016.
+Interestingly, the ways in which TIR1/AFBs are able to bind and mark Aux/IAAs for degradation are variable.
+TIR1 and AFB2 have slightly different speed at which they perform this function, with AFB2 being generally quicker, as shown by our lab’s yeast degradation system @Havens_Synthetic_2012.
+Finally, higher order TIR1/AFB mutants show AFB1 to have inconclusive organismal function.
+  Using the Natural Variation Webtool developed in our lab, I will characterize the natural variation in the TIR1/AFB family of auxin receptors by looking at the genomes of 1001 accessions of A.
+thaliana.
+This webtool allows me to	quantify and compare the nucleotide diversity of different genes in this family, as well as find relevant potentially function-altering variants.
+Using its diversity statistic function has shown me TIR1 and AFB1 both seem to be low diversity, hinting at their likely functional importance.
+AFB3 and 4 appear to be higher diversity, while their sisters are more conserved.
+This matches our current understanding of AFB3 and 4 as being minor players in the auxin signaling pathway, but brings up questions about the function of AFB1, which is inconclusive.
+Using this webtool, the AFB1 sequence can be aligned to the highly characterized TIR1 AA sequence showing locations of known non-synonymous variants.
+Although the most important functional regions do not contain variants in AFB1, there is a low diversity mutant, only found in the Can-1 accession, which is in the highly conserved TIR1/AFB oligomerization domain.
+Because variants eliciting significant changes in function often are not selected for, the fact that this variant is extremely low diversity holds promise that it is associated with a strong auxin signaling phenotype.
+I will order and characterize this accession in order to identify potential auxin signaling phenotypes.
+This would support that AFB1 functions through oligomerization with other TIR1/AFBs.
+  The AFB4 and AFB4 receptors are known to have an extension in their N-terminal domains.
+Using the alignment feature on this webtool has indicated this region to be the highest in diversity in these proteins, containing lots of missense variants (Fig.
+4).
+This suggests that this extension does not play an important functional role in these proteins.
 
-2) This web tool can be used to identify domains or residues of potential functional importance in any gene or gene family. (E.g. stretches of conservation in regions of unknown function).
+Figure 4: Alignment of A.
+thaliana TIR1/AFB F-box domains showing variants.
+Cyan marks the Fbox binding domain.
+Yellow, green, and magenta, mark missense and synonymous, synonymous only, and missense only variant locations respectively.
+Yellow squared E’s show Cul1-interacting residues, differing in AFB1.
 
->I was able to find that domains of iaa and phytate interaction are highly conserved, which is what we would expect.
->I was able to show that extra flanking domains of AFB4 and 5 are very mutated and therefore probably not important.
-
-3) It can help identify sites of relevant functional divergence between sister pairs and more broadly within gene families (E.g.
-one sister has much higher diversity than the other.
-What are the shared regions of variation?
-What regions are highly variable in the more diverse sister?)
-
->AFB4 the most diverse sister. All regions highly variable. Iaa binding mutated. likely not functional at all. All have higher variation before fbox binding domain and before dimerization domain.
-
-4) It can also identify accessions containing multiple nonsynonymous genes in a family or network that can be used for studying network evolution.
-(E.g.
-there is a nonsynonymous polymorphism in a key functional residue of a gene that you expect will alter its function, however there are many other polymorphisms in other genes in the pathway that may or may not compensate for this change.
-We can assess this by backcrossing this accession with Col-0 or other distantly related accessions and resequencing phenotypic segregants).
-
->Did not find this yet. None seem to be in a critical site, so I will be looking later on.
+@Parry_Complex_2009
+1.	Parry G, Calderon-Villalobos LI, Prigge M, et al.
+Complex regulation of the TIR1/AFB family of auxin receptors.
+Proceedings of the National Academy of Sciences of the United States of America.
+2009;106(52):22540-22545.
+doi:10.1073/pnas.0911967106.
+@Havens_Synthetic_2012
+2.	Kyle A. Havens, et al. A Synthetic Approach Reveals Extensive Tunability of Auxin Signaling. Plant Physiology Sep 2012, 160 (1) 135-142; doi: 10.1104/pp.112.202184
+@Guseman_Auxin-induced_2015
+3.	Guseman JM, et al. Auxin-induced degradation dynamics set the pace for lateral root development. Development (Cambridge, England). 2015;142(5):905-909. doi:10.1242/dev.117234.
+@Covington_Circadian_2007
+8.	Covington MF, Harmer SL. The Circadian Clock Regulates Auxin Signaling and Responses in Arabidopsis . Weigel D, ed. PLoS Biology. 2007;5(8):e222. doi:10.1371/journal.pbio.0050222.
+@Vidal_Nitrate_2010
+9.	Vidal EA, Araus V, Lu C, et al.
+Nitrate-responsive miR393/AFB3 regulatory module controls root system architecture in Arabidopsis thaliana.
+Proceedings of the National Academy of Sciences of the United States of America.
+2010;107(9):4477-4482.
+doi:10.1073/pnas.0909571107.
+@Prigge_Arabidopsis_2016
+10.	Prigge MJ, Greenham K, Zhang Y, et al.
+The Arabidopsis Auxin Receptor F-Box Proteins AFB4 and AFB5 Are Required for Response to the Synthetic Auxin Picloram.
+G3: Genes|Genomes|Genetics.
+2016;6(5):1383-1390.
+doi:10.1534/g3.115.025585.
+@Yu_Untethering_2015
+11.	Yu H, Zhang Y, Moss BL, et al.
+Untethering the TIR1 auxin receptor from the SCF complex increases its stability and inhibits auxin response.
+Nature plants.
+2015;1(3):14030.
+doi:10.1038/nplants.2014.30.
+>>>>>>> AFB
 ## Aux/IAA genes

### 2018-03-07 manuscript/manuscript.Rmd
Morgan's methods section

--- 2018-03-07 manuscript/manuscript.Rmd
@@
 ---
-title: "A comprehensive analysis of natural sequence variation within the <i>Arabidopsis thaliana</i> nuclear auxin signaling pathway"
-author: 
-  affiliation: University of Washington
-  name: "R. Clay Wright"
-
-date: "`r Sys.Date()`"
-
-output: 
-    html_document:
-      toc: true
-      fig_caption: yes
-
-keywords: r markdown
-fontsize: 11pt
-fontfamily: Helvetica
-
-bibliography: ["natural_variation.bib", "TIR1_AFB.bib"]
-csl: apa.csl
-link-citations: true
+title: "Untitled"
+author: "Morgan Hamm"
+date: "March 4, 2018"
+output: html_document
 ---
 
-```{r, echo = FALSE, message=FALSE, warning=FALSE}
-knitr::opts_chunk$set(warning = FALSE, collapse = TRUE, comment = "#>", tidy = TRUE, out.width = "90%", fig.height=4, fig.width=7, out.extra='style="margin: auto; display: block; padding-top: 15px;"')
-library(r1001genomes)
-library(DT)
-library(dplyr)
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
+```
 
-```
-<!--This is a CSS (Cascading Style Sheet).
- It controls how panels and images are formatted.></!-->
-<style>
-  img {
-    display: block;
-    margin: auto;
-    width: 90%;
-    height: auto;
-  }
-</style>
+## Nucleotide Diversity Calculation
+Nei  defined the nucleotide diversity statistic in his original paper as:
+“…the average number of nucleotide differences per site between two randomly chosen DNA sequences.” (Nei, 1979)
+And provided the equation:
 
-# List of working titles  
-* A comprehensive analysis of natural sequence variation within the <i>Arabidopsis thaliana</i> nuclear auxin signaling pathway  
-* *Add your suggestions*
+$\displaystyle \pi = \sum_{ij}x_{i}x_{j}\pi_{ij}$
 
-# Abstract
+Where Where $x_{i}$ is the frequency of the ith sequence in the population and $\pi_{ij}$ is the number of sites that are different between the ith and jth sequence divided by sequence length.
 
-* Motivating, high level statement  
-* Statement defining the cutting edge of your research area  
-* A statement of the form: "However, it remains unclear ..." that defines what problem you are  
-addressing.  
-* A statement of the form:"Here, we ..."  that concisely states your contribution.  
-* A series of 3?5 statement about your results.  
-* A concluding statement.
+a more general form, that treats each sequence in the population as unique can be written as:
 
-# Related work  
-What and how have we, as a field, learned from natural variation?  
-[@hollister_genomic_2014; @bielawski_detecting_2013; @kryazhimskiy_population_2008; @schmid_multilocus_2005; @nieduszynski_sequence_2011; @melamed_combining_2015; @weigel_population_2015; @schmitz_patterns_2013; @weigel_natural_2005; @hardtke_case_2010; @koornneef_naturally_2004; @mitchell-olds_genetic_2006; @joly-lopez_developing_2016; @wright_insights_2017a; @park_distributions_2017; @_135_2016a]
+$\displaystyle \pi = \frac{1}{L*n^2} \sum_{i=1}^n \sum_{j=1}^n \sum_{k=1}^L{\pi_{ijk}}
+\qquad
+\displaystyle \pi_{ijk} = \left\{\begin{array}{l}
+1 \quad \textrm{if} \quad N_{ik} \neq N_{jk} \\
+0 \quad \textrm{if} \quad N_{ik} = N_{jk}
+\end{array}\right.$
 
-# Methods
+where $N_{ik}$ is the nucelotide (A, T, C or G) at position k on the ith sequence of the population.
+$L$ is the lenght of the sequence (the same for all sequences in the population, 
+assumes only SNPs exist, no indels). n is the total number of sequences in the population.
 
-# Results  
-## TIR1/AFB genes  
-Romi
-[@gray_identification_1999; @kepinski_auxininduced_2004; @tan_mechanism_2007; @dezfulian_oligomerization_2016; @wu_evolution_2017b]
-@Terrile_Nitric_2012 nitrosylation residues C140 important in SCFTIR1 dimerization and interaction with Aux/IAAs
-@Wright_Insights_2017 residues in TIR1 and AFB2 relevant in their degradation of Aux/IAAs.
-@Yu_Untethering_2015 E12K E15K residues important in TIR1s interaction with CUL1 (SCF complex).
-@tan_mechanism_2007 R401 R484 K485 K74 K113 R114 R403 R344 All important residues binding phytate.
-R403 hydrogen bonds with auxin compounds
-E487 R489 I488 residues grab the non-polar rings of auxin in place. convex wall part.
-I107 P409 F465 F380 C405 P347 P350 F351 mostly non-polar residues interacting with IAAs. Part of Loop 14, 12, and 2.
-G142A, F143A, I151A, V117A and G147A mutants dimerization
+from this form we can re-arange summations to the form below:
 
-Auxin acts by binding to receptors that in turn target co-repressors for degradation.
-Auxin receptors (six in the model plant Arabidopsis thaliana) evolved through gene duplication and diversification early in the history of vascular plants @Parry_Complex_2009.
-The Nemhauser lab has recently shown that the rate of co-repressor degradation is determined by the identity of both the receptor and co-repressor @Havens_Synthetic_2012, and that this rate sets the pace of development3.
-Although it is unknown whether different receptors play different roles at the organismal level, a number of studies have pointed out differences in the ways the six different receptors in A.
-thaliana differ in their expression.
-The expression of AFB5 is induced in the presence of auxin, making this a circadian-regulated gene @Covington_Circadian_2007.
-It also appears AFB3 is more highly expressed in the roots in the presence of nitrate, allowing increased lateral root formation @Vidal_Nitrate_2010.
-Although TIR1/AFBs are expressed ubiquitously in A.
-thaliana tissues, TIR1, AFB2, and AFB3 have been shown to accumulate in the shoot and root meristems and leaf tissues, with slightly different expression patterns in TIR1 @Parry_Complex_2009.
-AFB1 has been found to have a mutation drastically reducing its ability to assemble into an SCF complex, contributing to it’s increased accumulation, as well as preventing it from degrading Aux/IAAs.
-Higher order TIR1/AFB mutants show AFB1 to have inconclusive organismal function.
-However, AFB1 is shown to be able to accumulate ubiquitously in all tissues @Yu_Untethering_2015.
-When it comes to binding auxins, AFB4 and AFB5 have been shown to preferentially and functionally bind the synthetic auxin picloram, while others have a preference for naturally occurring auxin 3-IAA @Prigge_Arabidopsis_2016.
-Interestingly, the ways in which TIR1/AFBs are able to bind and mark Aux/IAAs for degradation are variable.
-TIR1 and AFB2 have slightly different speed at which they perform this function, with AFB2 being generally quicker, as shown by our lab’s yeast degradation system @Havens_Synthetic_2012.
-Finally, higher order TIR1/AFB mutants show AFB1 to have inconclusive organismal function.
-  Using the Natural Variation Webtool developed in our lab, I will characterize the natural variation in the TIR1/AFB family of auxin receptors by looking at the genomes of 1001 accessions of A.
-thaliana.
-This webtool allows me to	quantify and compare the nucleotide diversity of different genes in this family, as well as find relevant potentially function-altering variants.
-Using its diversity statistic function has shown me TIR1 and AFB1 both seem to be low diversity, hinting at their likely functional importance.
-AFB3 and 4 appear to be higher diversity, while their sisters are more conserved.
-This matches our current understanding of AFB3 and 4 as being minor players in the auxin signaling pathway, but brings up questions about the function of AFB1, which is inconclusive.
-Using this webtool, the AFB1 sequence can be aligned to the highly characterized TIR1 AA sequence showing locations of known non-synonymous variants.
-Although the most important functional regions do not contain variants in AFB1, there is a low diversity mutant, only found in the Can-1 accession, which is in the highly conserved TIR1/AFB oligomerization domain.
-Because variants eliciting significant changes in function often are not selected for, the fact that this variant is extremely low diversity holds promise that it is associated with a strong auxin signaling phenotype.
-I will order and characterize this accession in order to identify potential auxin signaling phenotypes.
-This would support that AFB1 functions through oligomerization with other TIR1/AFBs.
-  The AFB4 and AFB4 receptors are known to have an extension in their N-terminal domains.
-Using the alignment feature on this webtool has indicated this region to be the highest in diversity in these proteins, containing lots of missense variants (Fig.
-4).
-This suggests that this extension does not play an important functional role in these proteins.
+$\displaystyle \pi = \frac{1}{L} \sum_{k=1}^L{\pi_{k}}
+\qquad
+\pi_{k} = \frac{1}{n^2} \sum_{i=1}^n \sum_{j=1}^n {\pi_{ijk}}$
 
-Figure 4: Alignment of A.
-thaliana TIR1/AFB F-box domains showing variants.
-Cyan marks the Fbox binding domain.
-Yellow, green, and magenta, mark missense and synonymous, synonymous only, and missense only variant locations respectively.
-Yellow squared E’s show Cul1-interacting residues, differing in AFB1.
+here $\pi_k$ can be thought of as the site-wise nucleotide diversity at postion k, and is equal to 
+the nucleotide diversity of a sequence of lenght 1 at location k. we can calculate $\pi_k$ for each
+site, then average those over the sequence lenght to calculate $\pi$, the nucleotide diversity 
+of the sequence.
 
-@Parry_Complex_2009
-1.	Parry G, Calderon-Villalobos LI, Prigge M, et al.
-Complex regulation of the TIR1/AFB family of auxin receptors.
-Proceedings of the National Academy of Sciences of the United States of America.
-2009;106(52):22540-22545.
-doi:10.1073/pnas.0911967106.
-@Havens_Synthetic_2012
-2.	Kyle A. Havens, et al. A Synthetic Approach Reveals Extensive Tunability of Auxin Signaling. Plant Physiology Sep 2012, 160 (1) 135-142; doi: 10.1104/pp.112.202184
-@Guseman_Auxin-induced_2015
-3.	Guseman JM, et al. Auxin-induced degradation dynamics set the pace for lateral root development. Development (Cambridge, England). 2015;142(5):905-909. doi:10.1242/dev.117234.
-@Covington_Circadian_2007
-8.	Covington MF, Harmer SL. The Circadian Clock Regulates Auxin Signaling and Responses in Arabidopsis . Weigel D, ed. PLoS Biology. 2007;5(8):e222. doi:10.1371/journal.pbio.0050222.
-@Vidal_Nitrate_2010
-9.	Vidal EA, Araus V, Lu C, et al.
-Nitrate-responsive miR393/AFB3 regulatory module controls root system architecture in Arabidopsis thaliana.
-Proceedings of the National Academy of Sciences of the United States of America.
-2010;107(9):4477-4482.
-doi:10.1073/pnas.0909571107.
-@Prigge_Arabidopsis_2016
-10.	Prigge MJ, Greenham K, Zhang Y, et al.
-The Arabidopsis Auxin Receptor F-Box Proteins AFB4 and AFB5 Are Required for Response to the Synthetic Auxin Picloram.
-G3: Genes|Genomes|Genetics.
-2016;6(5):1383-1390.
-doi:10.1534/g3.115.025585.
-@Yu_Untethering_2015
-11.	Yu H, Zhang Y, Moss BL, et al.
-Untethering the TIR1 auxin receptor from the SCF complex increases its stability and inhibits auxin response.
-Nature plants.
-2015;1(3):14030.
-doi:10.1038/nplants.2014.30.
-## Aux/IAA genes
+The function `Nucleotide_diversity` in the r1001genomes package calculates $\pi_k$ for each position in the gene or region that contains a variant.
+note, $\pi_k$ is equal to 0 at all locations without variants.
+This is also what is displayed in the Diversity Plot tab of the web app.
 
-### Group A Aux/IAAs
-Britney
+## Detailed $\pi_k$ calculaiton simplification.
 
-### Group B Aux/IAAs
-Mollye/Alex
+The formula for $\pi_k$ above requires comparing every sequence to everyother sequence at location k,
+however, we know there are only a few variant forms at each individual location.
 
-### Group C Aux/IAAs
-Mega/Alex
+so we can revert back to using Nei's original forula, modifying it slightly,
+replacing $x_i$ with $\frac{n_i}{n}$, $n_i$ being the number of sequences in the population of the form $N_i$ at location k: subscripts i and k are summed only over unique forms [need to clarify this]
 
-## TPL/TPR genes  
-Alex
+$\displaystyle \pi_k = \sum_{ij} \frac{n_i}{n}\frac{n_j}{n} \pi_{ij} = 
+\frac{1}{n^2}\sum_{ij} n_i n_j \pi_{ij}$
 
-## ARF genes
+$\pi_{ij(k)} = \left\{\begin{array}{l}
+1 \quad \textrm{if} \quad i \neq j \\
+0 \quad \textrm{if} \quad i = j
+\end{array}\right.$
 
-### Class A ARFs
-Amy
+we will define $n_{!i} = n - n_i$ as the number of sequences different from i at k.
 
-### Class B ARFs
-Andrew
+we can see the summed term will be zero if $i=j$, and $n_i n_j$ if $i\neq j$ so:
 
-### Class C ARFs
-Hardik
+$\displaystyle \pi_k = \frac{1}{n^2} \sum_{i} n_i n_{!i}$
+
+next we substitute our deffinition of $n_{!i}$
+
+$\displaystyle \pi_k = \frac{1}{n^2} \sum_{i} n_i (n - n_{i})$
+
+then distribute and split summation:
+
+$\displaystyle \pi_k = \frac{1}{n^2} (n \sum_{i} n_i - \sum_{i} n_{i}^2)$
+
+finally, summing $\displaystyle \sum_{i}n_i$ is equal to n:
+
+$\displaystyle \pi_k = \frac{1}{n^2} (n^2 - \sum_{i} n_{i}^2)$
+
+this simplified form for $\pi_k$ is used by the app, because the counts of unique 
+variants at a single nucleotide location can easily be sumaraized in R.

### 2018-03-07 manuscript/manuscript.Rmd

--- 2018-03-07 manuscript/manuscript.Rmd
@@
 # Methods
+
+## Nucleotide Diversity Calculation
+Nei  defined the nucleotide diversity statistic in his original paper as:
+“…the average number of nucleotide differences per site between two randomly chosen DNA sequences.” (Nei, 1979)
+And provided the equation:
+
+$\displaystyle \pi = \sum_{ij}x_{i}x_{j}\pi_{ij}$
+
+Where Where $x_{i}$ is the frequency of the ith sequence in the population and $\pi_{ij}$ is the number of sites that are different between the ith and jth sequence divided by sequence length.
+
+a more general form, that treats each sequence in the population as unique can be written as:
+
+$\displaystyle \pi = \frac{1}{L*n^2} \sum_{i=1}^n \sum_{j=1}^n \sum_{k=1}^L{\pi_{ijk}}
+\qquad
+\displaystyle \pi_{ijk} = \left\{\begin{array}{l}
+1 \quad \textrm{if} \quad N_{ik} \neq N_{jk} \\
+0 \quad \textrm{if} \quad N_{ik} = N_{jk}
+\end{array}\right.$
+
+where $N_{ik}$ is the nucelotide (A, T, C or G) at position k on the ith sequence of the population.
+$L$ is the lenght of the sequence (the same for all sequences in the population, 
+assumes only SNPs exist, no indels). n is the total number of sequences in the population.
+
+from this form we can re-arange summations to the form below:
+
+$\displaystyle \pi = \frac{1}{L} \sum_{k=1}^L{\pi_{k}}
+\qquad
+\pi_{k} = \frac{1}{n^2} \sum_{i=1}^n \sum_{j=1}^n {\pi_{ijk}}$
+
+here $\pi_k$ can be thought of as the site-wise nucleotide diversity at postion k, and is equal to 
+the nucleotide diversity of a sequence of lenght 1 at location k. we can calculate $\pi_k$ for each
+site, then average those over the sequence lenght to calculate $\pi$, the nucleotide diversity 
+of the sequence.
+
+The function `Nucleotide_diversity` in the r1001genomes package calculates $\pi_k$ for each position in the gene or region that contains a variant.
+note, $\pi_k$ is equal to 0 at all locations without variants.
+This is also what is displayed in the Diversity Plot tab of the web app.
+
+## Detailed $\pi_k$ calculaiton simplification.
+
+The formula for $\pi_k$ above requires comparing every sequence to everyother sequence at location k,
+however, we know there are only a few variant forms at each individual location.
+
+so we can revert back to using Nei's original forula, modifying it slightly,
+replacing $x_i$ with $\frac{n_i}{n}$, $n_i$ being the number of sequences in the population of the form $N_i$ at location k: subscripts i and k are summed only over unique forms [need to clarify this]
+
+$\displaystyle \pi_k = \sum_{ij} \frac{n_i}{n}\frac{n_j}{n} \pi_{ij} = 
+\frac{1}{n^2}\sum_{ij} n_i n_j \pi_{ij}$
+
+$\pi_{ij(k)} = \left\{\begin{array}{l}
+1 \quad \textrm{if} \quad i \neq j \\
+0 \quad \textrm{if} \quad i = j
+\end{array}\right.$
+
+we will define $n_{!i} = n - n_i$ as the number of sequences different from i at k.
+
+we can see the summed term will be zero if $i=j$, and $n_i n_j$ if $i\neq j$ so:
+
+$\displaystyle \pi_k = \frac{1}{n^2} \sum_{i} n_i n_{!i}$
+
+next we substitute our deffinition of $n_{!i}$
+
+$\displaystyle \pi_k = \frac{1}{n^2} \sum_{i} n_i (n - n_{i})$
+
+then distribute and split summation:
+
+$\displaystyle \pi_k = \frac{1}{n^2} (n \sum_{i} n_i - \sum_{i} n_{i}^2)$
+
+finally, summing $\displaystyle \sum_{i}n_i$ is equal to n:
+
+$\displaystyle \pi_k = \frac{1}{n^2} (n^2 - \sum_{i} n_{i}^2)$
+
+this simplified form for $\pi_k$ is used by the app, because the counts of unique 
+variants at a single nucleotide location can easily be sumaraized in R.
 

### 2018-03-07 manuscript/manuscript.Rmd
Revert "Intro section"
This reverts commit 5adfd9cb373391bbbc504fa5a125bf92f7eaa343.

--- 2018-03-07 manuscript/manuscript.Rmd
@@
 ---
-title: "Untitled"
-author: "Morgan Hamm"
-date: "March 4, 2018"
-output: html_document
+title: "A comprehensive analysis of natural sequence variation within the <i>Arabidopsis thaliana</i> nuclear auxin signaling pathway"
+author: 
+  affiliation: University of Washington
+  name: "R. Clay Wright"
+
+date: "`r Sys.Date()`"
+
+output: 
+    html_document:
+      toc: true
+      fig_caption: yes
+
+keywords: r markdown
+fontsize: 11pt
+fontfamily: Helvetica
+
+bibliography: ["natural_variation.bib", "TIR1_AFB.bib"]
+csl: apa.csl
+link-citations: true
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+```{r, echo = FALSE, message=FALSE, warning=FALSE}
+knitr::opts_chunk$set(warning = FALSE, collapse = TRUE, comment = "#>", tidy = TRUE, out.width = "90%", fig.height=4, fig.width=7, out.extra='style="margin: auto; display: block; padding-top: 15px;"')
+library(r1001genomes)
+library(DT)
+library(dplyr)
+
 ```
+<!--This is a CSS (Cascading Style Sheet).
+ It controls how panels and images are formatted.></!-->
+<style>
+  img {
+    display: block;
+    margin: auto;
+    width: 90%;
+    height: auto;
+  }
+</style>
 
-## Nucleotide Diversity Calculation
-Nei  defined the nucleotide diversity statistic in his original paper as:
-“…the average number of nucleotide differences per site between two randomly chosen DNA sequences.” (Nei, 1979)
-And provided the equation:
+# List of working titles  
+* A comprehensive analysis of natural sequence variation within the <i>Arabidopsis thaliana</i> nuclear auxin signaling pathway  
+* *Add your suggestions*
 
-$\displaystyle \pi = \sum_{ij}x_{i}x_{j}\pi_{ij}$
+# Abstract
 
-Where Where $x_{i}$ is the frequency of the ith sequence in the population and $\pi_{ij}$ is the number of sites that are different between the ith and jth sequence divided by sequence length.
+* Motivating, high level statement  
+* Statement defining the cutting edge of your research area  
+* A statement of the form: "However, it remains unclear ..." that defines what problem you are  
+addressing.  
+* A statement of the form:"Here, we ..."  that concisely states your contribution.  
+* A series of 3?5 statement about your results.  
+* A concluding statement.
 
-a more general form, that treats each sequence in the population as unique can be written as:
+# Related work  
+What and how have we, as a field, learned from natural variation?  
+[@hollister_genomic_2014; @bielawski_detecting_2013; @kryazhimskiy_population_2008; @schmid_multilocus_2005; @nieduszynski_sequence_2011; @melamed_combining_2015; @weigel_population_2015; @schmitz_patterns_2013; @weigel_natural_2005; @hardtke_case_2010; @koornneef_naturally_2004; @mitchell-olds_genetic_2006; @joly-lopez_developing_2016; @wright_insights_2017a; @park_distributions_2017; @_135_2016a]
 
-$\displaystyle \pi = \frac{1}{L*n^2} \sum_{i=1}^n \sum_{j=1}^n \sum_{k=1}^L{\pi_{ijk}}
-\qquad
-\displaystyle \pi_{ijk} = \left\{\begin{array}{l}
-1 \quad \textrm{if} \quad N_{ik} \neq N_{jk} \\
-0 \quad \textrm{if} \quad N_{ik} = N_{jk}
-\end{array}\right.$
+# Methods
 
-where $N_{ik}$ is the nucelotide (A, T, C or G) at position k on the ith sequence of the population.
-$L$ is the lenght of the sequence (the same for all sequences in the population, 
-assumes only SNPs exist, no indels). n is the total number of sequences in the population.
+# Results  
+## TIR1/AFB genes  
+Romi
+[@gray_identification_1999; @kepinski_auxininduced_2004; @tan_mechanism_2007; @dezfulian_oligomerization_2016; @wu_evolution_2017b]
 
-from this form we can re-arange summations to the form below:
+## Aux/IAA genes
 
-$\displaystyle \pi = \frac{1}{L} \sum_{k=1}^L{\pi_{k}}
-\qquad
-\pi_{k} = \frac{1}{n^2} \sum_{i=1}^n \sum_{j=1}^n {\pi_{ijk}}$
+### Group A Aux/IAAs
+Britney
 
-here $\pi_k$ can be thought of as the site-wise nucleotide diversity at postion k, and is equal to 
-the nucleotide diversity of a sequence of lenght 1 at location k. we can calculate $\pi_k$ for each
-site, then average those over the sequence lenght to calculate $\pi$, the nucleotide diversity 
-of the sequence.
+### Group B Aux/IAAs
+Mollye/Alex
 
-The function `Nucleotide_diversity` in the r1001genomes package calculates $\pi_k$ for each position in the gene or region that contains a variant.
-note, $\pi_k$ is equal to 0 at all locations without variants.
-This is also what is displayed in the Diversity Plot tab of the web app.
+### Group C Aux/IAAs
+Mega/Alex
 
-## Detailed $\pi_k$ calculaiton simplification.
+## TPL/TPR genes  
+Alex
 
-The formula for $\pi_k$ above requires comparing every sequence to everyother sequence at location k,
-however, we know there are only a few variant forms at each individual location.
+## ARF genes
 
-so we can revert back to using Nei's original forula, modifying it slightly,
-replacing $x_i$ with $\frac{n_i}{n}$, $n_i$ being the number of sequences in the population of the form $N_i$ at location k: subscripts i and k are summed only over unique forms [need to clarify this]
+### Class A ARFs
+Amy
 
-$\displaystyle \pi_k = \sum_{ij} \frac{n_i}{n}\frac{n_j}{n} \pi_{ij} = 
-\frac{1}{n^2}\sum_{ij} n_i n_j \pi_{ij}$
+### Class B ARFs
+Andrew
 
-$\pi_{ij(k)} = \left\{\begin{array}{l}
-1 \quad \textrm{if} \quad i \neq j \\
-0 \quad \textrm{if} \quad i = j
-\end{array}\right.$
-
-we will define $n_{!i} = n - n_i$ as the number of sequences different from i at k.
-
-we can see the summed term will be zero if $i=j$, and $n_i n_j$ if $i\neq j$ so:
-
-$\displaystyle \pi_k = \frac{1}{n^2} \sum_{i} n_i n_{!i}$
-
-next we substitute our deffinition of $n_{!i}$
-
-$\displaystyle \pi_k = \frac{1}{n^2} \sum_{i} n_i (n - n_{i})$
-
-then distribute and split summation:
-
-$\displaystyle \pi_k = \frac{1}{n^2} (n \sum_{i} n_i - \sum_{i} n_{i}^2)$
-
-finally, summing $\displaystyle \sum_{i}n_i$ is equal to n:
-
-$\displaystyle \pi_k = \frac{1}{n^2} (n^2 - \sum_{i} n_{i}^2)$
-
-this simplified form for $\pi_k$ is used by the app, because the counts of unique 
-variants at a single nucleotide location can easily be sumaraized in R.
+### Class C ARFs
+Hardik

### 2018-03-07 manuscript/manuscript.Rmd
intro section

@@
 ---
-title: "A comprehensive analysis of natural sequence variation within the <i>Arabidopsis thaliana</i> nuclear auxin signaling pathway"
+title: "A comprehensive analysis of natural sequence variation within the <i>Arabidopsis thaliana</i> nuclear auxin signaling pathway using an accessible web application"
 author: 
@@
 
+# Intro
+
+The first genome sequence of *Arabidopsis thaliana* facilitated rapid advancement of plant biology through molecular genetics.
+Since this initial genome, massive scaling of sequencing technology has allowed a pioneering group to survey the global genomic variation in natural *Arabidopsis thaliana* populations.
+This valuable population genetics resource has led to several associations of genetic loci with phenotypic traits.
+Beyond these genome-wide studies for gene- and variant- discovery this valuable dataset has seen little use.
+This natural genomic variation also provides a set of feasible functional variation at the gene, gene family, and gene network level that may provide insight into the function and evolution of genes, families, and networks.
+We present a web application and R-package through which plant molecular biologists with little to no bioinformatics experience can make use of this rich dataset of genetic variation to formulate hypotheses as to the sequence/function/phenotype relationships determined by the gene, family, or network of their interest.
+This candidate gene/family tool also allows family-wise alignments of variants to facilitate de novo functional domain identification and evolutionary analysis, easy access to lists of missense polymorphisms to facilitate biochemical assays of variant effects, as well as lists of accessions containing missense (or any type of) polymorphisms to facilitate genetic segregation analysis for deterimining the phenotypic effects of variants and the study of gene network evolution.
+Here, we have utilized this tool to analyze the genetic variation in the nuclear auxin signaling pathway, and formulate hypotheses regarding the functional implications of this variation and the evolution of the genes and gene families in this pathway.
+
 # Related work  

### 2018-03-07 manuscript/manuscript.Rmd

--- 2018-03-07 manuscript/manuscript.Rmd
@@
 * A concluding statement.
+
+# Intro
+
+The first genome sequence of *Arabidopsis thaliana* facilitated rapid advancement of plant biology through molecular genetics.
+Since this initial genome, massive scaling of sequencing technology has allowed a pioneering group to survey the global genomic variation in natural *Arabidopsis thaliana* populations.
+This valuable population genetics resource has led to several associations of genetic loci with phenotypic traits.
+Beyond these genome-wide studies for gene- and variant- discovery this valuable dataset has seen little use.
+This natural genomic variation also provides a set of feasible functional variation at the gene, gene family, and gene network level that may provide insight into the function and evolution of genes, families, and networks.
+We present a web application and R-package through which plant molecular biologists with little to no bioinformatics experience can make use of this rich dataset of genetic variation to formulate hypotheses as to the sequence/function/phenotype relationships determined by the gene, family, or network of their interest.
+This candidate gene/family tool also allows family-wise alignments of variants to facilitate de novo functional domain identification and evolutionary analysis, easy access to lists of missense polymorphisms to facilitate biochemical assays of variant effects, as well as lists of accessions containing missense (or any type of) polymorphisms to facilitate genetic segregation analysis for deterimining the phenotypic effects of variants and the study of gene network evolution.
+Here, we have utilized this tool to analyze the genetic variation in the nuclear auxin signaling pathway, and formulate hypotheses regarding the functional implications of this variation and the evolution of the genes and gene families in this pathway.
 

### 2018-03-08 manuscript/manuscript.Rmd
Initial edits, stopped at "Identification of new Loss-of-Function"

@@
 
-next we substitute our deffinition of $n_{!i}$
+next we substitute our definition of $n_{!i}$
 
@@
 Auxin acts by binding to receptors that in turn target co-repressors for degradation.
-Auxin receptors (six in the model plant Arabidopsis thaliana) evolved through gene duplication and diversification early in the history of vascular plants @Parry_Complex_2009.
-The Nemhauser lab has recently shown that the rate of co-repressor degradation is determined by the identity of both the receptor and co-repressor @Havens_Synthetic_2012, and that this rate sets the pace of development3.
+Auxin receptors (six in the model plant Arabidopsis thaliana) evolved through gene duplication and diversification early in the history of vascular plants [@Parry_Complex_2009].
+The rate of co-repressor degradation is determined by the identity of both the receptor and co-repressor [@Havens_Synthetic_2012], and that this rate sets the pace of development [@Guseman_Auxin-induced_2015].
+
 Although it is unknown whether different receptors play different roles at the organismal level, a number of studies have pointed out differences in the ways the six different receptors in A.
 thaliana differ in their expression.
-The expression of AFB5 is induced in the presence of auxin, making this a circadian-regulated gene @Covington_Circadian_2007.
-It also appears AFB3 is more highly expressed in the roots in the presence of nitrate, allowing increased lateral root formation @Vidal_Nitrate_2010.
+The expression of AFB5 is induced in the presence of auxin, making this a circadian-regulated gene [@Covington_Circadian_2007].
+Also, AFB3 is more highly expressed in the roots in the presence of nitrate, allowing increased lateral root formation [@Vidal_Nitrate_2010].
 Although TIR1/AFBs are expressed ubiquitously in A.
-thaliana tissues, TIR1, AFB2, and AFB3 have been shown to accumulate in the shoot and root meristems and leaf tissues, with slightly different expression patterns in TIR1 @Parry_Complex_2009.
-AFB1 has been found to have a mutation drastically reducing its ability to assemble into an SCF complex, contributing to it’s increased accumulation, as well as preventing it from degrading Aux/IAAs.
+thaliana tissues, TIR1, AFB2, and AFB3 have been shown to accumulate in the shoot and root meristems and leaf tissues, with slightly different expression patterns for TIR1 [@Parry_Complex_2009].
+AFB1 has a mutation drastically reducing its ability to assemble into an SCF complex, contributing to its increased accumulation, as well as preventing it from degrading Aux/IAAs[@Yu_Untethering_2015].
 Higher order TIR1/AFB mutants show AFB1 to have inconclusive organismal function.
-However, AFB1 is shown to be able to accumulate ubiquitously in all tissues @Yu_Untethering_2015.
-When it comes to binding auxins, AFB4 and AFB5 have been shown to preferentially and functionally bind the synthetic auxin picloram, while others have a preference for naturally occurring auxin 3-IAA @Prigge_Arabidopsis_2016.
-Interestingly, the ways in which TIR1/AFBs are able to bind and mark Aux/IAAs for degradation are variable.
-TIR1 and AFB2 have slightly different speed at which they perform this function, with AFB2 being generally quicker, as shown by our lab’s yeast degradation system @Havens_Synthetic_2012.
-Finally, higher order TIR1/AFB mutants show AFB1 to have inconclusive organismal function.
-  Using the Natural Variation Webtool developed in our lab, I will characterize the natural variation in the TIR1/AFB family of auxin receptors by looking at the genomes of 1001 accessions of A.
-thaliana.
-This webtool allows me to	quantify and compare the nucleotide diversity of different genes in this family, as well as find relevant potentially function-altering variants.
-Using its diversity statistic function has shown me TIR1 and AFB1 both seem to be low diversity, hinting at their likely functional importance.
-AFB3 and 4 appear to be higher diversity, while their sisters are more conserved.
-This matches our current understanding of AFB3 and 4 as being minor players in the auxin signaling pathway, but brings up questions about the function of AFB1, which is inconclusive.
-Using this webtool, the AFB1 sequence can be aligned to the highly characterized TIR1 AA sequence showing locations of known non-synonymous variants.
-Although the most important functional regions do not contain variants in AFB1, there is a low diversity mutant, only found in the Can-1 accession, which is in the highly conserved TIR1/AFB oligomerization domain.
-Because variants eliciting significant changes in function often are not selected for, the fact that this variant is extremely low diversity holds promise that it is associated with a strong auxin signaling phenotype.
-I will order and characterize this accession in order to identify potential auxin signaling phenotypes.
-This would support that AFB1 functions through oligomerization with other TIR1/AFBs.
-  The AFB4 and AFB4 receptors are known to have an extension in their N-terminal domains.
-Using the alignment feature on this webtool has indicated this region to be the highest in diversity in these proteins, containing lots of missense variants (Fig.
-4).
-This suggests that this extension does not play an important functional role in these proteins.
+However, AFB1 is shown to be able to accumulate ubiquitously in all tissues.AFB4 and AFB5 have been shown to preferentially and functionally bind the synthetic auxin picloram, while others have a preference for naturally occurring auxin indole-3-acetic acid [@Prigge_Arabidopsis_2016].
+Interestingly, the strength and rate with which TIR1/AFBs are able to bind and mark Aux/IAAs for degradation are variable [@Havens_Synthetic_2012].
+TIR1 and AFB2 have slightly different speed at which they perform this function, with AFB2 being generally faster.
+
+Examining the natural sequence variation across the AFB revealed that *TIR1* and *AFB1* both had very low nonsynonymous diversity, hinting at their likely functional importance.
+*AFB3* and *AFB4* had higher nonsynonymous diversity, while their sisters genes, *AFB2* and *AFB5* are more conserved.
+This matches our current understanding of *AFB3* and *AFB4* as being minor players in the auxin signaling pathway, but brings in to question the inconclusive role of *AFB1* in auxin signaling.
+Although most important functional regions do not contain variants in AFB1, there is a low diversity mutant, only found in the Can-1 accession, which is in the highly conserved TIR1/AFB oligomerization domain.
+Because variants eliciting significant changes in function often are not selected for, the fact that this variant is extremely low diversity may suggest that it is associated with a strong auxin signaling phenotype.
+Characterization of this accession may identify potential *AFB1* phenotypes and may also test the hypothesis that AFB1 functions through oligomerization with other TIR1/AFBs.
+  The AFB4 and AFB5 receptors have an N-terminal extension prior to the F-box domains.
+This extesion had very high nonsynonymous diversity(Fig.
+4), suggesting that this extension does not play an important functional role in these proteins.
 
@@
 
-@Parry_Complex_2009
-1.	Parry G, Calderon-Villalobos LI, Prigge M, et al.
-Complex regulation of the TIR1/AFB family of auxin receptors.
-Proceedings of the National Academy of Sciences of the United States of America.
-2009;106(52):22540-22545.
-doi:10.1073/pnas.0911967106.
-@Havens_Synthetic_2012
-2.	Kyle A. Havens, et al. A Synthetic Approach Reveals Extensive Tunability of Auxin Signaling. Plant Physiology Sep 2012, 160 (1) 135-142; doi: 10.1104/pp.112.202184
-@Guseman_Auxin-induced_2015
-3.	Guseman JM, et al. Auxin-induced degradation dynamics set the pace for lateral root development. Development (Cambridge, England). 2015;142(5):905-909. doi:10.1242/dev.117234.
-@Covington_Circadian_2007
-8.	Covington MF, Harmer SL. The Circadian Clock Regulates Auxin Signaling and Responses in Arabidopsis . Weigel D, ed. PLoS Biology. 2007;5(8):e222. doi:10.1371/journal.pbio.0050222.
-@Vidal_Nitrate_2010
-9.	Vidal EA, Araus V, Lu C, et al.
-Nitrate-responsive miR393/AFB3 regulatory module controls root system architecture in Arabidopsis thaliana.
-Proceedings of the National Academy of Sciences of the United States of America.
-2010;107(9):4477-4482.
-doi:10.1073/pnas.0909571107.
-@Prigge_Arabidopsis_2016
-10.	Prigge MJ, Greenham K, Zhang Y, et al.
-The Arabidopsis Auxin Receptor F-Box Proteins AFB4 and AFB5 Are Required for Response to the Synthetic Auxin Picloram.
-G3: Genes|Genomes|Genetics.
-2016;6(5):1383-1390.
-doi:10.1534/g3.115.025585.
-@Yu_Untethering_2015
-11.	Yu H, Zhang Y, Moss BL, et al.
-Untethering the TIR1 auxin receptor from the SCF complex increases its stability and inhibits auxin response.
-Nature plants.
-2015;1(3):14030.
-doi:10.1038/nplants.2014.30.
->>>>>>> AFB
 ## Aux/IAA genes
@@
 The Aux/IAA proteins (IAAs) mediate both auxin perception and transcriptional co-repression of auxin responsive genes.
-Auxin perception is carried out along with TIR1/AFB proteins, which together act as a coreceptor to perceive intracellular auxin.
-Interaction of IAAs with auxin and TIR1/AFBs is conferred by the highly conserved 13 amino acid degron (DII) motif [@tan_mechanism_2007].
-Stabilizing mutations within the degron can lead to dominant phenotypes [@berleth_auxin_2004; @yang_iaa1_2004].
-Degradation of IAAs relieves their repression upon ARF transcription factors.
-IAA degradation rate is influenced both by the degron sequence and by rate motifs found outside the degron, such as the KR motif [@moss_rate_2015; @dreher_arabidopsis_2006].
-IAA-mediated repression of ARFs depends upon an EAR domain that recruits co-repressors of the TOPLESS (TPL) family, and the PB1 domain that enables interaction with ARF PB1 domains [@nanao_structural_2014; @kagale_ear_2011].
+Auxin perception is carried out along with TIR1/AFB proteins, which together with IAAs act as a coreceptor to bind and perceive intracellular auxin.
+Interaction of IAAs with auxin and TIR1/AFBs is conferred by the highly conserved 13 amino acid degron motif [@tan_mechanism_2007].
+Formation of this auxin coreceptor complex triggers ubiquitination and subsequent degradation of IAAs.
+This degradation in turn relieves their repression upon ARF transcription factors.Stabilizing mutations within the degron can lead to dominant phenotypes [@berleth_auxin_2004; @yang_iaa1_2004].
 The Arabidopsis Aux/IAA gene family has 29 members, most of which possess a canonical degron.
-Notable exceptions are the Auxin insensitive IAA20 and IAA30 that lack a degron sequence entirely, while the closely related IAA31 retains a semi-functional degron sequence with a very slow degradation rate [@dreher_arabidopsis_2006; @havens_synthetic_2012].
-Using the Natural Variation Webtool, we found that for the canonical Aux/IAAs, no missense mutations were present in the KR motif or PB1 interaction faces, and very few missense mutations were observed in the EAR domain and degron (Supp.
+Notable exceptions are the auxin-insensitive IAA20 and IAA30 that lack a degron sequence entirely, while the closely related IAA31 retains a semi-functional degron sequence with a very slow degradation rate [@dreher_arabidopsis_2006; @havens_synthetic_2012].
+Additionally, IAA degradation rate is influenced by rate motifs found outside the degron, such as the KR motif [@moss_rate_2015; @dreher_arabidopsis_2006].
+IAA-mediated repression of ARFs depends upon both an EAR domain, N-terminal of the degron, that recruits co-repressors of the TOPLESS (TPL) family, and a Phox and Bem1 (PB1) domain that enables interaction via heteroligomerization with ARF PB1 domains [@nanao_structural_2014; @kagale_ear_2011].
+
+Within the available natural sequence variation, we found that for the canonical Aux/IAAs, no missense mutations were present in the KR motif or residues responsible for PB1-PB1 domain interactions (Supp.
 Fig.
 Alignment of canonical IAAs).
-However, as discussed below, there were several interesting missense mutations found in EAR domains and degrons.
-The majority of nonsynonymous mutations observed were in the variable regions between the conserved domains (Supp.
-Fig.
-Alignment of canonical IAAs).
+The majority of nonsynonymous mutations observed were in the variable regions between the conserved domains.
+However, several missense polymorphisms of interest were identified in EAR domains and degrons of several canonical IAAs.
 
 ###Sister Pair Diversification
-Previous research has pointed to genetic redundancy within closely related pairs or families of IAA proteins [@overvoorde_functional_2005].
+Previous research has ~~pointed to~~ *found evidence of* genetic redundancy within closely related pairs or groups of IAA proteins [@overvoorde_functional_2005].
 For example, the iaa8-1, iaa9-1 double mutant and the iaa5-1, iaa6-1, iaa19-1 triple mutant have wild-type phenotypes [@overvoorde_functional_2005].
-Despite this, most Arabidopsis IAA sister pairs have one member exhibiting higher diversity while the other sister is more conserved (Figure IAA phylogeny heat map).
+Despite this, most *Arabidopsis* IAA sister pairs have one member with high nonsynonymous diversity while the other sister is more conserved (Figure IAA phylogeny heat map).
 For example, despite having only minor differences in sequence (61.4% sequence identity), the IAA6/19 sister pair has significant differences in expression patterns, protein abundances and functions, suggesting they have undergone functional specialization since their divergence [@winkler_variation_2017].
-IAA19 was expressed at higher levels than IAA6 and showed less sequence divergence; IAA6 showed regions of increased divergence upstream of the degron and in the N-terminal half of the PB1 domain [@winkler_variation_2017].
-Consistent with these results, the Natural Variation Webtool (Figure IAA phylogeny heat map) revealed a lower natural variation for IAA19 (ᴨN/ᴨS = 0.55) compared to IAA6 (ᴨN/ᴨS = 2.3), and also detected high natural variation diversity within the same regions of IAA6 as seen in Winkler et al.
+In an interspecific comparison between *A.
+thaliana* abd *A.
+lyrata*, IAA19 is more conserved, while IAA6 showed regions of increased sequence divergence upstream of the degron and in the N-terminal half of the PB1 domain [@winkler_variation_2017].
+Consistent with these results, our analysis of intraspecific natural variation revealed higher conservation for IAA19 (${\pi_N/\pi_S}$ = 0.55) compared to IAA6 (${\pi_N/\pi_S}$ = 2.3) (Figure IAA phylogeny heat map), and also detected high diversity within the same regions of IAA6 as seen in Winkler et al.
 (Supp Fig IAA6 diversity plot).
-Within IAA Group B, there are four sister pairs--IAA10 and IAA11, IAA12/BDL and IAA13, IAA18 and IAA26/PAP1, and IAA20 and IAA30 (Fig X) [@remington_contrasting_2004].
-Interestingly, among these sister pairs only IAA12 and 13 have noticeable differences in diversity.
-IAA10 and IAA11 show little difference in ℼN/ℼS values, with ℼN/ℼS of 0.80 and 0.67 respectively, implicating the possibility that IAA10 and 11 play redundant roles.
-In rice however, it has been shown that IAA11 gain of function mutants block lateral root initiation [@zhu_gain--function_2012].
+
+Within the Group B IAA genes, there are four sister pairs--IAA10/IAA11, IAA12/IAA13, IAA18/IAA26, and IAA20/IAA30 (Fig X) [@remington_contrasting_2004].
+Interestingly, among these sister pairs only IAA12 and 13 have a noticeable difference in nonsynonymous diversity.
 The Arabidopsis EFP browser shows that IAA10 and 11 have almost identical expression patterns, with the exception of IAA10 being more highly expressed in the developing seed, and IAA11 being more highly expressed in root tissue during lateral root development.
-As discussed below, the Natural Variation Webtool provides a potentially expedited route to conducting experiments meant to elucidate the extent of redundancy between family members such as IAA10 and 11.
+In our analysis, IAA10 and IAA11 both showed functional conservation (${\pi_N/\pi_S}$ of 0.80 and 0.67 respectively) in support of IAA10 and IAA11 playing redundant roles.
+Accessions and associated variants identified here provide a potentially expedited route to elucidating the extent of redundancy between family members such as IAA10 and 11.
 

### 2018-03-08 manuscript/manuscript.Rmd
initial edits cont. "TPL/TPR"

@@
 ###Identification of new Loss-of-Function/Gain-of-Function mutations
-To test the extent of redundancy/specialization between sister pairs, the Natural Variation Webtool allows for identifications of accessions that can be potentially used to investigate the phenotypic effects in loss of function variants.
-For example, it has been shown that IAA10 has a missense mutation in the EAR domain, which converts the second leucine in the LxLxL motif to a serine (L49S), a mutation likely to reduce the ability of the IAA to recruit TOPLESS co-repressor [@lee_biological_2016].
-This mutation is found in three accessions in the southern UK.
-A second potentially stronger mutation, represented in one accession from northeastern Spain is a missense resulting in an early stop codon in the PB1 domain (need Supp.
+Accessions with functional variants can be potentially used to investigate the extent of redundancy/specialization within sister pairs.
+For example, it has been shown that IAA10 has a missense mutation in the EAR domain, which converts the second leucine in the LxLxL motif to a serine (L49S), a mutation likely to reduce the ability of the IAA to recruit the TOPLESS co-repressor [@lee_biological_2016].
+This mutation was found in three accessions from the southern UK.
+A second potentially stronger mutation, represented in one accession from northeastern Spain is a ~~missense~~ *frameshift?* resulting in an early stop codon in the PB1 domain (need Supp.
 Fig?).
-Such a mutation is likely to prevent IAA10 from interacting with an ARF, rendering it impotent.
-These alleles would be strong candidates for loss-of-function mutants, allowing researchers to discover the function of IAA10.
-In particular, it would make possible a straightforward path for creating an iaa10/11 double mutant using CRISPR/Cas9 directed mutagenesis, allowing for examination of the extent of redundancy between these family members.
+This likely prevents IAA10 from interacting with an ARF.
+~~These alleles would be strong candidates for loss-of-function mutants, allowing researchers to discover the function of IAA10.
+In particular, it would make possible a straightforward path for creating an iaa10/11 double mutant using CRISPR/Cas9 directed mutagenesis, allowing for examination of the extent of redundancy between these family members.~~ *I don't think we can say this.
+There are SALK lines in the 5' UTR and first intron of IAA10 and IAA11.*
 
-Several examples of possible gain-of-function degron missense mutations were found in IAA2, IAA3, IAA6, IAA7, IAA8, IAA13, IAA15, IAA18, IAA19, IAA27, IAA28, IAA29.
-Most degron mutations were found in one or two accessions represented in the database (Supplemental Table or Map?).
-Interestingly, almost all of the IAAs found to have accessions with degron missense mutations have a closely related sister, which may indicate a redundant function within the sister-pairs.
-IAA6 natural variation has two potentially dominant missense mutations found within the degron (Supp Fig.
+Several examples of possible gain-of-function degron missense variants were found in IAA2, IAA3, IAA6, IAA7, IAA8, IAA13, IAA15, IAA18, IAA19, IAA27, IAA28, IAA29.
+Most degron variants were only found in one or two accessions (Supplemental Table or Map?).
+~Interestingly, almost all of the IAAs found to have accessions with degron missense mutations have a closely related sister, which may indicate a redundant function within the sister-pairs.~ *But we know that iaa6 iaa19 iaa5 triple doesn't have a phenotype.* IAA6 has two potentially dominant missense mutations found within the degron (Supp Fig.
 alignment), one previously believed to lead to the shy1-1 phenotype [@kim_two_1996].
-However, this variant is widely represented in 31 accessions across europe (Supplemental Map Fig).
+Interestingly, this variant is widely represented in 31 accessions across europe (Supplemental Map Fig).
 This finding supports the proposition that iaa6[C78R] may not be the mutation causing the shy1-1 phenotype but rather represents natural variation in the degron sequence ([note from Mike Prigge in TAIR](https://www.arabidopsis.org/servlets/TairObject?type=notepad&id=11247)).
-These accessions could be used to confirm whether the shy1-1 phenotype exists within these plants.
-Additionally, an accession was identified with a V77M mutation in the IAA6 degron that may be useful to explore potential iaa6 gain-of-function phenotypes.
+These accessions could be used to confirm whether iaa6[C78R] is responsible for the SHY1-1D phenotype.
+Additionally, an accession was identified with the IAA6-V77M variant, also in the degron, that may be useful to further explore iaa6 gain-of-function phenotypes.
 
@@
 For example, recent detailed biochemical analysis identified putative lysine ubiquitination sites in members of the IAA6/19 sister pairs [@winkler_variation_2017]; Supp Fig IAA6/19 alignment).
-These sites show no natural sequence variation in Arabidopsis, while several other lysines are represented in the natural sequence diversity (Supp Figs alignment and map).
+These sites are conserved across natural sequence variation, while several other lysine residues vary (Supp Figs alignment and map).
 In IAA6, there are 17 lysines of which 6 are putative ubiquitination sites, one is in the KR motif, while two others show natural variation.
 In IAA19, there are 15 total lysines, with 8 as putative ubiquitination sites, 2 in the KR motif and degron, and 1 that exhibits natural variation.
-Natural variation sequence diversity may be used to narrow down a subset of lysines to assess for functional ubiquitination.
+**This sentence is not clear.
+Only 1 of the ubiquitination sites varies?** Natural sequence diversity may be used to narrow down a subset of lysines to assess for functional ubiquitination.
 
-The members of the IAA8/9/27 group have a unique N-terminal extension with a highly conserved region common to orthologs in several other species [@dreher_arabidopsis_2006].
-IAA8/9 have  redundant roles in leaf serration [@koenig_auxin_2009], while IAA8 important in lateral roots [@arase_iaa8_2012].
+The members of the IAA8/9/27 clade have a unique N-terminal extension with a highly conserved region common to orthologs in several other species [@dreher_arabidopsis_2006].
+IAA8/9 play redundant roles in leaf serration [@koenig_auxin_2009], while IAA8 important in lateral roots [@arase_iaa8_2012].
 IAA8 has been reported to uniquely localize in the cytosol [@arase_iaa8_2012] where it may interact with LSD1 protein [@coll_programmed_2011].
@@
 
-Members of the IAA17/7/14/16 group have a second EAR motif (LxLxL) with no observed missense mutations in IAA7/16/17 (Supp Fig IAA7/14/16/17 alignment).
+Members of the IAA17/7/14/16 group have a second EAR motif (LxLxL) with no observed missense mutations in IAA7, 16, or 17 (Supp Fig IAA7/14/16/17 alignment).
 This observation supports the experimental evidence for IAA7 that the second EAR motif is functional: it can interact with TPR1 corepressor and has a minor repressive role in auxin-related developmental phenotypes [lee_biological_2016].
-For IAA14, we did observe two missense mutations that may be of interest to probe functionality of the IAA14 EAR motif: one in the central “L” (L41F) that is represented in only one accession, and the other in the second “x” (K40N) that is found in 12 accessions around Copenhagen (Supp Fig.
+For IAA14, we did observe two missense mutations that may be of interest to probe functionality of the IAA14 EAR motif: one in the central lysine (L41F) that is represented in only one accession, and the other in the second “x” (K40N) that is found in 12 accessions around Copenhagen (Supp Fig.
 IAA14 map).

### 2018-03-10 manuscript/manuscript.Rmd
added a code chunk in sister pair diversification
last updates before sending to google drive

@@
 * A comprehensive analysis of natural sequence variation within the <i>Arabidopsis thaliana</i> nuclear auxin signaling pathway  
-* *Add your suggestions*
+* **Add your suggestions**
 
@@
 This valuable population genetics resource has led to several associations of genetic loci with phenotypic traits.
-Beyond these genome-wide studies for gene- and variant- discovery this valuable dataset has seen little use.
-This natural genomic variation also provides a set of feasible functional variation at the gene, gene family, and gene network level that may provide insight into the function and evolution of genes, families, and networks.
-We present a web application and R-package through which plant molecular biologists with little to no bioinformatics experience can make use of this rich dataset of genetic variation to formulate hypotheses as to the sequence/function/phenotype relationships determined by the gene, family, or network of their interest.
-This candidate gene/family tool also allows family-wise alignments of variants to facilitate de novo functional domain identification and evolutionary analysis, easy access to lists of missense polymorphisms to facilitate biochemical assays of variant effects, as well as lists of accessions containing missense (or any type of) polymorphisms to facilitate genetic segregation analysis for deterimining the phenotypic effects of variants and the study of gene network evolution.
-Here, we have utilized this tool to analyze the genetic variation in the nuclear auxin signaling pathway, and formulate hypotheses regarding the functional implications of this variation and the evolution of the genes and gene families in this pathway.
+Outside of genome-wide studies this valuable dataset has seen little use.
+Natural genomic variation also provides a set of feasible functional variation at the gene, gene family and gene network level that may provide insight into the function and evolution of genes, families and networks.
+Here, we present a web application and R-package through which plant molecular biologists with little to no bioinformatics experience can make use of this rich dataset of genetic variation to formulate hypotheses as to the sequence/function/phenotype relationships determined by the gene, family, or network of their interest.
+We demonstrate the utility of this tool through comprehensive analysis and identification of potential functional variation in the nuclear auxin signaling pathway.
 
-# Related work  
-What and how have we, as a field, learned from natural variation?  
-[@hollister_genomic_2014; @bielawski_detecting_2013; @kryazhimskiy_population_2008; @schmid_multilocus_2005; @nieduszynski_sequence_2011; @melamed_combining_2015; @weigel_population_2015; @schmitz_patterns_2013; @weigel_natural_2005; @hardtke_case_2010; @koornneef_naturally_2004; @mitchell-olds_genetic_2006; @joly-lopez_developing_2016; @wright_insights_2017a; @park_distributions_2017; @_135_2016a]
+This application allows easy access to 
+1. lists of missense polymorphisms to facilitate biochemical assays of variant effects [@starita_variant_2017],
+2. family-wise alignments of variants to facilitate de novo functional domain identification [@melamed_combining_2015], 
+3.
+lists of accessions containing missense (or any type of) polymorphisms to facilitate segregation analysis and measurment of the distribution of variant effects on phenotype [@park_distributions_2017],
+4. as well as the study of gene and network evolution [@delker_natural_2010; @kliebenstein_role_2008].
+
+Here, we have utilized this tool to analyze the genetic variation in the nuclear auxin signaling pathway and formulate hypotheses regarding the functional implications of this variation and the evolution of the genes and gene families in this pathway.
+Our analysis also provided further confirmation of much of the existing knowledge of these genes demonstrating the validity of this approach.
 
@@
 
-where $N_{ik}$ is the nucelotide (A, T, C or G) at position k on the ith sequence of the population.
-$L$ is the lenght of the sequence (the same for all sequences in the population, 
+where $N_{ik}$ is the nucleotide (A, T, C or G) at position k on the ith sequence of the population.
+$L$ is the length of the sequence (the same for all sequences in the population, 
 assumes only SNPs exist, no indels). n is the total number of sequences in the population.
 
-from this form we can re-arange summations to the form below:
+from this form we can re-arrange summations to the form below:
 
@@
 
-here $\pi_k$ can be thought of as the site-wise nucleotide diversity at postion k, and is equal to 
-the nucleotide diversity of a sequence of lenght 1 at location k. we can calculate $\pi_k$ for each
-site, then average those over the sequence lenght to calculate $\pi$, the nucleotide diversity 
+here $\pi_k$ can be thought of as the site-wise nucleotide diversity at position k, and is equal to 
+the nucleotide diversity of a sequence of length 1 at location k. we can calculate $\pi_k$ for each
+site, then average those over the sequence length to calculate $\pi$, the nucleotide diversity 
 of the sequence.
@@
 
-The formula for $\pi_k$ above requires comparing every sequence to everyother sequence at location k,
+The formula for $\pi_k$ above requires comparing every sequence to every other sequence at location k,
 however, we know there are only a few variant forms at each individual location.
 
-so we can revert back to using Nei's original forula, modifying it slightly,
+so we can revert back to using Nei's original formula, modifying it slightly,
 replacing $x_i$ with $\frac{n_i}{n}$, $n_i$ being the number of sequences in the population of the form $N_i$ at location k: subscripts i and k are summed only over unique forms [need to clarify this]
@@
 this simplified form for $\pi_k$ is used by the app, because the counts of unique 
-variants at a single nucleotide location can easily be sumaraized in R.
+variants at a single nucleotide location can easily be summarized in R.
 
@@
 
-Auxin acts by binding to receptors that in turn target co-repressors for degradation.
-Auxin receptors (six in the model plant Arabidopsis thaliana) evolved through gene duplication and diversification early in the history of vascular plants [@Parry_Complex_2009].
+Auxin acts by binding to receptors that in turn target co-corepressors for degradation.
+Auxin receptors (six in the model plant *Arabidopsis thaliana*) evolved through gene duplication and diversification early in the history of vascular plants [@Parry_Complex_2009].
 The rate of co-repressor degradation is determined by the identity of both the receptor and co-repressor [@Havens_Synthetic_2012], and that this rate sets the pace of development [@Guseman_Auxin-induced_2015].
@@
   The AFB4 and AFB5 receptors have an N-terminal extension prior to the F-box domains.
-This extesion had very high nonsynonymous diversity(Fig.
+This extension had very high nonsynonymous diversity(Fig.
 4), suggesting that this extension does not play an important functional role in these proteins.
@@
 thaliana TIR1/AFB F-box domains showing variants.
-Cyan marks the Fbox binding domain.
+Cyan marks the F-box binding domain.
 Yellow, green, and magenta, mark missense and synonymous, synonymous only, and missense only variant locations respectively.
@@
 
-###Introduction
 The Aux/IAA proteins (IAAs) mediate both auxin perception and transcriptional co-repression of auxin responsive genes.
-Auxin perception is carried out along with TIR1/AFB proteins, which together with IAAs act as a coreceptor to bind and perceive intracellular auxin.
+Auxin perception is carried out along with TIR1/AFB proteins, which together with IAAs act as a co-receptor to bind and perceive intracellular auxin.
 Interaction of IAAs with auxin and TIR1/AFBs is conferred by the highly conserved 13 amino acid degron motif [@tan_mechanism_2007].
-Formation of this auxin coreceptor complex triggers ubiquitination and subsequent degradation of IAAs.
+Formation of this auxin co-receptor complex triggers ubiquitination and subsequent degradation of IAAs.
 This degradation in turn relieves their repression upon ARF transcription factors.Stabilizing mutations within the degron can lead to dominant phenotypes [@berleth_auxin_2004; @yang_iaa1_2004].
-The Arabidopsis Aux/IAA gene family has 29 members, most of which possess a canonical degron.
+The *Arabidopsis thaliana* Aux/IAA gene family has 29 members, most of which possess a canonical degron.
 Notable exceptions are the auxin-insensitive IAA20 and IAA30 that lack a degron sequence entirely, while the closely related IAA31 retains a semi-functional degron sequence with a very slow degradation rate [@dreher_arabidopsis_2006; @havens_synthetic_2012].
 Additionally, IAA degradation rate is influenced by rate motifs found outside the degron, such as the KR motif [@moss_rate_2015; @dreher_arabidopsis_2006].
-IAA-mediated repression of ARFs depends upon both an EAR domain, N-terminal of the degron, that recruits co-repressors of the TOPLESS (TPL) family, and a Phox and Bem1 (PB1) domain that enables interaction via heteroligomerization with ARF PB1 domains [@nanao_structural_2014; @kagale_ear_2011].
+IAA-mediated repression of ARFs depends upon both an EAR domain, N-terminal of the degron, that recruits co-repressors of the TOPLESS (TPL) family, and a Phox and Bem1 (PB1) domain that enables interaction via hetero-oligomerization with ARF PB1 domains [@nanao_structural_2014; @kagale_ear_2011].
 
@@
 ###Sister Pair Diversification
-Previous research has ~~pointed to~~ *found evidence of* genetic redundancy within closely related pairs or groups of IAA proteins [@overvoorde_functional_2005].
-For example, the iaa8-1, iaa9-1 double mutant and the iaa5-1, iaa6-1, iaa19-1 triple mutant have wild-type phenotypes [@overvoorde_functional_2005].
-Despite this, most *Arabidopsis* IAA sister pairs have one member with high nonsynonymous diversity while the other sister is more conserved (Figure IAA phylogeny heat map).
-For example, despite having only minor differences in sequence (61.4% sequence identity), the IAA6/19 sister pair has significant differences in expression patterns, protein abundances and functions, suggesting they have undergone functional specialization since their divergence [@winkler_variation_2017].
+
+Previous research has ~~pointed to~~ **found evidence of** both broad genetic redundancy and specificity within closely related pairs or groups of IAA proteins [@overvoorde_functional_2005].
+For example, the iaa8-1, iaa9-1 double mutant and the iaa5-1, iaa6-1, iaa19-1 triple mutant have wild-type phenotypes [@overvoorde_functional_2005], yet the IAA6/19 sister pair has significant differences in expression patterns, protein abundances and functions, suggesting they have undergone functional specialization since their divergence [@winkler_variation_2017].
+
+In our analysis of the full IAA gene family, one member of most IAA sister pairs had high nonsynonymous diversity while the other sister was more conserved (Figure IAA phylogeny heat map).
 In an interspecific comparison between *A.
-thaliana* abd *A.
+thaliana* and *A.
 lyrata*, IAA19 is more conserved, while IAA6 showed regions of increased sequence divergence upstream of the degron and in the N-terminal half of the PB1 domain [@winkler_variation_2017].
@@
 Interestingly, among these sister pairs only IAA12 and 13 have a noticeable difference in nonsynonymous diversity.
-The Arabidopsis EFP browser shows that IAA10 and 11 have almost identical expression patterns, with the exception of IAA10 being more highly expressed in the developing seed, and IAA11 being more highly expressed in root tissue during lateral root development.
+The *Arabidopsis thaliana* EFP browser shows that IAA10 and 11 have almost identical expression patterns, with the exception of IAA10 being more highly expressed in the developing seed, and IAA11 being more highly expressed in root tissue during lateral root development.
 In our analysis, IAA10 and IAA11 both showed functional conservation (${\pi_N/\pi_S}$ of 0.80 and 0.67 respectively) in support of IAA10 and IAA11 playing redundant roles.
@@
 This mutation was found in three accessions from the southern UK.
-A second potentially stronger mutation, represented in one accession from northeastern Spain is a ~~missense~~ *frameshift?* resulting in an early stop codon in the PB1 domain (need Supp.
+A second potentially stronger mutation, represented in one accession from northeastern Spain is a ~~missense~~ **frameshift?** resulting in an early stop codon in the PB1 domain (need Supp.
 Fig?).
@@
 ~~These alleles would be strong candidates for loss-of-function mutants, allowing researchers to discover the function of IAA10.
-In particular, it would make possible a straightforward path for creating an iaa10/11 double mutant using CRISPR/Cas9 directed mutagenesis, allowing for examination of the extent of redundancy between these family members.~~ *I don't think we can say this.
-There are SALK lines in the 5' UTR and first intron of IAA10 and IAA11.*
+In particular, it would make possible a straightforward path for creating an iaa10/11 double mutant using CRISPR/Cas9 directed mutagenesis, allowing for examination of the extent of redundancy between these family members.~~ **I don't think we can say this.
+There are SALK lines in the 5' UTR and first intron of IAA10 and IAA11.**
 
@@
 Most degron variants were only found in one or two accessions (Supplemental Table or Map?).
-~Interestingly, almost all of the IAAs found to have accessions with degron missense mutations have a closely related sister, which may indicate a redundant function within the sister-pairs.~ *But we know that iaa6 iaa19 iaa5 triple doesn't have a phenotype.* IAA6 has two potentially dominant missense mutations found within the degron (Supp Fig.
+~Interestingly, almost all of the IAAs found to have accessions with degron missense mutations have a closely related sister, which may indicate a redundant function within the sister-pairs.~ **But we know that iaa6 iaa19 iaa5 triple doesn't have a phenotype.** IAA6 has two potentially dominant missense mutations found within the degron (Supp Fig.
 alignment), one previously believed to lead to the shy1-1 phenotype [@kim_two_1996].
-Interestingly, this variant is widely represented in 31 accessions across europe (Supplemental Map Fig).
+Interestingly, this variant is widely represented in 31 accessions across Europe (Supplemental Map Fig).
 This finding supports the proposition that iaa6[C78R] may not be the mutation causing the shy1-1 phenotype but rather represents natural variation in the degron sequence ([note from Mike Prigge in TAIR](https://www.arabidopsis.org/servlets/TairObject?type=notepad&id=11247)).
@@
 Members of the IAA17/7/14/16 group have a second EAR motif (LxLxL) with no observed missense mutations in IAA7, 16, or 17 (Supp Fig IAA7/14/16/17 alignment).
-This observation supports the experimental evidence for IAA7 that the second EAR motif is functional: it can interact with TPR1 corepressor and has a minor repressive role in auxin-related developmental phenotypes [lee_biological_2016].
+This observation supports the experimental evidence for IAA7 that the second EAR motif is functional: it can interact with TPR1 co-repressor and has a minor repressive role in auxin-related developmental phenotypes [lee_biological_2016].
 For IAA14, we did observe two missense mutations that may be of interest to probe functionality of the IAA14 EAR motif: one in the central lysine (L41F) that is represented in only one accession, and the other in the second “x” (K40N) that is found in 12 accessions around Copenhagen (Supp Fig.
@@
 
-  The Auxin signaling pathway utilizes the TOPLESS (TPL) family of Gro/TLE/TUP1 type co-repressor proteins to maintain auxin responsive genes in a transcriptionally-repressed state in the absence of auxin (cite).
-In Arabidopsis the TPL family includes TOPLESS and four TPL-related proteins (TPR1-4), which are comprised of three structural domains: an N-terminal TPL domain and two beta-propellar domains.
-TPL is recruited to the AUX/IAA proteins through interaction with a conserved protein motif known as an Ethylene-responsive element binding factor-associated amphiphilic repression (EAR) domain (cite).
-Canonical EAR domains have the amino acid sequence LxLxL, as is found in domain I of the AUX/IAAs (figure xA, cite).
-TPL co-repressors bind EAR domains through direct interaction with the TPL N-terminal C-terminal to LisH (CTLH) domain (citations of pre-structure founding papers/reviews).
-Recent structural analyses of the TPL N-terminal domain have highlighted the precise interaction interface between TPL and AUX/IAA EAR domains, as well as the TPL-TPL dimerization and tetramerization motifs (Avarelilo et al 2017, Ke et al, 2015), and even the residues required for higher-order multimers of TPL tetramers (Ma et al, 2017).
-The two tandem beta propellar domains of TPL/TPRs have as yet not been crystallized, however numerous high resolution crystal structures of beta propellar domains allow for structural prediction of the TPL c-terminal beta propellar domains and the locations of likely interaction faces (Figure, X).
+The Auxin signaling pathway utilizes the TOPLESS (TPL) and TOPLESS-related (TPR) family of Gro/TLE/TUP1 type co-repressor proteins to maintain auxin responsive genes in a transcriptionally-repressed state in the absence of auxin (cite).
+In *Arabidopsis thaliana* the TPL/TPR family includes *TPL* and four *TPR1-4*.
+The resulting proteins are comprised of three structural domains: an N-terminal TPL domain and two WD-40 domains.
+TPL/TPR proteins are recruited to the AUX/IAA proteins through interaction with the conserved Ethylene-responsive element binding factor-associated amphiphilic repression (EAR) domain (cite).
+Canonical EAR domains have the amino acid sequence LxLxL, as found in most AUX/IAAs (figure xA) [@overvoorde_functional_2005].
+TPL/TPR co-repressors bind EAR domains via their C-terminal to LisH (CTLH) domains found near their N-termini (citations of pre-structure founding papers/reviews).
+Recent structural analyses of the TPL N-terminal domain have highlighted the precise interaction interface between TPL and AUX/IAA EAR domains, as well as the TPL-TPL dimerization and tetramerization motifs (Avarelilo et al 2017, Ke et al, 2015).
+The residues required for higher-order multimers of TPL tetramers have also been identified (Ma et al, 2017).
+Additional interactions with transcriptional regulation and chromatin modifying machinery are likely mediated by two tandem beta propeller domains of TPL/TPRs.
+These domain have not yet been crystallized, however numerous high resolution crystal structures of beta propeller domains allow for structural prediction of the TPL c-terminal beta propeller domains and the locations of likely interaction faces (Figure, X).
 
-  The TOPLESS co-repressor family generally exhibits a high level of sequence conservation at the amino acid sequence level across the Arabidopsis accessions with all ${\pi_N/\pi_S}$ values below 1 (Table X).
-The closely related TPL and TPR1 have the highest ${\pi_N/\pi_S}$ values (TPL-0.425, TPR1-0.739), suggesting that these these two related genes tolerate a higher degree of sequence diversity compared to TPR2-4.
-The TPL/TPR family demonstrates a high level of sequence conservation across the N-terminal TPL domain, with the lowest of observed ${\pi_N/\pi_S}$ values (Figure X).
-Indeed, all non-synonymous mutations observed in this region are either in linker domains between alpha helices, or are highly conservative mutations within helices (i.e.
-Valine to Isoleucine), which would be predicted to exhibit little effect on TPL folding and activity.
+The TOPLESS co-repressor family generally exhibits a high level of sequence conservation at the amino acid sequence level across resequenced *Arabidopsis thaliana* accessions, with all ${\pi_N/\pi_S}$ values below 1 (Table X).
+The closely related TPL and TPR1 have the highest ${\pi_N/\pi_S}$ values (TPL-0.425, TPR1-0.739), suggesting that these these two related genes tolerate a higher degree of sequence and potentially functional diversity compared to TPR2-4.
+The N-terminal TPL domain of the TPL/TPR family is particularly conserved (Figure X).
+All non-synonymous polymorphisms observed in this region are either in the coils between helices or are highly conservative mutations within helices (i.e.
+Valine to Isoleucine), which would be predicted to exhibit little effect on TPL folding and function.
 
-  The increased N-terminal conservation underscores the importance of the N-terminal TPL domain.
-For example, the initial tpl-1 mutation (N176H) in the ninth helix is a dominant gain-of-function allele (Long et al, 2002, PMID12050130), which is capable of binding wild-type TPL protein and inducing protein agreggation (Ma, 2017, pmid28630893).
+The high degree of conservation in the entire N-terminal domain underscores its importance.
+For example, the initial tpl-1 mutation (N176H) in the ninth helix is a dominant gain-of-function allele (Long et al, 2002, PMID12050130), which is capable of binding wild-type TPL protein and inducing protein aggregation (Ma, 2017, pmid28630893).
 It is therefore understandable that very little diversity is observed in the ninth alpha helix of TPL/TPRs as mutations in this domain could act in a dominant negative fashion, rendering any de novo mutation in this region embryo lethal.
@@
 
-  We then compared the frequency of SNPs that cause nonsynonymous mutations across the TPL domains.
-Compared to the TPL domain, both the proline rich and WD-40 domains exhibit a higher rate of missense mutations accross both of the WD-40 repeat regions, here labeled WD40-1 and WD40-2 respectively.
-We utilized the Phyre2.0 modeling software to generate a rough model of the TPL/TPR WD-40 repeats stricly for the purpose of mapping non-synonymous mutations onto a visible protein chassis (Figure X).
-Here we can see both faces of each predicted beta-propellar domains, which are connected by a short linker region.
-Through utilization of protein modelling, it becomes clear that we observe conservation of amino acid sequence at the center of the beta-propellar domains.
-These observations could be used as the basis for forward mutagensis strategies to test whether this region of TPL functions similarly to its homolog TLE, which binds its target genes EH1 and the C-terminal WRPW/Y motif (Hairy/Hes/Runx) in the center of the beta-propeller domain (pmid 16762837).
+Compared to the TPL domain, both the proline rich and WD-40 domains exhibit a higher rate of missense mutations.
+We utilized the Phyre2.0 modeling software to predict the structure of the TPL/TPR WD-40 repeats in order to map non-synonymous mutations onto a visible protein chassis (Figure X).
+Here we can see both faces of each predicted beta-propeller structure of the WD-40 domains, which are connected by a short peptide linker.
+This model reveals amino acid conservation in the core and central residues of the WD-40 domains.
+~~These observations could be used as the basis for forward mutagenesis strategies to test whether this region of TPL functions similarly to its homolog TLE, which binds its target genes EH1 and the C-terminal WRPW/Y motif (Hairy/Hes/Runx) in the center of the WD-40 domain (pmid 16762837).~~ **Can we say something more general here?
+There is not enough background to support this.**
 
-  It is interesting to note that despite the presence of 5 TPL/TPR genes in the genome, which are considered to be redundant (pmid 16763149), very few SNPs are observed that are predicted to result in loss of function for any one TPL/TPR gene.
-In fact, only one deletion was observed (665delG) that creates a frameshift mutation in TPR4 in a single accession (Aiell-4 in Italy).
-This mutation occurs in codon 222, just downstream of the TPL N-terminal domain, and would generate an early stop codon 11 codons downstream of the deletion site.
+Despite the presence of 5 TPL/TPR genes in the genome, which are considered to be redundant (pmid 16763149), very few SNPs are observed that are predicted to result in loss of function for any TPL/TPR gene.
+Only one frameshift variant  was observed, TPR4-665delG, in a single accession (Aiell-4 from Italy).
+This deletion occurs in codon 222, just after the TPL N-terminal domain, and generates an early stop 11 codons downstream from the deletion.
 It is unclear whether this N-terminal fragment would be expressed at any appreciable level, as it is possible that the generation of an early stop codon would trigger nonsense-mediated decay.
-However, if expressed, it would be fascinating to examine whether this naturally occuring N-terminal truncation retains activity as a co-repressor.
+However, if expressed, it would be fascinating to examine whether this naturally occurring N-terminal truncation retains activity as a co-repressor.
 Given the observation that the TPL N-terminal domain is sufficient to bind histones (28630893), and repress transcription via synthetic assays in yeast (24979769), it is possible that this represents a semi-functional TPL variant.
-There has been no evidence for priotization of individual TPL/TPR family members in the auxin response pathway, but experiments modulating the gene copy number of TPR1 revealed the importance of maintaing the proper dosage of TPL/TPR in order to accurately trigger the salicyclic acid innate immune response (20647385).
-These results, taken together with the absence of TPL/TPR loss-of-function alleles in the Arabidopsis accessions suggest that perhaps the entire family is being conserved at the nucleotide level in the population to meet a gene dosage requirement, with little opportunity for neo- or sub-functionalization.
+There has been no evidence for prioritization of individual TPL/TPR family members in the auxin response pathway, but experiments modulating the gene copy number of TPR1 revealed the importance of maintaining the proper dosage of TPL/TPR in order to accurately trigger the salicylic acid innate immune response (20647385).
+These results, taken together with the absence of TPL/TPR loss-of-function alleles in the *Arabidopsis thaliana* accessions suggest that perhaps the entire family is being conserved at the nucleotide level in the population to meet a gene dosage requirement, with little opportunity for neo- or sub-functionalization.
 
@@
 
-Auxin response is mediated by the auxin responsive transcription factors, or ARFs.
-There are 23 ARFs in Arabidopsis that are divided into three phylogenetic classes.
-Class A ARFs, ARF5, ARF6, ARF7, ARF8, and ARF19, activate transcription.
-These ARFs have a Q-rich region in the middle of the protein that may mediate activation (Guilefoyle, 2007).
+Auxin response is ultimately mediated by the auxin responsive transcription factors (ARFs).
+There are 23 ARFs in *Arabidopsis thaliana* that are divided into three phylogenetic classes.
+Class A ARFs (ARF5, ARF6, ARF7, ARF8 and ARF19) activate transcription.
+These ARFs have a glutamine-rich region in the middle of the protein that may mediate activation (Guilefoyle, 2007).
 It has recently been shown that the middle region of ARF5 interacts with the SWI/SNF chromatin remodeling ATPases BRAMA and SPLAYED, possibly to reduce nucleosome occupancy and allow for the recruitment of transcription machinery (Wu, 2016).
-Additionally ARF7 can interact with Mediator subunits, directly tethering transcriptional proteins to its binding regions (Ito, 2016).
+Additionally, ARF7 interacts with Mediator subunits, directly tethering transcriptional activation machinery to its binding sites in the chromosome (Ito, 2016).
 Class B and C ARFs are historically categorized as repressor ARFs, though the mechanism through which they confer repression has not been identified.
-Their middle regions tend to be proline- and serine-rich.
+Their middle regions tend to be proline- and serine-rich (citation).
 
-ARFs are comprised of three major domains, recent structures of which have informed structure-function analysis of the ARFs (Boer, 2014; Nanao, 2014; Korasick, 2014).
+Canonical ARFs are comprised of three major domains.
+Recent crystallization of these domain have informed structure-function analysis of the ARFs (Boer, 2014; Nanao, 2014; Korasick, 2014).
 These domains are conserved throughout land plants.
@@
 Flanking this DNA-binding domain is a dimerization domain, which folds up into a single “taco-shaped” domain to allow for dimerization between ARFs.
-There is an auxillary domain that immediately follows and interacts with the dimerization domain.
+There is an auxiliary domain that immediately follows and interacts with the dimerization domain.
 The middle region is the most variable between ARFs, as mentioned above, but is characterized by repetitive units of glutamine (class A), serine, or proline residues (classes B and C).
-On the C-terminal end of the protein a protein-protein interaction domain mediates interactions among ARFs, between ARFs and other transcription factors, and between ARFs and the Aux/IAA repressors.
+
+The C-terminal domain of the ARF is a protein-protein interaction domain mediating interactions among ARFs, between ARFs and other transcription factors, and between ARFs and the Aux/IAA repressors.
 This interaction domain was recently characterized as a Phox and Bem1 (PB1) domain, which is comprised of a positive and negative face with conserved basic and acidic residues, respectively.
-Theoretically the dipolar nature of the PB1 domain can mediate multimerization by the pairwise interaction of these faces on different proteins, and indeed the ARF7 PB1 domain was crystallized as a multimer.
-It is unclear whether this multimerization occurs or plays a developmental significant role in vivo.
+The dipolar nature of the PB1 domain may mediate multimerization by the pairwise interaction of these faces on different proteins as the ARF7 PB1 domain was crystallized as a multimer [@korasick_].
+However, it is unclear whether this multimerization occurs or plays a developmental significant role in vivo.
 Interfering with ARF dimerization in either the DNA-binding proximal dimerization domain or the PB1 domain decreases the ability of class A ARFs to activate transcription in a heterologous yeast system (Pierre-Jerome, 2016).
@@
 The majority of domain variation among ARFs occurs in the large B-class subfamily.
-The liverwort Marcantia polymorpha has a single representative ARF of each class (Flores-Sandoval, 2015).
-The much expanded classes of the flowering plants arise from both whole genome and tandem duplication events (Remington, 2004), and may have allowed for the expansion of the quantity and complexity of loci regulated by the ARFs and a consequent expansion in their regulation of developmental processes.
-Here we examine natural variation in the three ARF classes and how this variation may be informative of conservation and function among the ARFs.
+The liverwort *Marchantia polymorpha* has a single representative ARF of each class (Flores-Sandoval, 2015).
+The expansion of these classes in flowering plants is the result of both whole genome and tandem duplication events (Remington, 2004).
+The growth of the ARF family may have allowed for the expansion of the quantity and complexity of loci regulated by the ARFs and subsequent expansion in their regulation of developmental processes.
 
-### Class A ARFs
-Amy
-
-Class A ARFs are the most well-studied ARF subfamily--the five family members all act as transcriptional activators and have well-characterized, distinct developmental targets.
-Class A ARFs all contain the canonical B3 DNA-binding domain and PB1 interaction domain, and all have glutamine-rich middle regions (Guilefoyle, 2007).
-Analysis of class A ARF nucleotide diversity showed that the majority of these ARFs had low nucleotide diveristy scores, all ${\pi_N/\pi_S}$ values substantially lower than 1 with the exception of ARF19, with a relatively high ${\pi_N/\pi_S}$ value of 1.8.
-Comparing diversity within sister pairs, the two class A ARF sister pairs had divergent diversity scores, with ARF19 and ARF8 having scores at least triple those of ARF7 and ARF6 respectively.
-This may suggest that ARF6 and ARF7 have more conserved function whereas ARF19 and ARF8 are more diverse.
+Class A ARFs are the most well-studied ARF subfamily---the five family members all act as transcriptional activators and have well-characterized, distinct developmental targets.
 Overall the diversity of class A ARFs was generally low, especially compared to the class B and C ARFs, suggesting that class A ARFs are central to auxin signal transduction and plant development.
+Analysis of class A ARF nonsynonymous diversity suggests that the majority of these ARFs are highly functionally conserved, with ${\pi_N/\pi_S}$ values much lower than 1 with the exception of ARF19, with ${\pi_N/\pi_S}$ value of 1.8.
+Comparing diversity within sister pairs, there is a similar trade-off as seen in most IAA sister pairs, with one sister being highly conserved and the other more divergent.
+ARF19 and ARF8 are the more divergent class A ARFs, with ${\pi_N/\pi_S}$ values at least three time those of their sisters, ARF7 and ARF6 respectively.
+This may suggest that ARF6 and ARF7 serve more essential purposes in plant development.
 
@@
 
+Class A ARFs all contain the canonical B3 DNA-binding domain, PB1 interaction domain and glutamine-rich middle region (Guilefoyle, 2007).
 Within the middle region, ARF7 had several expansions of polyglutamine sequences compared to the other class A ARFs.
-Polyglutamine regions are known to readily expand and contract throughout evolutionary time due to replication error, and variation in polyQ length can be acted on by natural selection and have phenotypic consequences (Press, 2014).
-
-A few missense variations in ARF5 highlighted potential functional variation among ecotypes.
-Two high-diversity changes at the positive face of the PB1 domain, very close to the conserved lysine residue, to phenylalanine were interstingly always found together in all ecotypes in which they occured: Tu-B2-3,Ru-2, Haes-1, Rd-0, 627ME-13Y1, Le-0,Bu-0, Bsch-0.
-These mutations to a rather bulky side chain may change the folding of the PB1 domain and interfere with interactions on the positive face.
-A high diversity polymorphism that results in change in the 17th amino acid of ARF5 from a leucine to a methionine is also intriguing.
-ARF5 has an N-terminal extension of unknown function relative to the other class A ARFs *how many amino acids*.
-It is possible that this methionine acts as an alternative start site in the accessions in which it is found yielding a truncated ARF5 variant without this N-terminal extension.
+Polyglutamine regions are known to readily expand and contract throughout evolutionary time due to replication error, and variation in polyglutamine length can be acted on by natural selection and have phenotypic consequences (Press, 2014).
+A few missense variants in ARF5 highlighted potential functional variation in the class A ARFs.
+Two highly represented variants resulting in substitutions of small amino acids for the bulky phenylalanine were found on the positive face of the PB1 domain, very close to the conserved lysine residue and may disrupt interactions with other ARF or Aux/IAA PB1 domains.
+Interestingly, these variants were always found together in accessions Tu-B2-3, Ru-2, Haes-1, Rd-0, 627ME-13Y1, Le-0, Bu-0 and Bsch-0.
+Another highly represented polymorphism in ARF5 results in change in the 17th amino acid from a leucine to a methionine.
+This position is in the N-terminal extension of ARF5 which is of unknown function.
+It is possible that this methionine acts as an alternative start site in the accessions.
 
@@
 These ARFs span the range of ${\pi_N/\pi_S}$ ratios relative to the other ARFs, ranging from 0.85 to 4.1.
-ARF13, with the highest ${\pi_N/\pi_S}$ of 4.1 has many nonsense variants of high diversity, perhaps suggesting it is pseudogenizing.
+ARF13, with the highest ${\pi_N/\pi_S}$ of 4.1 has many nonsense variants of high diversity.
 
 ARF14, ARF15, ARF16, ARF20, and ARF21 all do not have one or multiple of the conserved acidic residues in the PB1 domain, and themselves form a subclade of the B ARFs.
-Interestingly, no non-synonymous variation is seen in these amino acid residues of these ARFs, suggesting that the amino acids that are in the regions may still be important for PB1 function.
+Interestingly, no non-synonymous variation is seen in these amino acid residues of these ARFs, suggesting that the amino acids that are in the regions may still be important for the function of these proteins.
 
@@
 
-Class C ARFs shows low nucleotide diveristy scores, all pi_N/pi_S values substantially lower than 1.
-Comparing diversity within coding region of class C ARF, ARF16  has least diversity whereas its family members (ARF10, ARF17) shows  scores at least four times higher [@remington_contrasting_2004].
-Structurally, all three members of Class C ARFs contain both canonical B3 DNA-binding domain and whereas only ARF10 and ARF16 contains PB1 interaction domain.
-DNA binding domain exhibit overall low diversity and presence of largely synonymous changes for all the ARFs.
-Only ARF10, shows mis-sense variants (p.Ser147Thr/c.439T>A  and p.Ala157Ser/c.469G>T) DNA binding domains with higher diversity score for conservative variation (codon 147) compared to non conservation variation (codon 157) (Fig.
+Class C ARFs show low nucleotide diversity scores, all pi_N/pi_S values substantially lower than 1.
+ARF16 was the most conserved, whereas its clade members (ARF10, ARF17) had scores at least four times higher [@remington_contrasting_2004].
+
+Structurally, all three members of Class C ARFs contain a canonical B3 DNA-binding domain, but only ARF10 and ARF16 contain a PB1 domain.
+The DNA binding domains exhibit overall low diversity and presence of largely synonymous changes for all the ARFs.
+Only the ARF10 DNA binding domain contains missense variants (S147T and A157S) (Fig.
 ARF10_DBD Variants distribution).
-Of the PB1 domain containing class C ARFs, ARF16 exhibits several naturally occurring non-synonymous variants in contrast to ARF10 (Fig.
-Arf16_PB_variants distribution) and is sporadically distributed in location.
+**Any insights into whether these may affect function?
+They are quite conservative.** Of the PB1 domain containing class C ARFs, ARF16 exhibits several missense variants which are sporadically distributed, in contrast to the conserved PB1 domain of ARF10 (Fig.
+Arf16_PB_variants distribution) .
 
-Few members of class B ARFs (ARF 2,3,4) and C ARFs (ARF10,16, 17) distinguish from other ARFs on the basis of its ability to be post transcriptionally regulated by trans-acting small interfering RNA (tasiRNA) and microRNA (miRNA) [@rhoades_prediction_2002; @allen_micrornadirected_2005].
-(@AMY Arf 6 and 8 targeted by miRNA 167 - check for conservation).
-Interestingly, miRNA binding motif of the class C ARFs show high degree of conservation across the natural variation of 1001 genomes queried here (Fig.
+Several members of class B (ARF2, 3 and 4) and C ARFs (ARF10, 16 and 17) are distinguished from other ARFs on the basis of their post transcriptional regulation by trans-acting small interfering RNAs (tasiRNAs) and microRNAs (miRNAs) [@rhoades_prediction_2002; @allen_micrornadirected_2005].
+**(@AMY Arf 6 and 8 targeted by miRNA 167 - check for conservation).** The miRNA target of class C ARFs showed a high degree of conservation (Fig.
 ARF10_16_17_Alignment).
 This is consistent with perturbation studies that has shown drastic developmental phenotype for class C ARFs [@wang_control_2005; @mallory_micrornadirected_2005; @liu_repression_2007].
-For the class B ARFs (2,3,4) tasiRNA binding motif is highly conserved and show no variation suggesting strong regulatory constrains at the locus (Fig ARF2_3_4_Alignment).
+For the class B ARFs (2, 3 and 4) tasiRNA binding motif is also highly conserved suggesting strong regulatory constrains at the locus (Fig ARF2_3_4_Alignment).
 DNA binding domain of ARF4 exhibits non-synonymous variation in 4 accession and would be interesting to carry out phenotypic characterize them for developmental growth defects (Fig ARF4_DBD Variants distribution).
+
+#Discussion

### 2018-03-14 manuscript/manuscript.Rmd
made some figures for the AFB section, developed py code for mapping variants onto structures but this is still buggy

@@
 knitr::opts_chunk$set(warning = FALSE, collapse = TRUE, comment = "#>", tidy = TRUE, out.width = "90%", fig.height=4, fig.width=7, out.extra='style="margin: auto; display: block; padding-top: 15px;"')
+devtools::install_github("wrightrc/r1001genomes", ref = "auxin-natural-variation")
 library(r1001genomes)
 library(DT)
-library(dplyr)
-
+library(tidyverse)
+library(magrittr)
+library(RColorBrewer)
+library(ggpmisc)
+library(ggthemes)
 ```
@@
 
-Examining the natural sequence variation across the AFB revealed that *TIR1* and *AFB1* both had very low nonsynonymous diversity, hinting at their likely functional importance.
-*AFB3* and *AFB4* had higher nonsynonymous diversity, while their sisters genes, *AFB2* and *AFB5* are more conserved.
+```{r}
+gene_info <- geneInfoFromFile(
+  fname = system.file("shiny-app", "gene_family_data",
+                      "AFB_gene_ids.csv", package = "r1001genomes"))
+annotation <- readAnnotationFile(filename = system.file("shiny-app", "gene_family_data",
+                      "AFB_annotations.csv", package = "r1001genomes"), gene_info = gene_info)
+alignment <- alignCDS(gene_info$transcript_ID)
+# make an alignment data frame
+aln_df <- makeAlnDF(alignment[[2]])
+
+# make a collapsed alignment for translating positions
+collapsed_aln_df <- 
+  aln_df %>% dplyr::select(aln_pos, seq_pos, seq_name) %>%
+  tidyr::spread(key = seq_name, value = seq_pos)
+
+VCF <- readRDS(system.file("shiny-app", "gene_family_data",
+                      "AFBs_VCF.rds", package = "r1001genomes"))
+div_stats <- ldply(VCF, diversityStats, geneInfo=gene_info, .id="transcript_ID")
+formatRound(datatable(div_stats,
+                  colnames = c("transcript",
+                               "symbol",
+                               "&pi;<sub>N</sub>",
+                               "&pi;<sub>S</sub>",
+                               "&pi;<sub>N</sub>/&pi;<sub>S</sub>",
+                               "&pi; coding",
+                               "&pi; transcript"),
+                  rownames = FALSE, escape = FALSE,
+                  options = list(paging=FALSE, searching=FALSE)),
+                columns = 2:7, digits = 6)
+```
+
+Examining the natural sequence variation across the AFB family revealed that *TIR1* and *AFB1* both had very low nonsynonymous diversity, hinting at their likely functional importance.
+*AFB3* and *AFB4* had higher nonsynonymous diversity, while their sisters genes, *AFB2* and *AFB5* were more conserved.
 This matches our current understanding of *AFB3* and *AFB4* as being minor players in the auxin signaling pathway, but brings in to question the inconclusive role of *AFB1* in auxin signaling.
-Although most important functional regions do not contain variants in AFB1, there is a low diversity mutant, only found in the Can-1 accession, which is in the highly conserved TIR1/AFB oligomerization domain.
-Because variants eliciting significant changes in function often are not selected for, the fact that this variant is extremely low diversity may suggest that it is associated with a strong auxin signaling phenotype.
+
+Although most known functional regions are highly conserved in AFB1, there is a nonsynonymous polymorphism in the TIR1/AFB oligomerization domain, only found in the Can-1 accession.
+Mutations in this domain of TIR1 frequently have a semidominant effect on root phenotypes [@dezfulian_oligomerization_2016; @wright_insights_2017a].
 Characterization of this accession may identify potential *AFB1* phenotypes and may also test the hypothesis that AFB1 functions through oligomerization with other TIR1/AFBs.
+
+```{r AFB1 variants for PyMOL}
+coding_vcf <- ldply(.data = VCF, .fun = subset, 
+             !is.na(Transcript_ID) & gt_GT != "0|0")
+aln_df <- addSNPsToAlnDF(aln_df, coding_vcf, seq_name = Transcript_ID, 
+               seq_pos = Codon_Number)
+collapsed_aln_df %<>% left_join(y = subset(aln_df, seq_name == gene_info[gene_info$tair_symbol == "TIR1", "transcript_ID"]))
+
+#' Select variants for importing into PyMOL
+#'
+#' @param gene_info a gene info data frame
+#' @param aln_df an alignment data frame with attached SNPs
+#' @param structure_ref string, which tair_symbol or transcript_ID (depending on 
+#' \code{name_by_symbols}) corresponds to the protein structure to which 
+#' selections should be mapped
+#' @param effect_type string of the effect type to select from the alignment 
+#' data frame
+#' @param name_by_symbols logical, if true (default) tair_symbols will be used 
+#' to name the selections, otherwise transcript_IDs will be used
+#'
+#' @return
+#' @export
+#'
+#' @examples
+select_variants <- function(gene_info, aln_df, structure_ref, 
+                            effect_type = "missense", name_by_symbols = TRUE){
+  collapsed_aln_df <- aln_df %>% dplyr::select(aln_pos, seq_pos, seq_name) %>%
+    tidyr::spread(key = seq_name, value = seq_pos)
+  if(name_by_symbols)
+    column <- "tair_symbol"
+    else 
+      column <- "transcript_ID"
+
+  selections <- ldply(.data = gene_info[,column], .fun = function(gene){
+    collapsed_aln_df %<>% dplyr::left_join(y = subset(aln_df, seq_name ==  
+                      gene_info[gene_info[, column] == gene, "transcript_ID"]))
+    variants <- collapsed_aln_df[
+      grepl(effect_type, collapsed_aln_df$effects), 
+      gene_info[gene_info[, column] == structure_ref, "transcript_ID"]]
+    c(gene, paste0(na.omit(as.numeric(variants[[1]])), collapse = "+"))
+  })
+  return(selections)
+}
+
+selections <- select_variants(gene_info, aln_df, structure_ref = "TIR1")
+selections$V2 <- paste0("chain B and resi ", selections$V2)
+write.table(selections, file = "AFB_variant_selections", row.names = FALSE, 
+          col.names = FALSE)
+```
+
+```{python}
+#
+# -- readSelection.py
+#
+from pymol import cmd, stored
+
+def readSelection( fileName ):
+    '''
+DESCRIPTION
+
+    reads a csv file containing a column of selection names and a column of
+    selection strings and creates selection objects for each
+    '''
+	#read in the selection file
+	with open(filename) as file:
+		selections = file.readlines() 
+		#create a selection for each row
+		for selection in selections:
+			selection = line.split(",")
+			cmd.select(selection[1], selection[2])
+
+	file.close()
+
+cmd.extend( "readSelection", readSelection );
+
+```
+
+```{pymol}
+import readSelection
+
+```
+
+```{r AFB1 diversity plot}
+coding_variants <- getCodingDiv(VCF[[gene_info[gene_info$tair_symbol == "AFB1", "transcript_ID"]]])
+p <- plotCodingDiv(uniqueCodingVars = coding_variants)
+annotation <- addAlnPosToAnno(anno_df = annotation, aln_df)
+annotation
+# translate positions from TIR1 to AFB1
+
+AFB1_positions <- annotation$positions
+AFB1_positions %<>% left_join(y = select(collapsed_aln_df, aln_pos, gene_info[gene_info$tair_symbol == "AFB1", "transcript_ID"]), by = "aln_pos") 
+AFB1_positions$position <- AFB1_positions$AT4G03190.1
+
+p <- append_layers(p,list(
+        geom_rect(data = AFB1_positions,
+                   mapping = aes(xmin = as.integer(position)-1,
+                                 xmax = as.integer(position)+2,
+                                 fill = annotation),
+                  ymin = -Inf, ymax = Inf, inherit.aes = FALSE, alpha = 0.8)),
+      position = "bottom")
+p
+```
+
 The AFB4 and AFB5 receptors have an N-terminal extension prior to the F-box domains.
@@
 
-Within the Group B IAA genes, there are four sister pairs--IAA10/IAA11, IAA12/IAA13, IAA18/IAA26, and IAA20/IAA30 (Fig X) [@remington_contrasting_2004].
+Within the Group B IAA genes, there are four sister pairs---IAA10/IAA11, IAA12/IAA13, IAA18/IAA26, and IAA20/IAA30 (Fig X) [@remington_contrasting_2004].
 Interestingly, among these sister pairs only IAA12 and 13 have a noticeable difference in nonsynonymous diversity.
@@
 Accessions and associated variants identified here provide a potentially expedited route to elucidating the extent of redundancy between family members such as IAA10 and 11.
+
+```{r}
+
+```
 

### 2018-03-14 readSelection.py

@@ -1,0 +1,24 @@
+from pymol import cmd
+
+def readSelection (fileName):
+	'''
+	DESCRIPTION
+
+	reads a csv file containing a column of selection names and a column of
+	selection strings and creates selection objects for each
+	'''
+	line = []
+	#read in the selection file
+	with open(fileName) as f:
+		for line in f:
+			sel = line.split(",")
+			print sel[0] + " " + sel[1]
+			#cmd.select(selection[0], selection[1])
+
+	f.close()
+
+cmd.extend('readSelection',readSelection)
+
+
+
+

### 2018-03-14 updates_from_package.R

@@ -1,0 +1,24 @@
+plotCodingDiv <- function(uniqueCodingVars){
+
+  effectClasses <- readRDS(system.file("data", "effect_classes.rds", package="r1001genomes"))
+
+  classColors <- data.frame("color" = brewer.pal(n = 5, name = "RdYlBu")[5:1],
+                            "Class" = c("Synonymous", "Non_Coding", "Splice",
+                                        "Missense", "Nonsense"),
+                            "labels" = c("Synonymous", "Non_Coding", "Splice",
+                                         "Missense", "Nonsense"), stringsAsFactors=FALSE)
+
+  uniqueCodingVars <- dplyr::left_join(uniqueCodingVars, effectClasses, by = "Effect")
+  uniqueCodingVars <- dplyr::left_join(uniqueCodingVars, classColors, by = "Class")
+  classes <- classColors$Class %in% unique(uniqueCodingVars$Class)
+  #plot the diversity
+  plot <- ggplot(uniqueCodingVars, aes(x=Codon_Number,y=Diversity, colour=color, shape = Effect)) +
+    geom_point(size = 4, alpha = .9, position = position_jitter(height = 0.2)) +
+    scale_y_log10(breaks=c(0.001, 0.01, 0.1),limits=c(0.001, 1)) +
+    #scale_colour_manual(values=c(synonymous_diversity="blue", missense_diversity="red")) +
+    ylab("nucleotide diversity, log scale") + theme_few(base_size = 18) +
+    scale_color_identity("Class", breaks = classColors$color[classes],
+                         labels = classColors$labels[classes],
+                         guide = "legend")
+  return(plot)
+}

### 2018-05-22 manuscript/manuscript.Rmd
Added IAA tree figure

@@
 ---
-title: "A comprehensive analysis of natural sequence variation within the <i>Arabidopsis thaliana</i> nuclear auxin signaling pathway using an accessible web application"
+title: A comprehensive analysis of natural sequence variation within the <i>Arabidopsis
+  thaliana</i> nuclear auxin signaling pathway using an accessible web application
 author:
   affiliation: University of Washington
-  name: "R. Clay Wright"
-
+  name: R. Clay Wright
 date: "`r Sys.Date()`"
-
 output:
+  word_document:
+    toc: yes
   html_document:
-      toc: true
     fig_caption: yes
-
+    toc: yes
+fontfamily: Helvetica
+fontsize: 11pt
 keywords: r markdown
-fontsize: 11pt
-fontfamily: Helvetica
-
-bibliography: ["natural_variation.bib", "TIR1_AFB.bib", "ARF_C.bib", "AuxIAA_citations.bib"]
-
+link-citations: yes
 csl: apa.csl
-link-citations: true
+bibliography:
+- natural_variation.bib
+- TIR1_AFB.bib
+- ARF_C.bib
+- AuxIAA_citations.bib
 ---
 
-```{r, echo = FALSE, message=FALSE, warning=FALSE}
-knitr::opts_chunk$set(warning = FALSE, collapse = TRUE, comment = "#>", tidy = TRUE, out.width = "90%", fig.height=4, fig.width=7, out.extra='style="margin: auto; display: block; padding-top: 15px;"')
-devtools::install_github("wrightrc/r1001genomes", ref = "auxin-natural-variation")
+```{r setup echo = FALSE, message=FALSE, warning=FALSE}
+knitr::opts_chunk$set(warning = FALSE, collapse = TRUE, comment = "#>", tidy = TRUE, cache = TRUE, out.width = "90%", fig.height=4, fig.width=7, out.extra='style="margin: auto; display: block; padding-top: 15px;"')
+#devtools::install_github("wrightrc/r1001genomes", ref = "auxin-natural-variation")
 library(r1001genomes)
@@
 library(ggthemes)
+library(ggseqlogo)
+library(ggtree)
+library(ape)
+library(reticulate)
+library(scales)
+library(viridis)
+use_python(python = "/usr/local/bin/python")
 ```
@@
 AFB1 has a mutation drastically reducing its ability to assemble into an SCF complex, contributing to its increased accumulation, as well as preventing it from degrading Aux/IAAs[@Yu_Untethering_2015].
-Higher order TIR1/AFB mutants show AFB1 to have inconclusive organismal function.
-However, AFB1 is shown to be able to accumulate ubiquitously in all tissues.AFB4 and AFB5 have been shown to preferentially and functionally bind the synthetic auxin picloram, while others have a preference for naturally occurring auxin indole-3-acetic acid [@Prigge_Arabidopsis_2016].
+Higher order TIR1/AFB mutants show AFB1 to have inconclusive function.
+However, AFB1 is shown to be able to accumulate ubiquitously in all tissues [@Parry_Complex_2009].
+AFB4 and AFB5 have been shown to preferentially and functionally bind the synthetic auxin picloram, while others have a preference for naturally occurring auxin, indole-3-acetic acid [@Prigge_Arabidopsis_2016].
 Interestingly, the strength and rate with which TIR1/AFBs are able to bind and mark Aux/IAAs for degradation are variable [@Havens_Synthetic_2012].
@@
 
-```{r AFB1 variants for PyMOL}
+```{r AFB1 variants for PyMOL, eval=FALSE}
 coding_vcf <- ldply(.data = VCF, .fun = subset, 
@@
 write.table(selections, file = "AFB_variant_selections", row.names = FALSE, 
-          col.names = FALSE)
+          col.names = FALSE, sep = ",", quote = FALSE)
 ```
 
-```{python}
+```{python, eval=FALSE}
 #
@@
 
-def readSelection( fileName ):
+def readSelection (fileName):
 	'''
@@
 	#read in the selection file
-	with open(filename) as file:
-		selections = file.readlines() 
-		#create a selection for each row
-		for selection in selections:
-			selection = line.split(",")
-			cmd.select(selection[1], selection[2])
+	with open(fileName) as f:
+		lines = f.readlines()
+		for line in lines:
+			sel = line.split(",")
+			cmd.select(sel[0], sel[1])
 
-	file.close()
+	f.close()
 
-cmd.extend( "readSelection", readSelection );
-
+cmd.extend('readSelection',readSelection);
 ```
 
-```{pymol}
-import readSelection
+```{pymol, eval=FALSE}
+run ~/Desktop/natural-variation-webtool/manuscript/readSelection.py
+
+readSelection ~/Desktop/natural-variation-webtool/manuscript/AFB_variant_selections
 
@@
 Yellow squared E’s show Cul1-interacting residues, differing in AFB1.
+```{r TIR1/AFB Alignment}
+aln_df <- makeAlnDF(alignment[[2]])
+aln_df <- left_join(aln_df, dplyr::select(gene_info, "tair_locus",
+                                       "tair_symbol", "transcript_ID"),
+                        by = c("transcript_ID" = "transcript_ID"))
+aln_df$tair_symbol <- fct_rev(fct_relevel(aln_df$tair_symbol, "TIR1"))
+aln_df <- chunkAlnDF(aln_df = aln_df, chunk_width = 55)
+VCF <- ldply(.data = VCF, .fun = subset, gt_GT != "0|0")
+VCF <- getCodingDiv(VCF)
+aln_df <- addSNPsToAlnDF(aln_df, VCF)
+F_box <- subset(aln_df, aln_pos <= 96)
+F_box <- chunkAlnDF(aln_df = F_box, chunk_width = 80)
+p <-ggplot(F_box, aes(x = aln_pos, y = tair_symbol,
+                             group = seq_pos, text = variants))
+p <- p +
+  geom_tile(data = na.omit(F_box), mapping = aes(fill = effects),
+            width = 1, height = 1, alpha = 0.8) +
+  geom_text(aes(label=letter), alpha= 1, family = "Courier") +
+  scale_fill_brewer(type = "qual", palette = 1, direction = -1) +
+  scale_x_continuous(breaks=seq(1,max(F_box$aln_pos), by = 10)) +
+  scale_y_discrete() +
+  # expand increases distance from axis
+  labs(x = "", y = "", fill = "variant") +
+  theme_logo(base_family = "Helvetica") +
+  theme(panel.grid = element_blank(), panel.grid.minor = element_blank()) +
+  facet_wrap(facets = ~chunk, ncol = 1, scales = "free") +
+  theme(strip.background = element_blank(),
+        strip.text.x = element_blank(),
+        legend.box = "vertical")
+p
+```
 
@@
 IAA-mediated repression of ARFs depends upon both an EAR domain, N-terminal of the degron, that recruits co-repressors of the TOPLESS (TPL) family, and a Phox and Bem1 (PB1) domain that enables interaction via hetero-oligomerization with ARF PB1 domains [@nanao_structural_2014; @kagale_ear_2011].
+
+```{r, fig.height=11}
+gene_info <- geneInfoFromFile(
+  fname = system.file("shiny-app", "gene_family_data",
+                      "IAA_gene_ids.csv", package = "r1001genomes"))
+annotation <- readAnnotationFile(filename = system.file("shiny-app", "gene_family_data",
+                      "IAA_annotations.csv", package = "r1001genomes"), gene_info = gene_info)
+alignment <- alignCDS(gene_info$transcript_ID)
+# make an alignment data frame
+aln_df <- makeAlnDF(alignment[[2]])
+
+# make a collapsed alignment for translating positions
+collapsed_aln_df <- 
+  aln_df %>% dplyr::select(aln_pos, seq_pos, seq_name) %>%
+  tidyr::spread(key = seq_name, value = seq_pos)
+
+VCF <- readRDS(system.file("shiny-app", "gene_family_data",
+                      "IAAs_VCF.rds", package = "r1001genomes"))
+div_stats <- ldply(VCF, diversityStats, geneInfo=gene_info, .id="transcript_ID")
+div_stats$tair_symbol %<>% fct_reorder(.x = , div_stats$tair_symbol, 
+                                       .fun = function(x) 
+                                         {as.integer(str_remove(x, 
+                                              pattern = "IAA"))})
+formatRound(datatable(div_stats[order(div_stats$tair_symbol),],
+                  colnames = c("transcript",
+                               "symbol",
+                               "&pi;<sub>N</sub>",
+                               "&pi;<sub>S</sub>",
+                               "&pi;<sub>N</sub>/&pi;<sub>S</sub>",
+                               "&pi; coding",
+                               "&pi; transcript"),
+                  rownames = FALSE, escape = FALSE,
+                  options = list(paging=FALSE, searching=FALSE)),
+                columns = 2:7, digits = 6)
+```
 
@@
 thaliana* and *A.
-lyrata*, IAA19 is more conserved, while IAA6 showed regions of increased sequence divergence upstream of the degron and in the N-terminal half of the PB1 domain [@winkler_variation_2017].
+lyrata*, IAA19 is more conserved, while IAA6 has regions of increased sequence divergence upstream of the degron and in the N-terminal half of the PB1 domain [@winkler_variation_2017].
 Consistent with these results, our analysis of intraspecific natural variation revealed higher conservation for IAA19 (${\pi_N/\pi_S}$ = 0.55) compared to IAA6 (${\pi_N/\pi_S}$ = 2.3) (Figure IAA phylogeny heat map), and also detected high diversity within the same regions of IAA6 as seen in Winkler et al.
 (Supp Fig IAA6 diversity plot).
+```{r eval=FALSE}
+write.nexus.data(as.matrix(alignment[[2]]), file = "IAAs.nex", format = "protein")
+aliscore(as.AAbin(alignment[[2]]), path = "/Volumes/SSD/Applications/ALISCORE_v2.0/Aliscore_v.2.0/")
+id <- scan(file = "input.fas_List_random.txt", sep = " ", quiet = TRUE)
+AAAlign_no_gaps <- as.matrix(alignment[[2]])[, id]
+write.nexus.data(AAAlign_no_gaps, file = "~/Desktop/natural-variation-webtool/manuscript/IAAs.nex", format = "protein")
+```
+```{bash eval=FALSE}
+mb
+
+execute /Volumes/SSD/Users/clayw/Desktop/natural-variation-webtool/manuscript/IAAs.nex
+# set IAA33 as the outgroup
+outgroup AT5G57420.1
+
+lset nucmodel=protein Rates=invgamma
+# or include Nst=6  ngammacat=4 omegavar=M3 
+prset aamodelpr=fixed(jones) statefreqpr=fixed(empirical) 
+# initial run with aamodelpr=mixed yielded posterior prob of 1 for jones model
+propset ExtTBR$prob=0
+
+mcmcp ngen=250000 nchains=4 printfreq=1000 samplefreq=100 burnin=200
+
+mcmc
+sump
+sumt
+
+```
+
+```{r}
+tree <- treeio::read.beast(file = "~/Desktop/natural-variation-webtool/manuscript/IAAs.nex.con.tre")
+
+IAA_data <- read.csv("~/Dropbox/natural-variation-webtool/manuscript_IAA_A/Alex_Tree_work/IAAs_table.csv")
+#tip_labels need to be in the first column
+IAA_data %<>% select(name, everything())
+IAA_data$name %<>% as.character()
+IAA_data$AGI %<>% as.character()
+
+IAA_data <- IAA_data[match(IAA_data$AGI, gsub(x = tree@phylo$tip.label, pattern = "\\.1",replacement = "")), ]
+tree@phylo$tip.label <- IAA_data$name
+  IAA_data$AGI
+  IAA_data[match(IAA_data$AGI, gsub(x = tree@phylo$tip.label, pattern = "\\.1",replacement = "")), c("name", "AGI")]
+tree@phylo$tip.label
+
+```
+
+```{r IAA tree, fig.height=5, fig.width=5}
+p <- ggtree(tree, layout = "rectangular", color = "grey") %<+% IAA_data + theme_tree() + scale_x_continuous(expand = c(0, 1.2)) + geom_tiplab(hjust = -0.25) + 
+  geom_tippoint(mapping = aes(size = PiN_PiS, color = PiN_PiS), show.legend = TRUE) + 
+  scale_color_viridis(values = rescale(c(min(IAA_data$PiN_PiS), 1,
+                                        max(IAA_data$PiN_PiS)), 
+                                      to = c(0,1)), guide = "colourbar") + scale_size(guide = "none") + 
+  theme(legend.position = c(.9,.7)) + 
+  geom_nodelab(mapping = aes(label = round(prob,digits = 2)), nudge_x = -.3, nudge_y = .5, size = 3)
+p
+```
 

### 2018-05-22 manuscript/readSelection.py

@@ -1,4 +1,4 @@
-from pymol import cmd
+from pymol import cmd, stored
 
 def readSelection (fileName):
 	'''
@@ -7,17 +7,16 @@ def readSelection (fileName):
 	reads a csv file containing a column of selection names and a column of
 	selection strings and creates selection objects for each
 	'''
-	line = []
 	#read in the selection file
 	with open(fileName) as f:
-		for line in f:
+		lines = f.readlines()
+		for line in lines:
 			sel = line.split(",")
-			print sel[0] + " " + sel[1]
-			#cmd.select(selection[0], selection[1])
+			cmd.select(sel[0], sel[1])
 
 	f.close()
 
-cmd.extend('readSelection',readSelection)
+cmd.extend('readSelection',readSelection);
 
 
 

### 2018-05-23 manuscript/manuscript.Rmd
need to add IAA_annotations.csv to package

@@
                       "IAA_gene_ids.csv", package = "r1001genomes"))
-annotation <- readAnnotationFile(filename = system.file("shiny-app", "gene_family_data",
-                      "IAA_annotations.csv", package = "r1001genomes"), gene_info = gene_info)
+#annotation <- readAnnotationFile(filename = system.file("shiny-app", "gene_family_data",
+#                      "IAA_annotations.csv", package = "r1001genomes"), gene_info = gene_info)
 alignment <- alignCDS(gene_info$transcript_ID)
@@
 AAAlign_no_gaps <- as.matrix(alignment[[2]])[, id]
-write.nexus.data(AAAlign_no_gaps, file = "~/Desktop/natural-variation-webtool/manuscript/IAAs.nex", format = "protein")
+write.nexus.data(AAAlign_no_gaps, file = "manuscript/IAAs.nex", format = "protein")
 ```
@@
 
-execute /Volumes/SSD/Users/clayw/Desktop/natural-variation-webtool/manuscript/IAAs.nex
+execute manuscript/IAAs.nex
 # set IAA33 as the outgroup
@@
 ```{r}
-tree <- treeio::read.beast(file = "~/Desktop/natural-variation-webtool/manuscript/IAAs.nex.con.tre")
+tree <- treeio::read.beast(file = "IAAs.nex.con.tre")
 
-IAA_data <- read.csv("~/Dropbox/natural-variation-webtool/manuscript_IAA_A/Alex_Tree_work/IAAs_table.csv")
+IAA_data <- read.csv("IAAs_table.csv")
 #tip_labels need to be in the first column

### 2018-05-24 manuscript/manuscript.Rmd
fixed IAA tree tiplabels

@@
 
-IAA_data <- IAA_data[match(IAA_data$AGI, gsub(x = tree@phylo$tip.label, pattern = "\\.1",replacement = "")), ]
+IAA_data <- IAA_data[match(gsub(x = tree@phylo$tip.label, pattern = "\\.1",replacement = ""), IAA_data$AGI), ] 
 tree@phylo$tip.label <- IAA_data$name
-  IAA_data$AGI
-  IAA_data[match(IAA_data$AGI, gsub(x = tree@phylo$tip.label, pattern = "\\.1",replacement = "")), c("name", "AGI")]
-tree@phylo$tip.label
+ggtree(tree, layout = "rectangular", color = "grey") %<+% IAA_data + theme_tree() + scale_x_continuous(expand = c(0, 1.2)) + geom_tiplab(mapping = aes(color = Class), hjust = -0.25) + 
+  theme(legend.position = c(.9,.7)) + 
+  geom_nodelab(mapping = aes(label = round(prob,digits = 2)), nudge_x = -.3, nudge_y = .5, size = 3)
 

### 2018-05-24 manuscript/manuscript.Rmd
added AFB structure code

@@
     toc: yes
+always_allow_html: yes
 fontfamily: Helvetica
@@
 
-```{r setup echo = FALSE, message=FALSE, warning=FALSE}
+```{r setup}
 knitr::opts_chunk$set(warning = FALSE, collapse = TRUE, comment = "#>", tidy = TRUE, cache = TRUE, out.width = "90%", fig.height=4, fig.width=7, out.extra='style="margin: auto; display: block; padding-top: 15px;"')
+
 #devtools::install_github("wrightrc/r1001genomes", ref = "auxin-natural-variation")
@@
 readSelection ~/Desktop/natural-variation-webtool/manuscript/AFB_variant_selections
+
+### Natural variation
+```{r}
+gene_info <- geneInfoFromFile(
+  fname = system.file("shiny-app", "gene_family_data",
+                      "AFB_gene_ids.csv", package = "r1001genomes"))
+annotation <- readAnnotationFile(filename = system.file("shiny-app", "gene_family_data",
+                      "AFB_annotations.csv", package = "r1001genomes"), gene_info = gene_info)
+alignment <- alignCDS(gene_info$transcript_ID)
+# make an alignment data frame
+aln_df <- makeAlnDF(alignment[[2]])
+
+# make a collapsed alignment for translating positions
+collapsed_aln_df <- 
+  aln_df %>% dplyr::select(aln_pos, seq_pos, seq_name) %>%
+  tidyr::spread(key = seq_name, value = seq_pos)
+
+VCF <- readRDS(system.file("shiny-app", "gene_family_data",
+                      "AFBs_VCF.rds", package = "r1001genomes"))
+div_stats <- ldply(VCF, diversityStats, geneInfo=gene_info, .id="transcript_ID")
+```
+
+```{r AFB1 variants for PyMOL}
+coding_vcf <- ldply(.data = VCF, .fun = subset, 
+             !is.na(Transcript_ID) & gt_GT != "0|0")
+aln_df <- addSNPsToAlnDF(aln_df, coding_vcf, seq_name = Transcript_ID, 
+               seq_pos = Codon_Number)
+collapsed_aln_df %<>% left_join(y = subset(aln_df, seq_name == gene_info[gene_info$tair_symbol == "TIR1", "transcript_ID"]))
+
+#' Select variants for importing into PyMOL
+#'
+#' @param gene_info a gene info data frame
+#' @param aln_df an alignment data frame with attached SNPs
+#' @param structure_ref string, which tair_symbol or transcript_ID (depending on 
+#' \code{name_by_symbols}) corresponds to the protein structure to which 
+#' selections should be mapped
+#' @param effect_type string of the effect type to select from the alignment 
+#' data frame
+#' @param name_by_symbols logical, if true (default) tair_symbols will be used 
+#' to name the selections, otherwise transcript_IDs will be used
+#'
+#' @return
+#' @export
+#'
+#' @examples
+select_variants <- function(gene_info, aln_df, structure_ref, 
+                            effect_type = "missense", name_by_symbols = TRUE){
+  collapsed_aln_df <- aln_df %>% dplyr::select(aln_pos, seq_pos, seq_name) %>%
+    tidyr::spread(key = seq_name, value = seq_pos)
+  if(name_by_symbols)
+    column <- "tair_symbol"
+    else 
+      column <- "transcript_ID"
+
+  selections <- ldply(.data = gene_info[,column], .fun = function(gene){
+    collapsed_aln_df %<>% dplyr::left_join(y = subset(aln_df, seq_name ==  
+                      gene_info[gene_info[, column] == gene, "transcript_ID"]))
+    variants <- collapsed_aln_df[
+      grepl(effect_type, collapsed_aln_df$effects), 
+      gene_info[gene_info[, column] == structure_ref, "transcript_ID"]]
+    c(gene, paste0(na.omit(as.numeric(variants[[1]])), collapse = "+"))
+  })
+  return(selections)
+}
+
+selections <- select_variants(gene_info, aln_df, structure_ref = "TIR1")
+selections$V2 <- paste0("chain B and resi ", selections$V2)
+write.table(selections, file = "AFB_variant_selections", row.names = FALSE, 
+          col.names = FALSE, sep = ",", quote = FALSE)
+```
+
+```{python}
+#
+# -- readSelection.py
+#
+from pymol import cmd, stored
+
+def readSelection (fileName):
+	'''
+	DESCRIPTION
+
+	reads a csv file containing a column of selection names and a column of
+	selection strings and creates selection objects for each
+	'''
+	#read in the selection file
+	with open(fileName) as f:
+		lines = f.readlines()
+		for line in lines:
+			sel = line.split(",")
+			cmd.select(sel[0], sel[1])
+
+	f.close()
+
+cmd.extend('readSelection',readSelection);
+```
+
+```{python eval=FALSE}
+#from within pymol
+cmd.load("Dropbox/180705_NIH_renewal/Aim1-figures/2P1Q.pdb")
+hide everything
+
+run ~/Desktop/natural-variation-webtool/manuscript/readSelection.py
+
+readSelection ~/Desktop/natural-variation-webtool/manuscript/AFB_variant_selections
+
+# Define selections
+# select TIR1variants, (chain B and resi 154+239+546)
+select fbox, (chain B and resi 1-50)
+select LRR, (chain B and resi 50-576)
+select TIR1, (chain B and resi 1-576)
+# select AFB2variants, (chain B and resi 38+174+181+184+209+259+401+496)
+select tir1_1, (chain B and resi 147)
+select TIR1dm, (chain B and resi 170+473)
+select insPcontacts, (chain B and resi 74+78+113+114+344+401+403+436+460+484+485+509)
+select dimerNULL, (chain B and resi 142+143+151+117+147)
+select auxin, resn IAC
+select AUXcontacts, (chain B and resi 78+79+82+378+380+403+404+405+438+439+462+463+464+489)
+select insP, resn IHP
+select IAA7, chain C
+select IAA7contacts, (chain B and resi 81+82+84+139+347+349+350+351+352+380+405+406+407+409+464+465+489)
+select IAA7_N, chain C and resi 1
+select IAA7_C, chain C and resi 13
+select ASK1, chain A
+select Y, chain C and resi 11
+select Ycontacts, chain B and resi 165+166+195
+
+#Define color palette
+#set_color name, [ red-float, green-float, blue-float ]
+set_color blue1, [0.00000000, 0.20392157, 0.6980392]
+set_color blue2, [0.09803922, 0.36078431, 1.0000000]
+set_color blue3, [0.00000000, 0.29019608, 1.0000000]
+set_color teal,  [0.09803922, 1.00000000, 0.6980392]
+set_color LRR,   [1.00000000, 0.71764706, 0.0000000]
+set_color Fbox,  [0.69803922, 0.49411765, 0.0000000]
+set_color TIR1l, [0.67058824, 0.07843137, 0.8000000]
+set_color TIR1f, [0.63529412, 0.34117647, 0.6980392]
+set_color AFB2l, [0.37647059, 0.70196078, 1.0000000]
+set_color AFB2f, [0.07843137, 0.64313725, 0.8000000]
+
+show cartoon, 2P1Q
+
+color lrr, LRR
+
+color TIR1l, TIR1
+show sticks, TIR1
+color TIR1f, AFB1
+show sticks, AFB1
+color AFB2l, AFB2
+show sticks, AFB2
+color AFB2f, AFB3
+show sticks, AFB3
+
+set ray_trace_mode, 3
+
+# used get_view() to return the current view (position) of the molecule 
+# copy the code to be able to set the identical viewpoint later
+
+set_view (\
+     0.044671144,    0.083668359,   -0.995409012,\
+     0.774909019,    0.625905454,    0.087390237,\
+     0.630409002,   -0.775298953,   -0.036878210,\
+    -0.001916260,    0.005584979, -202.498916626,\
+    -5.538011074,   -8.507736206,   39.046234131,\
+  -219.237945557,  624.123413086,  -20.000000000 )
+
+#to render a high definition image use ray
+ray
 

### 2018-05-30 manuscript/manuscript.Rmd
Pymol variant selection for IAAs

@@
 
-```{r AFB1 variants for PyMOL}
+```{r AFB variants for PyMOL}
 coding_vcf <- ldply(.data = VCF, .fun = subset, 
@@
 # copy the code to be able to set the identical viewpoint later
+# to return the current view from Pymol
+# get_view
 
@@
 
+```{r}
+coding_vcf <- ldply(.data = VCF, .fun = subset, 
+             !is.na(Transcript_ID) & gt_GT != "0|0")
+aln_df <- addSNPsToAlnDF(aln_df, coding_vcf, seq_name = Transcript_ID, 
+               seq_pos = Codon_Number)
+collapsed_aln_df %<>% left_join(y = subset(aln_df, seq_name == gene_info[gene_info$tair_symbol == "TIR1", "transcript_ID"]))
+
+selections <- select_variants(gene_info, aln_df, structure_ref = "IAA17")
+selections$V2 <- paste0("chain X and resi ", selections$V2)
+write.table(selections, file = "IAA_variant_selections", row.names = FALSE, 
+          col.names = FALSE, sep = ",", quote = FALSE)
+```
+
 The members of the IAA8/9/27 clade have a unique N-terminal extension with a highly conserved region common to orthologs in several other species [@dreher_arabidopsis_2006].

### 2018-07-03 manuscript/manuscript.Rmd
- figure templates made - trees built - diversity across collapsed alignment for A ARFs - initial revision of text -- still need to cut and consolidate

@@
 ```{r setup}
-knitr::opts_chunk$set(warning = FALSE, collapse = TRUE, comment = "#>", tidy = TRUE, cache = TRUE, out.width = "90%", fig.height=4, fig.width=7, out.extra='style="margin: auto; display: block; padding-top: 15px;"')
+knitr::opts_chunk$set(warning = FALSE, collapse = TRUE, comment = "#>", tidy = TRUE, cache = TRUE, cache.lazy = FALSE, out.width = "90%", fig.height=4, fig.width=7, out.extra='style="margin: auto; display: block; padding-top: 15px;"')
 
 #devtools::install_github("wrightrc/r1001genomes", ref = "auxin-natural-variation")
+library(ape)
 library(r1001genomes)
@@
 library(ggtree)
-library(ape)
 library(reticulate)
@@
 library(viridis)
-use_python(python = "/usr/local/bin/python")
+library(ips)
+library(knitr)
+#use_python(python = "/usr/local/bin/python")
 ```
@@
 Since this initial genome, massive scaling of sequencing technology has allowed a pioneering group to survey the global genomic variation in natural *Arabidopsis thaliana* populations.
-This valuable population genetics resource has led to several associations of genetic loci with phenotypic traits.
-Outside of genome-wide studies this valuable dataset has seen little use.
+This valuable population genetics resource has led to several associations of genetic loci with phenotypic traits and provided insights into how selective pressure has and is influencing the evolution of plant genomes.
+Outside of genome-wide studies however, this valuable dataset has seen little use.
 Natural genomic variation also provides a set of feasible functional variation at the gene, gene family and gene network level that may provide insight into the function and evolution of genes, families and networks.
-Here, we present a web application and R-package through which plant molecular biologists with little to no bioinformatics experience can make use of this rich dataset of genetic variation to formulate hypotheses as to the sequence/function/phenotype relationships determined by the gene, family, or network of their interest.
+Here, we present a web application and R-package through which plant molecular biologists with little-to-no bioinformatics experience can make use of this rich dataset of genetic variation to formulate hypotheses as to the sequence/function/phenotype relationships determined by the gene, family, or network of their interest.
 We demonstrate the utility of this tool through comprehensive analysis and identification of potential functional variation in the nuclear auxin signaling pathway.
@@
 # Results  
-## TIR1/AFB genes
+## *TIR1/AFB* genes
 
-Auxin acts by binding to receptors that in turn target co-corepressors for degradation.
+Auxin acts by binding to receptors that in turn target co-repressors for degradation.
 Auxin receptors (six in the model plant *Arabidopsis thaliana*) evolved through gene duplication and diversification early in the history of vascular plants [@Parry_Complex_2009].
-The rate of co-repressor degradation is determined by the identity of both the receptor and co-repressor [@Havens_Synthetic_2012], and that this rate sets the pace of development [@Guseman_Auxin-induced_2015].
+The rate of co-repressor degradation is determined by the identity of both the receptor and co-repressor [@Havens_Synthetic_2012], and this rate sets the pace of lateral root development [@Guseman_Auxin-induced_2015].
 
-Although it is unknown whether different receptors play different roles at the organismal level, a number of studies have pointed out differences in the ways the six different receptors in A.
-thaliana differ in their expression.
-The expression of AFB5 is induced in the presence of auxin, making this a circadian-regulated gene [@Covington_Circadian_2007].
-Also, AFB3 is more highly expressed in the roots in the presence of nitrate, allowing increased lateral root formation [@Vidal_Nitrate_2010].
+Although it is unclear what unique roles each receptor plays in growth and development, a number of studies have pointed out differences in the ways the six different receptors in A.
+thaliana differ in biochemical function and expression domain[@dharmasiri_plant_2005; @parry_complex_2009; @prigge_arabidopsis_2016].
 Although TIR1/AFBs are expressed ubiquitously in A.
 thaliana tissues, TIR1, AFB2, and AFB3 have been shown to accumulate in the shoot and root meristems and leaf tissues, with slightly different expression patterns for TIR1 [@Parry_Complex_2009].
-AFB1 has a mutation drastically reducing its ability to assemble into an SCF complex, contributing to its increased accumulation, as well as preventing it from degrading Aux/IAAs[@Yu_Untethering_2015].
-Higher order TIR1/AFB mutants show AFB1 to have inconclusive function.
-However, AFB1 is shown to be able to accumulate ubiquitously in all tissues [@Parry_Complex_2009].
-AFB4 and AFB5 have been shown to preferentially and functionally bind the synthetic auxin picloram, while others have a preference for naturally occurring auxin, indole-3-acetic acid [@Prigge_Arabidopsis_2016].
-Interestingly, the strength and rate with which TIR1/AFBs are able to bind and mark Aux/IAAs for degradation are variable [@Havens_Synthetic_2012].
-TIR1 and AFB2 have slightly different speed at which they perform this function, with AFB2 being generally faster.
+Additionally, the expression of AFB5 is stronly circadian-regulated [@Covington_Circadian_2007] and AFB3 is more highly expressed in the roots in the presence of nitrate, allowing increased lateral root formation [@Vidal_Nitrate_2010], suggesting more broad environmental regulation of this gene family may exist.
+  Functionally, all members of this family have been shown to bind auxin and Aux/IAA proteins.
+However, AFB1 has drastically reduced ability to assemble into an SCF complex, due to the substitution E8K in its F-box domain, preventing it from inducing degradation of Aux/IAAs[@Yu_Untethering_2015].
+This lack of complexation may allow observed high ubiquitous AFB1 accumulation [@Parry_Complex_2009].
+Higher order mutants in the family containing *afb1* mutants suggest that *AFB1* has a moderate positive effect on auxin signaling.
+Additionally, AFB4 and AFB5 have been shown to preferentially and functionally bind the synthetic auxin picloram, while other family members preferentially bind indole-3-acetic acid [@Prigge_Arabidopsis_2016].
+Interestingly, the strength and rate with which TIR1/AFBs are able to bind and mark Aux/IAAs for degradation are variable [@calderonvillalobos_combinatorial_2012; @Havens_Synthetic_2012].
+AFB2 induced the degradation of certain Aux/IAA proteins at a faster rate than TIR1, suggesting some functional specificity has arisen since the initial duplication between the *TIR1/AFB1* and *AFB2/AFB3* clades.
 
-```{r}
+```{r TIR1_AFB_div_stats}
 gene_info <- geneInfoFromFile(
@@
 
-Examining the natural sequence variation across the AFB family revealed that *TIR1* and *AFB1* both had very low nonsynonymous diversity, hinting at their likely functional importance.
-*AFB3* and *AFB4* had higher nonsynonymous diversity, while their sisters genes, *AFB2* and *AFB5* were more conserved.
-This matches our current understanding of *AFB3* and *AFB4* as being minor players in the auxin signaling pathway, but brings in to question the inconclusive role of *AFB1* in auxin signaling.
+Examining the natural sequence variation across the *AFB* family revealed that *TIR1* and *AFB1* both had very low nonsynonymous diversity, hinting at their likely functional importance and bringing in to question the inconclusive role of *AFB1* in auxin signaling.
+*AFB3* and *AFB4* had higher nonsynonymous diversity, while their sister genes, *AFB2* and *AFB5* were more conserved.
+This matches our current understanding of *AFB3* as playing a minor role in the auxin signaling pathway and *AFB4* perhaps undergoing pseudogenization.
 
-Although most known functional regions are highly conserved in AFB1, there is a nonsynonymous polymorphism in the TIR1/AFB oligomerization domain, only found in the Can-1 accession.
-Mutations in this domain of TIR1 frequently have a semidominant effect on root phenotypes [@dezfulian_oligomerization_2016; @wright_insights_2017a].
+Although most known functional regions are highly conserved in *AFB1*, there is a nonsynonymous polymorphism  in the oligomerization domain, only found in the Can-0 accession.
+Mutations in this domain of *TIR1* frequently have a semidominant effect on root phenotypes [@dezfulian_oligomerization_2016; @wright_insights_2017a].
 Characterization of this accession may identify potential *AFB1* phenotypes and may also test the hypothesis that AFB1 functions through oligomerization with other TIR1/AFBs.
-
-```{r AFB1 variants for PyMOL, eval=FALSE}
-coding_vcf <- ldply(.data = VCF, .fun = subset, 
-             !is.na(Transcript_ID) & gt_GT != "0|0")
-aln_df <- addSNPsToAlnDF(aln_df, coding_vcf, seq_name = Transcript_ID, 
-               seq_pos = Codon_Number)
-collapsed_aln_df %<>% left_join(y = subset(aln_df, seq_name == gene_info[gene_info$tair_symbol == "TIR1", "transcript_ID"]))
-
-#' Select variants for importing into PyMOL
-#'
-#' @param gene_info a gene info data frame
-#' @param aln_df an alignment data frame with attached SNPs
-#' @param structure_ref string, which tair_symbol or transcript_ID (depending on 
-#' \code{name_by_symbols}) corresponds to the protein structure to which 
-#' selections should be mapped
-#' @param effect_type string of the effect type to select from the alignment 
-#' data frame
-#' @param name_by_symbols logical, if true (default) tair_symbols will be used 
-#' to name the selections, otherwise transcript_IDs will be used
-#'
-#' @return
-#' @export
-#'
-#' @examples
-select_variants <- function(gene_info, aln_df, structure_ref, 
-                            effect_type = "missense", name_by_symbols = TRUE){
-  collapsed_aln_df <- aln_df %>% dplyr::select(aln_pos, seq_pos, seq_name) %>%
-    tidyr::spread(key = seq_name, value = seq_pos)
-  if(name_by_symbols)
-    column <- "tair_symbol"
-    else 
-      column <- "transcript_ID"
-
-  selections <- ldply(.data = gene_info[,column], .fun = function(gene){
-    collapsed_aln_df %<>% dplyr::left_join(y = subset(aln_df, seq_name ==  
-                      gene_info[gene_info[, column] == gene, "transcript_ID"]))
-    variants <- collapsed_aln_df[
-      grepl(effect_type, collapsed_aln_df$effects), 
-      gene_info[gene_info[, column] == structure_ref, "transcript_ID"]]
-    c(gene, paste0(na.omit(as.numeric(variants[[1]])), collapse = "+"))
-  })
-  return(selections)
-}
-
-selections <- select_variants(gene_info, aln_df, structure_ref = "TIR1")
-selections$V2 <- paste0("chain B and resi ", selections$V2)
-write.table(selections, file = "AFB_variant_selections", row.names = FALSE, 
-          col.names = FALSE, sep = ",", quote = FALSE)
-```
-
-```{python, eval=FALSE}
-#
-# -- readSelection.py
-#
-from pymol import cmd, stored
-
-def readSelection (fileName):
-	'''
-	DESCRIPTION
-
-	reads a csv file containing a column of selection names and a column of
-	selection strings and creates selection objects for each
-	'''
-	#read in the selection file
-	with open(fileName) as f:
-		lines = f.readlines()
-		for line in lines:
-			sel = line.split(",")
-			cmd.select(sel[0], sel[1])
-
-	f.close()
-
-cmd.extend('readSelection',readSelection);
-```
-
-```{pymol, eval=FALSE}
-run ~/Desktop/natural-variation-webtool/manuscript/readSelection.py
-
-readSelection ~/Desktop/natural-variation-webtool/manuscript/AFB_variant_selections
-
-### Natural variation
-```{r}
-gene_info <- geneInfoFromFile(
-  fname = system.file("shiny-app", "gene_family_data",
-                      "AFB_gene_ids.csv", package = "r1001genomes"))
-annotation <- readAnnotationFile(filename = system.file("shiny-app", "gene_family_data",
-                      "AFB_annotations.csv", package = "r1001genomes"), gene_info = gene_info)
-alignment <- alignCDS(gene_info$transcript_ID)
-# make an alignment data frame
-aln_df <- makeAlnDF(alignment[[2]])
-
-# make a collapsed alignment for translating positions
-collapsed_aln_df <- 
-  aln_df %>% dplyr::select(aln_pos, seq_pos, seq_name) %>%
-  tidyr::spread(key = seq_name, value = seq_pos)
-
-VCF <- readRDS(system.file("shiny-app", "gene_family_data",
-                      "AFBs_VCF.rds", package = "r1001genomes"))
-div_stats <- ldply(VCF, diversityStats, geneInfo=gene_info, .id="transcript_ID")
-```
 
@@
 
-```{python}
+```{python readSelection, eval=FALSE}
 #
@@
 
-```{python eval=FALSE}
+```{PyMol, AFB_structure, eval=FALSE}
 #from within pymol
@@
 annotation <- addAlnPosToAnno(anno_df = annotation, aln_df)
-annotation
+
 # translate positions from TIR1 to AFB1
-
 AFB1_positions <- annotation$positions
@@
 4), suggesting that this extension does not play an important functional role in these proteins.
+Additionally, two frameshift variants and one stop-gained variant were observed in *AFB4* supporting its pseudogenization.
 
@@
 Yellow squared E’s show Cul1-interacting residues, differing in AFB1.
-```{r TIR1/AFB Alignment}
+```{r TIR1/AFB F-box Alignment}
 aln_df <- makeAlnDF(alignment[[2]])
@@
 
-## Aux/IAA genes
+## *Aux/IAA* genes
 
 The Aux/IAA proteins (IAAs) mediate both auxin perception and transcriptional co-repression of auxin responsive genes.
-Auxin perception is carried out along with TIR1/AFB proteins, which together with IAAs act as a co-receptor to bind and perceive intracellular auxin.
+Auxin perception is carried out along with TIR1/AFB proteins, which together with IAAs act as co-receptors to bind and perceive nuclear auxin.
 Interaction of IAAs with auxin and TIR1/AFBs is conferred by the highly conserved 13 amino acid degron motif [@tan_mechanism_2007].
 Formation of this auxin co-receptor complex triggers ubiquitination and subsequent degradation of IAAs.
-This degradation in turn relieves their repression upon ARF transcription factors.Stabilizing mutations within the degron can lead to dominant phenotypes [@berleth_auxin_2004; @yang_iaa1_2004].
-The *Arabidopsis thaliana* Aux/IAA gene family has 29 members, most of which possess a canonical degron.
+This degradation in turn relieves their repression upon ARF transcription factors.
+Stabilizing mutations within the degron can lead to dominant phenotypes [@berleth_auxin_2004; @yang_iaa1_2004].
+The *Arabidopsis thaliana* Aux/IAA family has 29 members, most of which possess a canonical auxin degron.
 Notable exceptions are the auxin-insensitive IAA20 and IAA30 that lack a degron sequence entirely, while the closely related IAA31 retains a semi-functional degron sequence with a very slow degradation rate [@dreher_arabidopsis_2006; @havens_synthetic_2012].
-Additionally, IAA degradation rate is influenced by rate motifs found outside the degron, such as the KR motif [@moss_rate_2015; @dreher_arabidopsis_2006].
+Additionally, IAA degradation rate is influenced by rate motifs found outside the degron, such as the KR motif [@dreher_arabidopsis_2006; @moss_rate_2015].
 IAA-mediated repression of ARFs depends upon both an EAR domain, N-terminal of the degron, that recruits co-repressors of the TOPLESS (TPL) family, and a Phox and Bem1 (PB1) domain that enables interaction via hetero-oligomerization with ARF PB1 domains [@nanao_structural_2014; @kagale_ear_2011].
 
-```{r, fig.height=11}
+```{r IAA diversity stats, fig.height=11}
 gene_info <- geneInfoFromFile(
@@
 #                      "IAA_annotations.csv", package = "r1001genomes"), gene_info = gene_info)
-alignment <- alignCDS(gene_info$transcript_ID)
-# make an alignment data frame
-aln_df <- makeAlnDF(alignment[[2]])
-
-# make a collapsed alignment for translating positions
-collapsed_aln_df <- 
-  aln_df %>% dplyr::select(aln_pos, seq_pos, seq_name) %>%
-  tidyr::spread(key = seq_name, value = seq_pos)
 
@@
                                               pattern = "IAA"))})
-formatRound(datatable(div_stats[order(div_stats$tair_symbol),],
+div_stats <- formatRound(datatable(div_stats[order(div_stats$tair_symbol),],
                   colnames = c("transcript",
@@
                 columns = 2:7, digits = 6)
+
+kable(div_stats$x$data)
+#saveWidget(div_stats, file = "IAA_div_stats.html", )
+
 ```
 
-Within the available natural sequence variation, we found that for the canonical Aux/IAAs, no missense mutations were present in the KR motif or residues responsible for PB1-PB1 domain interactions (Supp.
+Within the available natural sequence variation, we found several missense polymorphisms affecting the EAR domain or degron of several canonical IAAs.
+However, no missense polymorphisms were present in the KR motif or residues critical for PB1-PB1 domain interactions (Supp.
 Fig.
 Alignment of canonical IAAs).
-The majority of nonsynonymous mutations observed were in the variable regions between the conserved domains.
-However, several missense polymorphisms of interest were identified in EAR domains and degrons of several canonical IAAs.
+The majority of missense mutations observed were in the variable regions between the conserved domains.
 
@@
 
-Previous research has ~~pointed to~~ **found evidence of** both broad genetic redundancy and specificity within closely related pairs or groups of IAA proteins [@overvoorde_functional_2005].
-For example, the iaa8-1, iaa9-1 double mutant and the iaa5-1, iaa6-1, iaa19-1 triple mutant have wild-type phenotypes [@overvoorde_functional_2005], yet the IAA6/19 sister pair has significant differences in expression patterns, protein abundances and functions, suggesting they have undergone functional specialization since their divergence [@winkler_variation_2017].
+Previous research has found evidence of both broad genetic redundancy and specificity within closely related pairs or groups of IAA proteins [@overvoorde_functional_2005].
+For example, the *iaa8-1 iaa9-1* double mutant and the *iaa5-1 iaa6-1 iaa19-1* triple mutant have wild-type phenotypes [@overvoorde_functional_2005], yet the *IAA6/19* sister pair has significant differences in expression patterns, protein abundances and functions suggesting they have undergone functional specialization since their divergence [@winkler_variation_2017].
 
-In our analysis of the full IAA gene family, one member of most IAA sister pairs had high nonsynonymous diversity while the other sister was more conserved (Figure IAA phylogeny heat map).
+In our analysis of the *IAA* gene family, we frequently observed that one member of most *IAA* sister pairs had high nonsynonymous diversity while the other sister was more conserved (Figure IAA phylogeny heat map).
 In an interspecific comparison between *A.
 thaliana* and *A.
-lyrata*, IAA19 is more conserved, while IAA6 has regions of increased sequence divergence upstream of the degron and in the N-terminal half of the PB1 domain [@winkler_variation_2017].
-Consistent with these results, our analysis of intraspecific natural variation revealed higher conservation for IAA19 (${\pi_N/\pi_S}$ = 0.55) compared to IAA6 (${\pi_N/\pi_S}$ = 2.3) (Figure IAA phylogeny heat map), and also detected high diversity within the same regions of IAA6 as seen in Winkler et al.
+lyrata*, *IAA19* is more conserved, while *IAA6* has regions of increased sequence divergence upstream of the degron and in the N-terminal half of the PB1 domain [@winkler_variation_2017].
+Consistent with these results, our analysis of intraspecific natural variation revealed higher conservation for *IAA19* (${\pi_N/\pi_S}$ = 0.55) compared to *IAA6* (${\pi_N/\pi_S}$ = 2.3) (Figure IAA phylogeny heat map), and also detected high diversity within the same regions of *IAA6* as seen in Winkler et al.
 (Supp Fig IAA6 diversity plot).
-```{r eval=FALSE}
-write.nexus.data(as.matrix(alignment[[2]]), file = "IAAs.nex", format = "protein")
+```{r IAA alignment}
+alignment <- alignCDS(gene_info$transcript_ID)
+# make an alignment data frame
+aln_df <- makeAlnDF(alignment[[2]])
+
+# make a collapsed alignment for translating positions
+collapsed_aln_df <- 
+  aln_df %>% dplyr::select(aln_pos, seq_pos, seq_name) %>%
+  tidyr::spread(key = seq_name, value = seq_pos)
+```
+
+```{r Mask_IAA_alignment, eval=FALSE}
 aliscore(as.AAbin(alignment[[2]]), path = "/Volumes/SSD/Applications/ALISCORE_v2.0/Aliscore_v.2.0/")
@@
 ```
-```{bash eval=FALSE}
-mb
+```{bash IAA_phylogeny_construction, eval=FALSE}
+~/Desktop/natural-variation-webtool/manuscript
 
-execute manuscript/IAAs.nex
+/Applications/MrBayes/mb
+
+execute IAAs.nex
 # set IAA33 as the outgroup
@@
 
-```{r}
+```{r Load IAA phylogeny and data}
 tree <- treeio::read.beast(file = "IAAs.nex.con.tre")
@@
 tree@phylo$tip.label <- IAA_data$name
-ggtree(tree, layout = "rectangular", color = "grey") %<+% IAA_data + theme_tree() + scale_x_continuous(expand = c(0, 1.2)) + geom_tiplab(mapping = aes(color = Class), hjust = -0.25) + 
+```
+*I think IAA20, 30, and 15 should also be in class C. So there are really two large clades A and B and then there are truncations/loss of function/dominant variants scattered throughout, Class C.*
+
+```{r IAA_tree, fig.height=5, fig.width=5, fig.cap="(ref:IAA tree)"}
+p <- ggtree(tree, layout = "rectangular", color = "grey") %<+% IAA_data + theme_tree() + scale_x_continuous(expand = c(0, 1.2)) + 
+  geom_tippoint(mapping = aes(size = 4*PiN_PiS, fill = PiN_PiS), shape = 21, color = "white", show.legend = TRUE) + 
+  scale_fill_viridis(values = rescale(c(min(IAA_data$PiN_PiS), 1,
+                                        max(IAA_data$PiN_PiS)), 
+                                      to = c(0,1)), guide = "colourbar") + scale_size(guide = "none") + 
   theme(legend.position = c(.9,.7)) + 
-  geom_nodelab(mapping = aes(label = round(prob,digits = 2)), nudge_x = -.3, nudge_y = .5, size = 3)
+  geom_nodelab(mapping = aes(label = round(prob,digits = 2)), nudge_x = -.3, nudge_y = .5, size = 3) + 
+  geom_tiplab(mapping = aes(color = Class), hjust = -0.25) + 
+  scale_color_viridis(discrete = TRUE)
+p
+```
+
+(ref:IAA tree) IAA protein sequence tree mapped with ${\pi_N/\pi_S}$
+
+Within the Group B *IAA* genes, there are four sister pairs---*IAA10*/*IAA11*, *IAA12*/*IAA13*, *IAA18*/*IAA26*, and *IAA20*/*IAA30* \@ref(IAA tree) [@remington_contrasting_2004].
+Interestingly, among these sister pairs only *IAA12* and *IAA13* have a noticeable difference in nonsynonymous diversity.
+The *Arabidopsis thaliana* EFP browser shows that **IAA10** and *IAA11* have almost identical expression patterns, with the exception of *IAA10* being more highly expressed in the developing seed, and *IAA11* being more highly expressed in root tissue during lateral root development **need to cite original data source here**.
+In our analysis, *IAA10* and *IAA11* both showed functional conservation (${\pi_N/\pi_S}$ of 0.80 and 0.67 respectively) in support of *IAA10* and *IAA11* playing redundant roles.
+Accessions and associated variants identified here provide a potentially expedited route to elucidating the extent of redundancy between family members such as *IAA10* and *IAA11*.
+
+### New Loss-of-Function/Gain-of-Function alleles
+Accessions with functional variants can be potentially used to investigate the extent of redundancy/specialization within sister pairs.
+For example, three accessions have a missense variant affecting the EAR domain of IAA10 (HR-5, UKSE06-118 and UKSW06-240, all from southern UK).
+This variant (L49S) is in the second leucine of the LxLxL motif and likely reduces the ability of IAA10 to recruit the TOPLESS co-repressor [@lee_biological_2016].
+The nonsense variant (IAA10-Y219\*) resulting in truncated PB1 domain may be more indicative of *IAA10* null phenotype, as it likely prevents IAA10 from interacting with an ARF.
+This variant is only found in one accession from northeastern Spain (IP-Svi-0).
+These accessions and potential loss-of-function alleles may facilitate exploration of the function of *IAA10*.
+
+Several examples of possible gain-of-function missense variants were found in the degron of *IAA2*, *IAA3*, *IAA6*, *IAA7*, *IAA8*, *IAA13*, *IAA15*, *IAA18*, *IAA19*, *IAA27*, *IAA28* and *IAA29*.
+Most degron variants were very minor alleles in the population, only represented in one or two accessions.
+Interestingly, degron missense variants occured more frequently in *IAA* genes that have a closely related sister, which may indicate redundant function within the sister-pairs.
+*I think this is an interesting finding, but I'm not sure about this hypothesis of redundant function.
+If pairs are acting redundantly and we assume the degron variants are dominant, we would expect the auxin signal they transmit (their function) to be lower, in sum, as a pair.
+So I would actually hypothesize that there has been some divergence in function as a pair.
+*
+```{r Correlation between Distance to MRCA and prevalence of degron variants}
 
@@
 
-```{r IAA tree, fig.height=5, fig.width=5}
-p <- ggtree(tree, layout = "rectangular", color = "grey") %<+% IAA_data + theme_tree() + scale_x_continuous(expand = c(0, 1.2)) + geom_tiplab(hjust = -0.25) + 
-  geom_tippoint(mapping = aes(size = PiN_PiS, color = PiN_PiS), show.legend = TRUE) + 
-  scale_color_viridis(values = rescale(c(min(IAA_data$PiN_PiS), 1,
-                                        max(IAA_data$PiN_PiS)), 
-                                      to = c(0,1)), guide = "colourbar") + scale_size(guide = "none") + 
-  theme(legend.position = c(.9,.7)) + 
-  geom_nodelab(mapping = aes(label = round(prob,digits = 2)), nudge_x = -.3, nudge_y = .5, size = 3) 
-p
+```{r Correlation between Pi_N/Pi_S and k_5}
+
 ```
 
-Within the Group B IAA genes, there are four sister pairs---IAA10/IAA11, IAA12/IAA13, IAA18/IAA26, and IAA20/IAA30 (Fig X) [@remington_contrasting_2004].
-Interestingly, among these sister pairs only IAA12 and 13 have a noticeable difference in nonsynonymous diversity.
-The *Arabidopsis thaliana* EFP browser shows that IAA10 and 11 have almost identical expression patterns, with the exception of IAA10 being more highly expressed in the developing seed, and IAA11 being more highly expressed in root tissue during lateral root development.
-In our analysis, IAA10 and IAA11 both showed functional conservation (${\pi_N/\pi_S}$ of 0.80 and 0.67 respectively) in support of IAA10 and IAA11 playing redundant roles.
-Accessions and associated variants identified here provide a potentially expedited route to elucidating the extent of redundancy between family members such as IAA10 and 11.
+*IAA6* has two potentially dominant missense mutations found within the degron (Supp Fig.
+alignment), one previously believed to lead to the *shy1-1* phenotype [@kim_two_1996].
+Interestingly, this variant is widely represented in 31 accessions across Europe (Supplemental Map Fig).
+This finding supports the proposition that iaa6-C78R may not be the mutation causing the *shy1-1D* phenotype but rather represents natural variation in the degron sequence ([note from Mike Prigge in TAIR](https://www.arabidopsis.org/servlets/TairObject?type=notepad&id=11247)).
+These accessions could be used to confirm whether *iaa6-C78R* is responsible for the *shy1-1D* phenotype.
+Additionally, an accession was identified with the IAA6-V77M variant, also in the degron, that may be useful to further explore *iaa6* gain-of-function phenotypes.
+
+### Identifying novel functional sequences in IAAs
+An exciting possible use for the Natural Variation Webtool is to search for conserved, low-diversity sites that may represent possible sites of functional import, such as post-translational modification or protein-protein interaction.
+For example, recent detailed biochemical analysis identified putative lysine ubiquitination sites in members of the *IAA6/19* sister pairs [@winkler_variation_2017]; Supp Fig IAA6/19 alignment).
+These sites are conserved across natural sequence variation, while several other lysine residues vary (Supp Figs alignment and map).
+In IAA6, there are 17 lysines of which ~6 are putative ubiquitination sites~ *I think this should be 5, K3, K32, K33, K91, and K97.
+Missense variants occured in K34Q, K140N *.
+One is proximal to the KR motif, while two others show natural variation.
+In IAA19, there are 15 total lysines, 8 of which are putative ubiquitination sites (K3, K25, K68, L87, K93, K100, K111, and K141).
+Two of these ubiquitination sites are proximal to the KR motif and degron.
+**This sentence is not clear.
+Only 1 of the ubiquitination sites varies?
+or one lysine, I think you mean this only 1 lysine that varies, but it is not ubiquitinated.
+Ubiquitylated Missense K129T** Natural sequence diversity may be used to narrow down a subset of lysines to assess for functional ubiquitination.
 
@@
 
-###Identification of new Loss-of-Function/Gain-of-Function mutations
-Accessions with functional variants can be potentially used to investigate the extent of redundancy/specialization within sister pairs.
-For example, it has been shown that IAA10 has a missense mutation in the EAR domain, which converts the second leucine in the LxLxL motif to a serine (L49S), a mutation likely to reduce the ability of the IAA to recruit the TOPLESS co-repressor [@lee_biological_2016].
-This mutation was found in three accessions from the southern UK.
-A second potentially stronger mutation, represented in one accession from northeastern Spain is a ~~missense~~ **frameshift?** resulting in an early stop codon in the PB1 domain (need Supp.
-Fig?).
-This likely prevents IAA10 from interacting with an ARF.
-~~These alleles would be strong candidates for loss-of-function mutants, allowing researchers to discover the function of IAA10.
-In particular, it would make possible a straightforward path for creating an iaa10/11 double mutant using CRISPR/Cas9 directed mutagenesis, allowing for examination of the extent of redundancy between these family members.~~ **I don't think we can say this.
-There are SALK lines in the 5' UTR and first intron of IAA10 and IAA11.**
-
-Several examples of possible gain-of-function degron missense variants were found in IAA2, IAA3, IAA6, IAA7, IAA8, IAA13, IAA15, IAA18, IAA19, IAA27, IAA28, IAA29.
-Most degron variants were only found in one or two accessions (Supplemental Table or Map?).
-~Interestingly, almost all of the IAAs found to have accessions with degron missense mutations have a closely related sister, which may indicate a redundant function within the sister-pairs.~ **But we know that iaa6 iaa19 iaa5 triple doesn't have a phenotype.** IAA6 has two potentially dominant missense mutations found within the degron (Supp Fig.
-alignment), one previously believed to lead to the shy1-1 phenotype [@kim_two_1996].
-Interestingly, this variant is widely represented in 31 accessions across Europe (Supplemental Map Fig).
-This finding supports the proposition that iaa6[C78R] may not be the mutation causing the shy1-1 phenotype but rather represents natural variation in the degron sequence ([note from Mike Prigge in TAIR](https://www.arabidopsis.org/servlets/TairObject?type=notepad&id=11247)).
-These accessions could be used to confirm whether iaa6[C78R] is responsible for the SHY1-1D phenotype.
-Additionally, an accession was identified with the IAA6-V77M variant, also in the degron, that may be useful to further explore iaa6 gain-of-function phenotypes.
-
-###Identifying novel functional sequences in IAAs
-An exciting possible use for the Natural Variation Webtool is to search for conserved, low-diversity sites that may represent possible sites of post-translational modification or protein-protein interaction.
-For example, recent detailed biochemical analysis identified putative lysine ubiquitination sites in members of the IAA6/19 sister pairs [@winkler_variation_2017]; Supp Fig IAA6/19 alignment).
-These sites are conserved across natural sequence variation, while several other lysine residues vary (Supp Figs alignment and map).
-In IAA6, there are 17 lysines of which 6 are putative ubiquitination sites, one is in the KR motif, while two others show natural variation.
-In IAA19, there are 15 total lysines, with 8 as putative ubiquitination sites, 2 in the KR motif and degron, and 1 that exhibits natural variation.
-**This sentence is not clear.
-Only 1 of the ubiquitination sites varies?** Natural sequence diversity may be used to narrow down a subset of lysines to assess for functional ubiquitination.
-
-```{r}
+```{r IAA variants mapped to IAA17, eval=FALSE}
 coding_vcf <- ldply(.data = VCF, .fun = subset, 
@@
                seq_pos = Codon_Number)
-collapsed_aln_df %<>% left_join(y = subset(aln_df, seq_name == gene_info[gene_info$tair_symbol == "TIR1", "transcript_ID"]))
 
@@
 
-The members of the IAA8/9/27 clade have a unique N-terminal extension with a highly conserved region common to orthologs in several other species [@dreher_arabidopsis_2006].
-IAA8/9 play redundant roles in leaf serration [@koenig_auxin_2009], while IAA8 important in lateral roots [@arase_iaa8_2012].
+The members of the *IAA8/9/27* clade have a unique N-terminal extension with a highly conserved region common to orthologs in several other species [@dreher_arabidopsis_2006].
+*IAA8* and *IAA9* play redundant roles in leaf serration [@koenig_auxin_2009], while *IAA8* also plays an important role in lateral root development [@arase_iaa8_2012].
 IAA8 has been reported to uniquely localize in the cytosol [@arase_iaa8_2012] where it may interact with LSD1 protein [@coll_programmed_2011].
-The N-terminal extension has a conserved region shared between IAA8 and IAA9 with low natural sequence diversity (Supp Fig IAA8/9 alignment) that may represent a site of novel functionality, possibly mediating interaction with LSD1.
-There are a few accessions with missense alleles in this region in both IAA8 and IAA9 that may be valuable in probing the functionality of this region (Supplemental Fig IAA8/9 map).
+The N-terminal extension has low natural sequence diversity in *IAA8* and *IAA9* (Supp Fig IAA8/9 alignment) that may represent a site of novel functionality, possibly mediating interaction with LSD1.
+The few accessions with missense alleles in this region in both *IAA8* and *IAA9* may be valuable in probing the functionality of this region (Supplemental Fig IAA8/9 map).
 
-Members of the IAA17/7/14/16 group have a second EAR motif (LxLxL) with no observed missense mutations in IAA7, 16, or 17 (Supp Fig IAA7/14/16/17 alignment).
-This observation supports the experimental evidence for IAA7 that the second EAR motif is functional: it can interact with TPR1 co-repressor and has a minor repressive role in auxin-related developmental phenotypes [lee_biological_2016].
-For IAA14, we did observe two missense mutations that may be of interest to probe functionality of the IAA14 EAR motif: one in the central lysine (L41F) that is represented in only one accession, and the other in the second “x” (K40N) that is found in 12 accessions around Copenhagen (Supp Fig.
+```{r IAA8_9_27 alignment}
+
+```
+
+```{r IAA8_9 map}
+
+```
+
+Members of the *IAA17/7/14/16* clade have a second EAR motif (LxLxL) with no observed missense variants in *IAA7*, *IAA16*, or *IAA17* (Supp Fig IAA7/14/16/17 alignment).
+This observation supports the experimental evidence for IAA7 that the second EAR motif is functional, *i.e.* it can interact with TPR1 co-repressor and has a minor repressive role in auxin-related developmental phenotypes [lee_biological_2016].
+For *IAA14*, there were two missense variants that may be of interest to probe functionality of this second EAR motif.
+One variant, L41F, in the central lysine is represented in only one accession, and the other in the second “x” (K40N) is found in twelve accessions around Copenhagen (Supp Fig.
 IAA14 map).
 
-## TPL/TPR genes
+```{r IAA7_14_16_17 alignment}
 
-The Auxin signaling pathway utilizes the TOPLESS (TPL) and TOPLESS-related (TPR) family of Gro/TLE/TUP1 type co-repressor proteins to maintain auxin responsive genes in a transcriptionally-repressed state in the absence of auxin (cite).
-In *Arabidopsis thaliana* the TPL/TPR family includes *TPL* and four *TPR1-4*.
-The resulting proteins are comprised of three structural domains: an N-terminal TPL domain and two WD-40 domains.
-TPL/TPR proteins are recruited to the AUX/IAA proteins through interaction with the conserved Ethylene-responsive element binding factor-associated amphiphilic repression (EAR) domain (cite).
+```
+
+```{r IAA14 map}
+
+```
+
+## *TPL/TPR* genes
+
+The Auxin signaling pathway utilizes the TOPLESS (TPL) and TOPLESS-related (TPR) family of Gro/TLE/TUP1 type co-repressor proteins to maintain auxin responsive genes in a transcriptionally-repressed state in the absence of auxin [@szemenyei_topless_2008].
+In *Arabidopsis thaliana* the five member *TPL/TPR* family includes *TPL* and *TPR1-4*.
+The resulting proteins are comprised of three structural domains: an N-terminal TPL domain and two WD-40 domains [@long_topless_2006].
+TPL/TPR proteins are recruited to the AUX/IAA proteins through interaction with the conserved Ethylene-responsive element binding factor-associated amphiphilic repression (EAR) domain [@szemenyei_topless_2008].
 Canonical EAR domains have the amino acid sequence LxLxL, as found in most AUX/IAAs (figure xA) [@overvoorde_functional_2005].
 TPL/TPR co-repressors bind EAR domains via their C-terminal to LisH (CTLH) domains found near their N-termini (citations of pre-structure founding papers/reviews).
-Recent structural analyses of the TPL N-terminal domain have highlighted the precise interaction interface between TPL and AUX/IAA EAR domains, as well as the TPL-TPL dimerization and tetramerization motifs (Avarelilo et al 2017, Ke et al, 2015).
-The residues required for higher-order multimers of TPL tetramers have also been identified (Ma et al, 2017).
+Recent structural analyses of the TPL N-terminal domain have highlighted the precise interaction interface between TPL and AUX/IAA EAR domains, as well as the TPL-TPL dimerization and tetramerization motifs [@martin-arevalillo_structure_2017; @ke_structural_2015a].
+The residues required for higher-order multimers of TPL tetramers have also been identified [@ma_d53_2017].
 Additional interactions with transcriptional regulation and chromatin modifying machinery are likely mediated by two tandem beta propeller domains of TPL/TPRs.
-These domain have not yet been crystallized, however numerous high resolution crystal structures of beta propeller domains allow for structural prediction of the TPL c-terminal beta propeller domains and the locations of likely interaction faces (Figure, X).
+These domain have not yet been crystallized, however numerous high resolution crystal structures of beta propeller domains allow for structural prediction of the TPL C-terminal beta propeller domains and the locations of likely interaction faces (Figure, X).
 
 The TOPLESS co-repressor family generally exhibits a high level of sequence conservation at the amino acid sequence level across resequenced *Arabidopsis thaliana* accessions, with all ${\pi_N/\pi_S}$ values below 1 (Table X).
-The closely related TPL and TPR1 have the highest ${\pi_N/\pi_S}$ values (TPL-0.425, TPR1-0.739), suggesting that these these two related genes tolerate a higher degree of sequence and potentially functional diversity compared to TPR2-4.
+The closely related *TPL* and *TPR1* have the highest ${\pi_N/\pi_S}$ values (TPL-0.425, TPR1-0.739), suggesting that these these two related genes tolerate a higher degree of sequence and potentially functional diversity compared to *TPR2/3/4*.
 The N-terminal TPL domain of the TPL/TPR family is particularly conserved (Figure X).
-All non-synonymous polymorphisms observed in this region are either in the coils between helices or are highly conservative mutations within helices (i.e.
-Valine to Isoleucine), which would be predicted to exhibit little effect on TPL folding and function.
+All nonsynonymous polymorphisms observed in this region are either in the coils between helices or are highly conservative mutations within helices (i.e.
+Valine to Isoleucine), which would be predicted to exhibit little effect on folding and function.
 
-The high degree of conservation in the entire N-terminal domain underscores its importance.
-For example, the initial tpl-1 mutation (N176H) in the ninth helix is a dominant gain-of-function allele (Long et al, 2002, PMID12050130), which is capable of binding wild-type TPL protein and inducing protein aggregation (Ma, 2017, pmid28630893).
-It is therefore understandable that very little diversity is observed in the ninth alpha helix of TPL/TPRs as mutations in this domain could act in a dominant negative fashion, rendering any de novo mutation in this region embryo lethal.
-However, the entire N-terminal domain exhibits very low rates of nucleotide diversity, suggesting that each domain in this region is critical at the population level.
-Indeed, the crystal structure of TPL has revealed that the TPL domain is responsible for the multiple roles of TPL/TPR proteins: The LisH domain and CRA domain coordinate TPL dimerization and tetramerization respectively (26601214, 28698367), the CTLH domain is critical for interaction with DNA-binding transcription factors (26601214), and helix 9 within the CRA domain may be critical for the formation of higher-order TPL complexes (multimers of TPL tetramers, 28630893).
+```{r TPL SNP stats, fig.height=11}
+gene_info <- geneInfoFromFile(
+  fname = system.file("shiny-app", "gene_family_data",
+                      "TPL_gene_ids.csv", package = "r1001genomes"))
+#annotation <- readAnnotationFile(filename = system.file("shiny-app", "gene_family_data",
+#                      "IAA_annotations.csv", package = "r1001genomes"), gene_info = gene_info)
 
-Compared to the TPL domain, both the proline rich and WD-40 domains exhibit a higher rate of missense mutations.
-We utilized the Phyre2.0 modeling software to predict the structure of the TPL/TPR WD-40 repeats in order to map non-synonymous mutations onto a visible protein chassis (Figure X).
+VCF <- readRDS(system.file("shiny-app", "gene_family_data",
+                      "TPL_VCF.rds", package = "r1001genomes"))
+div_stats <- ldply(VCF, diversityStats, geneInfo=gene_info, .id="transcript_ID")
+div_stats$tair_symbol %<>% fct_reorder(.x = , div_stats$tair_symbol, 
+                                       .fun = function(x) 
+                                         {as.integer(str_remove(x, 
+                                              pattern = "IAA"))})
+div_stats <- formatRound(datatable(div_stats[order(div_stats$tair_symbol),],
+                  colnames = c("transcript",
+                               "symbol",
+                               "&pi;<sub>N</sub>",
+                               "&pi;<sub>S</sub>",
+                               "&pi;<sub>N</sub>/&pi;<sub>S</sub>",
+                               "&pi; coding",
+                               "&pi; transcript"),
+                  rownames = FALSE, escape = FALSE,
+                  options = list(paging=FALSE, searching=FALSE)),
+                columns = 2:7, digits = 6)
+
+kable(div_stats$x$data)
+```
+
+The high degree of conservation in the entire N-terminal domain underscores its importance in TPL/TPR function.
+For example, the initial *tpl-1* mutation (N176H) in the ninth helix is a dominant gain-of-function allele [@long_topless_2006], which is capable of binding wild-type TPL protein and inducing protein aggregation [@ma_d53_2017].
+It is therefore understandable that this helix had very low diversity as nonsynonymous variants in this domain could act in a dominant negative fashion.
+Addtionally, the crystal structure of TPL has revealed that the The LisH domain and CRA domain coordinate TPL dimerization and tetramerization respectively (26601214, 28698367), the CTLH domain is critical for interaction with DNA-binding transcription factors (26601214), and helix nine within the CRA domain may be critical for the formation of higher-order TPL complexes (multimers of TPL tetramers, 28630893).
+
+Compared to the TPL domain, both the proline rich and WD-40 domains exhibited a higher frequency of missense variants.
+We utilized Phyre2.0 modeling software to predict the structure of the TPL/TPR WD-40 repeats in order to map non-synonymous mutations to protein structure (Figure X).
 Here we can see both faces of each predicted beta-propeller structure of the WD-40 domains, which are connected by a short peptide linker.
@@
 ~~These observations could be used as the basis for forward mutagenesis strategies to test whether this region of TPL functions similarly to its homolog TLE, which binds its target genes EH1 and the C-terminal WRPW/Y motif (Hairy/Hes/Runx) in the center of the WD-40 domain (pmid 16762837).~~ **Can we say something more general here?
-There is not enough background to support this.**
+There is not enough background to support this.
+What is TLE from?
+What is EH1?
+Is TLE really binding to EH1 DNA?**
 
-Despite the presence of 5 TPL/TPR genes in the genome, which are considered to be redundant (pmid 16763149), very few SNPs are observed that are predicted to result in loss of function for any TPL/TPR gene.
-Only one frameshift variant  was observed, TPR4-665delG, in a single accession (Aiell-4 from Italy).
-This deletion occurs in codon 222, just after the TPL N-terminal domain, and generates an early stop 11 codons downstream from the deletion.
+```{r TPL_TPR alignment}
+
+```
+
+Despite the presence of five *TPL/TPR* genes in the genome, which are considered to be redundant (pmid 16763149), very few observed SNPs are predicted to result in loss of function for any *TPL/TPR* gene.
+Only one frameshift variant was observed, *TPR4-665delG*, in a single accession (Aiell-4 from Italy).
+This deletion occurs in codon 222, just after the TPL N-terminal domain, and generates an early stop eleven codons downstream from the deletion.
 It is unclear whether this N-terminal fragment would be expressed at any appreciable level, as it is possible that the generation of an early stop codon would trigger nonsense-mediated decay.
@@
 Given the observation that the TPL N-terminal domain is sufficient to bind histones (28630893), and repress transcription via synthetic assays in yeast (24979769), it is possible that this represents a semi-functional TPL variant.
-There has been no evidence for prioritization of individual TPL/TPR family members in the auxin response pathway, but experiments modulating the gene copy number of TPR1 revealed the importance of maintaining the proper dosage of TPL/TPR in order to accurately trigger the salicylic acid innate immune response (20647385).
-These results, taken together with the absence of TPL/TPR loss-of-function alleles in the *Arabidopsis thaliana* accessions suggest that perhaps the entire family is being conserved at the nucleotide level in the population to meet a gene dosage requirement, with little opportunity for neo- or sub-functionalization.
+There has been no evidence for prioritization of individual *TPL/TPR* family members in the auxin response pathway, but experiments modulating the gene copy number of *TPR1* revealed the importance of maintaining the proper dosage of *TPL/TPR* in order to accurately trigger the salicylic acid innate immune response (20647385).
+These results, taken together with the absence of *TPL/TPR* loss-of-function alleles in the *Arabidopsis thaliana* accessions suggest that perhaps the entire family is being conserved at the nucleotide level in the population to meet a gene dosage requirement, with little opportunity for neo- or sub-functionalization.
 
@@
 Class B and C ARFs are historically categorized as repressor ARFs, though the mechanism through which they confer repression has not been identified.
-Their middle regions tend to be proline- and serine-rich (citation).
+Their middle regions tend to be proline- and serine-rich (citation, Ulmasov?).
 
 Canonical ARFs are comprised of three major domains.
-Recent crystallization of these domain have informed structure-function analysis of the ARFs (Boer, 2014; Nanao, 2014; Korasick, 2014).
-These domains are conserved throughout land plants.
+Recent crystallization of these domains have informed structure-function analysis of the ARFs (Boer, 2014; Nanao, 2014; Korasick, 2014).
+These domains are conserved throughout land plants (Dolf's new paper).
 ARFs share an N-terminal B3 DNA binding domain.
@@
 The C-terminal domain of the ARF is a protein-protein interaction domain mediating interactions among ARFs, between ARFs and other transcription factors, and between ARFs and the Aux/IAA repressors.
-This interaction domain was recently characterized as a Phox and Bem1 (PB1) domain, which is comprised of a positive and negative face with conserved basic and acidic residues, respectively.
+This interaction domain was recently characterized as a Phox and Bem1 (PB1) domain, which is comprised of a positive and negative face with conserved basic and acidic residues, respectively (Nanao, 2014; Korasick, 2014).
 The dipolar nature of the PB1 domain may mediate multimerization by the pairwise interaction of these faces on different proteins as the ARF7 PB1 domain was crystallized as a multimer [@korasick_].
-However, it is unclear whether this multimerization occurs or plays a developmental significant role in vivo.
+However, it is unclear whether ARF multimerization occurs or plays a developmental significant role in vivo.
 Interfering with ARF dimerization in either the DNA-binding proximal dimerization domain or the PB1 domain decreases the ability of class A ARFs to activate transcription in a heterologous yeast system (Pierre-Jerome, 2016).
@@
 Analysis of class A ARF nonsynonymous diversity suggests that the majority of these ARFs are highly functionally conserved, with ${\pi_N/\pi_S}$ values much lower than 1 with the exception of ARF19, with ${\pi_N/\pi_S}$ value of 1.8.
-Comparing diversity within sister pairs, there is a similar trade-off as seen in most IAA sister pairs, with one sister being highly conserved and the other more divergent.
+Comparing diversity within sister pairs, there is a similar trade-off as seen in most *IAA* sister pairs, with one sister being highly conserved and the other more divergent.
 ARF19 and ARF8 are the more divergent class A ARFs, with ${\pi_N/\pi_S}$ values at least three time those of their sisters, ARF7 and ARF6 respectively.
@@
 
+```{r ARF diversity stats}
+gene_info <- geneInfoFromFile(
+  fname = system.file("shiny-app", "gene_family_data",
+                      "ARF_gene_ids.csv", package = "r1001genomes"))
+#annotation <- readAnnotationFile(filename = system.file("shiny-app", 
+  #"gene_family_data", "ARF_annotations.csv", 
+  #package = "r1001genomes"), gene_info = gene_info)
+
+VCF <- readRDS(system.file("shiny-app", "gene_family_data",
+                      "ARF_VCF.rds", package = "r1001genomes"))
+div_stats <- ldply(VCF, diversityStats, geneInfo=gene_info, .id="transcript_ID")
+div_stats$tair_symbol %<>% fct_reorder(.x =  div_stats$tair_symbol, 
+                                       .fun = function(x) 
+                                         {as.integer(str_remove(x, 
+                                              pattern = "ARF"))})
+div_stats <- formatRound(datatable(div_stats[order(div_stats$tair_symbol),],
+                  colnames = c("transcript",
+                               "symbol",
+                               "&pi;<sub>N</sub>",
+                               "&pi;<sub>S</sub>",
+                               "&pi;<sub>N</sub>/&pi;<sub>S</sub>",
+                               "&pi; coding",
+                               "&pi; transcript"),
+                  rownames = FALSE, escape = FALSE,
+                  options = list(paging=TRUE, searching=FALSE)),
+                columns = 2:7, digits = 6)
+
+kable(div_stats$x$data)
+#saveWidget(div_stats, file = "IAA_div_stats.html", )
+
+```
+
+```{r ARF alignment}
+alignment <- alignCDS(gene_info$transcript_ID)
+# make an alignment data frame
+aln_df <- makeAlnDF(alignment[[2]])
+
+# make a collapsed alignment for translating positions
+collapsed_aln_df <- 
+  aln_df %>% dplyr::select(aln_pos, seq_pos, seq_name) %>%
+  tidyr::spread(key = seq_name, value = seq_pos)
+```
+
+```{r Mask_ARF_alignment, eval=FALSE}
+aliscore(as.AAbin(alignment[[2]]), path = "/Volumes/SSD/Applications/ALISCORE_v2.0/Aliscore_v.2.0/")
+id <- scan(file = "/Volumes/SSD/Applications/ALISCORE_v2.0/Aliscore_v.2.0/input.fas_List_random.txt", sep = " ", quiet = TRUE)
+AAAlign_no_gaps <- as.matrix(alignment[[2]])[, id]
+write.nexus.data(AAAlign_no_gaps, file = "ARFs.nex", format = "protein")
+```
+```{bash ARF_phylogeny_construction, eval=FALSE}
+/Applications/MrBayes/mb
+
+execute ARFs.nex
+# set ARF17 as the outgroup
+outgroup AT1G77850.1
+
+lset nucmodel=protein Rates=invgamma
+# or include Nst=6  ngammacat=4 omegavar=M3 
+prset aamodelpr=fixed(jones) statefreqpr=fixed(empirical) 
+# initial run with aamodelpr=mixed yielded posterior prob of 1 for jones model
+propset ExtTBR$prob=0
+
+mcmcp ngen=250000 nchains=4 printfreq=1000 samplefreq=100 burnin=200
+
+mcmc
+sump
+sumt
+
+```
+
+```{r Load ARF phylogeny and data}
+tree <- treeio::read.beast(file = "ARFs.nex.con.tre")
+
+ARF_data <- read.csv(system.file("shiny-app", "gene_family_data",
+                      "ARF_gene_ids.csv", package = "r1001genomes"))
+ARF_data %<>% left_join(y = div_stats$x$data, by = c("name" = "tair_symbol"))
+#tip_labels need to be in the first column
+ARF_data %<>% select(name, everything())
+ARF_data$tair_locus %<>% as.character()
+
+ARF_data <- ARF_data[match(gsub(x = tree@phylo$tip.label, pattern = "\\.1",replacement = ""), ARF_data$tair_locus), ]
+
+tree@phylo$tip.label <- ARF_data$name
+
+```
+
+```{r Visualize ARF tree, fig.height=5, fig.width=5}
+p <- ggtree(tree, layout = "rectangular", color = "grey") %<+% ARF_data + theme_tree() + scale_x_continuous(expand = c(0, 1.2)) + geom_tiplab(hjust = -0.25) + 
+  geom_tippoint(mapping = aes(size = 4*Pi_NS_Ratio, fill = Pi_NS_Ratio), 
+                shape = 21, color = "white", show.legend = TRUE) + 
+  scale_fill_viridis(values = rescale(c(min(ARF_data$Pi_NS_Ratio), 1,
+                                        max(ARF_data$Pi_NS_Ratio)), 
+                                      to = c(0,1)), guide = "colourbar") + scale_size(guide = "none") + 
+  theme(legend.position = c(.1,.7)) + 
+  geom_nodelab(mapping = aes(label = round(prob,digits = 2)), nudge_x = -.3, nudge_y = .5, size = 3) +
+  geom_tiplab(mapping = aes(color = class), hjust = -0.25) + 
+  scale_color_viridis(discrete = TRUE)
+p
+```
+
+### Class A ARFs 
 For all class A ARFs, the middle region of the protein was the predominant high diversity region.
@@
 
+```{r A ARF Alignment}
+gene_info <- subset(gene_info, tair_symbol %in% c("ARF5", "ARF6", "ARF7", "ARF8", "ARF19"))
+alignment <- alignCDS(gene_info$transcript_ID)
+# make an alignment data frame
+aln_df <- makeAlnDF(alignment[[2]])
+#BrowseSeqs(alignment[[2]])
+aln_df <- makeAlnDF(alignment[[2]])
+aln_df <- left_join(aln_df, dplyr::select(gene_info, "tair_locus",
+                                       "tair_symbol", "transcript_ID"),
+                        by = c("transcript_ID" = "transcript_ID"))
+aln_df$tair_symbol %<>% as.factor()
+aln_df$tair_symbol %<>% fct_reorder(
+    {as.integer(str_remove(aln_df$tair_symbol, 
+                           pattern = "ARF"))}, 
+  .desc = TRUE)
+
+aln_df <- chunkAlnDF(aln_df = aln_df, chunk_width = 100)
+VCF <- ldply(.data = VCF, .fun = subset, gt_GT != "0|0")
+VCF <- getCodingDiv(VCF)
+aln_df <- addSNPsToAlnDF(aln_df, VCF)
+A_ARFs <- subset(aln_df, tair_symbol %in% c("ARF5", "ARF6", "ARF7", "ARF8", "ARF19"))
+p <-ggplot(A_ARFs, aes(x = aln_pos, y = tair_symbol,
+                             group = seq_pos, text = variants))
+p <- p +
+  geom_raster(data = na.omit(A_ARFs), mapping = aes(fill = effects)) +
+  scale_fill_viridis(option = "A", discrete = TRUE) +
+  scale_y_discrete() +
+  # expand increases distance from axis
+  labs(x = "", y = "", fill = "variant") +
+  theme_logo(base_family = "Helvetica") +
+  theme(panel.grid = element_blank(), panel.grid.minor = element_blank()) +
+  theme(strip.background = element_blank(),
+        strip.text.x = element_blank(),
+        legend.box = "vertical")
+p
+#ggsave(p,filename = "A_ARFs_alignment_variants.pdf", width = 12, height = 4)
+```
+
 Class A ARFs all contain the canonical B3 DNA-binding domain, PB1 interaction domain and glutamine-rich middle region (Guilefoyle, 2007).
-Within the middle region, ARF7 had several expansions of polyglutamine sequences compared to the other class A ARFs.
+In the analyzed natural variation, ARF7 had several expansions of polyglutamine sequences in the middle region.
 Polyglutamine regions are known to readily expand and contract throughout evolutionary time due to replication error, and variation in polyglutamine length can be acted on by natural selection and have phenotypic consequences (Press, 2014).
+
 A few missense variants in ARF5 highlighted potential functional variation in the class A ARFs.
-Two highly represented variants resulting in substitutions of small amino acids for the bulky phenylalanine were found on the positive face of the PB1 domain, very close to the conserved lysine residue and may disrupt interactions with other ARF or Aux/IAA PB1 domains.
-Interestingly, these variants were always found together in accessions Tu-B2-3, Ru-2, Haes-1, Rd-0, 627ME-13Y1, Le-0, Bu-0 and Bsch-0.
+A high frequency allele (*What are the variants?*), resulting in substitutions of small amino acids on the positive face of the PB1 domain for bulky phenylalanines, may disrupt interactions with other ARF or Aux/IAA PB1 domains.
+Interestingly, these variants were always found together in accessions Tu-B2-3, Ru-2, Haes-1, Rd-0, 627ME-13Y1, Le-0, Bu-0 and Bsch-0 *Where are these?*.
 Another highly represented polymorphism in ARF5 results in change in the 17th amino acid from a leucine to a methionine.
@@
 ### Class B ARFs
-Andrew
 
-ARF23, which is truncated in its DNA-binding region, has a high ${\pi_N/\pi_S}$ value of 4.1, and has many high-diversity valued nonsense variants, as may be expected of a gene that is pseudogenizing.
+*ARF23*, which has a truncated DNA-binding domain, has a high ${\pi_N/\pi_S}$ value of 4.1, and has many high-diversity nonsense variants, as may be expected of a gene that is pseudogenizing.
 
-ARF3, ARF13, and ARF17 all lack a PB1 C-terminal domain, which suggests they may be unable to interact with Aux/IAA repressors and may act as an auxin-insensitive mechanism of adding competitive inhibition to ARF-mediated transcription.
+ARF3, ARF13 and ARF17 all lack a C-terminal PB1 domain, which suggests they may be unable to interact with Aux/IAA repressors, possibly acting as auxin-insensitive competitive inhibitors of ARF-mediated transcription (Fig ARF3_13_17 alignment).
 These ARFs span the range of ${\pi_N/\pi_S}$ ratios relative to the other ARFs, ranging from 0.85 to 4.1.
-ARF13, with the highest ${\pi_N/\pi_S}$ of 4.1 has many nonsense variants of high diversity.
+ARF17 is more conserved than the activator ARF, ARF19.
+ARF13 has many nonsense variants of high diversity and may be pseudogenizing.
+
+```{r ARF3_13_17 alignment}
+
+```
 
@@
 ### Class C ARFs
-Hardik (All figure cite below are in folder "/Dropbox/natural-variation-webtool/manuscript_ARF_C/FIGURES")
 
-Class C ARFs show low nucleotide diversity scores, all pi_N/pi_S values substantially lower than 1.
+Class C ARFs show low nucleotide diversity scores, all ${\pi_N/\pi_S}$ values substantially lower than 1.
 ARF16 was the most conserved, whereas its clade members (ARF10, ARF17) had scores at least four times higher [@remington_contrasting_2004].
@@
 They are quite conservative.** Of the PB1 domain containing class C ARFs, ARF16 exhibits several missense variants which are sporadically distributed, in contrast to the conserved PB1 domain of ARF10 (Fig.
-Arf16_PB_variants distribution) .
+Arf16_PB_variants distribution).
+
+```{r ARF10_DBD Variants distribution}
+
+```
+
+```{r Arf16_PB_variants distribution}
+
+```
 
@@
 ARF10_16_17_Alignment).
-This is consistent with perturbation studies that has shown drastic developmental phenotype for class C ARFs [@wang_control_2005; @mallory_micrornadirected_2005; @liu_repression_2007].
-For the class B ARFs (2, 3 and 4) tasiRNA binding motif is also highly conserved suggesting strong regulatory constrains at the locus (Fig ARF2_3_4_Alignment).
-DNA binding domain of ARF4 exhibits non-synonymous variation in 4 accession and would be interesting to carry out phenotypic characterize them for developmental growth defects (Fig ARF4_DBD Variants distribution).
+This is consistent with perturbation studies that have shown drastic developmental phenotype for class C ARFs [@wang_control_2005; @mallory_micrornadirected_2005; @liu_repression_2007].
+For the class B ARFs (2, 3 and 4) the tasiRNA binding motif is also highly conserved suggesting strong regulatory constraints on these genes (Fig ARF2_3_4_Alignment).
+The DNA binding domain of ARF4 exhibits non-synonymous variation in 4 accession and would be interesting to carry out phenotypic characterization of these accessions for developmental growth defects (Fig ARF4_DBD Variants map).
+
+```{r ARF10_16_17_Alignment}
+
+```
+
+```{r ARF2_3_4_Alignment}
+
+```
+
+```{r ARF4_DBD Variants map}
+
+```
 

### 2018-07-03 manuscript/manuscript.Rmd
echo = FALSE

@@
 
-```{r setup}
-knitr::opts_chunk$set(warning = FALSE, collapse = TRUE, comment = "#>", tidy = TRUE, cache = TRUE, cache.lazy = FALSE, out.width = "90%", fig.height=4, fig.width=7, out.extra='style="margin: auto; display: block; padding-top: 15px;"')
+```{r setup, message=FALSE, warning=FALSE}
+knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE, collapse = TRUE, comment = "#>", tidy = TRUE, cache = TRUE, cache.lazy = FALSE, out.width = "90%", fig.height=4, fig.width=7, out.extra='style="margin: auto; display: block; padding-top: 15px;"')
 

### 2018-07-10 manuscript/flextable.Rmd
add kable display for div_stats in word documents

--- 2018-05-30 manuscript/manuscript.Rmd
@@
 ---
-title: A comprehensive analysis of natural sequence variation within the <i>Arabidopsis
-  thaliana</i> nuclear auxin signaling pathway using an accessible web application
-author:
-  affiliation: University of Washington
-  name: R. Clay Wright
-date: "`r Sys.Date()`"
+title: "Testing flextable"
+author: "R Clay Wright"
+date: "7/6/2018"
 output:
   word_document:
+    reference_docx: "template.docx"
     toc: yes
@@
     toc: yes
-always_allow_html: yes
-fontfamily: Helvetica
-fontsize: 11pt
-keywords: r markdown
-link-citations: yes
-csl: apa.csl
+  pdf_document:
+    toc: yes
 bibliography:
@@
 - AuxIAA_citations.bib
+#fontfamily: Helvetica
+fontsize: 11pt
+keywords: r markdown
+link-citations: yes
+csl: apa.csl
+always_allow_html: yes
 ---
 
-```{r setup}
-knitr::opts_chunk$set(warning = FALSE, collapse = TRUE, comment = "#>", tidy = TRUE, cache = TRUE, out.width = "90%", fig.height=4, fig.width=7, out.extra='style="margin: auto; display: block; padding-top: 15px;"')
+```{r setup, message=FALSE, warning=FALSE}
+knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE, collapse = TRUE, comment = "#>", tidy = TRUE,  out.width = "90%", fig.height=4, fig.width=7, out.extra='style="margin: auto; display: block; padding-top: 15px;"')
+#cache = TRUE, cache.lazy = FALSE,
 
 #devtools::install_github("wrightrc/r1001genomes", ref = "auxin-natural-variation")
+library(ape)
 library(r1001genomes)
@@
 library(viridis)
-use_python(python = "/usr/local/bin/python")
+library(ips)
+library(knitr)
+
+#use_python(python = "/usr/local/bin/python")
 ```
@@
                       "AFB_gene_ids.csv", package = "r1001genomes"))
-annotation <- readAnnotationFile(filename = system.file("shiny-app", "gene_family_data",
-                      "AFB_annotations.csv", package = "r1001genomes"), gene_info = gene_info)
-alignment <- alignCDS(gene_info$transcript_ID)
-# make an alignment data frame
-aln_df <- makeAlnDF(alignment[[2]])
-
-# make a collapsed alignment for translating positions
-collapsed_aln_df <- 
-  aln_df %>% dplyr::select(aln_pos, seq_pos, seq_name) %>%
-  tidyr::spread(key = seq_name, value = seq_pos)
 
@@
 div_stats <- ldply(VCF, diversityStats, geneInfo=gene_info, .id="transcript_ID")
-formatRound(datatable(div_stats,
+
+#DT table
+DT <- formatRound(datatable(div_stats,
                   colnames = c("transcript",
@@
                 columns = 2:7, digits = 6)
-```
-
-Examining the natural sequence variation across the AFB family revealed that *TIR1* and *AFB1* both had very low nonsynonymous diversity, hinting at their likely functional importance.
-*AFB3* and *AFB4* had higher nonsynonymous diversity, while their sisters genes, *AFB2* and *AFB5* were more conserved.
-This matches our current understanding of *AFB3* and *AFB4* as being minor players in the auxin signaling pathway, but brings in to question the inconclusive role of *AFB1* in auxin signaling.
-
-Although most known functional regions are highly conserved in AFB1, there is a nonsynonymous polymorphism in the TIR1/AFB oligomerization domain, only found in the Can-1 accession.
-Mutations in this domain of TIR1 frequently have a semidominant effect on root phenotypes [@dezfulian_oligomerization_2016; @wright_insights_2017a].
-Characterization of this accession may identify potential *AFB1* phenotypes and may also test the hypothesis that AFB1 functions through oligomerization with other TIR1/AFBs.
-
-```{r AFB1 variants for PyMOL, eval=FALSE}
-coding_vcf <- ldply(.data = VCF, .fun = subset, 
-             !is.na(Transcript_ID) & gt_GT != "0|0")
-aln_df <- addSNPsToAlnDF(aln_df, coding_vcf, seq_name = Transcript_ID, 
-               seq_pos = Codon_Number)
-collapsed_aln_df %<>% left_join(y = subset(aln_df, seq_name == gene_info[gene_info$tair_symbol == "TIR1", "transcript_ID"]))
-
-#' Select variants for importing into PyMOL
-#'
-#' @param gene_info a gene info data frame
-#' @param aln_df an alignment data frame with attached SNPs
-#' @param structure_ref string, which tair_symbol or transcript_ID (depending on 
-#' \code{name_by_symbols}) corresponds to the protein structure to which 
-#' selections should be mapped
-#' @param effect_type string of the effect type to select from the alignment 
-#' data frame
-#' @param name_by_symbols logical, if true (default) tair_symbols will be used 
-#' to name the selections, otherwise transcript_IDs will be used
-#'
-#' @return
-#' @export
-#'
-#' @examples
-select_variants <- function(gene_info, aln_df, structure_ref, 
-                            effect_type = "missense", name_by_symbols = TRUE){
-  collapsed_aln_df <- aln_df %>% dplyr::select(aln_pos, seq_pos, seq_name) %>%
-    tidyr::spread(key = seq_name, value = seq_pos)
-  if(name_by_symbols)
-    column <- "tair_symbol"
-    else 
-      column <- "transcript_ID"
-
-  selections <- ldply(.data = gene_info[,column], .fun = function(gene){
-    collapsed_aln_df %<>% dplyr::left_join(y = subset(aln_df, seq_name ==  
-                      gene_info[gene_info[, column] == gene, "transcript_ID"]))
-    variants <- collapsed_aln_df[
-      grepl(effect_type, collapsed_aln_df$effects), 
-      gene_info[gene_info[, column] == structure_ref, "transcript_ID"]]
-    c(gene, paste0(na.omit(as.numeric(variants[[1]])), collapse = "+"))
-  })
-  return(selections)
-}
-
-selections <- select_variants(gene_info, aln_df, structure_ref = "TIR1")
-selections$V2 <- paste0("chain B and resi ", selections$V2)
-write.table(selections, file = "AFB_variant_selections", row.names = FALSE, 
-          col.names = FALSE, sep = ",", quote = FALSE)
-```
-
-```{python, eval=FALSE}
-#
-# -- readSelection.py
-#
-from pymol import cmd, stored
-
-def readSelection (fileName):
-	'''
-	DESCRIPTION
-
-	reads a csv file containing a column of selection names and a column of
-	selection strings and creates selection objects for each
-	'''
-	#read in the selection file
-	with open(fileName) as f:
-		lines = f.readlines()
-		for line in lines:
-			sel = line.split(",")
-			cmd.select(sel[0], sel[1])
-
-	f.close()
-
-cmd.extend('readSelection',readSelection);
-```
-
-```{pymol, eval=FALSE}
-run ~/Desktop/natural-variation-webtool/manuscript/readSelection.py
-
-readSelection ~/Desktop/natural-variation-webtool/manuscript/AFB_variant_selections
-
-### Natural variation
-```{r}
-gene_info <- geneInfoFromFile(
-  fname = system.file("shiny-app", "gene_family_data",
-                      "AFB_gene_ids.csv", package = "r1001genomes"))
-annotation <- readAnnotationFile(filename = system.file("shiny-app", "gene_family_data",
-                      "AFB_annotations.csv", package = "r1001genomes"), gene_info = gene_info)
-alignment <- alignCDS(gene_info$transcript_ID)
-# make an alignment data frame
-aln_df <- makeAlnDF(alignment[[2]])
-
-# make a collapsed alignment for translating positions
-collapsed_aln_df <- 
-  aln_df %>% dplyr::select(aln_pos, seq_pos, seq_name) %>%
-  tidyr::spread(key = seq_name, value = seq_pos)
-
-VCF <- readRDS(system.file("shiny-app", "gene_family_data",
-                      "AFBs_VCF.rds", package = "r1001genomes"))
-div_stats <- ldply(VCF, diversityStats, geneInfo=gene_info, .id="transcript_ID")
-```
-
-```{r AFB variants for PyMOL}
-coding_vcf <- ldply(.data = VCF, .fun = subset, 
-             !is.na(Transcript_ID) & gt_GT != "0|0")
-aln_df <- addSNPsToAlnDF(aln_df, coding_vcf, seq_name = Transcript_ID, 
-               seq_pos = Codon_Number)
-collapsed_aln_df %<>% left_join(y = subset(aln_df, seq_name == gene_info[gene_info$tair_symbol == "TIR1", "transcript_ID"]))
-
-#' Select variants for importing into PyMOL
-#'
-#' @param gene_info a gene info data frame
-#' @param aln_df an alignment data frame with attached SNPs
-#' @param structure_ref string, which tair_symbol or transcript_ID (depending on 
-#' \code{name_by_symbols}) corresponds to the protein structure to which 
-#' selections should be mapped
-#' @param effect_type string of the effect type to select from the alignment 
-#' data frame
-#' @param name_by_symbols logical, if true (default) tair_symbols will be used 
-#' to name the selections, otherwise transcript_IDs will be used
-#'
-#' @return
-#' @export
-#'
-#' @examples
-select_variants <- function(gene_info, aln_df, structure_ref, 
-                            effect_type = "missense", name_by_symbols = TRUE){
-  collapsed_aln_df <- aln_df %>% dplyr::select(aln_pos, seq_pos, seq_name) %>%
-    tidyr::spread(key = seq_name, value = seq_pos)
-  if(name_by_symbols)
-    column <- "tair_symbol"
-    else 
-      column <- "transcript_ID"
-
-  selections <- ldply(.data = gene_info[,column], .fun = function(gene){
-    collapsed_aln_df %<>% dplyr::left_join(y = subset(aln_df, seq_name ==  
-                      gene_info[gene_info[, column] == gene, "transcript_ID"]))
-    variants <- collapsed_aln_df[
-      grepl(effect_type, collapsed_aln_df$effects), 
-      gene_info[gene_info[, column] == structure_ref, "transcript_ID"]]
-    c(gene, paste0(na.omit(as.numeric(variants[[1]])), collapse = "+"))
-  })
-  return(selections)
-}
-
-selections <- select_variants(gene_info, aln_df, structure_ref = "TIR1")
-selections$V2 <- paste0("chain B and resi ", selections$V2)
-write.table(selections, file = "AFB_variant_selections", row.names = FALSE, 
-          col.names = FALSE, sep = ",", quote = FALSE)
-```
-
-```{python}
-#
-# -- readSelection.py
-#
-from pymol import cmd, stored
-
-def readSelection (fileName):
-	'''
-	DESCRIPTION
-
-	reads a csv file containing a column of selection names and a column of
-	selection strings and creates selection objects for each
-	'''
-	#read in the selection file
-	with open(fileName) as f:
-		lines = f.readlines()
-		for line in lines:
-			sel = line.split(",")
-			cmd.select(sel[0], sel[1])
-
-	f.close()
-
-cmd.extend('readSelection',readSelection);
-```
-
-```{python eval=FALSE}
-#from within pymol
-cmd.load("Dropbox/180705_NIH_renewal/Aim1-figures/2P1Q.pdb")
-hide everything
-
-run ~/Desktop/natural-variation-webtool/manuscript/readSelection.py
-
-readSelection ~/Desktop/natural-variation-webtool/manuscript/AFB_variant_selections
-
-# Define selections
-# select TIR1variants, (chain B and resi 154+239+546)
-select fbox, (chain B and resi 1-50)
-select LRR, (chain B and resi 50-576)
-select TIR1, (chain B and resi 1-576)
-# select AFB2variants, (chain B and resi 38+174+181+184+209+259+401+496)
-select tir1_1, (chain B and resi 147)
-select TIR1dm, (chain B and resi 170+473)
-select insPcontacts, (chain B and resi 74+78+113+114+344+401+403+436+460+484+485+509)
-select dimerNULL, (chain B and resi 142+143+151+117+147)
-select auxin, resn IAC
-select AUXcontacts, (chain B and resi 78+79+82+378+380+403+404+405+438+439+462+463+464+489)
-select insP, resn IHP
-select IAA7, chain C
-select IAA7contacts, (chain B and resi 81+82+84+139+347+349+350+351+352+380+405+406+407+409+464+465+489)
-select IAA7_N, chain C and resi 1
-select IAA7_C, chain C and resi 13
-select ASK1, chain A
-select Y, chain C and resi 11
-select Ycontacts, chain B and resi 165+166+195
-
-#Define color palette
-#set_color name, [ red-float, green-float, blue-float ]
-set_color blue1, [0.00000000, 0.20392157, 0.6980392]
-set_color blue2, [0.09803922, 0.36078431, 1.0000000]
-set_color blue3, [0.00000000, 0.29019608, 1.0000000]
-set_color teal,  [0.09803922, 1.00000000, 0.6980392]
-set_color LRR,   [1.00000000, 0.71764706, 0.0000000]
-set_color Fbox,  [0.69803922, 0.49411765, 0.0000000]
-set_color TIR1l, [0.67058824, 0.07843137, 0.8000000]
-set_color TIR1f, [0.63529412, 0.34117647, 0.6980392]
-set_color AFB2l, [0.37647059, 0.70196078, 1.0000000]
-set_color AFB2f, [0.07843137, 0.64313725, 0.8000000]
-
-show cartoon, 2P1Q
-
-color lrr, LRR
-
-color TIR1l, TIR1
-show sticks, TIR1
-color TIR1f, AFB1
-show sticks, AFB1
-color AFB2l, AFB2
-show sticks, AFB2
-color AFB2f, AFB3
-show sticks, AFB3
-
-set ray_trace_mode, 3
-
-# used get_view() to return the current view (position) of the molecule 
-# copy the code to be able to set the identical viewpoint later
-# to return the current view from Pymol
-# get_view
-
-set_view (\
-     0.044671144,    0.083668359,   -0.995409012,\
-     0.774909019,    0.625905454,    0.087390237,\
-     0.630409002,   -0.775298953,   -0.036878210,\
-    -0.001916260,    0.005584979, -202.498916626,\
-    -5.538011074,   -8.507736206,   39.046234131,\
-  -219.237945557,  624.123413086,  -20.000000000 )
-
-#to render a high definition image use ray
-ray
 
@@
 
-```{r AFB1 diversity plot}
-coding_variants <- getCodingDiv(VCF[[gene_info[gene_info$tair_symbol == "AFB1", "transcript_ID"]]])
-p <- plotCodingDiv(uniqueCodingVars = coding_variants)
-annotation <- addAlnPosToAnno(anno_df = annotation, aln_df)
-annotation
-# translate positions from TIR1 to AFB1
-
-AFB1_positions <- annotation$positions
-AFB1_positions %<>% left_join(y = select(collapsed_aln_df, aln_pos, gene_info[gene_info$tair_symbol == "AFB1", "transcript_ID"]), by = "aln_pos") 
-AFB1_positions$position <- AFB1_positions$AT4G03190.1
-
-p <- append_layers(p,list(
-        geom_rect(data = AFB1_positions,
-                   mapping = aes(xmin = as.integer(position)-1,
-                                 xmax = as.integer(position)+2,
-                                 fill = annotation),
-                  ymin = -Inf, ymax = Inf, inherit.aes = FALSE, alpha = 0.8)),
-      position = "bottom")
-p
-```
-
-The AFB4 and AFB5 receptors have an N-terminal extension prior to the F-box domains.
-This extension had very high nonsynonymous diversity(Fig.
-4), suggesting that this extension does not play an important functional role in these proteins.
-
-Figure 4: Alignment of A.
-thaliana TIR1/AFB F-box domains showing variants.
-Cyan marks the F-box binding domain.
-Yellow, green, and magenta, mark missense and synonymous, synonymous only, and missense only variant locations respectively.
-Yellow squared E’s show Cul1-interacting residues, differing in AFB1.
-```{r TIR1/AFB Alignment}
-aln_df <- makeAlnDF(alignment[[2]])
-aln_df <- left_join(aln_df, dplyr::select(gene_info, "tair_locus",
-                                       "tair_symbol", "transcript_ID"),
-                        by = c("transcript_ID" = "transcript_ID"))
-aln_df$tair_symbol <- fct_rev(fct_relevel(aln_df$tair_symbol, "TIR1"))
-aln_df <- chunkAlnDF(aln_df = aln_df, chunk_width = 55)
-VCF <- ldply(.data = VCF, .fun = subset, gt_GT != "0|0")
-VCF <- getCodingDiv(VCF)
-aln_df <- addSNPsToAlnDF(aln_df, VCF)
-F_box <- subset(aln_df, aln_pos <= 96)
-F_box <- chunkAlnDF(aln_df = F_box, chunk_width = 80)
-p <-ggplot(F_box, aes(x = aln_pos, y = tair_symbol,
-                             group = seq_pos, text = variants))
-p <- p +
-  geom_tile(data = na.omit(F_box), mapping = aes(fill = effects),
-            width = 1, height = 1, alpha = 0.8) +
-  geom_text(aes(label=letter), alpha= 1, family = "Courier") +
-  scale_fill_brewer(type = "qual", palette = 1, direction = -1) +
-  scale_x_continuous(breaks=seq(1,max(F_box$aln_pos), by = 10)) +
-  scale_y_discrete() +
-  # expand increases distance from axis
-  labs(x = "", y = "", fill = "variant") +
-  theme_logo(base_family = "Helvetica") +
-  theme(panel.grid = element_blank(), panel.grid.minor = element_blank()) +
-  facet_wrap(facets = ~chunk, ncol = 1, scales = "free") +
-  theme(strip.background = element_blank(),
-        strip.text.x = element_blank(),
-        legend.box = "vertical")
-p
-```
-
-## Aux/IAA genes
-
-The Aux/IAA proteins (IAAs) mediate both auxin perception and transcriptional co-repression of auxin responsive genes.
-Auxin perception is carried out along with TIR1/AFB proteins, which together with IAAs act as a co-receptor to bind and perceive intracellular auxin.
-Interaction of IAAs with auxin and TIR1/AFBs is conferred by the highly conserved 13 amino acid degron motif [@tan_mechanism_2007].
-Formation of this auxin co-receptor complex triggers ubiquitination and subsequent degradation of IAAs.
-This degradation in turn relieves their repression upon ARF transcription factors.Stabilizing mutations within the degron can lead to dominant phenotypes [@berleth_auxin_2004; @yang_iaa1_2004].
-The *Arabidopsis thaliana* Aux/IAA gene family has 29 members, most of which possess a canonical degron.
-Notable exceptions are the auxin-insensitive IAA20 and IAA30 that lack a degron sequence entirely, while the closely related IAA31 retains a semi-functional degron sequence with a very slow degradation rate [@dreher_arabidopsis_2006; @havens_synthetic_2012].
-Additionally, IAA degradation rate is influenced by rate motifs found outside the degron, such as the KR motif [@moss_rate_2015; @dreher_arabidopsis_2006].
-IAA-mediated repression of ARFs depends upon both an EAR domain, N-terminal of the degron, that recruits co-repressors of the TOPLESS (TPL) family, and a Phox and Bem1 (PB1) domain that enables interaction via hetero-oligomerization with ARF PB1 domains [@nanao_structural_2014; @kagale_ear_2011].
-
-```{r, fig.height=11}
-gene_info <- geneInfoFromFile(
-  fname = system.file("shiny-app", "gene_family_data",
-                      "IAA_gene_ids.csv", package = "r1001genomes"))
-#annotation <- readAnnotationFile(filename = system.file("shiny-app", "gene_family_data",
-#                      "IAA_annotations.csv", package = "r1001genomes"), gene_info = gene_info)
-alignment <- alignCDS(gene_info$transcript_ID)
-# make an alignment data frame
-aln_df <- makeAlnDF(alignment[[2]])
-
-# make a collapsed alignment for translating positions
-collapsed_aln_df <- 
-  aln_df %>% dplyr::select(aln_pos, seq_pos, seq_name) %>%
-  tidyr::spread(key = seq_name, value = seq_pos)
-
-VCF <- readRDS(system.file("shiny-app", "gene_family_data",
-                      "IAAs_VCF.rds", package = "r1001genomes"))
-div_stats <- ldply(VCF, diversityStats, geneInfo=gene_info, .id="transcript_ID")
-div_stats$tair_symbol %<>% fct_reorder(.x = , div_stats$tair_symbol, 
-                                       .fun = function(x) 
-                                         {as.integer(str_remove(x, 
-                                              pattern = "IAA"))})
-formatRound(datatable(div_stats[order(div_stats$tair_symbol),],
-                  colnames = c("transcript",
+```{r using kable }
+dt <- div_stats
+names(dt) <- c("transcript",
                                "symbol",
-                               "&pi;<sub>N</sub>",
-                               "&pi;<sub>S</sub>",
-                               "&pi;<sub>N</sub>/&pi;<sub>S</sub>",
+                               "&pi;~N~",
+                               "&pi;~S~",
+                               "&pi;~N~/&pi;~S~",
                                "&pi; coding",
-                               "&pi; transcript"),
-                  rownames = FALSE, escape = FALSE,
-                  options = list(paging=FALSE, searching=FALSE)),
-                columns = 2:7, digits = 6)
-```
-
-Within the available natural sequence variation, we found that for the canonical Aux/IAAs, no missense mutations were present in the KR motif or residues responsible for PB1-PB1 domain interactions (Supp.
-Fig.
-Alignment of canonical IAAs).
-The majority of nonsynonymous mutations observed were in the variable regions between the conserved domains.
-However, several missense polymorphisms of interest were identified in EAR domains and degrons of several canonical IAAs.
-
-###Sister Pair Diversification
-
-Previous research has ~~pointed to~~ **found evidence of** both broad genetic redundancy and specificity within closely related pairs or groups of IAA proteins [@overvoorde_functional_2005].
-For example, the iaa8-1, iaa9-1 double mutant and the iaa5-1, iaa6-1, iaa19-1 triple mutant have wild-type phenotypes [@overvoorde_functional_2005], yet the IAA6/19 sister pair has significant differences in expression patterns, protein abundances and functions, suggesting they have undergone functional specialization since their divergence [@winkler_variation_2017].
-
-In our analysis of the full IAA gene family, one member of most IAA sister pairs had high nonsynonymous diversity while the other sister was more conserved (Figure IAA phylogeny heat map).
-In an interspecific comparison between *A.
-thaliana* and *A.
-lyrata*, IAA19 is more conserved, while IAA6 has regions of increased sequence divergence upstream of the degron and in the N-terminal half of the PB1 domain [@winkler_variation_2017].
-Consistent with these results, our analysis of intraspecific natural variation revealed higher conservation for IAA19 (${\pi_N/\pi_S}$ = 0.55) compared to IAA6 (${\pi_N/\pi_S}$ = 2.3) (Figure IAA phylogeny heat map), and also detected high diversity within the same regions of IAA6 as seen in Winkler et al.
-(Supp Fig IAA6 diversity plot).
-```{r eval=FALSE}
-write.nexus.data(as.matrix(alignment[[2]]), file = "IAAs.nex", format = "protein")
-aliscore(as.AAbin(alignment[[2]]), path = "/Volumes/SSD/Applications/ALISCORE_v2.0/Aliscore_v.2.0/")
-id <- scan(file = "input.fas_List_random.txt", sep = " ", quiet = TRUE)
-AAAlign_no_gaps <- as.matrix(alignment[[2]])[, id]
-write.nexus.data(AAAlign_no_gaps, file = "manuscript/IAAs.nex", format = "protein")
-```
-```{bash eval=FALSE}
-mb
-
-execute manuscript/IAAs.nex
-# set IAA33 as the outgroup
-outgroup AT5G57420.1
-
-lset nucmodel=protein Rates=invgamma
-# or include Nst=6  ngammacat=4 omegavar=M3 
-prset aamodelpr=fixed(jones) statefreqpr=fixed(empirical) 
-# initial run with aamodelpr=mixed yielded posterior prob of 1 for jones model
-propset ExtTBR$prob=0
-
-mcmcp ngen=250000 nchains=4 printfreq=1000 samplefreq=100 burnin=200
-
-mcmc
-sump
-sumt
+                               "&pi; transcript")
+dt %<>% mutate_if(is.numeric, round, 6)
+dt$symbol %<>% as.ordered()
+dt$symbol %<>% fct_relevel("TIR1")
+dt %<>% arrange(symbol)
+dt %<>% select(symbol, everything())
+names(dt) <- paste0("**", names(dt), "**")
+library(kableExtra)
+kable(dt) 
+# library(pander)
+# panderOptions('table.split.table', Inf)
+# panderOptions('table.alignment.default', function(df)
+#     ifelse(sapply(df, is.numeric), 'right', 'left'))
+# panderOptions('header.style', "atx")
+# panderOptions('keep.line.breaks', TRUE)
+# panderOptions('table.split.cells', Inf)
+#pander(dt)
 
@@
 ```{r}
-tree <- treeio::read.beast(file = "IAAs.nex.con.tre")
+# library(flextable)
+# #flextable
+# pi_format <- function(x){
+#   sprintf("%.6f", x)
+# }
+# 
+# ft <- regulartable(div_stats)
+# #names(div_stats)
+# ft <- set_header_labels(ft, transcript_ID = "transcript",
+#                                tair_symbol = "symbol",
+#                                Pi_non_syn = "&pi;<sub>N</sub>",
+#                                Pi_syn = "&pi;<sub>S</sub>",
+#                                Pi_NS_Ratio = "&pi;<sub>N</sub>/&pi;<sub>S</sub>",
+#                                Pi_coding = "&pi; coding",
+#                                Pi_transcript = "&pi; transcript")
+# ft
+#flextable doesn't work with symbols well 
+#perhaps huxtable will? then huxtable can be converted to flextable
+# library(huxtable)
+# ht <- as_hux(x = div_stats)
+# names(ht) <-  c("transcript",
+#                                "symbol",
+#                                "&pi;<sub>N</sub>",
+#                                "&pi;<sub>S</sub>",
+#                                "&pi;<sub>N</sub>/&pi;<sub>S</sub>",
+#                                "&pi; coding",
+#                                "&pi; transcript")
+# ht <- add_colnames(ht, rowname = NULL)
+# escape_contents(ht)[1,] <- FALSE
+# print_notebook(ht)
+# ft <- as_flextable(ht, colnames_to_header = TRUE)
+# ft
 
-IAA_data <- read.csv("IAAs_table.csv")
-#tip_labels need to be in the first column
-IAA_data %<>% select(name, everything())
-IAA_data$name %<>% as.character()
-IAA_data$AGI %<>% as.character()
-
-IAA_data <- IAA_data[match(gsub(x = tree@phylo$tip.label, pattern = "\\.1",replacement = ""), IAA_data$AGI), ] 
-tree@phylo$tip.label <- IAA_data$name
-ggtree(tree, layout = "rectangular", color = "grey") %<+% IAA_data + theme_tree() + scale_x_continuous(expand = c(0, 1.2)) + geom_tiplab(mapping = aes(color = Class), hjust = -0.25) + 
-  theme(legend.position = c(.9,.7)) + 
-  geom_nodelab(mapping = aes(label = round(prob,digits = 2)), nudge_x = -.3, nudge_y = .5, size = 3)
-
+## perhaps pander is the way to go
+# library(pander)
+# pt <- div_stats
+# names(pt) <- c("transcript",
+#                                "symbol",
+#                                "&pi;<sub>N</sub>",
+#                                "&pi;<sub>S</sub>",
+#                                "&pi;<sub>N</sub>/&pi;<sub>S</sub>",
+#                                "&pi; coding",
+#                                "&pi; transcript")
+# pandoc.table(pt, round = c(0,0,rep(6, 5)))
 ```
-
-```{r IAA tree, fig.height=5, fig.width=5}
-p <- ggtree(tree, layout = "rectangular", color = "grey") %<+% IAA_data + theme_tree() + scale_x_continuous(expand = c(0, 1.2)) + geom_tiplab(hjust = -0.25) + 
-  geom_tippoint(mapping = aes(size = PiN_PiS, color = PiN_PiS), show.legend = TRUE) + 
-  scale_color_viridis(values = rescale(c(min(IAA_data$PiN_PiS), 1,
-                                        max(IAA_data$PiN_PiS)), 
-                                      to = c(0,1)), guide = "colourbar") + scale_size(guide = "none") + 
-  theme(legend.position = c(.9,.7)) + 
-  geom_nodelab(mapping = aes(label = round(prob,digits = 2)), nudge_x = -.3, nudge_y = .5, size = 3) 
-p
-```
-
-Within the Group B IAA genes, there are four sister pairs---IAA10/IAA11, IAA12/IAA13, IAA18/IAA26, and IAA20/IAA30 (Fig X) [@remington_contrasting_2004].
-Interestingly, among these sister pairs only IAA12 and 13 have a noticeable difference in nonsynonymous diversity.
-The *Arabidopsis thaliana* EFP browser shows that IAA10 and 11 have almost identical expression patterns, with the exception of IAA10 being more highly expressed in the developing seed, and IAA11 being more highly expressed in root tissue during lateral root development.
-In our analysis, IAA10 and IAA11 both showed functional conservation (${\pi_N/\pi_S}$ of 0.80 and 0.67 respectively) in support of IAA10 and IAA11 playing redundant roles.
-Accessions and associated variants identified here provide a potentially expedited route to elucidating the extent of redundancy between family members such as IAA10 and 11.
-
-```{r}
-
-```
-
-###Identification of new Loss-of-Function/Gain-of-Function mutations
-Accessions with functional variants can be potentially used to investigate the extent of redundancy/specialization within sister pairs.
-For example, it has been shown that IAA10 has a missense mutation in the EAR domain, which converts the second leucine in the LxLxL motif to a serine (L49S), a mutation likely to reduce the ability of the IAA to recruit the TOPLESS co-repressor [@lee_biological_2016].
-This mutation was found in three accessions from the southern UK.
-A second potentially stronger mutation, represented in one accession from northeastern Spain is a ~~missense~~ **frameshift?** resulting in an early stop codon in the PB1 domain (need Supp.
-Fig?).
-This likely prevents IAA10 from interacting with an ARF.
-~~These alleles would be strong candidates for loss-of-function mutants, allowing researchers to discover the function of IAA10.
-In particular, it would make possible a straightforward path for creating an iaa10/11 double mutant using CRISPR/Cas9 directed mutagenesis, allowing for examination of the extent of redundancy between these family members.~~ **I don't think we can say this.
-There are SALK lines in the 5' UTR and first intron of IAA10 and IAA11.**
-
-Several examples of possible gain-of-function degron missense variants were found in IAA2, IAA3, IAA6, IAA7, IAA8, IAA13, IAA15, IAA18, IAA19, IAA27, IAA28, IAA29.
-Most degron variants were only found in one or two accessions (Supplemental Table or Map?).
-~Interestingly, almost all of the IAAs found to have accessions with degron missense mutations have a closely related sister, which may indicate a redundant function within the sister-pairs.~ **But we know that iaa6 iaa19 iaa5 triple doesn't have a phenotype.** IAA6 has two potentially dominant missense mutations found within the degron (Supp Fig.
-alignment), one previously believed to lead to the shy1-1 phenotype [@kim_two_1996].
-Interestingly, this variant is widely represented in 31 accessions across Europe (Supplemental Map Fig).
-This finding supports the proposition that iaa6[C78R] may not be the mutation causing the shy1-1 phenotype but rather represents natural variation in the degron sequence ([note from Mike Prigge in TAIR](https://www.arabidopsis.org/servlets/TairObject?type=notepad&id=11247)).
-These accessions could be used to confirm whether iaa6[C78R] is responsible for the SHY1-1D phenotype.
-Additionally, an accession was identified with the IAA6-V77M variant, also in the degron, that may be useful to further explore iaa6 gain-of-function phenotypes.
-
-###Identifying novel functional sequences in IAAs
-An exciting possible use for the Natural Variation Webtool is to search for conserved, low-diversity sites that may represent possible sites of post-translational modification or protein-protein interaction.
-For example, recent detailed biochemical analysis identified putative lysine ubiquitination sites in members of the IAA6/19 sister pairs [@winkler_variation_2017]; Supp Fig IAA6/19 alignment).
-These sites are conserved across natural sequence variation, while several other lysine residues vary (Supp Figs alignment and map).
-In IAA6, there are 17 lysines of which 6 are putative ubiquitination sites, one is in the KR motif, while two others show natural variation.
-In IAA19, there are 15 total lysines, with 8 as putative ubiquitination sites, 2 in the KR motif and degron, and 1 that exhibits natural variation.
-**This sentence is not clear.
-Only 1 of the ubiquitination sites varies?** Natural sequence diversity may be used to narrow down a subset of lysines to assess for functional ubiquitination.
-
-```{r}
-coding_vcf <- ldply(.data = VCF, .fun = subset, 
-             !is.na(Transcript_ID) & gt_GT != "0|0")
-aln_df <- addSNPsToAlnDF(aln_df, coding_vcf, seq_name = Transcript_ID, 
-               seq_pos = Codon_Number)
-collapsed_aln_df %<>% left_join(y = subset(aln_df, seq_name == gene_info[gene_info$tair_symbol == "TIR1", "transcript_ID"]))
-
-selections <- select_variants(gene_info, aln_df, structure_ref = "IAA17")
-selections$V2 <- paste0("chain X and resi ", selections$V2)
-write.table(selections, file = "IAA_variant_selections", row.names = FALSE, 
-          col.names = FALSE, sep = ",", quote = FALSE)
-```
-
-The members of the IAA8/9/27 clade have a unique N-terminal extension with a highly conserved region common to orthologs in several other species [@dreher_arabidopsis_2006].
-IAA8/9 play redundant roles in leaf serration [@koenig_auxin_2009], while IAA8 important in lateral roots [@arase_iaa8_2012].
-IAA8 has been reported to uniquely localize in the cytosol [@arase_iaa8_2012] where it may interact with LSD1 protein [@coll_programmed_2011].
-The N-terminal extension has a conserved region shared between IAA8 and IAA9 with low natural sequence diversity (Supp Fig IAA8/9 alignment) that may represent a site of novel functionality, possibly mediating interaction with LSD1.
-There are a few accessions with missense alleles in this region in both IAA8 and IAA9 that may be valuable in probing the functionality of this region (Supplemental Fig IAA8/9 map).
-
-Members of the IAA17/7/14/16 group have a second EAR motif (LxLxL) with no observed missense mutations in IAA7, 16, or 17 (Supp Fig IAA7/14/16/17 alignment).
-This observation supports the experimental evidence for IAA7 that the second EAR motif is functional: it can interact with TPR1 co-repressor and has a minor repressive role in auxin-related developmental phenotypes [lee_biological_2016].
-For IAA14, we did observe two missense mutations that may be of interest to probe functionality of the IAA14 EAR motif: one in the central lysine (L41F) that is represented in only one accession, and the other in the second “x” (K40N) that is found in 12 accessions around Copenhagen (Supp Fig.
-IAA14 map).
-
-## TPL/TPR genes
-
-The Auxin signaling pathway utilizes the TOPLESS (TPL) and TOPLESS-related (TPR) family of Gro/TLE/TUP1 type co-repressor proteins to maintain auxin responsive genes in a transcriptionally-repressed state in the absence of auxin (cite).
-In *Arabidopsis thaliana* the TPL/TPR family includes *TPL* and four *TPR1-4*.
-The resulting proteins are comprised of three structural domains: an N-terminal TPL domain and two WD-40 domains.
-TPL/TPR proteins are recruited to the AUX/IAA proteins through interaction with the conserved Ethylene-responsive element binding factor-associated amphiphilic repression (EAR) domain (cite).
-Canonical EAR domains have the amino acid sequence LxLxL, as found in most AUX/IAAs (figure xA) [@overvoorde_functional_2005].
-TPL/TPR co-repressors bind EAR domains via their C-terminal to LisH (CTLH) domains found near their N-termini (citations of pre-structure founding papers/reviews).
-Recent structural analyses of the TPL N-terminal domain have highlighted the precise interaction interface between TPL and AUX/IAA EAR domains, as well as the TPL-TPL dimerization and tetramerization motifs (Avarelilo et al 2017, Ke et al, 2015).
-The residues required for higher-order multimers of TPL tetramers have also been identified (Ma et al, 2017).
-Additional interactions with transcriptional regulation and chromatin modifying machinery are likely mediated by two tandem beta propeller domains of TPL/TPRs.
-These domain have not yet been crystallized, however numerous high resolution crystal structures of beta propeller domains allow for structural prediction of the TPL c-terminal beta propeller domains and the locations of likely interaction faces (Figure, X).
-
-The TOPLESS co-repressor family generally exhibits a high level of sequence conservation at the amino acid sequence level across resequenced *Arabidopsis thaliana* accessions, with all ${\pi_N/\pi_S}$ values below 1 (Table X).
-The closely related TPL and TPR1 have the highest ${\pi_N/\pi_S}$ values (TPL-0.425, TPR1-0.739), suggesting that these these two related genes tolerate a higher degree of sequence and potentially functional diversity compared to TPR2-4.
-The N-terminal TPL domain of the TPL/TPR family is particularly conserved (Figure X).
-All non-synonymous polymorphisms observed in this region are either in the coils between helices or are highly conservative mutations within helices (i.e.
-Valine to Isoleucine), which would be predicted to exhibit little effect on TPL folding and function.
-
-The high degree of conservation in the entire N-terminal domain underscores its importance.
-For example, the initial tpl-1 mutation (N176H) in the ninth helix is a dominant gain-of-function allele (Long et al, 2002, PMID12050130), which is capable of binding wild-type TPL protein and inducing protein aggregation (Ma, 2017, pmid28630893).
-It is therefore understandable that very little diversity is observed in the ninth alpha helix of TPL/TPRs as mutations in this domain could act in a dominant negative fashion, rendering any de novo mutation in this region embryo lethal.
-However, the entire N-terminal domain exhibits very low rates of nucleotide diversity, suggesting that each domain in this region is critical at the population level.
-Indeed, the crystal structure of TPL has revealed that the TPL domain is responsible for the multiple roles of TPL/TPR proteins: The LisH domain and CRA domain coordinate TPL dimerization and tetramerization respectively (26601214, 28698367), the CTLH domain is critical for interaction with DNA-binding transcription factors (26601214), and helix 9 within the CRA domain may be critical for the formation of higher-order TPL complexes (multimers of TPL tetramers, 28630893).
-
-Compared to the TPL domain, both the proline rich and WD-40 domains exhibit a higher rate of missense mutations.
-We utilized the Phyre2.0 modeling software to predict the structure of the TPL/TPR WD-40 repeats in order to map non-synonymous mutations onto a visible protein chassis (Figure X).
-Here we can see both faces of each predicted beta-propeller structure of the WD-40 domains, which are connected by a short peptide linker.
-This model reveals amino acid conservation in the core and central residues of the WD-40 domains.
-~~These observations could be used as the basis for forward mutagenesis strategies to test whether this region of TPL functions similarly to its homolog TLE, which binds its target genes EH1 and the C-terminal WRPW/Y motif (Hairy/Hes/Runx) in the center of the WD-40 domain (pmid 16762837).~~ **Can we say something more general here?
-There is not enough background to support this.**
-
-Despite the presence of 5 TPL/TPR genes in the genome, which are considered to be redundant (pmid 16763149), very few SNPs are observed that are predicted to result in loss of function for any TPL/TPR gene.
-Only one frameshift variant  was observed, TPR4-665delG, in a single accession (Aiell-4 from Italy).
-This deletion occurs in codon 222, just after the TPL N-terminal domain, and generates an early stop 11 codons downstream from the deletion.
-It is unclear whether this N-terminal fragment would be expressed at any appreciable level, as it is possible that the generation of an early stop codon would trigger nonsense-mediated decay.
-However, if expressed, it would be fascinating to examine whether this naturally occurring N-terminal truncation retains activity as a co-repressor.
-Given the observation that the TPL N-terminal domain is sufficient to bind histones (28630893), and repress transcription via synthetic assays in yeast (24979769), it is possible that this represents a semi-functional TPL variant.
-There has been no evidence for prioritization of individual TPL/TPR family members in the auxin response pathway, but experiments modulating the gene copy number of TPR1 revealed the importance of maintaining the proper dosage of TPL/TPR in order to accurately trigger the salicylic acid innate immune response (20647385).
-These results, taken together with the absence of TPL/TPR loss-of-function alleles in the *Arabidopsis thaliana* accessions suggest that perhaps the entire family is being conserved at the nucleotide level in the population to meet a gene dosage requirement, with little opportunity for neo- or sub-functionalization.
-
-## ARF genes
-
-Auxin response is ultimately mediated by the auxin responsive transcription factors (ARFs).
-There are 23 ARFs in *Arabidopsis thaliana* that are divided into three phylogenetic classes.
-Class A ARFs (ARF5, ARF6, ARF7, ARF8 and ARF19) activate transcription.
-These ARFs have a glutamine-rich region in the middle of the protein that may mediate activation (Guilefoyle, 2007).
-It has recently been shown that the middle region of ARF5 interacts with the SWI/SNF chromatin remodeling ATPases BRAMA and SPLAYED, possibly to reduce nucleosome occupancy and allow for the recruitment of transcription machinery (Wu, 2016).
-Additionally, ARF7 interacts with Mediator subunits, directly tethering transcriptional activation machinery to its binding sites in the chromosome (Ito, 2016).
-Class B and C ARFs are historically categorized as repressor ARFs, though the mechanism through which they confer repression has not been identified.
-Their middle regions tend to be proline- and serine-rich (citation).
-
-Canonical ARFs are comprised of three major domains.
-Recent crystallization of these domain have informed structure-function analysis of the ARFs (Boer, 2014; Nanao, 2014; Korasick, 2014).
-These domains are conserved throughout land plants.
-ARFs share an N-terminal B3 DNA binding domain.
-Flanking this DNA-binding domain is a dimerization domain, which folds up into a single “taco-shaped” domain to allow for dimerization between ARFs.
-There is an auxiliary domain that immediately follows and interacts with the dimerization domain.
-The middle region is the most variable between ARFs, as mentioned above, but is characterized by repetitive units of glutamine (class A), serine, or proline residues (classes B and C).
-
-The C-terminal domain of the ARF is a protein-protein interaction domain mediating interactions among ARFs, between ARFs and other transcription factors, and between ARFs and the Aux/IAA repressors.
-This interaction domain was recently characterized as a Phox and Bem1 (PB1) domain, which is comprised of a positive and negative face with conserved basic and acidic residues, respectively.
-The dipolar nature of the PB1 domain may mediate multimerization by the pairwise interaction of these faces on different proteins as the ARF7 PB1 domain was crystallized as a multimer [@korasick_].
-However, it is unclear whether this multimerization occurs or plays a developmental significant role in vivo.
-Interfering with ARF dimerization in either the DNA-binding proximal dimerization domain or the PB1 domain decreases the ability of class A ARFs to activate transcription in a heterologous yeast system (Pierre-Jerome, 2016).
-
-While domain architecture is broadly conserved among the ARFs, there are exceptional cases.
-Three ARFs do not contain a PB1 domain at all, ARF3, ARF13, and ARF17, and several more have lost the conserved acidic or basic residues in the PB1 domain, suggesting they may be reduced to a single interaction domain.
-Several ARFs additionally have an expanded conserved region within the DNA-binding domain, of unknown function.
-The majority of domain variation among ARFs occurs in the large B-class subfamily.
-The liverwort *Marchantia polymorpha* has a single representative ARF of each class (Flores-Sandoval, 2015).
-The expansion of these classes in flowering plants is the result of both whole genome and tandem duplication events (Remington, 2004).
-The growth of the ARF family may have allowed for the expansion of the quantity and complexity of loci regulated by the ARFs and subsequent expansion in their regulation of developmental processes.
-
-Class A ARFs are the most well-studied ARF subfamily---the five family members all act as transcriptional activators and have well-characterized, distinct developmental targets.
-Overall the diversity of class A ARFs was generally low, especially compared to the class B and C ARFs, suggesting that class A ARFs are central to auxin signal transduction and plant development.
-Analysis of class A ARF nonsynonymous diversity suggests that the majority of these ARFs are highly functionally conserved, with ${\pi_N/\pi_S}$ values much lower than 1 with the exception of ARF19, with ${\pi_N/\pi_S}$ value of 1.8.
-Comparing diversity within sister pairs, there is a similar trade-off as seen in most IAA sister pairs, with one sister being highly conserved and the other more divergent.
-ARF19 and ARF8 are the more divergent class A ARFs, with ${\pi_N/\pi_S}$ values at least three time those of their sisters, ARF7 and ARF6 respectively.
-This may suggest that ARF6 and ARF7 serve more essential purposes in plant development.
-
-For all class A ARFs, the middle region of the protein was the predominant high diversity region.
-The DNA-binding domain had very few, low-diversity missense mutations, as did the PB1 domain.
-Considering the necessity of their conserved functions, the low level of variation in these key DNA and protein-protein interaction domains is expected.
-
-Class A ARFs all contain the canonical B3 DNA-binding domain, PB1 interaction domain and glutamine-rich middle region (Guilefoyle, 2007).
-Within the middle region, ARF7 had several expansions of polyglutamine sequences compared to the other class A ARFs.
-Polyglutamine regions are known to readily expand and contract throughout evolutionary time due to replication error, and variation in polyglutamine length can be acted on by natural selection and have phenotypic consequences (Press, 2014).
-A few missense variants in ARF5 highlighted potential functional variation in the class A ARFs.
-Two highly represented variants resulting in substitutions of small amino acids for the bulky phenylalanine were found on the positive face of the PB1 domain, very close to the conserved lysine residue and may disrupt interactions with other ARF or Aux/IAA PB1 domains.
-Interestingly, these variants were always found together in accessions Tu-B2-3, Ru-2, Haes-1, Rd-0, 627ME-13Y1, Le-0, Bu-0 and Bsch-0.
-Another highly represented polymorphism in ARF5 results in change in the 17th amino acid from a leucine to a methionine.
-This position is in the N-terminal extension of ARF5 which is of unknown function.
-It is possible that this methionine acts as an alternative start site in the accessions.
-
-### Class B ARFs
-Andrew
-
-ARF23, which is truncated in its DNA-binding region, has a high ${\pi_N/\pi_S}$ value of 4.1, and has many high-diversity valued nonsense variants, as may be expected of a gene that is pseudogenizing.
-
-ARF3, ARF13, and ARF17 all lack a PB1 C-terminal domain, which suggests they may be unable to interact with Aux/IAA repressors and may act as an auxin-insensitive mechanism of adding competitive inhibition to ARF-mediated transcription.
-These ARFs span the range of ${\pi_N/\pi_S}$ ratios relative to the other ARFs, ranging from 0.85 to 4.1.
-ARF13, with the highest ${\pi_N/\pi_S}$ of 4.1 has many nonsense variants of high diversity.
-
-ARF14, ARF15, ARF16, ARF20, and ARF21 all do not have one or multiple of the conserved acidic residues in the PB1 domain, and themselves form a subclade of the B ARFs.
-Interestingly, no non-synonymous variation is seen in these amino acid residues of these ARFs, suggesting that the amino acids that are in the regions may still be important for the function of these proteins.
-
-### Class C ARFs
-Hardik (All figure cite below are in folder "/Dropbox/natural-variation-webtool/manuscript_ARF_C/FIGURES")
-
-Class C ARFs show low nucleotide diversity scores, all pi_N/pi_S values substantially lower than 1.
-ARF16 was the most conserved, whereas its clade members (ARF10, ARF17) had scores at least four times higher [@remington_contrasting_2004].
-
-Structurally, all three members of Class C ARFs contain a canonical B3 DNA-binding domain, but only ARF10 and ARF16 contain a PB1 domain.
-The DNA binding domains exhibit overall low diversity and presence of largely synonymous changes for all the ARFs.
-Only the ARF10 DNA binding domain contains missense variants (S147T and A157S) (Fig.
-ARF10_DBD Variants distribution).
-**Any insights into whether these may affect function?
-They are quite conservative.** Of the PB1 domain containing class C ARFs, ARF16 exhibits several missense variants which are sporadically distributed, in contrast to the conserved PB1 domain of ARF10 (Fig.
-Arf16_PB_variants distribution) .
-
-Several members of class B (ARF2, 3 and 4) and C ARFs (ARF10, 16 and 17) are distinguished from other ARFs on the basis of their post transcriptional regulation by trans-acting small interfering RNAs (tasiRNAs) and microRNAs (miRNAs) [@rhoades_prediction_2002; @allen_micrornadirected_2005].
-**(@AMY Arf 6 and 8 targeted by miRNA 167 - check for conservation).** The miRNA target of class C ARFs showed a high degree of conservation (Fig.
-ARF10_16_17_Alignment).
-This is consistent with perturbation studies that has shown drastic developmental phenotype for class C ARFs [@wang_control_2005; @mallory_micrornadirected_2005; @liu_repression_2007].
-For the class B ARFs (2, 3 and 4) tasiRNA binding motif is also highly conserved suggesting strong regulatory constrains at the locus (Fig ARF2_3_4_Alignment).
-DNA binding domain of ARF4 exhibits non-synonymous variation in 4 accession and would be interesting to carry out phenotypic characterize them for developmental growth defects (Fig ARF4_DBD Variants distribution).
-
-#Discussion

### 2018-07-10 manuscript/manuscript.Rmd
add kable display for div_stats in word documents

@@
 date: "`r Sys.Date()`"
+params: 
+  output: word_document
 output:
@@
 library(knitr)
+
 #use_python(python = "/usr/local/bin/python")
@@
 div_stats <- ldply(VCF, diversityStats, geneInfo=gene_info, .id="transcript_ID")
+div_stats$tair_symbol %<>% fct_reorder(.x =  div_stats$tair_symbol, 
+                                       .fun = function(x) 
+                                         {as.integer(str_remove(x, 
+                                              pattern = "AFB|TIR"))})
+div_stats$tair_symbol %<>% fct_relevel("TIR1")
+div_stats %<>% arrange(tair_symbol)
+div_stats %<>% select(tair_symbol, everything())
+
+if (params$output == "html_document") {
   formatRound(datatable(div_stats,
@@
                 columns = 2:7, digits = 6)
+  } else 
+{
+  dt <- div_stats
+  names(dt) <- c("transcript",
+                                 "symbol",
+                                 "&pi;~N~",
+                                 "&pi;~S~",
+                                 "&pi;~N~/&pi;~S~",
+                                 "&pi; coding",
+                                 "&pi; transcript")
+  dt %<>% mutate_if(is.numeric, round, 6)
+  names(dt) <- paste0("**", names(dt), "**")
+  kable(dt) 
+}
 ```
@@
                                               pattern = "IAA"))})
-div_stats <- formatRound(datatable(div_stats[order(div_stats$tair_symbol),],
+div_stats %<>% arrange(tair_symbol)
+div_stats %<>% select(tair_symbol, everything())
+
+if (params$output == "html_document") {
+  formatRound(datatable(div_stats,
                   colnames = c("transcript",
@@
                 columns = 2:7, digits = 6)
-
-kable(div_stats$x$data)
-#saveWidget(div_stats, file = "IAA_div_stats.html", )
-
+  } else 
+{
+  dt <- div_stats
+  names(dt) <- c("transcript",
+                                 "symbol",
+                                 "&pi;~N~",
+                                 "&pi;~S~",
+                                 "&pi;~N~/&pi;~S~",
+                                 "&pi; coding",
+                                 "&pi; transcript")
+  dt %<>% mutate_if(is.numeric, round, 6)
+  names(dt) <- paste0("**", names(dt), "**")
+  kable(dt) 
+}
 ```
@@
                                          {as.integer(str_remove(x, 
-                                              pattern = "IAA"))})
-div_stats <- formatRound(datatable(div_stats[order(div_stats$tair_symbol),],
+                                              pattern = "TPR"))})
+div_stats$tair_symbol %<>% fct_relevel("TPL")
+div_stats %<>% arrange(tair_symbol)
+div_stats %<>% select(tair_symbol, everything())
+
+if (params$output == "html_document") {
+  formatRound(datatable(div_stats,
                   colnames = c("transcript",
@@
                 columns = 2:7, digits = 6)
-
-kable(div_stats$x$data)
+  } else 
+{
+  dt <- div_stats
+  names(dt) <- c("transcript",
+                                 "symbol",
+                                 "&pi;~N~",
+                                 "&pi;~S~",
+                                 "&pi;~N~/&pi;~S~",
+                                 "&pi; coding",
+                                 "&pi; transcript")
+  dt %<>% mutate_if(is.numeric, round, 6)
+  names(dt) <- paste0("**", names(dt), "**")
+  kable(dt) 
+}
 ```
@@
                                               pattern = "ARF"))})
-div_stats <- formatRound(datatable(div_stats[order(div_stats$tair_symbol),],
+div_stats %<>% arrange(tair_symbol)
+div_stats %<>% select(tair_symbol, everything())
+
+if (params$output == "html_document") {
+  formatRound(datatable(div_stats,
                   colnames = c("transcript",
@@
                   rownames = FALSE, escape = FALSE,
-                  options = list(paging=TRUE, searching=FALSE)),
+                  options = list(paging=FALSE, searching=FALSE)),
                 columns = 2:7, digits = 6)
-
-kable(div_stats$x$data)
-#saveWidget(div_stats, file = "IAA_div_stats.html", )
-
+  } else 
+{
+  dt <- div_stats
+  names(dt) <- c("transcript",
+                                 "symbol",
+                                 "&pi;~N~",
+                                 "&pi;~S~",
+                                 "&pi;~N~/&pi;~S~",
+                                 "&pi; coding",
+                                 "&pi; transcript")
+  dt %<>% mutate_if(is.numeric, round, 6)
+  names(dt) <- paste0("**", names(dt), "**")
+  kable(dt) 
+}
 ```
@@
 #Discussion
+
+Analyzing natural diversity within the gene families constituting the auxin nuclear signaling pathway has confirmed existing knowledge, supported existing hypothses and generated new hypotheses.
+This analysis in particular has confirmed the importance of auxin-induced degradation of the Aux/IAA proteins as evidenced by the strong conservation of the *AFB* genes and the canonical degron region of the *Aux/IAA* genes.
+Despite this relative conservation, this analysis has also identified many accessions containing new alleles of interest in the *AFB* and *Aux/IAA* families.
+This analysis has also
+
+Importance of 
+-AFB
+-IAA degron
+-TPL
+-ARF DBD and PB1

### 2018-07-12 manuscript/01-intro.Rmd
-Cleaned up manuscript folder -moved to bookdown format -revised all sections

# Introduction {#intro}

The first genome sequence of *Arabidopsis thaliana* facilitated rapid advancement of plant biology through molecular genetics. Since this initial genome, massive scaling of sequencing technology has allowed a pioneering group to survey the genomic variation in natural *Arabidopsis thaliana* populations. This valuable population genetics resource has led to several associations of genetic loci with phenotypic traits and provided insights into how selective pressure has influenced the evolution of plant genomes. Outside of genome-wide studies however, this valuable resequencing dataset has seen little use.

Natural genomic variation also provides a set of feasible functional variation at the gene, gene family and gene network level. Characterizing this variation can provide insights into the function and evolution of genes, families and networks [@delker_natural_2010, @wright_insights_2017] . Here, we present a web application and R-package for gene, gene family and gene network analysis of this dataset. This application allows plant molecular biologists without bioinformatics experience to use genomic variation to formulate hypotheses as to the sequence/function/phenotype relationships within their gene, family, or network of their interest.

This application allows easy access to 
1. lists of missense polymorphisms to facilitate biochemical assays of variant effects [@starita_variant_2017],
2. family-wise alignments of variants to facilitate de novo functional domain identification [@melamed_combining_2015], 
3. lists of accessions containing missense (or any type of) polymorphisms to facilitate segregation analysis and measurment of the distribution of variant effects on phenotype [@park_distributions_2017], 
4. as well as the study of gene and network evolution [@delker_natural_2010; @kliebenstein_role_2008].

Here, we demonstrate the utility of this tool through comprehensive analysis of genomic variation in the nuclear auxin signaling pathway. We identify potential functional variation within the coding sequence of the genes within this network and formulate hypotheses regarding the functional implications of this variation and the evolution of these gene families. Our analysis provides further confirmation of much of the existing knowledge of these genes demonstrating the validity of this approach. We also identify natural alleles of several genes and accessions bearing these alleles to facilitate future characterization structure/function, genotype/phenotype, and epistatic relationships.

## bookdown example
You can label chapter and section titles using `{#label}` after them, e.g., we can reference Chapter \@ref(intro). If you do not manually label them, there will be automatic labels anyway, e.g., Chapter \@ref(methods).

Figures and tables with captions will be placed in `figure` and `table` environments, respectively.

```{r nice-fig, fig.cap='Here is a nice figure!', out.width='80%', fig.asp=.75, fig.align='center'}
par(mar = c(4, 4, .1, .1))
plot(pressure, type = 'b', pch = 19)
```

Reference a figure by its code chunk label with the `fig:` prefix, e.g., see Figure \@ref(fig:nice-fig). Similarly, you can reference tables generated from `knitr::kable()`, e.g., see Table \@ref(tab:nice-tab).

```{r nice-tab, tidy=FALSE}
knitr::kable(
  head(iris, 20), caption = 'Here is a nice table!',
  booktabs = TRUE
)
```

You can write citations, too. For example, we are using the **bookdown** package [@R-bookdown] in this sample book, which was built on top of R Markdown and **knitr** [@xie2015].

### 2018-07-12 manuscript/02-methods.Rmd
-Cleaned up manuscript folder -moved to bookdown format -revised all sections

--- 2018-07-10 manuscript/flextable.Rmd
@@
----
-title: "Testing flextable"
-author: "R Clay Wright"
-date: "7/6/2018"
-output:
-  word_document:
-    reference_docx: "template.docx"
-    toc: yes
-  html_document:
-    fig_caption: yes
-    toc: yes
-  pdf_document:
-    toc: yes
-bibliography:
-- natural_variation.bib
-- TIR1_AFB.bib
-- ARF_C.bib
-- AuxIAA_citations.bib
-#fontfamily: Helvetica
-fontsize: 11pt
-keywords: r markdown
-link-citations: yes
-csl: apa.csl
-always_allow_html: yes
----
-
-```{r setup, message=FALSE, warning=FALSE}
-knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE, collapse = TRUE, comment = "#>", tidy = TRUE,  out.width = "90%", fig.height=4, fig.width=7, out.extra='style="margin: auto; display: block; padding-top: 15px;"')
-#cache = TRUE, cache.lazy = FALSE,
-
-#devtools::install_github("wrightrc/r1001genomes", ref = "auxin-natural-variation")
-library(ape)
-library(r1001genomes)
-library(DT)
-library(tidyverse)
-library(magrittr)
-library(RColorBrewer)
-library(ggpmisc)
-library(ggthemes)
-library(ggseqlogo)
-library(ggtree)
-library(reticulate)
-library(scales)
-library(viridis)
-library(ips)
-library(knitr)
-
-#use_python(python = "/usr/local/bin/python")
-```
-<!--This is a CSS (Cascading Style Sheet).
- It controls how panels and images are formatted.></!-->
-<style>
-  img {
-    display: block;
-    margin: auto;
-    width: 90%;
-    height: auto;
-  }
-</style>
-
-# List of working titles  
-* A comprehensive analysis of natural sequence variation within the <i>Arabidopsis thaliana</i> nuclear auxin signaling pathway  
-* **Add your suggestions**
-
-# Abstract
-
-* Motivating, high level statement  
-* Statement defining the cutting edge of your research area  
-* A statement of the form: "However, it remains unclear ..." that defines what problem you are  
-addressing.  
-* A statement of the form:"Here, we ..."  that concisely states your contribution.  
-* A series of 3?5 statement about your results.  
-* A concluding statement.
-
-# Intro
-
-The first genome sequence of *Arabidopsis thaliana* facilitated rapid advancement of plant biology through molecular genetics.
-Since this initial genome, massive scaling of sequencing technology has allowed a pioneering group to survey the global genomic variation in natural *Arabidopsis thaliana* populations.
-This valuable population genetics resource has led to several associations of genetic loci with phenotypic traits and provided insights into how selective pressure has and is influencing the evolution of plant genomes.
-Outside of genome-wide studies however, this valuable dataset has seen little use.
-Natural genomic variation also provides a set of feasible functional variation at the gene, gene family and gene network level that may provide insight into the function and evolution of genes, families and networks.
-Here, we present a web application and R-package through which plant molecular biologists with little-to-no bioinformatics experience can make use of this rich dataset of genetic variation to formulate hypotheses as to the sequence/function/phenotype relationships determined by the gene, family, or network of their interest.
-We demonstrate the utility of this tool through comprehensive analysis and identification of potential functional variation in the nuclear auxin signaling pathway.
-
-This application allows easy access to 
-1. lists of missense polymorphisms to facilitate biochemical assays of variant effects [@starita_variant_2017],
-2. family-wise alignments of variants to facilitate de novo functional domain identification [@melamed_combining_2015], 
-3.
-lists of accessions containing missense (or any type of) polymorphisms to facilitate segregation analysis and measurment of the distribution of variant effects on phenotype [@park_distributions_2017],
-4. as well as the study of gene and network evolution [@delker_natural_2010; @kliebenstein_role_2008].
-
-Here, we have utilized this tool to analyze the genetic variation in the nuclear auxin signaling pathway and formulate hypotheses regarding the functional implications of this variation and the evolution of the genes and gene families in this pathway.
-Our analysis also provided further confirmation of much of the existing knowledge of these genes demonstrating the validity of this approach.
-
-# Methods
+# Methods {#methods}
 
@@
 variants at a single nucleotide location can easily be summarized in R.
-
-# Results  
-## *TIR1/AFB* genes
-
-Auxin acts by binding to receptors that in turn target co-repressors for degradation.
-Auxin receptors (six in the model plant *Arabidopsis thaliana*) evolved through gene duplication and diversification early in the history of vascular plants [@Parry_Complex_2009].
-The rate of co-repressor degradation is determined by the identity of both the receptor and co-repressor [@Havens_Synthetic_2012], and this rate sets the pace of lateral root development [@Guseman_Auxin-induced_2015].
-
-Although it is unclear what unique roles each receptor plays in growth and development, a number of studies have pointed out differences in the ways the six different receptors in A.
-thaliana differ in biochemical function and expression domain[@dharmasiri_plant_2005; @parry_complex_2009; @prigge_arabidopsis_2016].
-Although TIR1/AFBs are expressed ubiquitously in A.
-thaliana tissues, TIR1, AFB2, and AFB3 have been shown to accumulate in the shoot and root meristems and leaf tissues, with slightly different expression patterns for TIR1 [@Parry_Complex_2009].
-Additionally, the expression of AFB5 is stronly circadian-regulated [@Covington_Circadian_2007] and AFB3 is more highly expressed in the roots in the presence of nitrate, allowing increased lateral root formation [@Vidal_Nitrate_2010], suggesting more broad environmental regulation of this gene family may exist.
-  Functionally, all members of this family have been shown to bind auxin and Aux/IAA proteins.
-However, AFB1 has drastically reduced ability to assemble into an SCF complex, due to the substitution E8K in its F-box domain, preventing it from inducing degradation of Aux/IAAs[@Yu_Untethering_2015].
-This lack of complexation may allow observed high ubiquitous AFB1 accumulation [@Parry_Complex_2009].
-Higher order mutants in the family containing *afb1* mutants suggest that *AFB1* has a moderate positive effect on auxin signaling.
-Additionally, AFB4 and AFB5 have been shown to preferentially and functionally bind the synthetic auxin picloram, while other family members preferentially bind indole-3-acetic acid [@Prigge_Arabidopsis_2016].
-Interestingly, the strength and rate with which TIR1/AFBs are able to bind and mark Aux/IAAs for degradation are variable [@calderonvillalobos_combinatorial_2012; @Havens_Synthetic_2012].
-AFB2 induced the degradation of certain Aux/IAA proteins at a faster rate than TIR1, suggesting some functional specificity has arisen since the initial duplication between the *TIR1/AFB1* and *AFB2/AFB3* clades.
-
-```{r TIR1_AFB_div_stats}
-gene_info <- geneInfoFromFile(
-  fname = system.file("shiny-app", "gene_family_data",
-                      "AFB_gene_ids.csv", package = "r1001genomes"))
-
-VCF <- readRDS(system.file("shiny-app", "gene_family_data",
-                      "AFBs_VCF.rds", package = "r1001genomes"))
-div_stats <- ldply(VCF, diversityStats, geneInfo=gene_info, .id="transcript_ID")
-
-#DT table
-DT <- formatRound(datatable(div_stats,
-                  colnames = c("transcript",
-                               "symbol",
-                               "&pi;<sub>N</sub>",
-                               "&pi;<sub>S</sub>",
-                               "&pi;<sub>N</sub>/&pi;<sub>S</sub>",
-                               "&pi; coding",
-                               "&pi; transcript"),
-                  rownames = FALSE, escape = FALSE,
-                  options = list(paging=FALSE, searching=FALSE)),
-                columns = 2:7, digits = 6)
-
-```
-
-```{r using kable }
-dt <- div_stats
-names(dt) <- c("transcript",
-                               "symbol",
-                               "&pi;~N~",
-                               "&pi;~S~",
-                               "&pi;~N~/&pi;~S~",
-                               "&pi; coding",
-                               "&pi; transcript")
-dt %<>% mutate_if(is.numeric, round, 6)
-dt$symbol %<>% as.ordered()
-dt$symbol %<>% fct_relevel("TIR1")
-dt %<>% arrange(symbol)
-dt %<>% select(symbol, everything())
-names(dt) <- paste0("**", names(dt), "**")
-library(kableExtra)
-kable(dt) 
-# library(pander)
-# panderOptions('table.split.table', Inf)
-# panderOptions('table.alignment.default', function(df)
-#     ifelse(sapply(df, is.numeric), 'right', 'left'))
-# panderOptions('header.style', "atx")
-# panderOptions('keep.line.breaks', TRUE)
-# panderOptions('table.split.cells', Inf)
-#pander(dt)
-
-```
-
-```{r}
-# library(flextable)
-# #flextable
-# pi_format <- function(x){
-#   sprintf("%.6f", x)
-# }
-# 
-# ft <- regulartable(div_stats)
-# #names(div_stats)
-# ft <- set_header_labels(ft, transcript_ID = "transcript",
-#                                tair_symbol = "symbol",
-#                                Pi_non_syn = "&pi;<sub>N</sub>",
-#                                Pi_syn = "&pi;<sub>S</sub>",
-#                                Pi_NS_Ratio = "&pi;<sub>N</sub>/&pi;<sub>S</sub>",
-#                                Pi_coding = "&pi; coding",
-#                                Pi_transcript = "&pi; transcript")
-# ft
-#flextable doesn't work with symbols well 
-#perhaps huxtable will? then huxtable can be converted to flextable
-# library(huxtable)
-# ht <- as_hux(x = div_stats)
-# names(ht) <-  c("transcript",
-#                                "symbol",
-#                                "&pi;<sub>N</sub>",
-#                                "&pi;<sub>S</sub>",
-#                                "&pi;<sub>N</sub>/&pi;<sub>S</sub>",
-#                                "&pi; coding",
-#                                "&pi; transcript")
-# ht <- add_colnames(ht, rowname = NULL)
-# escape_contents(ht)[1,] <- FALSE
-# print_notebook(ht)
-# ft <- as_flextable(ht, colnames_to_header = TRUE)
-# ft
-
-## perhaps pander is the way to go
-# library(pander)
-# pt <- div_stats
-# names(pt) <- c("transcript",
-#                                "symbol",
-#                                "&pi;<sub>N</sub>",
-#                                "&pi;<sub>S</sub>",
-#                                "&pi;<sub>N</sub>/&pi;<sub>S</sub>",
-#                                "&pi; coding",
-#                                "&pi; transcript")
-# pandoc.table(pt, round = c(0,0,rep(6, 5)))
-```

### 2018-07-12 manuscript/03-AFBs.Rmd
-Cleaned up manuscript folder -moved to bookdown format -revised all sections

--- 2018-07-10 manuscript/manuscript.Rmd
@@
----
-title: A comprehensive analysis of natural sequence variation within the <i>Arabidopsis
-  thaliana</i> nuclear auxin signaling pathway using an accessible web application
-author:
-  affiliation: University of Washington
-  name: R. Clay Wright
-date: "`r Sys.Date()`"
-params: 
-  output: word_document
-output:
-  word_document:
-    toc: yes
-  html_document:
-    fig_caption: yes
-    toc: yes
-always_allow_html: yes
-fontfamily: Helvetica
-fontsize: 11pt
-keywords: r markdown
-link-citations: yes
-csl: apa.csl
-bibliography:
-- natural_variation.bib
-- TIR1_AFB.bib
-- ARF_C.bib
-- AuxIAA_citations.bib
----
+# *TIR1/AFB* genes {#AFBs}
 
-```{r setup, message=FALSE, warning=FALSE}
-knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE, collapse = TRUE, comment = "#>", tidy = TRUE, cache = TRUE, cache.lazy = FALSE, out.width = "90%", fig.height=4, fig.width=7, out.extra='style="margin: auto; display: block; padding-top: 15px;"')
-
-#devtools::install_github("wrightrc/r1001genomes", ref = "auxin-natural-variation")
-library(ape)
-library(r1001genomes)
-library(DT)
-library(tidyverse)
-library(magrittr)
-library(RColorBrewer)
-library(ggpmisc)
-library(ggthemes)
-library(ggseqlogo)
-library(ggtree)
-library(reticulate)
-library(scales)
-library(viridis)
-library(ips)
-library(knitr)
-
-#use_python(python = "/usr/local/bin/python")
-```
-<!--This is a CSS (Cascading Style Sheet).
- It controls how panels and images are formatted.></!-->
-<style>
-  img {
-    display: block;
-    margin: auto;
-    width: 90%;
-    height: auto;
-  }
-</style>
-
-# List of working titles  
-* A comprehensive analysis of natural sequence variation within the <i>Arabidopsis thaliana</i> nuclear auxin signaling pathway  
-* **Add your suggestions**
-
-# Abstract
-
-* Motivating, high level statement  
-* Statement defining the cutting edge of your research area  
-* A statement of the form: "However, it remains unclear ..." that defines what problem you are  
-addressing.  
-* A statement of the form:"Here, we ..."  that concisely states your contribution.  
-* A series of 3?5 statement about your results.  
-* A concluding statement.
-
-# Intro
-
-The first genome sequence of *Arabidopsis thaliana* facilitated rapid advancement of plant biology through molecular genetics.
-Since this initial genome, massive scaling of sequencing technology has allowed a pioneering group to survey the global genomic variation in natural *Arabidopsis thaliana* populations.
-This valuable population genetics resource has led to several associations of genetic loci with phenotypic traits and provided insights into how selective pressure has and is influencing the evolution of plant genomes.
-Outside of genome-wide studies however, this valuable dataset has seen little use.
-Natural genomic variation also provides a set of feasible functional variation at the gene, gene family and gene network level that may provide insight into the function and evolution of genes, families and networks.
-Here, we present a web application and R-package through which plant molecular biologists with little-to-no bioinformatics experience can make use of this rich dataset of genetic variation to formulate hypotheses as to the sequence/function/phenotype relationships determined by the gene, family, or network of their interest.
-We demonstrate the utility of this tool through comprehensive analysis and identification of potential functional variation in the nuclear auxin signaling pathway.
-
-This application allows easy access to 
-1. lists of missense polymorphisms to facilitate biochemical assays of variant effects [@starita_variant_2017],
-2. family-wise alignments of variants to facilitate de novo functional domain identification [@melamed_combining_2015], 
-3.
-lists of accessions containing missense (or any type of) polymorphisms to facilitate segregation analysis and measurment of the distribution of variant effects on phenotype [@park_distributions_2017],
-4. as well as the study of gene and network evolution [@delker_natural_2010; @kliebenstein_role_2008].
-
-Here, we have utilized this tool to analyze the genetic variation in the nuclear auxin signaling pathway and formulate hypotheses regarding the functional implications of this variation and the evolution of the genes and gene families in this pathway.
-Our analysis also provided further confirmation of much of the existing knowledge of these genes demonstrating the validity of this approach.
-
-# Methods
-
-## Nucleotide Diversity Calculation
-Nei  defined the nucleotide diversity statistic in his original paper as:
-“…the average number of nucleotide differences per site between two randomly chosen DNA sequences.” (Nei, 1979)
-And provided the equation:
-
-$\displaystyle \pi = \sum_{ij}x_{i}x_{j}\pi_{ij}$
-
-Where Where $x_{i}$ is the frequency of the ith sequence in the population and $\pi_{ij}$ is the number of sites that are different between the ith and jth sequence divided by sequence length.
-
-a more general form, that treats each sequence in the population as unique can be written as:
-
-$\displaystyle \pi = \frac{1}{L*n^2} \sum_{i=1}^n \sum_{j=1}^n \sum_{k=1}^L{\pi_{ijk}}
-\qquad
-\displaystyle \pi_{ijk} = \left\{\begin{array}{l}
-1 \quad \textrm{if} \quad N_{ik} \neq N_{jk} \\
-0 \quad \textrm{if} \quad N_{ik} = N_{jk}
-\end{array}\right.$
-
-where $N_{ik}$ is the nucleotide (A, T, C or G) at position k on the ith sequence of the population.
-$L$ is the length of the sequence (the same for all sequences in the population, 
-assumes only SNPs exist, no indels). n is the total number of sequences in the population.
-
-from this form we can re-arrange summations to the form below:
-
-$\displaystyle \pi = \frac{1}{L} \sum_{k=1}^L{\pi_{k}}
-\qquad
-\pi_{k} = \frac{1}{n^2} \sum_{i=1}^n \sum_{j=1}^n {\pi_{ijk}}$
-
-here $\pi_k$ can be thought of as the site-wise nucleotide diversity at position k, and is equal to 
-the nucleotide diversity of a sequence of length 1 at location k. we can calculate $\pi_k$ for each
-site, then average those over the sequence length to calculate $\pi$, the nucleotide diversity 
-of the sequence.
-
-The function `Nucleotide_diversity` in the r1001genomes package calculates $\pi_k$ for each position in the gene or region that contains a variant.
-note, $\pi_k$ is equal to 0 at all locations without variants.
-This is also what is displayed in the Diversity Plot tab of the web app.
-
-## Detailed $\pi_k$ calculaiton simplification.
-
-The formula for $\pi_k$ above requires comparing every sequence to every other sequence at location k,
-however, we know there are only a few variant forms at each individual location.
-
-so we can revert back to using Nei's original formula, modifying it slightly,
-replacing $x_i$ with $\frac{n_i}{n}$, $n_i$ being the number of sequences in the population of the form $N_i$ at location k: subscripts i and k are summed only over unique forms [need to clarify this]
-
-$\displaystyle \pi_k = \sum_{ij} \frac{n_i}{n}\frac{n_j}{n} \pi_{ij} = 
-\frac{1}{n^2}\sum_{ij} n_i n_j \pi_{ij}$
-
-$\pi_{ij(k)} = \left\{\begin{array}{l}
-1 \quad \textrm{if} \quad i \neq j \\
-0 \quad \textrm{if} \quad i = j
-\end{array}\right.$
-
-we will define $n_{!i} = n - n_i$ as the number of sequences different from i at k.
-
-we can see the summed term will be zero if $i=j$, and $n_i n_j$ if $i\neq j$ so:
-
-$\displaystyle \pi_k = \frac{1}{n^2} \sum_{i} n_i n_{!i}$
-
-next we substitute our definition of $n_{!i}$
-
-$\displaystyle \pi_k = \frac{1}{n^2} \sum_{i} n_i (n - n_{i})$
-
-then distribute and split summation:
-
-$\displaystyle \pi_k = \frac{1}{n^2} (n \sum_{i} n_i - \sum_{i} n_{i}^2)$
-
-finally, summing $\displaystyle \sum_{i}n_i$ is equal to n:
-
-$\displaystyle \pi_k = \frac{1}{n^2} (n^2 - \sum_{i} n_{i}^2)$
-
-this simplified form for $\pi_k$ is used by the app, because the counts of unique 
-variants at a single nucleotide location can easily be summarized in R.
-
-# Results  
-## *TIR1/AFB* genes
-
-Auxin acts by binding to receptors that in turn target co-repressors for degradation.
-Auxin receptors (six in the model plant *Arabidopsis thaliana*) evolved through gene duplication and diversification early in the history of vascular plants [@Parry_Complex_2009].
+Auxin acts by binding to receptors (Auxin-signaling F-Boxes, or AFBs) that in turn target co-repressors (Aux/IAAs) for degradation.
+The six auxin receptor genes in the model plant *Arabidopsis thaliana* , *TIR1* and *AFB1-5*, evolved through gene duplication and diversification early in the history of vascular plants [@Parry_Complex_2009].
 The rate of co-repressor degradation is determined by the identity of both the receptor and co-repressor [@Havens_Synthetic_2012], and this rate sets the pace of lateral root development [@Guseman_Auxin-induced_2015].
@@
 thaliana differ in biochemical function and expression domain[@dharmasiri_plant_2005; @parry_complex_2009; @prigge_arabidopsis_2016].
-Although TIR1/AFBs are expressed ubiquitously in A.
+*TIR1/AFB* genes are expressed ubiquitously in A.
 thaliana tissues, TIR1, AFB2, and AFB3 have been shown to accumulate in the shoot and root meristems and leaf tissues, with slightly different expression patterns for TIR1 [@Parry_Complex_2009].
 Additionally, the expression of AFB5 is stronly circadian-regulated [@Covington_Circadian_2007] and AFB3 is more highly expressed in the roots in the presence of nitrate, allowing increased lateral root formation [@Vidal_Nitrate_2010], suggesting more broad environmental regulation of this gene family may exist.
+
 Functionally, all members of this family have been shown to bind auxin and Aux/IAA proteins.
@@
 ```{r AFB variants for PyMOL}
+alignment <- alignCDS(gene_info$transcript_ID)
+# make an alignment data frame
+aln_df <- makeAlnDF(alignment[[2]])
+
+# make a collapsed alignment for translating positions
+collapsed_aln_df <- 
+  aln_df %>% dplyr::select(aln_pos, seq_pos, seq_name) %>%
+  tidyr::spread(key = seq_name, value = seq_pos)
 coding_vcf <- ldply(.data = VCF, .fun = subset, 
@@
 p <- plotCodingDiv(uniqueCodingVars = coding_variants)
+annotation <- readAnnotationFile(filename = system.file("shiny-app", "gene_family_data",
+                      "AFB_annotations.csv", package = "r1001genomes"), gene_info = gene_info)
 annotation <- addAlnPosToAnno(anno_df = annotation, aln_df)
@@
 ```
-
-## *Aux/IAA* genes
-
-The Aux/IAA proteins (IAAs) mediate both auxin perception and transcriptional co-repression of auxin responsive genes.
-Auxin perception is carried out along with TIR1/AFB proteins, which together with IAAs act as co-receptors to bind and perceive nuclear auxin.
-Interaction of IAAs with auxin and TIR1/AFBs is conferred by the highly conserved 13 amino acid degron motif [@tan_mechanism_2007].
-Formation of this auxin co-receptor complex triggers ubiquitination and subsequent degradation of IAAs.
-This degradation in turn relieves their repression upon ARF transcription factors.
-Stabilizing mutations within the degron can lead to dominant phenotypes [@berleth_auxin_2004; @yang_iaa1_2004].
-The *Arabidopsis thaliana* Aux/IAA family has 29 members, most of which possess a canonical auxin degron.
-Notable exceptions are the auxin-insensitive IAA20 and IAA30 that lack a degron sequence entirely, while the closely related IAA31 retains a semi-functional degron sequence with a very slow degradation rate [@dreher_arabidopsis_2006; @havens_synthetic_2012].
-Additionally, IAA degradation rate is influenced by rate motifs found outside the degron, such as the KR motif [@dreher_arabidopsis_2006; @moss_rate_2015].
-IAA-mediated repression of ARFs depends upon both an EAR domain, N-terminal of the degron, that recruits co-repressors of the TOPLESS (TPL) family, and a Phox and Bem1 (PB1) domain that enables interaction via hetero-oligomerization with ARF PB1 domains [@nanao_structural_2014; @kagale_ear_2011].
-
-```{r IAA diversity stats, fig.height=11}
-gene_info <- geneInfoFromFile(
-  fname = system.file("shiny-app", "gene_family_data",
-                      "IAA_gene_ids.csv", package = "r1001genomes"))
-#annotation <- readAnnotationFile(filename = system.file("shiny-app", "gene_family_data",
-#                      "IAA_annotations.csv", package = "r1001genomes"), gene_info = gene_info)
-
-VCF <- readRDS(system.file("shiny-app", "gene_family_data",
-                      "IAAs_VCF.rds", package = "r1001genomes"))
-div_stats <- ldply(VCF, diversityStats, geneInfo=gene_info, .id="transcript_ID")
-div_stats$tair_symbol %<>% fct_reorder(.x = , div_stats$tair_symbol, 
-                                       .fun = function(x) 
-                                         {as.integer(str_remove(x, 
-                                              pattern = "IAA"))})
-div_stats %<>% arrange(tair_symbol)
-div_stats %<>% select(tair_symbol, everything())
-
-if (params$output == "html_document") {
-  formatRound(datatable(div_stats,
-                  colnames = c("transcript",
-                               "symbol",
-                               "&pi;<sub>N</sub>",
-                               "&pi;<sub>S</sub>",
-                               "&pi;<sub>N</sub>/&pi;<sub>S</sub>",
-                               "&pi; coding",
-                               "&pi; transcript"),
-                  rownames = FALSE, escape = FALSE,
-                  options = list(paging=FALSE, searching=FALSE)),
-                columns = 2:7, digits = 6)
-  } else 
-{
-  dt <- div_stats
-  names(dt) <- c("transcript",
-                                 "symbol",
-                                 "&pi;~N~",
-                                 "&pi;~S~",
-                                 "&pi;~N~/&pi;~S~",
-                                 "&pi; coding",
-                                 "&pi; transcript")
-  dt %<>% mutate_if(is.numeric, round, 6)
-  names(dt) <- paste0("**", names(dt), "**")
-  kable(dt) 
-}
-```
-
-Within the available natural sequence variation, we found several missense polymorphisms affecting the EAR domain or degron of several canonical IAAs.
-However, no missense polymorphisms were present in the KR motif or residues critical for PB1-PB1 domain interactions (Supp.
-Fig.
-Alignment of canonical IAAs).
-The majority of missense mutations observed were in the variable regions between the conserved domains.
-
-###Sister Pair Diversification
-
-Previous research has found evidence of both broad genetic redundancy and specificity within closely related pairs or groups of IAA proteins [@overvoorde_functional_2005].
-For example, the *iaa8-1 iaa9-1* double mutant and the *iaa5-1 iaa6-1 iaa19-1* triple mutant have wild-type phenotypes [@overvoorde_functional_2005], yet the *IAA6/19* sister pair has significant differences in expression patterns, protein abundances and functions suggesting they have undergone functional specialization since their divergence [@winkler_variation_2017].
-
-In our analysis of the *IAA* gene family, we frequently observed that one member of most *IAA* sister pairs had high nonsynonymous diversity while the other sister was more conserved (Figure IAA phylogeny heat map).
-In an interspecific comparison between *A.
-thaliana* and *A.
-lyrata*, *IAA19* is more conserved, while *IAA6* has regions of increased sequence divergence upstream of the degron and in the N-terminal half of the PB1 domain [@winkler_variation_2017].
-Consistent with these results, our analysis of intraspecific natural variation revealed higher conservation for *IAA19* (${\pi_N/\pi_S}$ = 0.55) compared to *IAA6* (${\pi_N/\pi_S}$ = 2.3) (Figure IAA phylogeny heat map), and also detected high diversity within the same regions of *IAA6* as seen in Winkler et al.
-(Supp Fig IAA6 diversity plot).
-```{r IAA alignment}
-alignment <- alignCDS(gene_info$transcript_ID)
-# make an alignment data frame
-aln_df <- makeAlnDF(alignment[[2]])
-
-# make a collapsed alignment for translating positions
-collapsed_aln_df <- 
-  aln_df %>% dplyr::select(aln_pos, seq_pos, seq_name) %>%
-  tidyr::spread(key = seq_name, value = seq_pos)
-```
-
-```{r Mask_IAA_alignment, eval=FALSE}
-aliscore(as.AAbin(alignment[[2]]), path = "/Volumes/SSD/Applications/ALISCORE_v2.0/Aliscore_v.2.0/")
-id <- scan(file = "input.fas_List_random.txt", sep = " ", quiet = TRUE)
-AAAlign_no_gaps <- as.matrix(alignment[[2]])[, id]
-write.nexus.data(AAAlign_no_gaps, file = "manuscript/IAAs.nex", format = "protein")
-```
-```{bash IAA_phylogeny_construction, eval=FALSE}
-~/Desktop/natural-variation-webtool/manuscript
-
-/Applications/MrBayes/mb
-
-execute IAAs.nex
-# set IAA33 as the outgroup
-outgroup AT5G57420.1
-
-lset nucmodel=protein Rates=invgamma
-# or include Nst=6  ngammacat=4 omegavar=M3 
-prset aamodelpr=fixed(jones) statefreqpr=fixed(empirical) 
-# initial run with aamodelpr=mixed yielded posterior prob of 1 for jones model
-propset ExtTBR$prob=0
-
-mcmcp ngen=250000 nchains=4 printfreq=1000 samplefreq=100 burnin=200
-
-mcmc
-sump
-sumt
-
-```
-
-```{r Load IAA phylogeny and data}
-tree <- treeio::read.beast(file = "IAAs.nex.con.tre")
-
-IAA_data <- read.csv("IAAs_table.csv")
-#tip_labels need to be in the first column
-IAA_data %<>% select(name, everything())
-IAA_data$name %<>% as.character()
-IAA_data$AGI %<>% as.character()
-
-IAA_data <- IAA_data[match(gsub(x = tree@phylo$tip.label, pattern = "\\.1",replacement = ""), IAA_data$AGI), ] 
-tree@phylo$tip.label <- IAA_data$name
-```
-*I think IAA20, 30, and 15 should also be in class C. So there are really two large clades A and B and then there are truncations/loss of function/dominant variants scattered throughout, Class C.*
-
-```{r IAA_tree, fig.height=5, fig.width=5, fig.cap="(ref:IAA tree)"}
-p <- ggtree(tree, layout = "rectangular", color = "grey") %<+% IAA_data + theme_tree() + scale_x_continuous(expand = c(0, 1.2)) + 
-  geom_tippoint(mapping = aes(size = 4*PiN_PiS, fill = PiN_PiS), shape = 21, color = "white", show.legend = TRUE) + 
-  scale_fill_viridis(values = rescale(c(min(IAA_data$PiN_PiS), 1,
-                                        max(IAA_data$PiN_PiS)), 
-                                      to = c(0,1)), guide = "colourbar") + scale_size(guide = "none") + 
-  theme(legend.position = c(.9,.7)) + 
-  geom_nodelab(mapping = aes(label = round(prob,digits = 2)), nudge_x = -.3, nudge_y = .5, size = 3) + 
-  geom_tiplab(mapping = aes(color = Class), hjust = -0.25) + 
-  scale_color_viridis(discrete = TRUE)
-p
-```
-
-(ref:IAA tree) IAA protein sequence tree mapped with ${\pi_N/\pi_S}$
-
-Within the Group B *IAA* genes, there are four sister pairs---*IAA10*/*IAA11*, *IAA12*/*IAA13*, *IAA18*/*IAA26*, and *IAA20*/*IAA30* \@ref(IAA tree) [@remington_contrasting_2004].
-Interestingly, among these sister pairs only *IAA12* and *IAA13* have a noticeable difference in nonsynonymous diversity.
-The *Arabidopsis thaliana* EFP browser shows that **IAA10** and *IAA11* have almost identical expression patterns, with the exception of *IAA10* being more highly expressed in the developing seed, and *IAA11* being more highly expressed in root tissue during lateral root development **need to cite original data source here**.
-In our analysis, *IAA10* and *IAA11* both showed functional conservation (${\pi_N/\pi_S}$ of 0.80 and 0.67 respectively) in support of *IAA10* and *IAA11* playing redundant roles.
-Accessions and associated variants identified here provide a potentially expedited route to elucidating the extent of redundancy between family members such as *IAA10* and *IAA11*.
-
-### New Loss-of-Function/Gain-of-Function alleles
-Accessions with functional variants can be potentially used to investigate the extent of redundancy/specialization within sister pairs.
-For example, three accessions have a missense variant affecting the EAR domain of IAA10 (HR-5, UKSE06-118 and UKSW06-240, all from southern UK).
-This variant (L49S) is in the second leucine of the LxLxL motif and likely reduces the ability of IAA10 to recruit the TOPLESS co-repressor [@lee_biological_2016].
-The nonsense variant (IAA10-Y219\*) resulting in truncated PB1 domain may be more indicative of *IAA10* null phenotype, as it likely prevents IAA10 from interacting with an ARF.
-This variant is only found in one accession from northeastern Spain (IP-Svi-0).
-These accessions and potential loss-of-function alleles may facilitate exploration of the function of *IAA10*.
-
-Several examples of possible gain-of-function missense variants were found in the degron of *IAA2*, *IAA3*, *IAA6*, *IAA7*, *IAA8*, *IAA13*, *IAA15*, *IAA18*, *IAA19*, *IAA27*, *IAA28* and *IAA29*.
-Most degron variants were very minor alleles in the population, only represented in one or two accessions.
-Interestingly, degron missense variants occured more frequently in *IAA* genes that have a closely related sister, which may indicate redundant function within the sister-pairs.
-*I think this is an interesting finding, but I'm not sure about this hypothesis of redundant function.
-If pairs are acting redundantly and we assume the degron variants are dominant, we would expect the auxin signal they transmit (their function) to be lower, in sum, as a pair.
-So I would actually hypothesize that there has been some divergence in function as a pair.
-*
-```{r Correlation between Distance to MRCA and prevalence of degron variants}
-
-```
-
-```{r Correlation between Pi_N/Pi_S and k_5}
-
-```
-
-*IAA6* has two potentially dominant missense mutations found within the degron (Supp Fig.
-alignment), one previously believed to lead to the *shy1-1* phenotype [@kim_two_1996].
-Interestingly, this variant is widely represented in 31 accessions across Europe (Supplemental Map Fig).
-This finding supports the proposition that iaa6-C78R may not be the mutation causing the *shy1-1D* phenotype but rather represents natural variation in the degron sequence ([note from Mike Prigge in TAIR](https://www.arabidopsis.org/servlets/TairObject?type=notepad&id=11247)).
-These accessions could be used to confirm whether *iaa6-C78R* is responsible for the *shy1-1D* phenotype.
-Additionally, an accession was identified with the IAA6-V77M variant, also in the degron, that may be useful to further explore *iaa6* gain-of-function phenotypes.
-
-### Identifying novel functional sequences in IAAs
-An exciting possible use for the Natural Variation Webtool is to search for conserved, low-diversity sites that may represent possible sites of functional import, such as post-translational modification or protein-protein interaction.
-For example, recent detailed biochemical analysis identified putative lysine ubiquitination sites in members of the *IAA6/19* sister pairs [@winkler_variation_2017]; Supp Fig IAA6/19 alignment).
-These sites are conserved across natural sequence variation, while several other lysine residues vary (Supp Figs alignment and map).
-In IAA6, there are 17 lysines of which ~6 are putative ubiquitination sites~ *I think this should be 5, K3, K32, K33, K91, and K97.
-Missense variants occured in K34Q, K140N *.
-One is proximal to the KR motif, while two others show natural variation.
-In IAA19, there are 15 total lysines, 8 of which are putative ubiquitination sites (K3, K25, K68, L87, K93, K100, K111, and K141).
-Two of these ubiquitination sites are proximal to the KR motif and degron.
-**This sentence is not clear.
-Only 1 of the ubiquitination sites varies?
-or one lysine, I think you mean this only 1 lysine that varies, but it is not ubiquitinated.
-Ubiquitylated Missense K129T** Natural sequence diversity may be used to narrow down a subset of lysines to assess for functional ubiquitination.
-
-```{r}
-
-```
-
-```{r IAA variants mapped to IAA17, eval=FALSE}
-coding_vcf <- ldply(.data = VCF, .fun = subset, 
-             !is.na(Transcript_ID) & gt_GT != "0|0")
-aln_df <- addSNPsToAlnDF(aln_df, coding_vcf, seq_name = Transcript_ID, 
-               seq_pos = Codon_Number)
-
-selections <- select_variants(gene_info, aln_df, structure_ref = "IAA17")
-selections$V2 <- paste0("chain X and resi ", selections$V2)
-write.table(selections, file = "IAA_variant_selections", row.names = FALSE, 
-          col.names = FALSE, sep = ",", quote = FALSE)
-```
-
-The members of the *IAA8/9/27* clade have a unique N-terminal extension with a highly conserved region common to orthologs in several other species [@dreher_arabidopsis_2006].
-*IAA8* and *IAA9* play redundant roles in leaf serration [@koenig_auxin_2009], while *IAA8* also plays an important role in lateral root development [@arase_iaa8_2012].
-IAA8 has been reported to uniquely localize in the cytosol [@arase_iaa8_2012] where it may interact with LSD1 protein [@coll_programmed_2011].
-The N-terminal extension has low natural sequence diversity in *IAA8* and *IAA9* (Supp Fig IAA8/9 alignment) that may represent a site of novel functionality, possibly mediating interaction with LSD1.
-The few accessions with missense alleles in this region in both *IAA8* and *IAA9* may be valuable in probing the functionality of this region (Supplemental Fig IAA8/9 map).
-
-```{r IAA8_9_27 alignment}
-
-```
-
-```{r IAA8_9 map}
-
-```
-
-Members of the *IAA17/7/14/16* clade have a second EAR motif (LxLxL) with no observed missense variants in *IAA7*, *IAA16*, or *IAA17* (Supp Fig IAA7/14/16/17 alignment).
-This observation supports the experimental evidence for IAA7 that the second EAR motif is functional, *i.e.* it can interact with TPR1 co-repressor and has a minor repressive role in auxin-related developmental phenotypes [lee_biological_2016].
-For *IAA14*, there were two missense variants that may be of interest to probe functionality of this second EAR motif.
-One variant, L41F, in the central lysine is represented in only one accession, and the other in the second “x” (K40N) is found in twelve accessions around Copenhagen (Supp Fig.
-IAA14 map).
-
-```{r IAA7_14_16_17 alignment}
-
-```
-
-```{r IAA14 map}
-
-```
-
-## *TPL/TPR* genes
-
-The Auxin signaling pathway utilizes the TOPLESS (TPL) and TOPLESS-related (TPR) family of Gro/TLE/TUP1 type co-repressor proteins to maintain auxin responsive genes in a transcriptionally-repressed state in the absence of auxin [@szemenyei_topless_2008].
-In *Arabidopsis thaliana* the five member *TPL/TPR* family includes *TPL* and *TPR1-4*.
-The resulting proteins are comprised of three structural domains: an N-terminal TPL domain and two WD-40 domains [@long_topless_2006].
-TPL/TPR proteins are recruited to the AUX/IAA proteins through interaction with the conserved Ethylene-responsive element binding factor-associated amphiphilic repression (EAR) domain [@szemenyei_topless_2008].
-Canonical EAR domains have the amino acid sequence LxLxL, as found in most AUX/IAAs (figure xA) [@overvoorde_functional_2005].
-TPL/TPR co-repressors bind EAR domains via their C-terminal to LisH (CTLH) domains found near their N-termini (citations of pre-structure founding papers/reviews).
-Recent structural analyses of the TPL N-terminal domain have highlighted the precise interaction interface between TPL and AUX/IAA EAR domains, as well as the TPL-TPL dimerization and tetramerization motifs [@martin-arevalillo_structure_2017; @ke_structural_2015a].
-The residues required for higher-order multimers of TPL tetramers have also been identified [@ma_d53_2017].
-Additional interactions with transcriptional regulation and chromatin modifying machinery are likely mediated by two tandem beta propeller domains of TPL/TPRs.
-These domain have not yet been crystallized, however numerous high resolution crystal structures of beta propeller domains allow for structural prediction of the TPL C-terminal beta propeller domains and the locations of likely interaction faces (Figure, X).
-
-The TOPLESS co-repressor family generally exhibits a high level of sequence conservation at the amino acid sequence level across resequenced *Arabidopsis thaliana* accessions, with all ${\pi_N/\pi_S}$ values below 1 (Table X).
-The closely related *TPL* and *TPR1* have the highest ${\pi_N/\pi_S}$ values (TPL-0.425, TPR1-0.739), suggesting that these these two related genes tolerate a higher degree of sequence and potentially functional diversity compared to *TPR2/3/4*.
-The N-terminal TPL domain of the TPL/TPR family is particularly conserved (Figure X).
-All nonsynonymous polymorphisms observed in this region are either in the coils between helices or are highly conservative mutations within helices (i.e.
-Valine to Isoleucine), which would be predicted to exhibit little effect on folding and function.
-
-```{r TPL SNP stats, fig.height=11}
-gene_info <- geneInfoFromFile(
-  fname = system.file("shiny-app", "gene_family_data",
-                      "TPL_gene_ids.csv", package = "r1001genomes"))
-#annotation <- readAnnotationFile(filename = system.file("shiny-app", "gene_family_data",
-#                      "IAA_annotations.csv", package = "r1001genomes"), gene_info = gene_info)
-
-VCF <- readRDS(system.file("shiny-app", "gene_family_data",
-                      "TPL_VCF.rds", package = "r1001genomes"))
-div_stats <- ldply(VCF, diversityStats, geneInfo=gene_info, .id="transcript_ID")
-div_stats$tair_symbol %<>% fct_reorder(.x = , div_stats$tair_symbol, 
-                                       .fun = function(x) 
-                                         {as.integer(str_remove(x, 
-                                              pattern = "TPR"))})
-div_stats$tair_symbol %<>% fct_relevel("TPL")
-div_stats %<>% arrange(tair_symbol)
-div_stats %<>% select(tair_symbol, everything())
-
-if (params$output == "html_document") {
-  formatRound(datatable(div_stats,
-                  colnames = c("transcript",
-                               "symbol",
-                               "&pi;<sub>N</sub>",
-                               "&pi;<sub>S</sub>",
-                               "&pi;<sub>N</sub>/&pi;<sub>S</sub>",
-                               "&pi; coding",
-                               "&pi; transcript"),
-                  rownames = FALSE, escape = FALSE,
-                  options = list(paging=FALSE, searching=FALSE)),
-                columns = 2:7, digits = 6)
-  } else 
-{
-  dt <- div_stats
-  names(dt) <- c("transcript",
-                                 "symbol",
-                                 "&pi;~N~",
-                                 "&pi;~S~",
-                                 "&pi;~N~/&pi;~S~",
-                                 "&pi; coding",
-                                 "&pi; transcript")
-  dt %<>% mutate_if(is.numeric, round, 6)
-  names(dt) <- paste0("**", names(dt), "**")
-  kable(dt) 
-}
-```
-
-The high degree of conservation in the entire N-terminal domain underscores its importance in TPL/TPR function.
-For example, the initial *tpl-1* mutation (N176H) in the ninth helix is a dominant gain-of-function allele [@long_topless_2006], which is capable of binding wild-type TPL protein and inducing protein aggregation [@ma_d53_2017].
-It is therefore understandable that this helix had very low diversity as nonsynonymous variants in this domain could act in a dominant negative fashion.
-Addtionally, the crystal structure of TPL has revealed that the The LisH domain and CRA domain coordinate TPL dimerization and tetramerization respectively (26601214, 28698367), the CTLH domain is critical for interaction with DNA-binding transcription factors (26601214), and helix nine within the CRA domain may be critical for the formation of higher-order TPL complexes (multimers of TPL tetramers, 28630893).
-
-Compared to the TPL domain, both the proline rich and WD-40 domains exhibited a higher frequency of missense variants.
-We utilized Phyre2.0 modeling software to predict the structure of the TPL/TPR WD-40 repeats in order to map non-synonymous mutations to protein structure (Figure X).
-Here we can see both faces of each predicted beta-propeller structure of the WD-40 domains, which are connected by a short peptide linker.
-This model reveals amino acid conservation in the core and central residues of the WD-40 domains.
-~~These observations could be used as the basis for forward mutagenesis strategies to test whether this region of TPL functions similarly to its homolog TLE, which binds its target genes EH1 and the C-terminal WRPW/Y motif (Hairy/Hes/Runx) in the center of the WD-40 domain (pmid 16762837).~~ **Can we say something more general here?
-There is not enough background to support this.
-What is TLE from?
-What is EH1?
-Is TLE really binding to EH1 DNA?**
-
-```{r TPL_TPR alignment}
-
-```
-
-Despite the presence of five *TPL/TPR* genes in the genome, which are considered to be redundant (pmid 16763149), very few observed SNPs are predicted to result in loss of function for any *TPL/TPR* gene.
-Only one frameshift variant was observed, *TPR4-665delG*, in a single accession (Aiell-4 from Italy).
-This deletion occurs in codon 222, just after the TPL N-terminal domain, and generates an early stop eleven codons downstream from the deletion.
-It is unclear whether this N-terminal fragment would be expressed at any appreciable level, as it is possible that the generation of an early stop codon would trigger nonsense-mediated decay.
-However, if expressed, it would be fascinating to examine whether this naturally occurring N-terminal truncation retains activity as a co-repressor.
-Given the observation that the TPL N-terminal domain is sufficient to bind histones (28630893), and repress transcription via synthetic assays in yeast (24979769), it is possible that this represents a semi-functional TPL variant.
-There has been no evidence for prioritization of individual *TPL/TPR* family members in the auxin response pathway, but experiments modulating the gene copy number of *TPR1* revealed the importance of maintaining the proper dosage of *TPL/TPR* in order to accurately trigger the salicylic acid innate immune response (20647385).
-These results, taken together with the absence of *TPL/TPR* loss-of-function alleles in the *Arabidopsis thaliana* accessions suggest that perhaps the entire family is being conserved at the nucleotide level in the population to meet a gene dosage requirement, with little opportunity for neo- or sub-functionalization.
-
-## ARF genes
-
-Auxin response is ultimately mediated by the auxin responsive transcription factors (ARFs).
-There are 23 ARFs in *Arabidopsis thaliana* that are divided into three phylogenetic classes.
-Class A ARFs (ARF5, ARF6, ARF7, ARF8 and ARF19) activate transcription.
-These ARFs have a glutamine-rich region in the middle of the protein that may mediate activation (Guilefoyle, 2007).
-It has recently been shown that the middle region of ARF5 interacts with the SWI/SNF chromatin remodeling ATPases BRAMA and SPLAYED, possibly to reduce nucleosome occupancy and allow for the recruitment of transcription machinery (Wu, 2016).
-Additionally, ARF7 interacts with Mediator subunits, directly tethering transcriptional activation machinery to its binding sites in the chromosome (Ito, 2016).
-Class B and C ARFs are historically categorized as repressor ARFs, though the mechanism through which they confer repression has not been identified.
-Their middle regions tend to be proline- and serine-rich (citation, Ulmasov?).
-
-Canonical ARFs are comprised of three major domains.
-Recent crystallization of these domains have informed structure-function analysis of the ARFs (Boer, 2014; Nanao, 2014; Korasick, 2014).
-These domains are conserved throughout land plants (Dolf's new paper).
-ARFs share an N-terminal B3 DNA binding domain.
-Flanking this DNA-binding domain is a dimerization domain, which folds up into a single “taco-shaped” domain to allow for dimerization between ARFs.
-There is an auxiliary domain that immediately follows and interacts with the dimerization domain.
-The middle region is the most variable between ARFs, as mentioned above, but is characterized by repetitive units of glutamine (class A), serine, or proline residues (classes B and C).
-
-The C-terminal domain of the ARF is a protein-protein interaction domain mediating interactions among ARFs, between ARFs and other transcription factors, and between ARFs and the Aux/IAA repressors.
-This interaction domain was recently characterized as a Phox and Bem1 (PB1) domain, which is comprised of a positive and negative face with conserved basic and acidic residues, respectively (Nanao, 2014; Korasick, 2014).
-The dipolar nature of the PB1 domain may mediate multimerization by the pairwise interaction of these faces on different proteins as the ARF7 PB1 domain was crystallized as a multimer [@korasick_].
-However, it is unclear whether ARF multimerization occurs or plays a developmental significant role in vivo.
-Interfering with ARF dimerization in either the DNA-binding proximal dimerization domain or the PB1 domain decreases the ability of class A ARFs to activate transcription in a heterologous yeast system (Pierre-Jerome, 2016).
-
-While domain architecture is broadly conserved among the ARFs, there are exceptional cases.
-Three ARFs do not contain a PB1 domain at all, ARF3, ARF13, and ARF17, and several more have lost the conserved acidic or basic residues in the PB1 domain, suggesting they may be reduced to a single interaction domain.
-Several ARFs additionally have an expanded conserved region within the DNA-binding domain, of unknown function.
-The majority of domain variation among ARFs occurs in the large B-class subfamily.
-The liverwort *Marchantia polymorpha* has a single representative ARF of each class (Flores-Sandoval, 2015).
-The expansion of these classes in flowering plants is the result of both whole genome and tandem duplication events (Remington, 2004).
-The growth of the ARF family may have allowed for the expansion of the quantity and complexity of loci regulated by the ARFs and subsequent expansion in their regulation of developmental processes.
-
-Class A ARFs are the most well-studied ARF subfamily---the five family members all act as transcriptional activators and have well-characterized, distinct developmental targets.
-Overall the diversity of class A ARFs was generally low, especially compared to the class B and C ARFs, suggesting that class A ARFs are central to auxin signal transduction and plant development.
-Analysis of class A ARF nonsynonymous diversity suggests that the majority of these ARFs are highly functionally conserved, with ${\pi_N/\pi_S}$ values much lower than 1 with the exception of ARF19, with ${\pi_N/\pi_S}$ value of 1.8.
-Comparing diversity within sister pairs, there is a similar trade-off as seen in most *IAA* sister pairs, with one sister being highly conserved and the other more divergent.
-ARF19 and ARF8 are the more divergent class A ARFs, with ${\pi_N/\pi_S}$ values at least three time those of their sisters, ARF7 and ARF6 respectively.
-This may suggest that ARF6 and ARF7 serve more essential purposes in plant development.
-
-```{r ARF diversity stats}
-gene_info <- geneInfoFromFile(
-  fname = system.file("shiny-app", "gene_family_data",
-                      "ARF_gene_ids.csv", package = "r1001genomes"))
-#annotation <- readAnnotationFile(filename = system.file("shiny-app", 
-  #"gene_family_data", "ARF_annotations.csv", 
-  #package = "r1001genomes"), gene_info = gene_info)
-
-VCF <- readRDS(system.file("shiny-app", "gene_family_data",
-                      "ARF_VCF.rds", package = "r1001genomes"))
-div_stats <- ldply(VCF, diversityStats, geneInfo=gene_info, .id="transcript_ID")
-div_stats$tair_symbol %<>% fct_reorder(.x =  div_stats$tair_symbol, 
-                                       .fun = function(x) 
-                                         {as.integer(str_remove(x, 
-                                              pattern = "ARF"))})
-div_stats %<>% arrange(tair_symbol)
-div_stats %<>% select(tair_symbol, everything())
-
-if (params$output == "html_document") {
-  formatRound(datatable(div_stats,
-                  colnames = c("transcript",
-                               "symbol",
-                               "&pi;<sub>N</sub>",
-                               "&pi;<sub>S</sub>",
-                               "&pi;<sub>N</sub>/&pi;<sub>S</sub>",
-                               "&pi; coding",
-                               "&pi; transcript"),
-                  rownames = FALSE, escape = FALSE,
-                  options = list(paging=FALSE, searching=FALSE)),
-                columns = 2:7, digits = 6)
-  } else 
-{
-  dt <- div_stats
-  names(dt) <- c("transcript",
-                                 "symbol",
-                                 "&pi;~N~",
-                                 "&pi;~S~",
-                                 "&pi;~N~/&pi;~S~",
-                                 "&pi; coding",
-                                 "&pi; transcript")
-  dt %<>% mutate_if(is.numeric, round, 6)
-  names(dt) <- paste0("**", names(dt), "**")
-  kable(dt) 
-}
-```
-
-```{r ARF alignment}
-alignment <- alignCDS(gene_info$transcript_ID)
-# make an alignment data frame
-aln_df <- makeAlnDF(alignment[[2]])
-
-# make a collapsed alignment for translating positions
-collapsed_aln_df <- 
-  aln_df %>% dplyr::select(aln_pos, seq_pos, seq_name) %>%
-  tidyr::spread(key = seq_name, value = seq_pos)
-```
-
-```{r Mask_ARF_alignment, eval=FALSE}
-aliscore(as.AAbin(alignment[[2]]), path = "/Volumes/SSD/Applications/ALISCORE_v2.0/Aliscore_v.2.0/")
-id <- scan(file = "/Volumes/SSD/Applications/ALISCORE_v2.0/Aliscore_v.2.0/input.fas_List_random.txt", sep = " ", quiet = TRUE)
-AAAlign_no_gaps <- as.matrix(alignment[[2]])[, id]
-write.nexus.data(AAAlign_no_gaps, file = "ARFs.nex", format = "protein")
-```
-```{bash ARF_phylogeny_construction, eval=FALSE}
-/Applications/MrBayes/mb
-
-execute ARFs.nex
-# set ARF17 as the outgroup
-outgroup AT1G77850.1
-
-lset nucmodel=protein Rates=invgamma
-# or include Nst=6  ngammacat=4 omegavar=M3 
-prset aamodelpr=fixed(jones) statefreqpr=fixed(empirical) 
-# initial run with aamodelpr=mixed yielded posterior prob of 1 for jones model
-propset ExtTBR$prob=0
-
-mcmcp ngen=250000 nchains=4 printfreq=1000 samplefreq=100 burnin=200
-
-mcmc
-sump
-sumt
-
-```
-
-```{r Load ARF phylogeny and data}
-tree <- treeio::read.beast(file = "ARFs.nex.con.tre")
-
-ARF_data <- read.csv(system.file("shiny-app", "gene_family_data",
-                      "ARF_gene_ids.csv", package = "r1001genomes"))
-ARF_data %<>% left_join(y = div_stats$x$data, by = c("name" = "tair_symbol"))
-#tip_labels need to be in the first column
-ARF_data %<>% select(name, everything())
-ARF_data$tair_locus %<>% as.character()
-
-ARF_data <- ARF_data[match(gsub(x = tree@phylo$tip.label, pattern = "\\.1",replacement = ""), ARF_data$tair_locus), ]
-
-tree@phylo$tip.label <- ARF_data$name
-
-```
-
-```{r Visualize ARF tree, fig.height=5, fig.width=5}
-p <- ggtree(tree, layout = "rectangular", color = "grey") %<+% ARF_data + theme_tree() + scale_x_continuous(expand = c(0, 1.2)) + geom_tiplab(hjust = -0.25) + 
-  geom_tippoint(mapping = aes(size = 4*Pi_NS_Ratio, fill = Pi_NS_Ratio), 
-                shape = 21, color = "white", show.legend = TRUE) + 
-  scale_fill_viridis(values = rescale(c(min(ARF_data$Pi_NS_Ratio), 1,
-                                        max(ARF_data$Pi_NS_Ratio)), 
-                                      to = c(0,1)), guide = "colourbar") + scale_size(guide = "none") + 
-  theme(legend.position = c(.1,.7)) + 
-  geom_nodelab(mapping = aes(label = round(prob,digits = 2)), nudge_x = -.3, nudge_y = .5, size = 3) +
-  geom_tiplab(mapping = aes(color = class), hjust = -0.25) + 
-  scale_color_viridis(discrete = TRUE)
-p
-```
-
-### Class A ARFs 
-For all class A ARFs, the middle region of the protein was the predominant high diversity region.
-The DNA-binding domain had very few, low-diversity missense mutations, as did the PB1 domain.
-Considering the necessity of their conserved functions, the low level of variation in these key DNA and protein-protein interaction domains is expected.
-
-```{r A ARF Alignment}
-gene_info <- subset(gene_info, tair_symbol %in% c("ARF5", "ARF6", "ARF7", "ARF8", "ARF19"))
-alignment <- alignCDS(gene_info$transcript_ID)
-# make an alignment data frame
-aln_df <- makeAlnDF(alignment[[2]])
-#BrowseSeqs(alignment[[2]])
-aln_df <- makeAlnDF(alignment[[2]])
-aln_df <- left_join(aln_df, dplyr::select(gene_info, "tair_locus",
-                                       "tair_symbol", "transcript_ID"),
-                        by = c("transcript_ID" = "transcript_ID"))
-aln_df$tair_symbol %<>% as.factor()
-aln_df$tair_symbol %<>% fct_reorder(
-    {as.integer(str_remove(aln_df$tair_symbol, 
-                           pattern = "ARF"))}, 
-  .desc = TRUE)
-
-aln_df <- chunkAlnDF(aln_df = aln_df, chunk_width = 100)
-VCF <- ldply(.data = VCF, .fun = subset, gt_GT != "0|0")
-VCF <- getCodingDiv(VCF)
-aln_df <- addSNPsToAlnDF(aln_df, VCF)
-A_ARFs <- subset(aln_df, tair_symbol %in% c("ARF5", "ARF6", "ARF7", "ARF8", "ARF19"))
-p <-ggplot(A_ARFs, aes(x = aln_pos, y = tair_symbol,
-                             group = seq_pos, text = variants))
-p <- p +
-  geom_raster(data = na.omit(A_ARFs), mapping = aes(fill = effects)) +
-  scale_fill_viridis(option = "A", discrete = TRUE) +
-  scale_y_discrete() +
-  # expand increases distance from axis
-  labs(x = "", y = "", fill = "variant") +
-  theme_logo(base_family = "Helvetica") +
-  theme(panel.grid = element_blank(), panel.grid.minor = element_blank()) +
-  theme(strip.background = element_blank(),
-        strip.text.x = element_blank(),
-        legend.box = "vertical")
-p
-#ggsave(p,filename = "A_ARFs_alignment_variants.pdf", width = 12, height = 4)
-```
-
-Class A ARFs all contain the canonical B3 DNA-binding domain, PB1 interaction domain and glutamine-rich middle region (Guilefoyle, 2007).
-In the analyzed natural variation, ARF7 had several expansions of polyglutamine sequences in the middle region.
-Polyglutamine regions are known to readily expand and contract throughout evolutionary time due to replication error, and variation in polyglutamine length can be acted on by natural selection and have phenotypic consequences (Press, 2014).
-
-A few missense variants in ARF5 highlighted potential functional variation in the class A ARFs.
-A high frequency allele (*What are the variants?*), resulting in substitutions of small amino acids on the positive face of the PB1 domain for bulky phenylalanines, may disrupt interactions with other ARF or Aux/IAA PB1 domains.
-Interestingly, these variants were always found together in accessions Tu-B2-3, Ru-2, Haes-1, Rd-0, 627ME-13Y1, Le-0, Bu-0 and Bsch-0 *Where are these?*.
-Another highly represented polymorphism in ARF5 results in change in the 17th amino acid from a leucine to a methionine.
-This position is in the N-terminal extension of ARF5 which is of unknown function.
-It is possible that this methionine acts as an alternative start site in the accessions.
-
-### Class B ARFs
-
-*ARF23*, which has a truncated DNA-binding domain, has a high ${\pi_N/\pi_S}$ value of 4.1, and has many high-diversity nonsense variants, as may be expected of a gene that is pseudogenizing.
-
-ARF3, ARF13 and ARF17 all lack a C-terminal PB1 domain, which suggests they may be unable to interact with Aux/IAA repressors, possibly acting as auxin-insensitive competitive inhibitors of ARF-mediated transcription (Fig ARF3_13_17 alignment).
-These ARFs span the range of ${\pi_N/\pi_S}$ ratios relative to the other ARFs, ranging from 0.85 to 4.1.
-ARF17 is more conserved than the activator ARF, ARF19.
-ARF13 has many nonsense variants of high diversity and may be pseudogenizing.
-
-```{r ARF3_13_17 alignment}
-
-```
-
-ARF14, ARF15, ARF16, ARF20, and ARF21 all do not have one or multiple of the conserved acidic residues in the PB1 domain, and themselves form a subclade of the B ARFs.
-Interestingly, no non-synonymous variation is seen in these amino acid residues of these ARFs, suggesting that the amino acids that are in the regions may still be important for the function of these proteins.
-
-### Class C ARFs
-
-Class C ARFs show low nucleotide diversity scores, all ${\pi_N/\pi_S}$ values substantially lower than 1.
-ARF16 was the most conserved, whereas its clade members (ARF10, ARF17) had scores at least four times higher [@remington_contrasting_2004].
-
-Structurally, all three members of Class C ARFs contain a canonical B3 DNA-binding domain, but only ARF10 and ARF16 contain a PB1 domain.
-The DNA binding domains exhibit overall low diversity and presence of largely synonymous changes for all the ARFs.
-Only the ARF10 DNA binding domain contains missense variants (S147T and A157S) (Fig.
-ARF10_DBD Variants distribution).
-**Any insights into whether these may affect function?
-They are quite conservative.** Of the PB1 domain containing class C ARFs, ARF16 exhibits several missense variants which are sporadically distributed, in contrast to the conserved PB1 domain of ARF10 (Fig.
-Arf16_PB_variants distribution).
-
-```{r ARF10_DBD Variants distribution}
-
-```
-
-```{r Arf16_PB_variants distribution}
-
-```
-
-Several members of class B (ARF2, 3 and 4) and C ARFs (ARF10, 16 and 17) are distinguished from other ARFs on the basis of their post transcriptional regulation by trans-acting small interfering RNAs (tasiRNAs) and microRNAs (miRNAs) [@rhoades_prediction_2002; @allen_micrornadirected_2005].
-**(@AMY Arf 6 and 8 targeted by miRNA 167 - check for conservation).** The miRNA target of class C ARFs showed a high degree of conservation (Fig.
-ARF10_16_17_Alignment).
-This is consistent with perturbation studies that have shown drastic developmental phenotype for class C ARFs [@wang_control_2005; @mallory_micrornadirected_2005; @liu_repression_2007].
-For the class B ARFs (2, 3 and 4) the tasiRNA binding motif is also highly conserved suggesting strong regulatory constraints on these genes (Fig ARF2_3_4_Alignment).
-The DNA binding domain of ARF4 exhibits non-synonymous variation in 4 accession and would be interesting to carry out phenotypic characterization of these accessions for developmental growth defects (Fig ARF4_DBD Variants map).
-
-```{r ARF10_16_17_Alignment}
-
-```
-
-```{r ARF2_3_4_Alignment}
-
-```
-
-```{r ARF4_DBD Variants map}
-
-```
-
-#Discussion
-
-Analyzing natural diversity within the gene families constituting the auxin nuclear signaling pathway has confirmed existing knowledge, supported existing hypothses and generated new hypotheses.
-This analysis in particular has confirmed the importance of auxin-induced degradation of the Aux/IAA proteins as evidenced by the strong conservation of the *AFB* genes and the canonical degron region of the *Aux/IAA* genes.
-Despite this relative conservation, this analysis has also identified many accessions containing new alleles of interest in the *AFB* and *Aux/IAA* families.
-This analysis has also
-
-Importance of 
--AFB
--IAA degron
--TPL
--ARF DBD and PB1

### 2018-07-12 manuscript/04-IAAs.Rmd
-Cleaned up manuscript folder -moved to bookdown format -revised all sections

--- 2018-07-10 manuscript/manuscript.Rmd
@@
----
-title: A comprehensive analysis of natural sequence variation within the <i>Arabidopsis
-  thaliana</i> nuclear auxin signaling pathway using an accessible web application
-author:
-  affiliation: University of Washington
-  name: R. Clay Wright
-date: "`r Sys.Date()`"
-params: 
-  output: word_document
-output:
-  word_document:
-    toc: yes
-  html_document:
-    fig_caption: yes
-    toc: yes
-always_allow_html: yes
-fontfamily: Helvetica
-fontsize: 11pt
-keywords: r markdown
-link-citations: yes
-csl: apa.csl
-bibliography:
-- natural_variation.bib
-- TIR1_AFB.bib
-- ARF_C.bib
-- AuxIAA_citations.bib
----
-
-```{r setup, message=FALSE, warning=FALSE}
-knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE, collapse = TRUE, comment = "#>", tidy = TRUE, cache = TRUE, cache.lazy = FALSE, out.width = "90%", fig.height=4, fig.width=7, out.extra='style="margin: auto; display: block; padding-top: 15px;"')
-
-#devtools::install_github("wrightrc/r1001genomes", ref = "auxin-natural-variation")
-library(ape)
-library(r1001genomes)
-library(DT)
-library(tidyverse)
-library(magrittr)
-library(RColorBrewer)
-library(ggpmisc)
-library(ggthemes)
-library(ggseqlogo)
-library(ggtree)
-library(reticulate)
-library(scales)
-library(viridis)
-library(ips)
-library(knitr)
-
-#use_python(python = "/usr/local/bin/python")
-```
-<!--This is a CSS (Cascading Style Sheet).
- It controls how panels and images are formatted.></!-->
-<style>
-  img {
-    display: block;
-    margin: auto;
-    width: 90%;
-    height: auto;
-  }
-</style>
-
-# List of working titles  
-* A comprehensive analysis of natural sequence variation within the <i>Arabidopsis thaliana</i> nuclear auxin signaling pathway  
-* **Add your suggestions**
-
-# Abstract
-
-* Motivating, high level statement  
-* Statement defining the cutting edge of your research area  
-* A statement of the form: "However, it remains unclear ..." that defines what problem you are  
-addressing.  
-* A statement of the form:"Here, we ..."  that concisely states your contribution.  
-* A series of 3?5 statement about your results.  
-* A concluding statement.
-
-# Intro
-
-The first genome sequence of *Arabidopsis thaliana* facilitated rapid advancement of plant biology through molecular genetics.
-Since this initial genome, massive scaling of sequencing technology has allowed a pioneering group to survey the global genomic variation in natural *Arabidopsis thaliana* populations.
-This valuable population genetics resource has led to several associations of genetic loci with phenotypic traits and provided insights into how selective pressure has and is influencing the evolution of plant genomes.
-Outside of genome-wide studies however, this valuable dataset has seen little use.
-Natural genomic variation also provides a set of feasible functional variation at the gene, gene family and gene network level that may provide insight into the function and evolution of genes, families and networks.
-Here, we present a web application and R-package through which plant molecular biologists with little-to-no bioinformatics experience can make use of this rich dataset of genetic variation to formulate hypotheses as to the sequence/function/phenotype relationships determined by the gene, family, or network of their interest.
-We demonstrate the utility of this tool through comprehensive analysis and identification of potential functional variation in the nuclear auxin signaling pathway.
-
-This application allows easy access to 
-1. lists of missense polymorphisms to facilitate biochemical assays of variant effects [@starita_variant_2017],
-2. family-wise alignments of variants to facilitate de novo functional domain identification [@melamed_combining_2015], 
-3.
-lists of accessions containing missense (or any type of) polymorphisms to facilitate segregation analysis and measurment of the distribution of variant effects on phenotype [@park_distributions_2017],
-4. as well as the study of gene and network evolution [@delker_natural_2010; @kliebenstein_role_2008].
-
-Here, we have utilized this tool to analyze the genetic variation in the nuclear auxin signaling pathway and formulate hypotheses regarding the functional implications of this variation and the evolution of the genes and gene families in this pathway.
-Our analysis also provided further confirmation of much of the existing knowledge of these genes demonstrating the validity of this approach.
-
-# Methods
-
-## Nucleotide Diversity Calculation
-Nei  defined the nucleotide diversity statistic in his original paper as:
-“…the average number of nucleotide differences per site between two randomly chosen DNA sequences.” (Nei, 1979)
-And provided the equation:
-
-$\displaystyle \pi = \sum_{ij}x_{i}x_{j}\pi_{ij}$
-
-Where Where $x_{i}$ is the frequency of the ith sequence in the population and $\pi_{ij}$ is the number of sites that are different between the ith and jth sequence divided by sequence length.
-
-a more general form, that treats each sequence in the population as unique can be written as:
-
-$\displaystyle \pi = \frac{1}{L*n^2} \sum_{i=1}^n \sum_{j=1}^n \sum_{k=1}^L{\pi_{ijk}}
-\qquad
-\displaystyle \pi_{ijk} = \left\{\begin{array}{l}
-1 \quad \textrm{if} \quad N_{ik} \neq N_{jk} \\
-0 \quad \textrm{if} \quad N_{ik} = N_{jk}
-\end{array}\right.$
-
-where $N_{ik}$ is the nucleotide (A, T, C or G) at position k on the ith sequence of the population.
-$L$ is the length of the sequence (the same for all sequences in the population, 
-assumes only SNPs exist, no indels). n is the total number of sequences in the population.
-
-from this form we can re-arrange summations to the form below:
-
-$\displaystyle \pi = \frac{1}{L} \sum_{k=1}^L{\pi_{k}}
-\qquad
-\pi_{k} = \frac{1}{n^2} \sum_{i=1}^n \sum_{j=1}^n {\pi_{ijk}}$
-
-here $\pi_k$ can be thought of as the site-wise nucleotide diversity at position k, and is equal to 
-the nucleotide diversity of a sequence of length 1 at location k. we can calculate $\pi_k$ for each
-site, then average those over the sequence length to calculate $\pi$, the nucleotide diversity 
-of the sequence.
-
-The function `Nucleotide_diversity` in the r1001genomes package calculates $\pi_k$ for each position in the gene or region that contains a variant.
-note, $\pi_k$ is equal to 0 at all locations without variants.
-This is also what is displayed in the Diversity Plot tab of the web app.
-
-## Detailed $\pi_k$ calculaiton simplification.
-
-The formula for $\pi_k$ above requires comparing every sequence to every other sequence at location k,
-however, we know there are only a few variant forms at each individual location.
-
-so we can revert back to using Nei's original formula, modifying it slightly,
-replacing $x_i$ with $\frac{n_i}{n}$, $n_i$ being the number of sequences in the population of the form $N_i$ at location k: subscripts i and k are summed only over unique forms [need to clarify this]
-
-$\displaystyle \pi_k = \sum_{ij} \frac{n_i}{n}\frac{n_j}{n} \pi_{ij} = 
-\frac{1}{n^2}\sum_{ij} n_i n_j \pi_{ij}$
-
-$\pi_{ij(k)} = \left\{\begin{array}{l}
-1 \quad \textrm{if} \quad i \neq j \\
-0 \quad \textrm{if} \quad i = j
-\end{array}\right.$
-
-we will define $n_{!i} = n - n_i$ as the number of sequences different from i at k.
-
-we can see the summed term will be zero if $i=j$, and $n_i n_j$ if $i\neq j$ so:
-
-$\displaystyle \pi_k = \frac{1}{n^2} \sum_{i} n_i n_{!i}$
-
-next we substitute our definition of $n_{!i}$
-
-$\displaystyle \pi_k = \frac{1}{n^2} \sum_{i} n_i (n - n_{i})$
-
-then distribute and split summation:
-
-$\displaystyle \pi_k = \frac{1}{n^2} (n \sum_{i} n_i - \sum_{i} n_{i}^2)$
-
-finally, summing $\displaystyle \sum_{i}n_i$ is equal to n:
-
-$\displaystyle \pi_k = \frac{1}{n^2} (n^2 - \sum_{i} n_{i}^2)$
-
-this simplified form for $\pi_k$ is used by the app, because the counts of unique 
-variants at a single nucleotide location can easily be summarized in R.
-
-# Results  
-## *TIR1/AFB* genes
-
-Auxin acts by binding to receptors that in turn target co-repressors for degradation.
-Auxin receptors (six in the model plant *Arabidopsis thaliana*) evolved through gene duplication and diversification early in the history of vascular plants [@Parry_Complex_2009].
-The rate of co-repressor degradation is determined by the identity of both the receptor and co-repressor [@Havens_Synthetic_2012], and this rate sets the pace of lateral root development [@Guseman_Auxin-induced_2015].
-
-Although it is unclear what unique roles each receptor plays in growth and development, a number of studies have pointed out differences in the ways the six different receptors in A.
-thaliana differ in biochemical function and expression domain[@dharmasiri_plant_2005; @parry_complex_2009; @prigge_arabidopsis_2016].
-Although TIR1/AFBs are expressed ubiquitously in A.
-thaliana tissues, TIR1, AFB2, and AFB3 have been shown to accumulate in the shoot and root meristems and leaf tissues, with slightly different expression patterns for TIR1 [@Parry_Complex_2009].
-Additionally, the expression of AFB5 is stronly circadian-regulated [@Covington_Circadian_2007] and AFB3 is more highly expressed in the roots in the presence of nitrate, allowing increased lateral root formation [@Vidal_Nitrate_2010], suggesting more broad environmental regulation of this gene family may exist.
-  Functionally, all members of this family have been shown to bind auxin and Aux/IAA proteins.
-However, AFB1 has drastically reduced ability to assemble into an SCF complex, due to the substitution E8K in its F-box domain, preventing it from inducing degradation of Aux/IAAs[@Yu_Untethering_2015].
-This lack of complexation may allow observed high ubiquitous AFB1 accumulation [@Parry_Complex_2009].
-Higher order mutants in the family containing *afb1* mutants suggest that *AFB1* has a moderate positive effect on auxin signaling.
-Additionally, AFB4 and AFB5 have been shown to preferentially and functionally bind the synthetic auxin picloram, while other family members preferentially bind indole-3-acetic acid [@Prigge_Arabidopsis_2016].
-Interestingly, the strength and rate with which TIR1/AFBs are able to bind and mark Aux/IAAs for degradation are variable [@calderonvillalobos_combinatorial_2012; @Havens_Synthetic_2012].
-AFB2 induced the degradation of certain Aux/IAA proteins at a faster rate than TIR1, suggesting some functional specificity has arisen since the initial duplication between the *TIR1/AFB1* and *AFB2/AFB3* clades.
-
-```{r TIR1_AFB_div_stats}
-gene_info <- geneInfoFromFile(
-  fname = system.file("shiny-app", "gene_family_data",
-                      "AFB_gene_ids.csv", package = "r1001genomes"))
-annotation <- readAnnotationFile(filename = system.file("shiny-app", "gene_family_data",
-                      "AFB_annotations.csv", package = "r1001genomes"), gene_info = gene_info)
-alignment <- alignCDS(gene_info$transcript_ID)
-# make an alignment data frame
-aln_df <- makeAlnDF(alignment[[2]])
-
-# make a collapsed alignment for translating positions
-collapsed_aln_df <- 
-  aln_df %>% dplyr::select(aln_pos, seq_pos, seq_name) %>%
-  tidyr::spread(key = seq_name, value = seq_pos)
-
-VCF <- readRDS(system.file("shiny-app", "gene_family_data",
-                      "AFBs_VCF.rds", package = "r1001genomes"))
-div_stats <- ldply(VCF, diversityStats, geneInfo=gene_info, .id="transcript_ID")
-div_stats$tair_symbol %<>% fct_reorder(.x =  div_stats$tair_symbol, 
-                                       .fun = function(x) 
-                                         {as.integer(str_remove(x, 
-                                              pattern = "AFB|TIR"))})
-div_stats$tair_symbol %<>% fct_relevel("TIR1")
-div_stats %<>% arrange(tair_symbol)
-div_stats %<>% select(tair_symbol, everything())
-
-if (params$output == "html_document") {
-  formatRound(datatable(div_stats,
-                  colnames = c("transcript",
-                               "symbol",
-                               "&pi;<sub>N</sub>",
-                               "&pi;<sub>S</sub>",
-                               "&pi;<sub>N</sub>/&pi;<sub>S</sub>",
-                               "&pi; coding",
-                               "&pi; transcript"),
-                  rownames = FALSE, escape = FALSE,
-                  options = list(paging=FALSE, searching=FALSE)),
-                columns = 2:7, digits = 6)
-  } else 
-{
-  dt <- div_stats
-  names(dt) <- c("transcript",
-                                 "symbol",
-                                 "&pi;~N~",
-                                 "&pi;~S~",
-                                 "&pi;~N~/&pi;~S~",
-                                 "&pi; coding",
-                                 "&pi; transcript")
-  dt %<>% mutate_if(is.numeric, round, 6)
-  names(dt) <- paste0("**", names(dt), "**")
-  kable(dt) 
-}
-```
-
-Examining the natural sequence variation across the *AFB* family revealed that *TIR1* and *AFB1* both had very low nonsynonymous diversity, hinting at their likely functional importance and bringing in to question the inconclusive role of *AFB1* in auxin signaling.
-*AFB3* and *AFB4* had higher nonsynonymous diversity, while their sister genes, *AFB2* and *AFB5* were more conserved.
-This matches our current understanding of *AFB3* as playing a minor role in the auxin signaling pathway and *AFB4* perhaps undergoing pseudogenization.
-
-Although most known functional regions are highly conserved in *AFB1*, there is a nonsynonymous polymorphism  in the oligomerization domain, only found in the Can-0 accession.
-Mutations in this domain of *TIR1* frequently have a semidominant effect on root phenotypes [@dezfulian_oligomerization_2016; @wright_insights_2017a].
-Characterization of this accession may identify potential *AFB1* phenotypes and may also test the hypothesis that AFB1 functions through oligomerization with other TIR1/AFBs.
-
-```{r AFB variants for PyMOL}
-coding_vcf <- ldply(.data = VCF, .fun = subset, 
-             !is.na(Transcript_ID) & gt_GT != "0|0")
-aln_df <- addSNPsToAlnDF(aln_df, coding_vcf, seq_name = Transcript_ID, 
-               seq_pos = Codon_Number)
-collapsed_aln_df %<>% left_join(y = subset(aln_df, seq_name == gene_info[gene_info$tair_symbol == "TIR1", "transcript_ID"]))
-
-#' Select variants for importing into PyMOL
-#'
-#' @param gene_info a gene info data frame
-#' @param aln_df an alignment data frame with attached SNPs
-#' @param structure_ref string, which tair_symbol or transcript_ID (depending on 
-#' \code{name_by_symbols}) corresponds to the protein structure to which 
-#' selections should be mapped
-#' @param effect_type string of the effect type to select from the alignment 
-#' data frame
-#' @param name_by_symbols logical, if true (default) tair_symbols will be used 
-#' to name the selections, otherwise transcript_IDs will be used
-#'
-#' @return
-#' @export
-#'
-#' @examples
-select_variants <- function(gene_info, aln_df, structure_ref, 
-                            effect_type = "missense", name_by_symbols = TRUE){
-  collapsed_aln_df <- aln_df %>% dplyr::select(aln_pos, seq_pos, seq_name) %>%
-    tidyr::spread(key = seq_name, value = seq_pos)
-  if(name_by_symbols)
-    column <- "tair_symbol"
-    else 
-      column <- "transcript_ID"
-
-  selections <- ldply(.data = gene_info[,column], .fun = function(gene){
-    collapsed_aln_df %<>% dplyr::left_join(y = subset(aln_df, seq_name ==  
-                      gene_info[gene_info[, column] == gene, "transcript_ID"]))
-    variants <- collapsed_aln_df[
-      grepl(effect_type, collapsed_aln_df$effects), 
-      gene_info[gene_info[, column] == structure_ref, "transcript_ID"]]
-    c(gene, paste0(na.omit(as.numeric(variants[[1]])), collapse = "+"))
-  })
-  return(selections)
-}
-
-selections <- select_variants(gene_info, aln_df, structure_ref = "TIR1")
-selections$V2 <- paste0("chain B and resi ", selections$V2)
-write.table(selections, file = "AFB_variant_selections", row.names = FALSE, 
-          col.names = FALSE, sep = ",", quote = FALSE)
-```
-
-```{python readSelection, eval=FALSE}
-#
-# -- readSelection.py
-#
-from pymol import cmd, stored
-
-def readSelection (fileName):
-	'''
-	DESCRIPTION
-
-	reads a csv file containing a column of selection names and a column of
-	selection strings and creates selection objects for each
-	'''
-	#read in the selection file
-	with open(fileName) as f:
-		lines = f.readlines()
-		for line in lines:
-			sel = line.split(",")
-			cmd.select(sel[0], sel[1])
-
-	f.close()
-
-cmd.extend('readSelection',readSelection);
-```
-
-```{PyMol, AFB_structure, eval=FALSE}
-#from within pymol
-cmd.load("Dropbox/180705_NIH_renewal/Aim1-figures/2P1Q.pdb")
-hide everything
-
-run ~/Desktop/natural-variation-webtool/manuscript/readSelection.py
-
-readSelection ~/Desktop/natural-variation-webtool/manuscript/AFB_variant_selections
-
-# Define selections
-# select TIR1variants, (chain B and resi 154+239+546)
-select fbox, (chain B and resi 1-50)
-select LRR, (chain B and resi 50-576)
-select TIR1, (chain B and resi 1-576)
-# select AFB2variants, (chain B and resi 38+174+181+184+209+259+401+496)
-select tir1_1, (chain B and resi 147)
-select TIR1dm, (chain B and resi 170+473)
-select insPcontacts, (chain B and resi 74+78+113+114+344+401+403+436+460+484+485+509)
-select dimerNULL, (chain B and resi 142+143+151+117+147)
-select auxin, resn IAC
-select AUXcontacts, (chain B and resi 78+79+82+378+380+403+404+405+438+439+462+463+464+489)
-select insP, resn IHP
-select IAA7, chain C
-select IAA7contacts, (chain B and resi 81+82+84+139+347+349+350+351+352+380+405+406+407+409+464+465+489)
-select IAA7_N, chain C and resi 1
-select IAA7_C, chain C and resi 13
-select ASK1, chain A
-select Y, chain C and resi 11
-select Ycontacts, chain B and resi 165+166+195
-
-#Define color palette
-#set_color name, [ red-float, green-float, blue-float ]
-set_color blue1, [0.00000000, 0.20392157, 0.6980392]
-set_color blue2, [0.09803922, 0.36078431, 1.0000000]
-set_color blue3, [0.00000000, 0.29019608, 1.0000000]
-set_color teal,  [0.09803922, 1.00000000, 0.6980392]
-set_color LRR,   [1.00000000, 0.71764706, 0.0000000]
-set_color Fbox,  [0.69803922, 0.49411765, 0.0000000]
-set_color TIR1l, [0.67058824, 0.07843137, 0.8000000]
-set_color TIR1f, [0.63529412, 0.34117647, 0.6980392]
-set_color AFB2l, [0.37647059, 0.70196078, 1.0000000]
-set_color AFB2f, [0.07843137, 0.64313725, 0.8000000]
-
-show cartoon, 2P1Q
-
-color lrr, LRR
-
-color TIR1l, TIR1
-show sticks, TIR1
-color TIR1f, AFB1
-show sticks, AFB1
-color AFB2l, AFB2
-show sticks, AFB2
-color AFB2f, AFB3
-show sticks, AFB3
-
-set ray_trace_mode, 3
-
-# used get_view() to return the current view (position) of the molecule 
-# copy the code to be able to set the identical viewpoint later
-# to return the current view from Pymol
-# get_view
-
-set_view (\
-     0.044671144,    0.083668359,   -0.995409012,\
-     0.774909019,    0.625905454,    0.087390237,\
-     0.630409002,   -0.775298953,   -0.036878210,\
-    -0.001916260,    0.005584979, -202.498916626,\
-    -5.538011074,   -8.507736206,   39.046234131,\
-  -219.237945557,  624.123413086,  -20.000000000 )
-
-#to render a high definition image use ray
-ray
-
-```
-
-```{r AFB1 diversity plot}
-coding_variants <- getCodingDiv(VCF[[gene_info[gene_info$tair_symbol == "AFB1", "transcript_ID"]]])
-p <- plotCodingDiv(uniqueCodingVars = coding_variants)
-annotation <- addAlnPosToAnno(anno_df = annotation, aln_df)
-
-# translate positions from TIR1 to AFB1
-AFB1_positions <- annotation$positions
-AFB1_positions %<>% left_join(y = select(collapsed_aln_df, aln_pos, gene_info[gene_info$tair_symbol == "AFB1", "transcript_ID"]), by = "aln_pos") 
-AFB1_positions$position <- AFB1_positions$AT4G03190.1
-
-p <- append_layers(p,list(
-        geom_rect(data = AFB1_positions,
-                   mapping = aes(xmin = as.integer(position)-1,
-                                 xmax = as.integer(position)+2,
-                                 fill = annotation),
-                  ymin = -Inf, ymax = Inf, inherit.aes = FALSE, alpha = 0.8)),
-      position = "bottom")
-p
-```
-
-The AFB4 and AFB5 receptors have an N-terminal extension prior to the F-box domains.
-This extension had very high nonsynonymous diversity(Fig.
-4), suggesting that this extension does not play an important functional role in these proteins.
-Additionally, two frameshift variants and one stop-gained variant were observed in *AFB4* supporting its pseudogenization.
-
-Figure 4: Alignment of A.
-thaliana TIR1/AFB F-box domains showing variants.
-Cyan marks the F-box binding domain.
-Yellow, green, and magenta, mark missense and synonymous, synonymous only, and missense only variant locations respectively.
-Yellow squared E’s show Cul1-interacting residues, differing in AFB1.
-```{r TIR1/AFB F-box Alignment}
-aln_df <- makeAlnDF(alignment[[2]])
-aln_df <- left_join(aln_df, dplyr::select(gene_info, "tair_locus",
-                                       "tair_symbol", "transcript_ID"),
-                        by = c("transcript_ID" = "transcript_ID"))
-aln_df$tair_symbol <- fct_rev(fct_relevel(aln_df$tair_symbol, "TIR1"))
-aln_df <- chunkAlnDF(aln_df = aln_df, chunk_width = 55)
-VCF <- ldply(.data = VCF, .fun = subset, gt_GT != "0|0")
-VCF <- getCodingDiv(VCF)
-aln_df <- addSNPsToAlnDF(aln_df, VCF)
-F_box <- subset(aln_df, aln_pos <= 96)
-F_box <- chunkAlnDF(aln_df = F_box, chunk_width = 80)
-p <-ggplot(F_box, aes(x = aln_pos, y = tair_symbol,
-                             group = seq_pos, text = variants))
-p <- p +
-  geom_tile(data = na.omit(F_box), mapping = aes(fill = effects),
-            width = 1, height = 1, alpha = 0.8) +
-  geom_text(aes(label=letter), alpha= 1, family = "Courier") +
-  scale_fill_brewer(type = "qual", palette = 1, direction = -1) +
-  scale_x_continuous(breaks=seq(1,max(F_box$aln_pos), by = 10)) +
-  scale_y_discrete() +
-  # expand increases distance from axis
-  labs(x = "", y = "", fill = "variant") +
-  theme_logo(base_family = "Helvetica") +
-  theme(panel.grid = element_blank(), panel.grid.minor = element_blank()) +
-  facet_wrap(facets = ~chunk, ncol = 1, scales = "free") +
-  theme(strip.background = element_blank(),
-        strip.text.x = element_blank(),
-        legend.box = "vertical")
-p
-```
-
-## *Aux/IAA* genes
+# *Aux/IAA* genes {#IAAs}
 
@@
 ```
-
-## *TPL/TPR* genes
-
-The Auxin signaling pathway utilizes the TOPLESS (TPL) and TOPLESS-related (TPR) family of Gro/TLE/TUP1 type co-repressor proteins to maintain auxin responsive genes in a transcriptionally-repressed state in the absence of auxin [@szemenyei_topless_2008].
-In *Arabidopsis thaliana* the five member *TPL/TPR* family includes *TPL* and *TPR1-4*.
-The resulting proteins are comprised of three structural domains: an N-terminal TPL domain and two WD-40 domains [@long_topless_2006].
-TPL/TPR proteins are recruited to the AUX/IAA proteins through interaction with the conserved Ethylene-responsive element binding factor-associated amphiphilic repression (EAR) domain [@szemenyei_topless_2008].
-Canonical EAR domains have the amino acid sequence LxLxL, as found in most AUX/IAAs (figure xA) [@overvoorde_functional_2005].
-TPL/TPR co-repressors bind EAR domains via their C-terminal to LisH (CTLH) domains found near their N-termini (citations of pre-structure founding papers/reviews).
-Recent structural analyses of the TPL N-terminal domain have highlighted the precise interaction interface between TPL and AUX/IAA EAR domains, as well as the TPL-TPL dimerization and tetramerization motifs [@martin-arevalillo_structure_2017; @ke_structural_2015a].
-The residues required for higher-order multimers of TPL tetramers have also been identified [@ma_d53_2017].
-Additional interactions with transcriptional regulation and chromatin modifying machinery are likely mediated by two tandem beta propeller domains of TPL/TPRs.
-These domain have not yet been crystallized, however numerous high resolution crystal structures of beta propeller domains allow for structural prediction of the TPL C-terminal beta propeller domains and the locations of likely interaction faces (Figure, X).
-
-The TOPLESS co-repressor family generally exhibits a high level of sequence conservation at the amino acid sequence level across resequenced *Arabidopsis thaliana* accessions, with all ${\pi_N/\pi_S}$ values below 1 (Table X).
-The closely related *TPL* and *TPR1* have the highest ${\pi_N/\pi_S}$ values (TPL-0.425, TPR1-0.739), suggesting that these these two related genes tolerate a higher degree of sequence and potentially functional diversity compared to *TPR2/3/4*.
-The N-terminal TPL domain of the TPL/TPR family is particularly conserved (Figure X).
-All nonsynonymous polymorphisms observed in this region are either in the coils between helices or are highly conservative mutations within helices (i.e.
-Valine to Isoleucine), which would be predicted to exhibit little effect on folding and function.
-
-```{r TPL SNP stats, fig.height=11}
-gene_info <- geneInfoFromFile(
-  fname = system.file("shiny-app", "gene_family_data",
-                      "TPL_gene_ids.csv", package = "r1001genomes"))
-#annotation <- readAnnotationFile(filename = system.file("shiny-app", "gene_family_data",
-#                      "IAA_annotations.csv", package = "r1001genomes"), gene_info = gene_info)
-
-VCF <- readRDS(system.file("shiny-app", "gene_family_data",
-                      "TPL_VCF.rds", package = "r1001genomes"))
-div_stats <- ldply(VCF, diversityStats, geneInfo=gene_info, .id="transcript_ID")
-div_stats$tair_symbol %<>% fct_reorder(.x = , div_stats$tair_symbol, 
-                                       .fun = function(x) 
-                                         {as.integer(str_remove(x, 
-                                              pattern = "TPR"))})
-div_stats$tair_symbol %<>% fct_relevel("TPL")
-div_stats %<>% arrange(tair_symbol)
-div_stats %<>% select(tair_symbol, everything())
-
-if (params$output == "html_document") {
-  formatRound(datatable(div_stats,
-                  colnames = c("transcript",
-                               "symbol",
-                               "&pi;<sub>N</sub>",
-                               "&pi;<sub>S</sub>",
-                               "&pi;<sub>N</sub>/&pi;<sub>S</sub>",
-                               "&pi; coding",
-                               "&pi; transcript"),
-                  rownames = FALSE, escape = FALSE,
-                  options = list(paging=FALSE, searching=FALSE)),
-                columns = 2:7, digits = 6)
-  } else 
-{
-  dt <- div_stats
-  names(dt) <- c("transcript",
-                                 "symbol",
-                                 "&pi;~N~",
-                                 "&pi;~S~",
-                                 "&pi;~N~/&pi;~S~",
-                                 "&pi; coding",
-                                 "&pi; transcript")
-  dt %<>% mutate_if(is.numeric, round, 6)
-  names(dt) <- paste0("**", names(dt), "**")
-  kable(dt) 
-}
-```
-
-The high degree of conservation in the entire N-terminal domain underscores its importance in TPL/TPR function.
-For example, the initial *tpl-1* mutation (N176H) in the ninth helix is a dominant gain-of-function allele [@long_topless_2006], which is capable of binding wild-type TPL protein and inducing protein aggregation [@ma_d53_2017].
-It is therefore understandable that this helix had very low diversity as nonsynonymous variants in this domain could act in a dominant negative fashion.
-Addtionally, the crystal structure of TPL has revealed that the The LisH domain and CRA domain coordinate TPL dimerization and tetramerization respectively (26601214, 28698367), the CTLH domain is critical for interaction with DNA-binding transcription factors (26601214), and helix nine within the CRA domain may be critical for the formation of higher-order TPL complexes (multimers of TPL tetramers, 28630893).
-
-Compared to the TPL domain, both the proline rich and WD-40 domains exhibited a higher frequency of missense variants.
-We utilized Phyre2.0 modeling software to predict the structure of the TPL/TPR WD-40 repeats in order to map non-synonymous mutations to protein structure (Figure X).
-Here we can see both faces of each predicted beta-propeller structure of the WD-40 domains, which are connected by a short peptide linker.
-This model reveals amino acid conservation in the core and central residues of the WD-40 domains.
-~~These observations could be used as the basis for forward mutagenesis strategies to test whether this region of TPL functions similarly to its homolog TLE, which binds its target genes EH1 and the C-terminal WRPW/Y motif (Hairy/Hes/Runx) in the center of the WD-40 domain (pmid 16762837).~~ **Can we say something more general here?
-There is not enough background to support this.
-What is TLE from?
-What is EH1?
-Is TLE really binding to EH1 DNA?**
-
-```{r TPL_TPR alignment}
-
-```
-
-Despite the presence of five *TPL/TPR* genes in the genome, which are considered to be redundant (pmid 16763149), very few observed SNPs are predicted to result in loss of function for any *TPL/TPR* gene.
-Only one frameshift variant was observed, *TPR4-665delG*, in a single accession (Aiell-4 from Italy).
-This deletion occurs in codon 222, just after the TPL N-terminal domain, and generates an early stop eleven codons downstream from the deletion.
-It is unclear whether this N-terminal fragment would be expressed at any appreciable level, as it is possible that the generation of an early stop codon would trigger nonsense-mediated decay.
-However, if expressed, it would be fascinating to examine whether this naturally occurring N-terminal truncation retains activity as a co-repressor.
-Given the observation that the TPL N-terminal domain is sufficient to bind histones (28630893), and repress transcription via synthetic assays in yeast (24979769), it is possible that this represents a semi-functional TPL variant.
-There has been no evidence for prioritization of individual *TPL/TPR* family members in the auxin response pathway, but experiments modulating the gene copy number of *TPR1* revealed the importance of maintaining the proper dosage of *TPL/TPR* in order to accurately trigger the salicylic acid innate immune response (20647385).
-These results, taken together with the absence of *TPL/TPR* loss-of-function alleles in the *Arabidopsis thaliana* accessions suggest that perhaps the entire family is being conserved at the nucleotide level in the population to meet a gene dosage requirement, with little opportunity for neo- or sub-functionalization.
-
-## ARF genes
-
-Auxin response is ultimately mediated by the auxin responsive transcription factors (ARFs).
-There are 23 ARFs in *Arabidopsis thaliana* that are divided into three phylogenetic classes.
-Class A ARFs (ARF5, ARF6, ARF7, ARF8 and ARF19) activate transcription.
-These ARFs have a glutamine-rich region in the middle of the protein that may mediate activation (Guilefoyle, 2007).
-It has recently been shown that the middle region of ARF5 interacts with the SWI/SNF chromatin remodeling ATPases BRAMA and SPLAYED, possibly to reduce nucleosome occupancy and allow for the recruitment of transcription machinery (Wu, 2016).
-Additionally, ARF7 interacts with Mediator subunits, directly tethering transcriptional activation machinery to its binding sites in the chromosome (Ito, 2016).
-Class B and C ARFs are historically categorized as repressor ARFs, though the mechanism through which they confer repression has not been identified.
-Their middle regions tend to be proline- and serine-rich (citation, Ulmasov?).
-
-Canonical ARFs are comprised of three major domains.
-Recent crystallization of these domains have informed structure-function analysis of the ARFs (Boer, 2014; Nanao, 2014; Korasick, 2014).
-These domains are conserved throughout land plants (Dolf's new paper).
-ARFs share an N-terminal B3 DNA binding domain.
-Flanking this DNA-binding domain is a dimerization domain, which folds up into a single “taco-shaped” domain to allow for dimerization between ARFs.
-There is an auxiliary domain that immediately follows and interacts with the dimerization domain.
-The middle region is the most variable between ARFs, as mentioned above, but is characterized by repetitive units of glutamine (class A), serine, or proline residues (classes B and C).
-
-The C-terminal domain of the ARF is a protein-protein interaction domain mediating interactions among ARFs, between ARFs and other transcription factors, and between ARFs and the Aux/IAA repressors.
-This interaction domain was recently characterized as a Phox and Bem1 (PB1) domain, which is comprised of a positive and negative face with conserved basic and acidic residues, respectively (Nanao, 2014; Korasick, 2014).
-The dipolar nature of the PB1 domain may mediate multimerization by the pairwise interaction of these faces on different proteins as the ARF7 PB1 domain was crystallized as a multimer [@korasick_].
-However, it is unclear whether ARF multimerization occurs or plays a developmental significant role in vivo.
-Interfering with ARF dimerization in either the DNA-binding proximal dimerization domain or the PB1 domain decreases the ability of class A ARFs to activate transcription in a heterologous yeast system (Pierre-Jerome, 2016).
-
-While domain architecture is broadly conserved among the ARFs, there are exceptional cases.
-Three ARFs do not contain a PB1 domain at all, ARF3, ARF13, and ARF17, and several more have lost the conserved acidic or basic residues in the PB1 domain, suggesting they may be reduced to a single interaction domain.
-Several ARFs additionally have an expanded conserved region within the DNA-binding domain, of unknown function.
-The majority of domain variation among ARFs occurs in the large B-class subfamily.
-The liverwort *Marchantia polymorpha* has a single representative ARF of each class (Flores-Sandoval, 2015).
-The expansion of these classes in flowering plants is the result of both whole genome and tandem duplication events (Remington, 2004).
-The growth of the ARF family may have allowed for the expansion of the quantity and complexity of loci regulated by the ARFs and subsequent expansion in their regulation of developmental processes.
-
-Class A ARFs are the most well-studied ARF subfamily---the five family members all act as transcriptional activators and have well-characterized, distinct developmental targets.
-Overall the diversity of class A ARFs was generally low, especially compared to the class B and C ARFs, suggesting that class A ARFs are central to auxin signal transduction and plant development.
-Analysis of class A ARF nonsynonymous diversity suggests that the majority of these ARFs are highly functionally conserved, with ${\pi_N/\pi_S}$ values much lower than 1 with the exception of ARF19, with ${\pi_N/\pi_S}$ value of 1.8.
-Comparing diversity within sister pairs, there is a similar trade-off as seen in most *IAA* sister pairs, with one sister being highly conserved and the other more divergent.
-ARF19 and ARF8 are the more divergent class A ARFs, with ${\pi_N/\pi_S}$ values at least three time those of their sisters, ARF7 and ARF6 respectively.
-This may suggest that ARF6 and ARF7 serve more essential purposes in plant development.
-
-```{r ARF diversity stats}
-gene_info <- geneInfoFromFile(
-  fname = system.file("shiny-app", "gene_family_data",
-                      "ARF_gene_ids.csv", package = "r1001genomes"))
-#annotation <- readAnnotationFile(filename = system.file("shiny-app", 
-  #"gene_family_data", "ARF_annotations.csv", 
-  #package = "r1001genomes"), gene_info = gene_info)
-
-VCF <- readRDS(system.file("shiny-app", "gene_family_data",
-                      "ARF_VCF.rds", package = "r1001genomes"))
-div_stats <- ldply(VCF, diversityStats, geneInfo=gene_info, .id="transcript_ID")
-div_stats$tair_symbol %<>% fct_reorder(.x =  div_stats$tair_symbol, 
-                                       .fun = function(x) 
-                                         {as.integer(str_remove(x, 
-                                              pattern = "ARF"))})
-div_stats %<>% arrange(tair_symbol)
-div_stats %<>% select(tair_symbol, everything())
-
-if (params$output == "html_document") {
-  formatRound(datatable(div_stats,
-                  colnames = c("transcript",
-                               "symbol",
-                               "&pi;<sub>N</sub>",
-                               "&pi;<sub>S</sub>",
-                               "&pi;<sub>N</sub>/&pi;<sub>S</sub>",
-                               "&pi; coding",
-                               "&pi; transcript"),
-                  rownames = FALSE, escape = FALSE,
-                  options = list(paging=FALSE, searching=FALSE)),
-                columns = 2:7, digits = 6)
-  } else 
-{
-  dt <- div_stats
-  names(dt) <- c("transcript",
-                                 "symbol",
-                                 "&pi;~N~",
-                                 "&pi;~S~",
-                                 "&pi;~N~/&pi;~S~",
-                                 "&pi; coding",
-                                 "&pi; transcript")
-  dt %<>% mutate_if(is.numeric, round, 6)
-  names(dt) <- paste0("**", names(dt), "**")
-  kable(dt) 
-}
-```
-
-```{r ARF alignment}
-alignment <- alignCDS(gene_info$transcript_ID)
-# make an alignment data frame
-aln_df <- makeAlnDF(alignment[[2]])
-
-# make a collapsed alignment for translating positions
-collapsed_aln_df <- 
-  aln_df %>% dplyr::select(aln_pos, seq_pos, seq_name) %>%
-  tidyr::spread(key = seq_name, value = seq_pos)
-```
-
-```{r Mask_ARF_alignment, eval=FALSE}
-aliscore(as.AAbin(alignment[[2]]), path = "/Volumes/SSD/Applications/ALISCORE_v2.0/Aliscore_v.2.0/")
-id <- scan(file = "/Volumes/SSD/Applications/ALISCORE_v2.0/Aliscore_v.2.0/input.fas_List_random.txt", sep = " ", quiet = TRUE)
-AAAlign_no_gaps <- as.matrix(alignment[[2]])[, id]
-write.nexus.data(AAAlign_no_gaps, file = "ARFs.nex", format = "protein")
-```
-```{bash ARF_phylogeny_construction, eval=FALSE}
-/Applications/MrBayes/mb
-
-execute ARFs.nex
-# set ARF17 as the outgroup
-outgroup AT1G77850.1
-
-lset nucmodel=protein Rates=invgamma
-# or include Nst=6  ngammacat=4 omegavar=M3 
-prset aamodelpr=fixed(jones) statefreqpr=fixed(empirical) 
-# initial run with aamodelpr=mixed yielded posterior prob of 1 for jones model
-propset ExtTBR$prob=0
-
-mcmcp ngen=250000 nchains=4 printfreq=1000 samplefreq=100 burnin=200
-
-mcmc
-sump
-sumt
-
-```
-
-```{r Load ARF phylogeny and data}
-tree <- treeio::read.beast(file = "ARFs.nex.con.tre")
-
-ARF_data <- read.csv(system.file("shiny-app", "gene_family_data",
-                      "ARF_gene_ids.csv", package = "r1001genomes"))
-ARF_data %<>% left_join(y = div_stats$x$data, by = c("name" = "tair_symbol"))
-#tip_labels need to be in the first column
-ARF_data %<>% select(name, everything())
-ARF_data$tair_locus %<>% as.character()
-
-ARF_data <- ARF_data[match(gsub(x = tree@phylo$tip.label, pattern = "\\.1",replacement = ""), ARF_data$tair_locus), ]
-
-tree@phylo$tip.label <- ARF_data$name
-
-```
-
-```{r Visualize ARF tree, fig.height=5, fig.width=5}
-p <- ggtree(tree, layout = "rectangular", color = "grey") %<+% ARF_data + theme_tree() + scale_x_continuous(expand = c(0, 1.2)) + geom_tiplab(hjust = -0.25) + 
-  geom_tippoint(mapping = aes(size = 4*Pi_NS_Ratio, fill = Pi_NS_Ratio), 
-                shape = 21, color = "white", show.legend = TRUE) + 
-  scale_fill_viridis(values = rescale(c(min(ARF_data$Pi_NS_Ratio), 1,
-                                        max(ARF_data$Pi_NS_Ratio)), 
-                                      to = c(0,1)), guide = "colourbar") + scale_size(guide = "none") + 
-  theme(legend.position = c(.1,.7)) + 
-  geom_nodelab(mapping = aes(label = round(prob,digits = 2)), nudge_x = -.3, nudge_y = .5, size = 3) +
-  geom_tiplab(mapping = aes(color = class), hjust = -0.25) + 
-  scale_color_viridis(discrete = TRUE)
-p
-```
-
-### Class A ARFs 
-For all class A ARFs, the middle region of the protein was the predominant high diversity region.
-The DNA-binding domain had very few, low-diversity missense mutations, as did the PB1 domain.
-Considering the necessity of their conserved functions, the low level of variation in these key DNA and protein-protein interaction domains is expected.
-
-```{r A ARF Alignment}
-gene_info <- subset(gene_info, tair_symbol %in% c("ARF5", "ARF6", "ARF7", "ARF8", "ARF19"))
-alignment <- alignCDS(gene_info$transcript_ID)
-# make an alignment data frame
-aln_df <- makeAlnDF(alignment[[2]])
-#BrowseSeqs(alignment[[2]])
-aln_df <- makeAlnDF(alignment[[2]])
-aln_df <- left_join(aln_df, dplyr::select(gene_info, "tair_locus",
-                                       "tair_symbol", "transcript_ID"),
-                        by = c("transcript_ID" = "transcript_ID"))
-aln_df$tair_symbol %<>% as.factor()
-aln_df$tair_symbol %<>% fct_reorder(
-    {as.integer(str_remove(aln_df$tair_symbol, 
-                           pattern = "ARF"))}, 
-  .desc = TRUE)
-
-aln_df <- chunkAlnDF(aln_df = aln_df, chunk_width = 100)
-VCF <- ldply(.data = VCF, .fun = subset, gt_GT != "0|0")
-VCF <- getCodingDiv(VCF)
-aln_df <- addSNPsToAlnDF(aln_df, VCF)
-A_ARFs <- subset(aln_df, tair_symbol %in% c("ARF5", "ARF6", "ARF7", "ARF8", "ARF19"))
-p <-ggplot(A_ARFs, aes(x = aln_pos, y = tair_symbol,
-                             group = seq_pos, text = variants))
-p <- p +
-  geom_raster(data = na.omit(A_ARFs), mapping = aes(fill = effects)) +
-  scale_fill_viridis(option = "A", discrete = TRUE) +
-  scale_y_discrete() +
-  # expand increases distance from axis
-  labs(x = "", y = "", fill = "variant") +
-  theme_logo(base_family = "Helvetica") +
-  theme(panel.grid = element_blank(), panel.grid.minor = element_blank()) +
-  theme(strip.background = element_blank(),
-        strip.text.x = element_blank(),
-        legend.box = "vertical")
-p
-#ggsave(p,filename = "A_ARFs_alignment_variants.pdf", width = 12, height = 4)
-```
-
-Class A ARFs all contain the canonical B3 DNA-binding domain, PB1 interaction domain and glutamine-rich middle region (Guilefoyle, 2007).
-In the analyzed natural variation, ARF7 had several expansions of polyglutamine sequences in the middle region.
-Polyglutamine regions are known to readily expand and contract throughout evolutionary time due to replication error, and variation in polyglutamine length can be acted on by natural selection and have phenotypic consequences (Press, 2014).
-
-A few missense variants in ARF5 highlighted potential functional variation in the class A ARFs.
-A high frequency allele (*What are the variants?*), resulting in substitutions of small amino acids on the positive face of the PB1 domain for bulky phenylalanines, may disrupt interactions with other ARF or Aux/IAA PB1 domains.
-Interestingly, these variants were always found together in accessions Tu-B2-3, Ru-2, Haes-1, Rd-0, 627ME-13Y1, Le-0, Bu-0 and Bsch-0 *Where are these?*.
-Another highly represented polymorphism in ARF5 results in change in the 17th amino acid from a leucine to a methionine.
-This position is in the N-terminal extension of ARF5 which is of unknown function.
-It is possible that this methionine acts as an alternative start site in the accessions.
-
-### Class B ARFs
-
-*ARF23*, which has a truncated DNA-binding domain, has a high ${\pi_N/\pi_S}$ value of 4.1, and has many high-diversity nonsense variants, as may be expected of a gene that is pseudogenizing.
-
-ARF3, ARF13 and ARF17 all lack a C-terminal PB1 domain, which suggests they may be unable to interact with Aux/IAA repressors, possibly acting as auxin-insensitive competitive inhibitors of ARF-mediated transcription (Fig ARF3_13_17 alignment).
-These ARFs span the range of ${\pi_N/\pi_S}$ ratios relative to the other ARFs, ranging from 0.85 to 4.1.
-ARF17 is more conserved than the activator ARF, ARF19.
-ARF13 has many nonsense variants of high diversity and may be pseudogenizing.
-
-```{r ARF3_13_17 alignment}
-
-```
-
-ARF14, ARF15, ARF16, ARF20, and ARF21 all do not have one or multiple of the conserved acidic residues in the PB1 domain, and themselves form a subclade of the B ARFs.
-Interestingly, no non-synonymous variation is seen in these amino acid residues of these ARFs, suggesting that the amino acids that are in the regions may still be important for the function of these proteins.
-
-### Class C ARFs
-
-Class C ARFs show low nucleotide diversity scores, all ${\pi_N/\pi_S}$ values substantially lower than 1.
-ARF16 was the most conserved, whereas its clade members (ARF10, ARF17) had scores at least four times higher [@remington_contrasting_2004].
-
-Structurally, all three members of Class C ARFs contain a canonical B3 DNA-binding domain, but only ARF10 and ARF16 contain a PB1 domain.
-The DNA binding domains exhibit overall low diversity and presence of largely synonymous changes for all the ARFs.
-Only the ARF10 DNA binding domain contains missense variants (S147T and A157S) (Fig.
-ARF10_DBD Variants distribution).
-**Any insights into whether these may affect function?
-They are quite conservative.** Of the PB1 domain containing class C ARFs, ARF16 exhibits several missense variants which are sporadically distributed, in contrast to the conserved PB1 domain of ARF10 (Fig.
-Arf16_PB_variants distribution).
-
-```{r ARF10_DBD Variants distribution}
-
-```
-
-```{r Arf16_PB_variants distribution}
-
-```
-
-Several members of class B (ARF2, 3 and 4) and C ARFs (ARF10, 16 and 17) are distinguished from other ARFs on the basis of their post transcriptional regulation by trans-acting small interfering RNAs (tasiRNAs) and microRNAs (miRNAs) [@rhoades_prediction_2002; @allen_micrornadirected_2005].
-**(@AMY Arf 6 and 8 targeted by miRNA 167 - check for conservation).** The miRNA target of class C ARFs showed a high degree of conservation (Fig.
-ARF10_16_17_Alignment).
-This is consistent with perturbation studies that have shown drastic developmental phenotype for class C ARFs [@wang_control_2005; @mallory_micrornadirected_2005; @liu_repression_2007].
-For the class B ARFs (2, 3 and 4) the tasiRNA binding motif is also highly conserved suggesting strong regulatory constraints on these genes (Fig ARF2_3_4_Alignment).
-The DNA binding domain of ARF4 exhibits non-synonymous variation in 4 accession and would be interesting to carry out phenotypic characterization of these accessions for developmental growth defects (Fig ARF4_DBD Variants map).
-
-```{r ARF10_16_17_Alignment}
-
-```
-
-```{r ARF2_3_4_Alignment}
-
-```
-
-```{r ARF4_DBD Variants map}
-
-```
-
-#Discussion
-
-Analyzing natural diversity within the gene families constituting the auxin nuclear signaling pathway has confirmed existing knowledge, supported existing hypothses and generated new hypotheses.
-This analysis in particular has confirmed the importance of auxin-induced degradation of the Aux/IAA proteins as evidenced by the strong conservation of the *AFB* genes and the canonical degron region of the *Aux/IAA* genes.
-Despite this relative conservation, this analysis has also identified many accessions containing new alleles of interest in the *AFB* and *Aux/IAA* families.
-This analysis has also
-
-Importance of 
--AFB
--IAA degron
--TPL
--ARF DBD and PB1

### 2018-07-12 manuscript/05-TPLs.Rmd
-Cleaned up manuscript folder -moved to bookdown format -revised all sections

--- 2018-07-10 manuscript/manuscript.Rmd
@@
----
-title: A comprehensive analysis of natural sequence variation within the <i>Arabidopsis
-  thaliana</i> nuclear auxin signaling pathway using an accessible web application
-author:
-  affiliation: University of Washington
-  name: R. Clay Wright
-date: "`r Sys.Date()`"
-params: 
-  output: word_document
-output:
-  word_document:
-    toc: yes
-  html_document:
-    fig_caption: yes
-    toc: yes
-always_allow_html: yes
-fontfamily: Helvetica
-fontsize: 11pt
-keywords: r markdown
-link-citations: yes
-csl: apa.csl
-bibliography:
-- natural_variation.bib
-- TIR1_AFB.bib
-- ARF_C.bib
-- AuxIAA_citations.bib
----
-
-```{r setup, message=FALSE, warning=FALSE}
-knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE, collapse = TRUE, comment = "#>", tidy = TRUE, cache = TRUE, cache.lazy = FALSE, out.width = "90%", fig.height=4, fig.width=7, out.extra='style="margin: auto; display: block; padding-top: 15px;"')
-
-#devtools::install_github("wrightrc/r1001genomes", ref = "auxin-natural-variation")
-library(ape)
-library(r1001genomes)
-library(DT)
-library(tidyverse)
-library(magrittr)
-library(RColorBrewer)
-library(ggpmisc)
-library(ggthemes)
-library(ggseqlogo)
-library(ggtree)
-library(reticulate)
-library(scales)
-library(viridis)
-library(ips)
-library(knitr)
-
-#use_python(python = "/usr/local/bin/python")
-```
-<!--This is a CSS (Cascading Style Sheet).
- It controls how panels and images are formatted.></!-->
-<style>
-  img {
-    display: block;
-    margin: auto;
-    width: 90%;
-    height: auto;
-  }
-</style>
-
-# List of working titles  
-* A comprehensive analysis of natural sequence variation within the <i>Arabidopsis thaliana</i> nuclear auxin signaling pathway  
-* **Add your suggestions**
-
-# Abstract
-
-* Motivating, high level statement  
-* Statement defining the cutting edge of your research area  
-* A statement of the form: "However, it remains unclear ..." that defines what problem you are  
-addressing.  
-* A statement of the form:"Here, we ..."  that concisely states your contribution.  
-* A series of 3?5 statement about your results.  
-* A concluding statement.
-
-# Intro
-
-The first genome sequence of *Arabidopsis thaliana* facilitated rapid advancement of plant biology through molecular genetics.
-Since this initial genome, massive scaling of sequencing technology has allowed a pioneering group to survey the global genomic variation in natural *Arabidopsis thaliana* populations.
-This valuable population genetics resource has led to several associations of genetic loci with phenotypic traits and provided insights into how selective pressure has and is influencing the evolution of plant genomes.
-Outside of genome-wide studies however, this valuable dataset has seen little use.
-Natural genomic variation also provides a set of feasible functional variation at the gene, gene family and gene network level that may provide insight into the function and evolution of genes, families and networks.
-Here, we present a web application and R-package through which plant molecular biologists with little-to-no bioinformatics experience can make use of this rich dataset of genetic variation to formulate hypotheses as to the sequence/function/phenotype relationships determined by the gene, family, or network of their interest.
-We demonstrate the utility of this tool through comprehensive analysis and identification of potential functional variation in the nuclear auxin signaling pathway.
-
-This application allows easy access to 
-1. lists of missense polymorphisms to facilitate biochemical assays of variant effects [@starita_variant_2017],
-2. family-wise alignments of variants to facilitate de novo functional domain identification [@melamed_combining_2015], 
-3.
-lists of accessions containing missense (or any type of) polymorphisms to facilitate segregation analysis and measurment of the distribution of variant effects on phenotype [@park_distributions_2017],
-4. as well as the study of gene and network evolution [@delker_natural_2010; @kliebenstein_role_2008].
-
-Here, we have utilized this tool to analyze the genetic variation in the nuclear auxin signaling pathway and formulate hypotheses regarding the functional implications of this variation and the evolution of the genes and gene families in this pathway.
-Our analysis also provided further confirmation of much of the existing knowledge of these genes demonstrating the validity of this approach.
-
-# Methods
-
-## Nucleotide Diversity Calculation
-Nei  defined the nucleotide diversity statistic in his original paper as:
-“…the average number of nucleotide differences per site between two randomly chosen DNA sequences.” (Nei, 1979)
-And provided the equation:
-
-$\displaystyle \pi = \sum_{ij}x_{i}x_{j}\pi_{ij}$
-
-Where Where $x_{i}$ is the frequency of the ith sequence in the population and $\pi_{ij}$ is the number of sites that are different between the ith and jth sequence divided by sequence length.
-
-a more general form, that treats each sequence in the population as unique can be written as:
-
-$\displaystyle \pi = \frac{1}{L*n^2} \sum_{i=1}^n \sum_{j=1}^n \sum_{k=1}^L{\pi_{ijk}}
-\qquad
-\displaystyle \pi_{ijk} = \left\{\begin{array}{l}
-1 \quad \textrm{if} \quad N_{ik} \neq N_{jk} \\
-0 \quad \textrm{if} \quad N_{ik} = N_{jk}
-\end{array}\right.$
-
-where $N_{ik}$ is the nucleotide (A, T, C or G) at position k on the ith sequence of the population.
-$L$ is the length of the sequence (the same for all sequences in the population, 
-assumes only SNPs exist, no indels). n is the total number of sequences in the population.
-
-from this form we can re-arrange summations to the form below:
-
-$\displaystyle \pi = \frac{1}{L} \sum_{k=1}^L{\pi_{k}}
-\qquad
-\pi_{k} = \frac{1}{n^2} \sum_{i=1}^n \sum_{j=1}^n {\pi_{ijk}}$
-
-here $\pi_k$ can be thought of as the site-wise nucleotide diversity at position k, and is equal to 
-the nucleotide diversity of a sequence of length 1 at location k. we can calculate $\pi_k$ for each
-site, then average those over the sequence length to calculate $\pi$, the nucleotide diversity 
-of the sequence.
-
-The function `Nucleotide_diversity` in the r1001genomes package calculates $\pi_k$ for each position in the gene or region that contains a variant.
-note, $\pi_k$ is equal to 0 at all locations without variants.
-This is also what is displayed in the Diversity Plot tab of the web app.
-
-## Detailed $\pi_k$ calculaiton simplification.
-
-The formula for $\pi_k$ above requires comparing every sequence to every other sequence at location k,
-however, we know there are only a few variant forms at each individual location.
-
-so we can revert back to using Nei's original formula, modifying it slightly,
-replacing $x_i$ with $\frac{n_i}{n}$, $n_i$ being the number of sequences in the population of the form $N_i$ at location k: subscripts i and k are summed only over unique forms [need to clarify this]
-
-$\displaystyle \pi_k = \sum_{ij} \frac{n_i}{n}\frac{n_j}{n} \pi_{ij} = 
-\frac{1}{n^2}\sum_{ij} n_i n_j \pi_{ij}$
-
-$\pi_{ij(k)} = \left\{\begin{array}{l}
-1 \quad \textrm{if} \quad i \neq j \\
-0 \quad \textrm{if} \quad i = j
-\end{array}\right.$
-
-we will define $n_{!i} = n - n_i$ as the number of sequences different from i at k.
-
-we can see the summed term will be zero if $i=j$, and $n_i n_j$ if $i\neq j$ so:
-
-$\displaystyle \pi_k = \frac{1}{n^2} \sum_{i} n_i n_{!i}$
-
-next we substitute our definition of $n_{!i}$
-
-$\displaystyle \pi_k = \frac{1}{n^2} \sum_{i} n_i (n - n_{i})$
-
-then distribute and split summation:
-
-$\displaystyle \pi_k = \frac{1}{n^2} (n \sum_{i} n_i - \sum_{i} n_{i}^2)$
-
-finally, summing $\displaystyle \sum_{i}n_i$ is equal to n:
-
-$\displaystyle \pi_k = \frac{1}{n^2} (n^2 - \sum_{i} n_{i}^2)$
-
-this simplified form for $\pi_k$ is used by the app, because the counts of unique 
-variants at a single nucleotide location can easily be summarized in R.
-
-# Results  
-## *TIR1/AFB* genes
-
-Auxin acts by binding to receptors that in turn target co-repressors for degradation.
-Auxin receptors (six in the model plant *Arabidopsis thaliana*) evolved through gene duplication and diversification early in the history of vascular plants [@Parry_Complex_2009].
-The rate of co-repressor degradation is determined by the identity of both the receptor and co-repressor [@Havens_Synthetic_2012], and this rate sets the pace of lateral root development [@Guseman_Auxin-induced_2015].
-
-Although it is unclear what unique roles each receptor plays in growth and development, a number of studies have pointed out differences in the ways the six different receptors in A.
-thaliana differ in biochemical function and expression domain[@dharmasiri_plant_2005; @parry_complex_2009; @prigge_arabidopsis_2016].
-Although TIR1/AFBs are expressed ubiquitously in A.
-thaliana tissues, TIR1, AFB2, and AFB3 have been shown to accumulate in the shoot and root meristems and leaf tissues, with slightly different expression patterns for TIR1 [@Parry_Complex_2009].
-Additionally, the expression of AFB5 is stronly circadian-regulated [@Covington_Circadian_2007] and AFB3 is more highly expressed in the roots in the presence of nitrate, allowing increased lateral root formation [@Vidal_Nitrate_2010], suggesting more broad environmental regulation of this gene family may exist.
-  Functionally, all members of this family have been shown to bind auxin and Aux/IAA proteins.
-However, AFB1 has drastically reduced ability to assemble into an SCF complex, due to the substitution E8K in its F-box domain, preventing it from inducing degradation of Aux/IAAs[@Yu_Untethering_2015].
-This lack of complexation may allow observed high ubiquitous AFB1 accumulation [@Parry_Complex_2009].
-Higher order mutants in the family containing *afb1* mutants suggest that *AFB1* has a moderate positive effect on auxin signaling.
-Additionally, AFB4 and AFB5 have been shown to preferentially and functionally bind the synthetic auxin picloram, while other family members preferentially bind indole-3-acetic acid [@Prigge_Arabidopsis_2016].
-Interestingly, the strength and rate with which TIR1/AFBs are able to bind and mark Aux/IAAs for degradation are variable [@calderonvillalobos_combinatorial_2012; @Havens_Synthetic_2012].
-AFB2 induced the degradation of certain Aux/IAA proteins at a faster rate than TIR1, suggesting some functional specificity has arisen since the initial duplication between the *TIR1/AFB1* and *AFB2/AFB3* clades.
-
-```{r TIR1_AFB_div_stats}
-gene_info <- geneInfoFromFile(
-  fname = system.file("shiny-app", "gene_family_data",
-                      "AFB_gene_ids.csv", package = "r1001genomes"))
-annotation <- readAnnotationFile(filename = system.file("shiny-app", "gene_family_data",
-                      "AFB_annotations.csv", package = "r1001genomes"), gene_info = gene_info)
-alignment <- alignCDS(gene_info$transcript_ID)
-# make an alignment data frame
-aln_df <- makeAlnDF(alignment[[2]])
-
-# make a collapsed alignment for translating positions
-collapsed_aln_df <- 
-  aln_df %>% dplyr::select(aln_pos, seq_pos, seq_name) %>%
-  tidyr::spread(key = seq_name, value = seq_pos)
-
-VCF <- readRDS(system.file("shiny-app", "gene_family_data",
-                      "AFBs_VCF.rds", package = "r1001genomes"))
-div_stats <- ldply(VCF, diversityStats, geneInfo=gene_info, .id="transcript_ID")
-div_stats$tair_symbol %<>% fct_reorder(.x =  div_stats$tair_symbol, 
-                                       .fun = function(x) 
-                                         {as.integer(str_remove(x, 
-                                              pattern = "AFB|TIR"))})
-div_stats$tair_symbol %<>% fct_relevel("TIR1")
-div_stats %<>% arrange(tair_symbol)
-div_stats %<>% select(tair_symbol, everything())
-
-if (params$output == "html_document") {
-  formatRound(datatable(div_stats,
-                  colnames = c("transcript",
-                               "symbol",
-                               "&pi;<sub>N</sub>",
-                               "&pi;<sub>S</sub>",
-                               "&pi;<sub>N</sub>/&pi;<sub>S</sub>",
-                               "&pi; coding",
-                               "&pi; transcript"),
-                  rownames = FALSE, escape = FALSE,
-                  options = list(paging=FALSE, searching=FALSE)),
-                columns = 2:7, digits = 6)
-  } else 
-{
-  dt <- div_stats
-  names(dt) <- c("transcript",
-                                 "symbol",
-                                 "&pi;~N~",
-                                 "&pi;~S~",
-                                 "&pi;~N~/&pi;~S~",
-                                 "&pi; coding",
-                                 "&pi; transcript")
-  dt %<>% mutate_if(is.numeric, round, 6)
-  names(dt) <- paste0("**", names(dt), "**")
-  kable(dt) 
-}
-```
-
-Examining the natural sequence variation across the *AFB* family revealed that *TIR1* and *AFB1* both had very low nonsynonymous diversity, hinting at their likely functional importance and bringing in to question the inconclusive role of *AFB1* in auxin signaling.
-*AFB3* and *AFB4* had higher nonsynonymous diversity, while their sister genes, *AFB2* and *AFB5* were more conserved.
-This matches our current understanding of *AFB3* as playing a minor role in the auxin signaling pathway and *AFB4* perhaps undergoing pseudogenization.
-
-Although most known functional regions are highly conserved in *AFB1*, there is a nonsynonymous polymorphism  in the oligomerization domain, only found in the Can-0 accession.
-Mutations in this domain of *TIR1* frequently have a semidominant effect on root phenotypes [@dezfulian_oligomerization_2016; @wright_insights_2017a].
-Characterization of this accession may identify potential *AFB1* phenotypes and may also test the hypothesis that AFB1 functions through oligomerization with other TIR1/AFBs.
-
-```{r AFB variants for PyMOL}
-coding_vcf <- ldply(.data = VCF, .fun = subset, 
-             !is.na(Transcript_ID) & gt_GT != "0|0")
-aln_df <- addSNPsToAlnDF(aln_df, coding_vcf, seq_name = Transcript_ID, 
-               seq_pos = Codon_Number)
-collapsed_aln_df %<>% left_join(y = subset(aln_df, seq_name == gene_info[gene_info$tair_symbol == "TIR1", "transcript_ID"]))
-
-#' Select variants for importing into PyMOL
-#'
-#' @param gene_info a gene info data frame
-#' @param aln_df an alignment data frame with attached SNPs
-#' @param structure_ref string, which tair_symbol or transcript_ID (depending on 
-#' \code{name_by_symbols}) corresponds to the protein structure to which 
-#' selections should be mapped
-#' @param effect_type string of the effect type to select from the alignment 
-#' data frame
-#' @param name_by_symbols logical, if true (default) tair_symbols will be used 
-#' to name the selections, otherwise transcript_IDs will be used
-#'
-#' @return
-#' @export
-#'
-#' @examples
-select_variants <- function(gene_info, aln_df, structure_ref, 
-                            effect_type = "missense", name_by_symbols = TRUE){
-  collapsed_aln_df <- aln_df %>% dplyr::select(aln_pos, seq_pos, seq_name) %>%
-    tidyr::spread(key = seq_name, value = seq_pos)
-  if(name_by_symbols)
-    column <- "tair_symbol"
-    else 
-      column <- "transcript_ID"
-
-  selections <- ldply(.data = gene_info[,column], .fun = function(gene){
-    collapsed_aln_df %<>% dplyr::left_join(y = subset(aln_df, seq_name ==  
-                      gene_info[gene_info[, column] == gene, "transcript_ID"]))
-    variants <- collapsed_aln_df[
-      grepl(effect_type, collapsed_aln_df$effects), 
-      gene_info[gene_info[, column] == structure_ref, "transcript_ID"]]
-    c(gene, paste0(na.omit(as.numeric(variants[[1]])), collapse = "+"))
-  })
-  return(selections)
-}
-
-selections <- select_variants(gene_info, aln_df, structure_ref = "TIR1")
-selections$V2 <- paste0("chain B and resi ", selections$V2)
-write.table(selections, file = "AFB_variant_selections", row.names = FALSE, 
-          col.names = FALSE, sep = ",", quote = FALSE)
-```
-
-```{python readSelection, eval=FALSE}
-#
-# -- readSelection.py
-#
-from pymol import cmd, stored
-
-def readSelection (fileName):
-	'''
-	DESCRIPTION
-
-	reads a csv file containing a column of selection names and a column of
-	selection strings and creates selection objects for each
-	'''
-	#read in the selection file
-	with open(fileName) as f:
-		lines = f.readlines()
-		for line in lines:
-			sel = line.split(",")
-			cmd.select(sel[0], sel[1])
-
-	f.close()
-
-cmd.extend('readSelection',readSelection);
-```
-
-```{PyMol, AFB_structure, eval=FALSE}
-#from within pymol
-cmd.load("Dropbox/180705_NIH_renewal/Aim1-figures/2P1Q.pdb")
-hide everything
-
-run ~/Desktop/natural-variation-webtool/manuscript/readSelection.py
-
-readSelection ~/Desktop/natural-variation-webtool/manuscript/AFB_variant_selections
-
-# Define selections
-# select TIR1variants, (chain B and resi 154+239+546)
-select fbox, (chain B and resi 1-50)
-select LRR, (chain B and resi 50-576)
-select TIR1, (chain B and resi 1-576)
-# select AFB2variants, (chain B and resi 38+174+181+184+209+259+401+496)
-select tir1_1, (chain B and resi 147)
-select TIR1dm, (chain B and resi 170+473)
-select insPcontacts, (chain B and resi 74+78+113+114+344+401+403+436+460+484+485+509)
-select dimerNULL, (chain B and resi 142+143+151+117+147)
-select auxin, resn IAC
-select AUXcontacts, (chain B and resi 78+79+82+378+380+403+404+405+438+439+462+463+464+489)
-select insP, resn IHP
-select IAA7, chain C
-select IAA7contacts, (chain B and resi 81+82+84+139+347+349+350+351+352+380+405+406+407+409+464+465+489)
-select IAA7_N, chain C and resi 1
-select IAA7_C, chain C and resi 13
-select ASK1, chain A
-select Y, chain C and resi 11
-select Ycontacts, chain B and resi 165+166+195
-
-#Define color palette
-#set_color name, [ red-float, green-float, blue-float ]
-set_color blue1, [0.00000000, 0.20392157, 0.6980392]
-set_color blue2, [0.09803922, 0.36078431, 1.0000000]
-set_color blue3, [0.00000000, 0.29019608, 1.0000000]
-set_color teal,  [0.09803922, 1.00000000, 0.6980392]
-set_color LRR,   [1.00000000, 0.71764706, 0.0000000]
-set_color Fbox,  [0.69803922, 0.49411765, 0.0000000]
-set_color TIR1l, [0.67058824, 0.07843137, 0.8000000]
-set_color TIR1f, [0.63529412, 0.34117647, 0.6980392]
-set_color AFB2l, [0.37647059, 0.70196078, 1.0000000]
-set_color AFB2f, [0.07843137, 0.64313725, 0.8000000]
-
-show cartoon, 2P1Q
-
-color lrr, LRR
-
-color TIR1l, TIR1
-show sticks, TIR1
-color TIR1f, AFB1
-show sticks, AFB1
-color AFB2l, AFB2
-show sticks, AFB2
-color AFB2f, AFB3
-show sticks, AFB3
-
-set ray_trace_mode, 3
-
-# used get_view() to return the current view (position) of the molecule 
-# copy the code to be able to set the identical viewpoint later
-# to return the current view from Pymol
-# get_view
-
-set_view (\
-     0.044671144,    0.083668359,   -0.995409012,\
-     0.774909019,    0.625905454,    0.087390237,\
-     0.630409002,   -0.775298953,   -0.036878210,\
-    -0.001916260,    0.005584979, -202.498916626,\
-    -5.538011074,   -8.507736206,   39.046234131,\
-  -219.237945557,  624.123413086,  -20.000000000 )
-
-#to render a high definition image use ray
-ray
-
-```
-
-```{r AFB1 diversity plot}
-coding_variants <- getCodingDiv(VCF[[gene_info[gene_info$tair_symbol == "AFB1", "transcript_ID"]]])
-p <- plotCodingDiv(uniqueCodingVars = coding_variants)
-annotation <- addAlnPosToAnno(anno_df = annotation, aln_df)
-
-# translate positions from TIR1 to AFB1
-AFB1_positions <- annotation$positions
-AFB1_positions %<>% left_join(y = select(collapsed_aln_df, aln_pos, gene_info[gene_info$tair_symbol == "AFB1", "transcript_ID"]), by = "aln_pos") 
-AFB1_positions$position <- AFB1_positions$AT4G03190.1
-
-p <- append_layers(p,list(
-        geom_rect(data = AFB1_positions,
-                   mapping = aes(xmin = as.integer(position)-1,
-                                 xmax = as.integer(position)+2,
-                                 fill = annotation),
-                  ymin = -Inf, ymax = Inf, inherit.aes = FALSE, alpha = 0.8)),
-      position = "bottom")
-p
-```
-
-The AFB4 and AFB5 receptors have an N-terminal extension prior to the F-box domains.
-This extension had very high nonsynonymous diversity(Fig.
-4), suggesting that this extension does not play an important functional role in these proteins.
-Additionally, two frameshift variants and one stop-gained variant were observed in *AFB4* supporting its pseudogenization.
-
-Figure 4: Alignment of A.
-thaliana TIR1/AFB F-box domains showing variants.
-Cyan marks the F-box binding domain.
-Yellow, green, and magenta, mark missense and synonymous, synonymous only, and missense only variant locations respectively.
-Yellow squared E’s show Cul1-interacting residues, differing in AFB1.
-```{r TIR1/AFB F-box Alignment}
-aln_df <- makeAlnDF(alignment[[2]])
-aln_df <- left_join(aln_df, dplyr::select(gene_info, "tair_locus",
-                                       "tair_symbol", "transcript_ID"),
-                        by = c("transcript_ID" = "transcript_ID"))
-aln_df$tair_symbol <- fct_rev(fct_relevel(aln_df$tair_symbol, "TIR1"))
-aln_df <- chunkAlnDF(aln_df = aln_df, chunk_width = 55)
-VCF <- ldply(.data = VCF, .fun = subset, gt_GT != "0|0")
-VCF <- getCodingDiv(VCF)
-aln_df <- addSNPsToAlnDF(aln_df, VCF)
-F_box <- subset(aln_df, aln_pos <= 96)
-F_box <- chunkAlnDF(aln_df = F_box, chunk_width = 80)
-p <-ggplot(F_box, aes(x = aln_pos, y = tair_symbol,
-                             group = seq_pos, text = variants))
-p <- p +
-  geom_tile(data = na.omit(F_box), mapping = aes(fill = effects),
-            width = 1, height = 1, alpha = 0.8) +
-  geom_text(aes(label=letter), alpha= 1, family = "Courier") +
-  scale_fill_brewer(type = "qual", palette = 1, direction = -1) +
-  scale_x_continuous(breaks=seq(1,max(F_box$aln_pos), by = 10)) +
-  scale_y_discrete() +
-  # expand increases distance from axis
-  labs(x = "", y = "", fill = "variant") +
-  theme_logo(base_family = "Helvetica") +
-  theme(panel.grid = element_blank(), panel.grid.minor = element_blank()) +
-  facet_wrap(facets = ~chunk, ncol = 1, scales = "free") +
-  theme(strip.background = element_blank(),
-        strip.text.x = element_blank(),
-        legend.box = "vertical")
-p
-```
-
-## *Aux/IAA* genes
-
-The Aux/IAA proteins (IAAs) mediate both auxin perception and transcriptional co-repression of auxin responsive genes.
-Auxin perception is carried out along with TIR1/AFB proteins, which together with IAAs act as co-receptors to bind and perceive nuclear auxin.
-Interaction of IAAs with auxin and TIR1/AFBs is conferred by the highly conserved 13 amino acid degron motif [@tan_mechanism_2007].
-Formation of this auxin co-receptor complex triggers ubiquitination and subsequent degradation of IAAs.
-This degradation in turn relieves their repression upon ARF transcription factors.
-Stabilizing mutations within the degron can lead to dominant phenotypes [@berleth_auxin_2004; @yang_iaa1_2004].
-The *Arabidopsis thaliana* Aux/IAA family has 29 members, most of which possess a canonical auxin degron.
-Notable exceptions are the auxin-insensitive IAA20 and IAA30 that lack a degron sequence entirely, while the closely related IAA31 retains a semi-functional degron sequence with a very slow degradation rate [@dreher_arabidopsis_2006; @havens_synthetic_2012].
-Additionally, IAA degradation rate is influenced by rate motifs found outside the degron, such as the KR motif [@dreher_arabidopsis_2006; @moss_rate_2015].
-IAA-mediated repression of ARFs depends upon both an EAR domain, N-terminal of the degron, that recruits co-repressors of the TOPLESS (TPL) family, and a Phox and Bem1 (PB1) domain that enables interaction via hetero-oligomerization with ARF PB1 domains [@nanao_structural_2014; @kagale_ear_2011].
-
-```{r IAA diversity stats, fig.height=11}
-gene_info <- geneInfoFromFile(
-  fname = system.file("shiny-app", "gene_family_data",
-                      "IAA_gene_ids.csv", package = "r1001genomes"))
-#annotation <- readAnnotationFile(filename = system.file("shiny-app", "gene_family_data",
-#                      "IAA_annotations.csv", package = "r1001genomes"), gene_info = gene_info)
-
-VCF <- readRDS(system.file("shiny-app", "gene_family_data",
-                      "IAAs_VCF.rds", package = "r1001genomes"))
-div_stats <- ldply(VCF, diversityStats, geneInfo=gene_info, .id="transcript_ID")
-div_stats$tair_symbol %<>% fct_reorder(.x = , div_stats$tair_symbol, 
-                                       .fun = function(x) 
-                                         {as.integer(str_remove(x, 
-                                              pattern = "IAA"))})
-div_stats %<>% arrange(tair_symbol)
-div_stats %<>% select(tair_symbol, everything())
-
-if (params$output == "html_document") {
-  formatRound(datatable(div_stats,
-                  colnames = c("transcript",
-                               "symbol",
-                               "&pi;<sub>N</sub>",
-                               "&pi;<sub>S</sub>",
-                               "&pi;<sub>N</sub>/&pi;<sub>S</sub>",
-                               "&pi; coding",
-                               "&pi; transcript"),
-                  rownames = FALSE, escape = FALSE,
-                  options = list(paging=FALSE, searching=FALSE)),
-                columns = 2:7, digits = 6)
-  } else 
-{
-  dt <- div_stats
-  names(dt) <- c("transcript",
-                                 "symbol",
-                                 "&pi;~N~",
-                                 "&pi;~S~",
-                                 "&pi;~N~/&pi;~S~",
-                                 "&pi; coding",
-                                 "&pi; transcript")
-  dt %<>% mutate_if(is.numeric, round, 6)
-  names(dt) <- paste0("**", names(dt), "**")
-  kable(dt) 
-}
-```
-
-Within the available natural sequence variation, we found several missense polymorphisms affecting the EAR domain or degron of several canonical IAAs.
-However, no missense polymorphisms were present in the KR motif or residues critical for PB1-PB1 domain interactions (Supp.
-Fig.
-Alignment of canonical IAAs).
-The majority of missense mutations observed were in the variable regions between the conserved domains.
-
-###Sister Pair Diversification
-
-Previous research has found evidence of both broad genetic redundancy and specificity within closely related pairs or groups of IAA proteins [@overvoorde_functional_2005].
-For example, the *iaa8-1 iaa9-1* double mutant and the *iaa5-1 iaa6-1 iaa19-1* triple mutant have wild-type phenotypes [@overvoorde_functional_2005], yet the *IAA6/19* sister pair has significant differences in expression patterns, protein abundances and functions suggesting they have undergone functional specialization since their divergence [@winkler_variation_2017].
-
-In our analysis of the *IAA* gene family, we frequently observed that one member of most *IAA* sister pairs had high nonsynonymous diversity while the other sister was more conserved (Figure IAA phylogeny heat map).
-In an interspecific comparison between *A.
-thaliana* and *A.
-lyrata*, *IAA19* is more conserved, while *IAA6* has regions of increased sequence divergence upstream of the degron and in the N-terminal half of the PB1 domain [@winkler_variation_2017].
-Consistent with these results, our analysis of intraspecific natural variation revealed higher conservation for *IAA19* (${\pi_N/\pi_S}$ = 0.55) compared to *IAA6* (${\pi_N/\pi_S}$ = 2.3) (Figure IAA phylogeny heat map), and also detected high diversity within the same regions of *IAA6* as seen in Winkler et al.
-(Supp Fig IAA6 diversity plot).
-```{r IAA alignment}
-alignment <- alignCDS(gene_info$transcript_ID)
-# make an alignment data frame
-aln_df <- makeAlnDF(alignment[[2]])
-
-# make a collapsed alignment for translating positions
-collapsed_aln_df <- 
-  aln_df %>% dplyr::select(aln_pos, seq_pos, seq_name) %>%
-  tidyr::spread(key = seq_name, value = seq_pos)
-```
-
-```{r Mask_IAA_alignment, eval=FALSE}
-aliscore(as.AAbin(alignment[[2]]), path = "/Volumes/SSD/Applications/ALISCORE_v2.0/Aliscore_v.2.0/")
-id <- scan(file = "input.fas_List_random.txt", sep = " ", quiet = TRUE)
-AAAlign_no_gaps <- as.matrix(alignment[[2]])[, id]
-write.nexus.data(AAAlign_no_gaps, file = "manuscript/IAAs.nex", format = "protein")
-```
-```{bash IAA_phylogeny_construction, eval=FALSE}
-~/Desktop/natural-variation-webtool/manuscript
-
-/Applications/MrBayes/mb
-
-execute IAAs.nex
-# set IAA33 as the outgroup
-outgroup AT5G57420.1
-
-lset nucmodel=protein Rates=invgamma
-# or include Nst=6  ngammacat=4 omegavar=M3 
-prset aamodelpr=fixed(jones) statefreqpr=fixed(empirical) 
-# initial run with aamodelpr=mixed yielded posterior prob of 1 for jones model
-propset ExtTBR$prob=0
-
-mcmcp ngen=250000 nchains=4 printfreq=1000 samplefreq=100 burnin=200
-
-mcmc
-sump
-sumt
-
-```
-
-```{r Load IAA phylogeny and data}
-tree <- treeio::read.beast(file = "IAAs.nex.con.tre")
-
-IAA_data <- read.csv("IAAs_table.csv")
-#tip_labels need to be in the first column
-IAA_data %<>% select(name, everything())
-IAA_data$name %<>% as.character()
-IAA_data$AGI %<>% as.character()
-
-IAA_data <- IAA_data[match(gsub(x = tree@phylo$tip.label, pattern = "\\.1",replacement = ""), IAA_data$AGI), ] 
-tree@phylo$tip.label <- IAA_data$name
-```
-*I think IAA20, 30, and 15 should also be in class C. So there are really two large clades A and B and then there are truncations/loss of function/dominant variants scattered throughout, Class C.*
-
-```{r IAA_tree, fig.height=5, fig.width=5, fig.cap="(ref:IAA tree)"}
-p <- ggtree(tree, layout = "rectangular", color = "grey") %<+% IAA_data + theme_tree() + scale_x_continuous(expand = c(0, 1.2)) + 
-  geom_tippoint(mapping = aes(size = 4*PiN_PiS, fill = PiN_PiS), shape = 21, color = "white", show.legend = TRUE) + 
-  scale_fill_viridis(values = rescale(c(min(IAA_data$PiN_PiS), 1,
-                                        max(IAA_data$PiN_PiS)), 
-                                      to = c(0,1)), guide = "colourbar") + scale_size(guide = "none") + 
-  theme(legend.position = c(.9,.7)) + 
-  geom_nodelab(mapping = aes(label = round(prob,digits = 2)), nudge_x = -.3, nudge_y = .5, size = 3) + 
-  geom_tiplab(mapping = aes(color = Class), hjust = -0.25) + 
-  scale_color_viridis(discrete = TRUE)
-p
-```
-
-(ref:IAA tree) IAA protein sequence tree mapped with ${\pi_N/\pi_S}$
-
-Within the Group B *IAA* genes, there are four sister pairs---*IAA10*/*IAA11*, *IAA12*/*IAA13*, *IAA18*/*IAA26*, and *IAA20*/*IAA30* \@ref(IAA tree) [@remington_contrasting_2004].
-Interestingly, among these sister pairs only *IAA12* and *IAA13* have a noticeable difference in nonsynonymous diversity.
-The *Arabidopsis thaliana* EFP browser shows that **IAA10** and *IAA11* have almost identical expression patterns, with the exception of *IAA10* being more highly expressed in the developing seed, and *IAA11* being more highly expressed in root tissue during lateral root development **need to cite original data source here**.
-In our analysis, *IAA10* and *IAA11* both showed functional conservation (${\pi_N/\pi_S}$ of 0.80 and 0.67 respectively) in support of *IAA10* and *IAA11* playing redundant roles.
-Accessions and associated variants identified here provide a potentially expedited route to elucidating the extent of redundancy between family members such as *IAA10* and *IAA11*.
-
-### New Loss-of-Function/Gain-of-Function alleles
-Accessions with functional variants can be potentially used to investigate the extent of redundancy/specialization within sister pairs.
-For example, three accessions have a missense variant affecting the EAR domain of IAA10 (HR-5, UKSE06-118 and UKSW06-240, all from southern UK).
-This variant (L49S) is in the second leucine of the LxLxL motif and likely reduces the ability of IAA10 to recruit the TOPLESS co-repressor [@lee_biological_2016].
-The nonsense variant (IAA10-Y219\*) resulting in truncated PB1 domain may be more indicative of *IAA10* null phenotype, as it likely prevents IAA10 from interacting with an ARF.
-This variant is only found in one accession from northeastern Spain (IP-Svi-0).
-These accessions and potential loss-of-function alleles may facilitate exploration of the function of *IAA10*.
-
-Several examples of possible gain-of-function missense variants were found in the degron of *IAA2*, *IAA3*, *IAA6*, *IAA7*, *IAA8*, *IAA13*, *IAA15*, *IAA18*, *IAA19*, *IAA27*, *IAA28* and *IAA29*.
-Most degron variants were very minor alleles in the population, only represented in one or two accessions.
-Interestingly, degron missense variants occured more frequently in *IAA* genes that have a closely related sister, which may indicate redundant function within the sister-pairs.
-*I think this is an interesting finding, but I'm not sure about this hypothesis of redundant function.
-If pairs are acting redundantly and we assume the degron variants are dominant, we would expect the auxin signal they transmit (their function) to be lower, in sum, as a pair.
-So I would actually hypothesize that there has been some divergence in function as a pair.
-*
-```{r Correlation between Distance to MRCA and prevalence of degron variants}
-
-```
-
-```{r Correlation between Pi_N/Pi_S and k_5}
-
-```
-
-*IAA6* has two potentially dominant missense mutations found within the degron (Supp Fig.
-alignment), one previously believed to lead to the *shy1-1* phenotype [@kim_two_1996].
-Interestingly, this variant is widely represented in 31 accessions across Europe (Supplemental Map Fig).
-This finding supports the proposition that iaa6-C78R may not be the mutation causing the *shy1-1D* phenotype but rather represents natural variation in the degron sequence ([note from Mike Prigge in TAIR](https://www.arabidopsis.org/servlets/TairObject?type=notepad&id=11247)).
-These accessions could be used to confirm whether *iaa6-C78R* is responsible for the *shy1-1D* phenotype.
-Additionally, an accession was identified with the IAA6-V77M variant, also in the degron, that may be useful to further explore *iaa6* gain-of-function phenotypes.
-
-### Identifying novel functional sequences in IAAs
-An exciting possible use for the Natural Variation Webtool is to search for conserved, low-diversity sites that may represent possible sites of functional import, such as post-translational modification or protein-protein interaction.
-For example, recent detailed biochemical analysis identified putative lysine ubiquitination sites in members of the *IAA6/19* sister pairs [@winkler_variation_2017]; Supp Fig IAA6/19 alignment).
-These sites are conserved across natural sequence variation, while several other lysine residues vary (Supp Figs alignment and map).
-In IAA6, there are 17 lysines of which ~6 are putative ubiquitination sites~ *I think this should be 5, K3, K32, K33, K91, and K97.
-Missense variants occured in K34Q, K140N *.
-One is proximal to the KR motif, while two others show natural variation.
-In IAA19, there are 15 total lysines, 8 of which are putative ubiquitination sites (K3, K25, K68, L87, K93, K100, K111, and K141).
-Two of these ubiquitination sites are proximal to the KR motif and degron.
-**This sentence is not clear.
-Only 1 of the ubiquitination sites varies?
-or one lysine, I think you mean this only 1 lysine that varies, but it is not ubiquitinated.
-Ubiquitylated Missense K129T** Natural sequence diversity may be used to narrow down a subset of lysines to assess for functional ubiquitination.
-
-```{r}
-
-```
-
-```{r IAA variants mapped to IAA17, eval=FALSE}
-coding_vcf <- ldply(.data = VCF, .fun = subset, 
-             !is.na(Transcript_ID) & gt_GT != "0|0")
-aln_df <- addSNPsToAlnDF(aln_df, coding_vcf, seq_name = Transcript_ID, 
-               seq_pos = Codon_Number)
-
-selections <- select_variants(gene_info, aln_df, structure_ref = "IAA17")
-selections$V2 <- paste0("chain X and resi ", selections$V2)
-write.table(selections, file = "IAA_variant_selections", row.names = FALSE, 
-          col.names = FALSE, sep = ",", quote = FALSE)
-```
-
-The members of the *IAA8/9/27* clade have a unique N-terminal extension with a highly conserved region common to orthologs in several other species [@dreher_arabidopsis_2006].
-*IAA8* and *IAA9* play redundant roles in leaf serration [@koenig_auxin_2009], while *IAA8* also plays an important role in lateral root development [@arase_iaa8_2012].
-IAA8 has been reported to uniquely localize in the cytosol [@arase_iaa8_2012] where it may interact with LSD1 protein [@coll_programmed_2011].
-The N-terminal extension has low natural sequence diversity in *IAA8* and *IAA9* (Supp Fig IAA8/9 alignment) that may represent a site of novel functionality, possibly mediating interaction with LSD1.
-The few accessions with missense alleles in this region in both *IAA8* and *IAA9* may be valuable in probing the functionality of this region (Supplemental Fig IAA8/9 map).
-
-```{r IAA8_9_27 alignment}
-
-```
-
-```{r IAA8_9 map}
-
-```
-
-Members of the *IAA17/7/14/16* clade have a second EAR motif (LxLxL) with no observed missense variants in *IAA7*, *IAA16*, or *IAA17* (Supp Fig IAA7/14/16/17 alignment).
-This observation supports the experimental evidence for IAA7 that the second EAR motif is functional, *i.e.* it can interact with TPR1 co-repressor and has a minor repressive role in auxin-related developmental phenotypes [lee_biological_2016].
-For *IAA14*, there were two missense variants that may be of interest to probe functionality of this second EAR motif.
-One variant, L41F, in the central lysine is represented in only one accession, and the other in the second “x” (K40N) is found in twelve accessions around Copenhagen (Supp Fig.
-IAA14 map).
-
-```{r IAA7_14_16_17 alignment}
-
-```
-
-```{r IAA14 map}
-
-```
-
-## *TPL/TPR* genes
+# *TPL/TPR* genes {#TPLs}
 
@@
 These results, taken together with the absence of *TPL/TPR* loss-of-function alleles in the *Arabidopsis thaliana* accessions suggest that perhaps the entire family is being conserved at the nucleotide level in the population to meet a gene dosage requirement, with little opportunity for neo- or sub-functionalization.
-
-## ARF genes
-
-Auxin response is ultimately mediated by the auxin responsive transcription factors (ARFs).
-There are 23 ARFs in *Arabidopsis thaliana* that are divided into three phylogenetic classes.
-Class A ARFs (ARF5, ARF6, ARF7, ARF8 and ARF19) activate transcription.
-These ARFs have a glutamine-rich region in the middle of the protein that may mediate activation (Guilefoyle, 2007).
-It has recently been shown that the middle region of ARF5 interacts with the SWI/SNF chromatin remodeling ATPases BRAMA and SPLAYED, possibly to reduce nucleosome occupancy and allow for the recruitment of transcription machinery (Wu, 2016).
-Additionally, ARF7 interacts with Mediator subunits, directly tethering transcriptional activation machinery to its binding sites in the chromosome (Ito, 2016).
-Class B and C ARFs are historically categorized as repressor ARFs, though the mechanism through which they confer repression has not been identified.
-Their middle regions tend to be proline- and serine-rich (citation, Ulmasov?).
-
-Canonical ARFs are comprised of three major domains.
-Recent crystallization of these domains have informed structure-function analysis of the ARFs (Boer, 2014; Nanao, 2014; Korasick, 2014).
-These domains are conserved throughout land plants (Dolf's new paper).
-ARFs share an N-terminal B3 DNA binding domain.
-Flanking this DNA-binding domain is a dimerization domain, which folds up into a single “taco-shaped” domain to allow for dimerization between ARFs.
-There is an auxiliary domain that immediately follows and interacts with the dimerization domain.
-The middle region is the most variable between ARFs, as mentioned above, but is characterized by repetitive units of glutamine (class A), serine, or proline residues (classes B and C).
-
-The C-terminal domain of the ARF is a protein-protein interaction domain mediating interactions among ARFs, between ARFs and other transcription factors, and between ARFs and the Aux/IAA repressors.
-This interaction domain was recently characterized as a Phox and Bem1 (PB1) domain, which is comprised of a positive and negative face with conserved basic and acidic residues, respectively (Nanao, 2014; Korasick, 2014).
-The dipolar nature of the PB1 domain may mediate multimerization by the pairwise interaction of these faces on different proteins as the ARF7 PB1 domain was crystallized as a multimer [@korasick_].
-However, it is unclear whether ARF multimerization occurs or plays a developmental significant role in vivo.
-Interfering with ARF dimerization in either the DNA-binding proximal dimerization domain or the PB1 domain decreases the ability of class A ARFs to activate transcription in a heterologous yeast system (Pierre-Jerome, 2016).
-
-While domain architecture is broadly conserved among the ARFs, there are exceptional cases.
-Three ARFs do not contain a PB1 domain at all, ARF3, ARF13, and ARF17, and several more have lost the conserved acidic or basic residues in the PB1 domain, suggesting they may be reduced to a single interaction domain.
-Several ARFs additionally have an expanded conserved region within the DNA-binding domain, of unknown function.
-The majority of domain variation among ARFs occurs in the large B-class subfamily.
-The liverwort *Marchantia polymorpha* has a single representative ARF of each class (Flores-Sandoval, 2015).
-The expansion of these classes in flowering plants is the result of both whole genome and tandem duplication events (Remington, 2004).
-The growth of the ARF family may have allowed for the expansion of the quantity and complexity of loci regulated by the ARFs and subsequent expansion in their regulation of developmental processes.
-
-Class A ARFs are the most well-studied ARF subfamily---the five family members all act as transcriptional activators and have well-characterized, distinct developmental targets.
-Overall the diversity of class A ARFs was generally low, especially compared to the class B and C ARFs, suggesting that class A ARFs are central to auxin signal transduction and plant development.
-Analysis of class A ARF nonsynonymous diversity suggests that the majority of these ARFs are highly functionally conserved, with ${\pi_N/\pi_S}$ values much lower than 1 with the exception of ARF19, with ${\pi_N/\pi_S}$ value of 1.8.
-Comparing diversity within sister pairs, there is a similar trade-off as seen in most *IAA* sister pairs, with one sister being highly conserved and the other more divergent.
-ARF19 and ARF8 are the more divergent class A ARFs, with ${\pi_N/\pi_S}$ values at least three time those of their sisters, ARF7 and ARF6 respectively.
-This may suggest that ARF6 and ARF7 serve more essential purposes in plant development.
-
-```{r ARF diversity stats}
-gene_info <- geneInfoFromFile(
-  fname = system.file("shiny-app", "gene_family_data",
-                      "ARF_gene_ids.csv", package = "r1001genomes"))
-#annotation <- readAnnotationFile(filename = system.file("shiny-app", 
-  #"gene_family_data", "ARF_annotations.csv", 
-  #package = "r1001genomes"), gene_info = gene_info)
-
-VCF <- readRDS(system.file("shiny-app", "gene_family_data",
-                      "ARF_VCF.rds", package = "r1001genomes"))
-div_stats <- ldply(VCF, diversityStats, geneInfo=gene_info, .id="transcript_ID")
-div_stats$tair_symbol %<>% fct_reorder(.x =  div_stats$tair_symbol, 
-                                       .fun = function(x) 
-                                         {as.integer(str_remove(x, 
-                                              pattern = "ARF"))})
-div_stats %<>% arrange(tair_symbol)
-div_stats %<>% select(tair_symbol, everything())
-
-if (params$output == "html_document") {
-  formatRound(datatable(div_stats,
-                  colnames = c("transcript",
-                               "symbol",
-                               "&pi;<sub>N</sub>",
-                               "&pi;<sub>S</sub>",
-                               "&pi;<sub>N</sub>/&pi;<sub>S</sub>",
-                               "&pi; coding",
-                               "&pi; transcript"),
-                  rownames = FALSE, escape = FALSE,
-                  options = list(paging=FALSE, searching=FALSE)),
-                columns = 2:7, digits = 6)
-  } else 
-{
-  dt <- div_stats
-  names(dt) <- c("transcript",
-                                 "symbol",
-                                 "&pi;~N~",
-                                 "&pi;~S~",
-                                 "&pi;~N~/&pi;~S~",
-                                 "&pi; coding",
-                                 "&pi; transcript")
-  dt %<>% mutate_if(is.numeric, round, 6)
-  names(dt) <- paste0("**", names(dt), "**")
-  kable(dt) 
-}
-```
-
-```{r ARF alignment}
-alignment <- alignCDS(gene_info$transcript_ID)
-# make an alignment data frame
-aln_df <- makeAlnDF(alignment[[2]])
-
-# make a collapsed alignment for translating positions
-collapsed_aln_df <- 
-  aln_df %>% dplyr::select(aln_pos, seq_pos, seq_name) %>%
-  tidyr::spread(key = seq_name, value = seq_pos)
-```
-
-```{r Mask_ARF_alignment, eval=FALSE}
-aliscore(as.AAbin(alignment[[2]]), path = "/Volumes/SSD/Applications/ALISCORE_v2.0/Aliscore_v.2.0/")
-id <- scan(file = "/Volumes/SSD/Applications/ALISCORE_v2.0/Aliscore_v.2.0/input.fas_List_random.txt", sep = " ", quiet = TRUE)
-AAAlign_no_gaps <- as.matrix(alignment[[2]])[, id]
-write.nexus.data(AAAlign_no_gaps, file = "ARFs.nex", format = "protein")
-```
-```{bash ARF_phylogeny_construction, eval=FALSE}
-/Applications/MrBayes/mb
-
-execute ARFs.nex
-# set ARF17 as the outgroup
-outgroup AT1G77850.1
-
-lset nucmodel=protein Rates=invgamma
-# or include Nst=6  ngammacat=4 omegavar=M3 
-prset aamodelpr=fixed(jones) statefreqpr=fixed(empirical) 
-# initial run with aamodelpr=mixed yielded posterior prob of 1 for jones model
-propset ExtTBR$prob=0
-
-mcmcp ngen=250000 nchains=4 printfreq=1000 samplefreq=100 burnin=200
-
-mcmc
-sump
-sumt
-
-```
-
-```{r Load ARF phylogeny and data}
-tree <- treeio::read.beast(file = "ARFs.nex.con.tre")
-
-ARF_data <- read.csv(system.file("shiny-app", "gene_family_data",
-                      "ARF_gene_ids.csv", package = "r1001genomes"))
-ARF_data %<>% left_join(y = div_stats$x$data, by = c("name" = "tair_symbol"))
-#tip_labels need to be in the first column
-ARF_data %<>% select(name, everything())
-ARF_data$tair_locus %<>% as.character()
-
-ARF_data <- ARF_data[match(gsub(x = tree@phylo$tip.label, pattern = "\\.1",replacement = ""), ARF_data$tair_locus), ]
-
-tree@phylo$tip.label <- ARF_data$name
-
-```
-
-```{r Visualize ARF tree, fig.height=5, fig.width=5}
-p <- ggtree(tree, layout = "rectangular", color = "grey") %<+% ARF_data + theme_tree() + scale_x_continuous(expand = c(0, 1.2)) + geom_tiplab(hjust = -0.25) + 
-  geom_tippoint(mapping = aes(size = 4*Pi_NS_Ratio, fill = Pi_NS_Ratio), 
-                shape = 21, color = "white", show.legend = TRUE) + 
-  scale_fill_viridis(values = rescale(c(min(ARF_data$Pi_NS_Ratio), 1,
-                                        max(ARF_data$Pi_NS_Ratio)), 
-                                      to = c(0,1)), guide = "colourbar") + scale_size(guide = "none") + 
-  theme(legend.position = c(.1,.7)) + 
-  geom_nodelab(mapping = aes(label = round(prob,digits = 2)), nudge_x = -.3, nudge_y = .5, size = 3) +
-  geom_tiplab(mapping = aes(color = class), hjust = -0.25) + 
-  scale_color_viridis(discrete = TRUE)
-p
-```
-
-### Class A ARFs 
-For all class A ARFs, the middle region of the protein was the predominant high diversity region.
-The DNA-binding domain had very few, low-diversity missense mutations, as did the PB1 domain.
-Considering the necessity of their conserved functions, the low level of variation in these key DNA and protein-protein interaction domains is expected.
-
-```{r A ARF Alignment}
-gene_info <- subset(gene_info, tair_symbol %in% c("ARF5", "ARF6", "ARF7", "ARF8", "ARF19"))
-alignment <- alignCDS(gene_info$transcript_ID)
-# make an alignment data frame
-aln_df <- makeAlnDF(alignment[[2]])
-#BrowseSeqs(alignment[[2]])
-aln_df <- makeAlnDF(alignment[[2]])
-aln_df <- left_join(aln_df, dplyr::select(gene_info, "tair_locus",
-                                       "tair_symbol", "transcript_ID"),
-                        by = c("transcript_ID" = "transcript_ID"))
-aln_df$tair_symbol %<>% as.factor()
-aln_df$tair_symbol %<>% fct_reorder(
-    {as.integer(str_remove(aln_df$tair_symbol, 
-                           pattern = "ARF"))}, 
-  .desc = TRUE)
-
-aln_df <- chunkAlnDF(aln_df = aln_df, chunk_width = 100)
-VCF <- ldply(.data = VCF, .fun = subset, gt_GT != "0|0")
-VCF <- getCodingDiv(VCF)
-aln_df <- addSNPsToAlnDF(aln_df, VCF)
-A_ARFs <- subset(aln_df, tair_symbol %in% c("ARF5", "ARF6", "ARF7", "ARF8", "ARF19"))
-p <-ggplot(A_ARFs, aes(x = aln_pos, y = tair_symbol,
-                             group = seq_pos, text = variants))
-p <- p +
-  geom_raster(data = na.omit(A_ARFs), mapping = aes(fill = effects)) +
-  scale_fill_viridis(option = "A", discrete = TRUE) +
-  scale_y_discrete() +
-  # expand increases distance from axis
-  labs(x = "", y = "", fill = "variant") +
-  theme_logo(base_family = "Helvetica") +
-  theme(panel.grid = element_blank(), panel.grid.minor = element_blank()) +
-  theme(strip.background = element_blank(),
-        strip.text.x = element_blank(),
-        legend.box = "vertical")
-p
-#ggsave(p,filename = "A_ARFs_alignment_variants.pdf", width = 12, height = 4)
-```
-
-Class A ARFs all contain the canonical B3 DNA-binding domain, PB1 interaction domain and glutamine-rich middle region (Guilefoyle, 2007).
-In the analyzed natural variation, ARF7 had several expansions of polyglutamine sequences in the middle region.
-Polyglutamine regions are known to readily expand and contract throughout evolutionary time due to replication error, and variation in polyglutamine length can be acted on by natural selection and have phenotypic consequences (Press, 2014).
-
-A few missense variants in ARF5 highlighted potential functional variation in the class A ARFs.
-A high frequency allele (*What are the variants?*), resulting in substitutions of small amino acids on the positive face of the PB1 domain for bulky phenylalanines, may disrupt interactions with other ARF or Aux/IAA PB1 domains.
-Interestingly, these variants were always found together in accessions Tu-B2-3, Ru-2, Haes-1, Rd-0, 627ME-13Y1, Le-0, Bu-0 and Bsch-0 *Where are these?*.
-Another highly represented polymorphism in ARF5 results in change in the 17th amino acid from a leucine to a methionine.
-This position is in the N-terminal extension of ARF5 which is of unknown function.
-It is possible that this methionine acts as an alternative start site in the accessions.
-
-### Class B ARFs
-
-*ARF23*, which has a truncated DNA-binding domain, has a high ${\pi_N/\pi_S}$ value of 4.1, and has many high-diversity nonsense variants, as may be expected of a gene that is pseudogenizing.
-
-ARF3, ARF13 and ARF17 all lack a C-terminal PB1 domain, which suggests they may be unable to interact with Aux/IAA repressors, possibly acting as auxin-insensitive competitive inhibitors of ARF-mediated transcription (Fig ARF3_13_17 alignment).
-These ARFs span the range of ${\pi_N/\pi_S}$ ratios relative to the other ARFs, ranging from 0.85 to 4.1.
-ARF17 is more conserved than the activator ARF, ARF19.
-ARF13 has many nonsense variants of high diversity and may be pseudogenizing.
-
-```{r ARF3_13_17 alignment}
-
-```
-
-ARF14, ARF15, ARF16, ARF20, and ARF21 all do not have one or multiple of the conserved acidic residues in the PB1 domain, and themselves form a subclade of the B ARFs.
-Interestingly, no non-synonymous variation is seen in these amino acid residues of these ARFs, suggesting that the amino acids that are in the regions may still be important for the function of these proteins.
-
-### Class C ARFs
-
-Class C ARFs show low nucleotide diversity scores, all ${\pi_N/\pi_S}$ values substantially lower than 1.
-ARF16 was the most conserved, whereas its clade members (ARF10, ARF17) had scores at least four times higher [@remington_contrasting_2004].
-
-Structurally, all three members of Class C ARFs contain a canonical B3 DNA-binding domain, but only ARF10 and ARF16 contain a PB1 domain.
-The DNA binding domains exhibit overall low diversity and presence of largely synonymous changes for all the ARFs.
-Only the ARF10 DNA binding domain contains missense variants (S147T and A157S) (Fig.
-ARF10_DBD Variants distribution).
-**Any insights into whether these may affect function?
-They are quite conservative.** Of the PB1 domain containing class C ARFs, ARF16 exhibits several missense variants which are sporadically distributed, in contrast to the conserved PB1 domain of ARF10 (Fig.
-Arf16_PB_variants distribution).
-
-```{r ARF10_DBD Variants distribution}
-
-```
-
-```{r Arf16_PB_variants distribution}
-
-```
-
-Several members of class B (ARF2, 3 and 4) and C ARFs (ARF10, 16 and 17) are distinguished from other ARFs on the basis of their post transcriptional regulation by trans-acting small interfering RNAs (tasiRNAs) and microRNAs (miRNAs) [@rhoades_prediction_2002; @allen_micrornadirected_2005].
-**(@AMY Arf 6 and 8 targeted by miRNA 167 - check for conservation).** The miRNA target of class C ARFs showed a high degree of conservation (Fig.
-ARF10_16_17_Alignment).
-This is consistent with perturbation studies that have shown drastic developmental phenotype for class C ARFs [@wang_control_2005; @mallory_micrornadirected_2005; @liu_repression_2007].
-For the class B ARFs (2, 3 and 4) the tasiRNA binding motif is also highly conserved suggesting strong regulatory constraints on these genes (Fig ARF2_3_4_Alignment).
-The DNA binding domain of ARF4 exhibits non-synonymous variation in 4 accession and would be interesting to carry out phenotypic characterization of these accessions for developmental growth defects (Fig ARF4_DBD Variants map).
-
-```{r ARF10_16_17_Alignment}
-
-```
-
-```{r ARF2_3_4_Alignment}
-
-```
-
-```{r ARF4_DBD Variants map}
-
-```
-
-#Discussion
-
-Analyzing natural diversity within the gene families constituting the auxin nuclear signaling pathway has confirmed existing knowledge, supported existing hypothses and generated new hypotheses.
-This analysis in particular has confirmed the importance of auxin-induced degradation of the Aux/IAA proteins as evidenced by the strong conservation of the *AFB* genes and the canonical degron region of the *Aux/IAA* genes.
-Despite this relative conservation, this analysis has also identified many accessions containing new alleles of interest in the *AFB* and *Aux/IAA* families.
-This analysis has also
-
-Importance of 
--AFB
--IAA degron
--TPL
--ARF DBD and PB1

### 2018-07-12 manuscript/06-ARFs.Rmd
-Cleaned up manuscript folder -moved to bookdown format -revised all sections

--- 2018-07-10 manuscript/manuscript.Rmd
@@
----
-title: A comprehensive analysis of natural sequence variation within the <i>Arabidopsis
-  thaliana</i> nuclear auxin signaling pathway using an accessible web application
-author:
-  affiliation: University of Washington
-  name: R. Clay Wright
-date: "`r Sys.Date()`"
-params: 
-  output: word_document
-output:
-  word_document:
-    toc: yes
-  html_document:
-    fig_caption: yes
-    toc: yes
-always_allow_html: yes
-fontfamily: Helvetica
-fontsize: 11pt
-keywords: r markdown
-link-citations: yes
-csl: apa.csl
-bibliography:
-- natural_variation.bib
-- TIR1_AFB.bib
-- ARF_C.bib
-- AuxIAA_citations.bib
----
-
-```{r setup, message=FALSE, warning=FALSE}
-knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE, collapse = TRUE, comment = "#>", tidy = TRUE, cache = TRUE, cache.lazy = FALSE, out.width = "90%", fig.height=4, fig.width=7, out.extra='style="margin: auto; display: block; padding-top: 15px;"')
-
-#devtools::install_github("wrightrc/r1001genomes", ref = "auxin-natural-variation")
-library(ape)
-library(r1001genomes)
-library(DT)
-library(tidyverse)
-library(magrittr)
-library(RColorBrewer)
-library(ggpmisc)
-library(ggthemes)
-library(ggseqlogo)
-library(ggtree)
-library(reticulate)
-library(scales)
-library(viridis)
-library(ips)
-library(knitr)
-
-#use_python(python = "/usr/local/bin/python")
-```
-<!--This is a CSS (Cascading Style Sheet).
- It controls how panels and images are formatted.></!-->
-<style>
-  img {
-    display: block;
-    margin: auto;
-    width: 90%;
-    height: auto;
-  }
-</style>
-
-# List of working titles  
-* A comprehensive analysis of natural sequence variation within the <i>Arabidopsis thaliana</i> nuclear auxin signaling pathway  
-* **Add your suggestions**
-
-# Abstract
-
-* Motivating, high level statement  
-* Statement defining the cutting edge of your research area  
-* A statement of the form: "However, it remains unclear ..." that defines what problem you are  
-addressing.  
-* A statement of the form:"Here, we ..."  that concisely states your contribution.  
-* A series of 3?5 statement about your results.  
-* A concluding statement.
-
-# Intro
-
-The first genome sequence of *Arabidopsis thaliana* facilitated rapid advancement of plant biology through molecular genetics.
-Since this initial genome, massive scaling of sequencing technology has allowed a pioneering group to survey the global genomic variation in natural *Arabidopsis thaliana* populations.
-This valuable population genetics resource has led to several associations of genetic loci with phenotypic traits and provided insights into how selective pressure has and is influencing the evolution of plant genomes.
-Outside of genome-wide studies however, this valuable dataset has seen little use.
-Natural genomic variation also provides a set of feasible functional variation at the gene, gene family and gene network level that may provide insight into the function and evolution of genes, families and networks.
-Here, we present a web application and R-package through which plant molecular biologists with little-to-no bioinformatics experience can make use of this rich dataset of genetic variation to formulate hypotheses as to the sequence/function/phenotype relationships determined by the gene, family, or network of their interest.
-We demonstrate the utility of this tool through comprehensive analysis and identification of potential functional variation in the nuclear auxin signaling pathway.
-
-This application allows easy access to 
-1. lists of missense polymorphisms to facilitate biochemical assays of variant effects [@starita_variant_2017],
-2. family-wise alignments of variants to facilitate de novo functional domain identification [@melamed_combining_2015], 
-3.
-lists of accessions containing missense (or any type of) polymorphisms to facilitate segregation analysis and measurment of the distribution of variant effects on phenotype [@park_distributions_2017],
-4. as well as the study of gene and network evolution [@delker_natural_2010; @kliebenstein_role_2008].
-
-Here, we have utilized this tool to analyze the genetic variation in the nuclear auxin signaling pathway and formulate hypotheses regarding the functional implications of this variation and the evolution of the genes and gene families in this pathway.
-Our analysis also provided further confirmation of much of the existing knowledge of these genes demonstrating the validity of this approach.
-
-# Methods
-
-## Nucleotide Diversity Calculation
-Nei  defined the nucleotide diversity statistic in his original paper as:
-“…the average number of nucleotide differences per site between two randomly chosen DNA sequences.” (Nei, 1979)
-And provided the equation:
-
-$\displaystyle \pi = \sum_{ij}x_{i}x_{j}\pi_{ij}$
-
-Where Where $x_{i}$ is the frequency of the ith sequence in the population and $\pi_{ij}$ is the number of sites that are different between the ith and jth sequence divided by sequence length.
-
-a more general form, that treats each sequence in the population as unique can be written as:
-
-$\displaystyle \pi = \frac{1}{L*n^2} \sum_{i=1}^n \sum_{j=1}^n \sum_{k=1}^L{\pi_{ijk}}
-\qquad
-\displaystyle \pi_{ijk} = \left\{\begin{array}{l}
-1 \quad \textrm{if} \quad N_{ik} \neq N_{jk} \\
-0 \quad \textrm{if} \quad N_{ik} = N_{jk}
-\end{array}\right.$
-
-where $N_{ik}$ is the nucleotide (A, T, C or G) at position k on the ith sequence of the population.
-$L$ is the length of the sequence (the same for all sequences in the population, 
-assumes only SNPs exist, no indels). n is the total number of sequences in the population.
-
-from this form we can re-arrange summations to the form below:
-
-$\displaystyle \pi = \frac{1}{L} \sum_{k=1}^L{\pi_{k}}
-\qquad
-\pi_{k} = \frac{1}{n^2} \sum_{i=1}^n \sum_{j=1}^n {\pi_{ijk}}$
-
-here $\pi_k$ can be thought of as the site-wise nucleotide diversity at position k, and is equal to 
-the nucleotide diversity of a sequence of length 1 at location k. we can calculate $\pi_k$ for each
-site, then average those over the sequence length to calculate $\pi$, the nucleotide diversity 
-of the sequence.
-
-The function `Nucleotide_diversity` in the r1001genomes package calculates $\pi_k$ for each position in the gene or region that contains a variant.
-note, $\pi_k$ is equal to 0 at all locations without variants.
-This is also what is displayed in the Diversity Plot tab of the web app.
-
-## Detailed $\pi_k$ calculaiton simplification.
-
-The formula for $\pi_k$ above requires comparing every sequence to every other sequence at location k,
-however, we know there are only a few variant forms at each individual location.
-
-so we can revert back to using Nei's original formula, modifying it slightly,
-replacing $x_i$ with $\frac{n_i}{n}$, $n_i$ being the number of sequences in the population of the form $N_i$ at location k: subscripts i and k are summed only over unique forms [need to clarify this]
-
-$\displaystyle \pi_k = \sum_{ij} \frac{n_i}{n}\frac{n_j}{n} \pi_{ij} = 
-\frac{1}{n^2}\sum_{ij} n_i n_j \pi_{ij}$
-
-$\pi_{ij(k)} = \left\{\begin{array}{l}
-1 \quad \textrm{if} \quad i \neq j \\
-0 \quad \textrm{if} \quad i = j
-\end{array}\right.$
-
-we will define $n_{!i} = n - n_i$ as the number of sequences different from i at k.
-
-we can see the summed term will be zero if $i=j$, and $n_i n_j$ if $i\neq j$ so:
-
-$\displaystyle \pi_k = \frac{1}{n^2} \sum_{i} n_i n_{!i}$
-
-next we substitute our definition of $n_{!i}$
-
-$\displaystyle \pi_k = \frac{1}{n^2} \sum_{i} n_i (n - n_{i})$
-
-then distribute and split summation:
-
-$\displaystyle \pi_k = \frac{1}{n^2} (n \sum_{i} n_i - \sum_{i} n_{i}^2)$
-
-finally, summing $\displaystyle \sum_{i}n_i$ is equal to n:
-
-$\displaystyle \pi_k = \frac{1}{n^2} (n^2 - \sum_{i} n_{i}^2)$
-
-this simplified form for $\pi_k$ is used by the app, because the counts of unique 
-variants at a single nucleotide location can easily be summarized in R.
-
-# Results  
-## *TIR1/AFB* genes
-
-Auxin acts by binding to receptors that in turn target co-repressors for degradation.
-Auxin receptors (six in the model plant *Arabidopsis thaliana*) evolved through gene duplication and diversification early in the history of vascular plants [@Parry_Complex_2009].
-The rate of co-repressor degradation is determined by the identity of both the receptor and co-repressor [@Havens_Synthetic_2012], and this rate sets the pace of lateral root development [@Guseman_Auxin-induced_2015].
-
-Although it is unclear what unique roles each receptor plays in growth and development, a number of studies have pointed out differences in the ways the six different receptors in A.
-thaliana differ in biochemical function and expression domain[@dharmasiri_plant_2005; @parry_complex_2009; @prigge_arabidopsis_2016].
-Although TIR1/AFBs are expressed ubiquitously in A.
-thaliana tissues, TIR1, AFB2, and AFB3 have been shown to accumulate in the shoot and root meristems and leaf tissues, with slightly different expression patterns for TIR1 [@Parry_Complex_2009].
-Additionally, the expression of AFB5 is stronly circadian-regulated [@Covington_Circadian_2007] and AFB3 is more highly expressed in the roots in the presence of nitrate, allowing increased lateral root formation [@Vidal_Nitrate_2010], suggesting more broad environmental regulation of this gene family may exist.
-  Functionally, all members of this family have been shown to bind auxin and Aux/IAA proteins.
-However, AFB1 has drastically reduced ability to assemble into an SCF complex, due to the substitution E8K in its F-box domain, preventing it from inducing degradation of Aux/IAAs[@Yu_Untethering_2015].
-This lack of complexation may allow observed high ubiquitous AFB1 accumulation [@Parry_Complex_2009].
-Higher order mutants in the family containing *afb1* mutants suggest that *AFB1* has a moderate positive effect on auxin signaling.
-Additionally, AFB4 and AFB5 have been shown to preferentially and functionally bind the synthetic auxin picloram, while other family members preferentially bind indole-3-acetic acid [@Prigge_Arabidopsis_2016].
-Interestingly, the strength and rate with which TIR1/AFBs are able to bind and mark Aux/IAAs for degradation are variable [@calderonvillalobos_combinatorial_2012; @Havens_Synthetic_2012].
-AFB2 induced the degradation of certain Aux/IAA proteins at a faster rate than TIR1, suggesting some functional specificity has arisen since the initial duplication between the *TIR1/AFB1* and *AFB2/AFB3* clades.
-
-```{r TIR1_AFB_div_stats}
-gene_info <- geneInfoFromFile(
-  fname = system.file("shiny-app", "gene_family_data",
-                      "AFB_gene_ids.csv", package = "r1001genomes"))
-annotation <- readAnnotationFile(filename = system.file("shiny-app", "gene_family_data",
-                      "AFB_annotations.csv", package = "r1001genomes"), gene_info = gene_info)
-alignment <- alignCDS(gene_info$transcript_ID)
-# make an alignment data frame
-aln_df <- makeAlnDF(alignment[[2]])
-
-# make a collapsed alignment for translating positions
-collapsed_aln_df <- 
-  aln_df %>% dplyr::select(aln_pos, seq_pos, seq_name) %>%
-  tidyr::spread(key = seq_name, value = seq_pos)
-
-VCF <- readRDS(system.file("shiny-app", "gene_family_data",
-                      "AFBs_VCF.rds", package = "r1001genomes"))
-div_stats <- ldply(VCF, diversityStats, geneInfo=gene_info, .id="transcript_ID")
-div_stats$tair_symbol %<>% fct_reorder(.x =  div_stats$tair_symbol, 
-                                       .fun = function(x) 
-                                         {as.integer(str_remove(x, 
-                                              pattern = "AFB|TIR"))})
-div_stats$tair_symbol %<>% fct_relevel("TIR1")
-div_stats %<>% arrange(tair_symbol)
-div_stats %<>% select(tair_symbol, everything())
-
-if (params$output == "html_document") {
-  formatRound(datatable(div_stats,
-                  colnames = c("transcript",
-                               "symbol",
-                               "&pi;<sub>N</sub>",
-                               "&pi;<sub>S</sub>",
-                               "&pi;<sub>N</sub>/&pi;<sub>S</sub>",
-                               "&pi; coding",
-                               "&pi; transcript"),
-                  rownames = FALSE, escape = FALSE,
-                  options = list(paging=FALSE, searching=FALSE)),
-                columns = 2:7, digits = 6)
-  } else 
-{
-  dt <- div_stats
-  names(dt) <- c("transcript",
-                                 "symbol",
-                                 "&pi;~N~",
-                                 "&pi;~S~",
-                                 "&pi;~N~/&pi;~S~",
-                                 "&pi; coding",
-                                 "&pi; transcript")
-  dt %<>% mutate_if(is.numeric, round, 6)
-  names(dt) <- paste0("**", names(dt), "**")
-  kable(dt) 
-}
-```
-
-Examining the natural sequence variation across the *AFB* family revealed that *TIR1* and *AFB1* both had very low nonsynonymous diversity, hinting at their likely functional importance and bringing in to question the inconclusive role of *AFB1* in auxin signaling.
-*AFB3* and *AFB4* had higher nonsynonymous diversity, while their sister genes, *AFB2* and *AFB5* were more conserved.
-This matches our current understanding of *AFB3* as playing a minor role in the auxin signaling pathway and *AFB4* perhaps undergoing pseudogenization.
-
-Although most known functional regions are highly conserved in *AFB1*, there is a nonsynonymous polymorphism  in the oligomerization domain, only found in the Can-0 accession.
-Mutations in this domain of *TIR1* frequently have a semidominant effect on root phenotypes [@dezfulian_oligomerization_2016; @wright_insights_2017a].
-Characterization of this accession may identify potential *AFB1* phenotypes and may also test the hypothesis that AFB1 functions through oligomerization with other TIR1/AFBs.
-
-```{r AFB variants for PyMOL}
-coding_vcf <- ldply(.data = VCF, .fun = subset, 
-             !is.na(Transcript_ID) & gt_GT != "0|0")
-aln_df <- addSNPsToAlnDF(aln_df, coding_vcf, seq_name = Transcript_ID, 
-               seq_pos = Codon_Number)
-collapsed_aln_df %<>% left_join(y = subset(aln_df, seq_name == gene_info[gene_info$tair_symbol == "TIR1", "transcript_ID"]))
-
-#' Select variants for importing into PyMOL
-#'
-#' @param gene_info a gene info data frame
-#' @param aln_df an alignment data frame with attached SNPs
-#' @param structure_ref string, which tair_symbol or transcript_ID (depending on 
-#' \code{name_by_symbols}) corresponds to the protein structure to which 
-#' selections should be mapped
-#' @param effect_type string of the effect type to select from the alignment 
-#' data frame
-#' @param name_by_symbols logical, if true (default) tair_symbols will be used 
-#' to name the selections, otherwise transcript_IDs will be used
-#'
-#' @return
-#' @export
-#'
-#' @examples
-select_variants <- function(gene_info, aln_df, structure_ref, 
-                            effect_type = "missense", name_by_symbols = TRUE){
-  collapsed_aln_df <- aln_df %>% dplyr::select(aln_pos, seq_pos, seq_name) %>%
-    tidyr::spread(key = seq_name, value = seq_pos)
-  if(name_by_symbols)
-    column <- "tair_symbol"
-    else 
-      column <- "transcript_ID"
-
-  selections <- ldply(.data = gene_info[,column], .fun = function(gene){
-    collapsed_aln_df %<>% dplyr::left_join(y = subset(aln_df, seq_name ==  
-                      gene_info[gene_info[, column] == gene, "transcript_ID"]))
-    variants <- collapsed_aln_df[
-      grepl(effect_type, collapsed_aln_df$effects), 
-      gene_info[gene_info[, column] == structure_ref, "transcript_ID"]]
-    c(gene, paste0(na.omit(as.numeric(variants[[1]])), collapse = "+"))
-  })
-  return(selections)
-}
-
-selections <- select_variants(gene_info, aln_df, structure_ref = "TIR1")
-selections$V2 <- paste0("chain B and resi ", selections$V2)
-write.table(selections, file = "AFB_variant_selections", row.names = FALSE, 
-          col.names = FALSE, sep = ",", quote = FALSE)
-```
-
-```{python readSelection, eval=FALSE}
-#
-# -- readSelection.py
-#
-from pymol import cmd, stored
-
-def readSelection (fileName):
-	'''
-	DESCRIPTION
-
-	reads a csv file containing a column of selection names and a column of
-	selection strings and creates selection objects for each
-	'''
-	#read in the selection file
-	with open(fileName) as f:
-		lines = f.readlines()
-		for line in lines:
-			sel = line.split(",")
-			cmd.select(sel[0], sel[1])
-
-	f.close()
-
-cmd.extend('readSelection',readSelection);
-```
-
-```{PyMol, AFB_structure, eval=FALSE}
-#from within pymol
-cmd.load("Dropbox/180705_NIH_renewal/Aim1-figures/2P1Q.pdb")
-hide everything
-
-run ~/Desktop/natural-variation-webtool/manuscript/readSelection.py
-
-readSelection ~/Desktop/natural-variation-webtool/manuscript/AFB_variant_selections
-
-# Define selections
-# select TIR1variants, (chain B and resi 154+239+546)
-select fbox, (chain B and resi 1-50)
-select LRR, (chain B and resi 50-576)
-select TIR1, (chain B and resi 1-576)
-# select AFB2variants, (chain B and resi 38+174+181+184+209+259+401+496)
-select tir1_1, (chain B and resi 147)
-select TIR1dm, (chain B and resi 170+473)
-select insPcontacts, (chain B and resi 74+78+113+114+344+401+403+436+460+484+485+509)
-select dimerNULL, (chain B and resi 142+143+151+117+147)
-select auxin, resn IAC
-select AUXcontacts, (chain B and resi 78+79+82+378+380+403+404+405+438+439+462+463+464+489)
-select insP, resn IHP
-select IAA7, chain C
-select IAA7contacts, (chain B and resi 81+82+84+139+347+349+350+351+352+380+405+406+407+409+464+465+489)
-select IAA7_N, chain C and resi 1
-select IAA7_C, chain C and resi 13
-select ASK1, chain A
-select Y, chain C and resi 11
-select Ycontacts, chain B and resi 165+166+195
-
-#Define color palette
-#set_color name, [ red-float, green-float, blue-float ]
-set_color blue1, [0.00000000, 0.20392157, 0.6980392]
-set_color blue2, [0.09803922, 0.36078431, 1.0000000]
-set_color blue3, [0.00000000, 0.29019608, 1.0000000]
-set_color teal,  [0.09803922, 1.00000000, 0.6980392]
-set_color LRR,   [1.00000000, 0.71764706, 0.0000000]
-set_color Fbox,  [0.69803922, 0.49411765, 0.0000000]
-set_color TIR1l, [0.67058824, 0.07843137, 0.8000000]
-set_color TIR1f, [0.63529412, 0.34117647, 0.6980392]
-set_color AFB2l, [0.37647059, 0.70196078, 1.0000000]
-set_color AFB2f, [0.07843137, 0.64313725, 0.8000000]
-
-show cartoon, 2P1Q
-
-color lrr, LRR
-
-color TIR1l, TIR1
-show sticks, TIR1
-color TIR1f, AFB1
-show sticks, AFB1
-color AFB2l, AFB2
-show sticks, AFB2
-color AFB2f, AFB3
-show sticks, AFB3
-
-set ray_trace_mode, 3
-
-# used get_view() to return the current view (position) of the molecule 
-# copy the code to be able to set the identical viewpoint later
-# to return the current view from Pymol
-# get_view
-
-set_view (\
-     0.044671144,    0.083668359,   -0.995409012,\
-     0.774909019,    0.625905454,    0.087390237,\
-     0.630409002,   -0.775298953,   -0.036878210,\
-    -0.001916260,    0.005584979, -202.498916626,\
-    -5.538011074,   -8.507736206,   39.046234131,\
-  -219.237945557,  624.123413086,  -20.000000000 )
-
-#to render a high definition image use ray
-ray
-
-```
-
-```{r AFB1 diversity plot}
-coding_variants <- getCodingDiv(VCF[[gene_info[gene_info$tair_symbol == "AFB1", "transcript_ID"]]])
-p <- plotCodingDiv(uniqueCodingVars = coding_variants)
-annotation <- addAlnPosToAnno(anno_df = annotation, aln_df)
-
-# translate positions from TIR1 to AFB1
-AFB1_positions <- annotation$positions
-AFB1_positions %<>% left_join(y = select(collapsed_aln_df, aln_pos, gene_info[gene_info$tair_symbol == "AFB1", "transcript_ID"]), by = "aln_pos") 
-AFB1_positions$position <- AFB1_positions$AT4G03190.1
-
-p <- append_layers(p,list(
-        geom_rect(data = AFB1_positions,
-                   mapping = aes(xmin = as.integer(position)-1,
-                                 xmax = as.integer(position)+2,
-                                 fill = annotation),
-                  ymin = -Inf, ymax = Inf, inherit.aes = FALSE, alpha = 0.8)),
-      position = "bottom")
-p
-```
-
-The AFB4 and AFB5 receptors have an N-terminal extension prior to the F-box domains.
-This extension had very high nonsynonymous diversity(Fig.
-4), suggesting that this extension does not play an important functional role in these proteins.
-Additionally, two frameshift variants and one stop-gained variant were observed in *AFB4* supporting its pseudogenization.
-
-Figure 4: Alignment of A.
-thaliana TIR1/AFB F-box domains showing variants.
-Cyan marks the F-box binding domain.
-Yellow, green, and magenta, mark missense and synonymous, synonymous only, and missense only variant locations respectively.
-Yellow squared E’s show Cul1-interacting residues, differing in AFB1.
-```{r TIR1/AFB F-box Alignment}
-aln_df <- makeAlnDF(alignment[[2]])
-aln_df <- left_join(aln_df, dplyr::select(gene_info, "tair_locus",
-                                       "tair_symbol", "transcript_ID"),
-                        by = c("transcript_ID" = "transcript_ID"))
-aln_df$tair_symbol <- fct_rev(fct_relevel(aln_df$tair_symbol, "TIR1"))
-aln_df <- chunkAlnDF(aln_df = aln_df, chunk_width = 55)
-VCF <- ldply(.data = VCF, .fun = subset, gt_GT != "0|0")
-VCF <- getCodingDiv(VCF)
-aln_df <- addSNPsToAlnDF(aln_df, VCF)
-F_box <- subset(aln_df, aln_pos <= 96)
-F_box <- chunkAlnDF(aln_df = F_box, chunk_width = 80)
-p <-ggplot(F_box, aes(x = aln_pos, y = tair_symbol,
-                             group = seq_pos, text = variants))
-p <- p +
-  geom_tile(data = na.omit(F_box), mapping = aes(fill = effects),
-            width = 1, height = 1, alpha = 0.8) +
-  geom_text(aes(label=letter), alpha= 1, family = "Courier") +
-  scale_fill_brewer(type = "qual", palette = 1, direction = -1) +
-  scale_x_continuous(breaks=seq(1,max(F_box$aln_pos), by = 10)) +
-  scale_y_discrete() +
-  # expand increases distance from axis
-  labs(x = "", y = "", fill = "variant") +
-  theme_logo(base_family = "Helvetica") +
-  theme(panel.grid = element_blank(), panel.grid.minor = element_blank()) +
-  facet_wrap(facets = ~chunk, ncol = 1, scales = "free") +
-  theme(strip.background = element_blank(),
-        strip.text.x = element_blank(),
-        legend.box = "vertical")
-p
-```
-
-## *Aux/IAA* genes
-
-The Aux/IAA proteins (IAAs) mediate both auxin perception and transcriptional co-repression of auxin responsive genes.
-Auxin perception is carried out along with TIR1/AFB proteins, which together with IAAs act as co-receptors to bind and perceive nuclear auxin.
-Interaction of IAAs with auxin and TIR1/AFBs is conferred by the highly conserved 13 amino acid degron motif [@tan_mechanism_2007].
-Formation of this auxin co-receptor complex triggers ubiquitination and subsequent degradation of IAAs.
-This degradation in turn relieves their repression upon ARF transcription factors.
-Stabilizing mutations within the degron can lead to dominant phenotypes [@berleth_auxin_2004; @yang_iaa1_2004].
-The *Arabidopsis thaliana* Aux/IAA family has 29 members, most of which possess a canonical auxin degron.
-Notable exceptions are the auxin-insensitive IAA20 and IAA30 that lack a degron sequence entirely, while the closely related IAA31 retains a semi-functional degron sequence with a very slow degradation rate [@dreher_arabidopsis_2006; @havens_synthetic_2012].
-Additionally, IAA degradation rate is influenced by rate motifs found outside the degron, such as the KR motif [@dreher_arabidopsis_2006; @moss_rate_2015].
-IAA-mediated repression of ARFs depends upon both an EAR domain, N-terminal of the degron, that recruits co-repressors of the TOPLESS (TPL) family, and a Phox and Bem1 (PB1) domain that enables interaction via hetero-oligomerization with ARF PB1 domains [@nanao_structural_2014; @kagale_ear_2011].
-
-```{r IAA diversity stats, fig.height=11}
-gene_info <- geneInfoFromFile(
-  fname = system.file("shiny-app", "gene_family_data",
-                      "IAA_gene_ids.csv", package = "r1001genomes"))
-#annotation <- readAnnotationFile(filename = system.file("shiny-app", "gene_family_data",
-#                      "IAA_annotations.csv", package = "r1001genomes"), gene_info = gene_info)
-
-VCF <- readRDS(system.file("shiny-app", "gene_family_data",
-                      "IAAs_VCF.rds", package = "r1001genomes"))
-div_stats <- ldply(VCF, diversityStats, geneInfo=gene_info, .id="transcript_ID")
-div_stats$tair_symbol %<>% fct_reorder(.x = , div_stats$tair_symbol, 
-                                       .fun = function(x) 
-                                         {as.integer(str_remove(x, 
-                                              pattern = "IAA"))})
-div_stats %<>% arrange(tair_symbol)
-div_stats %<>% select(tair_symbol, everything())
-
-if (params$output == "html_document") {
-  formatRound(datatable(div_stats,
-                  colnames = c("transcript",
-                               "symbol",
-                               "&pi;<sub>N</sub>",
-                               "&pi;<sub>S</sub>",
-                               "&pi;<sub>N</sub>/&pi;<sub>S</sub>",
-                               "&pi; coding",
-                               "&pi; transcript"),
-                  rownames = FALSE, escape = FALSE,
-                  options = list(paging=FALSE, searching=FALSE)),
-                columns = 2:7, digits = 6)
-  } else 
-{
-  dt <- div_stats
-  names(dt) <- c("transcript",
-                                 "symbol",
-                                 "&pi;~N~",
-                                 "&pi;~S~",
-                                 "&pi;~N~/&pi;~S~",
-                                 "&pi; coding",
-                                 "&pi; transcript")
-  dt %<>% mutate_if(is.numeric, round, 6)
-  names(dt) <- paste0("**", names(dt), "**")
-  kable(dt) 
-}
-```
-
-Within the available natural sequence variation, we found several missense polymorphisms affecting the EAR domain or degron of several canonical IAAs.
-However, no missense polymorphisms were present in the KR motif or residues critical for PB1-PB1 domain interactions (Supp.
-Fig.
-Alignment of canonical IAAs).
-The majority of missense mutations observed were in the variable regions between the conserved domains.
-
-###Sister Pair Diversification
-
-Previous research has found evidence of both broad genetic redundancy and specificity within closely related pairs or groups of IAA proteins [@overvoorde_functional_2005].
-For example, the *iaa8-1 iaa9-1* double mutant and the *iaa5-1 iaa6-1 iaa19-1* triple mutant have wild-type phenotypes [@overvoorde_functional_2005], yet the *IAA6/19* sister pair has significant differences in expression patterns, protein abundances and functions suggesting they have undergone functional specialization since their divergence [@winkler_variation_2017].
-
-In our analysis of the *IAA* gene family, we frequently observed that one member of most *IAA* sister pairs had high nonsynonymous diversity while the other sister was more conserved (Figure IAA phylogeny heat map).
-In an interspecific comparison between *A.
-thaliana* and *A.
-lyrata*, *IAA19* is more conserved, while *IAA6* has regions of increased sequence divergence upstream of the degron and in the N-terminal half of the PB1 domain [@winkler_variation_2017].
-Consistent with these results, our analysis of intraspecific natural variation revealed higher conservation for *IAA19* (${\pi_N/\pi_S}$ = 0.55) compared to *IAA6* (${\pi_N/\pi_S}$ = 2.3) (Figure IAA phylogeny heat map), and also detected high diversity within the same regions of *IAA6* as seen in Winkler et al.
-(Supp Fig IAA6 diversity plot).
-```{r IAA alignment}
-alignment <- alignCDS(gene_info$transcript_ID)
-# make an alignment data frame
-aln_df <- makeAlnDF(alignment[[2]])
-
-# make a collapsed alignment for translating positions
-collapsed_aln_df <- 
-  aln_df %>% dplyr::select(aln_pos, seq_pos, seq_name) %>%
-  tidyr::spread(key = seq_name, value = seq_pos)
-```
-
-```{r Mask_IAA_alignment, eval=FALSE}
-aliscore(as.AAbin(alignment[[2]]), path = "/Volumes/SSD/Applications/ALISCORE_v2.0/Aliscore_v.2.0/")
-id <- scan(file = "input.fas_List_random.txt", sep = " ", quiet = TRUE)
-AAAlign_no_gaps <- as.matrix(alignment[[2]])[, id]
-write.nexus.data(AAAlign_no_gaps, file = "manuscript/IAAs.nex", format = "protein")
-```
-```{bash IAA_phylogeny_construction, eval=FALSE}
-~/Desktop/natural-variation-webtool/manuscript
-
-/Applications/MrBayes/mb
-
-execute IAAs.nex
-# set IAA33 as the outgroup
-outgroup AT5G57420.1
-
-lset nucmodel=protein Rates=invgamma
-# or include Nst=6  ngammacat=4 omegavar=M3 
-prset aamodelpr=fixed(jones) statefreqpr=fixed(empirical) 
-# initial run with aamodelpr=mixed yielded posterior prob of 1 for jones model
-propset ExtTBR$prob=0
-
-mcmcp ngen=250000 nchains=4 printfreq=1000 samplefreq=100 burnin=200
-
-mcmc
-sump
-sumt
-
-```
-
-```{r Load IAA phylogeny and data}
-tree <- treeio::read.beast(file = "IAAs.nex.con.tre")
-
-IAA_data <- read.csv("IAAs_table.csv")
-#tip_labels need to be in the first column
-IAA_data %<>% select(name, everything())
-IAA_data$name %<>% as.character()
-IAA_data$AGI %<>% as.character()
-
-IAA_data <- IAA_data[match(gsub(x = tree@phylo$tip.label, pattern = "\\.1",replacement = ""), IAA_data$AGI), ] 
-tree@phylo$tip.label <- IAA_data$name
-```
-*I think IAA20, 30, and 15 should also be in class C. So there are really two large clades A and B and then there are truncations/loss of function/dominant variants scattered throughout, Class C.*
-
-```{r IAA_tree, fig.height=5, fig.width=5, fig.cap="(ref:IAA tree)"}
-p <- ggtree(tree, layout = "rectangular", color = "grey") %<+% IAA_data + theme_tree() + scale_x_continuous(expand = c(0, 1.2)) + 
-  geom_tippoint(mapping = aes(size = 4*PiN_PiS, fill = PiN_PiS), shape = 21, color = "white", show.legend = TRUE) + 
-  scale_fill_viridis(values = rescale(c(min(IAA_data$PiN_PiS), 1,
-                                        max(IAA_data$PiN_PiS)), 
-                                      to = c(0,1)), guide = "colourbar") + scale_size(guide = "none") + 
-  theme(legend.position = c(.9,.7)) + 
-  geom_nodelab(mapping = aes(label = round(prob,digits = 2)), nudge_x = -.3, nudge_y = .5, size = 3) + 
-  geom_tiplab(mapping = aes(color = Class), hjust = -0.25) + 
-  scale_color_viridis(discrete = TRUE)
-p
-```
-
-(ref:IAA tree) IAA protein sequence tree mapped with ${\pi_N/\pi_S}$
-
-Within the Group B *IAA* genes, there are four sister pairs---*IAA10*/*IAA11*, *IAA12*/*IAA13*, *IAA18*/*IAA26*, and *IAA20*/*IAA30* \@ref(IAA tree) [@remington_contrasting_2004].
-Interestingly, among these sister pairs only *IAA12* and *IAA13* have a noticeable difference in nonsynonymous diversity.
-The *Arabidopsis thaliana* EFP browser shows that **IAA10** and *IAA11* have almost identical expression patterns, with the exception of *IAA10* being more highly expressed in the developing seed, and *IAA11* being more highly expressed in root tissue during lateral root development **need to cite original data source here**.
-In our analysis, *IAA10* and *IAA11* both showed functional conservation (${\pi_N/\pi_S}$ of 0.80 and 0.67 respectively) in support of *IAA10* and *IAA11* playing redundant roles.
-Accessions and associated variants identified here provide a potentially expedited route to elucidating the extent of redundancy between family members such as *IAA10* and *IAA11*.
-
-### New Loss-of-Function/Gain-of-Function alleles
-Accessions with functional variants can be potentially used to investigate the extent of redundancy/specialization within sister pairs.
-For example, three accessions have a missense variant affecting the EAR domain of IAA10 (HR-5, UKSE06-118 and UKSW06-240, all from southern UK).
-This variant (L49S) is in the second leucine of the LxLxL motif and likely reduces the ability of IAA10 to recruit the TOPLESS co-repressor [@lee_biological_2016].
-The nonsense variant (IAA10-Y219\*) resulting in truncated PB1 domain may be more indicative of *IAA10* null phenotype, as it likely prevents IAA10 from interacting with an ARF.
-This variant is only found in one accession from northeastern Spain (IP-Svi-0).
-These accessions and potential loss-of-function alleles may facilitate exploration of the function of *IAA10*.
-
-Several examples of possible gain-of-function missense variants were found in the degron of *IAA2*, *IAA3*, *IAA6*, *IAA7*, *IAA8*, *IAA13*, *IAA15*, *IAA18*, *IAA19*, *IAA27*, *IAA28* and *IAA29*.
-Most degron variants were very minor alleles in the population, only represented in one or two accessions.
-Interestingly, degron missense variants occured more frequently in *IAA* genes that have a closely related sister, which may indicate redundant function within the sister-pairs.
-*I think this is an interesting finding, but I'm not sure about this hypothesis of redundant function.
-If pairs are acting redundantly and we assume the degron variants are dominant, we would expect the auxin signal they transmit (their function) to be lower, in sum, as a pair.
-So I would actually hypothesize that there has been some divergence in function as a pair.
-*
-```{r Correlation between Distance to MRCA and prevalence of degron variants}
-
-```
-
-```{r Correlation between Pi_N/Pi_S and k_5}
-
-```
-
-*IAA6* has two potentially dominant missense mutations found within the degron (Supp Fig.
-alignment), one previously believed to lead to the *shy1-1* phenotype [@kim_two_1996].
-Interestingly, this variant is widely represented in 31 accessions across Europe (Supplemental Map Fig).
-This finding supports the proposition that iaa6-C78R may not be the mutation causing the *shy1-1D* phenotype but rather represents natural variation in the degron sequence ([note from Mike Prigge in TAIR](https://www.arabidopsis.org/servlets/TairObject?type=notepad&id=11247)).
-These accessions could be used to confirm whether *iaa6-C78R* is responsible for the *shy1-1D* phenotype.
-Additionally, an accession was identified with the IAA6-V77M variant, also in the degron, that may be useful to further explore *iaa6* gain-of-function phenotypes.
-
-### Identifying novel functional sequences in IAAs
-An exciting possible use for the Natural Variation Webtool is to search for conserved, low-diversity sites that may represent possible sites of functional import, such as post-translational modification or protein-protein interaction.
-For example, recent detailed biochemical analysis identified putative lysine ubiquitination sites in members of the *IAA6/19* sister pairs [@winkler_variation_2017]; Supp Fig IAA6/19 alignment).
-These sites are conserved across natural sequence variation, while several other lysine residues vary (Supp Figs alignment and map).
-In IAA6, there are 17 lysines of which ~6 are putative ubiquitination sites~ *I think this should be 5, K3, K32, K33, K91, and K97.
-Missense variants occured in K34Q, K140N *.
-One is proximal to the KR motif, while two others show natural variation.
-In IAA19, there are 15 total lysines, 8 of which are putative ubiquitination sites (K3, K25, K68, L87, K93, K100, K111, and K141).
-Two of these ubiquitination sites are proximal to the KR motif and degron.
-**This sentence is not clear.
-Only 1 of the ubiquitination sites varies?
-or one lysine, I think you mean this only 1 lysine that varies, but it is not ubiquitinated.
-Ubiquitylated Missense K129T** Natural sequence diversity may be used to narrow down a subset of lysines to assess for functional ubiquitination.
-
-```{r}
-
-```
-
-```{r IAA variants mapped to IAA17, eval=FALSE}
-coding_vcf <- ldply(.data = VCF, .fun = subset, 
-             !is.na(Transcript_ID) & gt_GT != "0|0")
-aln_df <- addSNPsToAlnDF(aln_df, coding_vcf, seq_name = Transcript_ID, 
-               seq_pos = Codon_Number)
-
-selections <- select_variants(gene_info, aln_df, structure_ref = "IAA17")
-selections$V2 <- paste0("chain X and resi ", selections$V2)
-write.table(selections, file = "IAA_variant_selections", row.names = FALSE, 
-          col.names = FALSE, sep = ",", quote = FALSE)
-```
-
-The members of the *IAA8/9/27* clade have a unique N-terminal extension with a highly conserved region common to orthologs in several other species [@dreher_arabidopsis_2006].
-*IAA8* and *IAA9* play redundant roles in leaf serration [@koenig_auxin_2009], while *IAA8* also plays an important role in lateral root development [@arase_iaa8_2012].
-IAA8 has been reported to uniquely localize in the cytosol [@arase_iaa8_2012] where it may interact with LSD1 protein [@coll_programmed_2011].
-The N-terminal extension has low natural sequence diversity in *IAA8* and *IAA9* (Supp Fig IAA8/9 alignment) that may represent a site of novel functionality, possibly mediating interaction with LSD1.
-The few accessions with missense alleles in this region in both *IAA8* and *IAA9* may be valuable in probing the functionality of this region (Supplemental Fig IAA8/9 map).
-
-```{r IAA8_9_27 alignment}
-
-```
-
-```{r IAA8_9 map}
-
-```
-
-Members of the *IAA17/7/14/16* clade have a second EAR motif (LxLxL) with no observed missense variants in *IAA7*, *IAA16*, or *IAA17* (Supp Fig IAA7/14/16/17 alignment).
-This observation supports the experimental evidence for IAA7 that the second EAR motif is functional, *i.e.* it can interact with TPR1 co-repressor and has a minor repressive role in auxin-related developmental phenotypes [lee_biological_2016].
-For *IAA14*, there were two missense variants that may be of interest to probe functionality of this second EAR motif.
-One variant, L41F, in the central lysine is represented in only one accession, and the other in the second “x” (K40N) is found in twelve accessions around Copenhagen (Supp Fig.
-IAA14 map).
-
-```{r IAA7_14_16_17 alignment}
-
-```
-
-```{r IAA14 map}
-
-```
-
-## *TPL/TPR* genes
-
-The Auxin signaling pathway utilizes the TOPLESS (TPL) and TOPLESS-related (TPR) family of Gro/TLE/TUP1 type co-repressor proteins to maintain auxin responsive genes in a transcriptionally-repressed state in the absence of auxin [@szemenyei_topless_2008].
-In *Arabidopsis thaliana* the five member *TPL/TPR* family includes *TPL* and *TPR1-4*.
-The resulting proteins are comprised of three structural domains: an N-terminal TPL domain and two WD-40 domains [@long_topless_2006].
-TPL/TPR proteins are recruited to the AUX/IAA proteins through interaction with the conserved Ethylene-responsive element binding factor-associated amphiphilic repression (EAR) domain [@szemenyei_topless_2008].
-Canonical EAR domains have the amino acid sequence LxLxL, as found in most AUX/IAAs (figure xA) [@overvoorde_functional_2005].
-TPL/TPR co-repressors bind EAR domains via their C-terminal to LisH (CTLH) domains found near their N-termini (citations of pre-structure founding papers/reviews).
-Recent structural analyses of the TPL N-terminal domain have highlighted the precise interaction interface between TPL and AUX/IAA EAR domains, as well as the TPL-TPL dimerization and tetramerization motifs [@martin-arevalillo_structure_2017; @ke_structural_2015a].
-The residues required for higher-order multimers of TPL tetramers have also been identified [@ma_d53_2017].
-Additional interactions with transcriptional regulation and chromatin modifying machinery are likely mediated by two tandem beta propeller domains of TPL/TPRs.
-These domain have not yet been crystallized, however numerous high resolution crystal structures of beta propeller domains allow for structural prediction of the TPL C-terminal beta propeller domains and the locations of likely interaction faces (Figure, X).
-
-The TOPLESS co-repressor family generally exhibits a high level of sequence conservation at the amino acid sequence level across resequenced *Arabidopsis thaliana* accessions, with all ${\pi_N/\pi_S}$ values below 1 (Table X).
-The closely related *TPL* and *TPR1* have the highest ${\pi_N/\pi_S}$ values (TPL-0.425, TPR1-0.739), suggesting that these these two related genes tolerate a higher degree of sequence and potentially functional diversity compared to *TPR2/3/4*.
-The N-terminal TPL domain of the TPL/TPR family is particularly conserved (Figure X).
-All nonsynonymous polymorphisms observed in this region are either in the coils between helices or are highly conservative mutations within helices (i.e.
-Valine to Isoleucine), which would be predicted to exhibit little effect on folding and function.
-
-```{r TPL SNP stats, fig.height=11}
-gene_info <- geneInfoFromFile(
-  fname = system.file("shiny-app", "gene_family_data",
-                      "TPL_gene_ids.csv", package = "r1001genomes"))
-#annotation <- readAnnotationFile(filename = system.file("shiny-app", "gene_family_data",
-#                      "IAA_annotations.csv", package = "r1001genomes"), gene_info = gene_info)
-
-VCF <- readRDS(system.file("shiny-app", "gene_family_data",
-                      "TPL_VCF.rds", package = "r1001genomes"))
-div_stats <- ldply(VCF, diversityStats, geneInfo=gene_info, .id="transcript_ID")
-div_stats$tair_symbol %<>% fct_reorder(.x = , div_stats$tair_symbol, 
-                                       .fun = function(x) 
-                                         {as.integer(str_remove(x, 
-                                              pattern = "TPR"))})
-div_stats$tair_symbol %<>% fct_relevel("TPL")
-div_stats %<>% arrange(tair_symbol)
-div_stats %<>% select(tair_symbol, everything())
-
-if (params$output == "html_document") {
-  formatRound(datatable(div_stats,
-                  colnames = c("transcript",
-                               "symbol",
-                               "&pi;<sub>N</sub>",
-                               "&pi;<sub>S</sub>",
-                               "&pi;<sub>N</sub>/&pi;<sub>S</sub>",
-                               "&pi; coding",
-                               "&pi; transcript"),
-                  rownames = FALSE, escape = FALSE,
-                  options = list(paging=FALSE, searching=FALSE)),
-                columns = 2:7, digits = 6)
-  } else 
-{
-  dt <- div_stats
-  names(dt) <- c("transcript",
-                                 "symbol",
-                                 "&pi;~N~",
-                                 "&pi;~S~",
-                                 "&pi;~N~/&pi;~S~",
-                                 "&pi; coding",
-                                 "&pi; transcript")
-  dt %<>% mutate_if(is.numeric, round, 6)
-  names(dt) <- paste0("**", names(dt), "**")
-  kable(dt) 
-}
-```
-
-The high degree of conservation in the entire N-terminal domain underscores its importance in TPL/TPR function.
-For example, the initial *tpl-1* mutation (N176H) in the ninth helix is a dominant gain-of-function allele [@long_topless_2006], which is capable of binding wild-type TPL protein and inducing protein aggregation [@ma_d53_2017].
-It is therefore understandable that this helix had very low diversity as nonsynonymous variants in this domain could act in a dominant negative fashion.
-Addtionally, the crystal structure of TPL has revealed that the The LisH domain and CRA domain coordinate TPL dimerization and tetramerization respectively (26601214, 28698367), the CTLH domain is critical for interaction with DNA-binding transcription factors (26601214), and helix nine within the CRA domain may be critical for the formation of higher-order TPL complexes (multimers of TPL tetramers, 28630893).
-
-Compared to the TPL domain, both the proline rich and WD-40 domains exhibited a higher frequency of missense variants.
-We utilized Phyre2.0 modeling software to predict the structure of the TPL/TPR WD-40 repeats in order to map non-synonymous mutations to protein structure (Figure X).
-Here we can see both faces of each predicted beta-propeller structure of the WD-40 domains, which are connected by a short peptide linker.
-This model reveals amino acid conservation in the core and central residues of the WD-40 domains.
-~~These observations could be used as the basis for forward mutagenesis strategies to test whether this region of TPL functions similarly to its homolog TLE, which binds its target genes EH1 and the C-terminal WRPW/Y motif (Hairy/Hes/Runx) in the center of the WD-40 domain (pmid 16762837).~~ **Can we say something more general here?
-There is not enough background to support this.
-What is TLE from?
-What is EH1?
-Is TLE really binding to EH1 DNA?**
-
-```{r TPL_TPR alignment}
-
-```
-
-Despite the presence of five *TPL/TPR* genes in the genome, which are considered to be redundant (pmid 16763149), very few observed SNPs are predicted to result in loss of function for any *TPL/TPR* gene.
-Only one frameshift variant was observed, *TPR4-665delG*, in a single accession (Aiell-4 from Italy).
-This deletion occurs in codon 222, just after the TPL N-terminal domain, and generates an early stop eleven codons downstream from the deletion.
-It is unclear whether this N-terminal fragment would be expressed at any appreciable level, as it is possible that the generation of an early stop codon would trigger nonsense-mediated decay.
-However, if expressed, it would be fascinating to examine whether this naturally occurring N-terminal truncation retains activity as a co-repressor.
-Given the observation that the TPL N-terminal domain is sufficient to bind histones (28630893), and repress transcription via synthetic assays in yeast (24979769), it is possible that this represents a semi-functional TPL variant.
-There has been no evidence for prioritization of individual *TPL/TPR* family members in the auxin response pathway, but experiments modulating the gene copy number of *TPR1* revealed the importance of maintaining the proper dosage of *TPL/TPR* in order to accurately trigger the salicylic acid innate immune response (20647385).
-These results, taken together with the absence of *TPL/TPR* loss-of-function alleles in the *Arabidopsis thaliana* accessions suggest that perhaps the entire family is being conserved at the nucleotide level in the population to meet a gene dosage requirement, with little opportunity for neo- or sub-functionalization.
-
-## ARF genes
+# *ARF* genes {#ARFs}
 
@@
                       "ARF_gene_ids.csv", package = "r1001genomes"))
-ARF_data %<>% left_join(y = div_stats$x$data, by = c("name" = "tair_symbol"))
+ARF_data %<>% left_join(y = div_stats, by = c("name" = "tair_symbol"))
 #tip_labels need to be in the first column
@@
 p
+#ggsave("ARF_tree.pdf", width = 6, height = 6)
 ```
@@
   # expand increases distance from axis
-  labs(x = "", y = "", fill = "variant") +
+  labs(x = "", y = "", fill = "") +
   theme_logo(base_family = "Helvetica") +
@@
 
-A few missense variants in ARF5 highlighted potential functional variation in the class A ARFs.
+A few missense variants in *ARF5* highlighted potential functional variation in the class A ARFs.
 A high frequency allele (*What are the variants?*), resulting in substitutions of small amino acids on the positive face of the PB1 domain for bulky phenylalanines, may disrupt interactions with other ARF or Aux/IAA PB1 domains.
@@
 
-*ARF23*, which has a truncated DNA-binding domain, has a high ${\pi_N/\pi_S}$ value of 4.1, and has many high-diversity nonsense variants, as may be expected of a gene that is pseudogenizing.
+*ARF23* has a truncated DNA-binding domain and had a high ${\pi_N/\pi_S}$ value of 4.1 and several high-frequency nonsense variants, as may be expected of a gene that is pseudogenizing.
 
@@
 
-ARF14, ARF15, ARF16, ARF20, and ARF21 all do not have one or multiple of the conserved acidic residues in the PB1 domain, and themselves form a subclade of the B ARFs.
+ARF14, ARF15, **ARF16 I don't think this is correct ARF 16 is class C**, ARF20, and ARF21 all do not have one or multiple of the conserved acidic residues in the PB1 domain, and themselves form a subclade of the B ARFs.
 Interestingly, no non-synonymous variation is seen in these amino acid residues of these ARFs, suggesting that the amino acids that are in the regions may still be important for the function of these proteins.
@@
 Structurally, all three members of Class C ARFs contain a canonical B3 DNA-binding domain, but only ARF10 and ARF16 contain a PB1 domain.
-The DNA binding domains exhibit overall low diversity and presence of largely synonymous changes for all the ARFs.
+The DNA binding domains exhibit overall low diversity.
 Only the ARF10 DNA binding domain contains missense variants (S147T and A157S) (Fig.
@@
 Arf16_PB_variants distribution).
+**Interesting that the PB1 domain of ARF10 is conserved and the DBD of ARF16 is conserved.
+It this suggestive of subfunctionalization in this family, where one retains DNA and/or dimerization specificity and the other retains specificity for IAAs?**
 
@@
 ARF10_16_17_Alignment).
-This is consistent with perturbation studies that have shown drastic developmental phenotype for class C ARFs [@wang_control_2005; @mallory_micrornadirected_2005; @liu_repression_2007].
+This is consistent with perturbation **what exactly does perturbation mean** studies that have shown drastic developmental phenotype for class C ARFs [@wang_control_2005; @mallory_micrornadirected_2005; @liu_repression_2007].
 For the class B ARFs (2, 3 and 4) the tasiRNA binding motif is also highly conserved suggesting strong regulatory constraints on these genes (Fig ARF2_3_4_Alignment).
@@
 ```
-
-#Discussion
-
-Analyzing natural diversity within the gene families constituting the auxin nuclear signaling pathway has confirmed existing knowledge, supported existing hypothses and generated new hypotheses.
-This analysis in particular has confirmed the importance of auxin-induced degradation of the Aux/IAA proteins as evidenced by the strong conservation of the *AFB* genes and the canonical degron region of the *Aux/IAA* genes.
-Despite this relative conservation, this analysis has also identified many accessions containing new alleles of interest in the *AFB* and *Aux/IAA* families.
-This analysis has also
-
-Importance of 
--AFB
--IAA degron
--TPL
--ARF DBD and PB1

### 2018-07-12 manuscript/07-discussion.Rmd
-Cleaned up manuscript folder -moved to bookdown format -revised all sections

--- 2018-07-10 manuscript/manuscript.Rmd
@@
----
-title: A comprehensive analysis of natural sequence variation within the <i>Arabidopsis
-  thaliana</i> nuclear auxin signaling pathway using an accessible web application
-author:
-  affiliation: University of Washington
-  name: R. Clay Wright
-date: "`r Sys.Date()`"
-params: 
-  output: word_document
-output:
-  word_document:
-    toc: yes
-  html_document:
-    fig_caption: yes
-    toc: yes
-always_allow_html: yes
-fontfamily: Helvetica
-fontsize: 11pt
-keywords: r markdown
-link-citations: yes
-csl: apa.csl
-bibliography:
-- natural_variation.bib
-- TIR1_AFB.bib
-- ARF_C.bib
-- AuxIAA_citations.bib
----
-
-```{r setup, message=FALSE, warning=FALSE}
-knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE, collapse = TRUE, comment = "#>", tidy = TRUE, cache = TRUE, cache.lazy = FALSE, out.width = "90%", fig.height=4, fig.width=7, out.extra='style="margin: auto; display: block; padding-top: 15px;"')
-
-#devtools::install_github("wrightrc/r1001genomes", ref = "auxin-natural-variation")
-library(ape)
-library(r1001genomes)
-library(DT)
-library(tidyverse)
-library(magrittr)
-library(RColorBrewer)
-library(ggpmisc)
-library(ggthemes)
-library(ggseqlogo)
-library(ggtree)
-library(reticulate)
-library(scales)
-library(viridis)
-library(ips)
-library(knitr)
-
-#use_python(python = "/usr/local/bin/python")
-```
-<!--This is a CSS (Cascading Style Sheet).
- It controls how panels and images are formatted.></!-->
-<style>
-  img {
-    display: block;
-    margin: auto;
-    width: 90%;
-    height: auto;
-  }
-</style>
-
-# List of working titles  
-* A comprehensive analysis of natural sequence variation within the <i>Arabidopsis thaliana</i> nuclear auxin signaling pathway  
-* **Add your suggestions**
-
-# Abstract
-
-* Motivating, high level statement  
-* Statement defining the cutting edge of your research area  
-* A statement of the form: "However, it remains unclear ..." that defines what problem you are  
-addressing.  
-* A statement of the form:"Here, we ..."  that concisely states your contribution.  
-* A series of 3?5 statement about your results.  
-* A concluding statement.
-
-# Intro
-
-The first genome sequence of *Arabidopsis thaliana* facilitated rapid advancement of plant biology through molecular genetics.
-Since this initial genome, massive scaling of sequencing technology has allowed a pioneering group to survey the global genomic variation in natural *Arabidopsis thaliana* populations.
-This valuable population genetics resource has led to several associations of genetic loci with phenotypic traits and provided insights into how selective pressure has and is influencing the evolution of plant genomes.
-Outside of genome-wide studies however, this valuable dataset has seen little use.
-Natural genomic variation also provides a set of feasible functional variation at the gene, gene family and gene network level that may provide insight into the function and evolution of genes, families and networks.
-Here, we present a web application and R-package through which plant molecular biologists with little-to-no bioinformatics experience can make use of this rich dataset of genetic variation to formulate hypotheses as to the sequence/function/phenotype relationships determined by the gene, family, or network of their interest.
-We demonstrate the utility of this tool through comprehensive analysis and identification of potential functional variation in the nuclear auxin signaling pathway.
-
-This application allows easy access to 
-1. lists of missense polymorphisms to facilitate biochemical assays of variant effects [@starita_variant_2017],
-2. family-wise alignments of variants to facilitate de novo functional domain identification [@melamed_combining_2015], 
-3.
-lists of accessions containing missense (or any type of) polymorphisms to facilitate segregation analysis and measurment of the distribution of variant effects on phenotype [@park_distributions_2017],
-4. as well as the study of gene and network evolution [@delker_natural_2010; @kliebenstein_role_2008].
-
-Here, we have utilized this tool to analyze the genetic variation in the nuclear auxin signaling pathway and formulate hypotheses regarding the functional implications of this variation and the evolution of the genes and gene families in this pathway.
-Our analysis also provided further confirmation of much of the existing knowledge of these genes demonstrating the validity of this approach.
-
-# Methods
-
-## Nucleotide Diversity Calculation
-Nei  defined the nucleotide diversity statistic in his original paper as:
-“…the average number of nucleotide differences per site between two randomly chosen DNA sequences.” (Nei, 1979)
-And provided the equation:
-
-$\displaystyle \pi = \sum_{ij}x_{i}x_{j}\pi_{ij}$
-
-Where Where $x_{i}$ is the frequency of the ith sequence in the population and $\pi_{ij}$ is the number of sites that are different between the ith and jth sequence divided by sequence length.
-
-a more general form, that treats each sequence in the population as unique can be written as:
-
-$\displaystyle \pi = \frac{1}{L*n^2} \sum_{i=1}^n \sum_{j=1}^n \sum_{k=1}^L{\pi_{ijk}}
-\qquad
-\displaystyle \pi_{ijk} = \left\{\begin{array}{l}
-1 \quad \textrm{if} \quad N_{ik} \neq N_{jk} \\
-0 \quad \textrm{if} \quad N_{ik} = N_{jk}
-\end{array}\right.$
-
-where $N_{ik}$ is the nucleotide (A, T, C or G) at position k on the ith sequence of the population.
-$L$ is the length of the sequence (the same for all sequences in the population, 
-assumes only SNPs exist, no indels). n is the total number of sequences in the population.
-
-from this form we can re-arrange summations to the form below:
-
-$\displaystyle \pi = \frac{1}{L} \sum_{k=1}^L{\pi_{k}}
-\qquad
-\pi_{k} = \frac{1}{n^2} \sum_{i=1}^n \sum_{j=1}^n {\pi_{ijk}}$
-
-here $\pi_k$ can be thought of as the site-wise nucleotide diversity at position k, and is equal to 
-the nucleotide diversity of a sequence of length 1 at location k. we can calculate $\pi_k$ for each
-site, then average those over the sequence length to calculate $\pi$, the nucleotide diversity 
-of the sequence.
-
-The function `Nucleotide_diversity` in the r1001genomes package calculates $\pi_k$ for each position in the gene or region that contains a variant.
-note, $\pi_k$ is equal to 0 at all locations without variants.
-This is also what is displayed in the Diversity Plot tab of the web app.
-
-## Detailed $\pi_k$ calculaiton simplification.
-
-The formula for $\pi_k$ above requires comparing every sequence to every other sequence at location k,
-however, we know there are only a few variant forms at each individual location.
-
-so we can revert back to using Nei's original formula, modifying it slightly,
-replacing $x_i$ with $\frac{n_i}{n}$, $n_i$ being the number of sequences in the population of the form $N_i$ at location k: subscripts i and k are summed only over unique forms [need to clarify this]
-
-$\displaystyle \pi_k = \sum_{ij} \frac{n_i}{n}\frac{n_j}{n} \pi_{ij} = 
-\frac{1}{n^2}\sum_{ij} n_i n_j \pi_{ij}$
-
-$\pi_{ij(k)} = \left\{\begin{array}{l}
-1 \quad \textrm{if} \quad i \neq j \\
-0 \quad \textrm{if} \quad i = j
-\end{array}\right.$
-
-we will define $n_{!i} = n - n_i$ as the number of sequences different from i at k.
-
-we can see the summed term will be zero if $i=j$, and $n_i n_j$ if $i\neq j$ so:
-
-$\displaystyle \pi_k = \frac{1}{n^2} \sum_{i} n_i n_{!i}$
-
-next we substitute our definition of $n_{!i}$
-
-$\displaystyle \pi_k = \frac{1}{n^2} \sum_{i} n_i (n - n_{i})$
-
-then distribute and split summation:
-
-$\displaystyle \pi_k = \frac{1}{n^2} (n \sum_{i} n_i - \sum_{i} n_{i}^2)$
-
-finally, summing $\displaystyle \sum_{i}n_i$ is equal to n:
-
-$\displaystyle \pi_k = \frac{1}{n^2} (n^2 - \sum_{i} n_{i}^2)$
-
-this simplified form for $\pi_k$ is used by the app, because the counts of unique 
-variants at a single nucleotide location can easily be summarized in R.
-
-# Results  
-## *TIR1/AFB* genes
-
-Auxin acts by binding to receptors that in turn target co-repressors for degradation.
-Auxin receptors (six in the model plant *Arabidopsis thaliana*) evolved through gene duplication and diversification early in the history of vascular plants [@Parry_Complex_2009].
-The rate of co-repressor degradation is determined by the identity of both the receptor and co-repressor [@Havens_Synthetic_2012], and this rate sets the pace of lateral root development [@Guseman_Auxin-induced_2015].
-
-Although it is unclear what unique roles each receptor plays in growth and development, a number of studies have pointed out differences in the ways the six different receptors in A.
-thaliana differ in biochemical function and expression domain[@dharmasiri_plant_2005; @parry_complex_2009; @prigge_arabidopsis_2016].
-Although TIR1/AFBs are expressed ubiquitously in A.
-thaliana tissues, TIR1, AFB2, and AFB3 have been shown to accumulate in the shoot and root meristems and leaf tissues, with slightly different expression patterns for TIR1 [@Parry_Complex_2009].
-Additionally, the expression of AFB5 is stronly circadian-regulated [@Covington_Circadian_2007] and AFB3 is more highly expressed in the roots in the presence of nitrate, allowing increased lateral root formation [@Vidal_Nitrate_2010], suggesting more broad environmental regulation of this gene family may exist.
-  Functionally, all members of this family have been shown to bind auxin and Aux/IAA proteins.
-However, AFB1 has drastically reduced ability to assemble into an SCF complex, due to the substitution E8K in its F-box domain, preventing it from inducing degradation of Aux/IAAs[@Yu_Untethering_2015].
-This lack of complexation may allow observed high ubiquitous AFB1 accumulation [@Parry_Complex_2009].
-Higher order mutants in the family containing *afb1* mutants suggest that *AFB1* has a moderate positive effect on auxin signaling.
-Additionally, AFB4 and AFB5 have been shown to preferentially and functionally bind the synthetic auxin picloram, while other family members preferentially bind indole-3-acetic acid [@Prigge_Arabidopsis_2016].
-Interestingly, the strength and rate with which TIR1/AFBs are able to bind and mark Aux/IAAs for degradation are variable [@calderonvillalobos_combinatorial_2012; @Havens_Synthetic_2012].
-AFB2 induced the degradation of certain Aux/IAA proteins at a faster rate than TIR1, suggesting some functional specificity has arisen since the initial duplication between the *TIR1/AFB1* and *AFB2/AFB3* clades.
-
-```{r TIR1_AFB_div_stats}
-gene_info <- geneInfoFromFile(
-  fname = system.file("shiny-app", "gene_family_data",
-                      "AFB_gene_ids.csv", package = "r1001genomes"))
-annotation <- readAnnotationFile(filename = system.file("shiny-app", "gene_family_data",
-                      "AFB_annotations.csv", package = "r1001genomes"), gene_info = gene_info)
-alignment <- alignCDS(gene_info$transcript_ID)
-# make an alignment data frame
-aln_df <- makeAlnDF(alignment[[2]])
-
-# make a collapsed alignment for translating positions
-collapsed_aln_df <- 
-  aln_df %>% dplyr::select(aln_pos, seq_pos, seq_name) %>%
-  tidyr::spread(key = seq_name, value = seq_pos)
-
-VCF <- readRDS(system.file("shiny-app", "gene_family_data",
-                      "AFBs_VCF.rds", package = "r1001genomes"))
-div_stats <- ldply(VCF, diversityStats, geneInfo=gene_info, .id="transcript_ID")
-div_stats$tair_symbol %<>% fct_reorder(.x =  div_stats$tair_symbol, 
-                                       .fun = function(x) 
-                                         {as.integer(str_remove(x, 
-                                              pattern = "AFB|TIR"))})
-div_stats$tair_symbol %<>% fct_relevel("TIR1")
-div_stats %<>% arrange(tair_symbol)
-div_stats %<>% select(tair_symbol, everything())
-
-if (params$output == "html_document") {
-  formatRound(datatable(div_stats,
-                  colnames = c("transcript",
-                               "symbol",
-                               "&pi;<sub>N</sub>",
-                               "&pi;<sub>S</sub>",
-                               "&pi;<sub>N</sub>/&pi;<sub>S</sub>",
-                               "&pi; coding",
-                               "&pi; transcript"),
-                  rownames = FALSE, escape = FALSE,
-                  options = list(paging=FALSE, searching=FALSE)),
-                columns = 2:7, digits = 6)
-  } else 
-{
-  dt <- div_stats
-  names(dt) <- c("transcript",
-                                 "symbol",
-                                 "&pi;~N~",
-                                 "&pi;~S~",
-                                 "&pi;~N~/&pi;~S~",
-                                 "&pi; coding",
-                                 "&pi; transcript")
-  dt %<>% mutate_if(is.numeric, round, 6)
-  names(dt) <- paste0("**", names(dt), "**")
-  kable(dt) 
-}
-```
-
-Examining the natural sequence variation across the *AFB* family revealed that *TIR1* and *AFB1* both had very low nonsynonymous diversity, hinting at their likely functional importance and bringing in to question the inconclusive role of *AFB1* in auxin signaling.
-*AFB3* and *AFB4* had higher nonsynonymous diversity, while their sister genes, *AFB2* and *AFB5* were more conserved.
-This matches our current understanding of *AFB3* as playing a minor role in the auxin signaling pathway and *AFB4* perhaps undergoing pseudogenization.
-
-Although most known functional regions are highly conserved in *AFB1*, there is a nonsynonymous polymorphism  in the oligomerization domain, only found in the Can-0 accession.
-Mutations in this domain of *TIR1* frequently have a semidominant effect on root phenotypes [@dezfulian_oligomerization_2016; @wright_insights_2017a].
-Characterization of this accession may identify potential *AFB1* phenotypes and may also test the hypothesis that AFB1 functions through oligomerization with other TIR1/AFBs.
-
-```{r AFB variants for PyMOL}
-coding_vcf <- ldply(.data = VCF, .fun = subset, 
-             !is.na(Transcript_ID) & gt_GT != "0|0")
-aln_df <- addSNPsToAlnDF(aln_df, coding_vcf, seq_name = Transcript_ID, 
-               seq_pos = Codon_Number)
-collapsed_aln_df %<>% left_join(y = subset(aln_df, seq_name == gene_info[gene_info$tair_symbol == "TIR1", "transcript_ID"]))
-
-#' Select variants for importing into PyMOL
-#'
-#' @param gene_info a gene info data frame
-#' @param aln_df an alignment data frame with attached SNPs
-#' @param structure_ref string, which tair_symbol or transcript_ID (depending on 
-#' \code{name_by_symbols}) corresponds to the protein structure to which 
-#' selections should be mapped
-#' @param effect_type string of the effect type to select from the alignment 
-#' data frame
-#' @param name_by_symbols logical, if true (default) tair_symbols will be used 
-#' to name the selections, otherwise transcript_IDs will be used
-#'
-#' @return
-#' @export
-#'
-#' @examples
-select_variants <- function(gene_info, aln_df, structure_ref, 
-                            effect_type = "missense", name_by_symbols = TRUE){
-  collapsed_aln_df <- aln_df %>% dplyr::select(aln_pos, seq_pos, seq_name) %>%
-    tidyr::spread(key = seq_name, value = seq_pos)
-  if(name_by_symbols)
-    column <- "tair_symbol"
-    else 
-      column <- "transcript_ID"
-
-  selections <- ldply(.data = gene_info[,column], .fun = function(gene){
-    collapsed_aln_df %<>% dplyr::left_join(y = subset(aln_df, seq_name ==  
-                      gene_info[gene_info[, column] == gene, "transcript_ID"]))
-    variants <- collapsed_aln_df[
-      grepl(effect_type, collapsed_aln_df$effects), 
-      gene_info[gene_info[, column] == structure_ref, "transcript_ID"]]
-    c(gene, paste0(na.omit(as.numeric(variants[[1]])), collapse = "+"))
-  })
-  return(selections)
-}
-
-selections <- select_variants(gene_info, aln_df, structure_ref = "TIR1")
-selections$V2 <- paste0("chain B and resi ", selections$V2)
-write.table(selections, file = "AFB_variant_selections", row.names = FALSE, 
-          col.names = FALSE, sep = ",", quote = FALSE)
-```
-
-```{python readSelection, eval=FALSE}
-#
-# -- readSelection.py
-#
-from pymol import cmd, stored
-
-def readSelection (fileName):
-	'''
-	DESCRIPTION
-
-	reads a csv file containing a column of selection names and a column of
-	selection strings and creates selection objects for each
-	'''
-	#read in the selection file
-	with open(fileName) as f:
-		lines = f.readlines()
-		for line in lines:
-			sel = line.split(",")
-			cmd.select(sel[0], sel[1])
-
-	f.close()
-
-cmd.extend('readSelection',readSelection);
-```
-
-```{PyMol, AFB_structure, eval=FALSE}
-#from within pymol
-cmd.load("Dropbox/180705_NIH_renewal/Aim1-figures/2P1Q.pdb")
-hide everything
-
-run ~/Desktop/natural-variation-webtool/manuscript/readSelection.py
-
-readSelection ~/Desktop/natural-variation-webtool/manuscript/AFB_variant_selections
-
-# Define selections
-# select TIR1variants, (chain B and resi 154+239+546)
-select fbox, (chain B and resi 1-50)
-select LRR, (chain B and resi 50-576)
-select TIR1, (chain B and resi 1-576)
-# select AFB2variants, (chain B and resi 38+174+181+184+209+259+401+496)
-select tir1_1, (chain B and resi 147)
-select TIR1dm, (chain B and resi 170+473)
-select insPcontacts, (chain B and resi 74+78+113+114+344+401+403+436+460+484+485+509)
-select dimerNULL, (chain B and resi 142+143+151+117+147)
-select auxin, resn IAC
-select AUXcontacts, (chain B and resi 78+79+82+378+380+403+404+405+438+439+462+463+464+489)
-select insP, resn IHP
-select IAA7, chain C
-select IAA7contacts, (chain B and resi 81+82+84+139+347+349+350+351+352+380+405+406+407+409+464+465+489)
-select IAA7_N, chain C and resi 1
-select IAA7_C, chain C and resi 13
-select ASK1, chain A
-select Y, chain C and resi 11
-select Ycontacts, chain B and resi 165+166+195
-
-#Define color palette
-#set_color name, [ red-float, green-float, blue-float ]
-set_color blue1, [0.00000000, 0.20392157, 0.6980392]
-set_color blue2, [0.09803922, 0.36078431, 1.0000000]
-set_color blue3, [0.00000000, 0.29019608, 1.0000000]
-set_color teal,  [0.09803922, 1.00000000, 0.6980392]
-set_color LRR,   [1.00000000, 0.71764706, 0.0000000]
-set_color Fbox,  [0.69803922, 0.49411765, 0.0000000]
-set_color TIR1l, [0.67058824, 0.07843137, 0.8000000]
-set_color TIR1f, [0.63529412, 0.34117647, 0.6980392]
-set_color AFB2l, [0.37647059, 0.70196078, 1.0000000]
-set_color AFB2f, [0.07843137, 0.64313725, 0.8000000]
-
-show cartoon, 2P1Q
-
-color lrr, LRR
-
-color TIR1l, TIR1
-show sticks, TIR1
-color TIR1f, AFB1
-show sticks, AFB1
-color AFB2l, AFB2
-show sticks, AFB2
-color AFB2f, AFB3
-show sticks, AFB3
-
-set ray_trace_mode, 3
-
-# used get_view() to return the current view (position) of the molecule 
-# copy the code to be able to set the identical viewpoint later
-# to return the current view from Pymol
-# get_view
-
-set_view (\
-     0.044671144,    0.083668359,   -0.995409012,\
-     0.774909019,    0.625905454,    0.087390237,\
-     0.630409002,   -0.775298953,   -0.036878210,\
-    -0.001916260,    0.005584979, -202.498916626,\
-    -5.538011074,   -8.507736206,   39.046234131,\
-  -219.237945557,  624.123413086,  -20.000000000 )
-
-#to render a high definition image use ray
-ray
-
-```
-
-```{r AFB1 diversity plot}
-coding_variants <- getCodingDiv(VCF[[gene_info[gene_info$tair_symbol == "AFB1", "transcript_ID"]]])
-p <- plotCodingDiv(uniqueCodingVars = coding_variants)
-annotation <- addAlnPosToAnno(anno_df = annotation, aln_df)
-
-# translate positions from TIR1 to AFB1
-AFB1_positions <- annotation$positions
-AFB1_positions %<>% left_join(y = select(collapsed_aln_df, aln_pos, gene_info[gene_info$tair_symbol == "AFB1", "transcript_ID"]), by = "aln_pos") 
-AFB1_positions$position <- AFB1_positions$AT4G03190.1
-
-p <- append_layers(p,list(
-        geom_rect(data = AFB1_positions,
-                   mapping = aes(xmin = as.integer(position)-1,
-                                 xmax = as.integer(position)+2,
-                                 fill = annotation),
-                  ymin = -Inf, ymax = Inf, inherit.aes = FALSE, alpha = 0.8)),
-      position = "bottom")
-p
-```
-
-The AFB4 and AFB5 receptors have an N-terminal extension prior to the F-box domains.
-This extension had very high nonsynonymous diversity(Fig.
-4), suggesting that this extension does not play an important functional role in these proteins.
-Additionally, two frameshift variants and one stop-gained variant were observed in *AFB4* supporting its pseudogenization.
-
-Figure 4: Alignment of A.
-thaliana TIR1/AFB F-box domains showing variants.
-Cyan marks the F-box binding domain.
-Yellow, green, and magenta, mark missense and synonymous, synonymous only, and missense only variant locations respectively.
-Yellow squared E’s show Cul1-interacting residues, differing in AFB1.
-```{r TIR1/AFB F-box Alignment}
-aln_df <- makeAlnDF(alignment[[2]])
-aln_df <- left_join(aln_df, dplyr::select(gene_info, "tair_locus",
-                                       "tair_symbol", "transcript_ID"),
-                        by = c("transcript_ID" = "transcript_ID"))
-aln_df$tair_symbol <- fct_rev(fct_relevel(aln_df$tair_symbol, "TIR1"))
-aln_df <- chunkAlnDF(aln_df = aln_df, chunk_width = 55)
-VCF <- ldply(.data = VCF, .fun = subset, gt_GT != "0|0")
-VCF <- getCodingDiv(VCF)
-aln_df <- addSNPsToAlnDF(aln_df, VCF)
-F_box <- subset(aln_df, aln_pos <= 96)
-F_box <- chunkAlnDF(aln_df = F_box, chunk_width = 80)
-p <-ggplot(F_box, aes(x = aln_pos, y = tair_symbol,
-                             group = seq_pos, text = variants))
-p <- p +
-  geom_tile(data = na.omit(F_box), mapping = aes(fill = effects),
-            width = 1, height = 1, alpha = 0.8) +
-  geom_text(aes(label=letter), alpha= 1, family = "Courier") +
-  scale_fill_brewer(type = "qual", palette = 1, direction = -1) +
-  scale_x_continuous(breaks=seq(1,max(F_box$aln_pos), by = 10)) +
-  scale_y_discrete() +
-  # expand increases distance from axis
-  labs(x = "", y = "", fill = "variant") +
-  theme_logo(base_family = "Helvetica") +
-  theme(panel.grid = element_blank(), panel.grid.minor = element_blank()) +
-  facet_wrap(facets = ~chunk, ncol = 1, scales = "free") +
-  theme(strip.background = element_blank(),
-        strip.text.x = element_blank(),
-        legend.box = "vertical")
-p
-```
-
-## *Aux/IAA* genes
-
-The Aux/IAA proteins (IAAs) mediate both auxin perception and transcriptional co-repression of auxin responsive genes.
-Auxin perception is carried out along with TIR1/AFB proteins, which together with IAAs act as co-receptors to bind and perceive nuclear auxin.
-Interaction of IAAs with auxin and TIR1/AFBs is conferred by the highly conserved 13 amino acid degron motif [@tan_mechanism_2007].
-Formation of this auxin co-receptor complex triggers ubiquitination and subsequent degradation of IAAs.
-This degradation in turn relieves their repression upon ARF transcription factors.
-Stabilizing mutations within the degron can lead to dominant phenotypes [@berleth_auxin_2004; @yang_iaa1_2004].
-The *Arabidopsis thaliana* Aux/IAA family has 29 members, most of which possess a canonical auxin degron.
-Notable exceptions are the auxin-insensitive IAA20 and IAA30 that lack a degron sequence entirely, while the closely related IAA31 retains a semi-functional degron sequence with a very slow degradation rate [@dreher_arabidopsis_2006; @havens_synthetic_2012].
-Additionally, IAA degradation rate is influenced by rate motifs found outside the degron, such as the KR motif [@dreher_arabidopsis_2006; @moss_rate_2015].
-IAA-mediated repression of ARFs depends upon both an EAR domain, N-terminal of the degron, that recruits co-repressors of the TOPLESS (TPL) family, and a Phox and Bem1 (PB1) domain that enables interaction via hetero-oligomerization with ARF PB1 domains [@nanao_structural_2014; @kagale_ear_2011].
-
-```{r IAA diversity stats, fig.height=11}
-gene_info <- geneInfoFromFile(
-  fname = system.file("shiny-app", "gene_family_data",
-                      "IAA_gene_ids.csv", package = "r1001genomes"))
-#annotation <- readAnnotationFile(filename = system.file("shiny-app", "gene_family_data",
-#                      "IAA_annotations.csv", package = "r1001genomes"), gene_info = gene_info)
-
-VCF <- readRDS(system.file("shiny-app", "gene_family_data",
-                      "IAAs_VCF.rds", package = "r1001genomes"))
-div_stats <- ldply(VCF, diversityStats, geneInfo=gene_info, .id="transcript_ID")
-div_stats$tair_symbol %<>% fct_reorder(.x = , div_stats$tair_symbol, 
-                                       .fun = function(x) 
-                                         {as.integer(str_remove(x, 
-                                              pattern = "IAA"))})
-div_stats %<>% arrange(tair_symbol)
-div_stats %<>% select(tair_symbol, everything())
-
-if (params$output == "html_document") {
-  formatRound(datatable(div_stats,
-                  colnames = c("transcript",
-                               "symbol",
-                               "&pi;<sub>N</sub>",
-                               "&pi;<sub>S</sub>",
-                               "&pi;<sub>N</sub>/&pi;<sub>S</sub>",
-                               "&pi; coding",
-                               "&pi; transcript"),
-                  rownames = FALSE, escape = FALSE,
-                  options = list(paging=FALSE, searching=FALSE)),
-                columns = 2:7, digits = 6)
-  } else 
-{
-  dt <- div_stats
-  names(dt) <- c("transcript",
-                                 "symbol",
-                                 "&pi;~N~",
-                                 "&pi;~S~",
-                                 "&pi;~N~/&pi;~S~",
-                                 "&pi; coding",
-                                 "&pi; transcript")
-  dt %<>% mutate_if(is.numeric, round, 6)
-  names(dt) <- paste0("**", names(dt), "**")
-  kable(dt) 
-}
-```
-
-Within the available natural sequence variation, we found several missense polymorphisms affecting the EAR domain or degron of several canonical IAAs.
-However, no missense polymorphisms were present in the KR motif or residues critical for PB1-PB1 domain interactions (Supp.
-Fig.
-Alignment of canonical IAAs).
-The majority of missense mutations observed were in the variable regions between the conserved domains.
-
-###Sister Pair Diversification
-
-Previous research has found evidence of both broad genetic redundancy and specificity within closely related pairs or groups of IAA proteins [@overvoorde_functional_2005].
-For example, the *iaa8-1 iaa9-1* double mutant and the *iaa5-1 iaa6-1 iaa19-1* triple mutant have wild-type phenotypes [@overvoorde_functional_2005], yet the *IAA6/19* sister pair has significant differences in expression patterns, protein abundances and functions suggesting they have undergone functional specialization since their divergence [@winkler_variation_2017].
-
-In our analysis of the *IAA* gene family, we frequently observed that one member of most *IAA* sister pairs had high nonsynonymous diversity while the other sister was more conserved (Figure IAA phylogeny heat map).
-In an interspecific comparison between *A.
-thaliana* and *A.
-lyrata*, *IAA19* is more conserved, while *IAA6* has regions of increased sequence divergence upstream of the degron and in the N-terminal half of the PB1 domain [@winkler_variation_2017].
-Consistent with these results, our analysis of intraspecific natural variation revealed higher conservation for *IAA19* (${\pi_N/\pi_S}$ = 0.55) compared to *IAA6* (${\pi_N/\pi_S}$ = 2.3) (Figure IAA phylogeny heat map), and also detected high diversity within the same regions of *IAA6* as seen in Winkler et al.
-(Supp Fig IAA6 diversity plot).
-```{r IAA alignment}
-alignment <- alignCDS(gene_info$transcript_ID)
-# make an alignment data frame
-aln_df <- makeAlnDF(alignment[[2]])
-
-# make a collapsed alignment for translating positions
-collapsed_aln_df <- 
-  aln_df %>% dplyr::select(aln_pos, seq_pos, seq_name) %>%
-  tidyr::spread(key = seq_name, value = seq_pos)
-```
-
-```{r Mask_IAA_alignment, eval=FALSE}
-aliscore(as.AAbin(alignment[[2]]), path = "/Volumes/SSD/Applications/ALISCORE_v2.0/Aliscore_v.2.0/")
-id <- scan(file = "input.fas_List_random.txt", sep = " ", quiet = TRUE)
-AAAlign_no_gaps <- as.matrix(alignment[[2]])[, id]
-write.nexus.data(AAAlign_no_gaps, file = "manuscript/IAAs.nex", format = "protein")
-```
-```{bash IAA_phylogeny_construction, eval=FALSE}
-~/Desktop/natural-variation-webtool/manuscript
-
-/Applications/MrBayes/mb
-
-execute IAAs.nex
-# set IAA33 as the outgroup
-outgroup AT5G57420.1
-
-lset nucmodel=protein Rates=invgamma
-# or include Nst=6  ngammacat=4 omegavar=M3 
-prset aamodelpr=fixed(jones) statefreqpr=fixed(empirical) 
-# initial run with aamodelpr=mixed yielded posterior prob of 1 for jones model
-propset ExtTBR$prob=0
-
-mcmcp ngen=250000 nchains=4 printfreq=1000 samplefreq=100 burnin=200
-
-mcmc
-sump
-sumt
-
-```
-
-```{r Load IAA phylogeny and data}
-tree <- treeio::read.beast(file = "IAAs.nex.con.tre")
-
-IAA_data <- read.csv("IAAs_table.csv")
-#tip_labels need to be in the first column
-IAA_data %<>% select(name, everything())
-IAA_data$name %<>% as.character()
-IAA_data$AGI %<>% as.character()
-
-IAA_data <- IAA_data[match(gsub(x = tree@phylo$tip.label, pattern = "\\.1",replacement = ""), IAA_data$AGI), ] 
-tree@phylo$tip.label <- IAA_data$name
-```
-*I think IAA20, 30, and 15 should also be in class C. So there are really two large clades A and B and then there are truncations/loss of function/dominant variants scattered throughout, Class C.*
-
-```{r IAA_tree, fig.height=5, fig.width=5, fig.cap="(ref:IAA tree)"}
-p <- ggtree(tree, layout = "rectangular", color = "grey") %<+% IAA_data + theme_tree() + scale_x_continuous(expand = c(0, 1.2)) + 
-  geom_tippoint(mapping = aes(size = 4*PiN_PiS, fill = PiN_PiS), shape = 21, color = "white", show.legend = TRUE) + 
-  scale_fill_viridis(values = rescale(c(min(IAA_data$PiN_PiS), 1,
-                                        max(IAA_data$PiN_PiS)), 
-                                      to = c(0,1)), guide = "colourbar") + scale_size(guide = "none") + 
-  theme(legend.position = c(.9,.7)) + 
-  geom_nodelab(mapping = aes(label = round(prob,digits = 2)), nudge_x = -.3, nudge_y = .5, size = 3) + 
-  geom_tiplab(mapping = aes(color = Class), hjust = -0.25) + 
-  scale_color_viridis(discrete = TRUE)
-p
-```
-
-(ref:IAA tree) IAA protein sequence tree mapped with ${\pi_N/\pi_S}$
-
-Within the Group B *IAA* genes, there are four sister pairs---*IAA10*/*IAA11*, *IAA12*/*IAA13*, *IAA18*/*IAA26*, and *IAA20*/*IAA30* \@ref(IAA tree) [@remington_contrasting_2004].
-Interestingly, among these sister pairs only *IAA12* and *IAA13* have a noticeable difference in nonsynonymous diversity.
-The *Arabidopsis thaliana* EFP browser shows that **IAA10** and *IAA11* have almost identical expression patterns, with the exception of *IAA10* being more highly expressed in the developing seed, and *IAA11* being more highly expressed in root tissue during lateral root development **need to cite original data source here**.
-In our analysis, *IAA10* and *IAA11* both showed functional conservation (${\pi_N/\pi_S}$ of 0.80 and 0.67 respectively) in support of *IAA10* and *IAA11* playing redundant roles.
-Accessions and associated variants identified here provide a potentially expedited route to elucidating the extent of redundancy between family members such as *IAA10* and *IAA11*.
-
-### New Loss-of-Function/Gain-of-Function alleles
-Accessions with functional variants can be potentially used to investigate the extent of redundancy/specialization within sister pairs.
-For example, three accessions have a missense variant affecting the EAR domain of IAA10 (HR-5, UKSE06-118 and UKSW06-240, all from southern UK).
-This variant (L49S) is in the second leucine of the LxLxL motif and likely reduces the ability of IAA10 to recruit the TOPLESS co-repressor [@lee_biological_2016].
-The nonsense variant (IAA10-Y219\*) resulting in truncated PB1 domain may be more indicative of *IAA10* null phenotype, as it likely prevents IAA10 from interacting with an ARF.
-This variant is only found in one accession from northeastern Spain (IP-Svi-0).
-These accessions and potential loss-of-function alleles may facilitate exploration of the function of *IAA10*.
-
-Several examples of possible gain-of-function missense variants were found in the degron of *IAA2*, *IAA3*, *IAA6*, *IAA7*, *IAA8*, *IAA13*, *IAA15*, *IAA18*, *IAA19*, *IAA27*, *IAA28* and *IAA29*.
-Most degron variants were very minor alleles in the population, only represented in one or two accessions.
-Interestingly, degron missense variants occured more frequently in *IAA* genes that have a closely related sister, which may indicate redundant function within the sister-pairs.
-*I think this is an interesting finding, but I'm not sure about this hypothesis of redundant function.
-If pairs are acting redundantly and we assume the degron variants are dominant, we would expect the auxin signal they transmit (their function) to be lower, in sum, as a pair.
-So I would actually hypothesize that there has been some divergence in function as a pair.
-*
-```{r Correlation between Distance to MRCA and prevalence of degron variants}
-
-```
-
-```{r Correlation between Pi_N/Pi_S and k_5}
-
-```
-
-*IAA6* has two potentially dominant missense mutations found within the degron (Supp Fig.
-alignment), one previously believed to lead to the *shy1-1* phenotype [@kim_two_1996].
-Interestingly, this variant is widely represented in 31 accessions across Europe (Supplemental Map Fig).
-This finding supports the proposition that iaa6-C78R may not be the mutation causing the *shy1-1D* phenotype but rather represents natural variation in the degron sequence ([note from Mike Prigge in TAIR](https://www.arabidopsis.org/servlets/TairObject?type=notepad&id=11247)).
-These accessions could be used to confirm whether *iaa6-C78R* is responsible for the *shy1-1D* phenotype.
-Additionally, an accession was identified with the IAA6-V77M variant, also in the degron, that may be useful to further explore *iaa6* gain-of-function phenotypes.
-
-### Identifying novel functional sequences in IAAs
-An exciting possible use for the Natural Variation Webtool is to search for conserved, low-diversity sites that may represent possible sites of functional import, such as post-translational modification or protein-protein interaction.
-For example, recent detailed biochemical analysis identified putative lysine ubiquitination sites in members of the *IAA6/19* sister pairs [@winkler_variation_2017]; Supp Fig IAA6/19 alignment).
-These sites are conserved across natural sequence variation, while several other lysine residues vary (Supp Figs alignment and map).
-In IAA6, there are 17 lysines of which ~6 are putative ubiquitination sites~ *I think this should be 5, K3, K32, K33, K91, and K97.
-Missense variants occured in K34Q, K140N *.
-One is proximal to the KR motif, while two others show natural variation.
-In IAA19, there are 15 total lysines, 8 of which are putative ubiquitination sites (K3, K25, K68, L87, K93, K100, K111, and K141).
-Two of these ubiquitination sites are proximal to the KR motif and degron.
-**This sentence is not clear.
-Only 1 of the ubiquitination sites varies?
-or one lysine, I think you mean this only 1 lysine that varies, but it is not ubiquitinated.
-Ubiquitylated Missense K129T** Natural sequence diversity may be used to narrow down a subset of lysines to assess for functional ubiquitination.
-
-```{r}
-
-```
-
-```{r IAA variants mapped to IAA17, eval=FALSE}
-coding_vcf <- ldply(.data = VCF, .fun = subset, 
-             !is.na(Transcript_ID) & gt_GT != "0|0")
-aln_df <- addSNPsToAlnDF(aln_df, coding_vcf, seq_name = Transcript_ID, 
-               seq_pos = Codon_Number)
-
-selections <- select_variants(gene_info, aln_df, structure_ref = "IAA17")
-selections$V2 <- paste0("chain X and resi ", selections$V2)
-write.table(selections, file = "IAA_variant_selections", row.names = FALSE, 
-          col.names = FALSE, sep = ",", quote = FALSE)
-```
-
-The members of the *IAA8/9/27* clade have a unique N-terminal extension with a highly conserved region common to orthologs in several other species [@dreher_arabidopsis_2006].
-*IAA8* and *IAA9* play redundant roles in leaf serration [@koenig_auxin_2009], while *IAA8* also plays an important role in lateral root development [@arase_iaa8_2012].
-IAA8 has been reported to uniquely localize in the cytosol [@arase_iaa8_2012] where it may interact with LSD1 protein [@coll_programmed_2011].
-The N-terminal extension has low natural sequence diversity in *IAA8* and *IAA9* (Supp Fig IAA8/9 alignment) that may represent a site of novel functionality, possibly mediating interaction with LSD1.
-The few accessions with missense alleles in this region in both *IAA8* and *IAA9* may be valuable in probing the functionality of this region (Supplemental Fig IAA8/9 map).
-
-```{r IAA8_9_27 alignment}
-
-```
-
-```{r IAA8_9 map}
-
-```
-
-Members of the *IAA17/7/14/16* clade have a second EAR motif (LxLxL) with no observed missense variants in *IAA7*, *IAA16*, or *IAA17* (Supp Fig IAA7/14/16/17 alignment).
-This observation supports the experimental evidence for IAA7 that the second EAR motif is functional, *i.e.* it can interact with TPR1 co-repressor and has a minor repressive role in auxin-related developmental phenotypes [lee_biological_2016].
-For *IAA14*, there were two missense variants that may be of interest to probe functionality of this second EAR motif.
-One variant, L41F, in the central lysine is represented in only one accession, and the other in the second “x” (K40N) is found in twelve accessions around Copenhagen (Supp Fig.
-IAA14 map).
-
-```{r IAA7_14_16_17 alignment}
-
-```
-
-```{r IAA14 map}
-
-```
-
-## *TPL/TPR* genes
-
-The Auxin signaling pathway utilizes the TOPLESS (TPL) and TOPLESS-related (TPR) family of Gro/TLE/TUP1 type co-repressor proteins to maintain auxin responsive genes in a transcriptionally-repressed state in the absence of auxin [@szemenyei_topless_2008].
-In *Arabidopsis thaliana* the five member *TPL/TPR* family includes *TPL* and *TPR1-4*.
-The resulting proteins are comprised of three structural domains: an N-terminal TPL domain and two WD-40 domains [@long_topless_2006].
-TPL/TPR proteins are recruited to the AUX/IAA proteins through interaction with the conserved Ethylene-responsive element binding factor-associated amphiphilic repression (EAR) domain [@szemenyei_topless_2008].
-Canonical EAR domains have the amino acid sequence LxLxL, as found in most AUX/IAAs (figure xA) [@overvoorde_functional_2005].
-TPL/TPR co-repressors bind EAR domains via their C-terminal to LisH (CTLH) domains found near their N-termini (citations of pre-structure founding papers/reviews).
-Recent structural analyses of the TPL N-terminal domain have highlighted the precise interaction interface between TPL and AUX/IAA EAR domains, as well as the TPL-TPL dimerization and tetramerization motifs [@martin-arevalillo_structure_2017; @ke_structural_2015a].
-The residues required for higher-order multimers of TPL tetramers have also been identified [@ma_d53_2017].
-Additional interactions with transcriptional regulation and chromatin modifying machinery are likely mediated by two tandem beta propeller domains of TPL/TPRs.
-These domain have not yet been crystallized, however numerous high resolution crystal structures of beta propeller domains allow for structural prediction of the TPL C-terminal beta propeller domains and the locations of likely interaction faces (Figure, X).
-
-The TOPLESS co-repressor family generally exhibits a high level of sequence conservation at the amino acid sequence level across resequenced *Arabidopsis thaliana* accessions, with all ${\pi_N/\pi_S}$ values below 1 (Table X).
-The closely related *TPL* and *TPR1* have the highest ${\pi_N/\pi_S}$ values (TPL-0.425, TPR1-0.739), suggesting that these these two related genes tolerate a higher degree of sequence and potentially functional diversity compared to *TPR2/3/4*.
-The N-terminal TPL domain of the TPL/TPR family is particularly conserved (Figure X).
-All nonsynonymous polymorphisms observed in this region are either in the coils between helices or are highly conservative mutations within helices (i.e.
-Valine to Isoleucine), which would be predicted to exhibit little effect on folding and function.
-
-```{r TPL SNP stats, fig.height=11}
-gene_info <- geneInfoFromFile(
-  fname = system.file("shiny-app", "gene_family_data",
-                      "TPL_gene_ids.csv", package = "r1001genomes"))
-#annotation <- readAnnotationFile(filename = system.file("shiny-app", "gene_family_data",
-#                      "IAA_annotations.csv", package = "r1001genomes"), gene_info = gene_info)
-
-VCF <- readRDS(system.file("shiny-app", "gene_family_data",
-                      "TPL_VCF.rds", package = "r1001genomes"))
-div_stats <- ldply(VCF, diversityStats, geneInfo=gene_info, .id="transcript_ID")
-div_stats$tair_symbol %<>% fct_reorder(.x = , div_stats$tair_symbol, 
-                                       .fun = function(x) 
-                                         {as.integer(str_remove(x, 
-                                              pattern = "TPR"))})
-div_stats$tair_symbol %<>% fct_relevel("TPL")
-div_stats %<>% arrange(tair_symbol)
-div_stats %<>% select(tair_symbol, everything())
-
-if (params$output == "html_document") {
-  formatRound(datatable(div_stats,
-                  colnames = c("transcript",
-                               "symbol",
-                               "&pi;<sub>N</sub>",
-                               "&pi;<sub>S</sub>",
-                               "&pi;<sub>N</sub>/&pi;<sub>S</sub>",
-                               "&pi; coding",
-                               "&pi; transcript"),
-                  rownames = FALSE, escape = FALSE,
-                  options = list(paging=FALSE, searching=FALSE)),
-                columns = 2:7, digits = 6)
-  } else 
-{
-  dt <- div_stats
-  names(dt) <- c("transcript",
-                                 "symbol",
-                                 "&pi;~N~",
-                                 "&pi;~S~",
-                                 "&pi;~N~/&pi;~S~",
-                                 "&pi; coding",
-                                 "&pi; transcript")
-  dt %<>% mutate_if(is.numeric, round, 6)
-  names(dt) <- paste0("**", names(dt), "**")
-  kable(dt) 
-}
-```
-
-The high degree of conservation in the entire N-terminal domain underscores its importance in TPL/TPR function.
-For example, the initial *tpl-1* mutation (N176H) in the ninth helix is a dominant gain-of-function allele [@long_topless_2006], which is capable of binding wild-type TPL protein and inducing protein aggregation [@ma_d53_2017].
-It is therefore understandable that this helix had very low diversity as nonsynonymous variants in this domain could act in a dominant negative fashion.
-Addtionally, the crystal structure of TPL has revealed that the The LisH domain and CRA domain coordinate TPL dimerization and tetramerization respectively (26601214, 28698367), the CTLH domain is critical for interaction with DNA-binding transcription factors (26601214), and helix nine within the CRA domain may be critical for the formation of higher-order TPL complexes (multimers of TPL tetramers, 28630893).
-
-Compared to the TPL domain, both the proline rich and WD-40 domains exhibited a higher frequency of missense variants.
-We utilized Phyre2.0 modeling software to predict the structure of the TPL/TPR WD-40 repeats in order to map non-synonymous mutations to protein structure (Figure X).
-Here we can see both faces of each predicted beta-propeller structure of the WD-40 domains, which are connected by a short peptide linker.
-This model reveals amino acid conservation in the core and central residues of the WD-40 domains.
-~~These observations could be used as the basis for forward mutagenesis strategies to test whether this region of TPL functions similarly to its homolog TLE, which binds its target genes EH1 and the C-terminal WRPW/Y motif (Hairy/Hes/Runx) in the center of the WD-40 domain (pmid 16762837).~~ **Can we say something more general here?
-There is not enough background to support this.
-What is TLE from?
-What is EH1?
-Is TLE really binding to EH1 DNA?**
-
-```{r TPL_TPR alignment}
-
-```
-
-Despite the presence of five *TPL/TPR* genes in the genome, which are considered to be redundant (pmid 16763149), very few observed SNPs are predicted to result in loss of function for any *TPL/TPR* gene.
-Only one frameshift variant was observed, *TPR4-665delG*, in a single accession (Aiell-4 from Italy).
-This deletion occurs in codon 222, just after the TPL N-terminal domain, and generates an early stop eleven codons downstream from the deletion.
-It is unclear whether this N-terminal fragment would be expressed at any appreciable level, as it is possible that the generation of an early stop codon would trigger nonsense-mediated decay.
-However, if expressed, it would be fascinating to examine whether this naturally occurring N-terminal truncation retains activity as a co-repressor.
-Given the observation that the TPL N-terminal domain is sufficient to bind histones (28630893), and repress transcription via synthetic assays in yeast (24979769), it is possible that this represents a semi-functional TPL variant.
-There has been no evidence for prioritization of individual *TPL/TPR* family members in the auxin response pathway, but experiments modulating the gene copy number of *TPR1* revealed the importance of maintaining the proper dosage of *TPL/TPR* in order to accurately trigger the salicylic acid innate immune response (20647385).
-These results, taken together with the absence of *TPL/TPR* loss-of-function alleles in the *Arabidopsis thaliana* accessions suggest that perhaps the entire family is being conserved at the nucleotide level in the population to meet a gene dosage requirement, with little opportunity for neo- or sub-functionalization.
-
-## ARF genes
-
-Auxin response is ultimately mediated by the auxin responsive transcription factors (ARFs).
-There are 23 ARFs in *Arabidopsis thaliana* that are divided into three phylogenetic classes.
-Class A ARFs (ARF5, ARF6, ARF7, ARF8 and ARF19) activate transcription.
-These ARFs have a glutamine-rich region in the middle of the protein that may mediate activation (Guilefoyle, 2007).
-It has recently been shown that the middle region of ARF5 interacts with the SWI/SNF chromatin remodeling ATPases BRAMA and SPLAYED, possibly to reduce nucleosome occupancy and allow for the recruitment of transcription machinery (Wu, 2016).
-Additionally, ARF7 interacts with Mediator subunits, directly tethering transcriptional activation machinery to its binding sites in the chromosome (Ito, 2016).
-Class B and C ARFs are historically categorized as repressor ARFs, though the mechanism through which they confer repression has not been identified.
-Their middle regions tend to be proline- and serine-rich (citation, Ulmasov?).
-
-Canonical ARFs are comprised of three major domains.
-Recent crystallization of these domains have informed structure-function analysis of the ARFs (Boer, 2014; Nanao, 2014; Korasick, 2014).
-These domains are conserved throughout land plants (Dolf's new paper).
-ARFs share an N-terminal B3 DNA binding domain.
-Flanking this DNA-binding domain is a dimerization domain, which folds up into a single “taco-shaped” domain to allow for dimerization between ARFs.
-There is an auxiliary domain that immediately follows and interacts with the dimerization domain.
-The middle region is the most variable between ARFs, as mentioned above, but is characterized by repetitive units of glutamine (class A), serine, or proline residues (classes B and C).
-
-The C-terminal domain of the ARF is a protein-protein interaction domain mediating interactions among ARFs, between ARFs and other transcription factors, and between ARFs and the Aux/IAA repressors.
-This interaction domain was recently characterized as a Phox and Bem1 (PB1) domain, which is comprised of a positive and negative face with conserved basic and acidic residues, respectively (Nanao, 2014; Korasick, 2014).
-The dipolar nature of the PB1 domain may mediate multimerization by the pairwise interaction of these faces on different proteins as the ARF7 PB1 domain was crystallized as a multimer [@korasick_].
-However, it is unclear whether ARF multimerization occurs or plays a developmental significant role in vivo.
-Interfering with ARF dimerization in either the DNA-binding proximal dimerization domain or the PB1 domain decreases the ability of class A ARFs to activate transcription in a heterologous yeast system (Pierre-Jerome, 2016).
-
-While domain architecture is broadly conserved among the ARFs, there are exceptional cases.
-Three ARFs do not contain a PB1 domain at all, ARF3, ARF13, and ARF17, and several more have lost the conserved acidic or basic residues in the PB1 domain, suggesting they may be reduced to a single interaction domain.
-Several ARFs additionally have an expanded conserved region within the DNA-binding domain, of unknown function.
-The majority of domain variation among ARFs occurs in the large B-class subfamily.
-The liverwort *Marchantia polymorpha* has a single representative ARF of each class (Flores-Sandoval, 2015).
-The expansion of these classes in flowering plants is the result of both whole genome and tandem duplication events (Remington, 2004).
-The growth of the ARF family may have allowed for the expansion of the quantity and complexity of loci regulated by the ARFs and subsequent expansion in their regulation of developmental processes.
-
-Class A ARFs are the most well-studied ARF subfamily---the five family members all act as transcriptional activators and have well-characterized, distinct developmental targets.
-Overall the diversity of class A ARFs was generally low, especially compared to the class B and C ARFs, suggesting that class A ARFs are central to auxin signal transduction and plant development.
-Analysis of class A ARF nonsynonymous diversity suggests that the majority of these ARFs are highly functionally conserved, with ${\pi_N/\pi_S}$ values much lower than 1 with the exception of ARF19, with ${\pi_N/\pi_S}$ value of 1.8.
-Comparing diversity within sister pairs, there is a similar trade-off as seen in most *IAA* sister pairs, with one sister being highly conserved and the other more divergent.
-ARF19 and ARF8 are the more divergent class A ARFs, with ${\pi_N/\pi_S}$ values at least three time those of their sisters, ARF7 and ARF6 respectively.
-This may suggest that ARF6 and ARF7 serve more essential purposes in plant development.
-
-```{r ARF diversity stats}
-gene_info <- geneInfoFromFile(
-  fname = system.file("shiny-app", "gene_family_data",
-                      "ARF_gene_ids.csv", package = "r1001genomes"))
-#annotation <- readAnnotationFile(filename = system.file("shiny-app", 
-  #"gene_family_data", "ARF_annotations.csv", 
-  #package = "r1001genomes"), gene_info = gene_info)
-
-VCF <- readRDS(system.file("shiny-app", "gene_family_data",
-                      "ARF_VCF.rds", package = "r1001genomes"))
-div_stats <- ldply(VCF, diversityStats, geneInfo=gene_info, .id="transcript_ID")
-div_stats$tair_symbol %<>% fct_reorder(.x =  div_stats$tair_symbol, 
-                                       .fun = function(x) 
-                                         {as.integer(str_remove(x, 
-                                              pattern = "ARF"))})
-div_stats %<>% arrange(tair_symbol)
-div_stats %<>% select(tair_symbol, everything())
-
-if (params$output == "html_document") {
-  formatRound(datatable(div_stats,
-                  colnames = c("transcript",
-                               "symbol",
-                               "&pi;<sub>N</sub>",
-                               "&pi;<sub>S</sub>",
-                               "&pi;<sub>N</sub>/&pi;<sub>S</sub>",
-                               "&pi; coding",
-                               "&pi; transcript"),
-                  rownames = FALSE, escape = FALSE,
-                  options = list(paging=FALSE, searching=FALSE)),
-                columns = 2:7, digits = 6)
-  } else 
-{
-  dt <- div_stats
-  names(dt) <- c("transcript",
-                                 "symbol",
-                                 "&pi;~N~",
-                                 "&pi;~S~",
-                                 "&pi;~N~/&pi;~S~",
-                                 "&pi; coding",
-                                 "&pi; transcript")
-  dt %<>% mutate_if(is.numeric, round, 6)
-  names(dt) <- paste0("**", names(dt), "**")
-  kable(dt) 
-}
-```
-
-```{r ARF alignment}
-alignment <- alignCDS(gene_info$transcript_ID)
-# make an alignment data frame
-aln_df <- makeAlnDF(alignment[[2]])
-
-# make a collapsed alignment for translating positions
-collapsed_aln_df <- 
-  aln_df %>% dplyr::select(aln_pos, seq_pos, seq_name) %>%
-  tidyr::spread(key = seq_name, value = seq_pos)
-```
-
-```{r Mask_ARF_alignment, eval=FALSE}
-aliscore(as.AAbin(alignment[[2]]), path = "/Volumes/SSD/Applications/ALISCORE_v2.0/Aliscore_v.2.0/")
-id <- scan(file = "/Volumes/SSD/Applications/ALISCORE_v2.0/Aliscore_v.2.0/input.fas_List_random.txt", sep = " ", quiet = TRUE)
-AAAlign_no_gaps <- as.matrix(alignment[[2]])[, id]
-write.nexus.data(AAAlign_no_gaps, file = "ARFs.nex", format = "protein")
-```
-```{bash ARF_phylogeny_construction, eval=FALSE}
-/Applications/MrBayes/mb
-
-execute ARFs.nex
-# set ARF17 as the outgroup
-outgroup AT1G77850.1
-
-lset nucmodel=protein Rates=invgamma
-# or include Nst=6  ngammacat=4 omegavar=M3 
-prset aamodelpr=fixed(jones) statefreqpr=fixed(empirical) 
-# initial run with aamodelpr=mixed yielded posterior prob of 1 for jones model
-propset ExtTBR$prob=0
-
-mcmcp ngen=250000 nchains=4 printfreq=1000 samplefreq=100 burnin=200
-
-mcmc
-sump
-sumt
-
-```
-
-```{r Load ARF phylogeny and data}
-tree <- treeio::read.beast(file = "ARFs.nex.con.tre")
-
-ARF_data <- read.csv(system.file("shiny-app", "gene_family_data",
-                      "ARF_gene_ids.csv", package = "r1001genomes"))
-ARF_data %<>% left_join(y = div_stats$x$data, by = c("name" = "tair_symbol"))
-#tip_labels need to be in the first column
-ARF_data %<>% select(name, everything())
-ARF_data$tair_locus %<>% as.character()
-
-ARF_data <- ARF_data[match(gsub(x = tree@phylo$tip.label, pattern = "\\.1",replacement = ""), ARF_data$tair_locus), ]
-
-tree@phylo$tip.label <- ARF_data$name
-
-```
-
-```{r Visualize ARF tree, fig.height=5, fig.width=5}
-p <- ggtree(tree, layout = "rectangular", color = "grey") %<+% ARF_data + theme_tree() + scale_x_continuous(expand = c(0, 1.2)) + geom_tiplab(hjust = -0.25) + 
-  geom_tippoint(mapping = aes(size = 4*Pi_NS_Ratio, fill = Pi_NS_Ratio), 
-                shape = 21, color = "white", show.legend = TRUE) + 
-  scale_fill_viridis(values = rescale(c(min(ARF_data$Pi_NS_Ratio), 1,
-                                        max(ARF_data$Pi_NS_Ratio)), 
-                                      to = c(0,1)), guide = "colourbar") + scale_size(guide = "none") + 
-  theme(legend.position = c(.1,.7)) + 
-  geom_nodelab(mapping = aes(label = round(prob,digits = 2)), nudge_x = -.3, nudge_y = .5, size = 3) +
-  geom_tiplab(mapping = aes(color = class), hjust = -0.25) + 
-  scale_color_viridis(discrete = TRUE)
-p
-```
-
-### Class A ARFs 
-For all class A ARFs, the middle region of the protein was the predominant high diversity region.
-The DNA-binding domain had very few, low-diversity missense mutations, as did the PB1 domain.
-Considering the necessity of their conserved functions, the low level of variation in these key DNA and protein-protein interaction domains is expected.
-
-```{r A ARF Alignment}
-gene_info <- subset(gene_info, tair_symbol %in% c("ARF5", "ARF6", "ARF7", "ARF8", "ARF19"))
-alignment <- alignCDS(gene_info$transcript_ID)
-# make an alignment data frame
-aln_df <- makeAlnDF(alignment[[2]])
-#BrowseSeqs(alignment[[2]])
-aln_df <- makeAlnDF(alignment[[2]])
-aln_df <- left_join(aln_df, dplyr::select(gene_info, "tair_locus",
-                                       "tair_symbol", "transcript_ID"),
-                        by = c("transcript_ID" = "transcript_ID"))
-aln_df$tair_symbol %<>% as.factor()
-aln_df$tair_symbol %<>% fct_reorder(
-    {as.integer(str_remove(aln_df$tair_symbol, 
-                           pattern = "ARF"))}, 
-  .desc = TRUE)
-
-aln_df <- chunkAlnDF(aln_df = aln_df, chunk_width = 100)
-VCF <- ldply(.data = VCF, .fun = subset, gt_GT != "0|0")
-VCF <- getCodingDiv(VCF)
-aln_df <- addSNPsToAlnDF(aln_df, VCF)
-A_ARFs <- subset(aln_df, tair_symbol %in% c("ARF5", "ARF6", "ARF7", "ARF8", "ARF19"))
-p <-ggplot(A_ARFs, aes(x = aln_pos, y = tair_symbol,
-                             group = seq_pos, text = variants))
-p <- p +
-  geom_raster(data = na.omit(A_ARFs), mapping = aes(fill = effects)) +
-  scale_fill_viridis(option = "A", discrete = TRUE) +
-  scale_y_discrete() +
-  # expand increases distance from axis
-  labs(x = "", y = "", fill = "variant") +
-  theme_logo(base_family = "Helvetica") +
-  theme(panel.grid = element_blank(), panel.grid.minor = element_blank()) +
-  theme(strip.background = element_blank(),
-        strip.text.x = element_blank(),
-        legend.box = "vertical")
-p
-#ggsave(p,filename = "A_ARFs_alignment_variants.pdf", width = 12, height = 4)
-```
-
-Class A ARFs all contain the canonical B3 DNA-binding domain, PB1 interaction domain and glutamine-rich middle region (Guilefoyle, 2007).
-In the analyzed natural variation, ARF7 had several expansions of polyglutamine sequences in the middle region.
-Polyglutamine regions are known to readily expand and contract throughout evolutionary time due to replication error, and variation in polyglutamine length can be acted on by natural selection and have phenotypic consequences (Press, 2014).
-
-A few missense variants in ARF5 highlighted potential functional variation in the class A ARFs.
-A high frequency allele (*What are the variants?*), resulting in substitutions of small amino acids on the positive face of the PB1 domain for bulky phenylalanines, may disrupt interactions with other ARF or Aux/IAA PB1 domains.
-Interestingly, these variants were always found together in accessions Tu-B2-3, Ru-2, Haes-1, Rd-0, 627ME-13Y1, Le-0, Bu-0 and Bsch-0 *Where are these?*.
-Another highly represented polymorphism in ARF5 results in change in the 17th amino acid from a leucine to a methionine.
-This position is in the N-terminal extension of ARF5 which is of unknown function.
-It is possible that this methionine acts as an alternative start site in the accessions.
-
-### Class B ARFs
-
-*ARF23*, which has a truncated DNA-binding domain, has a high ${\pi_N/\pi_S}$ value of 4.1, and has many high-diversity nonsense variants, as may be expected of a gene that is pseudogenizing.
-
-ARF3, ARF13 and ARF17 all lack a C-terminal PB1 domain, which suggests they may be unable to interact with Aux/IAA repressors, possibly acting as auxin-insensitive competitive inhibitors of ARF-mediated transcription (Fig ARF3_13_17 alignment).
-These ARFs span the range of ${\pi_N/\pi_S}$ ratios relative to the other ARFs, ranging from 0.85 to 4.1.
-ARF17 is more conserved than the activator ARF, ARF19.
-ARF13 has many nonsense variants of high diversity and may be pseudogenizing.
-
-```{r ARF3_13_17 alignment}
-
-```
-
-ARF14, ARF15, ARF16, ARF20, and ARF21 all do not have one or multiple of the conserved acidic residues in the PB1 domain, and themselves form a subclade of the B ARFs.
-Interestingly, no non-synonymous variation is seen in these amino acid residues of these ARFs, suggesting that the amino acids that are in the regions may still be important for the function of these proteins.
-
-### Class C ARFs
-
-Class C ARFs show low nucleotide diversity scores, all ${\pi_N/\pi_S}$ values substantially lower than 1.
-ARF16 was the most conserved, whereas its clade members (ARF10, ARF17) had scores at least four times higher [@remington_contrasting_2004].
-
-Structurally, all three members of Class C ARFs contain a canonical B3 DNA-binding domain, but only ARF10 and ARF16 contain a PB1 domain.
-The DNA binding domains exhibit overall low diversity and presence of largely synonymous changes for all the ARFs.
-Only the ARF10 DNA binding domain contains missense variants (S147T and A157S) (Fig.
-ARF10_DBD Variants distribution).
-**Any insights into whether these may affect function?
-They are quite conservative.** Of the PB1 domain containing class C ARFs, ARF16 exhibits several missense variants which are sporadically distributed, in contrast to the conserved PB1 domain of ARF10 (Fig.
-Arf16_PB_variants distribution).
-
-```{r ARF10_DBD Variants distribution}
-
-```
-
-```{r Arf16_PB_variants distribution}
-
-```
-
-Several members of class B (ARF2, 3 and 4) and C ARFs (ARF10, 16 and 17) are distinguished from other ARFs on the basis of their post transcriptional regulation by trans-acting small interfering RNAs (tasiRNAs) and microRNAs (miRNAs) [@rhoades_prediction_2002; @allen_micrornadirected_2005].
-**(@AMY Arf 6 and 8 targeted by miRNA 167 - check for conservation).** The miRNA target of class C ARFs showed a high degree of conservation (Fig.
-ARF10_16_17_Alignment).
-This is consistent with perturbation studies that have shown drastic developmental phenotype for class C ARFs [@wang_control_2005; @mallory_micrornadirected_2005; @liu_repression_2007].
-For the class B ARFs (2, 3 and 4) the tasiRNA binding motif is also highly conserved suggesting strong regulatory constraints on these genes (Fig ARF2_3_4_Alignment).
-The DNA binding domain of ARF4 exhibits non-synonymous variation in 4 accession and would be interesting to carry out phenotypic characterization of these accessions for developmental growth defects (Fig ARF4_DBD Variants map).
-
-```{r ARF10_16_17_Alignment}
-
-```
-
-```{r ARF2_3_4_Alignment}
-
-```
-
-```{r ARF4_DBD Variants map}
-
-```
-
 #Discussion

### 2018-07-12 manuscript/AFB_section.Rmd
-Cleaned up manuscript folder -moved to bookdown format -revised all sections

--- 2018-07-10 manuscript/manuscript.Rmd
@@
     toc: yes
+    fig_caption: yes
+    reference_docx: "template.docx"
   html_document:
@@
 
-# Methods
-
-## Nucleotide Diversity Calculation
-Nei  defined the nucleotide diversity statistic in his original paper as:
-“…the average number of nucleotide differences per site between two randomly chosen DNA sequences.” (Nei, 1979)
-And provided the equation:
-
-$\displaystyle \pi = \sum_{ij}x_{i}x_{j}\pi_{ij}$
-
-Where Where $x_{i}$ is the frequency of the ith sequence in the population and $\pi_{ij}$ is the number of sites that are different between the ith and jth sequence divided by sequence length.
-
-a more general form, that treats each sequence in the population as unique can be written as:
-
-$\displaystyle \pi = \frac{1}{L*n^2} \sum_{i=1}^n \sum_{j=1}^n \sum_{k=1}^L{\pi_{ijk}}
-\qquad
-\displaystyle \pi_{ijk} = \left\{\begin{array}{l}
-1 \quad \textrm{if} \quad N_{ik} \neq N_{jk} \\
-0 \quad \textrm{if} \quad N_{ik} = N_{jk}
-\end{array}\right.$
-
-where $N_{ik}$ is the nucleotide (A, T, C or G) at position k on the ith sequence of the population.
-$L$ is the length of the sequence (the same for all sequences in the population, 
-assumes only SNPs exist, no indels). n is the total number of sequences in the population.
-
-from this form we can re-arrange summations to the form below:
-
-$\displaystyle \pi = \frac{1}{L} \sum_{k=1}^L{\pi_{k}}
-\qquad
-\pi_{k} = \frac{1}{n^2} \sum_{i=1}^n \sum_{j=1}^n {\pi_{ijk}}$
-
-here $\pi_k$ can be thought of as the site-wise nucleotide diversity at position k, and is equal to 
-the nucleotide diversity of a sequence of length 1 at location k. we can calculate $\pi_k$ for each
-site, then average those over the sequence length to calculate $\pi$, the nucleotide diversity 
-of the sequence.
-
-The function `Nucleotide_diversity` in the r1001genomes package calculates $\pi_k$ for each position in the gene or region that contains a variant.
-note, $\pi_k$ is equal to 0 at all locations without variants.
-This is also what is displayed in the Diversity Plot tab of the web app.
-
-## Detailed $\pi_k$ calculaiton simplification.
-
-The formula for $\pi_k$ above requires comparing every sequence to every other sequence at location k,
-however, we know there are only a few variant forms at each individual location.
-
-so we can revert back to using Nei's original formula, modifying it slightly,
-replacing $x_i$ with $\frac{n_i}{n}$, $n_i$ being the number of sequences in the population of the form $N_i$ at location k: subscripts i and k are summed only over unique forms [need to clarify this]
-
-$\displaystyle \pi_k = \sum_{ij} \frac{n_i}{n}\frac{n_j}{n} \pi_{ij} = 
-\frac{1}{n^2}\sum_{ij} n_i n_j \pi_{ij}$
-
-$\pi_{ij(k)} = \left\{\begin{array}{l}
-1 \quad \textrm{if} \quad i \neq j \\
-0 \quad \textrm{if} \quad i = j
-\end{array}\right.$
-
-we will define $n_{!i} = n - n_i$ as the number of sequences different from i at k.
-
-we can see the summed term will be zero if $i=j$, and $n_i n_j$ if $i\neq j$ so:
-
-$\displaystyle \pi_k = \frac{1}{n^2} \sum_{i} n_i n_{!i}$
-
-next we substitute our definition of $n_{!i}$
-
-$\displaystyle \pi_k = \frac{1}{n^2} \sum_{i} n_i (n - n_{i})$
-
-then distribute and split summation:
-
-$\displaystyle \pi_k = \frac{1}{n^2} (n \sum_{i} n_i - \sum_{i} n_{i}^2)$
-
-finally, summing $\displaystyle \sum_{i}n_i$ is equal to n:
-
-$\displaystyle \pi_k = \frac{1}{n^2} (n^2 - \sum_{i} n_{i}^2)$
-
-this simplified form for $\pi_k$ is used by the app, because the counts of unique 
-variants at a single nucleotide location can easily be summarized in R.
-
 # Results  
@@
 
+(ref:TIR1_AFB_div_stats)
+
 ```{r TIR1_AFB_div_stats}
@@
 
-Figure 4: Alignment of A.
+(ref:AFB_F-box_align) Alignment of A.
 thaliana TIR1/AFB F-box domains showing variants.
@@
 Yellow squared E’s show Cul1-interacting residues, differing in AFB1.
-```{r TIR1/AFB F-box Alignment}
+
+```{r AFB_F-box_align, fig.cap='(ref:AFB_F-box_align)'}
 aln_df <- makeAlnDF(alignment[[2]])

### 2018-07-12 manuscript/index.Rmd
-Cleaned up manuscript folder -moved to bookdown format -revised all sections

--- 2018-07-12 manuscript/AFB_section.Rmd
@@
 --- 
-title: A comprehensive analysis of natural sequence variation within the <i>Arabidopsis
-  thaliana</i> nuclear auxin signaling pathway using an accessible web application
-author:
-  affiliation: University of Washington
-  name: R. Clay Wright
+title: "A Minimal Book Example"
+author: "Yihui Xie"
 date: "`r Sys.Date()`"
-params: 
-  output: word_document
-output:
-  word_document:
-    toc: yes
-    fig_caption: yes
-    reference_docx: "template.docx"
-  html_document:
-    fig_caption: yes
-    toc: yes
-always_allow_html: yes
-fontfamily: Helvetica
-fontsize: 11pt
-keywords: r markdown
+site: bookdown::bookdown_site
+documentclass: book
+bibliography: [book.bib, packages.bib]
+biblio-style: apalike
 link-citations: yes
-csl: apa.csl
-bibliography:
-- natural_variation.bib
-- TIR1_AFB.bib
-- ARF_C.bib
-- AuxIAA_citations.bib
+description: "This is a minimal example of using the bookdown package to write a book. The output format for this example is bookdown::gitbook."
 ---
@@
 ```
-<!--This is a CSS (Cascading Style Sheet).
- It controls how panels and images are formatted.></!-->
-<style>
-  img {
-    display: block;
-    margin: auto;
-    width: 90%;
-    height: auto;
-  }
-</style>
 
-# List of working titles  
+## List of working titles  
 * A comprehensive analysis of natural sequence variation within the <i>Arabidopsis thaliana</i> nuclear auxin signaling pathway  
@@
 
-# Intro
-
-The first genome sequence of *Arabidopsis thaliana* facilitated rapid advancement of plant biology through molecular genetics.
-Since this initial genome, massive scaling of sequencing technology has allowed a pioneering group to survey the global genomic variation in natural *Arabidopsis thaliana* populations.
-This valuable population genetics resource has led to several associations of genetic loci with phenotypic traits and provided insights into how selective pressure has and is influencing the evolution of plant genomes.
-Outside of genome-wide studies however, this valuable dataset has seen little use.
-Natural genomic variation also provides a set of feasible functional variation at the gene, gene family and gene network level that may provide insight into the function and evolution of genes, families and networks.
-Here, we present a web application and R-package through which plant molecular biologists with little-to-no bioinformatics experience can make use of this rich dataset of genetic variation to formulate hypotheses as to the sequence/function/phenotype relationships determined by the gene, family, or network of their interest.
-We demonstrate the utility of this tool through comprehensive analysis and identification of potential functional variation in the nuclear auxin signaling pathway.
-
-This application allows easy access to 
-1. lists of missense polymorphisms to facilitate biochemical assays of variant effects [@starita_variant_2017],
-2. family-wise alignments of variants to facilitate de novo functional domain identification [@melamed_combining_2015], 
-3.
-lists of accessions containing missense (or any type of) polymorphisms to facilitate segregation analysis and measurment of the distribution of variant effects on phenotype [@park_distributions_2017],
-4. as well as the study of gene and network evolution [@delker_natural_2010; @kliebenstein_role_2008].
-
-Here, we have utilized this tool to analyze the genetic variation in the nuclear auxin signaling pathway and formulate hypotheses regarding the functional implications of this variation and the evolution of the genes and gene families in this pathway.
-Our analysis also provided further confirmation of much of the existing knowledge of these genes demonstrating the validity of this approach.
-
-# Results  
-## *TIR1/AFB* genes
-
-Auxin acts by binding to receptors that in turn target co-repressors for degradation.
-Auxin receptors (six in the model plant *Arabidopsis thaliana*) evolved through gene duplication and diversification early in the history of vascular plants [@Parry_Complex_2009].
-The rate of co-repressor degradation is determined by the identity of both the receptor and co-repressor [@Havens_Synthetic_2012], and this rate sets the pace of lateral root development [@Guseman_Auxin-induced_2015].
-
-Although it is unclear what unique roles each receptor plays in growth and development, a number of studies have pointed out differences in the ways the six different receptors in A.
-thaliana differ in biochemical function and expression domain[@dharmasiri_plant_2005; @parry_complex_2009; @prigge_arabidopsis_2016].
-Although TIR1/AFBs are expressed ubiquitously in A.
-thaliana tissues, TIR1, AFB2, and AFB3 have been shown to accumulate in the shoot and root meristems and leaf tissues, with slightly different expression patterns for TIR1 [@Parry_Complex_2009].
-Additionally, the expression of AFB5 is stronly circadian-regulated [@Covington_Circadian_2007] and AFB3 is more highly expressed in the roots in the presence of nitrate, allowing increased lateral root formation [@Vidal_Nitrate_2010], suggesting more broad environmental regulation of this gene family may exist.
-  Functionally, all members of this family have been shown to bind auxin and Aux/IAA proteins.
-However, AFB1 has drastically reduced ability to assemble into an SCF complex, due to the substitution E8K in its F-box domain, preventing it from inducing degradation of Aux/IAAs[@Yu_Untethering_2015].
-This lack of complexation may allow observed high ubiquitous AFB1 accumulation [@Parry_Complex_2009].
-Higher order mutants in the family containing *afb1* mutants suggest that *AFB1* has a moderate positive effect on auxin signaling.
-Additionally, AFB4 and AFB5 have been shown to preferentially and functionally bind the synthetic auxin picloram, while other family members preferentially bind indole-3-acetic acid [@Prigge_Arabidopsis_2016].
-Interestingly, the strength and rate with which TIR1/AFBs are able to bind and mark Aux/IAAs for degradation are variable [@calderonvillalobos_combinatorial_2012; @Havens_Synthetic_2012].
-AFB2 induced the degradation of certain Aux/IAA proteins at a faster rate than TIR1, suggesting some functional specificity has arisen since the initial duplication between the *TIR1/AFB1* and *AFB2/AFB3* clades.
-
-(ref:TIR1_AFB_div_stats)
-
-```{r TIR1_AFB_div_stats}
-gene_info <- geneInfoFromFile(
-  fname = system.file("shiny-app", "gene_family_data",
-                      "AFB_gene_ids.csv", package = "r1001genomes"))
-
-VCF <- readRDS(system.file("shiny-app", "gene_family_data",
-                      "AFBs_VCF.rds", package = "r1001genomes"))
-div_stats <- ldply(VCF, diversityStats, geneInfo=gene_info, .id="transcript_ID")
-div_stats$tair_symbol %<>% fct_reorder(.x =  div_stats$tair_symbol, 
-                                       .fun = function(x) 
-                                         {as.integer(str_remove(x, 
-                                              pattern = "AFB|TIR"))})
-div_stats$tair_symbol %<>% fct_relevel("TIR1")
-div_stats %<>% arrange(tair_symbol)
-div_stats %<>% select(tair_symbol, everything())
-
-if (params$output == "html_document") {
-  formatRound(datatable(div_stats,
-                  colnames = c("transcript",
-                               "symbol",
-                               "&pi;<sub>N</sub>",
-                               "&pi;<sub>S</sub>",
-                               "&pi;<sub>N</sub>/&pi;<sub>S</sub>",
-                               "&pi; coding",
-                               "&pi; transcript"),
-                  rownames = FALSE, escape = FALSE,
-                  options = list(paging=FALSE, searching=FALSE)),
-                columns = 2:7, digits = 6)
-  } else 
-{
-  dt <- div_stats
-  names(dt) <- c("transcript",
-                                 "symbol",
-                                 "&pi;~N~",
-                                 "&pi;~S~",
-                                 "&pi;~N~/&pi;~S~",
-                                 "&pi; coding",
-                                 "&pi; transcript")
-  dt %<>% mutate_if(is.numeric, round, 6)
-  names(dt) <- paste0("**", names(dt), "**")
-  kable(dt) 
-}
+```{r include=FALSE}
+# automatically create a bib database for R packages
+knitr::write_bib(c(
+  .packages(), 'bookdown', 'knitr', 'rmarkdown'
+), 'packages.bib')
 ```
-
-Examining the natural sequence variation across the *AFB* family revealed that *TIR1* and *AFB1* both had very low nonsynonymous diversity, hinting at their likely functional importance and bringing in to question the inconclusive role of *AFB1* in auxin signaling.
-*AFB3* and *AFB4* had higher nonsynonymous diversity, while their sister genes, *AFB2* and *AFB5* were more conserved.
-This matches our current understanding of *AFB3* as playing a minor role in the auxin signaling pathway and *AFB4* perhaps undergoing pseudogenization.
-
-Although most known functional regions are highly conserved in *AFB1*, there is a nonsynonymous polymorphism  in the oligomerization domain, only found in the Can-0 accession.
-Mutations in this domain of *TIR1* frequently have a semidominant effect on root phenotypes [@dezfulian_oligomerization_2016; @wright_insights_2017a].
-Characterization of this accession may identify potential *AFB1* phenotypes and may also test the hypothesis that AFB1 functions through oligomerization with other TIR1/AFBs.
-
-```{r AFB variants for PyMOL}
-alignment <- alignCDS(gene_info$transcript_ID)
-# make an alignment data frame
-aln_df <- makeAlnDF(alignment[[2]])
-
-# make a collapsed alignment for translating positions
-collapsed_aln_df <- 
-  aln_df %>% dplyr::select(aln_pos, seq_pos, seq_name) %>%
-  tidyr::spread(key = seq_name, value = seq_pos)
-coding_vcf <- ldply(.data = VCF, .fun = subset, 
-             !is.na(Transcript_ID) & gt_GT != "0|0")
-aln_df <- addSNPsToAlnDF(aln_df, coding_vcf, seq_name = Transcript_ID, 
-               seq_pos = Codon_Number)
-collapsed_aln_df %<>% left_join(y = subset(aln_df, seq_name == gene_info[gene_info$tair_symbol == "TIR1", "transcript_ID"]))
-
-#' Select variants for importing into PyMOL
-#'
-#' @param gene_info a gene info data frame
-#' @param aln_df an alignment data frame with attached SNPs
-#' @param structure_ref string, which tair_symbol or transcript_ID (depending on 
-#' \code{name_by_symbols}) corresponds to the protein structure to which 
-#' selections should be mapped
-#' @param effect_type string of the effect type to select from the alignment 
-#' data frame
-#' @param name_by_symbols logical, if true (default) tair_symbols will be used 
-#' to name the selections, otherwise transcript_IDs will be used
-#'
-#' @return
-#' @export
-#'
-#' @examples
-select_variants <- function(gene_info, aln_df, structure_ref, 
-                            effect_type = "missense", name_by_symbols = TRUE){
-  collapsed_aln_df <- aln_df %>% dplyr::select(aln_pos, seq_pos, seq_name) %>%
-    tidyr::spread(key = seq_name, value = seq_pos)
-  if(name_by_symbols)
-    column <- "tair_symbol"
-    else 
-      column <- "transcript_ID"
-
-  selections <- ldply(.data = gene_info[,column], .fun = function(gene){
-    collapsed_aln_df %<>% dplyr::left_join(y = subset(aln_df, seq_name ==  
-                      gene_info[gene_info[, column] == gene, "transcript_ID"]))
-    variants <- collapsed_aln_df[
-      grepl(effect_type, collapsed_aln_df$effects), 
-      gene_info[gene_info[, column] == structure_ref, "transcript_ID"]]
-    c(gene, paste0(na.omit(as.numeric(variants[[1]])), collapse = "+"))
-  })
-  return(selections)
-}
-
-selections <- select_variants(gene_info, aln_df, structure_ref = "TIR1")
-selections$V2 <- paste0("chain B and resi ", selections$V2)
-write.table(selections, file = "AFB_variant_selections", row.names = FALSE, 
-          col.names = FALSE, sep = ",", quote = FALSE)
-```
-
-```{python readSelection, eval=FALSE}
-#
-# -- readSelection.py
-#
-from pymol import cmd, stored
-
-def readSelection (fileName):
-	'''
-	DESCRIPTION
-
-	reads a csv file containing a column of selection names and a column of
-	selection strings and creates selection objects for each
-	'''
-	#read in the selection file
-	with open(fileName) as f:
-		lines = f.readlines()
-		for line in lines:
-			sel = line.split(",")
-			cmd.select(sel[0], sel[1])
-
-	f.close()
-
-cmd.extend('readSelection',readSelection);
-```
-
-```{PyMol, AFB_structure, eval=FALSE}
-#from within pymol
-cmd.load("Dropbox/180705_NIH_renewal/Aim1-figures/2P1Q.pdb")
-hide everything
-
-run ~/Desktop/natural-variation-webtool/manuscript/readSelection.py
-
-readSelection ~/Desktop/natural-variation-webtool/manuscript/AFB_variant_selections
-
-# Define selections
-# select TIR1variants, (chain B and resi 154+239+546)
-select fbox, (chain B and resi 1-50)
-select LRR, (chain B and resi 50-576)
-select TIR1, (chain B and resi 1-576)
-# select AFB2variants, (chain B and resi 38+174+181+184+209+259+401+496)
-select tir1_1, (chain B and resi 147)
-select TIR1dm, (chain B and resi 170+473)
-select insPcontacts, (chain B and resi 74+78+113+114+344+401+403+436+460+484+485+509)
-select dimerNULL, (chain B and resi 142+143+151+117+147)
-select auxin, resn IAC
-select AUXcontacts, (chain B and resi 78+79+82+378+380+403+404+405+438+439+462+463+464+489)
-select insP, resn IHP
-select IAA7, chain C
-select IAA7contacts, (chain B and resi 81+82+84+139+347+349+350+351+352+380+405+406+407+409+464+465+489)
-select IAA7_N, chain C and resi 1
-select IAA7_C, chain C and resi 13
-select ASK1, chain A
-select Y, chain C and resi 11
-select Ycontacts, chain B and resi 165+166+195
-
-#Define color palette
-#set_color name, [ red-float, green-float, blue-float ]
-set_color blue1, [0.00000000, 0.20392157, 0.6980392]
-set_color blue2, [0.09803922, 0.36078431, 1.0000000]
-set_color blue3, [0.00000000, 0.29019608, 1.0000000]
-set_color teal,  [0.09803922, 1.00000000, 0.6980392]
-set_color LRR,   [1.00000000, 0.71764706, 0.0000000]
-set_color Fbox,  [0.69803922, 0.49411765, 0.0000000]
-set_color TIR1l, [0.67058824, 0.07843137, 0.8000000]
-set_color TIR1f, [0.63529412, 0.34117647, 0.6980392]
-set_color AFB2l, [0.37647059, 0.70196078, 1.0000000]
-set_color AFB2f, [0.07843137, 0.64313725, 0.8000000]
-
-show cartoon, 2P1Q
-
-color lrr, LRR
-
-color TIR1l, TIR1
-show sticks, TIR1
-color TIR1f, AFB1
-show sticks, AFB1
-color AFB2l, AFB2
-show sticks, AFB2
-color AFB2f, AFB3
-show sticks, AFB3
-
-set ray_trace_mode, 3
-
-# used get_view() to return the current view (position) of the molecule 
-# copy the code to be able to set the identical viewpoint later
-# to return the current view from Pymol
-# get_view
-
-set_view (\
-     0.044671144,    0.083668359,   -0.995409012,\
-     0.774909019,    0.625905454,    0.087390237,\
-     0.630409002,   -0.775298953,   -0.036878210,\
-    -0.001916260,    0.005584979, -202.498916626,\
-    -5.538011074,   -8.507736206,   39.046234131,\
-  -219.237945557,  624.123413086,  -20.000000000 )
-
-#to render a high definition image use ray
-ray
-
-```
-
-```{r AFB1 diversity plot}
-coding_variants <- getCodingDiv(VCF[[gene_info[gene_info$tair_symbol == "AFB1", "transcript_ID"]]])
-p <- plotCodingDiv(uniqueCodingVars = coding_variants)
-annotation <- readAnnotationFile(filename = system.file("shiny-app", "gene_family_data",
-                      "AFB_annotations.csv", package = "r1001genomes"), gene_info = gene_info)
-annotation <- addAlnPosToAnno(anno_df = annotation, aln_df)
-
-# translate positions from TIR1 to AFB1
-AFB1_positions <- annotation$positions
-AFB1_positions %<>% left_join(y = select(collapsed_aln_df, aln_pos, gene_info[gene_info$tair_symbol == "AFB1", "transcript_ID"]), by = "aln_pos") 
-AFB1_positions$position <- AFB1_positions$AT4G03190.1
-
-p <- append_layers(p,list(
-        geom_rect(data = AFB1_positions,
-                   mapping = aes(xmin = as.integer(position)-1,
-                                 xmax = as.integer(position)+2,
-                                 fill = annotation),
-                  ymin = -Inf, ymax = Inf, inherit.aes = FALSE, alpha = 0.8)),
-      position = "bottom")
-p
-```
-
-The AFB4 and AFB5 receptors have an N-terminal extension prior to the F-box domains.
-This extension had very high nonsynonymous diversity(Fig.
-4), suggesting that this extension does not play an important functional role in these proteins.
-Additionally, two frameshift variants and one stop-gained variant were observed in *AFB4* supporting its pseudogenization.
-
-(ref:AFB_F-box_align) Alignment of A.
-thaliana TIR1/AFB F-box domains showing variants.
-Cyan marks the F-box binding domain.
-Yellow, green, and magenta, mark missense and synonymous, synonymous only, and missense only variant locations respectively.
-Yellow squared E’s show Cul1-interacting residues, differing in AFB1.
-
-```{r AFB_F-box_align, fig.cap='(ref:AFB_F-box_align)'}
-aln_df <- makeAlnDF(alignment[[2]])
-aln_df <- left_join(aln_df, dplyr::select(gene_info, "tair_locus",
-                                       "tair_symbol", "transcript_ID"),
-                        by = c("transcript_ID" = "transcript_ID"))
-aln_df$tair_symbol <- fct_rev(fct_relevel(aln_df$tair_symbol, "TIR1"))
-aln_df <- chunkAlnDF(aln_df = aln_df, chunk_width = 55)
-VCF <- ldply(.data = VCF, .fun = subset, gt_GT != "0|0")
-VCF <- getCodingDiv(VCF)
-aln_df <- addSNPsToAlnDF(aln_df, VCF)
-F_box <- subset(aln_df, aln_pos <= 96)
-F_box <- chunkAlnDF(aln_df = F_box, chunk_width = 80)
-p <-ggplot(F_box, aes(x = aln_pos, y = tair_symbol,
-                             group = seq_pos, text = variants))
-p <- p +
-  geom_tile(data = na.omit(F_box), mapping = aes(fill = effects),
-            width = 1, height = 1, alpha = 0.8) +
-  geom_text(aes(label=letter), alpha= 1, family = "Courier") +
-  scale_fill_brewer(type = "qual", palette = 1, direction = -1) +
-  scale_x_continuous(breaks=seq(1,max(F_box$aln_pos), by = 10)) +
-  scale_y_discrete() +
-  # expand increases distance from axis
-  labs(x = "", y = "", fill = "variant") +
-  theme_logo(base_family = "Helvetica") +
-  theme(panel.grid = element_blank(), panel.grid.minor = element_blank()) +
-  facet_wrap(facets = ~chunk, ncol = 1, scales = "free") +
-  theme(strip.background = element_blank(),
-        strip.text.x = element_blank(),
-        legend.box = "vertical")
-p
-```

### 2018-07-12 manuscript/manuscript.Rmd
-Cleaned up manuscript folder -moved to bookdown format -revised all sections

@@
 
-## ARF genes
+## *ARF* genes
 

### 2018-07-12 manuscript/readSelection.py

@@ -1,23 +1,0 @@
-from pymol import cmd, stored
-
-def readSelection (fileName):
-	'''
-	DESCRIPTION
-
-	reads a csv file containing a column of selection names and a column of
-	selection strings and creates selection objects for each
-	'''
-	#read in the selection file
-	with open(fileName) as f:
-		lines = f.readlines()
-		for line in lines:
-			sel = line.split(",")
-			cmd.select(sel[0], sel[1])
-
-	f.close()
-
-cmd.extend('readSelection',readSelection);
-
-
-
-

### 2018-07-12 manuscript/structures/readSelection.py

@@ -1,0 +1,23 @@
+from pymol import cmd, stored
+
+def readSelection (fileName):
+	'''
+	DESCRIPTION
+
+	reads a csv file containing a column of selection names and a column of
+	selection strings and creates selection objects for each
+	'''
+	#read in the selection file
+	with open(fileName) as f:
+		lines = f.readlines()
+		for line in lines:
+			sel = line.split(",")
+			cmd.select(sel[0], sel[1])
+
+	f.close()
+
+cmd.extend('readSelection',readSelection);
+
+
+
+

### 2018-07-13 manuscript/03-AFBs.Rmd
-update gitignore for bookdown -update conditional table formatting -lots of revising -big rearrangement of repo structure

@@
 
-Although it is unclear what unique roles each receptor plays in growth and development, a number of studies have pointed out differences in the ways the six different receptors in A.
-thaliana differ in biochemical function and expression domain[@dharmasiri_plant_2005; @parry_complex_2009; @prigge_arabidopsis_2016].
-*TIR1/AFB* genes are expressed ubiquitously in A.
-thaliana tissues, TIR1, AFB2, and AFB3 have been shown to accumulate in the shoot and root meristems and leaf tissues, with slightly different expression patterns for TIR1 [@Parry_Complex_2009].
-Additionally, the expression of AFB5 is stronly circadian-regulated [@Covington_Circadian_2007] and AFB3 is more highly expressed in the roots in the presence of nitrate, allowing increased lateral root formation [@Vidal_Nitrate_2010], suggesting more broad environmental regulation of this gene family may exist.
-
-Functionally, all members of this family have been shown to bind auxin and Aux/IAA proteins.
+All members of this family have been shown to bind auxin and Aux/IAA proteins.
 However, AFB1 has drastically reduced ability to assemble into an SCF complex, due to the substitution E8K in its F-box domain, preventing it from inducing degradation of Aux/IAAs[@Yu_Untethering_2015].
@@
 
-```{r TIR1_AFB_div_stats}
+Although it is unclear what unique roles each receptor plays in growth and development, a number of studies have pointed out differences in the ways the six different receptors in A.
+thaliana differ in biochemical function and expression domain[@dharmasiri_plant_2005; @parry_complex_2009; @prigge_arabidopsis_2016].
+*TIR1/AFB* genes are expressed ubiquitously in A.
+thaliana tissues, TIR1, AFB2, and AFB3 have been shown to accumulate in the shoot and root meristems and leaf tissues, with slightly different expression patterns for TIR1 [@Parry_Complex_2009].
+Additionally, the expression of AFB5 is strongly circadian-regulated [@Covington_Circadian_2007] and AFB3 is more highly expressed in the roots in the presence of nitrate, facilitating increased lateral root formation [@Vidal_Nitrate_2010], suggesting more broad environmental regulation of this gene family may exist.
+
+All members of this family have been shown to bind auxin and Aux/IAA proteins.
+However, AFB1 has drastically reduced ability to assemble into an SCF complex, due to the substitution E8K in its F-box domain, preventing it from inducing degradation of Aux/IAAs[@Yu_Untethering_2015].
+This lack of complexation may be the cause of high, ubiquitous AFB1 accumulation [@Parry_Complex_2009].
+Higher order mutants in the family containing *afb1* mutants suggest that *AFB1* has a moderate positive effect on auxin signaling.
+Additionally, AFB4 and AFB5 have been shown to preferentially bind the synthetic auxin picloram, while other family members preferentially bind indole-3-acetic acid [@Prigge_Arabidopsis_2016].
+Interestingly, the strength and rate with which each TIR1/AFB is able to bind and mark an Aux/IAAs for degradation is variable [@calderonvillalobos_combinatorial_2012; @Havens_Synthetic_2012].
+
+(tab:TIR1_AFB_div_stats) **_TIR1/AFB_ nucleotide diversity.** The nucleotide diversity as Nei and Li's $\pi$ (the average number of nucleotide differences per site between all possible pairs of sequences at synonymous ($\pi_S$) and nonsynonymous ($\pi_N$) sites, as well as all sites in the coding sequence and full transcript of the *TIR1/AFB* family.
+```{r TIR1_AFB_div_stats, fig.cap = '(tab:TIR1_AFB_div_stats)'}
 gene_info <- geneInfoFromFile(
@@
 
-if (params$output == "html_document") {
+if (knitr::is_html_output()) {
   formatRound(datatable(div_stats,
@@
 
-Examining the natural sequence variation across the *AFB* family revealed that *TIR1* and *AFB1* both had very low nonsynonymous diversity, hinting at their likely functional importance and bringing in to question the inconclusive role of *AFB1* in auxin signaling.
+Examining the natural sequence variation across the *AFB* family \@ref(tab:TIR1_AFB_div_stats) revealed that *TIR1* and *AFB1* both had very low nonsynonymous diversity, hinting at their likely functional importance and bringing in to question the inconclusive role of *AFB1* in auxin signaling.
 *AFB3* and *AFB4* had higher nonsynonymous diversity, while their sister genes, *AFB2* and *AFB5* were more conserved.
@@
 Mutations in this domain of *TIR1* frequently have a semidominant effect on root phenotypes [@dezfulian_oligomerization_2016; @wright_insights_2017a].
-Characterization of this accession may identify potential *AFB1* phenotypes and may also test the hypothesis that AFB1 functions through oligomerization with other TIR1/AFBs.
+Characterization of this allele and accession may help determine the role of *AFB1* in this pathway.
 
-```{r AFB variants for PyMOL}
+```{r AFB variants for PyMOLx}
 alignment <- alignCDS(gene_info$transcript_ID)
@@
                             effect_type = "missense", name_by_symbols = TRUE){
-  collapsed_aln_df <- aln_df %>% dplyr::select(aln_pos, seq_pos, seq_name) %>%
+  collapsed_aln_df <- aln_df %>% as_tibble() %>% 
+    dplyr::select(aln_pos, seq_pos, seq_name) %>%
     tidyr::spread(key = seq_name, value = seq_pos)
@@
 selections$V2 <- paste0("chain B and resi ", selections$V2)
-write.table(selections, file = "AFB_variant_selections", row.names = FALSE, 
-          col.names = FALSE, sep = ",", quote = FALSE)
+#write.table(selections, file = "AFB_variant_selections", row.names = FALSE, 
+#          col.names = FALSE, sep = ",", quote = FALSE)
 ```
@@
 ray
+```
 
+(fig:AFB_structure_vars) A TIR1 structure (2P1Q) in gold, with residues 
+```{r AFB_structure_vars, fig.cap='(fig:AFB_structure_vars)'}
+knitr::include_graphics()
 ```

### 2018-07-13 manuscript/04-IAAs.Rmd
-update gitignore for bookdown -update conditional table formatting -lots of revising -big rearrangement of repo structure

@@
 
+(ref:IAA tree) IAA protein sequence tree mapped with ${\pi_N/\pi_S}$
+
 ```{r IAA_tree, fig.height=5, fig.width=5, fig.cap="(ref:IAA tree)"}
@@
 ```
-
-(ref:IAA tree) IAA protein sequence tree mapped with ${\pi_N/\pi_S}$
 
@@
 ### Identifying novel functional sequences in IAAs
+
 An exciting possible use for the Natural Variation Webtool is to search for conserved, low-diversity sites that may represent possible sites of functional import, such as post-translational modification or protein-protein interaction.

### 2018-07-13 manuscript/07-discussion.Rmd
-update gitignore for bookdown -update conditional table formatting -lots of revising -big rearrangement of repo structure

@@
 
-Analyzing natural diversity within the gene families constituting the auxin nuclear signaling pathway has confirmed existing knowledge, supported existing hypothses and generated new hypotheses.
-This analysis in particular has confirmed the importance of auxin-induced degradation of the Aux/IAA proteins as evidenced by the strong conservation of the *AFB* genes and the canonical degron region of the *Aux/IAA* genes.
+Our analysis of the natural diversity in the gene families that make up the auxin nuclear signaling pathway has confirmed existing knowledge, supported existing hypothses and generated new hypotheses.
+These results have been generated by a team of undergraduate and graduate students, postdoctoral researchers, and principal investigators as an alpha-test of the web application developed here.
+Our analysis has confirmed the importance of auxin-induced degradation of the Aux/IAA proteins as evidenced by the strong conservation of the *AFB* genes and the canonical degron region of the *Aux/IAA* genes.
 Despite this relative conservation, this analysis has also identified many accessions containing new alleles of interest in the *AFB* and *Aux/IAA* families.
-This analysis has also
+-interesting points from the AFBs
+-interesting points IAAs
+-TPLs
+-ARFs
+  -DBD and PB1
 

### 2018-07-13 manuscript/index.Rmd
-update gitignore for bookdown -update conditional table formatting -lots of revising -big rearrangement of repo structure

@@
 ---
-title: "A Minimal Book Example"
-author: "Yihui Xie"
+title: A comprehensive analysis of natural sequence variation within the <i>Arabidopsis
+  thaliana</i> nuclear auxin signaling pathway using an accessible web application
+author:
+  affiliation: University of Washington
+  name: R. Clay Wright
 date: "`r Sys.Date()`"
-site: bookdown::bookdown_site
-documentclass: book
-bibliography: [book.bib, packages.bib]
+
+abstract: "This will be the abstract"
+
+keywords: natural variation, Arabidopsis thaliana, genome diversity, structure, function, genotype, phenotype, bioconductor, bookdown, r markdown
+
+output:
+  word_document:
+    toc: yes
+  html_document:
+    fig_caption: yes
+    toc: yes
+always_allow_html: yes
+fontfamily: Helvetica
+fontsize: 11pt
+
+bibliography:
+- natural_variation.bib
+- TIR1_AFB.bib
+- ARF_C.bib
+- AuxIAA_citations.bib
 biblio-style: apalike
 link-citations: yes
-description: "This is a minimal example of using the bookdown package to write a book. The output format for this example is bookdown::gitbook."
+
 ---

### 2018-07-14 manuscript/03-AFBs.Rmd
correctly render AFB section as html

@@
 ```{r AFB_structure_vars, fig.cap='(fig:AFB_structure_vars)'}
-knitr::include_graphics()
+#knitr::include_graphics()
 ```

### 2018-07-14 manuscript/index.Rmd
correctly render AFB section as html

@@
 keywords: natural variation, Arabidopsis thaliana, genome diversity, structure, function, genotype, phenotype, bioconductor, bookdown, r markdown
-
-output:
-  word_document:
-    toc: yes
-  html_document:
-    fig_caption: yes
-    toc: yes
-always_allow_html: yes
-fontfamily: Helvetica
-fontsize: 11pt
 
@@
 library(knitr)
+library(dplyr)
 

### 2018-07-14 manuscript/02-methods.Rmd
-update gitignore for bookdown -update conditional table formatting -lots of revising -big rearrangement of repo structure

@@
 # Methods {#methods}
+
+## Data Sources
+
+### Variant Data
+Variant data was queried from the 1001 genomes project website (http://1001genomes.org) via URL requests to an (undocumented) API service that provides subset files of the whole-genome VCF.
+The whole-genome VCF file can be found on the project’s website at http://1001genomes.org/data/GMI-MPI/releases/v3.1/
+
+### Accession Information
+Details on each of the 1135 accessions including CS stock numbers, and geographic location where the samples were collected were taken from the 1001 genomes website at http://1001genomes.org/accessions.html, via the download link at the bottom of the page.
+This data file has been embedded in the R package.
+
+### Gene and transcript information
+Information on the genes and transcripts including chromosomal coordinates, start and end location, and transcript length were taken from either the TAIR10 or Araport11 databases.
+The TAIR10 database, found at http://arabidopsis.org, was accessed via the biomart protocol, using the R package biomaRt.
+The araport11 database, which can also be found on the arabidopsis.org website (https://www.arabidopsis.org/download/index-auto.jsp?dir=%2Fdownload_files%2FGenes%2FAraport11_genome_release), has been embedded in the R package.
 
@@
 
-$\displaystyle \pi = \sum_{ij}x_{i}x_{j}\pi_{ij}$
+\begin{equation}
+\displaystyle \pi = \sum_{ij}x_{i}x_{j}\pi_{ij}
+(\#eq:nei)
+\end{equation}
 
@@
 
-$\displaystyle \pi = \frac{1}{L*n^2} \sum_{i=1}^n \sum_{j=1}^n \sum_{k=1}^L{\pi_{ijk}}
+\begin{equation}
+\displaystyle \pi = \frac{1}{L*n^2} \sum_{i=1}^n \sum_{j=1}^n \sum_{k=1}^L{\pi_{ijk}}
 \qquad
@@
 0 \quad \textrm{if} \quad N_{ik} = N_{jk}
-\end{array}\right.$
+\end{array}\right.
+(\#eq:nei-gen)
+\end{equation}
 
@@
 
-$\displaystyle \pi = \frac{1}{L} \sum_{k=1}^L{\pi_{k}}
+\begin{equation}
+\displaystyle \pi = \frac{1}{L} \sum_{k=1}^L{\pi_{k}}
 \qquad
-\pi_{k} = \frac{1}{n^2} \sum_{i=1}^n \sum_{j=1}^n {\pi_{ijk}}$
+\pi_{k} = \frac{1}{n^2} \sum_{i=1}^n \sum_{j=1}^n {\pi_{ijk}}
+(\#eq:nei-rearrange)
+\end{equation}
 
@@
 
-so we can revert back to using Nei's original formula, modifying it slightly,
+so we can revert back to using Nei's original formula \@ref(eq:nei), modifying it slightly,
 replacing $x_i$ with $\frac{n_i}{n}$, $n_i$ being the number of sequences in the population of the form $N_i$ at location k: subscripts i and k are summed only over unique forms [need to clarify this]
 
-$\displaystyle \pi_k = \sum_{ij} \frac{n_i}{n}\frac{n_j}{n} \pi_{ij} = 
-\frac{1}{n^2}\sum_{ij} n_i n_j \pi_{ij}$
-
-$\pi_{ij(k)} = \left\{\begin{array}{l}
+\begin{equation}
+\displaystyle \pi_k = \sum_{ij} \frac{n_i}{n}\frac{n_j}{n} \pi_{ij} = 
+\frac{1}{n^2}\sum_{ij} n_i n_j \pi_{ij}\\
+\pi_{ij(k)} = \left\{\begin{array}{l}
 1 \quad \textrm{if} \quad i \neq j \\
 0 \quad \textrm{if} \quad i = j
-\end{array}\right.$
+\end{array}\right.
+(\#eq:nei-simp1)
+\end{equation}
 
@@
 
-$\displaystyle \pi_k = \frac{1}{n^2} \sum_{i} n_i n_{!i}$
+\begin{equation}
+\displaystyle \pi_k = \frac{1}{n^2} \sum_{i} n_i n_{!i}
+(\#eq:nei-simp2)
+\end{equation}
 
@@
 
-$\displaystyle \pi_k = \frac{1}{n^2} \sum_{i} n_i (n - n_{i})$
+\begin{equation}
+\displaystyle \pi_k = \frac{1}{n^2} \sum_{i} n_i (n - n_{i})
+(\#eq:nei-simp3)
+\end{equation}
 
@@
 
-$\displaystyle \pi_k = \frac{1}{n^2} (n \sum_{i} n_i - \sum_{i} n_{i}^2)$
+\begin{equation}
+\displaystyle \pi_k = \frac{1}{n^2} (n \sum_{i} n_i - \sum_{i} n_{i}^2)
+(\#eq:nei-simp4)
+\end{equation}
 
@@
 
-$\displaystyle \pi_k = \frac{1}{n^2} (n^2 - \sum_{i} n_{i}^2)$
+\begin{equation}
+\displaystyle \pi_k = \frac{1}{n^2} (n^2 - \sum_{i} n_{i}^2)
+(\#eq:nei-simp5)
+\end{equation}
 
@@
 variants at a single nucleotide location can easily be summarized in R.
+
+## Software
+
+The r1001genomes package has many software dependencies on other R packages, a few of the key bioinformatics packages used are listed below.
+
+**biomaRt:** The biomaRt package was used for accessing the TAIR10 database on arabidopsis.org
+
+**vcfR:** vcfR was used to read in the VCF files in a flat “tidy” format for easy manipulaiton
+
+**BSgenome:** The BSgenome package BSgenome.Athaliana.TAIR.TAIR9 was used as the source for the complete DNA string of the reference genome (Col-0).
+
+**DECIPHER:** DECIPHER was used to align nucleotide and amino acid sequences of homologous genes
+
+**GenomicFeatures:** The GenomicFeatures package was used for handling sequence annotations.
+
+**Biostrings:** Biostrings provides the underlying framework for the sequence manipulations used for generating and aligning sequences with BSgenome, Decipher, and GenomicFeatures

### 2018-07-16 manuscript/01-intro.Rmd
-Figure legends added in AFBs   -Table legends in HTML are same as figures -Fixed bibtex file updating

@@
 Natural genomic variation also provides a set of feasible functional variation at the gene, gene family and gene network level.
-Characterizing this variation can provide insights into the function and evolution of genes, families and networks [@delker_natural_2010, @wright_insights_2017] .
+Characterizing this variation can provide insights into the function and evolution of genes, families and networks [@delker_natural_2010; @wright_insights_2017] .
 Here, we present a web application and R-package for gene, gene family and gene network analysis of this dataset.
@@
 We also identify natural alleles of several genes and accessions bearing these alleles to facilitate future characterization structure/function, genotype/phenotype, and epistatic relationships.
-
-## bookdown example
-You can label chapter and section titles using `{#label}` after them, e.g., we can reference Chapter \@ref(intro).
-If you do not manually label them, there will be automatic labels anyway, e.g., Chapter \@ref(methods).
-
-Figures and tables with captions will be placed in `figure` and `table` environments, respectively.
-
-```{r nice-fig, fig.cap='Here is a nice figure!', out.width='80%', fig.asp=.75, fig.align='center'}
-par(mar = c(4, 4, .1, .1))
-plot(pressure, type = 'b', pch = 19)
-```
-
-Reference a figure by its code chunk label with the `fig:` prefix, e.g., see Figure \@ref(fig:nice-fig).
-Similarly, you can reference tables generated from `knitr::kable()`, e.g., see Table \@ref(tab:nice-tab).
-
-```{r nice-tab, tidy=FALSE}
-knitr::kable(
-  head(iris, 20), caption = 'Here is a nice table!',
-  booktabs = TRUE
-)
-```
-
-You can write citations, too. For example, we are using the **bookdown** package [@R-bookdown] in this sample book, which was built on top of R Markdown and **knitr** [@xie2015].

### 2018-07-16 manuscript/03-AFBs.Rmd
-Figure legends added in AFBs   -Table legends in HTML are same as figures -Fixed bibtex file updating

--- 2018-07-13 manuscript/03-AFBs.Rmd
@@
 Auxin acts by binding to receptors (Auxin-signaling F-Boxes, or AFBs) that in turn target co-repressors (Aux/IAAs) for degradation.
-The six auxin receptor genes in the model plant *Arabidopsis thaliana* , *TIR1* and *AFB1-5*, evolved through gene duplication and diversification early in the history of vascular plants [@Parry_Complex_2009].
-The rate of co-repressor degradation is determined by the identity of both the receptor and co-repressor [@Havens_Synthetic_2012], and this rate sets the pace of lateral root development [@Guseman_Auxin-induced_2015].
+The six auxin receptor genes in the model plant *Arabidopsis thaliana* , *TIR1* and *AFB1-5*, evolved through gene duplication and diversification early in the history of vascular plants [@parry_complex_2009].
+The rate of co-repressor degradation is determined by the identity of both the receptor and co-repressor [@havens_synthetic_2012], and this rate sets the pace of lateral root development [@guseman_auxininduced_2015].
 
 All members of this family have been shown to bind auxin and Aux/IAA proteins.
-However, AFB1 has drastically reduced ability to assemble into an SCF complex, due to the substitution E8K in its F-box domain, preventing it from inducing degradation of Aux/IAAs[@Yu_Untethering_2015].
-This lack of complexation may allow observed high ubiquitous AFB1 accumulation [@Parry_Complex_2009].
+However, AFB1 has drastically reduced ability to assemble into an SCF complex, due to the substitution E8K in its F-box domain, preventing it from inducing degradation of Aux/IAAs[@yu_untethering_2015].
+This lack of complexation may allow observed high ubiquitous AFB1 accumulation [@parry_complex_2009].
 Higher order mutants in the family containing *afb1* mutants suggest that *AFB1* has a moderate positive effect on auxin signaling.
-Additionally, AFB4 and AFB5 have been shown to preferentially and functionally bind the synthetic auxin picloram, while other family members preferentially bind indole-3-acetic acid [@Prigge_Arabidopsis_2016].
-Interestingly, the strength and rate with which TIR1/AFBs are able to bind and mark Aux/IAAs for degradation are variable [@calderonvillalobos_combinatorial_2012; @Havens_Synthetic_2012].
+Additionally, AFB4 and AFB5 have been shown to preferentially and functionally bind the synthetic auxin picloram, while other family members preferentially bind indole-3-acetic acid [@prigge_arabidopsis_2016].
+Interestingly, the strength and rate with which TIR1/AFBs are able to bind and mark Aux/IAAs for degradation are variable [@calderonvillalobos_combinatorial_2012; @havens_synthetic_2012].
 AFB2 induced the degradation of certain Aux/IAA proteins at a faster rate than TIR1, suggesting some functional specificity has arisen since the initial duplication between the *TIR1/AFB1* and *AFB2/AFB3* clades.
@@
 *TIR1/AFB* genes are expressed ubiquitously in A.
-thaliana tissues, TIR1, AFB2, and AFB3 have been shown to accumulate in the shoot and root meristems and leaf tissues, with slightly different expression patterns for TIR1 [@Parry_Complex_2009].
-Additionally, the expression of AFB5 is strongly circadian-regulated [@Covington_Circadian_2007] and AFB3 is more highly expressed in the roots in the presence of nitrate, facilitating increased lateral root formation [@Vidal_Nitrate_2010], suggesting more broad environmental regulation of this gene family may exist.
+thaliana tissues, TIR1, AFB2, and AFB3 have been shown to accumulate in the shoot and root meristems and leaf tissues, with slightly different expression patterns for TIR1 [@parry_complex_2009].
+Additionally, the expression of AFB5 is strongly circadian-regulated [@covington_circadian_2007] and AFB3 is more highly expressed in the roots in the presence of nitrate, facilitating increased lateral root formation [@vidal_nitrate_2014], suggesting more broad environmental regulation of this gene family may exist.
 
 All members of this family have been shown to bind auxin and Aux/IAA proteins.
-However, AFB1 has drastically reduced ability to assemble into an SCF complex, due to the substitution E8K in its F-box domain, preventing it from inducing degradation of Aux/IAAs[@Yu_Untethering_2015].
-This lack of complexation may be the cause of high, ubiquitous AFB1 accumulation [@Parry_Complex_2009].
+However, AFB1 has drastically reduced ability to assemble into an SCF complex, due to the substitution E8K in its F-box domain, preventing it from inducing degradation of Aux/IAAs[@yu_untethering_2015].
+This lack of complexation may be the cause of high, ubiquitous AFB1 accumulation [@parry_complex_2009].
 Higher order mutants in the family containing *afb1* mutants suggest that *AFB1* has a moderate positive effect on auxin signaling.
-Additionally, AFB4 and AFB5 have been shown to preferentially bind the synthetic auxin picloram, while other family members preferentially bind indole-3-acetic acid [@Prigge_Arabidopsis_2016].
-Interestingly, the strength and rate with which each TIR1/AFB is able to bind and mark an Aux/IAAs for degradation is variable [@calderonvillalobos_combinatorial_2012; @Havens_Synthetic_2012].
+Additionally, AFB4 and AFB5 have been shown to preferentially bind the synthetic auxin picloram, while other family members preferentially bind indole-3-acetic acid [@prigge_arabidopsis_2016].
+Interestingly, the strength and rate with which each TIR1/AFB is able to bind and mark an Aux/IAAs for degradation is variable [@calderonvillalobos_combinatorial_2012; @havens_synthetic_2012].
 
-(tab:TIR1_AFB_div_stats) **_TIR1/AFB_ nucleotide diversity.** The nucleotide diversity as Nei and Li's $\pi$ (the average number of nucleotide differences per site between all possible pairs of sequences at synonymous ($\pi_S$) and nonsynonymous ($\pi_N$) sites, as well as all sites in the coding sequence and full transcript of the *TIR1/AFB* family.
-```{r TIR1_AFB_div_stats, fig.cap = '(tab:TIR1_AFB_div_stats)'}
+(ref:TIR1AFBdivstats) **_TIR1/AFB_ nucleotide diversity.** The nucleotide diversity as Nei and Li's $\pi$ (the average number of nucleotide differences per site between all possible pairs of sequences at synonymous ($\pi_S$) and nonsynonymous ($\pi_N$) sites, as well as all sites in the coding sequence and full transcript of the *TIR1/AFB* family.
+
+```{r TIR1AFBdivstats, fig.cap="(ref:TIR1AFBdivstats)"}
 gene_info <- geneInfoFromFile(
@@
   names(dt) <- paste0("**", names(dt), "**")
-  kable(dt) 
+  kable(dt, caption = "(\\#tab:TIR1AFBdivstats)") 
 }
@@
 
-Examining the natural sequence variation across the *AFB* family \@ref(tab:TIR1_AFB_div_stats) revealed that *TIR1* and *AFB1* both had very low nonsynonymous diversity, hinting at their likely functional importance and bringing in to question the inconclusive role of *AFB1* in auxin signaling.
+Examining the natural sequence variation across the *AFB* family \@ref(fig:TIR1AFBdivstats) revealed that *TIR1* and *AFB1* both had very low nonsynonymous diversity, hinting at their likely functional importance and bringing in to question the inconclusive role of *AFB1* in auxin signaling.
 *AFB3* and *AFB4* had higher nonsynonymous diversity, while their sister genes, *AFB2* and *AFB5* were more conserved.
@@
 Although most known functional regions are highly conserved in *AFB1*, there is a nonsynonymous polymorphism  in the oligomerization domain, only found in the Can-0 accession.
-Mutations in this domain of *TIR1* frequently have a semidominant effect on root phenotypes [@dezfulian_oligomerization_2016; @wright_insights_2017a].
+Mutations in this domain of *TIR1* frequently have a semidominant effect on root phenotypes [@dezfulian_oligomerization_2016; @wright_insights_2017].
 Characterization of this allele and accession may help determine the role of *AFB1* in this pathway.
 
-```{r AFB variants for PyMOLx}
+```{r AFB variants for PyMOL}
 alignment <- alignCDS(gene_info$transcript_ID)
@@
 
-(fig:AFB_structure_vars) A TIR1 structure (2P1Q) in gold, with residues 
-```{r AFB_structure_vars, fig.cap='(fig:AFB_structure_vars)'}
-knitr::include_graphics()
+(ref:AFB_structure_vars) A TIR1 structure (2P1Q) in gold, with residues
+
+```{r AFB_structure_vars, fig.cap='(ref:AFB_structure_vars)'}
+#knitr::include_graphics()
 ```
@@
 The AFB4 and AFB5 receptors have an N-terminal extension prior to the F-box domains.
-This extension had very high nonsynonymous diversity(Fig.
-4), suggesting that this extension does not play an important functional role in these proteins.
+This extension had very high nonsynonymous diversity \@ref(fig:FBoxAlignment), suggesting that this extension does not play an important functional role in these proteins.
 Additionally, two frameshift variants and one stop-gained variant were observed in *AFB4* supporting its pseudogenization.
 
-Figure 4: Alignment of A.
-thaliana TIR1/AFB F-box domains showing variants.
-Cyan marks the F-box binding domain.
+(ref:FBoxAlignment) **Alignment of A.
+thaliana TIR1/AFB F-box domains showing variants.** Cyan marks the F-box binding domain.
 Yellow, green, and magenta, mark missense and synonymous, synonymous only, and missense only variant locations respectively.
 Yellow squared E’s show Cul1-interacting residues, differing in AFB1.
-```{r TIR1/AFB F-box Alignment}
+
+```{r FBoxAlignment, fig.cap = "(ref:FBoxAlignment)"}
 aln_df <- makeAlnDF(alignment[[2]])

### 2018-07-16 manuscript/04-IAAs.Rmd
-Figure legends added in AFBs   -Table legends in HTML are same as figures -Fixed bibtex file updating

@@
 
-```{r IAA diversity stats, fig.height=11}
+(ref:IAAdivstats) **_Aux/IAA_ family nucleotide diversity.** The nucleotide diversity as Nei and Li's $\pi$ (the average number of nucleotide differences per site between all possible pairs of sequences at synonymous ($\pi_S$) and nonsynonymous ($\pi_N$) sites, as well as all sites in the coding sequence and full transcript of the *Aux/IAA* family.
+
+```{r IAAdivstats, fig.height=11}
 gene_info <- geneInfoFromFile(
@@
 
-In our analysis of the *IAA* gene family, we frequently observed that one member of most *IAA* sister pairs had high nonsynonymous diversity while the other sister was more conserved (Figure IAA phylogeny heat map).
+In our analysis of the *IAA* gene family, we frequently observed that one member of most *IAA* sister pairs had high nonsynonymous diversity while the other sister was more conserved \@ref(fig:IAAtree).
 In an interspecific comparison between *A.
@@
 lyrata*, *IAA19* is more conserved, while *IAA6* has regions of increased sequence divergence upstream of the degron and in the N-terminal half of the PB1 domain [@winkler_variation_2017].
-Consistent with these results, our analysis of intraspecific natural variation revealed higher conservation for *IAA19* (${\pi_N/\pi_S}$ = 0.55) compared to *IAA6* (${\pi_N/\pi_S}$ = 2.3) (Figure IAA phylogeny heat map), and also detected high diversity within the same regions of *IAA6* as seen in Winkler et al.
+Consistent with these results, our analysis of intraspecific natural variation revealed higher conservation for *IAA19* (${\pi_N/\pi_S}$ = 0.55) compared to *IAA6* (${\pi_N/\pi_S}$ = 2.3) \@ref(fig:IAAtree), and also detected high diversity within the same regions of *IAA6* as seen in Winkler et al.
 (Supp Fig IAA6 diversity plot).
@@
 
-(ref:IAA tree) IAA protein sequence tree mapped with ${\pi_N/\pi_S}$
+(ref:IAAtree) **IAA protein sequence tree mapped with ${\pi_N/\pi_S}$.** Protein sequences were aligned [@wright_decipher_2015] and low information content regions were masked [@kuck_parametric_2010] prior to infering a phylogeny [@ronquist_mrbayes_2003].
+Tips of the tree are mapped with circles of diameter proportional to ${\pi_N/\pi_S}$ and also are colored according to ${\pi_N/\pi_S}$.
+Nodes are labeled with the poster probability of monophyly.
 
-```{r IAA_tree, fig.height=5, fig.width=5, fig.cap="(ref:IAA tree)"}
+```{r IAAtree, fig.height=5, fig.width=5, fig.cap="(ref:IAAtree)"}
 p <- ggtree(tree, layout = "rectangular", color = "grey") %<+% IAA_data + theme_tree() + scale_x_continuous(expand = c(0, 1.2)) + 

### 2018-07-16 manuscript/index.Rmd
-Figure legends added in AFBs   -Table legends in HTML are same as figures -Fixed bibtex file updating

@@
 bibliography:
-- natural_variation.bib
+- natural-variation.bib
 - TIR1_AFB.bib
-- ARF_C.bib
-- AuxIAA_citations.bib
+- ARF.bib
+- Aux_IAA.bib
 biblio-style: apalike
@@
 library(dplyr)
+library(bookdown)
+library(htmltools)
 

### 2018-07-20 manuscript/05-TPLs.Rmd
Added alignment to TPL section and re-worded some text in methods section

@@
 
-```{r TPL_TPR alignment}
+(ref:TPLsAlignment) **Alignment of A. thaliana TPL/TPR genes showing variants.** This alignment should probably be paired down to a key region rather than the eniter AA sequence.
 
+```{r TPL_TPR alignment, fig.height = 20, fig.cap = "(ref:TPLsAlignment)"}
+
+alignment <- alignCDS(gene_info$transcript_ID)
+
+aln_df <- makeAlnDF(alignment[[2]])
+aln_df <- left_join(aln_df, dplyr::select(gene_info, "tair_locus",
+                                       "tair_symbol", "transcript_ID"),
+                        by = c("transcript_ID" = "transcript_ID"))
+aln_df$tair_symbol <- fct_rev(aln_df$tair_symbol)
+aln_df <- chunkAlnDF(aln_df = aln_df, chunk_width = 55)
+VCF <- ldply(.data = VCF, .fun = subset, gt_GT != "0|0")
+VCF <- getCodingDiv(VCF)
+aln_df <- addSNPsToAlnDF(aln_df, VCF)
+
+# aln_df_chunk <- chunkAlnDF(aln_df = aln_df, chunk_width = 55)
+p <-ggplot(aln_df_chunk, aes(x = aln_pos, y = tair_symbol,
+                             group = seq_pos, text = variants))
+p <- p +
+  geom_tile(data = na.omit(aln_df_chunk), mapping = aes(fill = effects),
+            width = 1, height = 1, alpha = 0.8) +
+  geom_text(aes(label=letter), alpha= 1, family = "Courier") +
+  scale_fill_brewer(type = "qual", palette = 1, direction = -1) +
+  scale_x_continuous(breaks=seq(1,max(aln_df_chunk$aln_pos), by = 10)) +
+  scale_y_discrete() +
+  # expand increases distance from axis
+  labs(x = "", y = "", fill = "variant") +
+  theme_logo(base_family = "Helvetica") +
+  theme(panel.grid = element_blank(), panel.grid.minor = element_blank()) +
+  facet_wrap(facets = ~chunk, ncol = 1, scales = "free") +
+  theme(strip.background = element_blank(),
+        strip.text.x = element_blank(),
+        legend.box = "vertical")
+p
 ```

### 2018-09-09 manuscript/01-intro.Rmd
-TPL tree -AFB tree -Alignment plots
still some missing/broken chunks, but commit to sync to new computer

@@
 Characterizing this variation can provide insights into the function and evolution of genes, families and networks [@delker_natural_2010; @wright_insights_2017] .
-Here, we present a web application and R-package for gene, gene family and gene network analysis of this dataset.
+Here, we present ViVa a web application and R-package for Visualizing Variation within genes, gene families and gene networks.
+We apply this to analysis of the 1001 genomes dataset [@.
 This application allows plant molecular biologists without bioinformatics experience to use genomic variation to formulate hypotheses as to the sequence/function/phenotype relationships within their gene, family, or network of their interest.
@@
 lists of accessions containing missense (or any type of) polymorphisms to facilitate segregation analysis and measurment of the distribution of variant effects on phenotype [@park_distributions_2017],
-4. as well as the study of gene and network evolution [@delker_natural_2010; @kliebenstein_role_2008].
+4. as well as the study of gene, gene family and network evolution [@delker_natural_2010; @kliebenstein_role_2008].
 

### 2018-09-09 manuscript/03-AFBs.Rmd
-TPL tree -AFB tree -Alignment plots
still some missing/broken chunks, but commit to sync to new computer

@@
 div_stats %<>% arrange(tair_symbol)
-div_stats %<>% select(tair_symbol, everything())
+div_stats %<>% dplyr::select(tair_symbol, everything())
 
@@
 
-Although most known functional regions are highly conserved in *AFB1*, there is a nonsynonymous polymorphism  in the oligomerization domain, only found in the Can-0 accession.
-Mutations in this domain of *TIR1* frequently have a semidominant effect on root phenotypes [@dezfulian_oligomerization_2016; @wright_insights_2017].
-Characterization of this allele and accession may help determine the role of *AFB1* in this pathway.
-
-```{r AFB variants for PyMOL}
+```{r AFB_ alignment}
 alignment <- alignCDS(gene_info$transcript_ID)
@@
   tidyr::spread(key = seq_name, value = seq_pos)
+```
+
+```{r Mask_AFB_alignment, eval=FALSE}
+aliscore(as.AAbin(alignment[[2]]), path = "/Volumes/SSD/Applications/ALISCORE_v2.0/Aliscore_v.2.0/",  o = "AT2G39940")
+id <- scan(file = "/Volumes/SSD/Applications/ALISCORE_v2.0/Aliscore_v.2.0/input.fas_List_random.txt", sep = " ", quiet = TRUE)
+AAAlign_no_gaps <- as.matrix(alignment[[2]])[, id]
+write.nexus.data(AAAlign_no_gaps, file = "trees/AFBs.nex", format = "protein")
+```
+```{bash IAA_phylogeny_construction, eval=FALSE}
+cd ~/Desktop/natural-variation-webtool/manuscript/trees
+
+/Applications/MrBayes/mb
+
+execute AFBs.nex
+# set IAA33 as the outgroup
+outgroup AT2G39940.1
+
+lset nucmodel=protein Rates=invgamma
+# or include Nst=6  ngammacat=4 omegavar=M3 
+prset aamodelpr=fixed(jones) statefreqpr=fixed(empirical) 
+# initial run with aamodelpr=mixed yielded posterior prob of 1 for jones model
+propset ExtTBR$prob=0
+
+mcmcp ngen=250000 nchains=4 printfreq=1000 samplefreq=100 burnin=200
+
+mcmc
+sump
+sumt
+
+```
+
+```{r Load AFB phylogeny and data}
+tree <- treeio::read.beast(file = "trees/AFBs.nex.con.tre")
+
+AFB_data <- read.csv(system.file("shiny-app", "gene_family_data",
+                      "AFB_gene_ids.csv", package = "r1001genomes"))
+AFB_data %<>% left_join(y = div_stats, by = c("name" = "tair_symbol"))
+#tip_labels need to be in the first column
+AFB_data %<>% dplyr::select(name, everything())
+AFB_data$tair_locus %<>% as.character()
+
+AFB_data <- AFB_data[match(gsub(x = tree@phylo$tip.label, pattern = "\\.1",replacement = ""), AFB_data$tair_locus), ]
+
+tree@phylo$tip.label <- AFB_data$name
+```
+
+(ref:AFBtree) **AFB protein sequence tree mapped with ${\pi_N/\pi_S}$.** Protein sequences were aligned [@wright_decipher_2015] and low information content regions were masked [@kuck_parametric_2010] prior to infering a phylogeny [@ronquist_mrbayes_2003].
+Tips of the tree are mapped with circles of diameter proportional to ${\pi_N/\pi_S}$ and also are colored according to ${\pi_N/\pi_S}$.
+Nodes are labeled with the poster probability of monophyly.
+
+```{r AFBtree, fig.height=2, fig.width=3, fig.cap="(ref:AFBtree)"}
+p <- ggtree(tree, layout = "rectangular", color = "grey") %<+% AFB_data + theme_tree() + scale_x_continuous(expand = c(0, 1.2)) + 
+  geom_tippoint(mapping = aes(size = 4*Pi_NS_Ratio, fill = Pi_NS_Ratio), shape = 21, color = "white", show.legend = TRUE) + 
+  scale_fill_viridis(values = rescale(c(min(AFB_data$Pi_NS_Ratio),
+                                        max(AFB_data$Pi_NS_Ratio)), 
+                                      to = c(0,1)), guide = "colourbar") + scale_size(guide = "none") + 
+  theme(legend.position = c(.1,.7)) + 
+  geom_nodelab(mapping = aes(label = round(prob,digits = 2)), nudge_x = -.3, nudge_y = .5, size = 3) + 
+  geom_tiplab(hjust = -0.25) + 
+  scale_color_viridis(discrete = TRUE)
+p
+#ggsave("AFBs_tree.pdf", height = 3, width = 4)
+```
+
+```{r full-AFBs-alignment}
+AFB_order <- fct_inorder(na.exclude(p$data$label[base::order(p$data$y, decreasing = TRUE)]), ordered = TRUE)
+
+aln_df <- left_join(aln_df, dplyr::select(gene_info, "tair_locus",
+                                       "tair_symbol", "transcript_ID"),
+                        by = c("transcript_ID" = "transcript_ID"))
+aln_df$tair_symbol %<>% factor(levels = AFB_order, 
+                                       ordered = TRUE)
+aln_df$tair_symbol %<>% fct_rev()
+
+aln_df <- addSNPsToAlnDF(aln_df, VCF, effect_order = SNPeff_order)
+aln_df$gap <- aln_df$seq_pos == "-"
+
+p <-ggplot(aln_df, aes(x = aln_pos, y = tair_symbol)) +
+  geom_raster(data = aln_df, 
+             # mapping = aes(fill = effects)) +
+              mapping = aes(fill = strength,
+                            alpha = gap)) +
+  scale_alpha_manual(values = c("TRUE" = 0, "FALSE" = 1), 
+                     labels = c("aligned", "gap"), 
+                     guide = guide_legend(override.aes = 
+                                            list(fill = c("grey85", "white")),
+                       title = "alignment",
+                       direction = "vertical")) + 
+   scale_fill_viridis(breaks = range(effect_order$strength),
+                      limits = range(effect_order$strength),
+                      na.value = "grey85",
+                      direction = -1,
+                      labels = c("neutral", "deleterious"),
+                      option = "A", 
+                      guide = guide_colorbar(ticks = FALSE, 
+                                             title.vjust = .8, 
+                                             title.position = "top", 
+                                             title.hjust = 0.5)) +
+  scale_y_discrete(expand = c(0,0)) +
+  scale_x_continuous(expand = c(0,0)) +  # expand increases distance from axis
+  labs(x = "codon position in alignment", y = "", fill = "variant effect") +
+  theme_logo(base_family = "Helvetica") +
+  theme(panel.grid = element_blank(), 
+        panel.grid.minor = element_blank(),
+        strip.background = element_blank(),
+        strip.text.x = element_blank(),
+        #panel.background = element_rect(fill = "grey85", color = "white"),
+        legend.position = c("bottom"),
+        legend.key = element_rect(linetype = "solid"),
+        legend.spacing = unit(4, "char"), 
+        legend.box.spacing = unit(0.1,"line")) 
+p
+
+```
+
+```{r AFB variants for PyMOL}
 coding_vcf <- ldply(.data = VCF, .fun = subset, 
@@
 
-(ref:AFB_structure_vars) A TIR1 structure (2P1Q) in gold, with residues
+(ref:AFBstructure) A TIR1 structure (2P1Q) in gold in complex with ASK1 in grey and auxin, inositol phosphate and an Aux/IAA degron in teal.
+TIR1 residues or analagous residues containing nonsynonymous polymorphisms in TIR1 and AFB1 colored in purple and in AFB2 and AFB3 colored in blue.
 
-```{r AFB_structure_vars, fig.cap='(ref:AFB_structure_vars)'}
-#knitr::include_graphics()
+```{r AFBstructure, fig.cap='(ref:AFBstructure)'}
+knitr::include_graphics(path = "structures/natural_variation_TIR1AFB1-3.png")
 ```
+
+Although most known functional regions are highly conserved in *AFB1*, there is a nonsynonymous polymorphism  in the oligomerization domain, only found in the Can-0 accession.
+Mutations in this domain of *TIR1* frequently have a semidominant effect on root phenotypes [@dezfulian_oligomerization_2016; @wright_insights_2017].
+Characterization of this allele and accession may help determine the role of *AFB1* in this pathway.
+
+**not sure this is very informative, and a map doesn't seem very exciting, perhaps a domain scale alignment plot would be better here, or above, and would allow us to comment more on where the functional diversity lies within the family.**
 
@@
 AFB1_positions <- annotation$positions
-AFB1_positions %<>% left_join(y = select(collapsed_aln_df, aln_pos, gene_info[gene_info$tair_symbol == "AFB1", "transcript_ID"]), by = "aln_pos") 
+AFB1_positions %<>% left_join(y = dplyr::select(collapsed_aln_df, aln_pos, gene_info[gene_info$tair_symbol == "AFB1", "transcript_ID"]), by = "aln_pos") 
 AFB1_positions$position <- AFB1_positions$AT4G03190.1

### 2018-09-09 manuscript/04-IAAs.Rmd
-TPL tree -AFB tree -Alignment plots
still some missing/broken chunks, but commit to sync to new computer

@@
 
-```{r IAAdivstats, fig.height=11}
+```{r IAAdivstats, fig.cap="(ref:IAAdivstats)"}
 gene_info <- geneInfoFromFile(
@@
 
-Within the available natural sequence variation, we found several missense polymorphisms affecting the EAR domain or degron of several canonical IAAs.
-However, no missense polymorphisms were present in the KR motif or residues critical for PB1-PB1 domain interactions (Supp.
-Fig.
-Alignment of canonical IAAs).
-The majority of missense mutations observed were in the variable regions between the conserved domains.
+```{r IAA alignment}
+alignment <- alignCDS(gene_info$transcript_ID)
+# make an alignment data frame
+aln_df <- makeAlnDF(alignment[[2]])
+
+# make a collapsed alignment for translating positions
+collapsed_aln_df <- 
+  aln_df %>% dplyr::select(aln_pos, seq_pos, seq_name) %>%
+  tidyr::spread(key = seq_name, value = seq_pos)
+```
 
@@
 (Supp Fig IAA6 diversity plot).
-```{r IAA alignment}
-alignment <- alignCDS(gene_info$transcript_ID)
-# make an alignment data frame
-aln_df <- makeAlnDF(alignment[[2]])
-
-# make a collapsed alignment for translating positions
-collapsed_aln_df <- 
-  aln_df %>% dplyr::select(aln_pos, seq_pos, seq_name) %>%
-  tidyr::spread(key = seq_name, value = seq_pos)
-```
 
@@
 ```{r Load IAA phylogeny and data}
-tree <- treeio::read.beast(file = "IAAs.nex.con.tre")
+tree <- treeio::read.beast(file = "trees/IAAs.nex.con.tre")
 
-IAA_data <- read.csv("IAAs_table.csv")
+IAA_data <- read.csv("tables/IAAs_table.csv")
 #tip_labels need to be in the first column
-IAA_data %<>% select(name, everything())
+IAA_data %<>% dplyr::select(name, everything())
 IAA_data$name %<>% as.character()
@@
 ```
-*I think IAA20, 30, and 15 should also be in class C. So there are really two large clades A and B and then there are truncations/loss of function/dominant variants scattered throughout, Class C.*
+**I think IAA20, 30, and 15 should also be in class C. So there are really two large clades A and B and then there are truncations/loss of function/dominant variants scattered throughout, Class C.**
 
@@
 
-Within the Group B *IAA* genes, there are four sister pairs---*IAA10*/*IAA11*, *IAA12*/*IAA13*, *IAA18*/*IAA26*, and *IAA20*/*IAA30* \@ref(IAA tree) [@remington_contrasting_2004].
+Within the Group B *IAA* genes, there are four sister pairs---*IAA10*/*IAA11*, *IAA12*/*IAA13*, *IAA18*/*IAA26*, and *IAA20*/*IAA30* \@ref(fig:IAAtree) [@remington_contrasting_2004].
 Interestingly, among these sister pairs only *IAA12* and *IAA13* have a noticeable difference in nonsynonymous diversity.
-The *Arabidopsis thaliana* EFP browser shows that **IAA10** and *IAA11* have almost identical expression patterns, with the exception of *IAA10* being more highly expressed in the developing seed, and *IAA11* being more highly expressed in root tissue during lateral root development **need to cite original data source here**.
+The *Arabidopsis thaliana* EFP browser shows that *IAA10* and *IAA11* have almost identical expression patterns, with the exception of *IAA10* being more highly expressed in the developing seed, and *IAA11* being more highly expressed in root tissue during lateral root development **need to cite original data source here**.
 In our analysis, *IAA10* and *IAA11* both showed functional conservation (${\pi_N/\pi_S}$ of 0.80 and 0.67 respectively) in support of *IAA10* and *IAA11* playing redundant roles.
@@
 ### New Loss-of-Function/Gain-of-Function alleles
+
 Accessions with functional variants can be potentially used to investigate the extent of redundancy/specialization within sister pairs.
@@
 
+Within the available natural sequence variation, we found several missense polymorphisms affecting the EAR domain or degron of several canonical clade A IAAs.
+However, no missense polymorphisms were present in the KR motif or residues critical for PB1-PB1 domain interactions (\@ref(fig:A-IAA-alignment)).
+The majority of missense mutations observed were in the variable regions between the conserved domains.
+
+(ref:A-IAA-alignment) 
+```{r A-IAA-alignment, fig.cap="(ref:A-IAA-alignment)"}
+A_IAAs <- subset(gene_info, tair_symbol %in% IAA_data[which(IAA_data$class == "A"), "name"])
+A_ARFs <- alignCDS(A_IAAs$transcript_ID)
+# make an alignment data frame
+A_IAAs <- makeAlnDF(A_IAAs[[2]])
+A_IAAs <- left_join(A_IAAs, dplyr::select(gene_info, "tair_locus",
+                                       "tair_symbol", "transcript_ID"),
+                        by = c("transcript_ID" = "transcript_ID"))
+A_IAAs$tair_symbol %<>% as.factor()
+A_IAAs$tair_symbol %<>% fct_reorder(
+    {as.integer(str_remove(A_IAAs$tair_symbol, 
+                           pattern = "ARF"))}, 
+    .desc = TRUE)
+A_IAAs <- addSNPsToAlnDF(A_IAAs, VCF)
+
+p <-ggplot(A_IAAs, aes(x = aln_pos, y = tair_symbol,
+                             group = seq_pos, text = variants)) +
+  geom_raster(data = A_IAAs[A_IAAs$seq_pos == "-",], fill = "lightgrey") +
+  geom_raster(data = na.omit(A_IAAs), 
+             # mapping = aes(fill = effects)) +
+              mapping = aes(fill = as.integer(as.factor(effects)))) +
+  #scale_fill_viridis(option = "A", discrete = TRUE) +
+   scale_fill_viridis(breaks = range(
+     as.integer(as.factor(na.omit(A_IAAs$effects)))),
+                      labels = c("deleterious", "neutral"),
+                      option = "A", 
+                      guide = guide_colorbar(ticks = FALSE, 
+                                             title.vjust = .8)) +
+  scale_y_discrete() +
+  # expand increases distance from axis
+  labs(x = "codon position in alignment", y = "", fill = "variant effect") +
+  theme_logo(base_family = "Helvetica") +
+  theme(panel.grid = element_blank(), panel.grid.minor = element_blank()) +
+  theme(strip.background = element_blank(),
+        strip.text.x = element_blank(),
+        legend.position = "bottom")  
+p
+```
+
 Several examples of possible gain-of-function missense variants were found in the degron of *IAA2*, *IAA3*, *IAA6*, *IAA7*, *IAA8*, *IAA13*, *IAA15*, *IAA18*, *IAA19*, *IAA27*, *IAA28* and *IAA29*.
@@
 Interestingly, degron missense variants occured more frequently in *IAA* genes that have a closely related sister, which may indicate redundant function within the sister-pairs.
-*I think this is an interesting finding, but I'm not sure about this hypothesis of redundant function.
+**I think this is an interesting finding, but I'm not sure about this hypothesis of redundant function.
 If pairs are acting redundantly and we assume the degron variants are dominant, we would expect the auxin signal they transmit (their function) to be lower, in sum, as a pair.
-So I would actually hypothesize that there has been some divergence in function as a pair.
-*
+So I would actually hypothesize that there has been some divergence in function as a pair.**
 ```{r Correlation between Distance to MRCA and prevalence of degron variants}
@@
 These sites are conserved across natural sequence variation, while several other lysine residues vary (Supp Figs alignment and map).
-In IAA6, there are 17 lysines of which ~6 are putative ubiquitination sites~ *I think this should be 5, K3, K32, K33, K91, and K97.
+In IAA6, there are 17 lysines of which ~6 are putative ubiquitination sites~ **I think this should be 5, K3, K32, K33, K91, and K97.
 Missense variants occured in K34Q, K140N *.
@@
 Only 1 of the ubiquitination sites varies?
-or one lysine, I think you mean this only 1 lysine that varies, but it is not ubiquitinated.
-Ubiquitylated Missense K129T** Natural sequence diversity may be used to narrow down a subset of lysines to assess for functional ubiquitination.
+or one lysine, I think you mean this only 1 lysine that varies, but it is not ubiquitinated.** Natural sequence diversity may be used to narrow down a subset of lysines to assess for functional ubiquitination.
 
-```{r}
+```{r IAA6/19-alignment, fig.cap="(ref:IAA6/19-alignment"}
 

### 2018-09-09 manuscript/05-TPLs.Rmd
-TPL tree -AFB tree -Alignment plots
still some missing/broken chunks, but commit to sync to new computer

--- 2018-07-13 manuscript/05-TPLs.Rmd
@@
 TPL/TPR co-repressors bind EAR domains via their C-terminal to LisH (CTLH) domains found near their N-termini (citations of pre-structure founding papers/reviews).
-Recent structural analyses of the TPL N-terminal domain have highlighted the precise interaction interface between TPL and AUX/IAA EAR domains, as well as the TPL-TPL dimerization and tetramerization motifs [@martin-arevalillo_structure_2017; @ke_structural_2015a].
+Recent structural analyses of the TPL N-terminal domain have highlighted the precise interaction interface between TPL and AUX/IAA EAR domains, as well as the TPL-TPL dimerization and tetramerization motifs [@martin-arevalillo_structure_2017; @ke_structural_2015].
 The residues required for higher-order multimers of TPL tetramers have also been identified [@ma_d53_2017].
@@
 
-```{r TPL_TPR alignment}
+```{r TPL-alignment-coding-vars}
+alignment <- alignCDS(gene_info$transcript_ID)
+#possible outgroups  AtLUG, At4g32551; AtLUH, At2g32700
+# make an alignment data frame
+aln_df <- makeAlnDF(alignment[[2]])
+
+#get
+VCF <- ldply(.data = VCF, .fun = subset, gt_GT != "0|0")
+VCF <- getCodingDiv(VCF)
+effect_order <- read.csv("~/Desktop/SNPeff_order.csv")
+
+# make a collapsed alignment for translating positions
+collapsed_aln_df <- 
+  aln_df %>% dplyr::select(aln_pos, seq_pos, seq_name) %>%
+  tidyr::spread(key = seq_name, value = seq_pos)
+```
+
+```{r Mask_ARF_alignment, eval=FALSE}
+aliscore(as.AAbin(alignment[[2]]), path = "/Volumes/SSD/Applications/ALISCORE_v2.0/Aliscore_v.2.0/")
+id <- scan(file = "/Volumes/SSD/Applications/ALISCORE_v2.0/Aliscore_v.2.0/input.fas_List_random.txt", sep = " ", quiet = TRUE)
+AAAlign_no_gaps <- as.matrix(alignment[[2]])[, id]
+write.nexus.data(AAAlign_no_gaps, file = "TPLs.nex", format = "protein")
+```
+```{bash ARF_phylogeny_construction, eval=FALSE}
+/Applications/MrBayes/mb
+
+execute TPLs.nex
+# set ARF17 as the outgroup
+outgroup AT1G77850.1
+
+lset nucmodel=protein Rates=invgamma
+# or include Nst=6  ngammacat=4 omegavar=M3 
+prset aamodelpr=fixed(jones) statefreqpr=fixed(empirical) 
+# initial run with aamodelpr=mixed yielded posterior prob of 1 for jones model
+propset ExtTBR$prob=0
+
+mcmcp ngen=250000 nchains=4 printfreq=1000 samplefreq=100 burnin=200
+
+mcmc
+sump
+sumt
+
+```
+```{r Load TPL phylogeny and data}
+tree <- treeio::read.beast(file = "trees/TPLs.nex.con.tre")
+
+TPL_data <- read.csv(system.file("shiny-app", "gene_family_data",
+                      "TPL_gene_ids.csv", package = "r1001genomes"))
+TPL_data %<>% left_join(y = div_stats, by = c("name" = "tair_symbol"))
+#tip_labels need to be in the first column
+TPL_data %<>% dplyr::select(name, everything())
+TPL_data$tair_locus %<>% as.character()
+
+TPL_data <- TPL_data[match(gsub(x = tree@phylo$tip.label, pattern = "\\.1",replacement = ""), TPL_data$tair_locus), ]
+
+tree@phylo$tip.label <- TPL_data$name
+
+```
+
+```{r Visualize TPL tree, fig.height=5, fig.width=5}
+p <- ggtree(tree, layout = "rectangular", color = "grey") %<+% 
+  TPL_data + theme_tree() + scale_x_continuous(expand = c(0, 1.2)) +
+  geom_hilight(node = 28, extend = 1) + 
+  geom_hilight(node = 25, extend = 1) +
+  geom_tippoint(mapping = 
+                  aes(size = 4*Pi_NS_Ratio, fill = Pi_NS_Ratio),
+                shape = 21, color = "white", show.legend = TRUE,
+                position = position_nudge(x = 0.01)) + 
+  geom_rootpoint(color = "grey") + 
+  scale_fill_viridis(values = rescale(
+    c(min(TPL_data$Pi_NS_Ratio),
+      max(TPL_data$Pi_NS_Ratio)), 
+                                    to = c(0,1)), guide = "colourbar") +
+  scale_size(guide = "none") + 
+  theme(legend.position = c(.1,.7)) + 
+  geom_nodelab(mapping = aes(label = node), nudge_x = -.2, size = 3)
+  #geom_nodelab(mapping = aes(label = round(prob,digits = 2)), nudge_x = -.25, size = 3) +
+  #geom_tiplab(mapping = aes(color = class), hjust = -0.3) + 
+p
+```
+
+```{r full-ARFs-alignment}
+ARF_order <- fct_inorder(na.exclude(p$data$label[base::order(p$data$y, decreasing = TRUE)]), ordered = TRUE)
+
+aln_df <- left_join(aln_df, dplyr::select(gene_info, "tair_locus",
+                                       "tair_symbol", "transcript_ID"),
+                        by = c("transcript_ID" = "transcript_ID"))
+aln_df$tair_symbol %<>% factor(levels = ARF_order, 
+                                       ordered = TRUE)
+aln_df$tair_symbol %<>% fct_rev()
+
+aln_df <- addSNPsToAlnDF(aln_df, VCF, effect_order = effect_order)
+aln_df$gap <- aln_df$seq_pos == "-"
+
+p <-ggplot(aln_df, aes(x = aln_pos, y = tair_symbol)) +
+  geom_raster(data = aln_df, 
+             # mapping = aes(fill = effects)) +
+              mapping = aes(fill = strength,
+                            alpha = gap)) +
+  scale_alpha_manual(values = c("TRUE" = 0, "FALSE" = 1), 
+                     labels = c("aligned", "gap"), 
+                     guide = guide_legend(override.aes = 
+                                            list(fill = c("grey85", "white")),
+                       title = "alignment",
+                       direction = "vertical")) + 
+   scale_fill_viridis(breaks = range(effect_order$strength),
+                      limits = range(effect_order$strength),
+                      na.value = "grey85",
+                      direction = -1,
+                      labels = c("neutral", "deleterious"),
+                      option = "A", 
+                      guide = guide_colorbar(ticks = FALSE, 
+                                             title.vjust = .8, 
+                                             title.position = "top", 
+                                             title.hjust = 0.5)) +
+  scale_y_discrete(expand = c(0,0)) +
+  scale_x_continuous(expand = c(0,0)) +  # expand increases distance from axis
+  labs(x = "codon position in alignment", y = "", fill = "variant effect") +
+  theme_logo(base_family = "Helvetica") +
+  theme(panel.grid = element_blank(), 
+        panel.grid.minor = element_blank(),
+        strip.background = element_blank(),
+        strip.text.x = element_blank(),
+        #panel.background = element_rect(fill = "grey85", color = "white"),
+        legend.position = c("bottom"),
+        legend.key = element_rect(linetype = "solid"),
+        legend.spacing = unit(4, "char"), 
+        legend.box.spacing = unit(0.1,"line")) 
+p
 

### 2018-09-09 manuscript/06-ARFs.Rmd
-TPL tree -AFB tree -Alignment plots
still some missing/broken chunks, but commit to sync to new computer

@@
 Class A ARFs (ARF5, ARF6, ARF7, ARF8 and ARF19) activate transcription.
-These ARFs have a glutamine-rich region in the middle of the protein that may mediate activation (Guilefoyle, 2007).
-It has recently been shown that the middle region of ARF5 interacts with the SWI/SNF chromatin remodeling ATPases BRAMA and SPLAYED, possibly to reduce nucleosome occupancy and allow for the recruitment of transcription machinery (Wu, 2016).
-Additionally, ARF7 interacts with Mediator subunits, directly tethering transcriptional activation machinery to its binding sites in the chromosome (Ito, 2016).
+These ARFs have a glutamine-rich region in the middle of the protein that may mediate activation [@guilfoyle_auxin_2007].
+It has recently been shown that the middle region of ARF5 interacts with the SWI/SNF chromatin remodeling ATPases BRAMA and SPLAYED, possibly to reduce nucleosome occupancy and allow for the recruitment of transcription machinery [@wu_auxinregulated_2015].
+Additionally, ARF7 interacts with Mediator subunits, directly tethering transcriptional activation machinery to its binding sites in the chromosome [@ito_auxindependent_2016].
 Class B and C ARFs are historically categorized as repressor ARFs, though the mechanism through which they confer repression has not been identified.
@@
 Canonical ARFs are comprised of three major domains.
-Recent crystallization of these domains have informed structure-function analysis of the ARFs (Boer, 2014; Nanao, 2014; Korasick, 2014).
-These domains are conserved throughout land plants (Dolf's new paper).
+Recent crystallization of these domains have informed structure-function analysis of the ARFs [@boer_structural_2014; @korasick_molecular_2014; @nanao_structural_2014].
+These domains are conserved throughout land plants [@mutte_origin_2018].
 ARFs share an N-terminal B3 DNA binding domain.
@@
 The C-terminal domain of the ARF is a protein-protein interaction domain mediating interactions among ARFs, between ARFs and other transcription factors, and between ARFs and the Aux/IAA repressors.
-This interaction domain was recently characterized as a Phox and Bem1 (PB1) domain, which is comprised of a positive and negative face with conserved basic and acidic residues, respectively (Nanao, 2014; Korasick, 2014).
-The dipolar nature of the PB1 domain may mediate multimerization by the pairwise interaction of these faces on different proteins as the ARF7 PB1 domain was crystallized as a multimer [@korasick_].
+This interaction domain was recently characterized as a Phox and Bem1 (PB1) domain, which is comprised of a positive and negative face with conserved basic and acidic residues, respectively [@korasick_molecular_2014; @nanao_structural_2014].
+The dipolar nature of the PB1 domain may mediate multimerization by the pairwise interaction of these faces on different proteins as the ARF7 PB1 domain was crystallized as a multimer [@korasick_molecular_2014].
 However, it is unclear whether ARF multimerization occurs or plays a developmental significant role in vivo.
-Interfering with ARF dimerization in either the DNA-binding proximal dimerization domain or the PB1 domain decreases the ability of class A ARFs to activate transcription in a heterologous yeast system (Pierre-Jerome, 2016).
+Interfering with ARF dimerization in either the DNA-binding proximal dimerization domain or the PB1 domain decreases the ability of class A ARFs to activate transcription in a heterologous yeast system [@pierre-jerome_functional_2016].
 
@@
 The majority of domain variation among ARFs occurs in the large B-class subfamily.
-The liverwort *Marchantia polymorpha* has a single representative ARF of each class (Flores-Sandoval, 2015).
-The expansion of these classes in flowering plants is the result of both whole genome and tandem duplication events (Remington, 2004).
+The liverwort *Marchantia polymorpha* has a single representative ARF of each class [@flores-sandoval_simple_2015].
+The expansion of these classes in flowering plants is the result of both whole genome and tandem duplication events [@remington_contrasting_2004].
 The growth of the ARF family may have allowed for the expansion of the quantity and complexity of loci regulated by the ARFs and subsequent expansion in their regulation of developmental processes.
 
-Class A ARFs are the most well-studied ARF subfamily---the five family members all act as transcriptional activators and have well-characterized, distinct developmental targets.
-Overall the diversity of class A ARFs was generally low, especially compared to the class B and C ARFs, suggesting that class A ARFs are central to auxin signal transduction and plant development.
-Analysis of class A ARF nonsynonymous diversity suggests that the majority of these ARFs are highly functionally conserved, with ${\pi_N/\pi_S}$ values much lower than 1 with the exception of ARF19, with ${\pi_N/\pi_S}$ value of 1.8.
+Class A *ARFs* are the most well-studied ARF subfamily---the five family members all act as transcriptional activators and have well-characterized, distinct developmental targets.
+Overall the diversity of class A *ARFs* was generally low, especially compared to the class B and C *ARFs*, suggesting that class A *ARFs* are central to auxin signal transduction and plant development.
+Analysis of class A *ARF* nonsynonymous diversity suggests that the majority of these *ARFs* are highly functionally conserved, with ${\pi_N/\pi_S}$ values much lower than 1 with the exception of *ARF19*, with ${\pi_N/\pi_S}$ value of 1.8.
 Comparing diversity within sister pairs, there is a similar trade-off as seen in most *IAA* sister pairs, with one sister being highly conserved and the other more divergent.
-ARF19 and ARF8 are the more divergent class A ARFs, with ${\pi_N/\pi_S}$ values at least three time those of their sisters, ARF7 and ARF6 respectively.
+*ARF19* and *ARF8* are the more divergent class A *ARFs*, with ${\pi_N/\pi_S}$ values at least three time those of their sisters, *ARF7* and *ARF6* respectively.
 This may suggest that ARF6 and ARF7 serve more essential purposes in plant development.
@@
                       "ARF_gene_ids.csv", package = "r1001genomes"))
-#annotation <- readAnnotationFile(filename = system.file("shiny-app", 
-  #"gene_family_data", "ARF_annotations.csv", 
-  #package = "r1001genomes"), gene_info = gene_info)
+annotation <- readAnnotationFile(filename = system.file("shiny-app", 
+  "gene_family_data", "ARF_annotations.csv", 
+  package = "r1001genomes"), gene_info = gene_info)
 
@@
 
-```{r ARF alignment}
+```{r ARF-alignment-coding-vars}
 alignment <- alignCDS(gene_info$transcript_ID)
@@
 aln_df <- makeAlnDF(alignment[[2]])
+
+#get
+VCF <- ldply(.data = VCF, .fun = subset, gt_GT != "0|0")
+VCF <- getCodingDiv(VCF)
+effect_order <- read.csv("~/Desktop/SNPeff_order.csv")
 
@@
 ```{r Load ARF phylogeny and data}
-tree <- treeio::read.beast(file = "ARFs.nex.con.tre")
+tree <- treeio::read.beast(file = "trees/ARFs.nex.con.tre")
 
@@
 #tip_labels need to be in the first column
-ARF_data %<>% select(name, everything())
+ARF_data %<>% dplyr::select(name, everything())
 ARF_data$tair_locus %<>% as.character()
@@
 ```{r Visualize ARF tree, fig.height=5, fig.width=5}
-p <- ggtree(tree, layout = "rectangular", color = "grey") %<+% ARF_data + theme_tree() + scale_x_continuous(expand = c(0, 1.2)) + geom_tiplab(hjust = -0.25) + 
-  geom_tippoint(mapping = aes(size = 4*Pi_NS_Ratio, fill = Pi_NS_Ratio), 
-                shape = 21, color = "white", show.legend = TRUE) + 
+p <- ggtree(tree, layout = "rectangular", color = "grey") %<+% 
+  ARF_data + theme_tree() + scale_x_continuous(expand = c(0, 1.2)) +
+  geom_hilight(node = 28, extend = 1) + 
+  geom_hilight(node = 25, extend = 1) +
+  geom_tippoint(mapping = 
+                  aes(size = 4*Pi_NS_Ratio, fill = Pi_NS_Ratio),
+                shape = 21, show.legend = TRUE,
+                position = position_nudge(x = 0.01)) + 
+  geom_rootpoint(color = "grey") + 
   scale_fill_viridis(values = rescale(c(min(ARF_data$Pi_NS_Ratio), 1,
                                         max(ARF_data$Pi_NS_Ratio)), 
-                                      to = c(0,1)), guide = "colourbar") + scale_size(guide = "none") + 
+                                    to = c(0,1)), guide = "colourbar") +
+  scale_size(guide = "none") + 
   theme(legend.position = c(.1,.7)) + 
-  geom_nodelab(mapping = aes(label = round(prob,digits = 2)), nudge_x = -.3, nudge_y = .5, size = 3) +
-  geom_tiplab(mapping = aes(color = class), hjust = -0.25) + 
-  scale_color_viridis(discrete = TRUE)
+  geom_nodelab(mapping = aes(label = node), nudge_x = -.2, size = 3) +
+  #geom_nodelab(mapping = aes(label = round(prob,digits = 2)), nudge_x = -.25, size = 3) +
+  geom_tiplab(color = "black", hjust = -0.3) 
+p %<>% flip(28,32) %>% rotate(35)
 p
-#ggsave("ARF_tree.pdf", width = 6, height = 6)
 ```
 
-### Class A ARFs 
+```{r full-ARFs-alignment}
+ARF_order <- fct_inorder(na.exclude(p$data$label[base::order(p$data$y, decreasing = TRUE)]), ordered = TRUE)
+
+aln_df <- left_join(aln_df, dplyr::select(gene_info, "tair_locus",
+                                       "tair_symbol", "transcript_ID"),
+                        by = c("transcript_ID" = "transcript_ID"))
+aln_df$tair_symbol %<>% factor(levels = ARF_order, 
+                                       ordered = TRUE)
+aln_df$tair_symbol %<>% fct_rev()
+
+aln_df <- addSNPsToAlnDF(aln_df, VCF)
+aln_df$gap <- aln_df$seq_pos == "-"
+
+p <-ggplot(aln_df, aes(x = aln_pos, y = tair_symbol)) +
+  geom_raster(data = aln_df, 
+             # mapping = aes(fill = effects)) +
+              mapping = aes(fill = strength,
+                            alpha = gap)) +
+  scale_alpha_manual(values = c("TRUE" = 0, "FALSE" = 1), 
+                     labels = c("aligned", "gap"), 
+                     guide = guide_legend(override.aes = 
+                                            list(fill = c("grey85", "white")),
+                       title = "alignment",
+                       direction = "vertical")) + 
+   scale_fill_viridis(breaks = range(effect_order$strength),
+                      limits = range(effect_order$strength),
+                      na.value = "grey85",
+                      direction = -1,
+                      labels = c("neutral", "deleterious"),
+                      option = "A", 
+                      guide = guide_colorbar(ticks = FALSE, 
+                                             title.vjust = .8, 
+                                             title.position = "top", 
+                                             title.hjust = 0.5)) +
+  scale_y_discrete(expand = c(0,0)) +
+  scale_x_continuous(expand = c(0,0)) +  # expand increases distance from axis
+  labs(x = "codon position in alignment", y = "", fill = "variant effect") +
+  theme_logo(base_family = "Helvetica") +
+  theme(panel.grid = element_blank(), 
+        panel.grid.minor = element_blank(),
+        strip.background = element_blank(),
+        strip.text.x = element_blank(),
+        #panel.background = element_rect(fill = "grey85", color = "white"),
+        legend.position = c("bottom"),
+        legend.key = element_rect(linetype = "solid"),
+        legend.spacing = unit(4, "char"), 
+        legend.box.spacing = unit(0.1,"line")) 
+p
+
+```
+
+### Class A *ARFs* 
 For all class A ARFs, the middle region of the protein was the predominant high diversity region.
@@
 ```{r A ARF Alignment}
-gene_info <- subset(gene_info, tair_symbol %in% c("ARF5", "ARF6", "ARF7", "ARF8", "ARF19"))
-alignment <- alignCDS(gene_info$transcript_ID)
+A_ARFs <- subset(gene_info, tair_symbol %in% c("ARF5", "ARF6", "ARF7", "ARF8", "ARF19"))
+A_ARFs <- alignCDS(A_ARFs$transcript_ID)
 # make an alignment data frame
-aln_df <- makeAlnDF(alignment[[2]])
-#BrowseSeqs(alignment[[2]])
-aln_df <- makeAlnDF(alignment[[2]])
-aln_df <- left_join(aln_df, dplyr::select(gene_info, "tair_locus",
+A_ARFs <- makeAlnDF(A_ARFs[[2]])
+
+A_ARFs <- left_join(A_ARFs, dplyr::select(gene_info, "tair_locus",
                                        "tair_symbol", "transcript_ID"),
                         by = c("transcript_ID" = "transcript_ID"))
-aln_df$tair_symbol %<>% as.factor()
-aln_df$tair_symbol %<>% fct_reorder(
-    {as.integer(str_remove(aln_df$tair_symbol, 
-                           pattern = "ARF"))}, 
-  .desc = TRUE)
+A_ARFs$tair_symbol %<>% factor(levels = ARF_order, 
+                                       ordered = TRUE)
+A_ARFs$tair_symbol %<>% fct_rev()
 
-aln_df <- chunkAlnDF(aln_df = aln_df, chunk_width = 100)
-VCF <- ldply(.data = VCF, .fun = subset, gt_GT != "0|0")
-VCF <- getCodingDiv(VCF)
-aln_df <- addSNPsToAlnDF(aln_df, VCF)
-A_ARFs <- subset(aln_df, tair_symbol %in% c("ARF5", "ARF6", "ARF7", "ARF8", "ARF19"))
-p <-ggplot(A_ARFs, aes(x = aln_pos, y = tair_symbol,
-                             group = seq_pos, text = variants))
-p <- p +
-  geom_raster(data = na.omit(A_ARFs), mapping = aes(fill = effects)) +
-  scale_fill_viridis(option = "A", discrete = TRUE) +
-  scale_y_discrete() +
-  # expand increases distance from axis
-  labs(x = "", y = "", fill = "") +
+A_ARFs <- addSNPsToAlnDF(A_ARFs, VCF, effect_order = effect_order)
+A_ARFs$gap <- A_ARFs$seq_pos == "-"
+
+p <-ggplot(A_ARFs, aes(x = aln_pos, y = tair_symbol)) +
+  geom_raster(data = A_ARFs, 
+             # mapping = aes(fill = effects)) +
+              mapping = aes(fill = strength,
+                            alpha = gap)) +
+  scale_alpha_manual(values = c("TRUE" = 0, "FALSE" = 1), 
+                     labels = c("aligned", "gap"), 
+                     guide = guide_legend(override.aes = 
+                                            list(fill = c("grey85", "white")),
+                       title = "alignment",
+                       direction = "vertical")) + 
+   scale_fill_viridis(breaks = range(effect_order$strength),
+                      limits = range(effect_order$strength),
+                      na.value = "grey85",
+                      direction = -1,
+                      labels = c("neutral", "deleterious"),
+                      option = "A", 
+                      guide = guide_colorbar(ticks = FALSE, 
+                                             title.vjust = .8, 
+                                             title.position = "top", 
+                                             title.hjust = 0.5)) +
+  scale_y_discrete(expand = c(0,0)) +
+  scale_x_continuous(expand = c(0,0)) +  # expand increases distance from axis
+  labs(x = "codon position in alignment", y = "", fill = "variant effect") +
   theme_logo(base_family = "Helvetica") +
-  theme(panel.grid = element_blank(), panel.grid.minor = element_blank()) +
-  theme(strip.background = element_blank(),
+  theme(panel.grid = element_blank(), 
+        panel.grid.minor = element_blank(),
+        strip.background = element_blank(),
         strip.text.x = element_blank(),
-        legend.box = "vertical")
+        #panel.background = element_rect(fill = "grey85", color = "white"),
+        legend.position = c("bottom"),
+        legend.key = element_rect(linetype = "solid"),
+        legend.spacing = unit(4, "char"), 
+        legend.box.spacing = unit(0.1,"line")) 
 p
@@
 
-Class A ARFs all contain the canonical B3 DNA-binding domain, PB1 interaction domain and glutamine-rich middle region (Guilefoyle, 2007).
+Class A ARFs all contain the canonical B3 DNA-binding domain, PB1 interaction domain and glutamine-rich middle region [@guilfoyle_auxin_2007].
 In the analyzed natural variation, ARF7 had several expansions of polyglutamine sequences in the middle region.
-Polyglutamine regions are known to readily expand and contract throughout evolutionary time due to replication error, and variation in polyglutamine length can be acted on by natural selection and have phenotypic consequences (Press, 2014).
+Polyglutamine regions are known to readily expand and contract throughout evolutionary time due to replication error, and variation in polyglutamine length can be acted on by natural selection and have phenotypic consequences [@press_overdue_2014].
 
 A few missense variants in *ARF5* highlighted potential functional variation in the class A ARFs.
-A high frequency allele (*What are the variants?*), resulting in substitutions of small amino acids on the positive face of the PB1 domain for bulky phenylalanines, may disrupt interactions with other ARF or Aux/IAA PB1 domains.
-Interestingly, these variants were always found together in accessions Tu-B2-3, Ru-2, Haes-1, Rd-0, 627ME-13Y1, Le-0, Bu-0 and Bsch-0 *Where are these?*.
+A high frequency allele *(What are the variants?)*, resulting in substitutions of small amino acids on the positive face of the PB1 domain for bulky phenylalanines, may disrupt interactions with other ARF or Aux/IAA PB1 domains.
+Interestingly, these variants were always found together in accessions Tu-B2-3, Ru-2, Haes-1, Rd-0, 627ME-13Y1, Le-0, Bu-0 and Bsch-0 **Where are these?**.
 Another highly represented polymorphism in ARF5 results in change in the 17th amino acid from a leucine to a methionine.
@@
 
-### Class B ARFs
+### Class B *ARFs*
 
-*ARF23* has a truncated DNA-binding domain and had a high ${\pi_N/\pi_S}$ value of 4.1 and several high-frequency nonsense variants, as may be expected of a gene that is pseudogenizing.
+ARF23 has a truncated DNA-binding domain and had a high ${\pi_N/\pi_S}$ value of 4.1 and several high-frequency nonsense variants, as may be expected of a gene that is pseudogenizing.
 
@@
 
-```{r ARF3_13_17 alignment}
+```{r B-ARF-alignment, fig.cap="(ref:B-ARFalignment)"}
+B_ARFs <- subset(gene_info, tair_symbol %in% 
+                   ARF_data[which(ARF_data$class == "B"), "name"])
+B_ARFs <- alignCDS(B_ARFs$transcript_ID)
+# make an alignment data frame
+B_ARFs <- makeAlnDF(B_ARFs[[2]])
+B_ARFs <- left_join(B_ARFs, dplyr::select(gene_info, "tair_locus",
+                                       "tair_symbol", "transcript_ID"),
+                        by = c("transcript_ID" = "transcript_ID"))
+B_ARFs$tair_symbol %<>% factor(levels = ARF_order, 
+                               ordered = TRUE)
+B_ARFs$tair_symbol %<>% fct_rev()
 
+B_ARFs <- addSNPsToAlnDF(B_ARFs, VCF, effect_order = effect_order)
+B_ARFs$gap <- B_ARFs$seq_pos == "-"
+
+p <-ggplot(B_ARFs, aes(x = aln_pos, y = tair_symbol)) +
+  geom_raster(data = B_ARFs, 
+             # mapping = aes(fill = effects)) +
+              mapping = aes(fill = strength,
+                            alpha = gap)) +
+  scale_alpha_manual(values = c("TRUE" = 0, "FALSE" = 1), 
+                     labels = c("aligned", "gap"), 
+                     guide = guide_legend(override.aes = 
+                                            list(fill = c("grey85", "white")),
+                       title = "alignment",
+                       direction = "vertical")) + 
+   scale_fill_viridis(breaks = range(effect_order$strength),
+                      limits = range(effect_order$strength),
+                      na.value = "grey85",
+                      direction = -1,
+                      labels = c("neutral", "deleterious"),
+                      option = "A", 
+                      guide = guide_colorbar(ticks = FALSE, 
+                                             title.vjust = .8, 
+                                             title.position = "top", 
+                                             title.hjust = 0.5)) +
+  scale_y_discrete(expand = c(0,0)) +
+  scale_x_continuous(expand = c(0,0)) +  # expand increases distance from axis
+  labs(x = "codon position in alignment", y = "", fill = "variant effect") +
+  theme_logo(base_family = "Helvetica") +
+  theme(panel.grid = element_blank(), 
+        panel.grid.minor = element_blank(),
+        strip.background = element_blank(),
+        strip.text.x = element_blank(),
+        #panel.background = element_rect(fill = "grey85", color = "white"),
+        legend.position = c("bottom"),
+        legend.key = element_rect(linetype = "solid"),
+        legend.spacing = unit(4, "char"), 
+        legend.box.spacing = unit(0.1,"line"))
+
+p
 ```
 
-ARF14, ARF15, **ARF16 I don't think this is correct ARF 16 is class C**, ARF20, and ARF21 all do not have one or multiple of the conserved acidic residues in the PB1 domain, and themselves form a subclade of the B ARFs.
-Interestingly, no non-synonymous variation is seen in these amino acid residues of these ARFs, suggesting that the amino acids that are in the regions may still be important for the function of these proteins.
+ARF14, ARF15, **ARF16 I don't think this is correct ARF 16 is class C, perhaps this should be 12?**, ARF20, and ARF21 all do not have one or multiple of the conserved acidic residues in the PB1 domain, and all are within a subclade of the B ARFs.
+Interestingly, no non-synonymous variation is seen in these important residues for Aux/IAA and ARF interaction, suggesting that the amino acids that are in the regions may still be important for the function of these proteins.
+**Might cut this, it is interesting that perhaps homo/heterodimerization is not as important for this clade of ARFs but I think this conclusion mostly comes from looking at the alignments and not from the natural variation.
+Perhaps if it looks like these ARFs have higher than expected, or more deleterious variation across their PB1 domains**
 
-### Class C ARFs
+### Class C *ARFs*
 
@@
 ```{r ARF10_DBD Variants distribution}
+C_ARFs <- subset(gene_info, tair_symbol %in% ARF_data[which(ARF_data$class == "C"), "name"])
+C_ARFs <- alignCDS(C_ARFs$transcript_ID)
+# make an alignment data frame
+C_ARFs <- makeAlnDF(C_ARFs[[2]])
+C_ARFs <- left_join(C_ARFs, dplyr::select(gene_info, "tair_locus",
+                                       "tair_symbol", "transcript_ID"),
+                        by = c("transcript_ID" = "transcript_ID"))
+C_ARFs$tair_symbol %<>% factor(levels = ARF_order, 
+                               ordered = TRUE)
+C_ARFs$tair_symbol %<>% fct_rev()
 
-```
+C_ARFs <- addSNPsToAlnDF(C_ARFs, VCF, 
+                         effect_order = effect_order)
+C_ARFs$gap <- C_ARFs$seq_pos == "-"
 
-```{r Arf16_PB_variants distribution}
+p <-ggplot(C_ARFs, aes(x = aln_pos, y = tair_symbol)) +
+  geom_raster(mapping = aes(fill = strength,
+                            alpha = gap)) +
+  scale_alpha_manual(values = c("TRUE" = 0, "FALSE" = 1), 
+                     labels = c("aligned", "gap"), 
+                     guide = guide_legend(override.aes = 
+                                            list(fill = c("grey85", "white")),
+                       title = "alignment",
+                       direction = "vertical")) + 
+   scale_fill_viridis(breaks = range(effect_order$strength),
+                      limits = range(effect_order$strength),
+                      na.value = "grey85",
+                      direction = -1,
+                      labels = c("neutral", "deleterious"),
+                      option = "A", 
+                      guide = guide_colorbar(ticks = FALSE, 
+                                             title.vjust = .8, 
+                                             title.position = "top", 
+                                             title.hjust = 0.5)) +
+  scale_y_discrete(expand = c(0,0)) +
+  scale_x_continuous(expand = c(0,0)) +  # expand increases distance from axis
+  labs(x = "codon position in alignment", y = "", fill = "variant effect") +
+  theme_logo(base_family = "Helvetica") +
+  theme(panel.grid = element_blank(), 
+        panel.grid.minor = element_blank(),
+        strip.background = element_blank(),
+        strip.text.x = element_blank(),
+        #panel.background = element_rect(fill = "grey85", color = "white"),
+        legend.position = c("bottom"),
+        legend.key = element_rect(linetype = "solid"),
+        legend.spacing = unit(4, "char"), 
+        legend.box.spacing = unit(0.1,"line"))
 
+p
 ```
@@
 Several members of class B (ARF2, 3 and 4) and C ARFs (ARF10, 16 and 17) are distinguished from other ARFs on the basis of their post transcriptional regulation by trans-acting small interfering RNAs (tasiRNAs) and microRNAs (miRNAs) [@rhoades_prediction_2002; @allen_micrornadirected_2005].
-**(@AMY Arf 6 and 8 targeted by miRNA 167 - check for conservation).** The miRNA target of class C ARFs showed a high degree of conservation (Fig.
+**(AMY check for conservation of Arf 6 and 8 targeted by miRNA 167).** The miRNA target of class C ARFs showed a high degree of conservation (Fig.
 ARF10_16_17_Alignment).

### 2018-09-09 manuscript/index.Rmd
-TPL tree -AFB tree -Alignment plots
still some missing/broken chunks, but commit to sync to new computer

--- 2018-07-14 manuscript/index.Rmd
@@
 ---
-title: A comprehensive analysis of natural sequence variation within the <i>Arabidopsis
-  thaliana</i> nuclear auxin signaling pathway using an accessible web application
-author:
-  affiliation: University of Washington
-  name: R. Clay Wright
+title: Predicting structure--function relationships using natural genetic variation 
+author:  |
+  | R. Clay Wright^[Corresponding author: wrightrc@vt.edu] $^1$, 
+  | Morgan O. Hamm $^2$, 
+  | Jennifer Nemhauser $^2$
+  | $^1$Biological Systems Engineering, Virginia Tech, Blacksburg, VA 24061, 
+  | $^2$Deparment of Biology, University of Washington, Seattle, WA 98195
+
 date: "`r Sys.Date()`"
 
-abstract: "This will be the abstract"
+abstract: |
+  The way genes and genomes have evolved can provide information about which genes are critical to an organisms fitness, which genes faciliate adaptation, and how robust genes and gene networks are to perturbation.
+Thousands of genomes have now been sequenced representing a large swath of plant species and genetic variation within many species which has only begun to be explored in the last few years.
+However, much of this data remains inaccessible to researchers without significant bioinformatics experience.
+Here, we have developed a webtool ViVa (Visualization of Variation) to facilitate analysis, by any researcher, of The 1001 *Arabidopsis* Genomes Project dataset on a gene, gene family, or gene network level.
+We We have demonstrated the utility and accessibility of this tool by analyzing the natural variation in the well-studied nuclear auxin signaling pathway as a group of researchers from an array of experience levels.
+Our analysis has provided further confirmation of existing knowledge of this well studied pathway and has also helped generate new hypotheses.
+These results suggest there is much to be learned about less studied gene families and networks from similar analysis, especially when paired with biochemical and genetic characterization.
+This application provides access for every researcher allowing the community to extract maximum value from this large and expensive-to-generate dataset.
+ViVa is also extensible to any other database of genetic variation such as The 3,000 Rice Genomes Project.
 
@@
 bibliography:
-- natural_variation.bib
+- natural-variation.bib
 - TIR1_AFB.bib
-- ARF_C.bib
-- AuxIAA_citations.bib
+- Aux_IAA.bib
+- ARF.bib
+- TPL.bib
 biblio-style: apalike
@@
 
-```{r setup, message=FALSE, warning=FALSE}
+```{r setup, message=FALSE, warning=FALSE, include=FALSE}
 knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE, collapse = TRUE, comment = "#>", tidy = TRUE, cache = TRUE, cache.lazy = FALSE, out.width = "90%", fig.height=4, fig.width=7, out.extra='style="margin: auto; display: block; padding-top: 15px;"')
@@
 #devtools::install_github("wrightrc/r1001genomes", ref = "auxin-natural-variation")
+library(bookdown)
 library(ape)
@@
 ## List of working titles  
+* Predicting structure--function relationships using natural genetic variation 
+* Accelerating structure--function mapping using natural genetic variation 
 * A comprehensive analysis of natural sequence variation within the <i>Arabidopsis thaliana</i> nuclear auxin signaling pathway 
 * **Add your suggestions**
-
-# Abstract
-
-* Motivating, high level statement  
-* Statement defining the cutting edge of your research area  
-* A statement of the form: "However, it remains unclear ..." that defines what problem you are  
-addressing.  
-* A statement of the form:"Here, we ..."  that concisely states your contribution.  
-* A series of 3?5 statement about your results.  
-* A concluding statement.
 

### 2018-09-09 manuscript/02-methods.Rmd

@@
 so we can revert back to using Nei's original formula \@ref(eq:nei), modifying it slightly,
-replacing $x_i$ with $\frac{n_i}{n}$, $n_i$ being the number of sequences in the population of the form $N_i$ at location k: subscripts i and k are summed only over unique forms [need to clarify this]
+replacing $x_i$ with $\frac{n_i}{n}$, $n_i$ being the number of sequences in the population of the form $N_i$ at location k.
 
@@
 \end{equation}
+
+Note that in equation \@ref(eq:nei) subscripts i and j are summed over all sequences in the population, however in equation \@ref(eq:nei-simp1) i and j are only summed over unique variants at a particular location k.
 

### 2018-09-09 manuscript/05-TPLs.Rmd

@@
 What is EH1?
-Is TLE really binding to EH1 DNA?**
+Is
 
@@
 
-```{r Mask_ARF_alignment, eval=FALSE}
+(ref:TPLsAlignment) **Alignment of A. thaliana TPL/TPR genes showing variants.** This alignment should probably be paired down to a key region rather than the eniter AA sequence.
+
+```{r TPLsAlignment, fig.height = 20, fig.cap = "(ref:TPLsAlignment)"}
+
+alignment <- alignCDS(gene_info$transcript_ID)
+
+aln_df <- makeAlnDF(alignment[[2]])
+aln_df <- left_join(aln_df, dplyr::select(gene_info, "tair_locus",
+                                       "tair_symbol", "transcript_ID"),
+                        by = c("transcript_ID" = "transcript_ID"))
+aln_df$tair_symbol <- fct_rev(aln_df$tair_symbol)
+aln_df <- chunkAlnDF(aln_df = aln_df, chunk_width = 55)
+VCF <- ldply(.data = VCF, .fun = subset, gt_GT != "0|0")
+VCF <- getCodingDiv(VCF)
+aln_df <- addSNPsToAlnDF(aln_df, VCF)
+
+# aln_df_chunk <- chunkAlnDF(aln_df = aln_df, chunk_width = 55)
+p <-ggplot(aln_df_chunk, aes(x = aln_pos, y = tair_symbol,
+                             group = seq_pos, text = variants))
+p <- p +
+  geom_tile(data = na.omit(aln_df_chunk), mapping = aes(fill = effects),
+            width = 1, height = 1, alpha = 0.8) +
+  geom_text(aes(label=letter), alpha= 1, family = "Courier") +
+  scale_fill_brewer(type = "qual", palette = 1, direction = -1) +
+  scale_x_continuous(breaks=seq(1,max(aln_df_chunk$aln_pos), by = 10)) +
+  scale_y_discrete() +
+  # expand increases distance from axis
+  labs(x = "", y = "", fill = "variant") +
+  theme_logo(base_family = "Helvetica") +
+  theme(panel.grid = element_blank(), panel.grid.minor = element_blank()) +
+  facet_wrap(facets = ~chunk, ncol = 1, scales = "free") +
+  theme(strip.background = element_blank(),
+        strip.text.x = element_blank(),
+        legend.box = "vertical")
+p
+```
+
+```{r Mask_TPL_alignment, eval=FALSE}
 aliscore(as.AAbin(alignment[[2]]), path = "/Volumes/SSD/Applications/ALISCORE_v2.0/Aliscore_v.2.0/")

### 2018-10-19 manuscript/03-AFBs.Rmd
compiled to word and sent to Jennifer for feedback

@@
   names(dt) <- paste0("**", names(dt), "**")
-  kable(dt, caption = "(\\#tab:TIR1AFBdivstats)") 
+  kable(dt) 
 }
@@
 ```
-```{bash IAA_phylogeny_construction, eval=FALSE}
+```{bash AFB_phylogeny_construction, eval=FALSE}
 cd ~/Desktop/natural-variation-webtool/manuscript/trees
@@
                                       to = c(0,1)), guide = "colourbar") + scale_size(guide = "none") +
+  labs(fill = expression(pi [N] / pi [S])) + 
   theme(legend.position = c(.1,.7)) + 
-  geom_nodelab(mapping = aes(label = round(prob,digits = 2)), nudge_x = -.3, nudge_y = .5, size = 3) + 
+  #geom_nodelab(mapping = aes(label = node)) +
+  geom_nodelab(mapping = aes(label = round(as.numeric(prob),digits = 2)), nudge_x = -.2, nudge_y = .2, size = 3) + 
   geom_tiplab(hjust = -0.25) + 
   scale_color_viridis(discrete = TRUE)  
+p <- p %>% ggtree::rotate(12) %>% ggtree::rotate(9)
 p
@@
 aln_df$tair_symbol %<>% fct_rev()
+VCF <- ldply(.data = VCF, .fun = subset, gt_GT != "0|0")
+VCF <- getCodingDiv(VCF)
 
@@
 
-p <-ggplot(aln_df, aes(x = aln_pos, y = tair_symbol)) +
-  geom_raster(data = aln_df, 
-             # mapping = aes(fill = effects)) +
-              mapping = aes(fill = strength,
+p <-ggplot(subset(aln_df, tair_symbol != "COI1"), aes(x = aln_pos, y = tair_symbol)) +
+  geom_raster(mapping = aes(fill = strength,
                             alpha = gap)) +
@@
                        direction = "vertical")) + 
-   scale_fill_viridis(breaks = range(effect_order$strength),
-                      limits = range(effect_order$strength),
+   scale_fill_viridis(breaks = range(SNPeff_order$strength),
+                      limits = range(SNPeff_order$strength),
                       na.value = "grey85",
@@
 
-```{r AFB variants for PyMOL}
+```{r AFB variants for PyMOL, eval=FALSE}
 coding_vcf <- ldply(.data = VCF, .fun = subset, 
-             !is.na(Transcript_ID) & gt_GT != "0|0")
+             !is.na(transcript_ID) & gt_GT != "0|0")
 aln_df <- addSNPsToAlnDF(aln_df, coding_vcf, seq_name = Transcript_ID, 
@@
 ```{r AFB1 diversity plot}
-coding_variants <- getCodingDiv(VCF[[gene_info[gene_info$tair_symbol == "AFB1", "transcript_ID"]]])
+coding_variants <- VCF[VCF$Transcript_ID == gene_info[gene_info$tair_symbol == "AFB1", "transcript_ID"], ]
 p <- plotCodingDiv(uniqueCodingVars = coding_variants)
@@
 aln_df <- chunkAlnDF(aln_df = aln_df, chunk_width = 55)
-VCF <- ldply(.data = VCF, .fun = subset, gt_GT != "0|0")
-VCF <- getCodingDiv(VCF)
-aln_df <- addSNPsToAlnDF(aln_df, VCF)
+aln_df <- addSNPsToAlnDF(aln_df, VCF, effect_order = SNPeff_order)
 F_box <- subset(aln_df, aln_pos <= 96)
@@
   geom_tile(data = na.omit(F_box), mapping = aes(fill = effects),
-            width = 1, height = 1, alpha = 0.8) +
+            width = 1, height = 1, alpha = 0.5) +
   geom_text(aes(label=letter), alpha= 1, family = "Courier") +
-  scale_fill_brewer(type = "qual", palette = 1, direction = -1) +
   scale_x_continuous(breaks=seq(1,max(F_box$aln_pos), by = 10)) +
@@
   # expand increases distance from axis
-  labs(x = "", y = "", fill = "variant") +
+  labs(x = "", y = "") +
   theme_logo(base_family = "Helvetica") +
@@
         strip.text.x = element_blank(),
-        legend.box = "vertical")
+        legend.box = "vertical") + 
+  scale_fill_viridis(option = "A", discrete = TRUE)
 p

### 2018-10-19 manuscript/05-TPLs.Rmd
compiled to word and sent to Jennifer for feedback

--- 2018-09-09 manuscript/05-TPLs.Rmd
@@
 #annotation <- readAnnotationFile(filename = system.file("shiny-app", "gene_family_data",
-#                      "IAA_annotations.csv", package = "r1001genomes"), gene_info = gene_info)
+#                      "TPL_annotations.csv", package = "r1001genomes"), gene_info = gene_info)
 
@@
 VCF <- getCodingDiv(VCF)
-effect_order <- read.csv("~/Desktop/SNPeff_order.csv")
 
@@
 
-```{r Mask_ARF_alignment, eval=FALSE}
+```{r Mask_TPL_alignment, eval=FALSE}
 aliscore(as.AAbin(alignment[[2]]), path = "/Volumes/SSD/Applications/ALISCORE_v2.0/Aliscore_v.2.0/")
@@
 ```
-```{bash ARF_phylogeny_construction, eval=FALSE}
+```{bash TPL_phylogeny_construction, eval=FALSE}
 /Applications/MrBayes/mb
@@
                       "TPL_gene_ids.csv", package = "r1001genomes"))
+TPL_data$name <- as.character(TPL_data$name)
 TPL_data %<>% left_join(y = div_stats, by = c("name" = "tair_symbol"))
@@
 
-```{r Visualize TPL tree, fig.height=5, fig.width=5}
-p <- ggtree(tree, layout = "rectangular", color = "grey") %<+% 
-  TPL_data + theme_tree() + scale_x_continuous(expand = c(0, 1.2)) +
-  geom_hilight(node = 28, extend = 1) + 
-  geom_hilight(node = 25, extend = 1) +
-  geom_tippoint(mapping = 
-                  aes(size = 4*Pi_NS_Ratio, fill = Pi_NS_Ratio),
-                shape = 21, color = "white", show.legend = TRUE,
-                position = position_nudge(x = 0.01)) + 
-  geom_rootpoint(color = "grey") + 
-  scale_fill_viridis(values = rescale(
-    c(min(TPL_data$Pi_NS_Ratio),
+```{r Visualize TPL tree, fig.height=2, fig.width=3}
+p <- ggtree(tree, layout = "rectangular", color = "grey") %<+% TPL_data + theme_tree() + scale_x_continuous(expand = c(0, 1.2)) + 
+  geom_tippoint(mapping = aes(size = 4*Pi_NS_Ratio, fill = Pi_NS_Ratio), shape = 21, color = "white", show.legend = TRUE) + 
+  scale_fill_viridis(values = rescale(c(min(TPL_data$Pi_NS_Ratio),
                                         max(TPL_data$Pi_NS_Ratio)), 
-                                    to = c(0,1)), guide = "colourbar") +
-  scale_size(guide = "none") + 
+                                      to = c(0,1)), guide = "colourbar") + scale_size(guide = "none") +
+  labs(fill = expression(pi [N] / pi [S])) + 
   theme(legend.position = c(.1,.7)) + 
-  geom_nodelab(mapping = aes(label = node), nudge_x = -.2, size = 3)
-  #geom_nodelab(mapping = aes(label = round(prob,digits = 2)), nudge_x = -.25, size = 3) +
-  #geom_tiplab(mapping = aes(color = class), hjust = -0.3) + 
+  #geom_nodelab(mapping = aes(label = node)) +
+  geom_nodelab(mapping = aes(label = round(as.numeric(prob),digits = 2)), nudge_x = -.2, nudge_y = .2, size = 3) + 
+  geom_tiplab(hjust = -0.25) 
 p
@@
 
-```{r full-ARFs-alignment}
-ARF_order <- fct_inorder(na.exclude(p$data$label[base::order(p$data$y, decreasing = TRUE)]), ordered = TRUE)
+(ref:full-TPLs-alignment) **Alignment of A. thaliana TPL/TPR genes showing variants.** This alignment should probably be paired down to a key region rather than the eniter AA sequence.
+
+```{r full-TPLs-alignment, fig.cap= "(ref:full-TPLs-alignment)"}
+TPL_order <- fct_inorder(na.exclude(p$data$label[base::order(p$data$y, decreasing = TRUE)]), ordered = TRUE)
 
@@
                         by = c("transcript_ID" = "transcript_ID"))
-aln_df$tair_symbol %<>% factor(levels = ARF_order, 
+aln_df$tair_symbol %<>% factor(levels = TPL_order, 
                                        ordered = TRUE)
@@
 
-aln_df <- addSNPsToAlnDF(aln_df, VCF, effect_order = effect_order)
+aln_df <- addSNPsToAlnDF(aln_df, VCF, effect_order = SNPeff_order)
 aln_df$gap <- aln_df$seq_pos == "-"

### 2018-10-19 manuscript/06-ARFs.Rmd
compiled to word and sent to Jennifer for feedback

@@
   geom_tiplab(color = "black", hjust = -0.3) 
-p %<>% flip(28,32) %>% rotate(35)
+p %<>% ggtree::flip(28,32) %>% ggtree::rotate(35)
 p
@@
 
-```{r A ARF Alignment}
+```{r A-ARF-Alignment}
 A_ARFs <- subset(gene_info, tair_symbol %in% c("ARF5", "ARF6", "ARF7", "ARF8", "ARF19"))
@@
 
-A_ARFs <- addSNPsToAlnDF(A_ARFs, VCF, effect_order = effect_order)
+A_ARFs <- addSNPsToAlnDF(A_ARFs, VCF, effect_order = SNPeff_order)
 A_ARFs$gap <- A_ARFs$seq_pos == "-"
@@
 
-B_ARFs <- addSNPsToAlnDF(B_ARFs, VCF, effect_order = effect_order)
+B_ARFs <- addSNPsToAlnDF(B_ARFs, VCF, effect_order = SNPeff_order)
 B_ARFs$gap <- B_ARFs$seq_pos == "-"
@@
 C_ARFs <- addSNPsToAlnDF(C_ARFs, VCF, 
-                         effect_order = effect_order)
+                         effect_order = SNPeff_order)
 C_ARFs$gap <- C_ARFs$seq_pos == "-"

### 2018-10-19 manuscript/index.Rmd
compiled to word and sent to Jennifer for feedback

@@
 library(ggpmisc)
+library(gginnards)
 library(ggthemes)

### 2018-11-12 manuscript/01-intro.Rmd
clean up and rewrite of results section

--- 2018-07-16 manuscript/01-intro.Rmd
@@
 
-The first genome sequence of *Arabidopsis thaliana* facilitated rapid advancement of plant biology through molecular genetics.
-Since this initial genome, massive scaling of sequencing technology has allowed a pioneering group to survey the genomic variation in natural *Arabidopsis thaliana* populations.
-This valuable population genetics resource has led to several associations of genetic loci with phenotypic traits and provided insights into how selective pressure has influenced the evolution of plant genomes.
-Outside of genome-wide studies however, this valuable resequencing dataset has seen little use.
+The first genome sequence of *Arabidopsis thaliana* facilitated rapid advancement of plant biology [@thearabidopsisgenomeinitiative_analysis_2000].
+Since this initial genome, massive scaling of sequencing technology has allowed the survey the genomic variation in natural *Arabidopsis thaliana* populations [@nordborg_pattern_2005; @borevitz_genomewide_2007; @weigel_1001_2009].
+This valuable population genetics resource has led to several associations of genetic loci with phenotypic traits and provided insights into how selective pressure has influenced the evolution of plant genomes [@long_massive_2013; @atwell_genomewide_2010; @clark_common_2007].
+Despite the high value of this compendium of genetic variation, its use has been largely limited to investigations of population genetics and for genome-wide association studies.
 
-Natural genomic variation also provides a set of feasible functional variation at the gene, gene family and gene network level.
-Characterizing this variation can provide insights into the function and evolution of genes, families and networks [@delker_natural_2010; @wright_insights_2017] .
-Here, we present a web application and R-package for gene, gene family and gene network analysis of this dataset.
-This application allows plant molecular biologists without bioinformatics experience to use genomic variation to formulate hypotheses as to the sequence/function/phenotype relationships within their gene, family, or network of their interest.
+Natural genetic variation provides a catalog of permissible polymorphisms which may facilitate exploration and prediction of genotype/function/phenotype relationships at the gene, gene family and network scales [@joly-lopez_developing_2016].
+Outside of large effect variants, these relationships are difficult to map.
+Such studies in plants have the added obstacles of large gene families and the still quite time and resource-intensive process of allele replacement.
+Massively parallel assays of variant effects have revolutionized genetic diagnostics and personalized medicine [@starita_variant_2017; @gasperini_power_2016; @matreyek_platform_2017].
+We envision the use of plant natural variation datasets as a tool to similarly revolutionize breeding and genetic engineering of crop plants by rapidly advancing our understanding of genotype/function/phenotype relationships.
+Biochemical or synthetic assays of gene function and quantitative phenotyping of the germplasm of collections of natural variation can facilitate functional and phenotypic quantification [@wright_insights_2017].
+However, the high dimensionality of natural variation datasets often limits access to those with bioinformatics expertise.
 
-This application allows easy access to  
-1. lists of missense polymorphisms to facilitate biochemical assays of variant effects [@starita_variant_2017],  
-2. family-wise alignments of variants to facilitate de novo functional domain identification [@melamed_combining_2015],  
-3.
-lists of accessions containing missense (or any type of) polymorphisms to facilitate segregation analysis and measurment of the distribution of variant effects on phenotype [@park_distributions_2017],
-4. as well as the study of gene and network evolution [@delker_natural_2010; @kliebenstein_role_2008].
-
-Here, we demonstrate the utility of this tool through comprehensive analysis of genomic variation in the nuclear auxin signaling pathway.
-We identify potential functional variation within the coding sequence of the genes within this network and formulate hypotheses regarding the functional implications of this variation and the evolution of these gene families.
-Our analysis provides further confirmation of much of the existing knowledge of these genes demonstrating the validity of this approach.
-We also identify natural alleles of several genes and accessions bearing these alleles to facilitate future characterization structure/function, genotype/phenotype, and epistatic relationships.
+Here, we present ViVa: a web application and R-package for Visualizing Variation, which allows plant molecular biologists of any level access to the 1001 genomes database.
+ViVa facilitates formulation of hypotheses as to the genotype/function/phenotype relationships within any gene, family, or network.
+Using ViVa researchers may: 1.
+Identify polymorphisms to facilitate biochemical assays of variant effects [@starita_variant_2017; @wright_insights_2017]; 2.
+Produce family-wise alignments of variants to facilitate de novo functional domain identification [@melamed_combining_2015]; 3.
+Generate lists of accessions containing polymorphisms to facilitate phenotypic analysis of gene variant effects [@park_distributions_2017]; and 4.
+Quantify metrics of genetic diversity to facilitate the study of gene, gene family and network evolution [@delker_natural_2010; @kliebenstein_role_2008].

### 2018-11-12 manuscript/02-results.Rmd
clean up and rewrite of results section

# Results and Discussion {#rd}

## An introduction to ViVa 
**Morgan will provide screenshots**

## Visualizing Variation within the auxin signaling pathway

To test the useability and accessibility of this tool we assembled a group of alpha testers comprising postdoctoral, graduate, and undergraduate researchers. This team of researchers analyzed natural variation in the nuclear auxin signaling pathway. This signaling pathway has been well studied, revealing abundant functional knowledge and crystal structures of several protein domains, providing several points of ground truth reference for comparing structure-function variation and natural genetic variation.

Assuming there is selective pressure against mutations which alter a protein's function, nonsynonymous genetic variation in critical functional domains of a protein should be low. Therefore scanning gene coding sequences for regions of low nonsynonymous diversity, may predict functional domains. Within the auxin signaling pathway several domains have been identified as critical to transmitting auxin signals.

Perhaps the most critical domain is the degron of the Aux/IAA transcriptional repressor proteins. Mutations which prevent this domain from being recognized and marked for proteasomal degradation by SCF^TIR1/AFB^ are dominant and often result in obvious phenotypes [@berleth_auxin_2004; @yang_iaa1_2004]. As expected this critical functional region is highly conserved at the amino acid level \@ref(fig:A-IAA-alignment). Similarly, the EAR motif which facilitates Aux/IAA interaction with TPL/TPR transcriptional repressors, are also conserved.

```{r Load IAA gene info and VCF}
gene_info <- geneInfoFromFile(
  fname = system.file("shiny-app", "gene_family_data",
                      "IAA_gene_ids.csv", package = "r1001genomes"))
#annotation <- readAnnotationFile(filename = system.file("shiny-app", "gene_family_data",
#                      "IAA_annotations.csv", package = "r1001genomes"), gene_info = gene_info)

VCF <- readRDS(system.file("shiny-app", "gene_family_data",
                      "IAAs_VCF.rds", package = "r1001genomes"))
```
```{r IAA alignment}
alignment <- alignCDS(gene_info$transcript_ID)
# make an alignment data frame
aln_df <- makeAlnDF(alignment[[2]])

# make a collapsed alignment for translating positions
collapsed_aln_df <- 
  aln_df %>% dplyr::select(aln_pos, seq_pos, seq_name) %>%
  tidyr::spread(key = seq_name, value = seq_pos)
```
```{r Mask_IAA_alignment, eval=FALSE}
aliscore(as.AAbin(alignment[[2]]), path = "/Volumes/SSD/Applications/ALISCORE_v2.0/Aliscore_v.2.0/")
id <- scan(file = "input.fas_List_random.txt", sep = " ", quiet = TRUE)
AAAlign_no_gaps <- as.matrix(alignment[[2]])[, id]
write.nexus.data(AAAlign_no_gaps, file = "manuscript/IAAs.nex", format = "protein")
```
```{bash IAA_phylogeny_construction, eval=FALSE}
~/Desktop/natural-variation-webtool/manuscript

/Applications/MrBayes/mb

execute IAAs.nex
# set IAA33 as the outgroup
outgroup AT5G57420.1

lset nucmodel=protein Rates=invgamma
# or include Nst=6  ngammacat=4 omegavar=M3 
prset aamodelpr=fixed(jones) statefreqpr=fixed(empirical) 
# initial run with aamodelpr=mixed yielded posterior prob of 1 for jones model
propset ExtTBR$prob=0

mcmcp ngen=250000 nchains=4 printfreq=1000 samplefreq=100 burnin=200

mcmc
sump
sumt

```
```{r Load IAA phylogeny and data}
tree <- treeio::read.beast(file = "trees/IAAs.nex.con.tre")

IAA_data <- read.csv("tables/IAAs_table.csv")
#tip_labels need to be in the first column
IAA_data %<>% dplyr::select(name, everything())
IAA_data$name %<>% as.character()
IAA_data$AGI %<>% as.character()

IAA_data <- IAA_data[match(gsub(x = tree@phylo$tip.label, pattern = "\\.1",replacement = ""), IAA_data$AGI), ] 
tree@phylo$tip.label <- IAA_data$name
```

(ref:A-IAA-alignment) **Canonical IAAs have conserved Degron and EAR motifs** Protein sequences were aligned [@wright_decipher_2015] and variants were mapped to this alignment and colored according to the predicted functional effect of the most deleterious variant at that position, with light colors being less deleterious and darker colors being more deleterious. Red indicates nonsynonymous variants. Color scale is explained in {#methods}. EAR motif is highlighted by a black box including alignment positions 70-74. Degron domain is highlighted by a gray box including alignment positions 194-201.
```{r A-IAA-alignment, fig.cap="(ref:A-IAA-alignment)"}
A_IAAs <- subset(gene_info, tair_symbol %in% IAA_data[which(IAA_data$Class == "A"), "name"])
A_IAAs <- alignCDS(A_IAAs$transcript_ID)
# make an alignment data frame
A_IAAs <- makeAlnDF(A_IAAs[[2]])
A_IAAs <- left_join(A_IAAs, dplyr::select(gene_info, "tair_locus",
                                       "tair_symbol", "transcript_ID"),
                        by = c("transcript_ID" = "transcript_ID"))
A_IAAs$tair_symbol %<>% as.factor()
A_IAAs$tair_symbol %<>% fct_reorder(
    {as.integer(str_remove(A_IAAs$tair_symbol, 
                           pattern = "IAA"))}, 
    .desc = TRUE)
VCF <- ldply(.data = VCF, .fun = subset, gt_GT != "0|0")
coding_VCF <- getCodingDiv(VCF)
A_IAAs <- addSNPsToAlnDF(aln_df = A_IAAs,SNPs = coding_VCF, effect_order = SNPeff_order)
A_IAAs$gap <- A_IAAs$seq_pos == "-"

p <-ggplot(A_IAAs, aes(x = aln_pos, y = tair_symbol)) +
  geom_raster(data = A_IAAs, 
             # mapping = aes(fill = effects)) +
              mapping = aes(fill = strength,
                            alpha = gap)) +
  scale_alpha_manual(values = c("TRUE" = 0, "FALSE" = 1), 
                     labels = c("aligned", "gap"), 
                     guide = guide_legend(override.aes = 
                                            list(fill = c("grey85", "white")),
                       title = "alignment",
                       direction = "vertical")) + 
   scale_fill_viridis(breaks = range(SNPeff_order$strength),
                      limits = range(SNPeff_order$strength),
                      na.value = "grey85",
                      direction = -1,
                      labels = c("neutral", "deleterious"),
                      option = "A", 
                      guide = guide_colorbar(ticks = FALSE, 
                                             title.vjust = .8, 
                                             title.position = "top", 
                                             title.hjust = 0.5)) +
  scale_y_discrete(expand = c(0,0)) +
  scale_x_continuous(expand = c(0,0)) +  # expand increases distance from axis
  labs(x = "codon position in alignment", y = "", fill = "variant effect") +
  theme_logo(base_family = "Helvetica") +
  theme(panel.grid = element_blank(), 
        panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        strip.text.x = element_blank(),
        #panel.background = element_rect(fill = "grey85", color = "white"),
        legend.position = c("bottom"),
        legend.key = element_rect(linetype = "solid"),
        legend.spacing = unit(4, "char"), 
        legend.box.spacing = unit(0.1,"line")) 
p + geom_rect(mapping = aes(xmin = 193, xmax = 202, ymin = 0, ymax = length(levels((tair_symbol)))+1), color= "black", fill = NA) + geom_rect(mapping = aes(xmin = 69, xmax = 75, ymin = 0, ymax = length(levels((tair_symbol)))+1), color= "black", fill = NA)
```

The Aux/IAA proteins bind to ARF transcription factors allowing the Aux/IAA's to effect transcriptional repression at specific loci. Thus in the ARF transcription factor proteins we expect the DNA-binding function to be conserved. Indead, the DNA-binding domains of the ARFs had very few, low-diversity missense polymorphisms, as did the PB1 domain which facilitates ARF-Aux/IAA interaction \@ref(fig:A-ARF-Alignment).

```{r ARF diversity stats}
gene_info <- geneInfoFromFile(
  fname = system.file("shiny-app", "gene_family_data",
                      "ARF_gene_ids.csv", package = "r1001genomes"))

VCF <- readRDS(system.file("shiny-app", "gene_family_data",
                      "ARF_VCF.rds", package = "r1001genomes"))

VCF <- ldply(.data = VCF, .fun = subset, gt_GT != "0|0")
VCF <- getCodingDiv(VCF)
```

```{r Mask_ARF_alignment, eval=FALSE}
aliscore(as.AAbin(alignment[[2]]), path = "/Volumes/SSD/Applications/ALISCORE_v2.0/Aliscore_v.2.0/")
id <- scan(file = "/Volumes/SSD/Applications/ALISCORE_v2.0/Aliscore_v.2.0/input.fas_List_random.txt", sep = " ", quiet = TRUE)
AAAlign_no_gaps <- as.matrix(alignment[[2]])[, id]
write.nexus.data(AAAlign_no_gaps, file = "ARFs.nex", format = "protein")
```
```{bash ARF_phylogeny_construction, eval=FALSE}
/Applications/MrBayes/mb

execute ARFs.nex
# set ARF17 as the outgroup
outgroup AT1G77850.1

lset nucmodel=protein Rates=invgamma
# or include Nst=6  ngammacat=4 omegavar=M3 
prset aamodelpr=fixed(jones) statefreqpr=fixed(empirical) 
# initial run with aamodelpr=mixed yielded posterior prob of 1 for jones model
propset ExtTBR$prob=0

mcmcp ngen=250000 nchains=4 printfreq=1000 samplefreq=100 burnin=200

mcmc
sump
sumt

```

```{r Load ARF phylogeny and data}
tree <- treeio::read.beast(file = "trees/ARFs.nex.con.tre")

ARF_data <- read.csv(system.file("shiny-app", "gene_family_data",
                      "ARF_gene_ids.csv", package = "r1001genomes"))

#tip_labels need to be in the first column
ARF_data %<>% dplyr::select(name, everything())
ARF_data$tair_locus %<>% as.character()

ARF_data <- ARF_data[match(gsub(x = tree@phylo$tip.label, pattern = "\\.1",replacement = ""), ARF_data$tair_locus), ]

tree@phylo$tip.label <- ARF_data$name

```

(ref:A-ARF-Alignment) **Canonical ARFs have conserved DNA binding and PB1 domains** Protein sequences were aligned [@wright_decipher_2015] and variants were mapped to this alignment and colored according to the predicted functional effect of the most deleterious variant at that position, with light colors being less deleterious and darker colors being more deleterious. Red indicates nonsynonymous variants. Color scale is explained in {#methods}. 
```{r A-ARF-Alignment, fig.cap="(ref:A-ARF-Alignment)"}
A_ARFs <- subset(gene_info, tair_symbol %in% c("ARF5", "ARF6", "ARF7", "ARF8", "ARF19"))
A_ARFs <- alignCDS(A_ARFs$transcript_ID)
# make an alignment data frame
A_ARFs <- makeAlnDF(A_ARFs[[2]])

A_ARFs <- left_join(A_ARFs, dplyr::select(gene_info, "tair_locus",
                                       "tair_symbol", "transcript_ID"),
                        by = c("transcript_ID" = "transcript_ID"))
#A_ARFs$tair_symbol %<>% factor(levels = ARF_order, 
#                                       ordered = TRUE)
A_ARFs$tair_symbol %<>% fct_rev()

A_ARFs <- addSNPsToAlnDF(A_ARFs, VCF, effect_order = SNPeff_order)
A_ARFs$gap <- A_ARFs$seq_pos == "-"

p <-ggplot(A_ARFs, aes(x = aln_pos, y = tair_symbol)) +
  geom_raster(data = A_ARFs, 
             # mapping = aes(fill = effects)) +
              mapping = aes(fill = strength,
                            alpha = gap)) +
  scale_alpha_manual(values = c("TRUE" = 0, "FALSE" = 1), 
                     labels = c("aligned", "gap"), 
                     guide = guide_legend(override.aes = 
                                            list(fill = c("grey85", "white")),
                       title = "alignment",
                       direction = "vertical")) + 
   scale_fill_viridis(breaks = range(SNPeff_order$strength),
                      limits = range(SNPeff_order$strength),
                      na.value = "grey85",
                      direction = -1,
                      labels = c("neutral", "deleterious"),
                      option = "A", 
                      guide = guide_colorbar(ticks = FALSE, 
                                             title.vjust = .8, 
                                             title.position = "top", 
                                             title.hjust = 0.5)) +
  scale_y_discrete(expand = c(0,0)) +
  scale_x_continuous(expand = c(0,0)) +  # expand increases distance from axis
  labs(x = "codon position in alignment", y = "", fill = "variant effect") +
  theme_logo(base_family = "Helvetica") +
  theme(panel.grid = element_blank(), 
        panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        strip.text.x = element_blank(),
        #panel.background = element_rect(fill = "grey85", color = "white"),
        legend.position = c("bottom"),
        legend.key = element_rect(linetype = "solid"),
        legend.spacing = unit(4, "char"), 
        legend.box.spacing = unit(0.1,"line")) 
p
#ggsave(p,filename = "A_ARFs_alignment_variants.pdf", width = 12, height = 4)
```

Natural variation also provides a means to study how gene families are evolving. Mapping diversity at nonsynonymous variant sites relative to synonymous sites to gene trees allows comparison of recent functional divergence to the sequence divergence throughout the growth of the gene family. This visualization facilitates hypothesis generation regarding the functional redundancy or novelty of recent gene duplications.

Previous research has found evidence of both broad genetic redundancy and specificity within closely related pairs or groups of IAA proteins [@overvoorde_functional_2005]. For example, the *iaa8-1 iaa9-1* double mutant and the *iaa5-1 iaa6-1 iaa19-1* triple mutant have wild-type phenotypes [@overvoorde_functional_2005], yet the *IAA6/19* sister pair has significant differences in expression patterns, protein abundances and functions suggesting they have undergone functional specialization since their divergence [@winkler_variation_2017].

In our analysis of the *IAA* gene family, we frequently observed that one member of most *IAA* sister pairs had high nonsynonymous diversity while the other sister was more conserved \@ref(fig:IAAtree). In an interspecific comparison between *A. thaliana* and *A. lyrata*, *IAA19* is more conserved, while *IAA6* has regions of increased sequence divergence upstream of the degron and in the N-terminal half of the PB1 domain [@winkler_variation_2017]. Consistent with these results, our analysis of intraspecific natural variation revealed higher conservation for *IAA19* (${\pi_N/\pi_S}$ = 0.55) compared to *IAA6* (${\pi_N/\pi_S}$ = 2.3) \@ref(fig:IAAtree), and also detected high diversity within the same regions of *IAA6* as seen in Winkler *et al*.

(ref:IAAtree) **IAA protein sequence tree mapped with ${\pi_N/\pi_S}$.** Protein sequences were aligned [@wright_decipher_2015] and low information content regions were masked [@kuck_parametric_2010] prior to infering a phylogeny [@ronquist_mrbayes_2003]. Tips of the tree are mapped with circles of diameter proportional to ${\pi_N/\pi_S}$ and also are colored according to ${\pi_N/\pi_S}$. Nodes are labeled with the poster probability of monophyly.

```{r Load IAA phylogeny and data 2}
tree <- treeio::read.beast(file = "trees/IAAs.nex.con.tre")

IAA_data <- read.csv("tables/IAAs_table.csv")
#tip_labels need to be in the first column
IAA_data %<>% dplyr::select(name, everything())
IAA_data$name %<>% as.character()
IAA_data$AGI %<>% as.character()

IAA_data <- IAA_data[match(gsub(x = tree@phylo$tip.label, pattern = "\\.1",replacement = ""), IAA_data$AGI), ] 
tree@phylo$tip.label <- IAA_data$name
```
```{r IAAtree, fig.height=5, fig.width=5, fig.cap="(ref:IAAtree)"}
p <- ggtree(tree, layout = "rectangular", color = "grey") %<+% IAA_data + theme_tree() + scale_x_continuous(expand = c(0, 1.2)) + 
  geom_tippoint(mapping = aes(size = 4*PiN_PiS, fill = PiN_PiS), shape = 21, color = "white", show.legend = TRUE) + 
  scale_fill_viridis(values = rescale(c(min(IAA_data$PiN_PiS), 1,
                                        max(IAA_data$PiN_PiS)), 
                                      to = c(0,1)), guide = "colourbar") + scale_size(guide = "none") + 
  theme(legend.position = c(.9,.7)) + 
  geom_nodelab(mapping = aes(label = round(as.numeric(prob),digits = 2)), nudge_x = -.3, nudge_y = .5, size = 3) + 
  geom_tiplab(mapping = aes(color = Class), hjust = -0.25) + 
  scale_color_viridis(discrete = TRUE)
p
```

Within the Group B *IAA* genes, there are four sister pairs---*IAA10*/*IAA11*, *IAA12*/*IAA13*, *IAA18*/*IAA26*, and *IAA20*/*IAA30* \@ref(fig:IAAtree) [@remington_contrasting_2004]. Interestingly, among these sister pairs only *IAA12* and *IAA13* have a noticeable difference in nonsynonymous diversity. The *Arabidopsis thaliana* EFP browser shows that *IAA10* and *IAA11* have almost identical expression patterns, with the exception of *IAA10* being more highly expressed in the developing seed, and *IAA11* being more highly expressed in root tissue during lateral root development **need to cite original data source here**. In our analysis, *IAA10* and *IAA11* both showed functional conservation (${\pi_N/\pi_S}$ of 0.80 and 0.67 respectively) in support of *IAA10* and *IAA11* playing redundant roles. Accessions and associated variants identified here provide a potentially expedited route to elucidating the extent of redundancy between family members such as *IAA10* and *IAA11*.

# Conclusion

### 2018-11-12 manuscript/03-methods.Rmd
clean up and rewrite of results section

--- 2018-09-09 manuscript/02-methods.Rmd
@@
 ### Variant Data
-Variant data was queried from the 1001 genomes project website (http://1001genomes.org) via URL requests to an (undocumented) API service that provides subset files of the whole-genome VCF.
+Variant data was queried from the 1001 genomes project (http://1001genomes.org) via URL requests to their API service (http://tools.1001genomes.org/api/index.html).
+These queries provided subsets of the whole-genome VCF.
 The whole-genome VCF file can be found on the project’s website at http://1001genomes.org/data/GMI-MPI/releases/v3.1/

### 2018-11-12 manuscript/Old sections/IAAs.Rmd
clean up and rewrite of results section

--- 2018-09-09 manuscript/04-IAAs.Rmd
@@
   theme(legend.position = c(.9,.7)) + 
-  geom_nodelab(mapping = aes(label = round(prob,digits = 2)), nudge_x = -.3, nudge_y = .5, size = 3) + 
+  geom_nodelab(mapping = aes(label = round(as.numeric(prob),digits = 2)), nudge_x = -.3, nudge_y = .5, size = 3) + 
   geom_tiplab(mapping = aes(color = Class), hjust = -0.25) + 
@@
 ```{r A-IAA-alignment, fig.cap="(ref:A-IAA-alignment)"}
-A_IAAs <- subset(gene_info, tair_symbol %in% IAA_data[which(IAA_data$class == "A"), "name"])
-A_ARFs <- alignCDS(A_IAAs$transcript_ID)
+A_IAAs <- subset(gene_info, tair_symbol %in% IAA_data[which(IAA_data$Class == "A"), "name"])
+A_IAAs <- alignCDS(A_IAAs$transcript_ID)
 # make an alignment data frame
@@
     {as.integer(str_remove(A_IAAs$tair_symbol, 
-                           pattern = "ARF"))}, 
+                           pattern = "IAA"))}, 
     .desc = TRUE)
-A_IAAs <- addSNPsToAlnDF(A_IAAs, VCF)
+VCF <- ldply(.data = VCF, .fun = subset, gt_GT != "0|0")
+coding_VCF <- getCodingDiv(VCF)
+A_IAAs <- addSNPsToAlnDF(aln_df = A_IAAs,SNPs = coding_VCF, effect_order = SNPeff_order)
+A_IAAs$gap <- A_IAAs$seq_pos == "-"
 
-p <-ggplot(A_IAAs, aes(x = aln_pos, y = tair_symbol,
-                             group = seq_pos, text = variants)) +
-  geom_raster(data = A_IAAs[A_IAAs$seq_pos == "-",], fill = "lightgrey") +
-  geom_raster(data = na.omit(A_IAAs), 
+p <-ggplot(A_IAAs, aes(x = aln_pos, y = tair_symbol)) +
+  geom_raster(data = A_IAAs, 
              # mapping = aes(fill = effects)) +
-              mapping = aes(fill = as.integer(as.factor(effects)))) +
-  #scale_fill_viridis(option = "A", discrete = TRUE) +
-   scale_fill_viridis(breaks = range(
-     as.integer(as.factor(na.omit(A_IAAs$effects)))),
-                      labels = c("deleterious", "neutral"),
+              mapping = aes(fill = strength,
+                            alpha = gap)) +
+  scale_alpha_manual(values = c("TRUE" = 0, "FALSE" = 1), 
+                     labels = c("aligned", "gap"), 
+                     guide = guide_legend(override.aes = 
+                                            list(fill = c("grey85", "white")),
+                       title = "alignment",
+                       direction = "vertical")) + 
+   scale_fill_viridis(breaks = range(SNPeff_order$strength),
+                      limits = range(SNPeff_order$strength),
+                      na.value = "grey85",
+                      direction = -1,
+                      labels = c("neutral", "deleterious"),
                       option = "A", 
                       guide = guide_colorbar(ticks = FALSE, 
-                                             title.vjust = .8)) +
-  scale_y_discrete() +
-  # expand increases distance from axis
+                                             title.vjust = .8, 
+                                             title.position = "top", 
+                                             title.hjust = 0.5)) +
+  scale_y_discrete(expand = c(0,0)) +
+  scale_x_continuous(expand = c(0,0)) +  # expand increases distance from axis
   labs(x = "codon position in alignment", y = "", fill = "variant effect") +
   theme_logo(base_family = "Helvetica") +
-  theme(panel.grid = element_blank(), panel.grid.minor = element_blank()) +
-  theme(strip.background = element_blank(),
+  theme(panel.grid = element_blank(), 
+        panel.grid.minor = element_blank(),
+        strip.background = element_blank(),
         strip.text.x = element_blank(),
-        legend.position = "bottom")  
+        #panel.background = element_rect(fill = "grey85", color = "white"),
+        legend.position = c("bottom"),
+        legend.key = element_rect(linetype = "solid"),
+        legend.spacing = unit(4, "char"), 
+        legend.box.spacing = unit(0.1,"line")) 
 p
+
+ggsave("A_IAAs.png", width = 7, height = 4)
 ```

### 2018-11-12 manuscript/Old sections/discussion.Rmd
clean up and rewrite of results section

--- 2018-07-13 manuscript/07-discussion.Rmd
@@
 
-Our analysis of the natural diversity in the gene families that make up the auxin nuclear signaling pathway has confirmed existing knowledge, supported existing hypothses and generated new hypotheses.
-These results have been generated by a team of undergraduate and graduate students, postdoctoral researchers, and principal investigators as an alpha-test of the web application developed here.
+Our analysis of the natural diversity in the gene families that make up the auxin nuclear signaling pathway has confirmed existing knowledge, supported existing hypotheses and generated new hypotheses.
+These results have been generated by a team of undergraduate and graduate students, postdoctoral researchers, and principal investigators as an alpha-test of the ViVa web application developed here.
 Our analysis has confirmed the importance of auxin-induced degradation of the Aux/IAA proteins as evidenced by the strong conservation of the *AFB* genes and the canonical degron region of the *Aux/IAA* genes.

### 2018-11-12 manuscript/index.Rmd
clean up and rewrite of results section

--- 2018-09-09 manuscript/index.Rmd
@@
 ---
-title: Predicting structure--function relationships using natural genetic variation 
+title: Accelerating structure–function mapping using the ViVa webtool to mine natural variation 
 author:  |
@@
 
-abstract: |
-  The way genes and genomes have evolved can provide information about which genes are critical to an organisms fitness, which genes faciliate adaptation, and how robust genes and gene networks are to perturbation.
-Thousands of genomes have now been sequenced representing a large swath of plant species and genetic variation within many species which has only begun to be explored in the last few years.
-However, much of this data remains inaccessible to researchers without significant bioinformatics experience.
-Here, we have developed a webtool ViVa (Visualization of Variation) to facilitate analysis, by any researcher, of The 1001 *Arabidopsis* Genomes Project dataset on a gene, gene family, or gene network level.
-We We have demonstrated the utility and accessibility of this tool by analyzing the natural variation in the well-studied nuclear auxin signaling pathway as a group of researchers from an array of experience levels.
-Our analysis has provided further confirmation of existing knowledge of this well studied pathway and has also helped generate new hypotheses.
-These results suggest there is much to be learned about less studied gene families and networks from similar analysis, especially when paired with biochemical and genetic characterization.
-This application provides access for every researcher allowing the community to extract maximum value from this large and expensive-to-generate dataset.
-ViVa is also extensible to any other database of genetic variation such as The 3,000 Rice Genomes Project.
+abstract: Thousands of sequenced genomes are now publically available and represent a large swath of plant evolution; yet, much of this data remains inaccessible to researchers without significant bioinformatics experience.
+Here, we present a webtool called ViVa (Visualization of Variation) which aims to empower any researcher to take advantage of the amazing genetic resource collected in the Arabidopsis 1001 Genomes Project.
+ViVa facilates data mining on the gene, gene family or gene network level.
+To test the utility and accessibility of ViVa, we assembled a team with a range of expertise with biology and bioinformatics to analyze the natural variation in the well-studied nuclear auxin signaling pathway.
+Our analysis has provided further confirmation of existing knowledge and has also helped generate new hypotheses regarding this well studied pathway.
+These results highlight how natural variation can be used to generate and test hypotheses about less studied gene families and networks, especially when paired with biochemical and genetic characterization.
+ViVa is also readily extensible to newly emerging databases of genetic variation, such as the 3,000 Rice Genomes Project.
 
@@
 library(dplyr)
+library(stringr)
 
@@
 * A comprehensive analysis of natural sequence variation within the <i>Arabidopsis thaliana</i> nuclear auxin signaling pathway 
-* **Add your suggestions**
+* Visualizing natural variation with ViVa, a new webtool
+* Accelerating structure–function mapping using a new webtool to mine natural variation
 

### 2018-11-15 manuscript/02-results.Rmd
update viva use outline

@@
 ## An introduction to ViVa
+
+The current version of ViVa is specifically focused on visualizes natural variation in the coding sequences of genes.
+Future versions will expand upon this to examine noncoding sequences; however, indels are much more common in noncoding sequences, which limits the certainty of this resequencing data.
+
 **Morgan will provide screenshots**
+### Data input: Select the desired genes
+
+### SNPstats: Summary of gene information, structure, and diversity
+
+### Diversity plot: Visualize allelic diversity across the coding sequence
+
+### Accession mapping
+
+### Alignment
+
+### Gene tree
+
+### SNP browser
 
@@
 Mutations which prevent this domain from being recognized and marked for proteasomal degradation by SCF^TIR1/AFB^ are dominant and often result in obvious phenotypes [@berleth_auxin_2004; @yang_iaa1_2004].
-As expected this critical functional region is highly conserved at the amino acid level \@ref(fig:A-IAA-alignment).
+As expected this critical functional region is highly conserved at the amino acid level (\@ref(fig:A-IAA-alignment)).
 Similarly, the EAR motif which facilitates Aux/IAA interaction with TPL/TPR transcriptional repressors, are also conserved.
@@
 
-(ref:A-IAA-alignment) **Canonical IAAs have conserved Degron and EAR motifs** Protein sequences were aligned [@wright_decipher_2015] and variants were mapped to this alignment and colored according to the predicted functional effect of the most deleterious variant at that position, with light colors being less deleterious and darker colors being more deleterious.
-Red indicates nonsynonymous variants.
-Color scale is explained in {#methods}.
+(ref:A-IAA-alignment) **Canonical IAAs have conserved Degron and EAR motifs.** Protein sequences were aligned [@wright_decipher_2015] and variants were mapped to this alignment and colored according to the predicted functional effect of the most deleterious variant at that position, with light colors being less deleterious and darker colors being more deleterious.
+Red indicates missense variants.
+Color scale is explained in [Methods](#methods).
 EAR motif is highlighted by a black box including alignment positions 70-74.
@@
 Thus in the ARF transcription factor proteins we expect the DNA-binding function to be conserved.
-Indead, the DNA-binding domains of the ARFs had very few, low-diversity missense polymorphisms, as did the PB1 domain which facilitates ARF-Aux/IAA interaction \@ref(fig:A-ARF-Alignment).
+Indead, the DNA-binding domains of the ARFs had very few, low-diversity missense polymorphisms, as did the PB1 domain which facilitates ARF-Aux/IAA interaction (Figure \@ref(fig:A-ARF-Alignment)).
 
@@
 
-(ref:A-ARF-Alignment) **Canonical ARFs have conserved DNA binding and PB1 domains** Protein sequences were aligned [@wright_decipher_2015] and variants were mapped to this alignment and colored according to the predicted functional effect of the most deleterious variant at that position, with light colors being less deleterious and darker colors being more deleterious.
-Red indicates nonsynonymous variants.
-Color scale is explained in {#methods}.
+(ref:A-ARF-Alignment) **Canonical ARFs have conserved DNA binding and PB1 domains.** Protein sequences were aligned [@wright_decipher_2015] and variants were mapped to this alignment and colored according to the predicted functional effect of the most deleterious variant at that position, with light colors being less deleterious and darker colors being more deleterious.
+Red indicates missense variants.
+Color scale is explained in [Methods](#methods).
 ```{r A-ARF-Alignment, fig.cap="(ref:A-ARF-Alignment)"}

### 2018-11-15 manuscript/03-methods.Rmd
update viva use outline

@@
 
-where $N_{ik}$ is the nucleotide (A, T, C or G) at position k on the ith sequence of the population.
+where $N_{ik}$ is the nucleotide (A, T, C or G) at position k on the *i*th sequence of the population.
 $L$ is the length of the sequence (the same for all sequences in the population, 
@@
 The function `Nucleotide_diversity` in the r1001genomes package calculates $\pi_k$ for each position in the gene or region that contains a variant.
-note, $\pi_k$ is equal to 0 at all locations without variants.
+Note, $\pi_k$ is equal to 0 at all locations without variants.
 This is also what is displayed in the Diversity Plot tab of the web app.
 
-## Detailed $\pi_k$ calculaiton simplification.
+## Detailed $\pi_k$ calculation simplification.
 
@@
 \begin{equation}
-\displaystyle \pi_k = \sum_{ij} \frac{n_i}{n}\frac{n_j}{n} \pi_{ij} = 
-\frac{1}{n^2}\sum_{ij} n_i n_j \pi_{ij}\\
+\displaystyle \pi_k = \sum_{ij} \frac{n_i}{n}\frac{n_j}{n} \pi_{ij} = \frac{1}{n^2}\sum_{ij} n_i n_j \pi_{ij}
+\qquad
 \pi_{ij(k)} = \left\{\begin{array}{l}

### 2018-11-15 manuscript/index.Rmd
update viva use outline

@@
 abstract: Thousands of sequenced genomes are now publically available and represent a large swath of plant evolution; yet, much of this data remains inaccessible to researchers without significant bioinformatics experience.
-Here, we present a webtool called ViVa (Visualization of Variation) which aims to empower any researcher to take advantage of the amazing genetic resource collected in the Arabidopsis 1001 Genomes Project.
+Here, we present a webtool called ViVa (Visualizing Variation) which aims to empower any researcher to take advantage of the amazing genetic resource collected in the Arabidopsis 1001 Genomes Project.
 ViVa facilates data mining on the gene, gene family or gene network level.

### 2018-11-15 manuscript/02-results.Rmd
Filled in sections on ViVa app functionality.

@@
 **Morgan will provide screenshots**
+
 ### Data input: Select the desired genes
+At the top of the app, below the title but above the tab selection bar are two collapsible panels used for inputting the genes to query and additional annotation for those genes.
+The first collapsible panel is used to input the desired genes to investigate, either by typing in the TAIR gene IDs, or uploading a file.
+The second collapsible panel is optionally used to upload annotation files that will be plotted on some of the tabs of the apps second/analysis section.
 
 ### SNPstats: Summary of gene information, structure, and diversity
+Below the data input section, the rest of the app is divided into several tabs containing the interactive output of the program.
+
+The first tab is the SNP stats tab.
+This tab provides general information on the gene transcripts being queried, as well as some calculated counts/statistics on the content of variants found in the sample population.
+The first table of this section is basic information of the transcripts, including TAIR locus and symbol, the chromosomal start and end position of the transcript and the transcript length.
+This information is collected from the Araport 11 database stored in a local GFF file.
+
+The next two tables provide counts of occurrence of different effect types within the VCF dataset for each transcript.
+These counts are based on the “Effect” subfield of the “EFF” field in the VCF file.
+One table provides the total variant counts of each variant type from the reference genome (Col-0) in each accession.
+The unique allele count table only counts the number of unique variants/alleles, so if multiple accessions have the same variant allele relative the reference, that allele will only be counted once in the unique counts table.
+
+The final table on the SNP Stats tab provides Nei’s diversity statistic for the transcript and the coding region of each gene, as well as the diversity calculated for only synonymous and non-synonymous sites, and the ratio of the non-synonymous to synonymous diversity.
 
 ### Diversity plot: Visualize allelic diversity across the coding sequence
+The Diversity Plot tab shows Nei’s nucleotide diversity by nucleotide position of the protein coding region of a selected gene.
+Although the X-axis is marked by codon number from the N-terminus for interpretability, the diversity values are based on single nucleotide sites.
+The colors of markers on the plot identify the effect classes, based on fields from the VCF.
+If annotation files are provided, the background of the plot will be color coded by the annotated regions.
+If points on the plot are selected by clicking and dragging a box over them, the data for the selected points will appear in the grey box below the plot.
+Below this is a complete data table with the data provided for all points on the plot which can be downloaded as a .csv file.
 
 ### Accession mapping
+The SNP Mapping tab plots the location of accessions on a world map, and colors the points based on selected variants.
+
+After selecting the genes and filtering on the SNP type and level of nucleotide diversity, a checkbox group becomes available to select variant alleles to display on the map.
+The variant alleles are labeled with the Transcript_ID and Amino_Acid_Change fields, in the form [Transcript_ID|Amino_Acid_Change].
+
+After selecting the variant alleles and updating the map, the accessions on the map will be grouped and colored by each present unique combination of the selected alleles.
+Below the map is a table containing the accession details for all mapped accessions.
 
 ### Alignment
+The Alignment tab provides DNA and amino acid sequence alignments of selected genes, with loci colored by variants present.
+The content of this tab is most useful if the selected genes are all family members or have significant sequence homology.
+If annotation files are uploaded, the background of the sequences will be colored by annotation region.
+Hovering the mouse over sequence elements with polymorphisms will provide additional details about the SNP present at that loci.
 
@@
 ### SNP browser
+The Browser tab provides a way to search and filter the contents of the VCF file by different fields.
+After selecting the transcripts to include, a number of filters can be applied to the dataset to match text values, or set min and max limits on the values of numeric fields.
+When these filters are applied, the table below will update to only contain rows meeting the criteria for all filters.
 

### 2018-11-21 manuscript/02-results.Rmd
Added the app figure to the results section. also added a sub-section "ViVa R Package"

@@
 
-## An introduction to ViVa
+## An overview of ViVa
 
@@
 
-**Morgan will provide screenshots**
+```{r}
+appFigCaption <- "(ref:app_figure_elements) \n **Key elements of the web application** (a) The first section contains two collapsible panels and is used to input information about the genes to be investigated.
+(b) The SNP Stats tab provides gene level counts and statistics on SNPs.
+(c) The Diversity Plot tab plots the nucleotide diversity of SNP sites along the length of the coding region of a selected gene.
+(d) The SNP Mapping tab plots accessions on a world map colored to indicate what SNPs they have.
+(e) The  Browser tab allows searching for and filtering variants by text and numeric fields.
+(f) The Alignments tab aligns DNA and RNA sequences of homologous genes, and colors sequence elements based on SNPs and annotations."
+```
+
+```{r, echo = FALSE, fig.cap = appFigCaption, out.width="100%"}
+knitr::include_graphics("app_figure/App_figure.png")
+```
 
@@
 
-### Accession mapping
+### Accession mapping: View distribution of SNPs across the globe
 The SNP Mapping tab plots the location of accessions on a world map, and colors the points based on selected variants.
@@
 
-### Alignment
+### SNP browser: Filter and Search for Variants
+The Browser tab provides a way to search and filter the contents of the VCF file by different fields.
+After selecting the transcripts to include, a number of filters can be applied to the dataset to match text values, or set min and max limits on the values of numeric fields.
+When these filters are applied, the table below will update to only contain rows meeting the criteria for all filters.
+
+### Alignment: Visualize SNPs on alignments of homologous genes
 The Alignment tab provides DNA and amino acid sequence alignments of selected genes, with loci colored by variants present.
@@
 
-### Gene tree
-
-### SNP browser
-The Browser tab provides a way to search and filter the contents of the VCF file by different fields.
-After selecting the transcripts to include, a number of filters can be applied to the dataset to match text values, or set min and max limits on the values of numeric fields.
-When these filters are applied, the table below will update to only contain rows meeting the criteria for all filters.
+### ViVa R package: Programmatic access to ViVa's functionality
+All of the functionality of the ViVa web application are implemented through function calls to the ViVa R package.
+In addition to being able to  generate all figures and tables the web application can, users of the R package gain direct access to the underlying data structures and greater controll over parameters when processing and visualizing that data.
+The ViVa R package can be found at https://github.com/wrightrc/r1001genomes, or installed by typing the line
+`devtools::install_github("wrightrc/r1001genomes")` into an R console.
 

### 2018-11-21 manuscript/01-intro.Rmd
updates from JLN 181118

@@
 
-The first genome sequence of *Arabidopsis thaliana* facilitated rapid advancement of plant biology [@thearabidopsisgenomeinitiative_analysis_2000].
-Since this initial genome, massive scaling of sequencing technology has allowed the survey the genomic variation in natural *Arabidopsis thaliana* populations [@nordborg_pattern_2005; @borevitz_genomewide_2007; @weigel_1001_2009].
+The sequencing of the first *Arabidopsis thaliana* genome ushered in a new era of tool development and systematic functional annotation of plant genes [@thearabidopsisgenomeinitiative_analysis_2000].
+Since that landmark effort, massive scaling of sequencing technology has allowed for the survey of genomic variation in natural *Arabidopsis thaliana* populations [@nordborg_pattern_2005; @borevitz_genomewide_2007; @weigel_1001_2009].
 This valuable population genetics resource has led to several associations of genetic loci with phenotypic traits and provided insights into how selective pressure has influenced the evolution of plant genomes [@long_massive_2013; @atwell_genomewide_2010; @clark_common_2007].
-Despite the high value of this compendium of genetic variation, its use has been largely limited to investigations of population genetics and for genome-wide association studies.
 
-Natural genetic variation provides a catalog of permissible polymorphisms which may facilitate exploration and prediction of genotype/function/phenotype relationships at the gene, gene family and network scales [@joly-lopez_developing_2016].
-Outside of large effect variants, these relationships are difficult to map.
-Such studies in plants have the added obstacles of large gene families and the still quite time and resource-intensive process of allele replacement.
+Beyond its utility in gene discovery and understanding genome evolution, natural genetic variation provides a catalog of permissible polymorphisms that can facilitate the connection  of genotype to phenotype at the gene, gene family and network scales [@joly-lopez_developing_2016].
+This is an espcecially critica resource in large gene families where loss of function in individual genes may have little or no phenotypcial effect, and directed allele replacement remains time and resource-intensive.
 Massively parallel assays of variant effects have revolutionized genetic diagnostics and personalized medicine [@starita_variant_2017; @gasperini_power_2016; @matreyek_platform_2017].
 We envision the use of plant natural variation datasets as a tool to similarly revolutionize breeding and genetic engineering of crop plants by rapidly advancing our understanding of genotype/function/phenotype relationships.
-Biochemical or synthetic assays of gene function and quantitative phenotyping of the germplasm of collections of natural variation can facilitate functional and phenotypic quantification [@wright_insights_2017].
-However, the high dimensionality of natural variation datasets often limits access to those with bioinformatics expertise.
+A proof-of-principle survey of a relatively small subset of natural variants paired with a synthetic assay of gene function successfully mapped critical functional domains of auxin receptors [@wright_insights_2017].
 
-Here, we present ViVa: a web application and R-package for Visualizing Variation, which allows plant molecular biologists of any level access to the 1001 genomes database.
-ViVa facilitates formulation of hypotheses as to the genotype/function/phenotype relationships within any gene, family, or network.
-Using ViVa researchers may: 1.
-Identify polymorphisms to facilitate biochemical assays of variant effects [@starita_variant_2017; @wright_insights_2017]; 2.
-Produce family-wise alignments of variants to facilitate de novo functional domain identification [@melamed_combining_2015]; 3.
-Generate lists of accessions containing polymorphisms to facilitate phenotypic analysis of gene variant effects [@park_distributions_2017]; and 4.
-Quantify metrics of genetic diversity to facilitate the study of gene, gene family and network evolution [@delker_natural_2010; @kliebenstein_role_2008].
+So why is the survey of natural variants not as routine as a BLAST search?
+One reason may be the current requirement for a fairly high level of bioinformatics expertise to extract the desired information from the whole genome resequencing datasets.
+To address this concern, we created ViVa: a web application and R-package for Visualizing Variation, which allows plant molecular biologists of any level access to gene level data from the 1001 genomes database.
+Using ViVa researchers may: 1) Identify polymorphisms to facilitate biochemical assays of variant effects [@starita_variant_2017; @wright_insights_2017]; 2) Produce family-wise alignments of variants to facilitate de novo functional domain identification [@melamed_combining_2015]; 3) Generate lists of accessions containing polymorphisms to facilitate phenotypic analysis of gene variant effects [@park_distributions_2017]; and 4) Quantify metrics of genetic diversity to facilitate the study of gene, gene family and network evolution [@delker_natural_2010; @kliebenstein_role_2008].

### 2018-11-21 manuscript/02-results.Rmd
updates from JLN 181118

--- 2018-11-15 manuscript/02-results.Rmd
@@
 
-The current version of ViVa is specifically focused on visualizes natural variation in the coding sequences of genes.
-Future versions will expand upon this to examine noncoding sequences; however, indels are much more common in noncoding sequences, which limits the certainty of this resequencing data.
+ViVa is meant to visualize natural variation in the coding sequences of genes.
+For this first iteration, non-coding sequence variation is intentionally excluded from the analysis tools.
+This reflects challenges both in alignment of non-coding sequences and the increased difficulty in assessing variation in these regions [@alexandre_complex_2018].
 
 **Morgan will provide screenshots**
+
 ### Data input: Select the desired genes
@@
 
-To test the useability and accessibility of this tool we assembled a group of alpha testers comprising postdoctoral, graduate, and undergraduate researchers.
-This team of researchers analyzed natural variation in the nuclear auxin signaling pathway.
-This signaling pathway has been well studied, revealing abundant functional knowledge and crystal structures of several protein domains, providing several points of ground truth reference for comparing structure-function variation and natural genetic variation.
+To test the useability and accessibility of ViVa, we assembled a group of alpha testers comprising postdoctoral, graduate, and undergraduate researchers at a research university (University of Washington) and at a primarily-undergraduate institutions (Whitman College).
+Our testers focused their investigation of natural variation on the nuclear auxin signaling pathway.
+We selected this signaling pathway for multiple reasons including a wealth of functional data and solved structures of several domains or entire proteins.
+In this way, we were able to assess the predictive ability of the ViVa modules.
+Summary information about each family examined can be found in the supplement.
+Below, we describe the results for the Aux/IAA family in more detail.
 
-Assuming there is selective pressure against mutations which alter a protein's function, nonsynonymous genetic variation in critical functional domains of a protein should be low.
-Therefore scanning gene coding sequences for regions of low nonsynonymous diversity, may predict functional domains.
-Within the auxin signaling pathway several domains have been identified as critical to transmitting auxin signals.
-
-Perhaps the most critical domain is the degron of the Aux/IAA transcriptional repressor proteins.
-Mutations which prevent this domain from being recognized and marked for proteasomal degradation by SCF^TIR1/AFB^ are dominant and often result in obvious phenotypes [@berleth_auxin_2004; @yang_iaa1_2004].
-As expected this critical functional region is highly conserved at the amino acid level (\@ref(fig:A-IAA-alignment)).
-Similarly, the EAR motif which facilitates Aux/IAA interaction with TPL/TPR transcriptional repressors, are also conserved.
+In most cases, natural selection is expected to minimize the persistence of nonsynonymous mutations in sequences that encode critical functional domains  relative to their persistance in non-critical domains.
+Therefore, we reasoned that scanning gene coding sequences for regions of relatively low nonsynonymous diversity should highlight functional domains.
+This general principle can be seen clearly in the analysis of the Aux/IAA family of transcriptional co-repressors/co-receptors.
+Aux/IAAs have three major domains: domain I facilitates interaction with TPL/TPR transcriptional repressors, domain II facilitates interactions with the TIR1/AFB receptors in the presence of auxin, and the PB1 domain which failitates interactions with the ARF transcription factors.
+Domain 1 and Domain II can be readily identified (\@ref(fig:A-IAA-alignment)), perhaps because the multiple contributing residues are spread out in linear sequence space.
+It is worth noting that the most important residues within the PB1 domain also show very little variation.
 
@@
 EAR motif is highlighted by a black box including alignment positions 70-74.
-Degron domain is highlighted by a gray box including alignment positions 194-201.
+Degron domain is highlighted by a box including alignment positions 194-201.
+Critical residues for PB1 domain interactions are highlighted at alignment positions 246, 253, 316, and 320.
 ```{r A-IAA-alignment, fig.cap="(ref:A-IAA-alignment)"}
@@
         legend.box.spacing = unit(0.1,"line")) 
-p + geom_rect(mapping = aes(xmin = 193, xmax = 202, ymin = 0, ymax = length(levels((tair_symbol)))+1), color= "black", fill = NA) + geom_rect(mapping = aes(xmin = 69, xmax = 75, ymin = 0, ymax = length(levels((tair_symbol)))+1), color= "black", fill = NA)
-```
-
-The Aux/IAA proteins bind to ARF transcription factors allowing the Aux/IAA's to effect transcriptional repression at specific loci.
-Thus in the ARF transcription factor proteins we expect the DNA-binding function to be conserved.
-Indead, the DNA-binding domains of the ARFs had very few, low-diversity missense polymorphisms, as did the PB1 domain which facilitates ARF-Aux/IAA interaction (Figure \@ref(fig:A-ARF-Alignment)).
-
-```{r ARF diversity stats}
-gene_info <- geneInfoFromFile(
-  fname = system.file("shiny-app", "gene_family_data",
-                      "ARF_gene_ids.csv", package = "r1001genomes"))
-
-VCF <- readRDS(system.file("shiny-app", "gene_family_data",
-                      "ARF_VCF.rds", package = "r1001genomes"))
-
-VCF <- ldply(.data = VCF, .fun = subset, gt_GT != "0|0")
-VCF <- getCodingDiv(VCF)
-```
-
-```{r Mask_ARF_alignment, eval=FALSE}
-aliscore(as.AAbin(alignment[[2]]), path = "/Volumes/SSD/Applications/ALISCORE_v2.0/Aliscore_v.2.0/")
-id <- scan(file = "/Volumes/SSD/Applications/ALISCORE_v2.0/Aliscore_v.2.0/input.fas_List_random.txt", sep = " ", quiet = TRUE)
-AAAlign_no_gaps <- as.matrix(alignment[[2]])[, id]
-write.nexus.data(AAAlign_no_gaps, file = "ARFs.nex", format = "protein")
-```
-```{bash ARF_phylogeny_construction, eval=FALSE}
-/Applications/MrBayes/mb
-
-execute ARFs.nex
-# set ARF17 as the outgroup
-outgroup AT1G77850.1
-
-lset nucmodel=protein Rates=invgamma
-# or include Nst=6  ngammacat=4 omegavar=M3 
-prset aamodelpr=fixed(jones) statefreqpr=fixed(empirical) 
-# initial run with aamodelpr=mixed yielded posterior prob of 1 for jones model
-propset ExtTBR$prob=0
-
-mcmcp ngen=250000 nchains=4 printfreq=1000 samplefreq=100 burnin=200
-
-mcmc
-sump
-sumt
-
-```
-
-```{r Load ARF phylogeny and data}
-tree <- treeio::read.beast(file = "trees/ARFs.nex.con.tre")
-
-ARF_data <- read.csv(system.file("shiny-app", "gene_family_data",
-                      "ARF_gene_ids.csv", package = "r1001genomes"))
-
-#tip_labels need to be in the first column
-ARF_data %<>% dplyr::select(name, everything())
-ARF_data$tair_locus %<>% as.character()
-
-ARF_data <- ARF_data[match(gsub(x = tree@phylo$tip.label, pattern = "\\.1",replacement = ""), ARF_data$tair_locus), ]
-
-tree@phylo$tip.label <- ARF_data$name
-
-```
-
-(ref:A-ARF-Alignment) **Canonical ARFs have conserved DNA binding and PB1 domains.** Protein sequences were aligned [@wright_decipher_2015] and variants were mapped to this alignment and colored according to the predicted functional effect of the most deleterious variant at that position, with light colors being less deleterious and darker colors being more deleterious.
-Red indicates missense variants.
-Color scale is explained in [Methods](#methods).
-```{r A-ARF-Alignment, fig.cap="(ref:A-ARF-Alignment)"}
-A_ARFs <- subset(gene_info, tair_symbol %in% c("ARF5", "ARF6", "ARF7", "ARF8", "ARF19"))
-A_ARFs <- alignCDS(A_ARFs$transcript_ID)
-# make an alignment data frame
-A_ARFs <- makeAlnDF(A_ARFs[[2]])
-
-A_ARFs <- left_join(A_ARFs, dplyr::select(gene_info, "tair_locus",
-                                       "tair_symbol", "transcript_ID"),
-                        by = c("transcript_ID" = "transcript_ID"))
-#A_ARFs$tair_symbol %<>% factor(levels = ARF_order, 
-#                                       ordered = TRUE)
-A_ARFs$tair_symbol %<>% fct_rev()
-
-A_ARFs <- addSNPsToAlnDF(A_ARFs, VCF, effect_order = SNPeff_order)
-A_ARFs$gap <- A_ARFs$seq_pos == "-"
-
-p <-ggplot(A_ARFs, aes(x = aln_pos, y = tair_symbol)) +
-  geom_raster(data = A_ARFs, 
-             # mapping = aes(fill = effects)) +
-              mapping = aes(fill = strength,
-                            alpha = gap)) +
-  scale_alpha_manual(values = c("TRUE" = 0, "FALSE" = 1), 
-                     labels = c("aligned", "gap"), 
-                     guide = guide_legend(override.aes = 
-                                            list(fill = c("grey85", "white")),
-                       title = "alignment",
-                       direction = "vertical")) + 
-   scale_fill_viridis(breaks = range(SNPeff_order$strength),
-                      limits = range(SNPeff_order$strength),
-                      na.value = "grey85",
-                      direction = -1,
-                      labels = c("neutral", "deleterious"),
-                      option = "A", 
-                      guide = guide_colorbar(ticks = FALSE, 
-                                             title.vjust = .8, 
-                                             title.position = "top", 
-                                             title.hjust = 0.5)) +
-  scale_y_discrete(expand = c(0,0)) +
-  scale_x_continuous(expand = c(0,0)) +  # expand increases distance from axis
-  labs(x = "codon position in alignment", y = "", fill = "variant effect") +
-  theme_logo(base_family = "Helvetica") +
-  theme(panel.grid = element_blank(), 
-        panel.grid.minor = element_blank(),
-        strip.background = element_blank(),
-        strip.text.x = element_blank(),
-        #panel.background = element_rect(fill = "grey85", color = "white"),
-        legend.position = c("bottom"),
-        legend.key = element_rect(linetype = "solid"),
-        legend.spacing = unit(4, "char"), 
-        legend.box.spacing = unit(0.1,"line")) 
-p
-#ggsave(p,filename = "A_ARFs_alignment_variants.pdf", width = 12, height = 4)
+p + geom_rect(mapping = aes(xmin = 193, xmax = 202, ymin = 0, ymax = length(levels((tair_symbol)))+1), color= "black", fill = NA) + geom_rect(mapping = aes(xmin = 69, xmax = 75, ymin = 0, ymax = length(levels((tair_symbol)))+1), color= "black", fill = NA) + geom_rect(mapping = aes(xmin = 245, xmax = 247, ymin = 0, ymax = length(levels((tair_symbol)))+1), color = "black", fill = NA, size = .2) +
+geom_rect(mapping = aes(xmin = 319, xmax = 321, ymin = 0, ymax = length(levels((tair_symbol)))+1), color = "black", fill = NA, size = 0.2) +
+  geom_rect(mapping = aes(xmin = 252, xmax = 254, ymin = 0, ymax = length(levels((tair_symbol)))+1), color = "black", fill = NA, size = 0.2) +
+  geom_rect(mapping = aes(xmin = 315, xmax = 317, ymin = 0, ymax = length(levels((tair_symbol)))+1), color = "black", fill = NA, size = 0.2) 
 ```
@@
 Natural variation also provides a means to study how gene families are evolving.
-Mapping diversity at nonsynonymous variant sites relative to synonymous sites to gene trees allows comparison of recent functional divergence to the sequence divergence throughout the growth of the gene family.
-This visualization facilitates hypothesis generation regarding the functional redundancy or novelty of recent gene duplications.
+To do this, we used ViVa to map the diversity at nonsynonymous variant sites relative to synonymous sites to phylogenetic trees.
+This visualization makes it straightforward to compare rates of recent functional divergence within the context of rates of sequence divergence across the entire gene family.
+In this way, a researcher chould hone in on gene clades that are likely to exhibit high rates of functional redundancy or, conversely, gene clades where there is a greater likelihood that novel functions have emerged.
 
-Previous research has found evidence of both broad genetic redundancy and specificity within closely related pairs or groups of IAA proteins [@overvoorde_functional_2005].
+Previous research has found evidence of both broad genetic redundancy and specificity within closely related pairs or groups of IAA proteins [@overvoorde_functional_2005; @winkler_variation_2017].
 For example, the *iaa8-1 iaa9-1* double mutant and the *iaa5-1 iaa6-1 iaa19-1* triple mutant have wild-type phenotypes [@overvoorde_functional_2005], yet the *IAA6/19* sister pair has significant differences in expression patterns, protein abundances and functions suggesting they have undergone functional specialization since their divergence [@winkler_variation_2017].
-
-In our analysis of the *IAA* gene family, we frequently observed that one member of most *IAA* sister pairs had high nonsynonymous diversity while the other sister was more conserved \@ref(fig:IAAtree).
-In an interspecific comparison between *A.
+A closer examination of the *IAA19* and *IAA6* pair in *A.
 thaliana* and *A.
-lyrata*, *IAA19* is more conserved, while *IAA6* has regions of increased sequence divergence upstream of the degron and in the N-terminal half of the PB1 domain [@winkler_variation_2017].
-Consistent with these results, our analysis of intraspecific natural variation revealed higher conservation for *IAA19* (${\pi_N/\pi_S}$ = 0.55) compared to *IAA6* (${\pi_N/\pi_S}$ = 2.3) \@ref(fig:IAAtree), and also detected high diversity within the same regions of *IAA6* as seen in Winkler *et al*.
+lyrata* found that *IAA6* was diversifying far faster than [@winkler_variation_2017].
+Consistent with these results, ViVa revealed higher conservation for *IAA19* (${\pi_N/\pi_S}$ = 0.55) compared to *IAA6* (${\pi_N/\pi_S}$ = 2.3) (Figure \@ref(fig:IAAtree)), and also detected high diversity within the same regions of *IAA6* as seen in [@winkler_variation_2017].
+This pattern---one *Aux/IAA* sister showing high nonsynonymous diversity while the other sister was more conserved---was observed frequently across the family, suggesting this could be a recurring feature in *Aux/IAA* evolution supporting the large diversity in auxin functions.
 
@@
 
-Within the Group B *IAA* genes, there are four sister pairs---*IAA10*/*IAA11*, *IAA12*/*IAA13*, *IAA18*/*IAA26*, and *IAA20*/*IAA30* \@ref(fig:IAAtree) [@remington_contrasting_2004].
-Interestingly, among these sister pairs only *IAA12* and *IAA13* have a noticeable difference in nonsynonymous diversity.
-The *Arabidopsis thaliana* EFP browser shows that *IAA10* and *IAA11* have almost identical expression patterns, with the exception of *IAA10* being more highly expressed in the developing seed, and *IAA11* being more highly expressed in root tissue during lateral root development **need to cite original data source here**.
-In our analysis, *IAA10* and *IAA11* both showed functional conservation (${\pi_N/\pi_S}$ of 0.80 and 0.67 respectively) in support of *IAA10* and *IAA11* playing redundant roles.
-Accessions and associated variants identified here provide a potentially expedited route to elucidating the extent of redundancy between family members such as *IAA10* and *IAA11*.
+The notable exception to the patter of one conserved and one diersifying sister was observed within the Group B *IAA* genes (*IAA10*/*IAA11*, *IAA12*/*IAA13*, *IAA18*/*IAA26*, and *IAA20*/*IAA30*) [@remington_contrasting_2004].
+Among these sister pairs only *IAA12* and *IAA13* have a noticeable difference in nonsynonymous diversity.
+In our analysis, *IAA10* and *IAA11* both showed functional conservation (${\pi_N/\pi_S}$ of 0.80 and 0.67 respectively).
+In support of *IAA10* and *IAA11* playing largely redundant roles, the *Arabidopsis thaliana* EFP browser indicates that *IAA10* and *IAA11* have almost identical expression patterns.
 

### 2018-11-21 manuscript/index.Rmd
Added the app figure to the results section. also added a sub-section "ViVa R Package"
updates from JLN 181118

@@
 abstract: Thousands of sequenced genomes are now publically available and represent a large swath of plant evolution; yet, much of this data remains inaccessible to researchers without significant bioinformatics experience.
-Here, we present a webtool called ViVa (Visualizing Variation) which aims to empower any researcher to take advantage of the amazing genetic resource collected in the Arabidopsis 1001 Genomes Project.
+Here, we present a webtool called ViVa (Visualizing Variation) which aims to empower any researcher to take advantage of the amazing genetic resource collected in the Arabidopsis 1001 Genomes Project (http://1001genomes.org).
 ViVa facilates data mining on the gene, gene family or gene network level.
@@
 Our analysis has provided further confirmation of existing knowledge and has also helped generate new hypotheses regarding this well studied pathway.
-These results highlight how natural variation can be used to generate and test hypotheses about less studied gene families and networks, especially when paired with biochemical and genetic characterization.
-ViVa is also readily extensible to newly emerging databases of genetic variation, such as the 3,000 Rice Genomes Project.
+These results highlight how natural variation could be used to generate and test hypotheses about less studied gene families and networks, especially when paired with biochemical and genetic characterization.
+ViVa is also readily extensible to newly emerging databases of genetic variation, such as the 3,000 Rice Genomes Project (http://snp-seek.irri.org/).
 

### 2018-11-21 manuscript/02-results.Rmd
build after merge

@@
 
-**Morgan will provide screenshots**
+```{r}
+appFigCaption <- "(ref:app_figure_elements) \n **Key elements of the web application** (a) The first section contains two collapsible panels and is used to input information about the genes to be investigated.
+(b) The SNP Stats tab provides gene level counts and statistics on SNPs.
+(c) The Diversity Plot tab plots the nucleotide diversity of SNP sites along the length of the coding region of a selected gene.
+(d) The SNP Mapping tab plots accessions on a world map colored to indicate what SNPs they have.
+(e) The  Browser tab allows searching for and filtering variants by text and numeric fields.
+(f) The Alignments tab aligns DNA and RNA sequences of homologous genes, and colors sequence elements based on SNPs and annotations."
+```
+
+```{r, echo = FALSE, fig.cap = appFigCaption, out.width="100%"}
+knitr::include_graphics("app_figure/App_figure.png")
+```
 
 ### Data input: Select the desired genes
+At the top of the app, below the title but above the tab selection bar are two collapsible panels used for inputting the genes to query and additional annotation for those genes.
+The first collapsible panel is used to input the desired genes to investigate, either by typing in the TAIR gene IDs, or uploading a file.
+The second collapsible panel is optionally used to upload annotation files that will be plotted on some of the tabs of the apps second/analysis section.
 
 ### SNPstats: Summary of gene information, structure, and diversity
+Below the data input section, the rest of the app is divided into several tabs containing the interactive output of the program.
+
+The first tab is the SNP stats tab.
+This tab provides general information on the gene transcripts being queried, as well as some calculated counts/statistics on the content of variants found in the sample population.
+The first table of this section is basic information of the transcripts, including TAIR locus and symbol, the chromosomal start and end position of the transcript and the transcript length.
+This information is collected from the Araport 11 database stored in a local GFF file.
+
+The next two tables provide counts of occurrence of different effect types within the VCF dataset for each transcript.
+These counts are based on the “Effect” subfield of the “EFF” field in the VCF file.
+One table provides the total variant counts of each variant type from the reference genome (Col-0) in each accession.
+The unique allele count table only counts the number of unique variants/alleles, so if multiple accessions have the same variant allele relative the reference, that allele will only be counted once in the unique counts table.
+
+The final table on the SNP Stats tab provides Nei’s diversity statistic for the transcript and the coding region of each gene, as well as the diversity calculated for only synonymous and non-synonymous sites, and the ratio of the non-synonymous to synonymous diversity.
 
 ### Diversity plot: Visualize allelic diversity across the coding sequence
+The Diversity Plot tab shows Nei’s nucleotide diversity by nucleotide position of the protein coding region of a selected gene.
+Although the X-axis is marked by codon number from the N-terminus for interpretability, the diversity values are based on single nucleotide sites.
+The colors of markers on the plot identify the effect classes, based on fields from the VCF.
+If annotation files are provided, the background of the plot will be color coded by the annotated regions.
+If points on the plot are selected by clicking and dragging a box over them, the data for the selected points will appear in the grey box below the plot.
+Below this is a complete data table with the data provided for all points on the plot which can be downloaded as a .csv file.
 
-### Accession mapping
+### Accession mapping: View distribution of SNPs across the globe
+The SNP Mapping tab plots the location of accessions on a world map, and colors the points based on selected variants.
 
-### Alignment
+After selecting the genes and filtering on the SNP type and level of nucleotide diversity, a checkbox group becomes available to select variant alleles to display on the map.
+The variant alleles are labeled with the Transcript_ID and Amino_Acid_Change fields, in the form [Transcript_ID|Amino_Acid_Change].
 
-### Gene tree
+After selecting the variant alleles and updating the map, the accessions on the map will be grouped and colored by each present unique combination of the selected alleles.
+Below the map is a table containing the accession details for all mapped accessions.
 
-### SNP browser
+### SNP browser: Filter and Search for Variants
+The Browser tab provides a way to search and filter the contents of the VCF file by different fields.
+After selecting the transcripts to include, a number of filters can be applied to the dataset to match text values, or set min and max limits on the values of numeric fields.
+When these filters are applied, the table below will update to only contain rows meeting the criteria for all filters.
+
+### Alignment: Visualize SNPs on alignments of homologous genes
+The Alignment tab provides DNA and amino acid sequence alignments of selected genes, with loci colored by variants present.
+The content of this tab is most useful if the selected genes are all family members or have significant sequence homology.
+If annotation files are uploaded, the background of the sequences will be colored by annotation region.
+Hovering the mouse over sequence elements with polymorphisms will provide additional details about the SNP present at that loci.
+
+### ViVa R package: Programmatic access to ViVa's functionality
+All of the functionality of the ViVa web application are implemented through function calls to the ViVa R package.
+In addition to being able to  generate all figures and tables the web application can, users of the R package gain direct access to the underlying data structures and greater controll over parameters when processing and visualizing that data.
+The ViVa R package can be found at https://github.com/wrightrc/r1001genomes, or installed by typing the line
+`devtools::install_github("wrightrc/r1001genomes")` into an R console.
 

### 2018-12-06 manuscript/01-intro.Rmd
bioRxiv submission

@@
 Beyond its utility in gene discovery and understanding genome evolution, natural genetic variation provides a catalog of permissible polymorphisms that can facilitate the connection of genotype to phenotype at the gene, gene family and network scales [@joly-lopez_developing_2016].
-This is an espcecially critica resource in large gene families where loss of function in individual genes may have little or no phenotypcial effect, and directed allele replacement remains time and resource-intensive.
-Massively parallel assays of variant effects have revolutionized genetic diagnostics and personalized medicine [@starita_variant_2017; @gasperini_power_2016; @matreyek_platform_2017].
+This is an especially critical resource in large gene families where loss of function in individual genes may have little or no phenotypic effect and directed allele replacement remains time and resource-intensive.
+Massively parallel assays of variant effects in human clinical medicine stand to revolutionize genetic diagnostics and personalized medicine [@starita_variant_2017; @gasperini_power_2016; @matreyek_platform_2017].
 We envision the use of plant natural variation datasets as a tool to similarly revolutionize breeding and genetic engineering of crop plants by rapidly advancing our understanding of genotype/function/phenotype relationships.
-A proof-of-principle survey of a relatively small subset of natural variants paired with a synthetic assay of gene function successfully mapped critical functional domains of auxin receptors [@wright_insights_2017].
+A proof-of-principle survey of a relatively small subset of natural variants paired with a synthetic assay of gene function successfully mapped critical functional domains of auxin receptors and identified new alleles which affect plant phenotype [@wright_insights_2017].
 
-So why is the survey of natural variants not as routine as a BLAST search?
-One reason may be the current requirement for a fairly high level of bioinformatics expertise to extract the desired information from the whole genome resequencing datasets.
-To address this concern, we created ViVa: a web application and R-package for Visualizing Variation, which allows plant molecular biologists of any level access to gene level data from the 1001 genomes database.
-Using ViVa researchers may: 1) Identify polymorphisms to facilitate biochemical assays of variant effects [@starita_variant_2017; @wright_insights_2017]; 2) Produce family-wise alignments of variants to facilitate de novo functional domain identification [@melamed_combining_2015]; 3) Generate lists of accessions containing polymorphisms to facilitate phenotypic analysis of gene variant effects [@park_distributions_2017]; and 4) Quantify metrics of genetic diversity to facilitate the study of gene, gene family and network evolution [@delker_natural_2010; @kliebenstein_role_2008].
+Why is the survey of natural variants not as routine as a BLAST search or ordering T-DNA insertion mutants?
+One reason may be the current requirement for a fairly high level of bioinformatics expertise to extract the desired information from whole genome resequencing datasets.
+While existing resources such as the 1001 Proteomes website [@joshi_1001_2012] and ePlant [@waese_eplant_2017] facilitate access to this data at the gene scale, they do not provide summaries or visualizations of variation at the gene family and network scales.
+To address this concern, we created ViVa: a webtool and R-package for **Vi**sualizing **Va**riation, which allows plant molecular biologists of any level access to gene-level data from the 1001 Genomes database.
+Using ViVa researchers may: 1) Identify polymorphisms to facilitate biochemical assays of variant effects [@starita_variant_2017; @wright_insights_2017]; 2) Produce family-wise alignments of variants to facilitate *de novo* functional domain identification [@melamed_combining_2015]; 3) Generate lists of accessions containing polymorphisms to facilitate phenotypic analysis of gene variant effects [@park_distributions_2017]; and 4) Quantify metrics of genetic diversity to facilitate the study of gene, gene family and network evolution [@delker_natural_2010; @kliebenstein_role_2008].

### 2018-12-06 manuscript/02-results.Rmd
bioRxiv submission

--- 2018-11-21 manuscript/02-results.Rmd
@@
 
-## An introduction to ViVa
+## An overview of ViVa
 
-ViVa is meant to visualize natural variation in the coding sequences of genes.
-For this first iteration, non-coding sequence variation is intentionally excluded from the analysis tools.
+ViVa, in this first iteration, is meant to visualize natural variation in the coding sequences of genes.
+Non-coding sequence variation is intentionally excluded from the analysis tools.
 This reflects challenges both in alignment of non-coding sequences and the increased difficulty in assessing variation in these regions [@alexandre_complex_2018].
 
-**Morgan will provide screenshots**
+The development version of ViVa can be accessed as a Docker container https://hub.docker.com/r/wrightrc/r1001genomes/ or as an R-package at https://github.com/wrightrc/r1001genomes.
+ViVa will be hosted at https://www.plantsynbiolab.bse.vt.edu/ViVa/ upon release.
 
-### Data input: Select the desired genes
+(ref:app-figure-elements) **Key elements of the webtool** (a) The first section contains two collapsible panels, Gene Select and Annotation Files, which are used to input information about the genes to be investigated.
+(b) The SNP Stats tab provides gene-structure level counts and statistics on SNPs.
+(c) The Diversity Plot tab plots the nucleotide diversity of SNP sites along the length of the coding region of a selected gene.
+(d) The SNP Mapping tab plots accessions on a world map colored according to the selected set of SNPs.
+(e) The SNP Browser tab allows variants and accessions to be filtered by any combination of text and numeric fields.
+(f) The Alignments tab aligns DNA and amino acid sequences of homologous genes, and colors sequence elements based on SNPs and annotations.
+```{r app-figure-elements, fig.cap = "(ref:app-figure-elements)"}
+knitr::include_graphics("app_figure/App_figure.png")
+```
 
-### SNPstats: Summary of gene information, structure, and diversity
+### Gene Select and Annotation Files  
+At the top of the ViVa webtool are two collapsible panels used for entering the genes to query and custom annotations for those genes (Figure \@ref(fig:app-figure-elements)a).
+The Gene Select panel permits gene input by either typing in the genes' AGI/TAIR locus identifiers, or uploading a .csv file.
+The Annotation Files panel is optionally used to upload an annotation file containing coordinates of domains, mutations, or any other sequence knowledge that will be plotted on some of the tabs of the webtools analysis section.
+
+Below the data input section, the rest of the webtool is divided into several tabs containing the interactive output of the program.
+
+### SNP Stats: Summary of gene information, structure, and diversity
+
+The SNP Stats tab provides general information on the gene transcripts being queried, as well as calculated counts/statistics on the content of variants found in the sample population (Figure \@ref(fig:app-figure-elements)b).
+The first table of this section is basic information about the transcripts, including TAIR locus and symbol, the chromosomal start and end position and the transcript length.
+This information was collected from the Araport11 Official Release (06/2016) annotation dataset [@cheng_araport11_2017].
+
+The next two tables provide counts of SNPs across the gene body for each transcript.
+The Total Polymorphism Counts table provides the total number of observations of non-reference allele counts of each variant type (the Col-0 accession TAIR9 genome is the reference genome for this dataset).
+The Unique Allele Count table only counts the number of unique variants and alleles within the population of accessions (*e.g.* if multiple accessions have the same variant allele, that allele will only be counted once).
+
+The Nucleotide Diversity Statistics table provides a nucleotide diversity statistic (${\pi}$) for the transcript and the coding sequence of each gene [@nei_mathematical_1979].
+Nucleotide diversity is also calculated for the set of only synonymous (${\pi_S}$) and only non-synonymous sites (${\pi_N}$).
+The ratio of the presence of non-synonymous to synonymous polymorphism provides a measure of the potential for functional diversity [@firnberg_genetic_2013; @whitehead_optimization_2012].
+We present ${\pi_N/\pi_S}$ here as an correlate for functional diversity throughout ViVa [@nelson_snpgenie_2015; @hughes_adaptive_2000].
+While imperfect, this metric may be suggestive of functional constraint when ${\pi_N/\pi_S << 1}$ and functional diversity when ${\pi_N/\pi_S >> 1}$ [@hughes_adaptive_1999].
 
 ### Diversity plot: Visualize allelic diversity across the coding sequence
+The Diversity Plot tab shows the nucleotide diversity of each variant in the coding region of a selected gene (Figure \@ref(fig:app-figure-elements)c).
+Although the X-axis is marked by codon number from the N-terminus for interpretability, the diversity values are based on single nucleotide sites.
+The colors of markers on the plot identify the effect of the polymorphism.
+If annotation files are provided, the background of the plot will be color coded by the annotated regions.
+If points on the plot are selected by clicking and dragging a box over them, the data for the selected points will appear in the grey box below the plot.
+Below this is a complete data table containing all points on the plot which can be downloaded as a .csv file.
+This tab allows users to identify regions of high diversity as well as isolate polymorphisms that may affect gene function and exist in multiple accessions, facilitating phenotypic analysis.
 
-### Accession mapping
+### SNP Mapping: View distributions of SNPs across the globe
+The SNP Mapping tab plots the accessions collection locations on a world map, and colors the points based on selected variant alleles (Figure \@ref(fig:app-figure-elements)d).
+After selecting the genes and filtering on the SNP type and level of nucleotide diversity, a group of checkboxes becomes available to select variant alleles to display on the map.
+The variant alleles are labeled with the Transcript_ID and Amino_Acid_Change fields, in the form [Transcript_ID|Amino_Acid_Change].
+After selecting the variant alleles and updating the map, the accessions on the map will be colored by each present unique combination of the selected alleles.
+Below the map is a table containing the accession details for all mapped accessions.
+This tab may help users formulate hypotheses about the relatedness of accessions sharing a common allele and environments in which that allele may be favorable.
 
-### Alignment
+### SNP Browser: Filter and search for variants
+The SNP Browser tab provides a way to search and filter the variant data by different fields (Figure \@ref(fig:app-figure-elements)e).
+After selecting the transcripts to include, a number of filters can be applied to the dataset to match text values (e.g.
+gene name or variant effect), or set minimum and maximum limits on the values of numeric fields (e.g.
+nucleotide diversity).
+When these filters are applied, the table below will update to only contain rows meeting the criteria for all filters.
+This tab can be useful for identifying all accessions with a particular SNP allele, or any non-reference alleles in a particular region of a gene that may not have been easily accessible in another tab.
 
-### Gene tree
+### Alignments: Visualize SNPs on alignments of homologous genes
+The Alignments tab provides DNA and amino acid sequence alignments of selected genes, colored according to the variant allele with the strongest functional effect at each position (Figure \@ref(fig:app-figure-elements)f and \@ref(fig:A-IAA-alignment)).
+The content of this tab is most useful if the selected genes are all family members or have significant sequence homology.
+If annotation files are uploaded, the background of the sequences will be colored by annotation region.
+Hovering the cursor over variants will provide additional details about the alleles present at that locus.
+This tab facilitates family-wise analysis of functional conservation, allowing users to identify potential functional regions and alleles which may be useful in deciphering this function.
 
-### SNP browser
+### Gene Tree: Visualize functional diversity and sequence divergence of a gene family
+The Gene Tree tab provides a neighbor joining tree (or uploaded tree created by the user) for the selected genes with the tips of the tree mapped with predicted functional diversity as represented by ${\pi_N/\pi_S}$ in the 1001 Genomes dataset (Figure \@ref(fig:IAAtree)).
+This tab allows users to generate hypotheses regarding functional diversity and redundancy within the context of the predicted evolution of the gene family.
+
+### ViVa R package: Programmatic access to ViVa's functionality
+All of the functionalities of the ViVa webtool are implemented through function calls to the ViVa R package.
+In addition to being able to generate the same sets of figures and tables as in the webtool, users of the R package also gain direct access to the underlying data structures, providing greater control over parameters when processing and visualizing the data.
+The ViVa R package can be found at https://github.com/wrightrc/r1001genomes and can be installed in your R environment via the `devtools` package:
+`devtools::install_github("wrightrc/r1001genomes")`.
 
@@
 
-To test the useability and accessibility of ViVa, we assembled a group of alpha testers comprising postdoctoral, graduate, and undergraduate researchers at a research university (University of Washington) and at a primarily-undergraduate institutions (Whitman College).
+To test the usability and accessibility of ViVa, we assembled a group of alpha testers comprising postdoctoral, graduate, and undergraduate researchers at a research university (University of Washington) and at a primarily-undergraduate institution (Whitman College).
 Our testers focused their investigation of natural variation on the nuclear auxin signaling pathway.
 We selected this signaling pathway for multiple reasons including a wealth of functional data and solved structures of several domains or entire proteins.
-In this way, we were able to assess the predictive ability of the ViVa modules.
-Summary information about each family examined can be found in the supplement.
-Below, we describe the results for the Aux/IAA family in more detail.
+Using this exitsting knowledge, we were able to qualitatively assess the predictive ability of the ViVa modules.
+Summary information about each nuclear auxin signaling gene family examined can be found in the [Supplemental Data](#si).
+Below, we describe the results for the *Aux/IAA* family in more detail.
 
-In most cases, natural selection is expected to minimize the persistence of nonsynonymous mutations in sequences that encode critical functional domains  relative to their persistance in non-critical domains.
+In most cases, natural selection is expected to minimize the persistence of nonsynonymous mutations in sequences that encode critical functional domains relative to their persistence in non-critical domains [@hughes_adaptive_2000].
 Therefore, we reasoned that scanning gene coding sequences for regions of relatively low nonsynonymous diversity should highlight functional domains.
-This general principle can be seen clearly in the analysis of the Aux/IAA family of transcriptional co-repressors/co-receptors.
-Aux/IAAs have three major domains: domain I facilitates interaction with TPL/TPR transcriptional repressors, domain II facilitates interactions with the TIR1/AFB receptors in the presence of auxin, and the PB1 domain which failitates interactions with the ARF transcription factors.
-Domain 1 and Domain II can be readily identified (\@ref(fig:A-IAA-alignment)), perhaps because the multiple contributing residues are spread out in linear sequence space.
-It is worth noting that the most important residues within the PB1 domain also show very little variation.
+This general principle can be seen clearly in the analysis of the *Aux/IAA* family of transcriptional co-repressors/co-receptors.
+Aux/IAAs have three major domains.
+Domain I contains an EAR motif that facilitates interaction with TOPLESS (TPL) and TOPLESS-related (TPR) transcriptional repressors [@tiwari_aux_2004; @szemenyei_topless_2008].
+Domain II, the degron, facilitates interactions with the TIR1/AFB receptors in the presence of auxin [@tan_mechanism_2007].
+Domain III (which was originally considered domains III and IV) is a PB1 domain and facilitates interactions with the ARF transcription factors [@ulmasov_aux_1997; @guilfoyle_getting_2012; @nanao_structural_2014; @korasick_molecular_2014].
+The EAR motif and degron can be readily identified by the drop in nonsynonymous variation, as visualized by the lack of stronger functional effects in Figure \@ref(fig:A-IAA-alignment).
+The PB1 domain is not as readily identified, perhaps because the multiple contributing residues are spread out in linear sequence space.
+It is worth noting that the charged residues that facilitate electrostatic PB1-PB1 interaction show little variation.
 
@@
 
-(ref:A-IAA-alignment) **Canonical IAAs have conserved Degron and EAR motifs.** Protein sequences were aligned [@wright_decipher_2015] and variants were mapped to this alignment and colored according to the predicted functional effect of the most deleterious variant at that position, with light colors being less deleterious and darker colors being more deleterious.
+(ref:A-IAA-alignment) **Canonical Aux/IAAs have conserved Degron and EAR motifs.** Protein sequences were aligned with DECIPHER [@wright_decipher_2015] and variants were mapped to this alignment and colored according to the predicted functional effect of the allele of strongest effect at that position, with light colors having weaker effects on function and darker colors stronger effects.
 Red indicates missense variants.
 Color scale is explained in [Methods](#methods).
-EAR motif is highlighted by a black box including alignment positions 70-74.
-Degron domain is highlighted by a box including alignment positions 194-201.
-Critical residues for PB1 domain interactions are highlighted at alignment positions 246, 253, 316, and 320.
+
 ```{r A-IAA-alignment, fig.cap="(ref:A-IAA-alignment)"}
@@
                       direction = -1,
-                      labels = c("neutral", "deleterious"),
+                      labels = c("weak", "strong"),
                       option = "A", 
@@
                                              title.hjust = 0.5)) +
-  scale_y_discrete(expand = c(0,0)) +
+  scale_y_discrete(expand = expand_scale(add = c(0,2.5))) +
   scale_x_continuous(expand = c(0,0)) +  # expand increases distance from axis
-  labs(x = "codon position in alignment", y = "", fill = "variant effect") +
+  labs(x = "codon position in alignment", y = "", 
+       fill = "variant functional effect", title = "") +
   theme_logo(base_family = "Helvetica") +
@@
         legend.box.spacing = unit(0.1,"line")) 
-p + geom_rect(mapping = aes(xmin = 193, xmax = 202, ymin = 0, ymax = length(levels((tair_symbol)))+1), color= "black", fill = NA) + geom_rect(mapping = aes(xmin = 69, xmax = 75, ymin = 0, ymax = length(levels((tair_symbol)))+1), color= "black", fill = NA) + geom_rect(mapping = aes(xmin = 245, xmax = 247, ymin = 0, ymax = length(levels((tair_symbol)))+1), color = "black", fill = NA, size = .2) +
-geom_rect(mapping = aes(xmin = 319, xmax = 321, ymin = 0, ymax = length(levels((tair_symbol)))+1), color = "black", fill = NA, size = 0.2) +
-  geom_rect(mapping = aes(xmin = 252, xmax = 254, ymin = 0, ymax = length(levels((tair_symbol)))+1), color = "black", fill = NA, size = 0.2) +
-  geom_rect(mapping = aes(xmin = 315, xmax = 317, ymin = 0, ymax = length(levels((tair_symbol)))+1), color = "black", fill = NA, size = 0.2) 
+p + annotate("rect", xmin = 69, xmax = 75, ymin = 0.4, 
+             ymax = length(levels((A_IAAs$tair_symbol))) + 0.6, 
+             color= "black", fill = NA) + 
+  annotate("text", x = 72, y = length(levels((A_IAAs$tair_symbol))) + 1.2, 
+                          label = "EAR domain", family="Arial") + 
+  annotate("rect", xmin = 193, xmax = 202, ymin = 0.4, 
+           ymax = length(levels(A_IAAs$tair_symbol)) + 0.6, color= "black", 
+           fill = NA) + 
+  annotate("text", x = (202-193)/2+193, 
+           y = length(levels(A_IAAs$tair_symbol)) + 1.2, label = "degron", 
+           family="Arial") + 
+  annotate("rect", xmin = 245, xmax = 247, ymin = 0.4, 
+           ymax = length(levels(A_IAAs$tair_symbol))+0.6, color = "black", 
+           fill = NA) +
+  annotate("rect", xmin = 252, xmax = 254, ymin = 0.4, 
+           ymax = length(levels(A_IAAs$tair_symbol)) + 0.6, color = "black", 
+           fill = NA) +
+  annotate("rect", xmin = 315, xmax = 317, ymin = 0.4, 
+           ymax = length(levels(A_IAAs$tair_symbol)) + 0.6, color = "black", 
+           fill = NA) +
+  annotate("rect", xmin = 319, xmax = 321, ymin = 0.4, 
+           ymax = length(levels(A_IAAs$tair_symbol)) + 0.6, color = "black", 
+           fill = NA) + 
+  annotate("text", x = (321-245)/2+245, y= length(levels(A_IAAs$tair_symbol)) + 1.2, label = "PB1 domain", family="Arial")
+
+#ggsave("A-IAAs_alignment.png", width = 6, height = 4)
 ```
@@
 Natural variation also provides a means to study how gene families are evolving.
-To do this, we used ViVa to map the diversity at nonsynonymous variant sites relative to synonymous sites to phylogenetic trees.
-This visualization makes it straightforward to compare rates of recent functional divergence within the context of rates of sequence divergence across the entire gene family.
-In this way, a researcher chould hone in on gene clades that are likely to exhibit high rates of functional redundancy or, conversely, gene clades where there is a greater likelihood that novel functions have emerged.
+To do this, we used ViVa to map onto the Aux/IAA phylogenetic tree the diversity at nonsynonymous variant sites relative to synonymous sites (Figure \@ref(fig:IAAtree)).
+This visualization enables straightforward comparison of rates of recent functional divergence within the context of rates of sequence divergence across the entire gene family.
+By comparing nonsynonymous diversity gene clades can be identified that are likely to exhibit high rates of functional conservation and possible redundancy or conversely, where there is a possibility for recent emergent novel function or pseudogenization.
 
-Previous research has found evidence of both broad genetic redundancy and specificity within closely related pairs or groups of IAA proteins [@overvoorde_functional_2005; @winkler_variation_2017].
-For example, the *iaa8-1 iaa9-1* double mutant and the *iaa5-1 iaa6-1 iaa19-1* triple mutant have wild-type phenotypes [@overvoorde_functional_2005], yet the *IAA6/19* sister pair has significant differences in expression patterns, protein abundances and functions suggesting they have undergone functional specialization since their divergence [@winkler_variation_2017].
-A closer examination of the *IAA19* and *IAA6* pair in *A.
-thaliana* and *A.
-lyrata* found that *IAA6* was diversifying far faster than [@winkler_variation_2017].
-Consistent with these results, ViVa revealed higher conservation for *IAA19* (${\pi_N/\pi_S}$ = 0.55) compared to *IAA6* (${\pi_N/\pi_S}$ = 2.3) (Figure \@ref(fig:IAAtree)), and also detected high diversity within the same regions of *IAA6* as seen in [@winkler_variation_2017].
-This pattern---one *Aux/IAA* sister showing high nonsynonymous diversity while the other sister was more conserved---was observed frequently across the family, suggesting this could be a recurring feature in *Aux/IAA* evolution supporting the large diversity in auxin functions.
+Previous research has found evidence of both broad genetic redundancy among the *Aux/IAAs* and also specificity within closely related pairs or groups of Aux/IAA proteins [@overvoorde_functional_2005; @winkler_variation_2017].
+For example, the *iaa8-1 iaa9-1* double mutant and the *iaa5-1 iaa6-1 iaa19-1* triple mutant have wild-type phenotypes [@overvoorde_functional_2005], yet the *IAA6/IAA19* sister pair has significant differences in expression patterns, protein abundances and functions suggesting they have undergone functional specialization since their divergence [@winkler_variation_2017].
+A closer examination of the *IAA19* and *IAA6* pair within *Brassicaceae* found evidence for positive selection and possible subfunctionalization of *IAA6* relative to *IAA19* [@winkler_variation_2017].
+Consistent with these results, ViVa revealed higher conservation for *IAA19* (${\pi_N/\pi_S}$ = 0.55) compared to *IAA6* (${\pi_N/\pi_S}$ = 2.3) (Figure \@ref(fig:IAAtree)), and also detected high nonsynonymous diversity within the same regions of *IAA6* as seen by Winkler et al.
+(Figure \@ref(fig:IAA6-div-plot)).
+This pattern---one sister showing high nonsynonymous diversity while the other sister was more conserved---was observed frequently across the *Aux/IAA* as well as the *AFB* and *ARF* families (Figure \@ref(fig:AFBtree) and \@ref(fig:ARFtree)), suggesting this could be a recurring feature in the evolution of these families supporting the large diversity in auxin functions.
 
-(ref:IAAtree) **IAA protein sequence tree mapped with ${\pi_N/\pi_S}$.** Protein sequences were aligned [@wright_decipher_2015] and low information content regions were masked [@kuck_parametric_2010] prior to infering a phylogeny [@ronquist_mrbayes_2003].
-Tips of the tree are mapped with circles of diameter proportional to ${\pi_N/\pi_S}$ and also are colored according to ${\pi_N/\pi_S}$.
+(ref:IAAtree) **IAA protein sequence tree mapped with ${\pi_N/\pi_S}$.** Protein sequences were aligned with DECIPHER [@wright_decipher_2015] and low information content regions were masked [@kuck_parametric_2010] prior to inferring a phylogeny [@ronquist_mrbayes_2003].
+Tips of the tree are mapped with circles of color and diameter proportional to ${\pi_N/\pi_S}$.
+${\pi_N/\pi_S}$ provides a prediction of functional diversity.
 Nodes are labeled with the poster probability of monophyly.
+There are two distinct clades of Aux/IAAs represented by the majority of the A and B classes.
+C class Aux/IAAs are missing one or more of the canonical Aux/IAA domains.
 
@@
 ```
-```{r IAAtree, fig.height=5, fig.width=5, fig.cap="(ref:IAAtree)"}
+```{r IAAtree, fig.height=5.5, fig.width=5, fig.cap="(ref:IAAtree)"}
 p <- ggtree(tree, layout = "rectangular", color = "grey") %<+% IAA_data + theme_tree() + scale_x_continuous(expand = c(0, 1.2)) + 
@@
                                         max(IAA_data$PiN_PiS)), 
-                                      to = c(0,1)), guide = "colourbar") + scale_size(guide = "none") + 
-  theme(legend.position = c(.9,.7)) + 
+                                      to = c(0,1)), guide = "colourbar", breaks = c(min(IAA_data$PiN_PiS), 1,
+                                        max(IAA_data$PiN_PiS)), labels = c(min(IAA_data$PiN_PiS), 1,
+                                        max(IAA_data$PiN_PiS))) + scale_size(guide = "none") + 
+  theme(legend.position = c(.8,.7)) + 
   geom_nodelab(mapping = aes(label = round(as.numeric(prob),digits = 2)), nudge_x = -.3, nudge_y = .5, size = 3) + 
   geom_tiplab(mapping = aes(color = Class), hjust = -0.25) + 
-  scale_color_viridis(discrete = TRUE)
+  scale_color_viridis(end = .9, discrete = TRUE, option = "A") +  
+  labs(fill = expression(pi [N] / pi [S]), color = "IAA class")
 p
+
+#ggsave("IAA-tree.png", width = 6, height = 5.5)
 ```
 
-The notable exception to the patter of one conserved and one diersifying sister was observed within the Group B *IAA* genes (*IAA10*/*IAA11*, *IAA12*/*IAA13*, *IAA18*/*IAA26*, and *IAA20*/*IAA30*) [@remington_contrasting_2004].
-Among these sister pairs only *IAA12* and *IAA13* have a noticeable difference in nonsynonymous diversity.
-In our analysis, *IAA10* and *IAA11* both showed functional conservation (${\pi_N/\pi_S}$ of 0.80 and 0.67 respectively).
-In support of *IAA10* and *IAA11* playing largely redundant roles, the *Arabidopsis thaliana* EFP browser indicates that *IAA10* and *IAA11* have almost identical expression patterns.
+```{r IAA-tree-order}
+IAA_order <- fct_inorder(na.exclude(p$data$label[base::order(p$data$y, decreasing = TRUE)]), ordered = TRUE)
+```
 
-# Conclusion
+The *Aux/IAA* phylogeny clusters into two distinct clades represented by the A and B classes [@remington_contrasting_2004].
+The C class Aux/IAAs are missing one or more of the canonical Aux/IAA domains.
+We found notable exceptions to pattern of diversification and conservation between sister pairs within the Class B *Aux/IAA* genes.
+The *IAA10*/*IAA11*, *IAA18*/*IAA26*, and *IAA20*/*IAA30* pairs showed similar levels of nonsynonymous diversity.
+For example, *IAA10* and *IAA11* both showed functional conservation (${\pi_N/\pi_S}$ of 0.80 and 0.67 respectively).
+In support of this strong conservation of both *IAA10* and *IAA11*, the *Arabidopsis thaliana* ePlant browser indicates that *IAA10* and *IAA11* have almost identical expression patterns [@waese_eplant_2017].
+Together this evidence suggests a strong dosage requirement for these genes or that they have taken on novel functions since their emergence.

### 2018-12-06 manuscript/03-conclusion.Rmd
bioRxiv submission

# Conclusion {#conclusion}

ViVa has allowed our team of testers from various skill levels and backgrounds to meaningfully access the 1001 genomes dataset. The visualizations of natural variation further supported much of the existing structure-function knowledge of this well studied signaling pathway and facilitated the generation of new hypotheses. Application of ViVa to less studied genes and gene families promises to yield more novel hypotheses, which can be evaluated with mutagenesis and functional assays to glean novel structure/function knowledge from this rich dataset.

ViVa results are intended to inform and inspire hypothesis generation, not be taken as absolute evidence of trends in gene or gene family evolution. Among the cautions worth noting in interpreting results are limitations of short-read sequencing that lead to regions of missing data where low read quality may have prevented variant calls. We have assumed these missing variants are reference alleles, leading to undercounting in ViVa's diversity estimations. Recent advances in sequencing technologies have been combined to generate extremely high quality genomes [@michael_high_2018], and will reduce this source of uncertainty in future resequencing datasets. Another limitation is that the geographic coverage of accessions in the 1001 Genomes dataset is far from uniform, and thus diversity scores may not accurately reflect the allelic distributions of the global *Arabidopsis thaliana* population.

We hope that ViVa will advance understanding of genotype-phenotype relationships by allowing all researchers access to large resequencing datasets. In the future, we intend to expand ViVa beyond the plant genetics workhorse, *Arabidopsis thaliana*, to more agriculturally relevant species with existing resequencing projects, such as rice [@wang_genomic_2018] and soybean [@zhou_resequencing_2015]. Indeed, the ViVa framework is readily adaptable to any source of targeted resequencing data. If François Jacob’s metaphor holds true, and evolution is indeed a tinkerer and not an engineer [@jacob_evolution_1977], it is only by examining the largest possible number of nature’s solutions that we may eventually decipher the principles constraining innovations in form and function.

### 2018-12-06 manuscript/04-methods.Rmd
bioRxiv submission

# Methods {#methods}

## Data Sources

### Variant Data
Variant data was queried from the 1001 genomes project (http://1001genomes.org) via URL requests to their API service (http://tools.1001genomes.org/api/index.html). These queries returned subsets of the whole-genome variant call format (VCF) file as SnpEFF VCF files. The whole-genome VCF file can be found on the project’s website at http://1001genomes.org/data/GMI-MPI/releases/v3.1/.

### Germplasm Accession Information
A dataset of each of the 1135 accessions including CS stock numbers and geographic location where the samples were collected was retrieved from the 1001 Genomes website at http://1001genomes.org/accessions.html, via the download link at the bottom of the page. This data file has been embedded in the R package as `accessions`.

### Gene and Transcript Accession Information
Information on the genes and transcripts including chromosomal coordinates, start and end location, and transcript length were pulled from Araport11 [@cheng_araport11_2017]. The TAIR10 database, found at http://arabidopsis.org, was accessed via the `biomart` function, using the biomaRt R-package. The Araport11 full genome general feature format file, which can also be found on the TAIR website (https://www.arabidopsis.org/download/index-auto.jsp?dir=%2Fdownload_files%2FGenes%2FAraport11_genome_release), has been embedded in the R-package as GRanges object, `gr`. Gene identifiers used in this study are in Table \@ref(tab:gene-ids-table).
```{r gene-ids-table}
AFBs <- read.csv(system.file("shiny-app", "gene_family_data",
                      "AFB_gene_ids.csv", package = "r1001genomes"))
IAAs <- read.csv(system.file("shiny-app", "gene_family_data",
                      "IAA_gene_ids.csv", package = "r1001genomes"))
TPLs <- read.csv(system.file("shiny-app", "gene_family_data",
                      "TPL_gene_ids.csv", package = "r1001genomes"))
ARFs <- read.csv(system.file("shiny-app", "gene_family_data",
                      "ARF_gene_ids.csv", package = "r1001genomes"))
gene_ids <- bind_rows(AFBs, IAAs, TPLs, ARFs)
kable(gene_ids, col.names = c("Arabidopsis AGI locus identifier", "Gene symbol", "Clade/class"),caption = "Full list of genes used in this study, by identifier, symbol and classification.")
```

## Ranking of variant functional effects
Alignments were colored according to the strongest effect variant allele occurring at any frequency at that position as reported in the SnpEFF "effect" field, per the scale in Figure \@ref(fig:SNPeff-order).

(ref:SNPeff-order) **Rank order of the strength of functional effects** Effect classes were ordered by subjective prediction of average strength of effect on gene function. Strength was then assigned to each effect on an integer scale.
```{r SNPeff-order, fig.cap="(ref:SNPeff-order)"}
SNPeff_order$effect <- factor(as.character(SNPeff_order$effect), levels = as.character(SNPeff_order$effect), ordered = TRUE)

ggplot(data = SNPeff_order, aes(x = effect, y = 1)) + 
  geom_raster(mapping = aes(fill = strength)) +
  scale_fill_viridis(breaks = range(SNPeff_order$strength),
                      limits = range(SNPeff_order$strength),
                      na.value = "grey85",
                      direction = -1,
                      labels = c("weak", "strong"),
                      option = "A", 
                      guide = "none") + 
  coord_flip() + 
  theme_minimal() + theme(axis.text.x = element_blank()) + 
  geom_text(mapping = aes(y = 1, label = strength, color = strength)) +
  scale_color_gradient(low = "black", high = "white", guide = "none") + 
  labs(y = "strength of functional effect", x = "variant effect")

```

## Nucleotide Diversity Calculation
Nei and Li defined the nucleotide diversity statistic in their original paper as:
"the average number of nucleotide differences per site between two randomly chosen DNA sequences” [@nei_mathematical_1979], and provided the equation:

\begin{equation}
\displaystyle \pi = \sum_{ij}x_{i}x_{j}\pi_{ij}.
(\#eq:nei)
\end{equation}

Where $x_{i}$ is the frequency of the ith sequence in the population and $\pi_{ij}$ is the number of sites that are different between the $i$th and $j$th sequence divided by sequence length.

A more general form, that treats each sequence in the population as unique can be written as:

\begin{equation}
\displaystyle \pi = \frac{1}{L*n^2} \sum_{i=1}^n \sum_{j=1}^n \sum_{k=1}^L{\pi_{ijk}}
\qquad
\displaystyle \pi_{ijk} = \left\{\begin{array}{l}
1 \quad \textrm{if} \quad N_{ik} \neq N_{jk} \\
0 \quad \textrm{if} \quad N_{ik} = N_{jk}
\end{array}\right.
(\#eq:nei-gen)
\end{equation}

where $N_{ik}$ is the nucleotide (A, T, C or G) at position $k$ on the *i*th sequence of the population.
$L$ is the length of the sequence. Indels are excluded from the diversity calculation leading to a single $L$ for the population. $n$ is the total number of sequences in the population.

From this form we can re-arrange summations to the form below:

\begin{equation}
\displaystyle \pi = \frac{1}{L} \sum_{k=1}^L{\pi_{k}}
\qquad
\pi_{k} = \frac{1}{n^2} \sum_{i=1}^n \sum_{j=1}^n {\pi_{ijk}}
(\#eq:nei-rearrange)
\end{equation}

where $\pi_k$ can be thought of as the site-wise nucleotide diversity at position $k$, and is equal to 
the nucleotide diversity of a sequence of length 1 at location $k$. We can calculate $\pi_k$ for each
site, then average those over the sequence length to calculate $\pi$, the nucleotide diversity 
of the sequence.

The function `Nucleotide_diversity` in the r1001genomes package calculates $\pi_k$ for each position in the gene or region that contains a variant. Note, $\pi_k$ is equal to 0 at all locations without variants.
This is also what is displayed in the Diversity Plot tab of the webtool.

## Detailed $\pi_k$ calculation simplification.

The formula for $\pi_k$ above requires comparing every sequence to every other sequence at location k,
however, we know there are only a few variant forms at each individual location.

So, we can revert back to using Nei and Li's original formula \@ref(eq:nei), modifying it slightly,
replacing $x_i$ with $\frac{n_i}{n}$, $n_i$ being the number of sequences in the population with nucleotide $N_i$ at location $k$:

\begin{equation}
\displaystyle \pi_k = \sum_{ij} \frac{n_i}{n}\frac{n_j}{n} \pi_{ij} = \frac{1}{n^2}\sum_{ij} n_i n_j \pi_{ij}
\qquad
\pi_{ij(k)} = \left\{\begin{array}{l}
1 \quad \textrm{if} \quad i \neq j \\
0 \quad \textrm{if} \quad i = j
\end{array}\right.
(\#eq:nei-simp1)
\end{equation}

Note that in equation \@ref(eq:nei) subscripts $i$ and $j$ are summed over all sequences in the population, however in equation \@ref(eq:nei-simp1) $i$ and $j$ are only summed over unique variants at a particular location k.

We will define $n_{!i} = n - n_i$ as the number of sequences different from $i$ at position $k$. We can also see that the summed term will be zero if $i=j$, and $n_i n_j$ if $i\neq j$. Therefore:

\begin{equation}
\displaystyle \pi_k = \frac{1}{n^2} \sum_{i} n_i n_{!i}
(\#eq:nei-simp2)
\end{equation}

Next we substitute our definition of $n_{!i}$:

\begin{equation}
\displaystyle \pi_k = \frac{1}{n^2} \sum_{i} n_i (n - n_{i})
(\#eq:nei-simp3)
\end{equation}

Distributing and splitting summation yields:

\begin{equation}
\displaystyle \pi_k = \frac{1}{n^2} (n \sum_{i} n_i - \sum_{i} n_{i}^2)
(\#eq:nei-simp4)
\end{equation}

Finally, summing $\displaystyle \sum_{i}n_i$ is equal to $n$:

\begin{equation}
\displaystyle \pi_k = \frac{1}{n^2} (n^2 - \sum_{i} n_{i}^2)
(\#eq:nei-simp5)
\end{equation}

This simplified form for $\pi_k$ is used by the app, because the counts of unique variants at a single nucleotide location can easily be summarized in R.

## Software

The r1001genomes package has many software dependencies on other R packages, a few of the key bioinformatics packages used are listed below.

**biomaRt:** used for accessing the TAIR10 database on arabidopsis.org

**vcfR:** used to read in the VCF files in a flat “tidy” format for easy manipulation

**BSgenome:** used as the source for the complete DNA string of the reference genome (Col-0).

**DECIPHER:** used to align nucleotide and amino acid sequences of homologous genes

**GenomicFeatures:** used for handling sequence annotations.

**Biostrings:** provides the underlying framework for the sequence manipulations used for generating and aligning sequences with BSgenome, Decipher, and GenomicFeatures

Other packages that were critical to building ViVa and/or writing this document include: [@R-ape; @R-base; @R-bindrcpp; @R-BiocGenerics; @R-Biostrings; @R-bookdown; @R-colorspace; @R-DECIPHER; @R-dplyr; @R-DT; @R-forcats; @R-gginnards; @R-ggplot2; @R-ggpmisc; @R-ggseqlogo; @R-ggthemes; @R-ggtree; @R-ips; @R-IRanges; @R-knitr; @R-magrittr; @R-plyr; @R-purrr; @R-r1001genomes; @R-RColorBrewer; @R-readr; @R-reshape2; @R-reticulate; @R-rmarkdown; @R-RSQLite; @R-S4Vectors; @R-scales; @R-stringr; @R-tibble; @R-tidyr; @R-tidyverse; @R-treeio; @R-viridis; @R-viridisLite; @R-XML; @R-XVector]

### 2018-12-06 manuscript/05-acknowledgments.Rmd
bioRxiv submission

# Acknowledgements

The authors would like to thank Oghenemega Okolo for assistance testing the ViVa software, and Song Li and Bo Zhang for helpful comments on the manuscript. This work was supported by the National Institute of Health (R01-GM107084) and the Howard Hughes Medical Institute. R.C.W. received fellowship support from the National Science Foundation
(DBI-1402222). B.L.M. and H.K. received support from the M.J. Murdock Charitable Trust. A.R.L. is a Simons Foundation Fellow of the Life Sciences Research Foundation.

### 2018-12-06 manuscript/06-supplement.Rmd
bioRxiv submission

# Supplemental Data {#si}

## *TIR1/AFB* genes {#AFBs} 
Auxin acts by binding to receptors (Auxin-signaling F-Boxes, or AFBs) that in turn target co-repressors (Aux/IAAs) for degradation. The six auxin receptor genes in the model plant *Arabidopsis thaliana*, *TIR1* and *AFB1-5*, evolved through gene duplication and diversification early in the history of vascular plants [@parry_complex_2009]. The rate of co-repressor degradation is determined by the identity of both the receptor and co-repressor [@havens_synthetic_2012], and this rate sets the pace of lateral root development [@guseman_auxininduced_2015].

All members of this family have been shown to bind auxin and Aux/IAA proteins. However, AFB1 has drastically reduced ability to assemble into an SCF complex, due to the substitution E8K in its F-box domain, preventing it from inducing degradation of Aux/IAAs [@yu_untethering_2015]. This lack of SCF formation may allow for the high and ubiquitous AFB1 accumulation observed in *Arabidopsis* tissues [@parry_complex_2009]. Higher order receptor mutants in the family containing *afb1* mutants suggest that *AFB1* has a moderate positive effect on auxin signaling [@dharmasiri_plant_2005]. Additionally, AFB4 and AFB5 have been shown to preferentially and functionally bind the synthetic auxin picloram, while other family members preferentially bind indole-3-acetic acid [@prigge_arabidopsis_2016]. Interestingly, the strength and rate with which TIR1/AFBs are able to bind and mark Aux/IAAs for degradation are variable [@calderonvillalobos_combinatorial_2012; @havens_synthetic_2012]. AFB2 induces the degradation of certain Aux/IAA proteins at a faster rate than TIR1, suggesting some functional specificity has arisen since the initial duplication between the *TIR1/AFB1* and *AFB2/AFB3* clades.

```{r TIR1AFBdivstats}
gene_info <- geneInfoFromFile(
  fname = system.file("shiny-app", "gene_family_data",
                      "AFB_gene_ids.csv", package = "r1001genomes"))

VCF <- readRDS(system.file("shiny-app", "gene_family_data",
                      "AFBs_VCF.rds", package = "r1001genomes"))
div_stats <- ldply(VCF, diversityStats, geneInfo=gene_info, .id="transcript_ID")
div_stats$tair_symbol %<>% fct_reorder(.x =  div_stats$tair_symbol, 
                                       .fun = function(x) 
                                         {as.integer(str_remove(x, 
                                              pattern = "AFB|TIR"))})
div_stats$tair_symbol %<>% fct_relevel("TIR1")
div_stats %<>% arrange(tair_symbol)
div_stats %<>% dplyr::select(tair_symbol, everything())

```

```{r AFB_ alignment}
alignment <- alignCDS(gene_info$transcript_ID)
# make an alignment data frame
aln_df <- makeAlnDF(alignment[[2]])

# make a collapsed alignment for translating positions
collapsed_aln_df <- 
  aln_df %>% dplyr::select(aln_pos, seq_pos, seq_name) %>%
  tidyr::spread(key = seq_name, value = seq_pos)
```

```{r Mask_AFB_alignment, eval=FALSE}
aliscore(as.AAbin(alignment[[2]]), path = "/Volumes/SSD/Applications/ALISCORE_v2.0/Aliscore_v.2.0/",  o = "AT2G39940")
id <- scan(file = "/Volumes/SSD/Applications/ALISCORE_v2.0/Aliscore_v.2.0/input.fas_List_random.txt", sep = " ", quiet = TRUE)
AAAlign_no_gaps <- as.matrix(alignment[[2]])[, id]
write.nexus.data(AAAlign_no_gaps, file = "trees/AFBs.nex", format = "protein")
```
```{bash AFB_phylogeny_construction, eval=FALSE}
cd ~/Desktop/natural-variation-webtool/manuscript/trees

/Applications/MrBayes/mb

execute AFBs.nex
# set IAA33 as the outgroup
outgroup AT2G39940.1

lset nucmodel=protein Rates=invgamma
# or include Nst=6  ngammacat=4 omegavar=M3 
prset aamodelpr=fixed(jones) statefreqpr=fixed(empirical) 
# initial run with aamodelpr=mixed yielded posterior prob of 1 for jones model
propset ExtTBR$prob=0

mcmcp ngen=250000 nchains=4 printfreq=1000 samplefreq=100 burnin=200

mcmc
sump
sumt

```

```{r Load AFB phylogeny and data}
tree <- treeio::read.beast(file = "trees/AFBs.nex.con.tre")

AFB_data <- read.csv(system.file("shiny-app", "gene_family_data",
                      "AFB_gene_ids.csv", package = "r1001genomes"))
AFB_data %<>% left_join(y = div_stats, by = c("name" = "tair_symbol"))
#tip_labels need to be in the first column
AFB_data %<>% dplyr::select(name, everything())
AFB_data$tair_locus %<>% as.character()

AFB_data <- AFB_data[match(gsub(x = tree@phylo$tip.label, pattern = "\\.1",replacement = ""), AFB_data$tair_locus), ]

tree@phylo$tip.label <- AFB_data$name
```

```{r AFBtree, fig.cap="(ref:AFBtree)"}
p <- ggtree(tree, layout = "rectangular", color = "grey") %<+% AFB_data + theme_tree() + scale_x_continuous(expand = c(0, 1.2)) + 
  geom_tippoint(mapping = aes(size = 4*Pi_NS_Ratio, fill = Pi_NS_Ratio), shape = 21, color = "white", show.legend = TRUE) + 
  scale_fill_viridis(values = rescale(c(min(AFB_data$Pi_NS_Ratio),
                                        max(AFB_data$Pi_NS_Ratio)), 
                                      to = c(0,1)), guide = "colourbar", 
                     breaks = round(c(min(AFB_data$Pi_NS_Ratio),
                                        max(AFB_data$Pi_NS_Ratio), 1)
                                    ,2), 
                     labels = round(c(min(AFB_data$Pi_NS_Ratio),
                                        max(AFB_data$Pi_NS_Ratio), 1),
                                    2))+
  scale_size(guide = "none") +
  labs(fill = expression(pi [N] / pi [S])) + 
  theme(legend.position = c(.1,.5)) + 
  #geom_nodelab(mapping = aes(label = node)) +
  geom_nodelab(mapping = aes(label = round(as.numeric(prob),digits = 2)), nudge_x = -.2, nudge_y = .2, size = 3) + 
  geom_tiplab(hjust = -0.25) + 
  scale_color_viridis(discrete = TRUE)  
p <- p %>% ggtree::rotate(12) %>% ggtree::rotate(9)
p
#ggsave("AFBs_tree.pdf", height = 3, width = 4)
```
(ref:AFBtree) **AFB protein sequence tree mapped with ${\pi_N/\pi_S}$.** Protein sequences were aligned with DECIPHER [@wright_decipher_2015] and low information content regions were masked with Aliscore [@kuck_parametric_2010] prior to inferring a phylogeny with MrBayes [@ronquist_mrbayes_2003]. Tips of the tree are mapped with circles of diameter proportional to ${\pi_N/\pi_S}$ and also are colored according to ${\pi_N/\pi_S}$. Nodes are labeled with the posterior probability of monophyly.

Examining the natural sequence variation across the *AFB* family \@ref(fig:AFBtree) revealed that *TIR1* and *AFB1* both had very low nonsynonymous diversity, hinting at their likely functional importance and bringing in to question the inconclusive role of *AFB1* in auxin signaling. *AFB3* and *AFB4* had higher nonsynonymous diversity, while their sister genes, *AFB2* and *AFB5* were more conserved. This matches our current understanding of *AFB3* as playing a minor role in the auxin signaling pathway [@dharmasiri_plant_2005] and suggests *AFB4* may be undergoing pseudogenization, especially when paired with its low expression levels [@prigge_arabidopsis_2016].

```{r full-AFBs-alignment, fig.cap="(ref:full-AFBs-alignment)"}
AFB_order <- fct_inorder(na.exclude(p$data$label[base::order(p$data$y, decreasing = TRUE)]), ordered = TRUE)

aln_df <- left_join(aln_df, dplyr::select(gene_info, "tair_locus",
                                       "tair_symbol", "transcript_ID"),
                        by = c("transcript_ID" = "transcript_ID"))
aln_df$tair_symbol %<>% factor(levels = AFB_order, 
                                       ordered = TRUE)
aln_df$tair_symbol %<>% fct_rev()
VCF <- ldply(.data = VCF, .fun = subset, gt_GT != "0|0")
VCF <- getCodingDiv(VCF)

aln_df <- addSNPsToAlnDF(aln_df, VCF, effect_order = SNPeff_order)
aln_df$gap <- aln_df$seq_pos == "-"

p <-ggplot(aln_df, aes(x = aln_pos, y = tair_symbol)) +
  geom_raster(mapping = aes(fill = strength,
                            alpha = gap)) +
  scale_alpha_manual(values = c("TRUE" = 0, "FALSE" = 1), 
                     labels = c("aligned", "gap"), 
                     guide = guide_legend(override.aes = 
                                            list(fill = c("grey85", "white")),
                       title = "alignment",
                       direction = "vertical")) + 
   scale_fill_viridis(breaks = range(SNPeff_order$strength),
                      limits = range(SNPeff_order$strength),
                      na.value = "grey85",
                      direction = -1,
                      labels = c("weak", "strong"),
                      option = "A", 
                      guide = guide_colorbar(ticks = FALSE, 
                                             title.vjust = .8, 
                                             title.position = "top", 
                                             title.hjust = 0.5)) +
  scale_y_discrete(expand = c(0,0)) +
  scale_x_continuous(expand = c(0,0)) +  # expand increases distance from axis
  labs(x = "codon position in alignment", y = "", fill = "variant functional effect") +
  theme_logo(base_family = "Helvetica") +
  theme(panel.grid = element_blank(), 
        panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        strip.text.x = element_blank(),
        #panel.background = element_rect(fill = "grey85", color = "white"),
        legend.position = c("bottom"),
        legend.key = element_rect(linetype = "solid"),
        legend.spacing = unit(4, "char"), 
        legend.box.spacing = unit(0.1,"line")) 
p

```

(ref:full-AFBs-alignment) **Alignment of the AFB family** Protein sequences were aligned with DECIPHER [@wright_decipher_2015] and variants were mapped to this alignment and colored according to the predicted functional effect of the allele of strongest effect at that position, with light colors having weaker effects on function and darker colors stronger effects. Red indicates missense variants. Color scale is explained in [Methods](#methods).

Although most known functional regions are highly conserved in *AFB1*, there are some nonsynonymous polymorphism in the oligomerization domain that are only present in single accessions (F125E in Can-0 and I163N in Pu2-23). Mutations in this domain of *TIR1* frequently have a semidominant effect on root phenotypes [@dezfulian_oligomerization_2016; @wright_insights_2017]. Characterization of this allele and accession may help determine the role of *AFB1* in this pathway.

The AFB4 and AFB5 receptors have an N-terminal extension prior to their F-box domains. This extension had very high nonsynonymous diversity (Figure \@ref(fig:full-AFBs-alignment), suggesting that this extension does not play an important functional role in these proteins. Additionally, two frameshift variants and one stop-gained variant were observed in *AFB4* supporting its pseudogenization.

## *Aux/IAA* genes {#IAAs}

For simplicity we have included only the alignment of the class A Aux/IAAs in the main manuscript. For completion we include here the complete alignment of the family.

```{r IAAdivstats}
gene_info <- geneInfoFromFile(
  fname = system.file("shiny-app", "gene_family_data",
                      "IAA_gene_ids.csv", package = "r1001genomes"))
#annotation <- readAnnotationFile(filename = system.file("shiny-app", "gene_family_data",
#                      "IAA_annotations.csv", package = "r1001genomes"), gene_info = gene_info)

VCF <- readRDS(system.file("shiny-app", "gene_family_data",
                      "IAAs_VCF.rds", package = "r1001genomes"))
div_stats <- ldply(VCF, diversityStats, geneInfo=gene_info, .id="transcript_ID")
div_stats$tair_symbol %<>% fct_reorder(.x = , div_stats$tair_symbol, 
                                       .fun = function(x) 
                                         {as.integer(str_remove(x, 
                                              pattern = "IAA"))})
div_stats %<>% arrange(tair_symbol)
div_stats %<>% dplyr::select(tair_symbol, everything())

```

```{r IAA alignment sup}
alignment <- alignCDS(gene_info$transcript_ID)
# make an alignment data frame
aln_df <- makeAlnDF(alignment[[2]])

# make a collapsed alignment for translating positions
collapsed_aln_df <- 
  aln_df %>% dplyr::select(aln_pos, seq_pos, seq_name) %>%
  tidyr::spread(key = seq_name, value = seq_pos)
```

```{r Mask_IAA_alignment_sup, eval=FALSE}
aliscore(as.AAbin(alignment[[2]]), path = "/Volumes/SSD/Applications/ALISCORE_v2.0/Aliscore_v.2.0/")
id <- scan(file = "input.fas_List_random.txt", sep = " ", quiet = TRUE)
AAAlign_no_gaps <- as.matrix(alignment[[2]])[, id]
write.nexus.data(AAAlign_no_gaps, file = "manuscript/IAAs.nex", format = "protein")
```
```{bash IAA_phylogeny_construction_sup, eval=FALSE}
~/Desktop/natural-variation-webtool/manuscript

/Applications/MrBayes/mb

execute IAAs.nex
# set IAA33 as the outgroup
outgroup AT5G57420.1

lset nucmodel=protein Rates=invgamma
# or include Nst=6  ngammacat=4 omegavar=M3 
prset aamodelpr=fixed(jones) statefreqpr=fixed(empirical) 
# initial run with aamodelpr=mixed yielded posterior prob of 1 for jones model
propset ExtTBR$prob=0

mcmcp ngen=250000 nchains=4 printfreq=1000 samplefreq=100 burnin=200

mcmc
sump
sumt

```

```{r Load IAA phylogeny and data_sup}
tree <- treeio::read.beast(file = "trees/IAAs.nex.con.tre")

IAA_data <- read.csv("tables/IAAs_table.csv")
#tip_labels need to be in the first column
IAA_data %<>% dplyr::select(name, everything())
IAA_data$name %<>% as.character()
IAA_data$AGI %<>% as.character()

IAA_data <- IAA_data[match(gsub(x = tree@phylo$tip.label, pattern = "\\.1",replacement = ""), IAA_data$AGI), ] 
tree@phylo$tip.label <- IAA_data$name
```

(ref:full-IAA-alignment) **Alignment of the complete Aux/IAA family.** Protein sequences were aligned with DECIPHER [@wright_decipher_2015] and variants were mapped to this alignment and colored according to the predicted functional effect of the allele of strongest effect at that position, with light colors having weaker effects on function and darker colors stronger effects. Red indicates missense variants. Color scale is explained in [Methods](#methods).

```{r full-IAA-alignment, fig.cap="(ref:full-IAA-alignment)", fig.height=8}
IAAs <- alignCDS(gene_info$transcript_ID)
# make an alignment data frame
IAAs <- makeAlnDF(IAAs[[2]])
IAAs <- left_join(IAAs, dplyr::select(gene_info, "tair_locus",
                                       "tair_symbol", "transcript_ID"),
                        by = c("transcript_ID" = "transcript_ID"))
IAAs$tair_symbol %<>% as.factor()
IAAs$tair_symbol %<>% factor(levels = IAA_order, 
                                       ordered = TRUE)
IAAs$tair_symbol %<>% fct_rev()
VCF <- ldply(.data = VCF, .fun = subset, gt_GT != "0|0")
coding_VCF <- getCodingDiv(VCF)
IAAs <- addSNPsToAlnDF(aln_df = IAAs,SNPs = coding_VCF, effect_order = SNPeff_order)
IAAs$gap <- IAAs$seq_pos == "-"

p <-ggplot(IAAs, aes(x = aln_pos, y = tair_symbol)) +
  geom_raster(data = IAAs, 
             # mapping = aes(fill = effects)) +
              mapping = aes(fill = strength,
                            alpha = gap)) +
  scale_alpha_manual(values = c("TRUE" = 0, "FALSE" = 1), 
                     labels = c("aligned", "gap"), 
                     guide = guide_legend(override.aes = 
                                            list(fill = c("grey85", "white")),
                       title = "alignment",
                       direction = "vertical")) + 
   scale_fill_viridis(breaks = range(SNPeff_order$strength),
                      limits = range(SNPeff_order$strength),
                      na.value = "grey85",
                      direction = -1,
                      labels = c("weak", "strong"),
                      option = "A", 
                      guide = guide_colorbar(ticks = FALSE, 
                                             title.vjust = .8, 
                                             title.position = "top", 
                                             title.hjust = 0.5)) +
  scale_y_discrete(expand = c(0,0)) +
  scale_x_continuous(expand = c(0,0)) +  # expand increases distance from axis
  labs(x = "codon position in alignment", y = "", fill = "variant functional effect") +
  theme_logo(base_family = "Helvetica") +
  theme(panel.grid = element_blank(), 
        panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        strip.text.x = element_blank(),
        #panel.background = element_rect(fill = "grey85", color = "white"),
        legend.position = c("bottom"),
        legend.key = element_rect(linetype = "solid"),
        legend.spacing = unit(4, "char"), 
        legend.box.spacing = unit(0.1,"line")) 
p

#ggsave("IAAs.png", width = 7, height = 4)
```

```{r, IAA6-div-plot, fig.cap="(ref:IAA6-div-plot)"}
coding_variants <- VCF[VCF$Transcript_ID == gene_info[gene_info$tair_symbol == "IAA6", "transcript_ID"], ]
p <- plotCodingDiv(uniqueCodingVars = coding_variants)
p + annotate("rect", xmin = 106, xmax = 127, ymin = 0.001, ymax = 1, color = "lightgrey", alpha =.1) + annotate("text", x = 116.5, y = 0.5, label = "region of positive\nselection identified\nin Winkler et al.")
```
(ref:IAA6-div-plot) **IAA6 diversity plot.** Nucleotide diversity of variant positions throughout the IAA6 coding sequence are plotted and colored according to the effect of the variant alleles at each position. The region of positive selection identified by Winkler et al. is highlighed.

## *TPL/TPR* genes {#TPLs}

The auxin signaling pathway utilizes the TOPLESS (TPL) and TOPLESS-related (TPR) family of Gro/TLE/TUP1 type co-repressor proteins to maintain auxin responsive genes in a transcriptionally-repressed state in the absence of auxin [@szemenyei_topless_2008]. In *Arabidopsis thaliana* the five member *TPL/TPR* family includes *TPL* and *TPR1-4*. The resulting proteins are comprised of three structural domains: an N-terminal TPL domain and two WD-40 domains [@long_topless_2006]. TPL/TPR proteins are recruited to the AUX/IAA proteins through interaction with the conserved Ethylene-responsive element binding factor-associated amphiphilic repression (EAR) domain [@szemenyei_topless_2008]. Canonical EAR domains have the amino acid sequence LxLxL, as found in most AUX/IAAs [@overvoorde_functional_2005]. TPL/TPR co-repressors bind EAR domains via their C-terminal to LisH (CTLH) domains found near their N-termini (citations of pre-structure founding papers/reviews). Recent structural analyses of the TPL N-terminal domain have highlighted the precise interaction interface between TPL and AUX/IAA EAR domains, as well as the TPL-TPL dimerization and tetramerization motifs [@martin-arevalillo_structure_2017; @ke_structural_2015]. The residues required for higher-order multimers of TPL tetramers have also been identified [@ma_d53_2017]. Additional interactions with transcriptional regulation and chromatin modifying machinery are likely mediated by two tandem beta propeller domains of TPL/TPRs.

The TOPLESS co-repressor family generally exhibits a high level of sequence conservation at the amino acid sequence level across resequenced *Arabidopsis thaliana* accessions, with all ${\pi_N/\pi_S}$ values below 1 (Figure \@ref(fig:TPLtree)). The closely related *TPL* and *TPR1* have the highest ${\pi_N/\pi_S}$ values, suggesting that these these two related genes tolerate a higher degree of sequence and potentially functional diversity compared to *TPR2/3/4*. The N-terminal TPL domain of the TPL/TPR family is particularly conserved. All nonsynonymous polymorphisms observed in this region are either in the coils between helices or are highly conservative mutations within helices (i.e. valine to isoleucine), which would be predicted to exhibit little effect on folding and function.

```{r TPL-SNPstats}
gene_info <- geneInfoFromFile(
  fname = system.file("shiny-app", "gene_family_data",
                      "TPL_gene_ids.csv", package = "r1001genomes"))
#annotation <- readAnnotationFile(filename = system.file("shiny-app", "gene_family_data",
#                      "TPL_annotations.csv", package = "r1001genomes"), gene_info = gene_info)

VCF <- readRDS(system.file("shiny-app", "gene_family_data",
                      "TPL_VCF.rds", package = "r1001genomes"))
div_stats <- ldply(VCF, diversityStats, geneInfo=gene_info, .id="transcript_ID")
div_stats$tair_symbol %<>% fct_reorder(.x = , div_stats$tair_symbol, 
                                       .fun = function(x) 
                                         {as.integer(str_remove(x, 
                                              pattern = "TPR"))})
div_stats$tair_symbol %<>% fct_relevel("TPL")
div_stats %<>% arrange(tair_symbol)
div_stats %<>% dplyr::select(tair_symbol, everything())
```

```{r TPL-alignment-coding-vars}
alignment <- alignCDS(gene_info$transcript_ID)
#possible outgroups  AtLUG, At4g32551; AtLUH, At2g32700
# make an alignment data frame
aln_df <- makeAlnDF(alignment[[2]])

#get
VCF <- ldply(.data = VCF, .fun = subset, gt_GT != "0|0")
VCF <- getCodingDiv(VCF)

# make a collapsed alignment for translating positions
collapsed_aln_df <- 
  aln_df %>% dplyr::select(aln_pos, seq_pos, seq_name) %>%
  tidyr::spread(key = seq_name, value = seq_pos)
```

```{r Mask_TPL_alignment, eval=FALSE}
aliscore(as.AAbin(alignment[[2]]), path = "/Volumes/SSD/Applications/ALISCORE_v2.0/Aliscore_v.2.0/")
id <- scan(file = "/Volumes/SSD/Applications/ALISCORE_v2.0/Aliscore_v.2.0/input.fas_List_random.txt", sep = " ", quiet = TRUE)
AAAlign_no_gaps <- as.matrix(alignment[[2]])[, id]
write.nexus.data(AAAlign_no_gaps, file = "TPLs.nex", format = "protein")
```
```{bash TPL_phylogeny_construction, eval=FALSE}
/Applications/MrBayes/mb

execute TPLs.nex
# set ARF17 as the outgroup
outgroup AT1G77850.1

lset nucmodel=protein Rates=invgamma
# or include Nst=6  ngammacat=4 omegavar=M3 
prset aamodelpr=fixed(jones) statefreqpr=fixed(empirical) 
# initial run with aamodelpr=mixed yielded posterior prob of 1 for jones model
propset ExtTBR$prob=0

mcmcp ngen=250000 nchains=4 printfreq=1000 samplefreq=100 burnin=200

mcmc
sump
sumt

```
```{r Load TPL phylogeny and data}
tree <- treeio::read.beast(file = "trees/TPLs.nex.con.tre")

TPL_data <- read.csv(system.file("shiny-app", "gene_family_data",
                      "TPL_gene_ids.csv", package = "r1001genomes"))
TPL_data$name <- as.character(TPL_data$name)
TPL_data %<>% left_join(y = div_stats, by = c("name" = "tair_symbol"))
#tip_labels need to be in the first column
TPL_data %<>% dplyr::select(name, everything())
TPL_data$tair_locus %<>% as.character()

TPL_data <- TPL_data[match(gsub(x = tree@phylo$tip.label, pattern = "\\.1",replacement = ""), TPL_data$tair_locus), ]

tree@phylo$tip.label <- TPL_data$name

```

```{r TPLtree, fig.cap="(ref:TPLtree)"}
p <- ggtree(tree, layout = "rectangular", color = "grey") %<+% TPL_data + theme_tree() + scale_x_continuous(expand = c(0, 1.2)) + 
  geom_tippoint(mapping = aes(size = 4*Pi_NS_Ratio, fill = Pi_NS_Ratio), shape = 21, color = "white", show.legend = TRUE) + 
  scale_fill_viridis(values = rescale(c(min(TPL_data$Pi_NS_Ratio),
                                        max(TPL_data$Pi_NS_Ratio),1), 
                                      to = c(0,1)), guide = "colourbar", 
                     breaks = c(min(TPL_data$Pi_NS_Ratio),
                                        max(TPL_data$Pi_NS_Ratio), 1), 
                     labels = round(c(min(TPL_data$Pi_NS_Ratio),
                                        max(TPL_data$Pi_NS_Ratio), 1),
                                    2)) + scale_size(guide = "none") +
  labs(fill = expression(pi [N] / pi [S])) + 
  theme(legend.position = c(.1,.5)) + 
  #geom_nodelab(mapping = aes(label = node)) +
  geom_nodelab(mapping = aes(label = round(as.numeric(prob),digits = 2)), nudge_x = -.2, nudge_y = .2, size = 3) + 
  geom_tiplab(hjust = -0.25) 
p
```

(ref:TPLtree) **TPL protein sequence tree mapped with ${\pi_N/\pi_S}$.** Protein sequences were aligned with DECIPHER [@wright_decipher_2015] and low information content regions were masked with Aliscore [@kuck_parametric_2010] prior to inferring a phylogeny with MrBayes [@ronquist_mrbayes_2003]. Tips of the tree are mapped with circles of diameter proportional to ${\pi_N/\pi_S}$ and also are colored according to ${\pi_N/\pi_S}$. Nodes are labeled with the poster probability of monophyly.

The high degree of conservation in the entire N-terminal domain underscores its importance in TPL/TPR function (Figure \@ref(fig:full-TPLs-alignment). For example, the initial *tpl-1* mutation (N176H) in the ninth helix is a dominant gain-of-function allele [@long_topless_2006], which is capable of binding wild-type TPL protein and inducing protein aggregation [@ma_d53_2017]. It is therefore understandable that this helix had very low diversity as nonsynonymous variants in this domain could act in a dominant negative fashion.

(ref:full-TPLs-alignment) **Alignment of the TPL/TPR family.** Protein sequences were aligned with DECIPHER [@wright_decipher_2015] and variants were mapped to this alignment and colored according to the predicted functional effect of the allele of strongest effect at that position, with light colors having weaker effects on function and darker colors stronger effects. Red indicates missense variants. Color scale is explained in [Methods](#methods).
```{r full-TPLs-alignment, fig.cap= "(ref:full-TPLs-alignment)"}
TPL_order <- fct_inorder(na.exclude(p$data$label[base::order(p$data$y, decreasing = TRUE)]), ordered = TRUE)

aln_df <- left_join(aln_df, dplyr::select(gene_info, "tair_locus",
                                       "tair_symbol", "transcript_ID"),
                        by = c("transcript_ID" = "transcript_ID"))
aln_df$tair_symbol %<>% factor(levels = TPL_order, 
                                       ordered = TRUE)
aln_df$tair_symbol %<>% fct_rev()

aln_df <- addSNPsToAlnDF(aln_df, VCF, effect_order = SNPeff_order)
aln_df$gap <- aln_df$seq_pos == "-"

p <-ggplot(aln_df, aes(x = aln_pos, y = tair_symbol)) +
  geom_raster(data = aln_df, 
             # mapping = aes(fill = effects)) +
              mapping = aes(fill = strength,
                            alpha = gap)) +
  scale_alpha_manual(values = c("TRUE" = 0, "FALSE" = 1), 
                     labels = c("aligned", "gap"), 
                     guide = guide_legend(override.aes = 
                                            list(fill = c("grey85", "white")),
                       title = "alignment",
                       direction = "vertical")) + 
   scale_fill_viridis(breaks = range(SNPeff_order$strength),
                      limits = range(SNPeff_order$strength),
                      na.value = "grey85",
                      direction = -1,
                      labels = c("weak", "strong"),
                      option = "A", 
                      guide = guide_colorbar(ticks = FALSE, 
                                             title.vjust = .8, 
                                             title.position = "top", 
                                             title.hjust = 0.5)) +
  scale_y_discrete(expand = c(0,0)) +
  scale_x_continuous(expand = c(0,0)) +  # expand increases distance from axis
  labs(x = "codon position in alignment", y = "", fill = "variant functional effect") +
  theme_logo(base_family = "Helvetica") +
  theme(panel.grid = element_blank(), 
        panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        strip.text.x = element_blank(),
        #panel.background = element_rect(fill = "grey85", color = "white"),
        legend.position = c("bottom"),
        legend.key = element_rect(linetype = "solid"),
        legend.spacing = unit(4, "char"), 
        legend.box.spacing = unit(0.1,"line")) 
p

```

# *ARF* genes {#ARFs}

Auxin response is mediated by the auxin responsive transcription factors (ARFs). There are 23 *ARF*s in *Arabidopsis thaliana* that are divided into three phylogenetic classes. Class A ARFs (ARF5, ARF6, ARF7, ARF8 and ARF19) activate transcription. These ARFs have a glutamine-rich region in the middle of the protein that may mediate activation [@guilfoyle_auxin_2007]. It has recently been shown that the middle region of ARF5 interacts with the SWI/SNF chromatin remodeling ATPases BRAMA and SPLAYED, possibly to reduce nucleosome occupancy and allow for the recruitment of transcription machinery [@wu_auxinregulated_2015]. Additionally, ARF7 interacts with Mediator subunits, directly tethering transcriptional activation machinery to its binding sites in the chromosome [@ito_auxindependent_2016]. Class B and C *ARF*s are historically categorized as repressor *ARF*s, though the mechanism through which they confer repression has not been identified. Their middle regions tend to be proline- and serine-rich.

Canonical ARFs are comprised of three major domains. Recent crystallization of these domains have informed structure-function analysis of the ARFs [@boer_structural_2014; @korasick_molecular_2014; @nanao_structural_2014]. These domains are conserved throughout land plants [@mutte_origin_2018]. ARFs share an N-terminal B3 DNA binding domain. Flanking this DNA-binding domain is a dimerization domain, which folds up into a single “taco-shaped” domain to allow for dimerization between ARFs. There is an auxiliary domain that immediately follows and interacts with the dimerization domain. The middle region is the most variable between ARFs, as mentioned above, but is characterized by repetitive units of glutamine (class A), serine, or proline residues (classes B and C). The C-terminal domain of canonical ARFs is a PB1 protein-protein interaction domain mediating interactions among ARFs, between ARFs and other transcription factors, and between ARFs and the Aux/IAA repressors. This interaction domain was recently characterized as a Phox and Bem1 (PB1) domain, which is comprised of a positive and negative face with conserved basic and acidic residues, respectively [@korasick_molecular_2014; @nanao_structural_2014]. The dipolar nature of the PB1 domain may mediate multimerization by the pairwise interaction of these faces on different proteins as the ARF7 PB1 domain was crystallized as a multimer [@korasick_molecular_2014]. However, it is unclear whether ARF multimerization occurs or plays a significant role *in vivo*. Interfering with ARF dimerization in either the DNA-binding proximal dimerization domain or the PB1 domain decreases the ability of class A ARFs to activate transcription in a heterologous yeast system [@pierre-jerome_functional_2016].

While domain architecture is broadly conserved among the ARFs, there are exceptional cases. Three ARFs do not contain a PB1 domain at all, ARF3, ARF13, and ARF17, and several more have lost the conserved acidic or basic residues in the PB1 domain, suggesting they may be reduced to a single interaction domain. Several ARFs additionally have an expanded conserved region within the DNA-binding domain, of unknown function. The majority of domain variation among ARFs occurs in the large B-class subfamily. The liverwort *Marchantia polymorpha* has a single representative ARF of each class [@flores-sandoval_simple_2015]. The expansion of these classes in flowering plants is the result of both whole genome and tandem duplication events [@remington_contrasting_2004]. The growth of the ARF family may have allowed for the expansion of the quantity and complexity of loci regulated by the ARFs and subsequent expansion in their regulation of developmental processes [@mutte_origin_2018].

Class A ARFs are the most well-studied ARF subfamily---the five family members all act as transcriptional activators and have well-characterized, distinct developmental targets. Overall the diversity of class A *ARF*s was generally low, especially compared to the class B and C *ARF*s (Figure \@ref(fig:ARFtree)), suggesting that class A *ARF*s are central to auxin signal transduction and plant development. Analysis of class A *ARF* nonsynonymous diversity suggests that the majority of these *ARF*s are highly functionally conserved, with ${\pi_N/\pi_S}$ values much lower than 1 with the exception of *ARF19*, with ${\pi_N/\pi_S}$ value of 1.8. Comparing diversity within sister pairs, there is a similar trade-off as seen in most *IAA* sister pairs, with one sister being highly conserved and the other more divergent. *ARF19* and *ARF8* are the more divergent class A *ARFs*, with ${\pi_N/\pi_S}$ values at least three time those of their sisters, *ARF7* and *ARF6* respectively. This may suggest that ARF6 and ARF7 serve more essential purposes in plant development.

```{r ARF diversity stats}
gene_info <- geneInfoFromFile(
  fname = system.file("shiny-app", "gene_family_data",
                      "ARF_gene_ids.csv", package = "r1001genomes"))
annotation <- readAnnotationFile(filename = system.file("shiny-app", 
  "gene_family_data", "ARF_annotations.csv", 
  package = "r1001genomes"), gene_info = gene_info)

VCF <- readRDS(system.file("shiny-app", "gene_family_data",
                      "ARF_VCF.rds", package = "r1001genomes"))
div_stats <- ldply(VCF, diversityStats, geneInfo=gene_info, .id="transcript_ID")
div_stats$tair_symbol %<>% fct_reorder(.x =  div_stats$tair_symbol, 
                                       .fun = function(x) 
                                         {as.integer(str_remove(x, 
                                              pattern = "ARF"))})
div_stats %<>% arrange(tair_symbol)
div_stats %<>% dplyr::select(tair_symbol, everything())

```

```{r ARF-alignment-coding-vars}
alignment <- alignCDS(gene_info$transcript_ID)
# make an alignment data frame
aln_df <- makeAlnDF(alignment[[2]])

#get
VCF <- ldply(.data = VCF, .fun = subset, gt_GT != "0|0")
VCF <- getCodingDiv(VCF)

# make a collapsed alignment for translating positions
collapsed_aln_df <- 
  aln_df %>% dplyr::select(aln_pos, seq_pos, seq_name) %>%
  tidyr::spread(key = seq_name, value = seq_pos)
```

```{r Mask_ARF_alignment, eval=FALSE}
aliscore(as.AAbin(alignment[[2]]), path = "/Volumes/SSD/Applications/ALISCORE_v2.0/Aliscore_v.2.0/")
id <- scan(file = "/Volumes/SSD/Applications/ALISCORE_v2.0/Aliscore_v.2.0/input.fas_List_random.txt", sep = " ", quiet = TRUE)
AAAlign_no_gaps <- as.matrix(alignment[[2]])[, id]
write.nexus.data(AAAlign_no_gaps, file = "ARFs.nex", format = "protein")
```
```{bash ARF_phylogeny_construction, eval=FALSE}
/Applications/MrBayes/mb

execute ARFs.nex
# set ARF17 as the outgroup
outgroup AT1G77850.1

lset nucmodel=protein Rates=invgamma
# or include Nst=6  ngammacat=4 omegavar=M3 
prset aamodelpr=fixed(jones) statefreqpr=fixed(empirical) 
# initial run with aamodelpr=mixed yielded posterior prob of 1 for jones model
propset ExtTBR$prob=0

mcmcp ngen=250000 nchains=4 printfreq=1000 samplefreq=100 burnin=200

mcmc
sump
sumt

```

```{r Load ARF phylogeny and data}
tree <- treeio::read.beast(file = "trees/ARFs.nex.con.tre")

ARF_data <- read.csv(system.file("shiny-app", "gene_family_data",
                      "ARF_gene_ids.csv", package = "r1001genomes"))
ARF_data %<>% left_join(y = div_stats, by = c("name" = "tair_symbol"))
#tip_labels need to be in the first column
ARF_data %<>% dplyr::select(name, everything())
ARF_data$tair_locus %<>% as.character()

ARF_data <- ARF_data[match(gsub(x = tree@phylo$tip.label, pattern = "\\.1",replacement = ""), ARF_data$tair_locus), ]

tree@phylo$tip.label <- ARF_data$name

```

```{r ARFtree, fig.cap="(ref:ARFtree)"}
p <- ggtree(tree, layout = "rectangular", color = "grey") %<+% 
  ARF_data + theme_tree() + scale_x_continuous(expand = c(0, 1.2)) +
  geom_tippoint(mapping = 
                  aes(size = 4*Pi_NS_Ratio, fill = Pi_NS_Ratio),
                shape = 21, show.legend = TRUE, color = "white",
                position = position_nudge(x = 0.1)) + 
  geom_rootpoint(color = "grey") + 
  geom_nodelab(mapping = aes(label = round(as.numeric(prob), 
                                           digits = 2)), 
               nudge_x = -.3, nudge_y = .5, size = 3) +
  scale_fill_viridis(values = rescale(c(min(ARF_data$Pi_NS_Ratio), 1,
                                        max(ARF_data$Pi_NS_Ratio)), 
                                      to = c(0,1)), guide = "colourbar",
                     breaks = c(min(ARF_data$Pi_NS_Ratio), 1,
                                        max(ARF_data$Pi_NS_Ratio)), 
                     labels = round(c(min(ARF_data$Pi_NS_Ratio), 1,
                                        max(ARF_data$Pi_NS_Ratio)), 2)) +
  scale_size(guide = "none") + 
  theme(legend.position = c(.1,.4)) + 
  geom_tiplab(mapping = aes(color = class), hjust = -0.4) + 
  scale_color_viridis(end = .9, discrete = TRUE, option = "A") + 
  labs(fill = expression(pi [N] / pi [S]), color = "ARF class")
p %<>% ggtree::flip(28,32) %>% ggtree::rotate(35)
p 
```

(ref:ARFtree) **ARF protein sequence tree mapped with ${\pi_N/\pi_S}$.** Protein sequences were aligned with DECIPHER [@wright_decipher_2015] and low information content regions were masked with Aliscore [@kuck_parametric_2010] prior to inferring a phylogeny with MrBayes [@ronquist_mrbayes_2003]. Tips of the tree are mapped with circles of diameter proportional to ${\pi_N/\pi_S}$ and also are colored according to ${\pi_N/\pi_S}$. Nodes are labeled with the poster probability of monophyly.

For all class A *ARF*s, the middle region of the protein was the predominant high diversity region (Figure \@ref(fig:full-ARFs-alignment)). In the analyzed natural variation, ARF7 had several expansions of polyglutamine sequences in the middle region. Polyglutamine regions are known to readily expand and contract throughout evolutionary time due to replication error, and variation in polyglutamine length can be acted on by natural selection and have phenotypic consequences [@press_overdue_2014]. The ARF DNA-binding domain had very few, low-diversity missense mutations, as did the critical residues of the PB1 domain. Considering the necessity of their conserved functions, the low level of variation in these key DNA and protein-protein interaction domains is expected.

```{r full-ARFs-alignment, fig.cap="(ref:full-ARFs-alignment)", fig.height=8}
ARF_order <- fct_inorder(na.exclude(p$data$label[base::order(p$data$y, decreasing = TRUE)]), ordered = TRUE)

aln_df <- left_join(aln_df, dplyr::select(gene_info, "tair_locus",
                                       "tair_symbol", "transcript_ID"),
                        by = c("transcript_ID" = "transcript_ID"))
aln_df$tair_symbol %<>% factor(levels = ARF_order, 
                                       ordered = TRUE)
aln_df$tair_symbol %<>% fct_rev()

aln_df <- addSNPsToAlnDF(aln_df, VCF)
aln_df$gap <- aln_df$seq_pos == "-"

p <-ggplot(aln_df, aes(x = aln_pos, y = tair_symbol)) +
  geom_raster(data = aln_df, 
             # mapping = aes(fill = effects)) +
              mapping = aes(fill = strength,
                            alpha = gap)) +
  scale_alpha_manual(values = c("TRUE" = 0, "FALSE" = 1), 
                     labels = c("aligned", "gap"), 
                     guide = guide_legend(override.aes = 
                                            list(fill = c("grey85", "white")),
                       title = "alignment",
                       direction = "vertical")) + 
   scale_fill_viridis(breaks = range(SNPeff_order$strength),
                      limits = range(SNPeff_order$strength),
                      na.value = "grey85",
                      direction = -1,
                      labels = c("weak", "strong"),
                      option = "A", 
                      guide = guide_colorbar(ticks = FALSE, 
                                             title.vjust = .8, 
                                             title.position = "top", 
                                             title.hjust = 0.5)) +
  scale_y_discrete(expand = c(0,0)) +
  scale_x_continuous(expand = c(0,0)) +  # expand increases distance from axis
  labs(x = "codon position in alignment", y = "", fill = "variant functional effect") +
  theme_logo(base_family = "Helvetica") +
  theme(panel.grid = element_blank(), 
        panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        strip.text.x = element_blank(),
        #panel.background = element_rect(fill = "grey85", color = "white"),
        legend.position = c("bottom"),
        legend.key = element_rect(linetype = "solid"),
        legend.spacing = unit(4, "char"), 
        legend.box.spacing = unit(0.1,"line")) 
p

```

(ref:full-ARFs-alignment) **Alignment of the full ARF family.** Protein sequences were aligned with DECIPHER [@wright_decipher_2015] and variants were mapped to this alignment and colored according to the predicted functional effect of the allele of strongest effect at that position, with light colors having weaker effects on function and darker colors stronger effects. Red indicates missense variants. Color scale is explained in [Methods](#methods).

Many of the class B *ARF*s have very high ${\pi_N/\pi_S}$ ratios relative to the other *ARF*s. ARF23 has a truncated DNA-binding domain and had a high ${\pi_N/\pi_S}$ value of 4.1. ARF13 has many high-diversity nonsense variants and lacks a C-terminal PB1 domain. This high level of diversity, prevelance of high-frequency nonsense variants and frequent loss of critical domains, may suggest that several genes in this class are undergoing pseudogenization.

There are also a few highly conserved class B *ARF*s. The high conservation of *ARF1* and *ARF2* is expected as they play critical, redundant roles in senescence and abscission [@ellis_auxin_2005]. Little is known about *ARF9* however, and its low nonsynonymous diversity maybe worthy of investigation.

Class C ARFs show low nucleotide diversity scores, with all ${\pi_N/\pi_S}$ values substantially lower than 1. *ARF16* was the most conserved, whereas its clade members (*ARF10*, *ARF17*) had scores at least four times higher. Structurally, all three members of Class C ARFs contain a canonical B3 DNA-binding domain, but only ARF10 and ARF16 contain a PB1 domain. The DNA binding domains exhibit overall low diversity. Of the PB1 domain containing class C ARFs, ARF16 exhibits several missense variants which are sporadically distributed, in contrast to the conserved PB1 domain of ARF10. This conservation in the PB1 domain of ARF10 and the DBD of ARF16 may suggest subfunctionalization in this family.

### 2018-12-06 manuscript/Supplement/AFBs.Rmd
bioRxiv submission

--- 2018-10-19 manuscript/03-AFBs.Rmd
@@
 # *TIR1/AFB* genes {#AFBs}
-
-Auxin acts by binding to receptors (Auxin-signaling F-Boxes, or AFBs) that in turn target co-repressors (Aux/IAAs) for degradation.
-The six auxin receptor genes in the model plant *Arabidopsis thaliana* , *TIR1* and *AFB1-5*, evolved through gene duplication and diversification early in the history of vascular plants [@parry_complex_2009].
-The rate of co-repressor degradation is determined by the identity of both the receptor and co-repressor [@havens_synthetic_2012], and this rate sets the pace of lateral root development [@guseman_auxininduced_2015].
-
-All members of this family have been shown to bind auxin and Aux/IAA proteins.
-However, AFB1 has drastically reduced ability to assemble into an SCF complex, due to the substitution E8K in its F-box domain, preventing it from inducing degradation of Aux/IAAs[@yu_untethering_2015].
-This lack of complexation may allow observed high ubiquitous AFB1 accumulation [@parry_complex_2009].
-Higher order mutants in the family containing *afb1* mutants suggest that *AFB1* has a moderate positive effect on auxin signaling.
-Additionally, AFB4 and AFB5 have been shown to preferentially and functionally bind the synthetic auxin picloram, while other family members preferentially bind indole-3-acetic acid [@prigge_arabidopsis_2016].
-Interestingly, the strength and rate with which TIR1/AFBs are able to bind and mark Aux/IAAs for degradation are variable [@calderonvillalobos_combinatorial_2012; @havens_synthetic_2012].
-AFB2 induced the degradation of certain Aux/IAA proteins at a faster rate than TIR1, suggesting some functional specificity has arisen since the initial duplication between the *TIR1/AFB1* and *AFB2/AFB3* clades.
-
-Although it is unclear what unique roles each receptor plays in growth and development, a number of studies have pointed out differences in the ways the six different receptors in A.
-thaliana differ in biochemical function and expression domain[@dharmasiri_plant_2005; @parry_complex_2009; @prigge_arabidopsis_2016].
-*TIR1/AFB* genes are expressed ubiquitously in A.
-thaliana tissues, TIR1, AFB2, and AFB3 have been shown to accumulate in the shoot and root meristems and leaf tissues, with slightly different expression patterns for TIR1 [@parry_complex_2009].
-Additionally, the expression of AFB5 is strongly circadian-regulated [@covington_circadian_2007] and AFB3 is more highly expressed in the roots in the presence of nitrate, facilitating increased lateral root formation [@vidal_nitrate_2014], suggesting more broad environmental regulation of this gene family may exist.
-
-All members of this family have been shown to bind auxin and Aux/IAA proteins.
-However, AFB1 has drastically reduced ability to assemble into an SCF complex, due to the substitution E8K in its F-box domain, preventing it from inducing degradation of Aux/IAAs[@yu_untethering_2015].
-This lack of complexation may be the cause of high, ubiquitous AFB1 accumulation [@parry_complex_2009].
-Higher order mutants in the family containing *afb1* mutants suggest that *AFB1* has a moderate positive effect on auxin signaling.
-Additionally, AFB4 and AFB5 have been shown to preferentially bind the synthetic auxin picloram, while other family members preferentially bind indole-3-acetic acid [@prigge_arabidopsis_2016].
-Interestingly, the strength and rate with which each TIR1/AFB is able to bind and mark an Aux/IAAs for degradation is variable [@calderonvillalobos_combinatorial_2012; @havens_synthetic_2012].
-
-(ref:TIR1AFBdivstats) **_TIR1/AFB_ nucleotide diversity.** The nucleotide diversity as Nei and Li's $\pi$ (the average number of nucleotide differences per site between all possible pairs of sequences at synonymous ($\pi_S$) and nonsynonymous ($\pi_N$) sites, as well as all sites in the coding sequence and full transcript of the *TIR1/AFB* family.
 
@@
 ```
-
-Examining the natural sequence variation across the *AFB* family \@ref(fig:TIR1AFBdivstats) revealed that *TIR1* and *AFB1* both had very low nonsynonymous diversity, hinting at their likely functional importance and bringing in to question the inconclusive role of *AFB1* in auxin signaling.
-*AFB3* and *AFB4* had higher nonsynonymous diversity, while their sister genes, *AFB2* and *AFB5* were more conserved.
-This matches our current understanding of *AFB3* as playing a minor role in the auxin signaling pathway and *AFB4* perhaps undergoing pseudogenization.
 
@@
 ```
-
-```{r AFB variants for PyMOL, eval=FALSE}
-coding_vcf <- ldply(.data = VCF, .fun = subset, 
-             !is.na(transcript_ID) & gt_GT != "0|0")
-aln_df <- addSNPsToAlnDF(aln_df, coding_vcf, seq_name = Transcript_ID, 
-               seq_pos = Codon_Number)
-collapsed_aln_df %<>% left_join(y = subset(aln_df, seq_name == gene_info[gene_info$tair_symbol == "TIR1", "transcript_ID"]))
-
-#' Select variants for importing into PyMOL
-#'
-#' @param gene_info a gene info data frame
-#' @param aln_df an alignment data frame with attached SNPs
-#' @param structure_ref string, which tair_symbol or transcript_ID (depending on 
-#' \code{name_by_symbols}) corresponds to the protein structure to which 
-#' selections should be mapped
-#' @param effect_type string of the effect type to select from the alignment 
-#' data frame
-#' @param name_by_symbols logical, if true (default) tair_symbols will be used 
-#' to name the selections, otherwise transcript_IDs will be used
-#'
-#' @return
-#' @export
-#'
-#' @examples
-select_variants <- function(gene_info, aln_df, structure_ref, 
-                            effect_type = "missense", name_by_symbols = TRUE){
-  collapsed_aln_df <- aln_df %>% as_tibble() %>% 
-    dplyr::select(aln_pos, seq_pos, seq_name) %>%
-    tidyr::spread(key = seq_name, value = seq_pos)
-  if(name_by_symbols)
-    column <- "tair_symbol"
-    else 
-      column <- "transcript_ID"
-
-  selections <- ldply(.data = gene_info[,column], .fun = function(gene){
-    collapsed_aln_df %<>% dplyr::left_join(y = subset(aln_df, seq_name ==  
-                      gene_info[gene_info[, column] == gene, "transcript_ID"]))
-    variants <- collapsed_aln_df[
-      grepl(effect_type, collapsed_aln_df$effects), 
-      gene_info[gene_info[, column] == structure_ref, "transcript_ID"]]
-    c(gene, paste0(na.omit(as.numeric(variants[[1]])), collapse = "+"))
-  })
-  return(selections)
-}
-
-selections <- select_variants(gene_info, aln_df, structure_ref = "TIR1")
-selections$V2 <- paste0("chain B and resi ", selections$V2)
-#write.table(selections, file = "AFB_variant_selections", row.names = FALSE, 
-#          col.names = FALSE, sep = ",", quote = FALSE)
-```
-
-```{python readSelection, eval=FALSE}
-#
-# -- readSelection.py
-#
-from pymol import cmd, stored
-
-def readSelection (fileName):
-	'''
-	DESCRIPTION
-
-	reads a csv file containing a column of selection names and a column of
-	selection strings and creates selection objects for each
-	'''
-	#read in the selection file
-	with open(fileName) as f:
-		lines = f.readlines()
-		for line in lines:
-			sel = line.split(",")
-			cmd.select(sel[0], sel[1])
-
-	f.close()
-
-cmd.extend('readSelection',readSelection);
-```
-
-```{PyMol, AFB_structure, eval=FALSE}
-#from within pymol
-cmd.load("Dropbox/180705_NIH_renewal/Aim1-figures/2P1Q.pdb")
-hide everything
-
-run ~/Desktop/natural-variation-webtool/manuscript/readSelection.py
-
-readSelection ~/Desktop/natural-variation-webtool/manuscript/AFB_variant_selections
-
-# Define selections
-# select TIR1variants, (chain B and resi 154+239+546)
-select fbox, (chain B and resi 1-50)
-select LRR, (chain B and resi 50-576)
-select TIR1, (chain B and resi 1-576)
-# select AFB2variants, (chain B and resi 38+174+181+184+209+259+401+496)
-select tir1_1, (chain B and resi 147)
-select TIR1dm, (chain B and resi 170+473)
-select insPcontacts, (chain B and resi 74+78+113+114+344+401+403+436+460+484+485+509)
-select dimerNULL, (chain B and resi 142+143+151+117+147)
-select auxin, resn IAC
-select AUXcontacts, (chain B and resi 78+79+82+378+380+403+404+405+438+439+462+463+464+489)
-select insP, resn IHP
-select IAA7, chain C
-select IAA7contacts, (chain B and resi 81+82+84+139+347+349+350+351+352+380+405+406+407+409+464+465+489)
-select IAA7_N, chain C and resi 1
-select IAA7_C, chain C and resi 13
-select ASK1, chain A
-select Y, chain C and resi 11
-select Ycontacts, chain B and resi 165+166+195
-
-#Define color palette
-#set_color name, [ red-float, green-float, blue-float ]
-set_color blue1, [0.00000000, 0.20392157, 0.6980392]
-set_color blue2, [0.09803922, 0.36078431, 1.0000000]
-set_color blue3, [0.00000000, 0.29019608, 1.0000000]
-set_color teal,  [0.09803922, 1.00000000, 0.6980392]
-set_color LRR,   [1.00000000, 0.71764706, 0.0000000]
-set_color Fbox,  [0.69803922, 0.49411765, 0.0000000]
-set_color TIR1l, [0.67058824, 0.07843137, 0.8000000]
-set_color TIR1f, [0.63529412, 0.34117647, 0.6980392]
-set_color AFB2l, [0.37647059, 0.70196078, 1.0000000]
-set_color AFB2f, [0.07843137, 0.64313725, 0.8000000]
-
-show cartoon, 2P1Q
-
-color lrr, LRR
-
-color TIR1l, TIR1
-show sticks, TIR1
-color TIR1f, AFB1
-show sticks, AFB1
-color AFB2l, AFB2
-show sticks, AFB2
-color AFB2f, AFB3
-show sticks, AFB3
-
-set ray_trace_mode, 3
-
-# used get_view() to return the current view (position) of the molecule 
-# copy the code to be able to set the identical viewpoint later
-# to return the current view from Pymol
-# get_view
-
-set_view (\
-     0.044671144,    0.083668359,   -0.995409012,\
-     0.774909019,    0.625905454,    0.087390237,\
-     0.630409002,   -0.775298953,   -0.036878210,\
-    -0.001916260,    0.005584979, -202.498916626,\
-    -5.538011074,   -8.507736206,   39.046234131,\
-  -219.237945557,  624.123413086,  -20.000000000 )
-
-#to render a high definition image use ray
-ray
-```
-
-(ref:AFBstructure) A TIR1 structure (2P1Q) in gold in complex with ASK1 in grey and auxin, inositol phosphate and an Aux/IAA degron in teal.
-TIR1 residues or analagous residues containing nonsynonymous polymorphisms in TIR1 and AFB1 colored in purple and in AFB2 and AFB3 colored in blue.
-
-```{r AFBstructure, fig.cap='(ref:AFBstructure)'}
-knitr::include_graphics(path = "structures/natural_variation_TIR1AFB1-3.png")
-```
-
-Although most known functional regions are highly conserved in *AFB1*, there is a nonsynonymous polymorphism  in the oligomerization domain, only found in the Can-0 accession.
-Mutations in this domain of *TIR1* frequently have a semidominant effect on root phenotypes [@dezfulian_oligomerization_2016; @wright_insights_2017].
-Characterization of this allele and accession may help determine the role of *AFB1* in this pathway.
-
-**not sure this is very informative, and a map doesn't seem very exciting, perhaps a domain scale alignment plot would be better here, or above, and would allow us to comment more on where the functional diversity lies within the family.**
-
-```{r AFB1 diversity plot}
-coding_variants <- VCF[VCF$Transcript_ID == gene_info[gene_info$tair_symbol == "AFB1", "transcript_ID"], ]
-p <- plotCodingDiv(uniqueCodingVars = coding_variants)
-annotation <- readAnnotationFile(filename = system.file("shiny-app", "gene_family_data",
-                      "AFB_annotations.csv", package = "r1001genomes"), gene_info = gene_info)
-annotation <- addAlnPosToAnno(anno_df = annotation, aln_df)
-
-# translate positions from TIR1 to AFB1
-AFB1_positions <- annotation$positions
-AFB1_positions %<>% left_join(y = dplyr::select(collapsed_aln_df, aln_pos, gene_info[gene_info$tair_symbol == "AFB1", "transcript_ID"]), by = "aln_pos") 
-AFB1_positions$position <- AFB1_positions$AT4G03190.1
-
-p <- append_layers(p,list(
-        geom_rect(data = AFB1_positions,
-                   mapping = aes(xmin = as.integer(position)-1,
-                                 xmax = as.integer(position)+2,
-                                 fill = annotation),
-                  ymin = -Inf, ymax = Inf, inherit.aes = FALSE, alpha = 0.8)),
-      position = "bottom")
-p
-```
-
-The AFB4 and AFB5 receptors have an N-terminal extension prior to the F-box domains.
-This extension had very high nonsynonymous diversity \@ref(fig:FBoxAlignment), suggesting that this extension does not play an important functional role in these proteins.
-Additionally, two frameshift variants and one stop-gained variant were observed in *AFB4* supporting its pseudogenization.
-
-(ref:FBoxAlignment) **Alignment of A.
-thaliana TIR1/AFB F-box domains showing variants.** Cyan marks the F-box binding domain.
-Yellow, green, and magenta, mark missense and synonymous, synonymous only, and missense only variant locations respectively.
-Yellow squared E’s show Cul1-interacting residues, differing in AFB1.
-
-```{r FBoxAlignment, fig.cap = "(ref:FBoxAlignment)"}
-aln_df <- makeAlnDF(alignment[[2]])
-aln_df <- left_join(aln_df, dplyr::select(gene_info, "tair_locus",
-                                       "tair_symbol", "transcript_ID"),
-                        by = c("transcript_ID" = "transcript_ID"))
-aln_df$tair_symbol <- fct_rev(fct_relevel(aln_df$tair_symbol, "TIR1"))
-aln_df <- chunkAlnDF(aln_df = aln_df, chunk_width = 55)
-aln_df <- addSNPsToAlnDF(aln_df, VCF, effect_order = SNPeff_order)
-F_box <- subset(aln_df, aln_pos <= 96)
-F_box <- chunkAlnDF(aln_df = F_box, chunk_width = 80)
-p <-ggplot(F_box, aes(x = aln_pos, y = tair_symbol,
-                             group = seq_pos, text = variants))
-p <- p +
-  geom_tile(data = na.omit(F_box), mapping = aes(fill = effects),
-            width = 1, height = 1, alpha = 0.5) +
-  geom_text(aes(label=letter), alpha= 1, family = "Courier") +
-  scale_x_continuous(breaks=seq(1,max(F_box$aln_pos), by = 10)) +
-  scale_y_discrete() +
-  # expand increases distance from axis
-  labs(x = "", y = "") +
-  theme_logo(base_family = "Helvetica") +
-  theme(panel.grid = element_blank(), panel.grid.minor = element_blank()) +
-  facet_wrap(facets = ~chunk, ncol = 1, scales = "free") +
-  theme(strip.background = element_blank(),
-        strip.text.x = element_blank(),
-        legend.box = "vertical") + 
-  scale_fill_viridis(option = "A", discrete = TRUE)
-p
-```

### 2018-12-06 manuscript/Supplement/ARFs.Rmd
bioRxiv submission

--- 2018-10-19 manuscript/06-ARFs.Rmd
@@
 Class B and C ARFs are historically categorized as repressor ARFs, though the mechanism through which they confer repression has not been identified.
-Their middle regions tend to be proline- and serine-rich (citation, Ulmasov?).
+Their middle regions tend to be proline- and serine-rich.
 
@@
 
-A few missense variants in *ARF5* highlighted potential functional variation in the class A ARFs.
-A high frequency allele *(What are the variants?)*, resulting in substitutions of small amino acids on the positive face of the PB1 domain for bulky phenylalanines, may disrupt interactions with other ARF or Aux/IAA PB1 domains.
-Interestingly, these variants were always found together in accessions Tu-B2-3, Ru-2, Haes-1, Rd-0, 627ME-13Y1, Le-0, Bu-0 and Bsch-0 **Where are these?**.
-Another highly represented polymorphism in ARF5 results in change in the 17th amino acid from a leucine to a methionine.
-This position is in the N-terminal extension of ARF5 which is of unknown function.
-It is possible that this methionine acts as an alternative start site in the accessions.
-
 ### Class B *ARFs*
@@
 
-ARF14, ARF15, **ARF16 I don't think this is correct ARF 16 is class C, perhaps this should be 12?**, ARF20, and ARF21 all do not have one or multiple of the conserved acidic residues in the PB1 domain, and all are within a subclade of the B ARFs.
-Interestingly, no non-synonymous variation is seen in these important residues for Aux/IAA and ARF interaction, suggesting that the amino acids that are in the regions may still be important for the function of these proteins.
-**Might cut this, it is interesting that perhaps homo/heterodimerization is not as important for this clade of ARFs but I think this conclusion mostly comes from looking at the alignments and not from the natural variation.
-Perhaps if it looks like these ARFs have higher than expected, or more deleterious variation across their PB1 domains**
-
 ### Class C *ARFs*
@@
 
-```{r ARF10_DBD Variants distribution}
+```{r C-ARFs-alignment}
 C_ARFs <- subset(gene_info, tair_symbol %in% ARF_data[which(ARF_data$class == "C"), "name"])
@@
 ```
-
-Several members of class B (ARF2, 3 and 4) and C ARFs (ARF10, 16 and 17) are distinguished from other ARFs on the basis of their post transcriptional regulation by trans-acting small interfering RNAs (tasiRNAs) and microRNAs (miRNAs) [@rhoades_prediction_2002; @allen_micrornadirected_2005].
-**(AMY check for conservation of Arf 6 and 8 targeted by miRNA 167).** The miRNA target of class C ARFs showed a high degree of conservation (Fig.
-ARF10_16_17_Alignment).
-This is consistent with perturbation **what exactly does perturbation mean** studies that have shown drastic developmental phenotype for class C ARFs [@wang_control_2005; @mallory_micrornadirected_2005; @liu_repression_2007].
-For the class B ARFs (2, 3 and 4) the tasiRNA binding motif is also highly conserved suggesting strong regulatory constraints on these genes (Fig ARF2_3_4_Alignment).
-The DNA binding domain of ARF4 exhibits non-synonymous variation in 4 accession and would be interesting to carry out phenotypic characterization of these accessions for developmental growth defects (Fig ARF4_DBD Variants map).
-
-```{r ARF10_16_17_Alignment}
-
-```
-
-```{r ARF2_3_4_Alignment}
-
-```
-
-```{r ARF4_DBD Variants map}
-
-```

### 2018-12-06 manuscript/Supplement/IAAs.Rmd
bioRxiv submission

--- 2018-09-09 manuscript/04-IAAs.Rmd
@@
 # *Aux/IAA* genes {#IAAs}
-
-The Aux/IAA proteins (IAAs) mediate both auxin perception and transcriptional co-repression of auxin responsive genes.
-Auxin perception is carried out along with TIR1/AFB proteins, which together with IAAs act as co-receptors to bind and perceive nuclear auxin.
-Interaction of IAAs with auxin and TIR1/AFBs is conferred by the highly conserved 13 amino acid degron motif [@tan_mechanism_2007].
-Formation of this auxin co-receptor complex triggers ubiquitination and subsequent degradation of IAAs.
-This degradation in turn relieves their repression upon ARF transcription factors.
-Stabilizing mutations within the degron can lead to dominant phenotypes [@berleth_auxin_2004; @yang_iaa1_2004].
-The *Arabidopsis thaliana* Aux/IAA family has 29 members, most of which possess a canonical auxin degron.
-Notable exceptions are the auxin-insensitive IAA20 and IAA30 that lack a degron sequence entirely, while the closely related IAA31 retains a semi-functional degron sequence with a very slow degradation rate [@dreher_arabidopsis_2006; @havens_synthetic_2012].
-Additionally, IAA degradation rate is influenced by rate motifs found outside the degron, such as the KR motif [@dreher_arabidopsis_2006; @moss_rate_2015].
-IAA-mediated repression of ARFs depends upon both an EAR domain, N-terminal of the degron, that recruits co-repressors of the TOPLESS (TPL) family, and a Phox and Bem1 (PB1) domain that enables interaction via hetero-oligomerization with ARF PB1 domains [@nanao_structural_2014; @kagale_ear_2011].
 
@@
 
-###Sister Pair Diversification
-
-Previous research has found evidence of both broad genetic redundancy and specificity within closely related pairs or groups of IAA proteins [@overvoorde_functional_2005].
-For example, the *iaa8-1 iaa9-1* double mutant and the *iaa5-1 iaa6-1 iaa19-1* triple mutant have wild-type phenotypes [@overvoorde_functional_2005], yet the *IAA6/19* sister pair has significant differences in expression patterns, protein abundances and functions suggesting they have undergone functional specialization since their divergence [@winkler_variation_2017].
-
-In our analysis of the *IAA* gene family, we frequently observed that one member of most *IAA* sister pairs had high nonsynonymous diversity while the other sister was more conserved \@ref(fig:IAAtree).
-In an interspecific comparison between *A.
-thaliana* and *A.
-lyrata*, *IAA19* is more conserved, while *IAA6* has regions of increased sequence divergence upstream of the degron and in the N-terminal half of the PB1 domain [@winkler_variation_2017].
-Consistent with these results, our analysis of intraspecific natural variation revealed higher conservation for *IAA19* (${\pi_N/\pi_S}$ = 0.55) compared to *IAA6* (${\pi_N/\pi_S}$ = 2.3) \@ref(fig:IAAtree), and also detected high diversity within the same regions of *IAA6* as seen in Winkler et al.
-(Supp Fig IAA6 diversity plot).
-
 ```{r Mask_IAA_alignment, eval=FALSE}
@@
 ```
-**I think IAA20, 30, and 15 should also be in class C. So there are really two large clades A and B and then there are truncations/loss of function/dominant variants scattered throughout, Class C.**
 
-(ref:IAAtree) **IAA protein sequence tree mapped with ${\pi_N/\pi_S}$.** Protein sequences were aligned [@wright_decipher_2015] and low information content regions were masked [@kuck_parametric_2010] prior to infering a phylogeny [@ronquist_mrbayes_2003].
-Tips of the tree are mapped with circles of diameter proportional to ${\pi_N/\pi_S}$ and also are colored according to ${\pi_N/\pi_S}$.
-Nodes are labeled with the poster probability of monophyly.
-
-```{r IAAtree, fig.height=5, fig.width=5, fig.cap="(ref:IAAtree)"}
-p <- ggtree(tree, layout = "rectangular", color = "grey") %<+% IAA_data + theme_tree() + scale_x_continuous(expand = c(0, 1.2)) + 
-  geom_tippoint(mapping = aes(size = 4*PiN_PiS, fill = PiN_PiS), shape = 21, color = "white", show.legend = TRUE) + 
-  scale_fill_viridis(values = rescale(c(min(IAA_data$PiN_PiS), 1,
-                                        max(IAA_data$PiN_PiS)), 
-                                      to = c(0,1)), guide = "colourbar") + scale_size(guide = "none") + 
-  theme(legend.position = c(.9,.7)) + 
-  geom_nodelab(mapping = aes(label = round(prob,digits = 2)), nudge_x = -.3, nudge_y = .5, size = 3) + 
-  geom_tiplab(mapping = aes(color = Class), hjust = -0.25) + 
-  scale_color_viridis(discrete = TRUE)
-p
-```
-
-Within the Group B *IAA* genes, there are four sister pairs---*IAA10*/*IAA11*, *IAA12*/*IAA13*, *IAA18*/*IAA26*, and *IAA20*/*IAA30* \@ref(fig:IAAtree) [@remington_contrasting_2004].
-Interestingly, among these sister pairs only *IAA12* and *IAA13* have a noticeable difference in nonsynonymous diversity.
-The *Arabidopsis thaliana* EFP browser shows that *IAA10* and *IAA11* have almost identical expression patterns, with the exception of *IAA10* being more highly expressed in the developing seed, and *IAA11* being more highly expressed in root tissue during lateral root development **need to cite original data source here**.
-In our analysis, *IAA10* and *IAA11* both showed functional conservation (${\pi_N/\pi_S}$ of 0.80 and 0.67 respectively) in support of *IAA10* and *IAA11* playing redundant roles.
-Accessions and associated variants identified here provide a potentially expedited route to elucidating the extent of redundancy between family members such as *IAA10* and *IAA11*.
-
-### New Loss-of-Function/Gain-of-Function alleles
-
-Accessions with functional variants can be potentially used to investigate the extent of redundancy/specialization within sister pairs.
-For example, three accessions have a missense variant affecting the EAR domain of IAA10 (HR-5, UKSE06-118 and UKSW06-240, all from southern UK).
-This variant (L49S) is in the second leucine of the LxLxL motif and likely reduces the ability of IAA10 to recruit the TOPLESS co-repressor [@lee_biological_2016].
-The nonsense variant (IAA10-Y219\*) resulting in truncated PB1 domain may be more indicative of *IAA10* null phenotype, as it likely prevents IAA10 from interacting with an ARF.
-This variant is only found in one accession from northeastern Spain (IP-Svi-0).
-These accessions and potential loss-of-function alleles may facilitate exploration of the function of *IAA10*.
-
-Within the available natural sequence variation, we found several missense polymorphisms affecting the EAR domain or degron of several canonical clade A IAAs.
-However, no missense polymorphisms were present in the KR motif or residues critical for PB1-PB1 domain interactions (\@ref(fig:A-IAA-alignment)).
-The majority of missense mutations observed were in the variable regions between the conserved domains.
-
-(ref:A-IAA-alignment) 
-```{r A-IAA-alignment, fig.cap="(ref:A-IAA-alignment)"}
-A_IAAs <- subset(gene_info, tair_symbol %in% IAA_data[which(IAA_data$class == "A"), "name"])
-A_ARFs <- alignCDS(A_IAAs$transcript_ID)
+(ref:full-IAA-alignment) 
+```{r full-IAA-alignment, fig.cap="(ref:full-IAA-alignment)"}
+IAAs <- alignCDS(gene_info$transcript_ID)
 # make an alignment data frame
-A_IAAs <- makeAlnDF(A_IAAs[[2]])
-A_IAAs <- left_join(A_IAAs, dplyr::select(gene_info, "tair_locus",
+IAAs <- makeAlnDF(IAAs[[2]])
+IAAs <- left_join(IAAs, dplyr::select(gene_info, "tair_locus",
                                        "tair_symbol", "transcript_ID"),
                         by = c("transcript_ID" = "transcript_ID"))
-A_IAAs$tair_symbol %<>% as.factor()
-A_IAAs$tair_symbol %<>% fct_reorder(
-    {as.integer(str_remove(A_IAAs$tair_symbol, 
-                           pattern = "ARF"))}, 
+IAAs$tair_symbol %<>% as.factor()
+IAAs$tair_symbol %<>% fct_reorder(
+    {as.integer(str_remove(IAAs$tair_symbol, 
+                           pattern = "IAA"))}, 
     .desc = TRUE)
-A_IAAs <- addSNPsToAlnDF(A_IAAs, VCF)
+VCF <- ldply(.data = VCF, .fun = subset, gt_GT != "0|0")
+coding_VCF <- getCodingDiv(VCF)
+IAAs <- addSNPsToAlnDF(aln_df = IAAs,SNPs = coding_VCF, effect_order = SNPeff_order)
+IAAs$gap <- IAAs$seq_pos == "-"
 
-p <-ggplot(A_IAAs, aes(x = aln_pos, y = tair_symbol,
-                             group = seq_pos, text = variants)) +
-  geom_raster(data = A_IAAs[A_IAAs$seq_pos == "-",], fill = "lightgrey") +
-  geom_raster(data = na.omit(A_IAAs), 
+p <-ggplot(IAAs, aes(x = aln_pos, y = tair_symbol)) +
+  geom_raster(data = IAAs, 
              # mapping = aes(fill = effects)) +
-              mapping = aes(fill = as.integer(as.factor(effects)))) +
-  #scale_fill_viridis(option = "A", discrete = TRUE) +
-   scale_fill_viridis(breaks = range(
-     as.integer(as.factor(na.omit(A_IAAs$effects)))),
-                      labels = c("deleterious", "neutral"),
+              mapping = aes(fill = strength,
+                            alpha = gap)) +
+  scale_alpha_manual(values = c("TRUE" = 0, "FALSE" = 1), 
+                     labels = c("aligned", "gap"), 
+                     guide = guide_legend(override.aes = 
+                                            list(fill = c("grey85", "white")),
+                       title = "alignment",
+                       direction = "vertical")) + 
+   scale_fill_viridis(breaks = range(SNPeff_order$strength),
+                      limits = range(SNPeff_order$strength),
+                      na.value = "grey85",
+                      direction = -1,
+                      labels = c("neutral", "deleterious"),
                       option = "A", 
                       guide = guide_colorbar(ticks = FALSE, 
-                                             title.vjust = .8)) +
-  scale_y_discrete() +
-  # expand increases distance from axis
+                                             title.vjust = .8, 
+                                             title.position = "top", 
+                                             title.hjust = 0.5)) +
+  scale_y_discrete(expand = c(0,0)) +
+  scale_x_continuous(expand = c(0,0)) +  # expand increases distance from axis
   labs(x = "codon position in alignment", y = "", fill = "variant effect") +
   theme_logo(base_family = "Helvetica") +
-  theme(panel.grid = element_blank(), panel.grid.minor = element_blank()) +
-  theme(strip.background = element_blank(),
+  theme(panel.grid = element_blank(), 
+        panel.grid.minor = element_blank(),
+        strip.background = element_blank(),
         strip.text.x = element_blank(),
-        legend.position = "bottom")  
+        #panel.background = element_rect(fill = "grey85", color = "white"),
+        legend.position = c("bottom"),
+        legend.key = element_rect(linetype = "solid"),
+        legend.spacing = unit(4, "char"), 
+        legend.box.spacing = unit(0.1,"line")) 
 p
+
+#ggsave("IAAs.png", width = 7, height = 4)
 ```
-
-Several examples of possible gain-of-function missense variants were found in the degron of *IAA2*, *IAA3*, *IAA6*, *IAA7*, *IAA8*, *IAA13*, *IAA15*, *IAA18*, *IAA19*, *IAA27*, *IAA28* and *IAA29*.
-Most degron variants were very minor alleles in the population, only represented in one or two accessions.
-Interestingly, degron missense variants occured more frequently in *IAA* genes that have a closely related sister, which may indicate redundant function within the sister-pairs.
-**I think this is an interesting finding, but I'm not sure about this hypothesis of redundant function.
-If pairs are acting redundantly and we assume the degron variants are dominant, we would expect the auxin signal they transmit (their function) to be lower, in sum, as a pair.
-So I would actually hypothesize that there has been some divergence in function as a pair.**
-```{r Correlation between Distance to MRCA and prevalence of degron variants}
-
-```
-
-```{r Correlation between Pi_N/Pi_S and k_5}
-
-```
-
-*IAA6* has two potentially dominant missense mutations found within the degron (Supp Fig.
-alignment), one previously believed to lead to the *shy1-1* phenotype [@kim_two_1996].
-Interestingly, this variant is widely represented in 31 accessions across Europe (Supplemental Map Fig).
-This finding supports the proposition that iaa6-C78R may not be the mutation causing the *shy1-1D* phenotype but rather represents natural variation in the degron sequence ([note from Mike Prigge in TAIR](https://www.arabidopsis.org/servlets/TairObject?type=notepad&id=11247)).
-These accessions could be used to confirm whether *iaa6-C78R* is responsible for the *shy1-1D* phenotype.
-Additionally, an accession was identified with the IAA6-V77M variant, also in the degron, that may be useful to further explore *iaa6* gain-of-function phenotypes.
-
-### Identifying novel functional sequences in IAAs
-
-An exciting possible use for the Natural Variation Webtool is to search for conserved, low-diversity sites that may represent possible sites of functional import, such as post-translational modification or protein-protein interaction.
-For example, recent detailed biochemical analysis identified putative lysine ubiquitination sites in members of the *IAA6/19* sister pairs [@winkler_variation_2017]; Supp Fig IAA6/19 alignment).
-These sites are conserved across natural sequence variation, while several other lysine residues vary (Supp Figs alignment and map).
-In IAA6, there are 17 lysines of which ~6 are putative ubiquitination sites~ **I think this should be 5, K3, K32, K33, K91, and K97.
-Missense variants occured in K34Q, K140N *.
-One is proximal to the KR motif, while two others show natural variation.
-In IAA19, there are 15 total lysines, 8 of which are putative ubiquitination sites (K3, K25, K68, L87, K93, K100, K111, and K141).
-Two of these ubiquitination sites are proximal to the KR motif and degron.
-**This sentence is not clear.
-Only 1 of the ubiquitination sites varies?
-or one lysine, I think you mean this only 1 lysine that varies, but it is not ubiquitinated.** Natural sequence diversity may be used to narrow down a subset of lysines to assess for functional ubiquitination.
-
-```{r IAA6/19-alignment, fig.cap="(ref:IAA6/19-alignment"}
-
-```
-
-```{r IAA variants mapped to IAA17, eval=FALSE}
-coding_vcf <- ldply(.data = VCF, .fun = subset, 
-             !is.na(Transcript_ID) & gt_GT != "0|0")
-aln_df <- addSNPsToAlnDF(aln_df, coding_vcf, seq_name = Transcript_ID, 
-               seq_pos = Codon_Number)
-
-selections <- select_variants(gene_info, aln_df, structure_ref = "IAA17")
-selections$V2 <- paste0("chain X and resi ", selections$V2)
-write.table(selections, file = "IAA_variant_selections", row.names = FALSE, 
-          col.names = FALSE, sep = ",", quote = FALSE)
-```
-
-The members of the *IAA8/9/27* clade have a unique N-terminal extension with a highly conserved region common to orthologs in several other species [@dreher_arabidopsis_2006].
-*IAA8* and *IAA9* play redundant roles in leaf serration [@koenig_auxin_2009], while *IAA8* also plays an important role in lateral root development [@arase_iaa8_2012].
-IAA8 has been reported to uniquely localize in the cytosol [@arase_iaa8_2012] where it may interact with LSD1 protein [@coll_programmed_2011].
-The N-terminal extension has low natural sequence diversity in *IAA8* and *IAA9* (Supp Fig IAA8/9 alignment) that may represent a site of novel functionality, possibly mediating interaction with LSD1.
-The few accessions with missense alleles in this region in both *IAA8* and *IAA9* may be valuable in probing the functionality of this region (Supplemental Fig IAA8/9 map).
-
-```{r IAA8_9_27 alignment}
-
-```
-
-```{r IAA8_9 map}
-
-```
-
-Members of the *IAA17/7/14/16* clade have a second EAR motif (LxLxL) with no observed missense variants in *IAA7*, *IAA16*, or *IAA17* (Supp Fig IAA7/14/16/17 alignment).
-This observation supports the experimental evidence for IAA7 that the second EAR motif is functional, *i.e.* it can interact with TPR1 co-repressor and has a minor repressive role in auxin-related developmental phenotypes [lee_biological_2016].
-For *IAA14*, there were two missense variants that may be of interest to probe functionality of this second EAR motif.
-One variant, L41F, in the central lysine is represented in only one accession, and the other in the second “x” (K40N) is found in twelve accessions around Copenhagen (Supp Fig.
-IAA14 map).
-
-```{r IAA7_14_16_17 alignment}
-
-```
-
-```{r IAA14 map}
-
-```

### 2018-12-06 manuscript/Supplement/TPLs.Rmd
bioRxiv submission

--- 2018-10-19 manuscript/05-TPLs.Rmd
@@
 The TOPLESS co-repressor family generally exhibits a high level of sequence conservation at the amino acid sequence level across resequenced *Arabidopsis thaliana* accessions, with all ${\pi_N/\pi_S}$ values below 1 (Table X).
-The closely related *TPL* and *TPR1* have the highest ${\pi_N/\pi_S}$ values (TPL-0.425, TPR1-0.739), suggesting that these these two related genes tolerate a higher degree of sequence and potentially functional diversity compared to *TPR2/3/4*.
+The closely related *TPL* and *TPR1* have the highest ${\pi_N/\pi_S}$ values, suggesting that these these two related genes tolerate a higher degree of sequence and potentially functional diversity compared to *TPR2/3/4*.
 The N-terminal TPL domain of the TPL/TPR family is particularly conserved (Figure X).
@@
 Addtionally, the crystal structure of TPL has revealed that the The LisH domain and CRA domain coordinate TPL dimerization and tetramerization respectively (26601214, 28698367), the CTLH domain is critical for interaction with DNA-binding transcription factors (26601214), and helix nine within the CRA domain may be critical for the formation of higher-order TPL complexes (multimers of TPL tetramers, 28630893).
-
-Compared to the TPL domain, both the proline rich and WD-40 domains exhibited a higher frequency of missense variants.
-We utilized Phyre2.0 modeling software to predict the structure of the TPL/TPR WD-40 repeats in order to map non-synonymous mutations to protein structure (Figure X).
-Here we can see both faces of each predicted beta-propeller structure of the WD-40 domains, which are connected by a short peptide linker.
-This model reveals amino acid conservation in the core and central residues of the WD-40 domains.
-~~These observations could be used as the basis for forward mutagenesis strategies to test whether this region of TPL functions similarly to its homolog TLE, which binds its target genes EH1 and the C-terminal WRPW/Y motif (Hairy/Hes/Runx) in the center of the WD-40 domain (pmid 16762837).~~ **Can we say something more general here?
-There is not enough background to support this.
-What is TLE from?
-What is EH1?
-Is
 
@@
 
-(ref:full-TPLs-alignment) **Alignment of A. thaliana TPL/TPR genes showing variants.** This alignment should probably be paired down to a key region rather than the eniter AA sequence.
-
+(ref:full-TPLs-alignment) **Alignment of A. thaliana TPL/TPR genes** 
 ```{r full-TPLs-alignment, fig.cap= "(ref:full-TPLs-alignment)"}
@@
 ```
-
-Despite the presence of five *TPL/TPR* genes in the genome, which are considered to be redundant (pmid 16763149), very few observed SNPs are predicted to result in loss of function for any *TPL/TPR* gene.
-Only one frameshift variant was observed, *TPR4-665delG*, in a single accession (Aiell-4 from Italy).
-This deletion occurs in codon 222, just after the TPL N-terminal domain, and generates an early stop eleven codons downstream from the deletion.
-It is unclear whether this N-terminal fragment would be expressed at any appreciable level, as it is possible that the generation of an early stop codon would trigger nonsense-mediated decay.
-However, if expressed, it would be fascinating to examine whether this naturally occurring N-terminal truncation retains activity as a co-repressor.
-Given the observation that the TPL N-terminal domain is sufficient to bind histones (28630893), and repress transcription via synthetic assays in yeast (24979769), it is possible that this represents a semi-functional TPL variant.
-There has been no evidence for prioritization of individual *TPL/TPR* family members in the auxin response pathway, but experiments modulating the gene copy number of *TPR1* revealed the importance of maintaining the proper dosage of *TPL/TPR* in order to accurately trigger the salicylic acid innate immune response (20647385).
-These results, taken together with the absence of *TPL/TPR* loss-of-function alleles in the *Arabidopsis thaliana* accessions suggest that perhaps the entire family is being conserved at the nucleotide level in the population to meet a gene dosage requirement, with little opportunity for neo- or sub-functionalization.

### 2018-12-06 manuscript/index.Rmd
bioRxiv submission

@@
 ---
-title: Accelerating structure–function mapping using the ViVa webtool to mine natural variation 
-author:  |
-  | R. Clay Wright^[Corresponding author: wrightrc@vt.edu] $^1$, 
-  | Morgan O. Hamm $^2$,
-  | Jennifer Nemhauser $^2$
-  | $^1$Biological Systems Engineering, Virginia Tech, Blacksburg, VA 24061, 
-  | $^2$Deparment of Biology, University of Washington, Seattle, WA 98195
+title: Accelerating structure-function mapping using the ViVa webtool to mine natural variation 
+shorttitle: 'ViVa: Visualizing Variation' 
+author:
+  - Morgan O.
+Hamm^1^, Britney L.
+Moss^2^, Alexander R.
+Leydon^1^,  Hardik P.
+Gala^1^,  Amy Lanctot^1^, Romi Ramos^1^, Hannah Klaeser^2^, Andrew C.
+Lemmex^1^, Mollye L.
+Zahler^1^, Jennifer L.
+Nemhauser^1^, R.
+Clay Wright^3^$*$
+  - ^1^Department of Biology, University of Washington, Seattle, WA, USA.
+  - ^2^Department of Biology, Whitman College, Walla Walla, WA, USA. 
+  - ^3^Biological Systems Engineering, Virginia Tech, Blacksburg, VA, USA. Previous affiliation$:$ Department of Biology, University of Washington, Seattle, WA, USA.
+  - $*$Direct correspondence to$:$ wrightrc@vt.edu.
 
 date: "`r Sys.Date()`"
-
-abstract: Thousands of sequenced genomes are now publically available and represent a large swath of plant evolution; yet, much of this data remains inaccessible to researchers without significant bioinformatics experience.
-Here, we present a webtool called ViVa (Visualizing Variation) which aims to empower any researcher to take advantage of the amazing genetic resource collected in the Arabidopsis 1001 Genomes Project (http://1001genomes.org).
-ViVa facilates data mining on the gene, gene family or gene network level.
-To test the utility and accessibility of ViVa, we assembled a team with a range of expertise with biology and bioinformatics to analyze the natural variation in the well-studied nuclear auxin signaling pathway.
+abstract: "Thousands of sequenced genomes are now publicly available capturing a significant amount of natural variation within plant species; yet, much of this data remains inaccessible to researchers without significant bioinformatics experience.
+Here, we present a webtool called ViVa (Visualizing Variation) which aims to empower any researcher to take advantage of the amazing genetic resource collected in the *Arabidopsis thaliana* 1001 Genomes Project (http://1001genomes.org).
+ViVa facilitates data mining on the gene, gene family or gene network level.
+To test the utility and accessibility of ViVa, we assembled a team with a range of expertise within biology and bioinformatics to analyze the natural variation within the well-studied nuclear auxin signaling pathway.
 Our analysis has provided further confirmation of existing knowledge and has also helped generate new hypotheses regarding this well studied pathway.
 These results highlight how natural variation could be used to generate and test hypotheses about less studied gene families and networks, especially when paired with biochemical and genetic characterization.
-ViVa is also readily extensible to newly emerging databases of genetic variation, such as the 3,000 Rice Genomes Project (http://snp-seek.irri.org/).
+ViVa is also readily extensible to databases of interspecific genetic variation in plants as well as other organisms, such as the 3,000 Rice Genomes Project (http://snp-seek.irri.org/) and human genetic variation (https://www.ncbi.nlm.nih.gov/clinvar/).
 
-keywords: natural variation, Arabidopsis thaliana, genome diversity, structure, function, genotype, phenotype, bioconductor, bookdown, r markdown
+Keywords: natural variation, Arabidopsis thaliana, genome diversity, structure, function, genotype, phenotype, bioconductor, bookdown, r
+
+"
 
@@
 - natural-variation.bib 
-- TIR1_AFB.bib
-- Aux_IAA.bib
-- ARF.bib
-- TPL.bib
+- packages.bib
+
 biblio-style: apalike
 link-citations: yes
-
+fig_caption: yes
 ---
@@
 ```{r setup, message=FALSE, warning=FALSE, include=FALSE}
-knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE, collapse = TRUE, comment = "#>", tidy = TRUE, cache = TRUE, cache.lazy = FALSE, out.width = "90%", fig.height=4, fig.width=7, out.extra='style="margin: auto; display: block; padding-top: 15px;"')
+knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE, collapse = TRUE, comment = "#>", tidy = TRUE, cache = TRUE, cache.lazy = FALSE, fig.height=4, fig.width=7, dpi = 600, out.extra='style="margin: auto; display: block; padding-top: 15px;"')
 
@@
 library(stringr)
+library(treeio)
 
@@
 ```
-
-## List of working titles  
-* Predicting structure--function relationships using natural genetic variation 
-* Accelerating structure--function mapping using natural genetic variation 
-* A comprehensive analysis of natural sequence variation within the <i>Arabidopsis thaliana</i> nuclear auxin signaling pathway 
-* Visualizing natural variation with ViVa, a new webtool
-* Accelerating structure–function mapping using a new webtool to mine natural variation
 
@@
 ```
+
+\newpage

### 2019-01-10 manuscript/05-acknowledgments.Rmd
submitted

@@
 The authors would like to thank Oghenemega Okolo for assistance testing the ViVa software, and Song Li and Bo Zhang for helpful comments on the manuscript.
-This work was supported by the National Institute of Health (R01-GM107084) and the Howard Hughes Medical Institute.
+This work was supported by the National Institute of Health (R01-GM107084), the National Science Foundation (IOS-1546873) and the Howard Hughes Medical Institute.
 R.C.W.
-received fellowship support from the National Science Foundation
-(DBI-1402222). B.L.M. and H.K. received support from the M.J. Murdock Charitable Trust. A.R.L. is a Simons Foundation Fellow of the Life Sciences Research Foundation.
+received fellowship support from the National Science Foundation (DBI-1402222).
+B.L.M.
+and H.K.
+received support from the M.J.
+Murdock Charitable Trust.
+A.R.L.
+is a Simons Foundation Fellow of the Life Sciences Research Foundation.
+A.L.
+was supported by an NSF Graduate Research Fellowship DGE-1256082.

### 2019-01-10 manuscript/index.Rmd
submitted

@@
 Leydon^1^,  Hardik P.
-Gala^1^,  Amy Lanctot^1^, Romi Ramos^1^, Hannah Klaeser^2^, Andrew C.
+Gala^1^,  Amy Lanctot^1^, Román Ramos^1^, Hannah Klaeser^2^, Andrew C.
 Lemmex^1^, Mollye L.
@@
 ```{r setup, message=FALSE, warning=FALSE, include=FALSE}
-knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE, collapse = TRUE, comment = "#>", tidy = TRUE, cache = TRUE, cache.lazy = FALSE, fig.height=4, fig.width=7, dpi = 600, out.extra='style="margin: auto; display: block; padding-top: 15px;"')
+knitr::opts_chunk$set(
+  warning = FALSE, 
+  message = FALSE, 
+  echo = FALSE,
+  collapse = TRUE, 
+  comment = "#>", 
+  tidy = TRUE, 
+  cache = TRUE, 
+  cache.lazy = FALSE, 
+  fig.height=4, 
+  fig.width=7, 
+  dpi = 600, 
+  out.extra='style="margin: auto; display: block; padding-top: 15px;"')
 

### 2019-04-02 _main.Rmd
revisions done, back to coauthors

--- 2018-01-05 example.Rmd
@@
 <!--=========================================================================-->
+
+<!--chapter:end:example.Rmd-->
+
+---
+title: "Collaborative analysis and writing in R"
+author: "R Clay Wright"
+date: "12/28/2017"
+output: 
+  html_document:
+    toc: true
+
+---
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
+```
+
+<!--This is a CSS (Cascading Style Sheet).
+It controls how panels and images are formatted.></!-->
+<style>
+  img {
+    display: block;
+    margin: auto;
+    width: 90%;
+    height: auto;
+
+  }
+</style>
+
+This tutorial will (hopefully) gently guide you through how to use the tools docker and Rmarkdown to analyze the natural variation in your set of genes and collaborate on our manuscript, tentatively titled: *A comprehensive analysis of natural sequence variation within the Arabidopsis thaliana nuclear auxin response pathway* (running title: "natural variation in the nuclear auxin response").
+I assume that you have a basic working knowledge of how to download and install software as well as R and RStudio,
+but if you feel you could use a refresher check out [this introduction](https://github.com/wrightrc/Intro_to_R_RStudio_and_ANOVA).
+We'll use docker to share with you a version of R and Rstudio that already has all of the necessary packages to do the analysis.
+We'll use Rmarkdown to write the manuscript and make figures and tables by integrating prose, R code and results from the code.
+
+So in this tutorial I will guide you through:  
+1. Installing docker  
+2. Starting RStudio and r1001genomes in docker  
+3. Working in docker/RStudio  
+4. Saving your work in docker/RStudio  
+5. Using Rmarkdown to format text  
+6. Using Rmarkdown to insert R code and results into text  
+7. Citations in Rmarkdown
+
+If you have any troubles with anything, find something unclear or even find a typo please email me and/or make the change in your version of these documents and I will integrate them in the next version.
+
+Ready to get started?
+
+## Docker
+
+Docker is a program that allows software configurations to be shared across computers and time.
+We'll use docker to share the RStudio configuration that will allow you to run the `r1001genomes` package and shiny app without having to upgrade your computer, your R and Rstudio or deal with installing the packages `r1001genomes` depends on.
+Docker is kind of like a virtual machine that you may have used in the past to run Windows on your Mac or *visa versa*, but docker is better, faster, stronger than virtual machines.
+
+First things first, let's install docker!
+
+### Installing docker
+
+Docker requires a 64bit processor and operating system.
+For Macs, you can check your processor by clicking the "apple" in the menu bar and selecting "About this Mac".
+If "Processor" says Core Duo, you unfortunately have a 32 bit processor, otherwise anything else is a new 64bit processor.
+On a machine running Windows, go to the start menu and settings, then click on "System" and finally "About".
+If under "System type" you see "32-bit operating system,..." you can't install docker.
+If you see "32-bit operating system, x64-based processor", you *can* update Windows to 64-bit and then install docker, but this update requires a clean install, meaning all of your files will be deleted.
+
+Anyway, if you have 32-bit woes let me know and we'll come up with a way to allow you to collaborate. I've installed docker on the Macs in the lab so you can always work those.
+
+#### Mac  
+If you are using a 2010 or newer Mac with OS X Yosemite (10.10) or newer you can simply [download the Community Edition of docker](https://download.docker.com/mac/stable/Docker.dmg)(click the link to begin download) and install it.
+
+If you are using an older Mac or OS you will need to install the older [docker Toolbox](https://download.docker.com/mac/stable/DockerToolbox.pkg)(click the link to begin download).
+[Follow these directions](https://docs.docker.com/toolbox/toolbox_install_mac/) to help you get everything installed.
+
+#### Windows  
+If you are using Windows 10 Pro, Enterprise or Education you can simply [download the Community Edition of docker](https://download.docker.com/win/stable/Docker%20for%20Windows%20Installer.exe)(click the link to begin download) and install it.
+
+If you are using an older version of Windows you will need to install the older [docker Toolbox](https://download.docker.com/mac/stable/DockerToolbox.pkg)(click the link to begin download).
+[Follow these directions](https://docs.docker.com/toolbox/toolbox_install_windows/) to help you get everything installed.
+
+### Using docker
+
+Before we jump in let's learn some docker vocabulary.
+The initial software setup that is a stable snapshot and can be shared is called an "image".
+Whenever you run that image, you create a "container", which is just an instance of that container.
+From the docker website:
+
+>An **image** is a lightweight, stand-alone, executable package that includes everything needed to run a piece of software, including the code, a runtime, libraries, environment variables, and config files.
+>  
+>A **container** is a runtime instance of an image—what the image becomes in memory when actually executed.
+It runs completely isolated from the host environment by default, only accessing host files and ports if configured to do so.
+
+So I have shared with you (and the rest of the world actually) the docker image that will allow you to easily run the r1001genomes package and Shiny app.
+When you run that image you will create your own container of that image that you can do work in, change settings, build on, save or erase.
+
+Docker has recently built a graphical interface for managing images and containers called kitematic.
+We'll use this to simplify setting up your container from the image and allowing you to transfer files to and from the container.
+
+#### Installing kitematic
+
+If you click on the docker icon in the menu bar you will see kitematic listed as an option.  
+![](TUTORIAL_screenshots/Screenshot 2017-12-29 16.01.58.png)  
+Select "Kitematic" to open a window with a link to download and install kitematic. Once you have installed kitematic selecting the "Kitematic" link will open the kitematic interface.  
+![](TUTORIAL_screenshots/Screenshot 2017-12-29 16.50.28.png)  
+The containers that you currently have running on your system are listed in the left panel.
+You can see I have a "nuclear-auxin-natural-variation" container running, that docker has given the creative name "distracted-nobel".
+In the main pane of kitematic there are several popular images that we can create new containers from.
+
+The image that we are going to work with here is called `r1001genomes`.
+Docker images build on previous images (like all programming does, and all science does, for that matter).
+I built the `r1001genomes` image on top of a previous image called `verse` from the `rocker` project.
+(The rocker project and R Open Science also have a [great docker tutorial](http://ropenscilabs.github.io/r-docker-tutorial/).) The `verse` image contains R, Rstudio, the [tidyverse](www.tidyverse.org) packages, Rmarkdown and all of the dependencies for these things.
+To this, I added  the necessary packages that `r1001genomes` builds from and `r1001genomes`.
+Let's make a new container for this image!
+
+You can search in the box at the top of kitematic for "r1001genomes", then click "create" to download the image and create a container.
+This will take a **while**.
+You are downloading linux, R, Rstudio and all of those packages.
+Fortunately docker will reuse all of these components whenever you download a related image.
+![](TUTORIAL_screenshots/Screenshot 2017-12-29 21.08.01.png)
+
+Once the image is downloaded and the container is created you'll see a log and some other information "IP & PORTS" and "VOLUMES".  
+![](TUTORIAL_screenshots/Screenshot 2017-12-29 21.13.24.png)
+
+As it says: "You can access this container using the following IP address and port".
+This means that if you point your browser to the "ACCESS URL" you will see the version of Rstudio that we've built.
+The username and password are both "rstudio".
+
+Go ahead give it a try!  
+![](TUTORIAL_screenshots/Screenshot 2017-12-29 21.34.22.png)  
+Now we need some data to play with in our Rstudio container!
+You'll notice that the only folder available in the "Files" pane is "kitematic", which matches the folder in the "VOLUMES" pane back in kitematic.
+To make use of this folder we need to map it to a folder on our computer.
+Click on the "gear" settings icon in the "VOLUMES" pane, then click "CHANGE" and select a folder to be able to access it within your container.
+Go ahead and select the "natural-variation-webtool" folder that this tutorial is in.
+
+Kitematic will then restart the container and the access URL might be changed.
+So go back to the "Home" tab and paste the new URL into your browser or, if the access URL didn't change just reload the page.
+
+Now when you click on the "kitematic" folder you will see the contents of this folder.
+Any files you save in this folder will be saved to the folder you mapped on your computer.
+You can also open files from this folder in RStudio.
+You can save your data and the state of your RStudio session into this folder and it will stay on your computer.
+If you quit your browser you can navigate back to the same access url, because the docker container will be accessible unless you stop it in kitematic.
+
+In the RStudio "Files" pane, navigate to the "manuscript" folder and set it as your working directory (using the "More" button). Click on "manuscript.Rmd" to open the manuscript markdown file.
+
+Time for some R markdown!
+
+## R Markdown
+
+The idea behind the creation of markdown and R markdown was to make html (hyper-text markUP language) more low key.
+So the goals of markDOWN are to be easy to write and easy to read but to still have a lot of the powers of HTML (text formatting, layout and templating).
+R Markdown also allows us to generate PDF, Word, and Presentation documents in addition to HTML.
+
+This document was is actually written in R Markdown.
+If you haven't already, open the "code" markdown version of this document ("TUTORIAL.Rmd") in Rstudio, and navigate down to the "R Markdown" section using the navigation bar at the bottom of the "Source" pane (see the figure).
+![](TUTORIAL_screenshots/Screenshot 2017-12-29 22.45.46.png)
+
+You can follow along with the code and the document side-by-side to learn how markdown text formatting works. Here's some example markdown from the R markdown file template.
+
+Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
+
+When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
+You can embed an R code chunk like this:
+
+```{r cars}
+summary(cars)
+```
+
+You can insert code chunks easily using the "Insert" button at the top of the source window, Code>Insert Chunk in the menu bar or hotkeys specified next to this menu bar command.
+
+### Including Plots and Code
+
+You can also embed plots, for example:
+
+```{r pressure, echo=FALSE}
+plot(pressure)
+```
+
+Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot (check out the markdown code in TUTORIAL.Rmd).
+
+You can also incorporate results from code chunks in your prose. First, we need a result .
+
+```{r}
+set.seed(42)
+group1 <- rnorm(n = 10, mean = 1, sd = 0.5)
+group2 <- rnorm(n = 10, mean = 2, sd = 0.5)
+t_test <- t.test(x = group1, y = group2)
+t_test
+```
+
+Now we can refer to these results in our code by creating an inline code chunk with syntax `r #code`.
+So we could say that we have some support that the means of the two groups we simulated are different (`r t_test$method`, t-value = `r t_test$statistic`, p = `r t_test$p.value`)
+
+Don't get bogged down in formatting the text, but I would highly recommend checking out the <http://rmarkdown.rstudio.com> link and don't miss the [cheatsheet](http://rmarkdown.rstudio.com/lesson-15.html)!
+You can also find this and other cheatsheets and a Markdown reference in the RStudio "Help" menu.
+![](TUTORIAL_screenshots/Screenshot 2017-12-29 22.57.15.png)
+
+A few tips to help your R Markdown experience stay low key:  
+1. Turn on whitespace characters so you know where spaces, tabs and returns are. This is in Preferences>Code>Display.  
+2. Include 2 spaces at the end of each paragraph or whenever you want to start a new line (like in this numbered list).  
+3.
+Use and empty line between paragraphs, headings, offset quotes and any other sort of formatted elements like lists, tables and code chunks.
+This offset helps improve readability and makes sure that these elements are formatted the way you want them.
+
+**However, keep in mind your foremost focus should be on the content. We'll worry about formatting in revisions.**
+
+In the example analysis you can also see some examples of how to use R markdown, DT (data table) and kable to format and insert tables into our document.
+But, one last thing before you move on to the example analysis and your own analysis...
+
+### Citations in R markdown
+
+You can include citations in markdown documents using the following syntax: `[@citation_key]`.
+The citation key refers to a unique identifier generated when exporting a bibliography from a reference manager.
+For the manuscript you will need to export the references from your reference editor in one of many formats, but I would recommend BibTex (.bib).
+For other possible formats see [the markdown documentation](http://rmarkdown.rstudio.com/authoring_pandoc_markdown.html#citations).
+
+Let's take a look at an example, open the "references.bib" file.
+You'll notice each reference begins with an @-symbol and the type of reference.
+Then within curly braces are all of the data about the reference.
+The first item is the citation key, which in this case is "gray_identification_1999".
+So to cite this paper in markdown you would add `[@gray_identification_1999]` to the end of a sentence.
+My process for citing-while-writing is to first write, then when I want to cite something, find the paper in my reference manager and copy it to a folder for the project I'm working on.
+This will allow me to export the folder as a .bib file later).
+Then I copy the citation key for the reference and paste it into my writing.
+When I'm ready to knit the document, I'll export the folder of references as a .bib (with Better Bibtex you can sync this file so that it is automatically updated).
+This bibliography file must be in the same folder with the .Rmd file, along with a .csl file that specifies the format of the citations and bibliography in the final document (available for most journals [via Zotero](https://www.zotero.org/styles)).
+These files also must be specified in the header of the document.
+![](TUTORIAL_screenshots/Screenshot 2017-12-30 14.04.14.png)
+
+Most common reference managers allow you to export a bibliography as a bibtex file, and copy and paste citation keys, however I would highly recommend downloading [Zotero](www.zotero.org) and importing your reference library from whatever reference editor you currently use.
+Zotero is free, open-source and integrates very well with all popular web browsers and writing tools.
+If you want to work in a different citation manager, here are some links to help you create and copy and paste citation keys as well as export BibTex bibliography files.
+Note that most of these tutorials are for Latex, which uses a slightly different syntax from Rmarkdown.
+In Rmarkdown insert citations via `[@citation_key]`.
+The `citation_key` string does have some stipulations,
+
+>The citation key must begin with a letter, digit, or \_, and may contain alphanumerics, _, and internal punctuation characters (:.#$%&-+?<>~/).
+
+This can cause some issues with BibTex files exported from some reference managers, that is why I recommend Zotero with the Better BibTex add-on.
+
+*To insert multiple citations separate the two citations by semicolons, e.g.* `Blah blah [@smith04; @doe99].`
+
+[Endnote](https://www.reed.edu/cis/help/LaTeX/EndNote.html): Export as [bibtex with labels](http://www.endnote.com/downloads/style/bibtex-export-using-en-label-field).  
+Endnote seems to have some issues generating unique citation keys for each reference.
+It might be best to export your library as an Endnote XML and import the whole thing into Zotero, if you have issues with the above solutions.
+[Mendeley](https://blog.mendeley.com/2011/10/25/howto-use-mendeley-to-create-citations-using-latex-and-bibtex/)  
+[Zotero](https://github.com/bbest/rmarkdown-example#writing-with-rmarkdown)
+
+Once you are done writing and citing and have exported your BibTex file, you can add your bibliography file to the manuscript folder and add the name of your file to the document header (comma separated).
+This should allow your version of the manuscript to knit, with perhaps some errors if there are duplicate citation keys.
+Don't worry we'll sort out duplicates later, once I have merged everyones sections' together.
+
+#### Also print out your key citations and we will make a binder for everyones use in the computer room! Also don't forget to send me any changes or suggestions you have for this tutorial.
+
+Now you are ready to work through example.Rmd (and/or example.html) and begin thinking about and analyzing natural variation in your gene family!
+
+Thanks and happy collaborative writing!
+
+<!--chapter:end:TUTORIAL.Rmd-->

### 2019-04-02 manuscript/01-intro.Rmd
revisions done, back to coauthors

@@
 
-Beyond its utility in gene discovery and understanding genome evolution, natural genetic variation provides a catalog of permissible polymorphisms that can facilitate the connection of genotype to phenotype at the gene, gene family and network scales [@joly-lopez_developing_2016].
-This is an especially critical resource in large gene families where loss of function in individual genes may have little or no phenotypic effect and directed allele replacement remains time and resource-intensive.
-Massively parallel assays of variant effects in human clinical medicine stand to revolutionize genetic diagnostics and personalized medicine [@starita_variant_2017; @gasperini_power_2016; @matreyek_platform_2017].
-We envision the use of plant natural variation datasets as a tool to similarly revolutionize breeding and genetic engineering of crop plants by rapidly advancing our understanding of genotype/function/phenotype relationships.
+Beyond its utility in gene discovery and understanding genome evolution, natural genetic variation provides a catalog of permissible polymorphisms that can facilitate the connection of genotype to phenotype at the gene, gene family and network scales [@nieduszynski_sequence_2011; @joly-lopez_developing_2016].
+This is an especially critical resource in large gene families where loss of function in individual genes may have little or no phenotypic effect [@dharmasiri_plant_2005; @moore_evolutionary_2005; @guo_gene_2013] and directed allele replacement remains time and resource-intensive [@chen_crispr_2019].
+Natural variation datasets provide novel alleles and germplasm which can be examined with biochemical and genetic approaches to map sequence to function and genotype to phenotype.
+In human clinical medicine, massively parallel assays of variant effects  stand to revolutionize genetic diagnostics and personalized medicine [@starita_variant_2017; @gasperini_power_2016; @matreyek_platform_2017].
+Similarly, we envision the use of plant natural variation datasets as a tool to revolutionize breeding and genetic engineering of crop plants by rapidly advancing our understanding of genotype/function/phenotype relationships.
 A proof-of-principle survey of a relatively small subset of natural variants paired with a synthetic assay of gene function successfully mapped critical functional domains of auxin receptors and identified new alleles which affect plant phenotype [@wright_insights_2017].
@@
 One reason may be the current requirement for a fairly high level of bioinformatics expertise to extract the desired information from whole genome resequencing datasets.
-While existing resources such as the 1001 Proteomes website [@joshi_1001_2012] and ePlant [@waese_eplant_2017] facilitate access to this data at the gene scale, they do not provide summaries or visualizations of variation at the gene family and network scales.
+While existing resources such as the 1001 Proteomes website [@joshi_1001_2012] and ePlant [@waese_eplant_2017] facilitate access to this data at the gene scale, they do not provide comparative summaries or visualizations of variation at the gene family or network scales.
 To address this concern, we created ViVa: a webtool and R-package for **Vi**sualizing **Va**riation, which allows plant molecular biologists of any level access to gene-level data from the 1001 Genomes database.
-Using ViVa researchers may: 1) Identify polymorphisms to facilitate biochemical assays of variant effects [@starita_variant_2017; @wright_insights_2017]; 2) Produce family-wise alignments of variants to facilitate *de novo* functional domain identification [@melamed_combining_2015]; 3) Generate lists of accessions containing polymorphisms to facilitate phenotypic analysis of gene variant effects [@park_distributions_2017]; and 4) Quantify metrics of genetic diversity to facilitate the study of gene, gene family and network evolution [@delker_natural_2010; @kliebenstein_role_2008].
+Using ViVa researchers may: 1) Identify polymorphisms to facilitate biochemical assays of variant effects [@starita_variant_2017; @wright_insights_2017]; 2) Produce family-wise alignments of variants to facilitate *de novo* functional domain identification [@melamed_combining_2015]; 3) Generate lists of accessions containing polymorphisms to facilitate phenotypic analysis of gene variant effects [@park_distributions_2017]; and 4) Quantify metrics of genetic diversity to facilitate the study of gene, gene family, and network evolution [@delker_natural_2010; @kliebenstein_role_2008].
+Here we present a summary of the functionality of ViVa along with an analysis of the natural variation in the nuclear auxin signaling pathway using ViVa.
+To demonstrate ViVa usage we focus on the analysis of the *Aux/IAA* family and reserve the remainder of our analysis for the interested reader as supplemental material.

### 2019-04-02 manuscript/02-results.Rmd
revisions done, back to coauthors

@@
 
-ViVa, in this first iteration, is meant to visualize natural variation in the coding sequences of genes.
+ViVa, in this first iteration, is meant to visualize natural variation in the coding sequences of genes or gene families.
 Non-coding sequence variation is intentionally excluded from the analysis tools.
-This reflects challenges both in alignment of non-coding sequences and the increased difficulty in assessing variation in these regions [@alexandre_complex_2018].
+This reflects challenges both in alignment of non-coding sequences and the increased difficulty in assessing the effects of variation in these regions [@alexandre_complex_2018].
+These analyses can be conjoined, as was done here, to understand variation within gene networks.
+Visualizations of variation within whole gene networks are planned for future iterations.
 
@@
 (e) The SNP Browser tab allows variants and accessions to be filtered by any combination of text and numeric fields.
-(f) The Alignments tab aligns DNA and amino acid sequences of homologous genes, and colors sequence elements based on SNPs and annotations.
+(f) The Alignments tab aligns DNA and amino acid sequences of homologous genes and colors sequence elements based on SNPs and annotations.
 ```{r app-figure-elements, fig.cap = "(ref:app-figure-elements)"}
@@
 
+(ref:app-workflow) **ViVa workflow** (a) Workflow diagram of ViVa.
+Blue indicates user actions, orange indicates processing steps by the application.
+(b) Detailed look at mapping tab parameters: User selects which genes to look at (1) then clicks the submit button (2).
+The Allele selection pane is then filled in with all non-reference variants meeting the criteria.
+The user can then adjust the range of nucleotide diversity and the type of SNP with a slider and radio buttons (3) to make the list of variants a manageable size.
+The list of variants will update as changes are made to these controls.
+The user then selects variants to display on the map from the allele selection pane (4).
+Clicking the update map button (5) will populate the map below with points located at the collection coordinates of each accession, colored by the selected non-reference alleles that are present in the sample.
+(c) Detailed look at the browser tab options: The user first selects which genes analyze and clicks submit (1).
+The “hide 0|0 genotype?” checkbox (2) removes rows from the table containing the reference allele.
+Four configurable filters of two types are provided.
+The first filter type is text matching (3); the user selects a column to filter from a drop-down menu then enters one or more text strings to match in that column.
+The second type of filter is numeric range matching (4); the user again selects a column, then specifies a range to match by typing or selecting minimum and maximum values.
+After configuring the filters, the user then clicks the apply filters button (5) to update the table.
+The table will be updated such that only rows that meet all filter conditions are displayed.
+```{r app-workflow, fig.cap = "(ref:app-workflow)"}
+knitr::include_graphics("app_figure/workflow_fig.png")
+```
+
 ### Gene Select and Annotation Files  
-At the top of the ViVa webtool are two collapsible panels used for entering the genes to query and custom annotations for those genes (Figure \@ref(fig:app-figure-elements)a).
+At the top of the ViVa webtool are two collapsible panels used for entering the genes to query and custom annotations for those genes (Figure \@ref(fig:app-figure-elements)a and \@ref(fig:app-workflow)a).
 The Gene Select panel permits gene input by either typing in the genes' AGI/TAIR locus identifiers, or uploading a .csv file.
@@
 
-Below the data input section, the rest of the webtool is divided into several tabs containing the interactive output of the program.
+Below the data input section, the rest of the webtool is divided into several tabs containing interactive output.
 
@@
 The SNP Stats tab provides general information on the gene transcripts being queried, as well as calculated counts/statistics on the content of variants found in the sample population (Figure \@ref(fig:app-figure-elements)b).
-The first table of this section is basic information about the transcripts, including TAIR locus and symbol, the chromosomal start and end position and the transcript length.
+The first table of this section is basic information about the transcripts, including TAIR locus and symbol, the chromosomal start and end position, and the transcript length.
 This information was collected from the Araport11 Official Release (06/2016) annotation dataset [@cheng_araport11_2017].
@@
 The next two tables provide counts of SNPs across the gene body for each transcript.
-The Total Polymorphism Counts table provides the total number of observations of non-reference allele counts of each variant type (the Col-0 accession TAIR9 genome is the reference genome for this dataset).
-The Unique Allele Count table only counts the number of unique variants and alleles within the population of accessions (*e.g.* if multiple accessions have the same variant allele, that allele will only be counted once).
+The Total Polymorphism Counts table provides the total number of variant observations of non-reference alleles categorized by type and location (the Col-0 accession TAIR9 genome is the reference genome for this dataset).
+The Unique Allele Count table only counts the number of unique variant alleles within the population of accessions (*e.g.* if multiple accessions have the same variant, these will be counted as a single allele at that position).
 
 The Nucleotide Diversity Statistics table provides a nucleotide diversity statistic (${\pi}$) for the transcript and the coding sequence of each gene [@nei_mathematical_1979].
-Nucleotide diversity is also calculated for the set of only synonymous (${\pi_S}$) and only non-synonymous sites (${\pi_N}$).
-The ratio of the presence of non-synonymous to synonymous polymorphism provides a measure of the potential for functional diversity [@firnberg_genetic_2013; @whitehead_optimization_2012].
-We present ${\pi_N/\pi_S}$ here as an correlate for functional diversity throughout ViVa [@nelson_snpgenie_2015; @hughes_adaptive_2000].
+Nucleotide diversity is also calculated for the set of only synonymous (${\pi_S}$) and only nonsynonymous sites (${\pi_N}$).
+The ratio of the presence of nonsynonymous to synonymous polymorphism provides a measure of the potential for functional diversity [@firnberg_genetic_2013; @whitehead_optimization_2012].
+We present ${\pi_N/\pi_S}$, the ratio of nonsynonymous to synonymous diversity here as a correlate for functional diversity throughout ViVa [@nelson_snpgenie_2015; @hughes_adaptive_2000].
 While imperfect, this metric may be suggestive of functional constraint when ${\pi_N/\pi_S << 1}$ and functional diversity when ${\pi_N/\pi_S >> 1}$ [@hughes_adaptive_1999].
@@
 ### Diversity plot: Visualize allelic diversity across the coding sequence
-The Diversity Plot tab shows the nucleotide diversity of each variant in the coding region of a selected gene (Figure \@ref(fig:app-figure-elements)c).
+The Diversity Plot tab shows the nucleotide diversity of each variant in the coding region of a selected gene (Figure \@ref(fig:app-figure-elements)c and \@ref(fig:IAA6-div-plot)).
 Although the X-axis is marked by codon number from the N-terminus for interpretability, the diversity values are based on single nucleotide sites.
@@
 ### SNP Mapping: View distributions of SNPs across the globe
-The SNP Mapping tab plots the accessions collection locations on a world map, and colors the points based on selected variant alleles (Figure \@ref(fig:app-figure-elements)d).
-After selecting the genes and filtering on the SNP type and level of nucleotide diversity, a group of checkboxes becomes available to select variant alleles to display on the map.
+The SNP Mapping tab plots the accessions collection locations on a world map and colors the points based on selected variant alleles (Figure \@ref(fig:app-figure-elements)d).
+After selecting the genes and filtering on the SNP type and level of nucleotide diversity, a group of checkboxes becomes available to select variant alleles to display on the map (Figure \@ref(fig:app-workflow)b and \@ref(fig:AFB1-map)).
 The variant alleles are labeled with the Transcript_ID and Amino_Acid_Change fields, in the form [Transcript_ID|Amino_Acid_Change].
@@
 gene name or variant effect), or set minimum and maximum limits on the values of numeric fields (e.g.
-nucleotide diversity).
+nucleotide diversity) (\@ref(fig:app-workflow)c).
 When these filters are applied, the table below will update to only contain rows meeting the criteria for all filters.
@@
 ### Alignments: Visualize SNPs on alignments of homologous genes
-The Alignments tab provides DNA and amino acid sequence alignments of selected genes, colored according to the variant allele with the strongest functional effect at each position (Figure \@ref(fig:app-figure-elements)f and \@ref(fig:A-IAA-alignment)).
+The Alignments tab provides DNA and amino acid sequence alignments of selected genes, colored according to the variant allele with the strongest functional effect at each position (Figure \@ref(fig:app-figure-elements)f and \@ref(fig:A-IAA-alignment), for description of color scale see Figure \@ref(fig:SNPeff-order)).
 The content of this tab is most useful if the selected genes are all family members or have significant sequence homology.
-If annotation files are uploaded, the background of the sequences will be colored by annotation region.
+If annotation files are uploaded, open boxes will be overlayed on the alignment, colored by annotation.
 Hovering the cursor over variants will provide additional details about the alleles present at that locus.
@@
 ### ViVa R package: Programmatic access to ViVa's functionality
-All of the functionalities of the ViVa webtool are implemented through function calls to the ViVa R package.
+All of the functionalities of the ViVa webtool are implemented through functions within the ViVa R package.
 In addition to being able to generate the same sets of figures and tables as in the webtool, users of the R package also gain direct access to the underlying data structures, providing greater control over parameters when processing and visualizing the data.
+The ViVa R package is intended for users familiar with R programming who want to extend the capabilities of the webtool.
 The ViVa R package can be found at https://github.com/wrightrc/r1001genomes and can be installed in your R environment via the `devtools` package:
@@
 We selected this signaling pathway for multiple reasons including a wealth of functional data and solved structures of several domains or entire proteins.
-Using this exitsting knowledge, we were able to qualitatively assess the predictive ability of the ViVa modules.
-Summary information about each nuclear auxin signaling gene family examined can be found in the [Supplemental Data](#si).
-Below, we describe the results for the *Aux/IAA* family in more detail.
+Using this existing knowledge, we were able to qualitatively assess the predictive ability of the ViVa modules.
+Below, we describe the results for the *Aux/IAA* family in more detail as a brief vignette of ViVa use.
+Analysis of each gene family in the nuclear auxin signaling network can be found in the [Supplemental Analysis](#si).
 
-In most cases, natural selection is expected to minimize the persistence of nonsynonymous mutations in sequences that encode critical functional domains relative to their persistence in non-critical domains [@hughes_adaptive_2000].
+The Aux/IAA family plays a critical role in transmitting auxin signals, as their degradation is triggered by auxin accumulation.
+This degradation is in response to ubiquitination via a SKP-Cullin-F-box ubitquitin ligase complex containing an Auxin-signaling F-box (AFB) auxin receptor protein.
+Variation in the *AFB* family is presented in [Supplemental Analysis](#si) Figures \@ref(fig:AFBtree), \@ref(fig:full-AFBs-alignment), and \@ref(fig:AFB1-map)).
+The Aux/IAAs repress transcription of Auxin Response Factor (ARF) bound genes, via recruitment of TOPLESS (TPL) and TOPLESS-related (TPR) co-repressors.
+Variation in the *TPL/TPR* family is presented in [Supplemental Analysis](#si) Figures \@ref(fig:TPLtree) and \@ref(fig:full-TPLs-alignment)).
+Thus Aux/IAA degradation, relieves this repression allowing ARFs to activate transcription of auxin response genes.
+Variation in the *ARF* family is presented in [Supplemental Analysis](#si) Figures \@ref(fig:ARFtree) and \@ref(fig:full-ARFs-alignment)).
+
+To predict the functional impact of variation in coding sequences of genes ViVa uses the frequency of nonsynonymous and synonymous polymorphisms.
+In most cases, nonsynonymous polymorphisms in critical functional domains are likely to have deleterious effects on gene function [@hughes_adaptive_2000].
+This will result in less accumulation of nonsynonymous polymorphism in domains which are critical to plant fitness.
 Therefore, we reasoned that scanning gene coding sequences for regions of relatively low nonsynonymous diversity should highlight functional domains.
 This general principle can be seen clearly in the analysis of the *Aux/IAA* family of transcriptional co-repressors/co-receptors.
-Aux/IAAs have three major domains.
-Domain I contains an EAR motif that facilitates interaction with TOPLESS (TPL) and TOPLESS-related (TPR) transcriptional repressors [@tiwari_aux_2004; @szemenyei_topless_2008].
+Most, Aux/IAAs have three major domains.
+Domain I contains an EAR motif that facilitates interaction with TPL/TPR transcriptional co-repressors [@tiwari_aux_2004; @szemenyei_topless_2008].
 Domain II, the degron, facilitates interactions with the TIR1/AFB receptors in the presence of auxin [@tan_mechanism_2007].
 Domain III (which was originally considered domains III and IV) is a PB1 domain and facilitates interactions with the ARF transcription factors [@ulmasov_aux_1997; @guilfoyle_getting_2012; @nanao_structural_2014; @korasick_molecular_2014].
-The EAR motif and degron can be readily identified by the drop in nonsynonymous variation, as visualized by the lack of stronger functional effects in Figure \@ref(fig:A-IAA-alignment).
+
+Here, we analyze just the most conserved A class Aux/IAAs, which contain all of these domains, by mapping the natural genetic variation onto an alignment of the coding sequences via the ViVa Alignment tab.
+An similar visualization of the full family is presented in the [Supplemental Analysis](#si) (Figure \@ref(fig:full-IAA-alignment)).
+In the alignment of A class Aux/IAAs, the EAR motif and degron can be readily identified by the drop in nonsynonymous variation, as visualized by the lack of stronger functional effects in Figure \@ref(fig:A-IAA-alignment).
 The PB1 domain is not as readily identified, perhaps because the multiple contributing residues are spread out in linear sequence space.
-It is worth noting that the charged residues that facilitate electrostatic PB1-PB1 interaction show little variation.
+It is worth noting that the charged residues that facilitate electrostatic PB1-PB1 interactions show little variation.
 
@@
 Color scale is explained in [Methods](#methods).
+In grayscale underneath the plot, alignment consensus is shown as measured by Evolutionary trace [@pluciennik_balcony_2018; @wilkins_evolutionary_2012], with high consensus positions in black and low consensus in white.
+Key functional domains are outlined in black and labeled above the plot.
+The EAR domain spans codon alignment positions 70 to 74, corresponding to IAA1 amino acids LRLGL, 14-18.
+The degron domain spans alignment positions 194-201, corresponding to IAA1 amino acids QIVGWPPV, 55-62.
+The charged residues of the PB1 domain correspond to alignment positions 246, 256, 316, 318, 320, and 326 corresponding to IAA1 amino acids K77, R88, D133, D135, D137, and D143.
 
@@
 A_IAAs <- alignCDS(A_IAAs$transcript_ID)
+consensus <- BALCONY::RealValET_conservativity(as.alignment(Biostrings::as.matrix(A_IAAs[[2]])))
+
 # make an alignment data frame
@@
                                              title.hjust = 0.5)) +
-  scale_y_discrete(expand = expand_scale(add = c(0,2.5))) +
+  scale_y_discrete(expand = expand_scale(add = c(1,2.5))) +
   scale_x_continuous(expand = c(0,0)) +  # expand increases distance from axis
@@
         legend.spacing = unit(4, "char"), 
-        legend.box.spacing = unit(0.1,"line")) 
+        legend.box.spacing = unit(0.1,"line")) +
+  annotation_raster(matrix(gray(rescale(consensus, to = c(0,1))), nrow = 1),  xmin = -Inf, xmax = Inf, ymin = -10, ymax = 0.4)
+
 p + annotate("rect", xmin = 69, xmax = 75, ymin = 0.4, 
@@
            fill = NA) +
-  annotate("rect", xmin = 252, xmax = 254, ymin = 0.4, 
-           ymax = length(levels(A_IAAs$tair_symbol)) + 0.6, color = "black", 
+  annotate("rect", xmin = 255, xmax = 257, ymin = 0.4, 
+           ymax = length(levels(A_IAAs$tair_symbol))+0.6, color = "black", 
          fill = NA) +
@@
            fill = NA) +
+  annotate("rect", xmin = 317, xmax = 319, ymin = 0.4, 
+           ymax = length(levels(A_IAAs$tair_symbol))+0.6, color = "black", 
+           fill = NA) +
   annotate("rect", xmin = 319, xmax = 321, ymin = 0.4, 
@@
            fill = NA) + 
-  annotate("text", x = (321-245)/2+245, y= length(levels(A_IAAs$tair_symbol)) + 1.2, label = "PB1 domain", family="Arial")
+  annotate("rect", xmin = 325, xmax = 327, ymin = 0.4, 
+           ymax = length(levels(A_IAAs$tair_symbol))+0.6, color = "black", 
+           fill = NA) +
+  annotate("text", x = (327-245)/2+245, y= length(levels(A_IAAs$tair_symbol)) + 1.2, label = "PB1 domain", family="Arial")
 
@@
 Natural variation also provides a means to study how gene families are evolving.
-To do this, we used ViVa to map onto the Aux/IAA phylogenetic tree the diversity at nonsynonymous variant sites relative to synonymous sites (Figure \@ref(fig:IAAtree)).
-This visualization enables straightforward comparison of rates of recent functional divergence within the context of rates of sequence divergence across the entire gene family.
-By comparing nonsynonymous diversity gene clades can be identified that are likely to exhibit high rates of functional conservation and possible redundancy or conversely, where there is a possibility for recent emergent novel function or pseudogenization.
+To visualize this, we used ViVa to map the diversity at nonsynonymous variant sites relative to synonymous sites onto the Aux/IAA phylogenetic tree  (Figure \@ref(fig:IAAtree)).
+This visualization enables straightforward comparison of rates of recent functional divergence within natural variation in the context of the sequence divergence throughout the history of gene family.
+Clades and individual genes exhibiting low nonsynonymous diversity, and are therefore likely functionally conserved, can be identified.
+Conversely, genes with high nonsynonymous diversity are more likely to be under relaxed functional selection, indicating a possibility of functional drift, emergence of novel function, or pseudogenization.
 
@@
 For example, the *iaa8-1 iaa9-1* double mutant and the *iaa5-1 iaa6-1 iaa19-1* triple mutant have wild-type phenotypes [@overvoorde_functional_2005], yet the *IAA6/IAA19* sister pair has significant differences in expression patterns, protein abundances and functions suggesting they have undergone functional specialization since their divergence [@winkler_variation_2017].
-A closer examination of the *IAA19* and *IAA6* pair within *Brassicaceae* found evidence for positive selection and possible subfunctionalization of *IAA6* relative to *IAA19* [@winkler_variation_2017].
-Consistent with these results, ViVa revealed higher conservation for *IAA19* (${\pi_N/\pi_S}$ = 0.55) compared to *IAA6* (${\pi_N/\pi_S}$ = 2.3) (Figure \@ref(fig:IAAtree)), and also detected high nonsynonymous diversity within the same regions of *IAA6* as seen by Winkler et al.
+A closer examination of the *IAA19* and *IAA6* pair within *Brassicaceae* found evidence for positive selection and subfunctionalization of *IAA6* relative to *IAA19* [@winkler_variation_2017].
+Consistent with these results, ViVa revealed higher conservation (*i.e.* lower ratio of nonsynonymous to synonymous diversity) for *IAA19* (${\pi_N/\pi_S}$ = 0.55) compared to *IAA6* (${\pi_N/\pi_S}$ = 2.3) (Figure \@ref(fig:IAAtree)), and also detected high nonsynonymous diversity within the same regions of *IAA6* as seen by Winkler et al.
 (Figure \@ref(fig:IAA6-div-plot)).
@@
 
-(ref:IAAtree) **IAA protein sequence tree mapped with ${\pi_N/\pi_S}$.** Protein sequences were aligned with DECIPHER [@wright_decipher_2015] and low information content regions were masked [@kuck_parametric_2010] prior to inferring a phylogeny [@ronquist_mrbayes_2003].
+(ref:IAAtree) **IAA protein sequence tree mapped with ${\pi_N/\pi_S}$.** Protein sequences were aligned with DECIPHER [@wright_decipher_2015] and low information content regions were masked with Aliscore [@kuck_parametric_2010] prior to inferring a phylogeny with MrBayes [@ronquist_mrbayes_2003].
 Tips of the tree are mapped with circles of color and diameter proportional to ${\pi_N/\pi_S}$.
@@
 The C class Aux/IAAs are missing one or more of the canonical Aux/IAA domains.
-We found notable exceptions to pattern of diversification and conservation between sister pairs within the Class B *Aux/IAA* genes.
+We found notable exceptions to the pattern of diversification and conservation between sister pairs within the Class B *Aux/IAA* genes.
 The *IAA10*/*IAA11*, *IAA18*/*IAA26*, and *IAA20*/*IAA30* pairs showed similar levels of nonsynonymous diversity.
@@
 In support of this strong conservation of both *IAA10* and *IAA11*, the *Arabidopsis thaliana* ePlant browser indicates that *IAA10* and *IAA11* have almost identical expression patterns [@waese_eplant_2017].
-Together this evidence suggests a strong dosage requirement for these genes or that they have taken on novel functions since their emergence.
+Together this evidence suggests a strong dosage requirement for these genes or that they have taken on novel functions since their emergence, with both genes contributing similarly to plant fitness.

### 2019-04-02 manuscript/04-methods.Rmd
revisions done, back to coauthors

@@
 ### Gene and Transcript Accession Information
-Information on the genes and transcripts including chromosomal coordinates, start and end location, and transcript length were pulled from Araport11 [@cheng_araport11_2017].
+Information on the genes and transcripts including chromosomal coordinates, start and end location, and transcript length were downloaded from Araport11 as a General Feature Format (GFF) file [@cheng_araport11_2017].
+The Araport11 full genome general feature format file, which can also be found on the TAIR website (https://www.arabidopsis.org/download/index-auto.jsp?dir=%2Fdownload_files%2FGenes%2FAraport11_genome_release), has been embedded in the R-package as GRanges object, `gr`.
 The TAIR10 database, found at http://arabidopsis.org, was accessed via the `biomart` function, using the biomaRt R-package.
-The Araport11 full genome general feature format file, which can also be found on the TAIR website (https://www.arabidopsis.org/download/index-auto.jsp?dir=%2Fdownload_files%2FGenes%2FAraport11_genome_release), has been embedded in the R-package as GRanges object, `gr`.
 Gene identifiers used in this study are in Table \@ref(tab:gene-ids-table).
@@
 Other packages that were critical to building ViVa and/or writing this document include: [@R-ape; @R-base; @R-bindrcpp; @R-BiocGenerics; @R-Biostrings; @R-bookdown; @R-colorspace; @R-DECIPHER; @R-dplyr; @R-DT; @R-forcats; @R-gginnards; @R-ggplot2; @R-ggpmisc; @R-ggseqlogo; @R-ggthemes; @R-ggtree; @R-ips; @R-IRanges; @R-knitr; @R-magrittr; @R-plyr; @R-purrr; @R-r1001genomes; @R-RColorBrewer; @R-readr; @R-reshape2; @R-reticulate; @R-rmarkdown; @R-RSQLite; @R-S4Vectors; @R-scales; @R-stringr; @R-tibble; @R-tidyr; @R-tidyverse; @R-treeio; @R-viridis; @R-viridisLite; @R-XML; @R-XVector]
+
+## Testing
+A group consisting of undergraduate and graduate students, postdoctoral and postbaccalaureate researchers (the authors of this publication) were assembled to test the functionality of ViVa.
+Testers were asked to use the ViVa web interface to analyze the genetic variation within a gene family or clade within a family of the auxin nuclear signaling pathway.
+Testers were provided with a brief vignette on how to use ViVa to formulate new hypotheses or support existing hypotheses about gene function and evolution, which has been expanded to the "An Overview of ViVa" section.
+The testers had background knowledge of these genes as members of laboratories studying auxin nuclear signaling.
+Testers met weekly with the developers to discuss their user experience and their findings.
+Tester experiences, issues, and suggestions were incorporated into the ViVa software.
+The results collected by these analyses are summarized in the "Visualizing Variation within the auxin signaling pathway" section and [Supplemental Data](#si).

### 2019-04-02 manuscript/06-supplement.Rmd
revisions done, back to coauthors

@@
 alignment <- alignCDS(gene_info$transcript_ID)
+consensus <- BALCONY::RealValET_conservativity(as.alignment(Biostrings::as.matrix(alignment[[2]])))
 # make an alignment data frame
@@
                                              title.hjust = 0.5)) +
-  scale_y_discrete(expand = c(0,0)) +
+  scale_y_discrete(expand = expand_scale(add = c(1,2.5))) +
   scale_x_continuous(expand = c(0,0)) +  # expand increases distance from axis
@@
         legend.spacing = unit(4, "char"), 
-        legend.box.spacing = unit(0.1,"line")) 
+        legend.box.spacing = unit(0.1,"line"))  +
+  annotation_raster(matrix(gray(rescale(consensus, to = c(0,1))), nrow = 1),  xmin = -Inf, xmax = Inf, ymin = -10, ymax = 0.4)
 p
@@
 Color scale is explained in [Methods](#methods).
+In grayscale underneath the plot, alignment consensus is shown as measured by Evolutionary trace [@pluciennik_balcony_2018; @wilkins_evolutionary_2012], with high consensus positions in black and low consensus in white.
 
@@
 
+```{r AFB1-map, fig.cap="(ref: AFB1-map)"}
+knitr::include_graphics("app_figure/Map_figure.png")
+```
+(ref: AFB1-map) **Map of AFB1 Oligomerization domain variant accessions** (a) Map showing the two accessions with variants in the AFB1 oligomerization domain (b) The user selectable parameters of the mapping tab used to generate the map are provided as an example of using the SNP Mapping tab in the ViVa webtool.
+See Figure \@ref(fig:app-workflow) for details on filling in the parameters.
+
 The AFB4 and AFB5 receptors have an N-terminal extension prior to their F-box domains.
-This extension had very high nonsynonymous diversity (Figure \@ref(fig:full-AFBs-alignment), suggesting that this extension does not play an important functional role in these proteins.
+This extension had very high nonsynonymous diversity (Figure \@ref(fig:full-AFBs-alignment)), suggesting that this extension does not play an important functional role in these proteins.
 Additionally, two frameshift variants and one stop-gained variant were observed in *AFB4* supporting its pseudogenization.
-
-## *Aux/IAA* genes {#IAAs}
-
-For simplicity we have included only the alignment of the class A Aux/IAAs in the main manuscript. For completion we include here the complete alignment of the family.
-
-```{r IAAdivstats}
-gene_info <- geneInfoFromFile(
-  fname = system.file("shiny-app", "gene_family_data",
-                      "IAA_gene_ids.csv", package = "r1001genomes"))
-#annotation <- readAnnotationFile(filename = system.file("shiny-app", "gene_family_data",
-#                      "IAA_annotations.csv", package = "r1001genomes"), gene_info = gene_info)
-
-VCF <- readRDS(system.file("shiny-app", "gene_family_data",
-                      "IAAs_VCF.rds", package = "r1001genomes"))
-div_stats <- ldply(VCF, diversityStats, geneInfo=gene_info, .id="transcript_ID")
-div_stats$tair_symbol %<>% fct_reorder(.x = , div_stats$tair_symbol, 
-                                       .fun = function(x) 
-                                         {as.integer(str_remove(x, 
-                                              pattern = "IAA"))})
-div_stats %<>% arrange(tair_symbol)
-div_stats %<>% dplyr::select(tair_symbol, everything())
-
-```
-
-```{r IAA alignment sup}
-alignment <- alignCDS(gene_info$transcript_ID)
-# make an alignment data frame
-aln_df <- makeAlnDF(alignment[[2]])
-
-# make a collapsed alignment for translating positions
-collapsed_aln_df <- 
-  aln_df %>% dplyr::select(aln_pos, seq_pos, seq_name) %>%
-  tidyr::spread(key = seq_name, value = seq_pos)
-```
-
-```{r Mask_IAA_alignment_sup, eval=FALSE}
-aliscore(as.AAbin(alignment[[2]]), path = "/Volumes/SSD/Applications/ALISCORE_v2.0/Aliscore_v.2.0/")
-id <- scan(file = "input.fas_List_random.txt", sep = " ", quiet = TRUE)
-AAAlign_no_gaps <- as.matrix(alignment[[2]])[, id]
-write.nexus.data(AAAlign_no_gaps, file = "manuscript/IAAs.nex", format = "protein")
-```
-```{bash IAA_phylogeny_construction_sup, eval=FALSE}
-~/Desktop/natural-variation-webtool/manuscript
-
-/Applications/MrBayes/mb
-
-execute IAAs.nex
-# set IAA33 as the outgroup
-outgroup AT5G57420.1
-
-lset nucmodel=protein Rates=invgamma
-# or include Nst=6  ngammacat=4 omegavar=M3 
-prset aamodelpr=fixed(jones) statefreqpr=fixed(empirical) 
-# initial run with aamodelpr=mixed yielded posterior prob of 1 for jones model
-propset ExtTBR$prob=0
-
-mcmcp ngen=250000 nchains=4 printfreq=1000 samplefreq=100 burnin=200
-
-mcmc
-sump
-sumt
-
-```
-
-```{r Load IAA phylogeny and data_sup}
-tree <- treeio::read.beast(file = "trees/IAAs.nex.con.tre")
-
-IAA_data <- read.csv("tables/IAAs_table.csv")
-#tip_labels need to be in the first column
-IAA_data %<>% dplyr::select(name, everything())
-IAA_data$name %<>% as.character()
-IAA_data$AGI %<>% as.character()
-
-IAA_data <- IAA_data[match(gsub(x = tree@phylo$tip.label, pattern = "\\.1",replacement = ""), IAA_data$AGI), ] 
-tree@phylo$tip.label <- IAA_data$name
-```
-
-(ref:full-IAA-alignment) **Alignment of the complete Aux/IAA family.** Protein sequences were aligned with DECIPHER [@wright_decipher_2015] and variants were mapped to this alignment and colored according to the predicted functional effect of the allele of strongest effect at that position, with light colors having weaker effects on function and darker colors stronger effects.
-Red indicates missense variants.
-Color scale is explained in [Methods](#methods).
-
-```{r full-IAA-alignment, fig.cap="(ref:full-IAA-alignment)", fig.height=8}
-IAAs <- alignCDS(gene_info$transcript_ID)
-# make an alignment data frame
-IAAs <- makeAlnDF(IAAs[[2]])
-IAAs <- left_join(IAAs, dplyr::select(gene_info, "tair_locus",
-                                       "tair_symbol", "transcript_ID"),
-                        by = c("transcript_ID" = "transcript_ID"))
-IAAs$tair_symbol %<>% as.factor()
-IAAs$tair_symbol %<>% factor(levels = IAA_order, 
-                                       ordered = TRUE)
-IAAs$tair_symbol %<>% fct_rev()
-VCF <- ldply(.data = VCF, .fun = subset, gt_GT != "0|0")
-coding_VCF <- getCodingDiv(VCF)
-IAAs <- addSNPsToAlnDF(aln_df = IAAs,SNPs = coding_VCF, effect_order = SNPeff_order)
-IAAs$gap <- IAAs$seq_pos == "-"
-
-p <-ggplot(IAAs, aes(x = aln_pos, y = tair_symbol)) +
-  geom_raster(data = IAAs, 
-             # mapping = aes(fill = effects)) +
-              mapping = aes(fill = strength,
-                            alpha = gap)) +
-  scale_alpha_manual(values = c("TRUE" = 0, "FALSE" = 1), 
-                     labels = c("aligned", "gap"), 
-                     guide = guide_legend(override.aes = 
-                                            list(fill = c("grey85", "white")),
-                       title = "alignment",
-                       direction = "vertical")) + 
-   scale_fill_viridis(breaks = range(SNPeff_order$strength),
-                      limits = range(SNPeff_order$strength),
-                      na.value = "grey85",
-                      direction = -1,
-                      labels = c("weak", "strong"),
-                      option = "A", 
-                      guide = guide_colorbar(ticks = FALSE, 
-                                             title.vjust = .8, 
-                                             title.position = "top", 
-                                             title.hjust = 0.5)) +
-  scale_y_discrete(expand = c(0,0)) +
-  scale_x_continuous(expand = c(0,0)) +  # expand increases distance from axis
-  labs(x = "codon position in alignment", y = "", fill = "variant functional effect") +
-  theme_logo(base_family = "Helvetica") +
-  theme(panel.grid = element_blank(), 
-        panel.grid.minor = element_blank(),
-        strip.background = element_blank(),
-        strip.text.x = element_blank(),
-        #panel.background = element_rect(fill = "grey85", color = "white"),
-        legend.position = c("bottom"),
-        legend.key = element_rect(linetype = "solid"),
-        legend.spacing = unit(4, "char"), 
-        legend.box.spacing = unit(0.1,"line")) 
-p
-
-#ggsave("IAAs.png", width = 7, height = 4)
-```
-
-```{r, IAA6-div-plot, fig.cap="(ref:IAA6-div-plot)"}
-coding_variants <- VCF[VCF$Transcript_ID == gene_info[gene_info$tair_symbol == "IAA6", "transcript_ID"], ]
-p <- plotCodingDiv(uniqueCodingVars = coding_variants)
-p + annotate("rect", xmin = 106, xmax = 127, ymin = 0.001, ymax = 1, color = "lightgrey", alpha =.1) + annotate("text", x = 116.5, y = 0.5, label = "region of positive\nselection identified\nin Winkler et al.")
-```
-(ref:IAA6-div-plot) **IAA6 diversity plot.** Nucleotide diversity of variant positions throughout the IAA6 coding sequence are plotted and colored according to the effect of the variant alleles at each position.
-The region of positive selection identified by Winkler et al.
-is highlighed.
 
@@
 alignment <- alignCDS(gene_info$transcript_ID)
+consensus <- BALCONY::RealValET_conservativity(as.alignment(Biostrings::as.matrix(alignment[[2]])))
 #possible outgroups  AtLUG, At4g32551; AtLUH, At2g32700
@@
 Color scale is explained in [Methods](#methods).
+In grayscale underneath the plot, alignment consensus is shown as measured by Evolutionary trace [@pluciennik_balcony_2018; @wilkins_evolutionary_2012], with high consensus positions in black and low consensus in white.
 ```{r full-TPLs-alignment, fig.cap= "(ref:full-TPLs-alignment)"}
@@
         legend.spacing = unit(4, "char"), 
-        legend.box.spacing = unit(0.1,"line")) 
+        legend.box.spacing = unit(0.1,"line")) +
+  annotation_raster(matrix(gray(rescale(consensus, to = c(0,1))), nrow = 1),  xmin = -Inf, xmax = Inf, ymin = -10, ymax = 0.4)
 p
@@
 This conservation in the PB1 domain of ARF10 and the DBD of ARF16 may suggest subfunctionalization in this family.
+
+## *Aux/IAA* genes {#IAAs}
+
+For simplicity we have included only the alignment of the class A Aux/IAAs in the main manuscript.
+For completion we include here the complete alignment of the family (Figure \@ref(fig:full-IAA-alignment)).
+
+```{r IAAdivstats}
+gene_info <- geneInfoFromFile(
+  fname = system.file("shiny-app", "gene_family_data",
+                      "IAA_gene_ids.csv", package = "r1001genomes"))
+#annotation <- readAnnotationFile(filename = system.file("shiny-app", "gene_family_data",
+#                      "IAA_annotations.csv", package = "r1001genomes"), gene_info = gene_info)
+
+VCF <- readRDS(system.file("shiny-app", "gene_family_data",
+                      "IAAs_VCF.rds", package = "r1001genomes"))
+div_stats <- ldply(VCF, diversityStats, geneInfo=gene_info, .id="transcript_ID")
+div_stats$tair_symbol %<>% fct_reorder(.x = , div_stats$tair_symbol, 
+                                       .fun = function(x) 
+                                         {as.integer(str_remove(x, 
+                                              pattern = "IAA"))})
+div_stats %<>% arrange(tair_symbol)
+div_stats %<>% dplyr::select(tair_symbol, everything())
+
+```
+
+```{r IAA alignment sup}
+alignment <- alignCDS(gene_info$transcript_ID)
+# make an alignment data frame
+aln_df <- makeAlnDF(alignment[[2]])
+
+# make a collapsed alignment for translating positions
+collapsed_aln_df <- 
+  aln_df %>% dplyr::select(aln_pos, seq_pos, seq_name) %>%
+  tidyr::spread(key = seq_name, value = seq_pos)
+```
+
+```{r Mask_IAA_alignment_sup, eval=FALSE}
+aliscore(as.AAbin(alignment[[2]]), path = "/Volumes/SSD/Applications/ALISCORE_v2.0/Aliscore_v.2.0/")
+id <- scan(file = "input.fas_List_random.txt", sep = " ", quiet = TRUE)
+AAAlign_no_gaps <- as.matrix(alignment[[2]])[, id]
+write.nexus.data(AAAlign_no_gaps, file = "manuscript/IAAs.nex", format = "protein")
+```
+```{bash IAA_phylogeny_construction_sup, eval=FALSE}
+~/Desktop/natural-variation-webtool/manuscript
+
+/Applications/MrBayes/mb
+
+execute IAAs.nex
+# set IAA33 as the outgroup
+outgroup AT5G57420.1
+
+lset nucmodel=protein Rates=invgamma
+# or include Nst=6  ngammacat=4 omegavar=M3 
+prset aamodelpr=fixed(jones) statefreqpr=fixed(empirical) 
+# initial run with aamodelpr=mixed yielded posterior prob of 1 for jones model
+propset ExtTBR$prob=0
+
+mcmcp ngen=250000 nchains=4 printfreq=1000 samplefreq=100 burnin=200
+
+mcmc
+sump
+sumt
+
+```
+
+```{r Load IAA phylogeny and data_sup}
+tree <- treeio::read.beast(file = "trees/IAAs.nex.con.tre")
+
+IAA_data <- read.csv("tables/IAAs_table.csv")
+#tip_labels need to be in the first column
+IAA_data %<>% dplyr::select(name, everything())
+IAA_data$name %<>% as.character()
+IAA_data$AGI %<>% as.character()
+
+IAA_data <- IAA_data[match(gsub(x = tree@phylo$tip.label, pattern = "\\.1",replacement = ""), IAA_data$AGI), ] 
+tree@phylo$tip.label <- IAA_data$name
+```
+
+(ref:full-IAA-alignment) **Alignment of the complete Aux/IAA family.** Protein sequences were aligned with DECIPHER [@wright_decipher_2015] and variants were mapped to this alignment and colored according to the predicted functional effect of the allele of strongest effect at that position, with light colors having weaker effects on function and darker colors stronger effects.
+Red indicates missense variants.
+Color scale is explained in [Methods](#methods).
+In grayscale underneath the plot, alignment consensus is shown as measured by Evolutionary trace [@pluciennik_balcony_2018; @wilkins_evolutionary_2012], with high consensus positions in black and low consensus in white.
+
+```{r full-IAA-alignment, fig.cap="(ref:full-IAA-alignment)", fig.height=8}
+IAAs <- alignCDS(gene_info$transcript_ID)
+consensus <- BALCONY::RealValET_conservativity(as.alignment(Biostrings::as.matrix(IAAs[[2]])))
+# make an alignment data frame
+IAAs <- makeAlnDF(IAAs[[2]])
+IAAs <- left_join(IAAs, dplyr::select(gene_info, "tair_locus",
+                                       "tair_symbol", "transcript_ID"),
+                        by = c("transcript_ID" = "transcript_ID"))
+IAAs$tair_symbol %<>% as.factor()
+IAAs$tair_symbol %<>% factor(levels = IAA_order, 
+                                       ordered = TRUE)
+IAAs$tair_symbol %<>% fct_rev()
+VCF <- ldply(.data = VCF, .fun = subset, gt_GT != "0|0")
+coding_VCF <- getCodingDiv(VCF)
+IAAs <- addSNPsToAlnDF(aln_df = IAAs,SNPs = coding_VCF, effect_order = SNPeff_order)
+IAAs$gap <- IAAs$seq_pos == "-"
+
+p <-ggplot(IAAs, aes(x = aln_pos, y = tair_symbol)) +
+  geom_raster(data = IAAs, 
+             # mapping = aes(fill = effects)) +
+              mapping = aes(fill = strength,
+                            alpha = gap)) +
+  scale_alpha_manual(values = c("TRUE" = 0, "FALSE" = 1), 
+                     labels = c("aligned", "gap"), 
+                     guide = guide_legend(override.aes = 
+                                            list(fill = c("grey85", "white")),
+                       title = "alignment",
+                       direction = "vertical")) + 
+   scale_fill_viridis(breaks = range(SNPeff_order$strength),
+                      limits = range(SNPeff_order$strength),
+                      na.value = "grey85",
+                      direction = -1,
+                      labels = c("weak", "strong"),
+                      option = "A", 
+                      guide = guide_colorbar(ticks = FALSE, 
+                                             title.vjust = .8, 
+                                             title.position = "top", 
+                                             title.hjust = 0.5)) +
+  scale_y_discrete(expand = expand_scale(add = c(1,2.5))) +
+  scale_x_continuous(expand = c(0,0)) +  # expand increases distance from axis
+  labs(x = "codon position in alignment", y = "", fill = "variant functional effect") +
+  theme_logo(base_family = "Helvetica") +
+  theme(panel.grid = element_blank(), 
+        panel.grid.minor = element_blank(),
+        strip.background = element_blank(),
+        strip.text.x = element_blank(),
+        #panel.background = element_rect(fill = "grey85", color = "white"),
+        legend.position = c("bottom"),
+        legend.key = element_rect(linetype = "solid"),
+        legend.spacing = unit(4, "char"), 
+        legend.box.spacing = unit(0.1,"line"))  +
+  annotation_raster(matrix(gray(rescale(consensus, to = c(0,1))), nrow = 1),  xmin = -Inf, xmax = Inf, ymin = -10, ymax = 0.4)
+p
+
+#ggsave("IAAs.png", width = 7, height = 4)
+```
+
+```{r IAA6-div-plot, fig.cap="(ref:IAA6-div-plot)"}
+coding_variants <- VCF[VCF$Transcript_ID == gene_info[gene_info$tair_symbol == "IAA6", "transcript_ID"], ]
+p <- plotCodingDiv(uniqueCodingVars = coding_variants)
+p + annotate("rect", xmin = 106, xmax = 127, ymin = 0.001, ymax = 1, color = "lightgrey", alpha =.1) + annotate("text", x = 116.5, y = 0.5, label = "region of positive\nselection identified\nin Winkler et al.")
+```
+(ref:IAA6-div-plot) **IAA6 diversity plot.** Nucleotide diversity of variant positions throughout the IAA6 coding sequence are plotted and colored according to the effect of the variant alleles at each position.
+The region of positive selection identified by Winkler et al.
+is highlighted.

### 2019-04-02 manuscript/index.Rmd
revisions done, back to coauthors

@@
 ViVa is also readily extensible to databases of interspecific genetic variation in plants as well as other organisms, such as the 3,000 Rice Genomes Project (http://snp-seek.irri.org/) and human genetic variation (https://www.ncbi.nlm.nih.gov/clinvar/).
+"
 
 Keywords: natural variation, Arabidopsis thaliana, genome diversity, structure, function, genotype, phenotype, bioconductor, bookdown, r
-
-"
 
@@
 library(bookdown)
+library(BALCONY)
 library(ape)

### 2019-04-21 manuscript/01-intro.Rmd
submitted revisions

@@
 Beyond its utility in gene discovery and understanding genome evolution, natural genetic variation provides a catalog of permissible polymorphisms that can facilitate the connection of genotype to phenotype at the gene, gene family and network scales [@nieduszynski_sequence_2011; @joly-lopez_developing_2016].
-This is an especially critical resource in large gene families where loss of function in individual genes may have little or no phenotypic effect [@dharmasiri_plant_2005; @moore_evolutionary_2005; @guo_gene_2013] and directed allele replacement remains time and resource-intensive [@chen_crispr_2019].
+This is an especially critical resource for studying large gene families where loss of function in individual genes may have little or no phenotypic effect [@dharmasiri_plant_2005; @moore_evolutionary_2005; @guo_gene_2013] and directed allele replacement remains time and resource-intensive [@chen_crispr_2019].
 Natural variation datasets provide novel alleles and germplasm which can be examined with biochemical and genetic approaches to map sequence to function and genotype to phenotype.
@@
 One reason may be the current requirement for a fairly high level of bioinformatics expertise to extract the desired information from whole genome resequencing datasets.
-While existing resources such as the 1001 Proteomes website [@joshi_1001_2012] and ePlant [@waese_eplant_2017] facilitate access to this data at the gene scale, they do not provide comparative summaries or visualizations of variation at the gene family or network scales.
+While existing resources such as the 1001 Proteomes website [@joshi_1001_2012] and ePlant [@waese_eplant_2017] facilitate access to this data at the gene scale, they do not provide comparative summaries or visualizations of variation at the gene family scale.
 To address this concern, we created ViVa: a webtool and R-package for **Vi**sualizing **Va**riation, which allows plant molecular biologists of any level access to gene-level data from the 1001 Genomes database.
 Using ViVa researchers may: 1) Identify polymorphisms to facilitate biochemical assays of variant effects [@starita_variant_2017; @wright_insights_2017]; 2) Produce family-wise alignments of variants to facilitate *de novo* functional domain identification [@melamed_combining_2015]; 3) Generate lists of accessions containing polymorphisms to facilitate phenotypic analysis of gene variant effects [@park_distributions_2017]; and 4) Quantify metrics of genetic diversity to facilitate the study of gene, gene family, and network evolution [@delker_natural_2010; @kliebenstein_role_2008].
-Here we present a summary of the functionality of ViVa along with an analysis of the natural variation in the nuclear auxin signaling pathway using ViVa.
-To demonstrate ViVa usage we focus on the analysis of the *Aux/IAA* family and reserve the remainder of our analysis for the interested reader as supplemental material.
+Here we present a summary of the functionality of ViVa and an analysis of the natural variation in the nuclear auxin signaling network using ViVa.
+To succinctly demonstrate the use of ViVa, we focus on the analysis of the *Aux/IAA* family; similar analyses were performed for the other nuclear auxin signaling gene families and are provided as [Supplemental Analysis](#si) for the interested reader.

### 2019-04-21 manuscript/02-results.Rmd
submitted revisions

@@
 This reflects challenges both in alignment of non-coding sequences and the increased difficulty in assessing the effects of variation in these regions [@alexandre_complex_2018].
-These analyses can be conjoined, as was done here, to understand variation within gene networks.
-Visualizations of variation within whole gene networks are planned for future iterations.
 
+The stable version of ViVa is hosted at https://www.plantsynbiolab.bse.vt.edu/ViVa/.
 The development version of ViVa can be accessed as a Docker container https://hub.docker.com/r/wrightrc/r1001genomes/ or as an R-package at https://github.com/wrightrc/r1001genomes.
-ViVa will be hosted at https://www.plantsynbiolab.bse.vt.edu/ViVa/ upon release.
 
@@
 (ref:app-workflow) **ViVa workflow** (a) Workflow diagram of ViVa.
-Blue indicates user actions, orange indicates processing steps by the application.
-(b) Detailed look at mapping tab parameters: User selects which genes to look at (1) then clicks the submit button (2).
-The Allele selection pane is then filled in with all non-reference variants meeting the criteria.
-The user can then adjust the range of nucleotide diversity and the type of SNP with a slider and radio buttons (3) to make the list of variants a manageable size.
-The list of variants will update as changes are made to these controls.
-The user then selects variants to display on the map from the allele selection pane (4).
-Clicking the update map button (5) will populate the map below with points located at the collection coordinates of each accession, colored by the selected non-reference alleles that are present in the sample.
-(c) Detailed look at the browser tab options: The user first selects which genes analyze and clicks submit (1).
+Blue indicates user actions, yellow indicates processing steps performed by the application.
+(b) Detailed look at mapping tab parameters: User selects which genes to look at (1) then clicks the Submit button (2).
+The "Allele selection" pane is then filled in with all non-reference variants meeting the criteria.
+The user can adjust the range of nucleotide diversity and the type of SNP with a slider and radio buttons (3) to make the list of variants a manageable size.
+The list of variants is updated as changes are made to these controls.
+The user then selects variants to display on the map (4).
+Clicking the Update Map button (5) populates the map below with points located at the collection coordinates of each accession and colored by the selected alleles.
+(c) Detailed look at the browser tab options: The user first selects which genes to analyze and clicks submit (1).
 The “hide 0|0 genotype?” checkbox (2) removes rows from the table containing the reference allele.
@@
 The second type of filter is numeric range matching (4); the user again selects a column, then specifies a range to match by typing or selecting minimum and maximum values.
-After configuring the filters, the user then clicks the apply filters button (5) to update the table.
-The table will be updated such that only rows that meet all filter conditions are displayed.
+After configuring the filters, the user clicks the Apply Filters button (5) to update the table.
+The table is updated such that only rows that meet all filter conditions are displayed.
 ```{r app-workflow, fig.cap = "(ref:app-workflow)"}
@@
 At the top of the ViVa webtool are two collapsible panels used for entering the genes to query and custom annotations for those genes (Figure \@ref(fig:app-figure-elements)a and \@ref(fig:app-workflow)a).
-The Gene Select panel permits gene input by either typing in the genes' AGI/TAIR locus identifiers, or uploading a .csv file.
-The Annotation Files panel is optionally used to upload an annotation file containing coordinates of domains, mutations, or any other sequence knowledge that will be plotted on some of the tabs of the webtools analysis section.
+The Gene Select panel permits gene input by either typing in or uploading a .csv file of AGI/TAIR locus identifiers.
+The Annotation Files panel is optionally used to upload an annotation file containing coordinates of domains, mutations, or any other sequence knowledge that can be plotted on some of the tabs of the ViVa analysis tabs.
 
@@
 The Total Polymorphism Counts table provides the total number of variant observations of non-reference alleles categorized by type and location (the Col-0 accession TAIR9 genome is the reference genome for this dataset).
-The Unique Allele Count table only counts the number of unique variant alleles within the population of accessions (*e.g.* if multiple accessions have the same variant, these will be counted as a single allele at that position).
+The Unique Allele Count table only counts the number of unique variant alleles within the population of accessions (*e.g.* if multiple accessions have the same variant, these are counted as a single allele at that position).
 
-The Nucleotide Diversity Statistics table provides a nucleotide diversity statistic (${\pi}$) for the transcript and the coding sequence of each gene [@nei_mathematical_1979].
-Nucleotide diversity is also calculated for the set of only synonymous (${\pi_S}$) and only nonsynonymous sites (${\pi_N}$).
-The ratio of the presence of nonsynonymous to synonymous polymorphism provides a measure of the potential for functional diversity [@firnberg_genetic_2013; @whitehead_optimization_2012].
+The Nucleotide Diversity Statistics table provides nucleotide diversity statistic (${\pi}$) values for the transcript and the coding sequence of each gene [@nei_mathematical_1979].
+Given a set of nucleotide sequences from a population, ${\pi}$ is the average number of nucleotide differences per site.
+Nucleotide diversity is also calculated for the set of only synonymous (${\pi_S}$) and only nonsynonymous polymorphisms (${\pi_N}$).
+As nonsynonymous polymorphisms are more likely to give rise to functional change than synonymous polymorphisms, the ratio of the presence of nonsynonymous to synonymous polymorphisms provides a measure of the potential for functional diversity [@firnberg_genetic_2013; @whitehead_optimization_2012].
 We present ${\pi_N/\pi_S}$, the ratio of nonsynonymous to synonymous diversity here as a correlate for functional diversity throughout ViVa [@nelson_snpgenie_2015; @hughes_adaptive_2000].
@@
 The colors of markers on the plot identify the effect of the polymorphism.
-If annotation files are provided, the background of the plot will be color coded by the annotated regions.
-If points on the plot are selected by clicking and dragging a box over them, the data for the selected points will appear in the grey box below the plot.
+If annotation files are provided, the background of the plot is color-coded by the annotated regions.
+If points on the plot are selected by clicking and dragging a box over them, the data for the selected points appears in the grey box below the plot.
 Below this is a complete data table containing all points on the plot which can be downloaded as a .csv file.
@@
 ### SNP Mapping: View distributions of SNPs across the globe
-The SNP Mapping tab plots the accessions collection locations on a world map and colors the points based on selected variant alleles (Figure \@ref(fig:app-figure-elements)d).
-After selecting the genes and filtering on the SNP type and level of nucleotide diversity, a group of checkboxes becomes available to select variant alleles to display on the map (Figure \@ref(fig:app-workflow)b and \@ref(fig:AFB1-map)).
+The SNP Mapping tab plots the accessions' collection locations on a world map and colors the points based on selected variant alleles (Figure \@ref(fig:app-figure-elements)d).
+After the user selects genes and filters based on the SNP type and level of nucleotide diversity, a group of checkboxes becomes available to select variant alleles to display on the map (Figure \@ref(fig:app-workflow)b and \@ref(fig:AFB1-map)).
 The variant alleles are labeled with the Transcript_ID and Amino_Acid_Change fields, in the form [Transcript_ID|Amino_Acid_Change].
-After selecting the variant alleles and updating the map, the accessions on the map will be colored by each present unique combination of the selected alleles.
+After selecting the variant alleles and updating the map, the accessions are plotted on the map colored by each unique combination of the selected alleles.
 Below the map is a table containing the accession details for all mapped accessions.
@@
 After selecting the transcripts to include, a number of filters can be applied to the dataset to match text values (e.g.
-gene name or variant effect), or set minimum and maximum limits on the values of numeric fields (e.g.
+gene name or variant effect) or set minimum and maximum limits on the values of numeric fields (e.g.
 nucleotide diversity) (\@ref(fig:app-workflow)c).
-When these filters are applied, the table below will update to only contain rows meeting the criteria for all filters.
-This tab can be useful for identifying all accessions with a particular SNP allele, or any non-reference alleles in a particular region of a gene that may not have been easily accessible in another tab.
+When these filters are applied, the table below is updated to only contain rows meeting the criteria for all filters.
+This tab can be useful for identifying all accessions with a particular allele, or any non-reference alleles in a particular region of a gene that may not have been easily accessible in another tab.
 
 ### Alignments: Visualize SNPs on alignments of homologous genes
-The Alignments tab provides DNA and amino acid sequence alignments of selected genes, colored according to the variant allele with the strongest functional effect at each position (Figure \@ref(fig:app-figure-elements)f and \@ref(fig:A-IAA-alignment), for description of color scale see Figure \@ref(fig:SNPeff-order)).
+The Alignments tab provides DNA and amino acid sequence alignments of selected genes, colored according to the variant allele with the strongest functional effect at each position (Figure \@ref(fig:app-figure-elements)f and \@ref(fig:A-IAA-alignment), for a full description of the color scale see Figure \@ref(fig:SNPeff-order)).
 The content of this tab is most useful if the selected genes are all family members or have significant sequence homology.
-If annotation files are uploaded, open boxes will be overlayed on the alignment, colored by annotation.
+If annotation files are uploaded, open boxes will be overlaid on the alignment, colored by annotation.
 Hovering the cursor over variants will provide additional details about the alleles present at that locus.
@@
 The ViVa R package can be found at https://github.com/wrightrc/r1001genomes and can be installed in your R environment via the `devtools` package:
-`devtools::install_github("wrightrc/r1001genomes")`.
+`remotes::install_github("wrightrc/r1001genomes")`.
 
@@
 Using this existing knowledge, we were able to qualitatively assess the predictive ability of the ViVa modules.
-Below, we describe the results for the *Aux/IAA* family in more detail as a brief vignette of ViVa use.
+Below, we describe the results for detailed analysis of the *Aux/IAA* family in more detail as a brief vignette of ViVa use.
 Analysis of each gene family in the nuclear auxin signaling network can be found in the [Supplemental Analysis](#si).
 
-The Aux/IAA family plays a critical role in transmitting auxin signals, as their degradation is triggered by auxin accumulation.
-This degradation is in response to ubiquitination via a SKP-Cullin-F-box ubitquitin ligase complex containing an Auxin-signaling F-box (AFB) auxin receptor protein.
+The Aux/IAA family plays a critical role in transmitting auxin signals.
+Aux/IAA degradation is triggered by auxin accumulation [@zenser_auxin_2001] and is mediated by ubiquitination via a SKP-Cullin-F-box ubiquitin ligase complex containing an Auxin-signaling F-box (AFB) auxin receptor protein [@gray_auxin_2001].
 Variation in the *AFB* family is presented in [Supplemental Analysis](#si) Figures \@ref(fig:AFBtree), \@ref(fig:full-AFBs-alignment), and \@ref(fig:AFB1-map)).
-The Aux/IAAs repress transcription of Auxin Response Factor (ARF) bound genes, via recruitment of TOPLESS (TPL) and TOPLESS-related (TPR) co-repressors.
+The Aux/IAAs repress transcription of Auxin Response Factor (ARF) bound genes [@tiwari_aux_2004], via recruitment of TOPLESS (TPL) and TOPLESS-related (TPR) co-repressors [@szemenyei_topless_2008].
 Variation in the *TPL/TPR* family is presented in [Supplemental Analysis](#si) Figures \@ref(fig:TPLtree) and \@ref(fig:full-TPLs-alignment)).
-Thus Aux/IAA degradation, relieves this repression allowing ARFs to activate transcription of auxin response genes.
+Thus Aux/IAA degradation, relieves this repression allowing ARFs to activate transcription of auxin response genes [@tiwari_roles_2003].
 Variation in the *ARF* family is presented in [Supplemental Analysis](#si) Figures \@ref(fig:ARFtree) and \@ref(fig:full-ARFs-alignment)).
 
-To predict the functional impact of variation in coding sequences of genes ViVa uses the frequency of nonsynonymous and synonymous polymorphisms.
+To predict the functional impact of variation in gene coding sequences, ViVa uses the frequency of nonsynonymous and synonymous polymorphisms.
 In most cases, nonsynonymous polymorphisms in critical functional domains are likely to have deleterious effects on gene function [@hughes_adaptive_2000].
-This will result in less accumulation of nonsynonymous polymorphism in domains which are critical to plant fitness.
-Therefore, we reasoned that scanning gene coding sequences for regions of relatively low nonsynonymous diversity should highlight functional domains.
+Therefore, domains which are critical to plant fitness will accumulate fewer nonsynonymous polymorphism than regions which are non-critical domains.
+Thus, we reasoned that scanning gene coding sequences for regions of relatively low nonsynonymous diversity should highlight functional domains.
 This general principle can be seen clearly in the analysis of the *Aux/IAA* family of transcriptional co-repressors/co-receptors.
-Most, Aux/IAAs have three major domains.
+Most Aux/IAAs have three major domains.
 Domain I contains an EAR motif that facilitates interaction with TPL/TPR transcriptional co-repressors [@tiwari_aux_2004; @szemenyei_topless_2008].
@@
 
-Here, we analyze just the most conserved A class Aux/IAAs, which contain all of these domains, by mapping the natural genetic variation onto an alignment of the coding sequences via the ViVa Alignment tab.
-An similar visualization of the full family is presented in the [Supplemental Analysis](#si) (Figure \@ref(fig:full-IAA-alignment)).
-In the alignment of A class Aux/IAAs, the EAR motif and degron can be readily identified by the drop in nonsynonymous variation, as visualized by the lack of stronger functional effects in Figure \@ref(fig:A-IAA-alignment).
-The PB1 domain is not as readily identified, perhaps because the multiple contributing residues are spread out in linear sequence space.
-It is worth noting that the charged residues that facilitate electrostatic PB1-PB1 interactions show little variation.
+We focused our analysis on the most conserved A class Aux/IAAs, which possess all three of these domains.
+We began by mapping the natural genetic variation onto an alignment of the coding sequences via the ViVa Alignment tab.
+A similar visualization of the full gene family is presented in the [Supplemental Analysis](#si) (Figure \@ref(fig:full-IAA-alignment)).
+In the alignment of A class Aux/IAAs, the EAR motif and degron can be readily identified by the drop in nonsynonymous variation, as visualized by the lack of strong variant functional effects (Figure \@ref(fig:A-IAA-alignment)).
+The PB1 domain is not as readily identified, perhaps because the multiple key residues are spread out in linear sequence space.
+It is worth noting that these key residues that facilitate electrostatic PB1-PB1 interactions show little variation.
 
@@
 ```
-
-(ref:A-IAA-alignment) **Canonical Aux/IAAs have conserved Degron and EAR motifs.** Protein sequences were aligned with DECIPHER [@wright_decipher_2015] and variants were mapped to this alignment and colored according to the predicted functional effect of the allele of strongest effect at that position, with light colors having weaker effects on function and darker colors stronger effects.
-Red indicates missense variants.
-Color scale is explained in [Methods](#methods).
-In grayscale underneath the plot, alignment consensus is shown as measured by Evolutionary trace [@pluciennik_balcony_2018; @wilkins_evolutionary_2012], with high consensus positions in black and low consensus in white.
-Key functional domains are outlined in black and labeled above the plot.
-The EAR domain spans codon alignment positions 70 to 74, corresponding to IAA1 amino acids LRLGL, 14-18.
-The degron domain spans alignment positions 194-201, corresponding to IAA1 amino acids QIVGWPPV, 55-62.
-The charged residues of the PB1 domain correspond to alignment positions 246, 256, 316, 318, 320, and 326 corresponding to IAA1 amino acids K77, R88, D133, D135, D137, and D143.
 
@@
 
+(ref:A-IAA-alignment) **Critical functional domains of the conserved *Aux/IAA* genes show low nonsynonymous variation compared to regions of unknown functional importance.** Protein sequences were aligned with DECIPHER [@wright_decipher_2015] and variants were mapped to this alignment and colored according to the predicted functional effect of the allele of strongest effect at that position, with light colors having weaker effects on function and darker colors stronger effects.
+Red indicates missense variants.
+Color scale is further explained in [Methods](#methods).
+Alignment consensus is shown in grayscale underneath the plot as measured by Evolutionary trace [@pluciennik_balcony_2018; @wilkins_evolutionary_2012], with high consensus positions in black and low consensus in white.
+Key functional domains are outlined in black and labeled above the plot.
+The EAR domain spans codon alignment positions 70 to 74, corresponding to IAA1 amino acids LRLGL, 14-18.
+The degron domain spans alignment positions 194-201, corresponding to IAA1 amino acids QIVGWPPV, 55-62.
+The charged residues of the PB1 domain correspond to alignment positions 246, 256, 316, 318, 320, and 326 corresponding to IAA1 amino acids K77, R88, D133, D135, D137, and D143.
+
 Natural variation also provides a means to study how gene families are evolving.
-To visualize this, we used ViVa to map the diversity at nonsynonymous variant sites relative to synonymous sites onto the Aux/IAA phylogenetic tree  (Figure \@ref(fig:IAAtree)).
-This visualization enables straightforward comparison of rates of recent functional divergence within natural variation in the context of the sequence divergence throughout the history of gene family.
-Clades and individual genes exhibiting low nonsynonymous diversity, and are therefore likely functionally conserved, can be identified.
-Conversely, genes with high nonsynonymous diversity are more likely to be under relaxed functional selection, indicating a possibility of functional drift, emergence of novel function, or pseudogenization.
+To visualize this, we used ViVa to map the diversity at nonsynonymous variant sites relative to synonymous variant sites (${\pi_N/\pi_S}$) onto the Aux/IAA phylogenetic tree  (Figure \@ref(fig:IAAtree)).
+This visualization enables straightforward comparison of rates of recent functional divergence as predicted by diversity within natural variation in the context of the sequence divergence throughout the history of a gene family.
+Clades and individual genes exhibiting low nonsynonymous diversity are likely functionally conserved.
+Conversely, genes with high nonsynonymous diversity are more likely to be under relaxed functional selection, indicating the possibility of functional drift, emergence of novel function, or pseudogenization.
 
@@
 
-(ref:IAAtree) **IAA protein sequence tree mapped with ${\pi_N/\pi_S}$.** Protein sequences were aligned with DECIPHER [@wright_decipher_2015] and low information content regions were masked with Aliscore [@kuck_parametric_2010] prior to inferring a phylogeny with MrBayes [@ronquist_mrbayes_2003].
+(ref:IAAtree) **IAA protein sequence phylogenetic tree mapped with ${\pi_N/\pi_S}$ reveals patterns of sister pair diversity/conservation.** Protein sequences were aligned with DECIPHER [@wright_decipher_2015] and low information content regions were masked with Aliscore [@kuck_parametric_2010] prior to inferring a phylogeny with MrBayes [@ronquist_mrbayes_2003].
 Tips of the tree are mapped with circles of color and diameter proportional to ${\pi_N/\pi_S}$.
-${\pi_N/\pi_S}$ provides a prediction of functional diversity.
-Nodes are labeled with the poster probability of monophyly.
+The ${\pi_N/\pi_S}$ statistic provides a prediction of functional diversity.
+Nodes are labeled with the posterior probability of monophyly, a measure of confidence in the branch assignment, with one representing high confidence and zero, low confidence.
 There are two distinct clades of Aux/IAAs represented by the majority of the A and B classes.
 C class Aux/IAAs are missing one or more of the canonical Aux/IAA domains.
+These classes are represented by the text color of the gene name.
 
@@
                                         max(IAA_data$PiN_PiS))) + scale_size(guide = "none") + 
-  theme(legend.position = c(.8,.7)) + 
+  theme(legend.position = c(.85,.7)) + 
   geom_nodelab(mapping = aes(label = round(as.numeric(prob),digits = 2)), nudge_x = -.3, nudge_y = .5, size = 3) + 
@@
   scale_color_viridis(end = .9, discrete = TRUE, option = "A") +  
-  labs(fill = expression(pi [N] / pi [S]), color = "IAA class")
+  labs(fill = expression(pi [N] / pi [S]), color = "Aux/IAA class")
 p
@@
 For example, *IAA10* and *IAA11* both showed functional conservation (${\pi_N/\pi_S}$ of 0.80 and 0.67 respectively).
-In support of this strong conservation of both *IAA10* and *IAA11*, the *Arabidopsis thaliana* ePlant browser indicates that *IAA10* and *IAA11* have almost identical expression patterns [@waese_eplant_2017].
+The *Arabidopsis thaliana* ePlant browser indicates that *IAA10* and *IAA11* have almost identical expression patterns [@waese_eplant_2017].
 Together this evidence suggests a strong dosage requirement for these genes or that they have taken on novel functions since their emergence, with both genes contributing similarly to plant fitness.
+In support of novel function, expression and mutant analysis during embryogenesis suggest that IAA10 is required for suspensor-hypophysis transition, while IAA11 is involved in later cell fate transitions [@rademacher_different_2012].

### 2019-04-21 manuscript/03-conclusion.Rmd
submitted revisions

@@
 
-ViVa has allowed our team of testers from various skill levels and backgrounds to meaningfully access the 1001 genomes dataset.
-The visualizations of natural variation further supported much of the existing structure-function knowledge of this well studied signaling pathway and facilitated the generation of new hypotheses.
-Application of ViVa to less studied genes and gene families promises to yield more novel hypotheses, which can be evaluated with mutagenesis and functional assays to glean novel structure/function knowledge from this rich dataset.
+ViVa has allowed our team of testers from various skill levels and backgrounds to meaningfully access and mine the 1001 genomes dataset.
+The visualizations of natural variation further supported much of the existing structure-function knowledge of the well-studied nuclear auxin signaling pathway and facilitated the generation of new hypotheses.
+Gene and gene family analyses can be combined, as was done here for nuclear auxin signaling, to understand variation within gene networks.
+Visualizations of variation within whole gene networks are planned for future iterations of ViVa.
+Application of ViVa to less studied genes and gene families promises to yield more novel hypotheses, which can be evaluated via genetic and functional assays to glean novel structure/function knowledge from this rich dataset.
 
@@
 We have assumed these missing variants are reference alleles, leading to undercounting in ViVa's diversity estimations.
-Recent advances in sequencing technologies have been combined to generate extremely high quality genomes [@michael_high_2018], and will reduce this source of uncertainty in future resequencing datasets.
+Visualizations of this uncertainty will be added to a future version of ViVa.
+Recent advances in sequencing technologies have been combined to generate extremely high-quality genomes [@michael_high_2018], and will reduce this source of uncertainty in future resequencing datasets.
 Another limitation is that the geographic coverage of accessions in the 1001 Genomes dataset is far from uniform, and thus diversity scores may not accurately reflect the allelic distributions of the global *Arabidopsis thaliana* population.

### 2019-04-21 manuscript/04-methods.Rmd
submitted revisions

@@
 ### Germplasm Accession Information
-A dataset of each of the 1135 accessions including CS stock numbers and geographic location where the samples were collected was retrieved from the 1001 Genomes website at http://1001genomes.org/accessions.html, via the download link at the bottom of the page.
+A dataset of each of the 1135 accessions including CS stock numbers and geographic locations where the samples were collected was retrieved from the 1001 Genomes website at http://1001genomes.org/accessions.html, via the download link at the bottom of the page.
 This data file has been embedded in the R package as `accessions`.
@@
 
-(ref:SNPeff-order) **Rank order of the strength of functional effects** Effect classes were ordered by subjective prediction of average strength of effect on gene function.
+(ref:SNPeff-order) **Rank order of the strength of functional effects** Variant effect classes were ordered by subjective prediction of average strength of effect on gene function.
 Strength was then assigned to each effect on an integer scale.
@@
 ## Testing
-A group consisting of undergraduate and graduate students, postdoctoral and postbaccalaureate researchers (the authors of this publication) were assembled to test the functionality of ViVa.
-Testers were asked to use the ViVa web interface to analyze the genetic variation within a gene family or clade within a family of the auxin nuclear signaling pathway.
+A group consisting of undergraduate and graduate students, post-doctoral and post-baccalaureate researchers (the authors of this publication) were assembled to test the functionality of ViVa.
+Testers were asked to use the ViVa web interface to analyze the genetic variation at the clade or whole gene family level of the auxin nuclear signaling pathway.
 Testers were provided with a brief vignette on how to use ViVa to formulate new hypotheses or support existing hypotheses about gene function and evolution, which has been expanded to the "An Overview of ViVa" section.
@@
 Tester experiences, issues, and suggestions were incorporated into the ViVa software.
-The results collected by these analyses are summarized in the "Visualizing Variation within the auxin signaling pathway" section and [Supplemental Data](#si).
+The results collected by these analyses are summarized in the "Visualizing Variation within the auxin signaling pathway" section and [Supplemental Analysis](#si).

### 2019-04-21 manuscript/06-supplement.Rmd
submitted revisions

@@
-# Supplemental Data {#si}
+# Supplemental Analysis {#si}
 
-## *TIR1/AFB* genes {#AFBs} 
+## Additional natural variation in the *Aux/IAA* genes {#IAAs}
+
+For simplicity we have included only the alignment of the class A Aux/IAAs in the main manuscript. We include here the complete alignment of the family (Figure \@ref(fig:full-IAA-alignment)).
+
+```{r IAAdivstats}
+gene_info <- geneInfoFromFile(
+  fname = system.file("shiny-app", "gene_family_data",
+                      "IAA_gene_ids.csv", package = "r1001genomes"))
+#annotation <- readAnnotationFile(filename = system.file("shiny-app", "gene_family_data",
+#                      "IAA_annotations.csv", package = "r1001genomes"), gene_info = gene_info)
+
+VCF <- readRDS(system.file("shiny-app", "gene_family_data",
+                      "IAAs_VCF.rds", package = "r1001genomes"))
+div_stats <- ldply(VCF, diversityStats, geneInfo=gene_info, .id="transcript_ID")
+div_stats$tair_symbol %<>% fct_reorder(.x = , div_stats$tair_symbol, 
+                                       .fun = function(x) 
+                                         {as.integer(str_remove(x, 
+                                              pattern = "IAA"))})
+div_stats %<>% arrange(tair_symbol)
+div_stats %<>% dplyr::select(tair_symbol, everything())
+
+```
+
+```{r IAA alignment sup}
+alignment <- alignCDS(gene_info$transcript_ID)
+# make an alignment data frame
+aln_df <- makeAlnDF(alignment[[2]])
+
+# make a collapsed alignment for translating positions
+collapsed_aln_df <- 
+  aln_df %>% dplyr::select(aln_pos, seq_pos, seq_name) %>%
+  tidyr::spread(key = seq_name, value = seq_pos)
+```
+
+```{r Mask_IAA_alignment_sup, eval=FALSE}
+aliscore(as.AAbin(alignment[[2]]), path = "/Volumes/SSD/Applications/ALISCORE_v2.0/Aliscore_v.2.0/")
+id <- scan(file = "input.fas_List_random.txt", sep = " ", quiet = TRUE)
+AAAlign_no_gaps <- as.matrix(alignment[[2]])[, id]
+write.nexus.data(AAAlign_no_gaps, file = "manuscript/IAAs.nex", format = "protein")
+```
+```{bash IAA_phylogeny_construction_sup, eval=FALSE}
+~/Desktop/natural-variation-webtool/manuscript
+
+/Applications/MrBayes/mb
+
+execute IAAs.nex
+# set IAA33 as the outgroup
+outgroup AT5G57420.1
+
+lset nucmodel=protein Rates=invgamma
+# or include Nst=6  ngammacat=4 omegavar=M3 
+prset aamodelpr=fixed(jones) statefreqpr=fixed(empirical) 
+# initial run with aamodelpr=mixed yielded posterior prob of 1 for jones model
+propset ExtTBR$prob=0
+
+mcmcp ngen=250000 nchains=4 printfreq=1000 samplefreq=100 burnin=200
+
+mcmc
+sump
+sumt
+
+```
+
+```{r Load IAA phylogeny and data_sup}
+tree <- treeio::read.beast(file = "trees/IAAs.nex.con.tre")
+
+IAA_data <- read.csv("tables/IAAs_table.csv")
+#tip_labels need to be in the first column
+IAA_data %<>% dplyr::select(name, everything())
+IAA_data$name %<>% as.character()
+IAA_data$AGI %<>% as.character()
+
+IAA_data <- IAA_data[match(gsub(x = tree@phylo$tip.label, pattern = "\\.1",replacement = ""), IAA_data$AGI), ] 
+tree@phylo$tip.label <- IAA_data$name
+```
+
+(ref:full-IAA-alignment) **Alignment of the complete Aux/IAA family.** Protein sequences were aligned with DECIPHER [@wright_decipher_2015] and variants were mapped to this alignment and colored according to the predicted functional effect of the allele of strongest effect at that position, with light colors having weaker effects on function and darker colors stronger effects.
+Red indicates missense variants.
+Color scale is further explained in [Methods](#methods).
+Alignment consensus is shown in grayscale underneath the plot as measured by Evolutionary trace [@pluciennik_balcony_2018; @wilkins_evolutionary_2012], with high consensus positions in black and low consensus in white.
+
+```{r full-IAA-alignment, fig.cap="(ref:full-IAA-alignment)", fig.height=8}
+IAAs <- alignCDS(gene_info$transcript_ID)
+consensus <- BALCONY::RealValET_conservativity(as.alignment(Biostrings::as.matrix(IAAs[[2]])))
+# make an alignment data frame
+IAAs <- makeAlnDF(IAAs[[2]])
+IAAs <- left_join(IAAs, dplyr::select(gene_info, "tair_locus",
+                                       "tair_symbol", "transcript_ID"),
+                        by = c("transcript_ID" = "transcript_ID"))
+IAAs$tair_symbol %<>% as.factor()
+IAAs$tair_symbol %<>% factor(levels = IAA_order, 
+                                       ordered = TRUE)
+IAAs$tair_symbol %<>% fct_rev()
+VCF <- ldply(.data = VCF, .fun = subset, gt_GT != "0|0")
+coding_VCF <- getCodingDiv(VCF)
+IAAs <- addSNPsToAlnDF(aln_df = IAAs,SNPs = coding_VCF, effect_order = SNPeff_order)
+IAAs$gap <- IAAs$seq_pos == "-"
+
+p <-ggplot(IAAs, aes(x = aln_pos, y = tair_symbol)) +
+  geom_raster(data = IAAs, 
+             # mapping = aes(fill = effects)) +
+              mapping = aes(fill = strength,
+                            alpha = gap)) +
+  scale_alpha_manual(values = c("TRUE" = 0, "FALSE" = 1), 
+                     labels = c("aligned", "gap"), 
+                     guide = guide_legend(override.aes = 
+                                            list(fill = c("grey85", "white")),
+                       title = "alignment",
+                       direction = "vertical")) + 
+   scale_fill_viridis(breaks = range(SNPeff_order$strength),
+                      limits = range(SNPeff_order$strength),
+                      na.value = "grey85",
+                      direction = -1,
+                      labels = c("weak", "strong"),
+                      option = "A", 
+                      guide = guide_colorbar(ticks = FALSE, 
+                                             title.vjust = .8, 
+                                             title.position = "top", 
+                                             title.hjust = 0.5)) +
+  scale_y_discrete(expand = expand_scale(add = c(1,2.5))) +
+  scale_x_continuous(expand = c(0,0)) +  # expand increases distance from axis
+  labs(x = "codon position in alignment", y = "", fill = "variant functional effect") +
+  theme_logo(base_family = "Helvetica") +
+  theme(panel.grid = element_blank(), 
+        panel.grid.minor = element_blank(),
+        strip.background = element_blank(),
+        strip.text.x = element_blank(),
+        #panel.background = element_rect(fill = "grey85", color = "white"),
+        legend.position = c("bottom"),
+        legend.key = element_rect(linetype = "solid"),
+        legend.spacing = unit(4, "char"), 
+        legend.box.spacing = unit(0.1,"line"))  +
+  annotation_raster(matrix(gray(rescale(consensus, to = c(0,1))), nrow = 1),  xmin = -Inf, xmax = Inf, ymin = -10, ymax = 0.4)
+p
+
+#ggsave("IAAs.png", width = 7, height = 4)
+```
+
+```{r IAA6-div-plot, fig.cap="(ref:IAA6-div-plot)", fig.height=4}
+coding_variants <- VCF[VCF$Transcript_ID == gene_info[gene_info$tair_symbol == "IAA6", "transcript_ID"], ]
+p <- plotCodingDiv(uniqueCodingVars = coding_variants)
+p + annotate("rect", xmin = 106, xmax = 127, ymin = 0.001, ymax = 1, 
+             color = "lightgrey", alpha =.1) + 
+  annotate("text", x = 116.5, y = 0.1, 
+           label = 
+             "region of positive selection\nidentified in Winkler et al.") + labs(color = "", y = "nucleotide diversity,\nlog scale") + theme(legend.position = "bottom", legend.text = element_text(size = 8), axis.title = element_text(size = 12)) + guides(color=guide_legend(nrow=5))
+```
+(ref:IAA6-div-plot) **IAA6 diversity plot.** Nucleotide diversity of variant positions throughout the IAA6 coding sequence are plotted and colored according to the effect of the variant alleles at each position.
+The region of positive selection identified by Winkler et al.
+is highlighted.
+
+## Natural variation in the *TIR1/AFB* genes {#AFBs} 
 Auxin acts by binding to receptors (Auxin-signaling F-Boxes, or AFBs) that in turn target co-repressors (Aux/IAAs) for degradation.
@@
 All members of this family have been shown to bind auxin and Aux/IAA proteins.
-However, AFB1 has drastically reduced ability to assemble into an SCF complex, due to the substitution E8K in its F-box domain, preventing it from inducing degradation of Aux/IAAs [@yu_untethering_2015].
+However, AFB1 has a drastically reduced ability to assemble into an SCF complex, due to the substitution E8K in its F-box domain, preventing it from inducing degradation of Aux/IAAs [@yu_untethering_2015].
 This lack of SCF formation may allow for the high and ubiquitous AFB1 accumulation observed in *Arabidopsis* tissues [@parry_complex_2009].
@@
 
-Examining the natural sequence variation across the *AFB* family \@ref(fig:AFBtree) revealed that *TIR1* and *AFB1* both had very low nonsynonymous diversity, hinting at their likely functional importance and bringing in to question the inconclusive role of *AFB1* in auxin signaling.
+Examining the natural sequence variation across the *AFB* family revealed that *TIR1* and *AFB1* both had very low nonsynonymous diversity (Figure \@ref(fig:AFBtree)), hinting at their likely functional importance and bringing in to question the inconclusive role of *AFB1* in auxin signaling.
 *AFB3* and *AFB4* had higher nonsynonymous diversity, while their sister genes, *AFB2* and *AFB5* were more conserved.
-This matches our current understanding of *AFB3* as playing a minor role in the auxin signaling pathway [@dharmasiri_plant_2005] and suggests *AFB4* may be undergoing pseudogenization, especially when paired with its low expression levels [@prigge_arabidopsis_2016].
+This matches our current understanding of *AFB3* as playing a minor role in the auxin signaling pathway [@dharmasiri_plant_2005].
+Two frameshift variants and one stop-gained (nonsense) variant were observed in *AFB4* supporting its pseudogenization., suggesting that *AFB4* may be undergoing pseudogenization, especially when paired with its low expression levels [@prigge_arabidopsis_2016].
+AFB4 and AFB5 have an N-terminal extension prior to their F-box domains.
+This extension had very high nonsynonymous diversity (Figure \@ref(fig:full-AFBs-alignment)), suggesting that this extension does not play an important functional role in these proteins.
 
@@
 
-Although most known functional regions are highly conserved in *AFB1*, there are some nonsynonymous polymorphism in the oligomerization domain that are only present in single accessions (F125E in Can-0 and I163N in Pu2-23).
+Although most known functional regions are highly conserved in *AFB1*, there are some nonsynonymous polymorphism in the oligomerization domain that are only present in single accessions (F125E in Can-0 and I163N in Pu2-23, as shown in Figure \@ref(fig:AFB1-map)).
 Mutations in this domain of *TIR1* frequently have a semidominant effect on root phenotypes [@dezfulian_oligomerization_2016; @wright_insights_2017].
@@
 
-```{r AFB1-map, fig.cap="(ref: AFB1-map)"}
+```{r AFB1-map, fig.cap="(ref:AFB1-map)"}
 knitr::include_graphics("app_figure/Map_figure.png")
 ```
-(ref: AFB1-map) **Map of AFB1 Oligomerization domain variant accessions** (a) Map showing the two accessions with variants in the AFB1 oligomerization domain (b) The user selectable parameters of the mapping tab used to generate the map are provided as an example of using the SNP Mapping tab in the ViVa webtool.
+
+(ref:AFB1-map) **Map of AFB1 Oligomerization domain variant accessions.** (a) Map showing the two accessions with variants in the AFB1 oligomerization domain (b) The user selectable parameters of the mapping tab used to generate the map are provided as an example of using the SNP Mapping tab in the ViVa webtool.
 See Figure \@ref(fig:app-workflow) for details on filling in the parameters.
 
-The AFB4 and AFB5 receptors have an N-terminal extension prior to their F-box domains.
-This extension had very high nonsynonymous diversity (Figure \@ref(fig:full-AFBs-alignment)), suggesting that this extension does not play an important functional role in these proteins.
-Additionally, two frameshift variants and one stop-gained variant were observed in *AFB4* supporting its pseudogenization.
-
-## *TPL/TPR* genes {#TPLs}
+## Natural variation in the *TPL/TPR* genes {#TPLs}
 
@@
 Canonical EAR domains have the amino acid sequence LxLxL, as found in most AUX/IAAs [@overvoorde_functional_2005].
-TPL/TPR co-repressors bind EAR domains via their C-terminal to LisH (CTLH) domains found near their N-termini (citations of pre-structure founding papers/reviews).
+TPL/TPR co-repressors bind EAR domains via their C-terminal to LisH (CTLH) domains found near their N-termini [@long_topless_2006].
 Recent structural analyses of the TPL N-terminal domain have highlighted the precise interaction interface between TPL and AUX/IAA EAR domains, as well as the TPL-TPL dimerization and tetramerization motifs [@martin-arevalillo_structure_2017; @ke_structural_2015].
@@
 
-The high degree of conservation in the entire N-terminal domain underscores its importance in TPL/TPR function (Figure \@ref(fig:full-TPLs-alignment).
+The high degree of conservation in the entire N-terminal domain underscores its importance in TPL/TPR function (Figure \@ref(fig:full-TPLs-alignment)).
 For example, the initial *tpl-1* mutation (N176H) in the ninth helix is a dominant gain-of-function allele [@long_topless_2006], which is capable of binding wild-type TPL protein and inducing protein aggregation [@ma_d53_2017].
@@
 
-# *ARF* genes {#ARFs}
+## Natural Variation in the *ARF* genes {#ARFs}
 
@@
 In the analyzed natural variation, ARF7 had several expansions of polyglutamine sequences in the middle region.
-Polyglutamine regions are known to readily expand and contract throughout evolutionary time due to replication error, and variation in polyglutamine length can be acted on by natural selection and have phenotypic consequences [@press_overdue_2014].
+Polyglutamine regions are known to readily expand and contract throughout evolutionary time due to replication error, and variation in polyglutamine length can have phenotypic consequences and be acted on by natural selection [@press_overdue_2014].
 The ARF DNA-binding domain had very few, low-diversity missense mutations, as did the critical residues of the PB1 domain.
@@
 ARF13 has many high-diversity nonsense variants and lacks a C-terminal PB1 domain.
-This high level of diversity, prevelance of high-frequency nonsense variants and frequent loss of critical domains, may suggest that several genes in this class are undergoing pseudogenization.
+This high level of diversity, prevalance of high-frequency nonsense variants and frequent loss of critical domains, may suggest that several genes in this class are undergoing pseudogenization.
 
@@
 Class C ARFs show low nucleotide diversity scores, with all ${\pi_N/\pi_S}$ values substantially lower than 1.
-*ARF16* was the most conserved, whereas its clade members (*ARF10*, *ARF17*) had scores at least four times higher.
+*ARF16* was the most conserved, whereas the other clade members (*ARF10*, *ARF17*) had scores at least four times higher.
 Structurally, all three members of Class C ARFs contain a canonical B3 DNA-binding domain, but only ARF10 and ARF16 contain a PB1 domain.
-The DNA binding domains exhibit overall low diversity.
-Of the PB1 domain containing class C ARFs, ARF16 exhibits several missense variants which are sporadically distributed, in contrast to the conserved PB1 domain of ARF10.
-This conservation in the PB1 domain of ARF10 and the DBD of ARF16 may suggest subfunctionalization in this family.
-
-## *Aux/IAA* genes {#IAAs}
-
-For simplicity we have included only the alignment of the class A Aux/IAAs in the main manuscript.
-For completion we include here the complete alignment of the family (Figure \@ref(fig:full-IAA-alignment)).
-
-```{r IAAdivstats}
-gene_info <- geneInfoFromFile(
-  fname = system.file("shiny-app", "gene_family_data",
-                      "IAA_gene_ids.csv", package = "r1001genomes"))
-#annotation <- readAnnotationFile(filename = system.file("shiny-app", "gene_family_data",
-#                      "IAA_annotations.csv", package = "r1001genomes"), gene_info = gene_info)
-
-VCF <- readRDS(system.file("shiny-app", "gene_family_data",
-                      "IAAs_VCF.rds", package = "r1001genomes"))
-div_stats <- ldply(VCF, diversityStats, geneInfo=gene_info, .id="transcript_ID")
-div_stats$tair_symbol %<>% fct_reorder(.x = , div_stats$tair_symbol, 
-                                       .fun = function(x) 
-                                         {as.integer(str_remove(x, 
-                                              pattern = "IAA"))})
-div_stats %<>% arrange(tair_symbol)
-div_stats %<>% dplyr::select(tair_symbol, everything())
-
-```
-
-```{r IAA alignment sup}
-alignment <- alignCDS(gene_info$transcript_ID)
-# make an alignment data frame
-aln_df <- makeAlnDF(alignment[[2]])
-
-# make a collapsed alignment for translating positions
-collapsed_aln_df <- 
-  aln_df %>% dplyr::select(aln_pos, seq_pos, seq_name) %>%
-  tidyr::spread(key = seq_name, value = seq_pos)
-```
-
-```{r Mask_IAA_alignment_sup, eval=FALSE}
-aliscore(as.AAbin(alignment[[2]]), path = "/Volumes/SSD/Applications/ALISCORE_v2.0/Aliscore_v.2.0/")
-id <- scan(file = "input.fas_List_random.txt", sep = " ", quiet = TRUE)
-AAAlign_no_gaps <- as.matrix(alignment[[2]])[, id]
-write.nexus.data(AAAlign_no_gaps, file = "manuscript/IAAs.nex", format = "protein")
-```
-```{bash IAA_phylogeny_construction_sup, eval=FALSE}
-~/Desktop/natural-variation-webtool/manuscript
-
-/Applications/MrBayes/mb
-
-execute IAAs.nex
-# set IAA33 as the outgroup
-outgroup AT5G57420.1
-
-lset nucmodel=protein Rates=invgamma
-# or include Nst=6  ngammacat=4 omegavar=M3 
-prset aamodelpr=fixed(jones) statefreqpr=fixed(empirical) 
-# initial run with aamodelpr=mixed yielded posterior prob of 1 for jones model
-propset ExtTBR$prob=0
-
-mcmcp ngen=250000 nchains=4 printfreq=1000 samplefreq=100 burnin=200
-
-mcmc
-sump
-sumt
-
-```
-
-```{r Load IAA phylogeny and data_sup}
-tree <- treeio::read.beast(file = "trees/IAAs.nex.con.tre")
-
-IAA_data <- read.csv("tables/IAAs_table.csv")
-#tip_labels need to be in the first column
-IAA_data %<>% dplyr::select(name, everything())
-IAA_data$name %<>% as.character()
-IAA_data$AGI %<>% as.character()
-
-IAA_data <- IAA_data[match(gsub(x = tree@phylo$tip.label, pattern = "\\.1",replacement = ""), IAA_data$AGI), ] 
-tree@phylo$tip.label <- IAA_data$name
-```
-
-(ref:full-IAA-alignment) **Alignment of the complete Aux/IAA family.** Protein sequences were aligned with DECIPHER [@wright_decipher_2015] and variants were mapped to this alignment and colored according to the predicted functional effect of the allele of strongest effect at that position, with light colors having weaker effects on function and darker colors stronger effects.
-Red indicates missense variants.
-Color scale is explained in [Methods](#methods).
-In grayscale underneath the plot, alignment consensus is shown as measured by Evolutionary trace [@pluciennik_balcony_2018; @wilkins_evolutionary_2012], with high consensus positions in black and low consensus in white.
-
-```{r full-IAA-alignment, fig.cap="(ref:full-IAA-alignment)", fig.height=8}
-IAAs <- alignCDS(gene_info$transcript_ID)
-consensus <- BALCONY::RealValET_conservativity(as.alignment(Biostrings::as.matrix(IAAs[[2]])))
-# make an alignment data frame
-IAAs <- makeAlnDF(IAAs[[2]])
-IAAs <- left_join(IAAs, dplyr::select(gene_info, "tair_locus",
-                                       "tair_symbol", "transcript_ID"),
-                        by = c("transcript_ID" = "transcript_ID"))
-IAAs$tair_symbol %<>% as.factor()
-IAAs$tair_symbol %<>% factor(levels = IAA_order, 
-                                       ordered = TRUE)
-IAAs$tair_symbol %<>% fct_rev()
-VCF <- ldply(.data = VCF, .fun = subset, gt_GT != "0|0")
-coding_VCF <- getCodingDiv(VCF)
-IAAs <- addSNPsToAlnDF(aln_df = IAAs,SNPs = coding_VCF, effect_order = SNPeff_order)
-IAAs$gap <- IAAs$seq_pos == "-"
-
-p <-ggplot(IAAs, aes(x = aln_pos, y = tair_symbol)) +
-  geom_raster(data = IAAs, 
-             # mapping = aes(fill = effects)) +
-              mapping = aes(fill = strength,
-                            alpha = gap)) +
-  scale_alpha_manual(values = c("TRUE" = 0, "FALSE" = 1), 
-                     labels = c("aligned", "gap"), 
-                     guide = guide_legend(override.aes = 
-                                            list(fill = c("grey85", "white")),
-                       title = "alignment",
-                       direction = "vertical")) + 
-   scale_fill_viridis(breaks = range(SNPeff_order$strength),
-                      limits = range(SNPeff_order$strength),
-                      na.value = "grey85",
-                      direction = -1,
-                      labels = c("weak", "strong"),
-                      option = "A", 
-                      guide = guide_colorbar(ticks = FALSE, 
-                                             title.vjust = .8, 
-                                             title.position = "top", 
-                                             title.hjust = 0.5)) +
-  scale_y_discrete(expand = expand_scale(add = c(1,2.5))) +
-  scale_x_continuous(expand = c(0,0)) +  # expand increases distance from axis
-  labs(x = "codon position in alignment", y = "", fill = "variant functional effect") +
-  theme_logo(base_family = "Helvetica") +
-  theme(panel.grid = element_blank(), 
-        panel.grid.minor = element_blank(),
-        strip.background = element_blank(),
-        strip.text.x = element_blank(),
-        #panel.background = element_rect(fill = "grey85", color = "white"),
-        legend.position = c("bottom"),
-        legend.key = element_rect(linetype = "solid"),
-        legend.spacing = unit(4, "char"), 
-        legend.box.spacing = unit(0.1,"line"))  +
-  annotation_raster(matrix(gray(rescale(consensus, to = c(0,1))), nrow = 1),  xmin = -Inf, xmax = Inf, ymin = -10, ymax = 0.4)
-p
-
-#ggsave("IAAs.png", width = 7, height = 4)
-```
-
-```{r IAA6-div-plot, fig.cap="(ref:IAA6-div-plot)"}
-coding_variants <- VCF[VCF$Transcript_ID == gene_info[gene_info$tair_symbol == "IAA6", "transcript_ID"], ]
-p <- plotCodingDiv(uniqueCodingVars = coding_variants)
-p + annotate("rect", xmin = 106, xmax = 127, ymin = 0.001, ymax = 1, color = "lightgrey", alpha =.1) + annotate("text", x = 116.5, y = 0.5, label = "region of positive\nselection identified\nin Winkler et al.")
-```
-(ref:IAA6-div-plot) **IAA6 diversity plot.** Nucleotide diversity of variant positions throughout the IAA6 coding sequence are plotted and colored according to the effect of the variant alleles at each position.
-The region of positive selection identified by Winkler et al.
-is highlighted.
+The DNA binding domains exhibited overall low diversity.
+Of the PB1 domain containing class C ARFs, ARF16's PB1 domain exhibited several missense variants which are sporadically distributed, in contrast to the conserved PB1 domain of ARF10.
+This conservation in the PB1 domain of ARF10 and the DBD of ARF16 may suggest subfunctionalization in this class.

### 2019-04-21 manuscript/index.Rmd
submitted revisions

@@
   fig.width=7, 
-  dpi = 600, 
+  dpi = 800, 
   out.extra='style="margin: auto; display: block; padding-top: 15px;"')

### 2019-09-26 manuscript/02-results.Rmd
final submission, minus a few copyedits

@@
 knitr::include_graphics("app_figure/workflow_fig.png")
+
 ```
@@
 
-#ggsave("A-IAAs_alignment.png", width = 6, height = 4)
+ggsave("A-IAAs_alignment.tiff", dpi = 800)
 ```
@@
 
-#ggsave("IAA-tree.png", width = 6, height = 5.5)
+ggsave("IAA-tree.tiff", dpi = 800)
 ```

### 2019-09-26 manuscript/06-supplement.Rmd
final submission, minus a few copyedits

@@
 
-#ggsave("IAAs.png", width = 7, height = 4)
+ggsave("full-IAA-alignment.tiff", dpi = 800)
 ```
@@
              "region of positive selection\nidentified in Winkler et al.") + labs(color = "", y = "nucleotide diversity,\nlog scale") + theme(legend.position = "bottom", legend.text = element_text(size = 8), axis.title = element_text(size = 12)) + guides(color=guide_legend(nrow=5))
+
+ggsave("IAA6-div-plot.tiff", dpi = 800)
 ```
@@
 p
-#ggsave("AFBs_tree.pdf", height = 3, width = 4)
+
+ggsave("AFB_tree.tiff", dpi = 800)
 ```
@@
 
+ggsave("full-AFBs-alignment.tiff", dpi = 800)
 ```
@@
 p
+
+ggsave("TPLtree.tiff", dpi = 800)
 ```
@@
 
+ggsave("full-TPLs-alignment.tiff", dpi = 800)
 ```
@@
 p
+
+ggsave("ARFtree.tiff", dpi = 800)
 ```
@@
 
+ggsave("full-ARFs-alignment.tiff", dpi = 800)
+
 ```
